### 2015-01-16 HGmodel_draft3.tex
add new manu

% \documentclass[rmp,aps,floatfix,authordate1-4,preprint]{revtex4}
% \documentclass[prb,aps,floatfix,authordate1-4,preprint]{revtex4}

\documentclass[pre,aps,floatfix,authordate1-4,twocolumn]{revtex4-1}
%\documentclass[pre,aps,floatfix,authordate1-4]{revtex4-1}

%\documentclass[aps,prl,preprint,superscriptaddress]{revtex4}

%\documentclass[aps,prl,preprint,groupedaddress]{revtex4}

\usepackage{rotating} 
\usepackage{times}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[obeyFinal]{easy-todo}
%\usepackage{scrtime}
\usepackage[multiple]{footmisc}
%\usepackage[sort&compress]{natbib}
%\bibpunct{(}{)}{,}{n}{}{}
%\renewcommand{\bibnumfmt}[1]{#1.}
\usepackage[normalem]{ulem}

\begin{document}
%\singlespacing

\title{Towards atomistic resolution structure of phosphatidylcholine glycerol backbone and choline headgroup at different ambient conditions}

\author{Alexandru Botan}
\thanks{The authors are listed in alphaphetical order.}
\thanks{The author list is not completed.}
\thanks{Institut Lumi\`ere Mati\`ere, UMR5306 Universit\'e Lyon 1-CNRS, Universit\'e de Lyon 69622 Villeurbanne, France}
%
\author{Andrea Catte}
\thanks{University of East Anglia, Norwich, United Kingdom}
\author{Fernando Favela}
\author{Patrick Fuchs}
\thanks{Institut Jacques Monod, CNRS, Université Paris Diderot, Sorbonne Paris Cité, Paris, France}
%\thanks{Université Paris Diderot, Sorbonne Paris Cité, Paris, France}
\author{Matti Javanainen}
\thanks{Tampere University of Technology, Tampere, Finland}
\author{Waldemar Kulig}
\thanks{Tampere University of Technology, Tampere, Finland}
\author{Antti Lamberg}
\thanks{Department of Chemical Engineering, Kyoto University, Kyoto, Japan}
%\author{Alexander Lyubartsev}
\author{Markus S. Miettinen}
\thanks{Fachbereich Physik, Freie Universität Berlin, Berlin, Germany}
\author{Luca Monticelli}
\thanks{5IBCP, CNRS UMR 5086, Lyon, France}
\author{Jukka M\"a\"att\"a}
\thanks{Aalto University, Espoo, Finland}
%\setcounter{page}{1}
\author{Vasily S. Oganesyan}
\thanks{University of East Anglia, Norwich, United Kingdom}
\author{O. H. Samuli Ollila} 
\thanks{{\bf Author to whom correspondence may be addressed. E-mail: samuli.ollila@aalto.fi.}}
\thanks{Helsinki Biophysics and Biomembrane Group, Department of Biomedical Engineering and Computational Science, Aalto University, Espoo, Finland}
\author{Marius Retegan}
\thanks{Max Planck Institute for Chemical Energy Conversion, Mülheim an der Ruhr, Germany}
\author{Hubert Santuz}
\thanks{INSERM, U1134, DSIMB; Institut National de la Transfusion Sanguine (INTS); Laboratoire d'Excellence GR-Ex, Paris, France}
\thanks{Université Paris Diderot, Sorbonne Paris Cité, Paris, France}
\author{Joona Tynkkynen}
\thanks{Tampere University of Technology, Tampere, Finland}
%\author{Mark Wilson}
%\author{Alexander Vogel}

\begin{abstract}
Phospholipids are essential building blocks of biological membranes.
Despite of vast amount of accurate experimental data the atomistic resolution structures sampled by the glycerol backbone and choline headgroup
in phoshatidylcholine bilayers are not known. Atomistic resolution molecular dynamics simulation model 
would automatically resolve the structures giving an interpretation of experimental results, if the model
would reproduce the experimental data. In this work we compare the C-H bond vector order
parameters for glycerol backbone and choline headgroup between 14 different atomistic resolution
models and experiments in fully hydrated lipid bilayer. The current models are not accurately enough to resolve the structure.
However, closer inspection of three best performing models (CHARMM36, GAFFlipid and MacRog) suggest
that improvements in the sampled dihedral angle distributions would potentilly lead to the model which
would resolve the structure. Despite of the inaccuracy in the fully hydrated
structures, the response to the dehydration, i.e. P-N vector tilting more parallel to membrane normal, 
is qualitatively correct in all models. The CHARMM36 and MacRog models describe the interactions
between lipids and cholesterol better than Berger/H\"oltje model.
This work has been, and continues to be, progressed and discussed through the blog: nmrlipids.blogspot.fi. 
Everyone is invited to join the discussion and make contributions through the blog. 
The manuscript will be eventually submitted to an appropriate scientific journal. 
Everyone who has contributed to the work through the blog will be offered 
coauthorship. For more details see: nmrlipids.blogspot.fi.

\end{abstract}

\maketitle

~\vspace{0.3cm}\\
{\it \bf}

\section{Introduction}

Phospholipids containing different polar headgroups and different acyl chains are essential building blocks of 
biological membranes. Lamellar phospholipid bilayer structures have been widely studied with various experimental 
and theoretical techniques as a simple model for the biological membranes~\cite{lipowsky95,tieleman97,klauda08,edholm08,tieleman10,piggot12,rabinovich13,marsh13}. 
Phospholipid molecules are composed of hydrophobic acyl chains and hydrophilic headgroup, which are connected by glycerol backbone,
see Fig.~\ref{POPCstructure} for the structure of 1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
The behaviour of the acyl chains in a bilayer is relatively well understood~\cite{Israelachvili80,lipowsky95,tieleman97,klauda08,edholm08,tieleman10,marsh13}. 
The structures sampled by the glycerol backbone and choline in liquid bilayer state, however, are not fully 
resolved since even the most accurate scattering and Nuclear Magnetic Resonance (NMR)
techniques give only a set of values that the structure has to fulfill, but
there is no unique way to derive the actual structure from these parameters~\cite{seelig77b,skarjune79,Israelachvili80,jacobs80,davis83,strenk85,akutsu91,hong95b,hong96,semchyschyn04}.
Some structural details have been extracted from crystal structure, $^1$H NMR studies and Raman spectroscopy~\cite{hauser80,hauser81,hauser81b,akutsu81b,pascher92,hauser88,marsh06}
but general consensus about the structures sampled in the liquid state have not been reached~\cite{seelig77b,skarjune79,Israelachvili80,jacobs80,davis83,strenk85,hauser88,akutsu91,hong95b,hong96,semchyschyn04,marsh06}. 
On the other hand, the glycerol backbone structures are similar for various biologically
relevant lipid species (phosphatidylcholine (PC), phosphatidylethanolamine (PE) and phosphatidylglycerol (PG)) 
in various environments~\cite{gally81} and the headgroup choline structures are similar in model membranes and
real cells (mouse fibroblast L-M cell)~\cite{scherer87}.
Thus, the resolvation of phosphatidylcholine glycerol backbone and choline structures would be 
useful for understanding wide range of different biological membranes.

Classical atomistic resolution molecular dynamics simulations have been widely used to study  
lipid bilayers~\cite{tieleman97,klauda08,edholm08,tieleman10,piggot12,rabinovich13}. As these models provide an atomistic
resolution description of the whole lipid molecule, they have potential to resolve the glycerol backbone and 
headgroup structures. In particular, the experimental C-H bond order parameters for the glycerol backbone 
(g$_1$, g$_2$ and g$_3$) and choline ($\alpha$ and $\beta$) segments (see Fig.~\ref{POPCstructure} for definitions) are among the main parameters used in
attempts to derive the structures from experimental data~\cite{seelig77b,skarjune79,jacobs80,davis83,akutsu91,hong95b,semchyschyn04}.
These parameters are also routinely compared between experiments and simulations for the acyl chains~\cite{tieleman97,klauda08,edholm08,tieleman10,piggot12}.
Thus, the structures sampled in a simulation model that reproduces these and other experimental parameters, automatically
give an interpretation of the experiments, in other words they can be considered as reasonable atomistic resolution descriptions of
the behaviour of lipid molecules in a bilayer.

The glycerol backbone and choline headgroup order parameters have been compared between simulations and experiments
in some studies~\cite{shinoda97,hogberg08,castro08,klauda10,kapla12,dickson12,poger12,ferreira13,chowdhary13,maciejewski14}, 
however much less frequently than for acyl tail chains~\cite{tieleman97,klauda08,edholm08,tieleman10,piggot12}.
The main reason is probably that the existing experimental data for the glycerol backbone
and choline headgroups is scattered over many publications and published in a format that is difficult to understand without some NMR expertise. 
In addition to the order parameters, also dihedral angles for glycerol backbone and headgroup estimated from experiments have been sometimes used to 
assess the quality of simulation model~\cite{robinson94,essex94,kothekar96,hyvonen97,shinoda97,duong99}.

In this work we first review the most relevant experimental data for the glycerol backbone and choline headgroup order parameters
in a phosphatidylcholine lipid bilayer. Then the available atomistic resolution lipid models are carefully compared to the 
experimental data. The comparison reveals that the CHARMM36~\cite{klauda10}, GAFFlipid~\cite{dickson12} and MacRog models~\cite{maciejewski14}
have the most realistic glycerol backbone and choline structures. We also compare the glycerol backbone and choline 
structures between the most often used (Berger based) lipid model~\cite{berger97} 
\todo{Based on the results and discussions with Olle Edholm and Peter Tieleman in the blog, it seems that 
the glycerol backbone dihedrals are not the same in the model used in the publication by Berger et al.~\cite{berger97}
and in the model shared in the webpage of Peter Tieleman (http://wcm.ucalgary.ca/tieleman/downloads).
In this work (as it is quite often done in other works as well) we currently refer to the model in the webpage as ''Berger'' model.
This is misleading. We have to call it something else. One suggestion could be to call it ''BergerM'', because it was firt time
used in Marrink et al. Biophys. J. 74,931-943 (1998). Any other suggestions?\\
COMMENT: Patrick Fuchs suggested ''Berger-T''. This is better. \\
COMMENT2: Current trial, based in blog discussion, is to use Berger-MOLECULE-YEAR [citation].
}
and the best performing models, to demonstrate that by using the 
order parameters we can distinguish the more reasonable structures from the less reasonable ones. However, none of the current models 
is accurate enough to resolve the atomistic resolution stuctures.

In addition to the fully--hydrated single--component lipid bilayers, the glycerol backbone and choline order parameters
have been measured under a large number of different conditions. For example, as a function of hydration level~\cite{bechinger91,ulrich94,dvinskikh05b}, cholesterol content~\cite{brown78,ferreira13}
ion concentration~\cite{brown77,akutsu81,altenbach84,roux90,roux91}, temperature~\cite{gally75}, charged lipid content~\cite{roux90,roux91}, charged surfactant content~\cite{scherer89}, 
drug molecule concentration~\cite{browning82,kelusky84,castro08}, and protein content~\cite{roux89,kuchinka89} (listing only the publications most relevant for this work and the pioneering studies).
Awareness of the existence of this type of data allows the comparison of structural responses to varying conditions between simulations and experiments,
which can be used to validate the simulation models and to interpret the original experiments. 
In this publication we demonstrate the power of this approach for understanding the behaviour of a bilayer as a function of hydration level and cholesterol concentration.
Choline headgroup order parameters as function of ion concentration, and their relation to the ion binding affinity, are discussed elsewhere~\cite{ionpaper}.

  \begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{POPCstructure.eps}

  \caption{\label{POPCstructure}
    Chemical structure of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).}

\end{figure}

\section{methods}

\subsection{Open collaboration}
\todo{Should we write more about this?}

This work has been done as an open collaboration by using nmrlipids.blogspot.fi as an communication platform.
The approach is inspired by the Polymath project~\cite{gowers09}, however there are some essential differences. The manuscript pointing out problems in the glycerol backbone and
headgroup structure in the most used molecular dynamics simulation model for lipid bilayers (Berger based models) was used as a starting point~\cite{ollila13}.
After publishing the initial manuscript, the blog was opened for further contributions and discussion from anyone interested. 
All the contributions were done publicly through the blog and the blog contributors were offered coauthorship. The final author list
is based on the self-assesment of the authors scientific contribution to the project.

Large portion of the used simulation files and scripts are shared through the GitHub organization https://github.com/NMRlipids
and Zenodo community https://zenodo.org/collection/user-nmrlipids.

\subsection{Analysis}
\todo{ We should mention the sign issue and refer to the blog post:
http://nmrlipids.blogspot.fi/2014/04/on-signs-of-order-parameters.html somewhere.
Maybe we should change the name of this section II.B. to: "Analysis and terminology of order parameters from experiments and simulations" or similar, and also move the "forking" explanation here.}
The order parameter of a hydrocarbon C-H vector is defined as $S_{{\rm CH}}=\frac{3}{2}\langle \cos^2 \theta-1 \rangle$, where
the average is an ensemble average over the sampled conformations, and $\theta$ is the angle between the C-H bond and the membrane normal.
The order parameters can be measured by detecting quadrupolar splitting with $^2$H NMR~\cite{seelig77c} or by detecting dipolar 
splitting with $^{13}$C NMR~\cite{hong95a,gross97,dvinskikh05a,ferreira13}. The measurements are based on
different physical interactions and also the connection between order parameters and quadrupolar or dipolar splitting
are different. The order parameters from the measured quadrupolar splitting $\Delta \nu_Q$ ($^2$H NMR) are calculated using 
the equation $|S_{{\rm CD}}|=\frac{4}{3} \frac{e^2qQ}{h} \Delta \nu_Q$, where the value for the static quadrupole
splitting constant is estimated from various experiments to be 170~kHz leading to a numerical relation $|S_{{\rm CD}}|=0.00784 \times \Delta \nu_Q$~\cite{seelig77c}. 
The order parameters from the measured dipolar splitting $d_\mathrm{CH}$ ($^{13}$C NMR) are calculated using equation
$|S_{{\rm CH}}|=\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c} d_\mathrm{CH}$, where
values between 20.2-22.7 kHz are used for $\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c}$,
depending on the original authors~\cite{hong95a,gross97,dvinskikh05a,ferreira13}.
It is important to note that the order parameters measured with different techniques based on different physical interactions are in good agreement
with each other (see Results and Discussion), indicating very high quantitative accuracy of the measurements.
For more detailed discussion see http://nmrlipids.blogspot.fi/2014/02/accuracy-of-order-parameter-measurements.html
\todo{This is temporarily linked to the blog. Currently I think that the posts (with comments included) to which we want to cite are put into 
the Figshare in pdf format as they are in when the publication is submitted. Figshare allows commenting to continue the discussion.}.

The order parameters from simulations were calculated directly using the definition.
%$S_{{\rm CH}}=\frac{3}{2} \langle \cos^2 \theta-1 \rangle$. 
For the united atom models the hydrogen locations were generated 
post-simulationally using the positions of the heavy atoms in the simulation trajectories.
The statistical error estimates were calculated for the best performing simulation
models by calculating the error of the mean for the average over individual lipids in the system.

It has been recently pointed out that the sampling of individual dihedral angles might be very
slow compared to the typical simulation timescales~\cite{vogel12}. On the other hand, another recent
study shows that the slowest rotational correlation functions of C-H bond (g$_1$) reaches plateau (S$_{CH}^2$)
after 200ns in the Berger-POPC-07~\cite{ollila07a} model, and that the dynamics of this segment is significantly too slow in simulations
compared to the experiments~\cite{ferreira15}. In practise, less than 200~ns data set is enough for the order parameter
calculation due to the average over different lipid molecules. In conclusion, if the sampling with typical simulation times
is not enough for the convergense of the order parameters, then the simulation models has significantly too slow dynamics.

\subsection{Simulated systems}
All simulations are ran with a standard setup for planar lipid bilayer in zero tension and constant temperature
with periodic boundary conditions in all directions by using Gromacs software package~\cite{hess08} 
(version numbers 4.5-4.6) or LAMMPS~\cite{plimpton95}.
The number of molecules, temperatures and the length of simulations for all the simulated systems 
are listed in Tables~\ref{systems},~\ref{systemsDEHYD} and~\ref{systemsCHOL}. Full simulation
details are given in the Supplementary Information (SI) or in the original publications if the
data is used previously. For some systems also the simulation related files and trajectories are publicly
available. The references pointing to simulation details and files are also listed in Tables~\ref{systems},~\ref{systemsDEHYD} and~\ref{systemsCHOL}.

\begin{table*}[htb]
\centering
\caption{Simulated single component lipid bilayers with full hydration. The simulation file data sets marked with $^*$ include also part of the trajectory.
  If simulation data from previously published work has been directly used, the original publication is cited for simulation details. For other systems the simulation details
  are given in Supplementary Information. 
}\label{systems}
\begin{tabular}{c c c c c c c c c}
%\hline
Force field & lipid  & \footnote{The number of lipid molecules}N$_{\rm l}$   &  \footnote{The number of water molecules}N$_{\rm w}$ & \footnote{Simulation temperature}T (K)  & \footnote{The total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time frames used in the analysis}t$_{{\rm anal}}$ (ns) & \footnote{Reference link for the downloadable simulation files}Files  &  \footnote{Reference for the full simulation details} Details\\
\hline
Berger-POPC-07~\cite{ollila07a}          &   POPC & 128 & 7290  & 298  & 270 & 240 & \cite{bergerFILESpopc}$^*$ & \cite{ferreira15} \\
\sout {Berger-POPC-07~\cite{ollila07a}}\todoi{Not included in the Fig. 2, to be removed}          &   POPC & 128 & 7290  & 323  & 56 & 56  & ? & SI \\
Berger-DPPC-98~\cite{marrink98}          &   DPPC & 72 & 2864  & 323  & 140 & 100  & \cite{bergerDPPCfiles} & SI \\
Berger-DMPC-04~\cite{gurtovenko04}          &   DMPC & 128 & 5097  & 323  & 130 & 100  & \cite{dmpcFILES} & \cite{miettinen09} \\
CHARMM36\cite{klauda10}       & POPC   & 72  &  2242 & 303 & 30 & 20  & \cite{charmm36filesSHORT}$^*$ & SI \\
CHARMM36\cite{klauda10}\todoi{Hubert Santuz, these files are already in Gitbub, however to get doi citation it should be put to Zenodo}        & POPC   & 128 &  5120    & 303 & 150 & 100  & ? & SI \\
CHARMM36\cite{klauda10}\todoi{Samuli, put to Zenodo}        & DPPC   & 72  &  2189 & 323 & 30 & 25  & ? & SI \\
CHARMM36\cite{klauda10}\todoi{Not included in the Fig. 2. Should we remove or include?}        & DOPC   & 128 &  5120    & 303 & 150 & 100  & ? & SI \\
MacRog\cite{maciejewski14}  & POPC & 288  & 12600 & 310 & 100 & 80  & \cite{macrogFILES}$^*$ & SI  \\
GAFFlipid\cite{dickson12}       & POPC & 126  & 3948  & 303 & 137 & 32  & \cite{GAFFlipidFILES}$^*$ & SI \\
Lipid14\cite{dickson14}         & POPC  & 72 & 2234 & 303 & 100 & 50  & \cite{lipid14files}$^*$ & SI \\
Poger\cite{poger10}             & DPPC  & 128 & 5841 & 323 & 2$\times$100 & 2$\times$50 & \cite{pogerFILES} & SI \\
Slipid\cite{jambeck12}          & DPPC & 128 & 3840 & 323 & 150 & 100 & \cite{slipidsFILES}$^*$ & SI \\
Kukol\cite{kukol09}          & POPC   & 512 & 20564 & 298 & 50 & 30  & \cite{kukolFILES}$^*$ & SI \\
Chiu et al.\cite{chiu09} \todoi{Samuli, put to Zenodo}     & POPC  & 128 & 3552  & 298 & 56 & 50  & ? & SI \\
Rabinovich et al.\cite{rabinovich14}\todoi{Alexander Luybartsev, let us know if you share the files.}  & POPC   &  128 & 3840  & 303 & 100 & 80  & ? & \cite{rabinovich14}  \\
H\"ogberg et al.\cite{hogberg08}\todoi{Alexander Luybartsev, let us know if you share the files.}  & DMPC   &  98 & 3840  & 303 & 75 & 50 & ? & \cite{hogberg08} \\
Ulmschneider\cite{Ulmschneider09}    & POPC  & 128 & 3328 & 310 & 100 & 50 & \cite{ulmschneiderFILES}$^*$ & SI \\
Tj\"ornhammar et al.\cite{tjornhammar14}   & DPPC  & 144 & 7056 & 323 & 200 & 100 & \cite{tjornhammarfiles}$^*$ & \cite{tjornhammar14} \\
%\sout{OPLS-AA}\cite{rog09b}\todoi{We have this only with 150mM of NaCl delivered by Joona Tynkkynen. Options are to remove these results from this publicatio, run the simulations without ions (if not yet available), 
%  or include the results with ions. In my understanding this is a proto version of MacRog so it could be left out as well. However, for historical reasons and 
%to understand literature it might be useful to include also these results. COMMENT: Matti Javanainen suggested that we would remove these results. If there are no objections and the
%data without ions will not be delivered, we will remove this.}         &   &  &  & & & & & \\
CHARMM36-UA~\cite{henin08,lee14}     & DLPC   & 128  & 3840  & 323 & 30 & 20 & \cite{charmmUAfiles} & SI \\
\end{tabular}
\end{table*}

\begin{table*}[htb]
\centering
\caption{Simulated single component lipid bilayers with varying hydration levels. The simulation file data sets marked with $^*$ include also part of the trajectory.
}\label{systemsDEHYD}
\begin{tabular}{c c c c c c c c c c}
%\hline
Force field & lipid & \footnote{Water/lipid molar ratio} n (w/l)   & \footnote{The number of lipid molecules}N$_{\rm l}$   &  \footnote{The number of water molecules}N$_{\rm w}$ & \footnote{Simulation temperature}T (K)  & \footnote{The total simulation time}t$_{{\rm sim}}$(ns)  & \footnote{Time frames used in the analysis}t$_{{\rm anal}}$ (ns)& \footnote{Reference link for the downloadable simulation files}Files  &  \footnote{Reference for the full simulation details} Details\\
\hline
Berger-POPC-07~\cite{ollila07a}          &   POPC & 57  &128 & 7290  & 298  & 270 & 240 & \cite{bergerFILESpopc}$^*$ & SI \\
                                        &   POPC & 7  &128 & 896   & 298  & 60 & 50 & \cite{bergerDEHYDfiles}$^*$ & SI \\
CHARMM36\cite{klauda10}              & POPC   & 31 & 72  &  2242 & 303 & 30 & 20 & \cite{charmm36filesSHORT}$^*$ & SI \\
                               & POPC   & 15 & 72 &  1080  & 303 & 59 & 40 & \cite{charmm36files15wPERl}$^*$ & SI \\
                            & POPC   & 7  & 72  &  504  & 303 & 60 & 20 & \cite{charmm36files7wPERl}$^*$ & SI \\
MacRog\cite{maciejewski14}     & POPC   & 50 & 288  & 14400 & 310 & 90 & 40 & \cite{macrogdehydFILES}$^*$ & SI \\
                               & POPC   & 25 & 288  & 7200 & 310 & 100 & 50 & \cite{macrogdehydFILES}$^*$ & SI \\
                                & POPC   & 20 & 288  & 5760 & 310 & 100 & 50 & \cite{macrogdehydFILES}$^*$ & SI \\
                                & POPC   & 15 & 288  & 4320 & 310 & 100 & 50 & \cite{macrogdehydFILES}$^*$ & SI \\
                                & POPC   & 10 & 288  & 2880 & 310 & 100 & 50 & \cite{macrogdehydFILES}$^*$ & SI \\
                                & POPC   & 5 & 288   & 1440 & 310 & 100 & 50 & \cite{macrogdehydFILES}$^*$ & SI \\
GAFFlipid\cite{dickson12}      & POPC   & 31& 126  & 3948  & 303 & 137 & 32 & \cite{GAFFlipidFILES}$^*$ & SI  \\
                               & POPC   & 7& 126  & 896   & 303 & 130 & 40 & \cite{gaffDEHYDfiles}$^*$ & SI \\
\end{tabular}
\end{table*}

\begin{table*}[htb]
\centering
\caption{Simulated lipid bilayers with cholesterol. The simulation file data sets marked with $^*$ include also part of the trajectory.
}\label{systemsCHOL}
\begin{tabular}{c c c c c c c c c c}
%\hline
Force field & lipid  & \footnote{The number of lipid molecules}N$_{\rm l}$  & \footnote{number of cholesterol molecules}N$_{\rm chol}$   &  \footnote{The number of water molecules}N$_{\rm w}$ & \footnote{Simulation temperature}T (K)  & \footnote{The total simulation time}t$_{{\rm sim}}$(ns)  & \footnote{Time frames used in the analysis}t$_{{\rm anal}}$ (ns)\\
\hline
Berger-POPC-07~\cite{ollila07a}/H\"oltje-CHOL-13~\cite{holtje01,ferreira13}          &   POPC &128 & 0 & 7290  & 298  & 270 & 240 & \cite{bergerFILESpopc}$^*$ & \cite{ferreira15} \\
                               &   POPC &120 & 8 & 7290   & 298  & 100 & 80 & \cite{bergerFILESpopc7chol}$^*$ & \cite{ferreira13} \\
                               &   POPC &110 & 18 & 8481  & 298  & 100 & 80 & \cite{bergerFILESpopc15chol}$^*$ & \cite{ferreira13}  \\
                               &   POPC &84 & 44 & 6794   & 298  & 100 & 80 & \cite{bergerFILESpopc34chol}$^*$ & \cite{ferreira13} \\
                               &   POPC &64 & 64 & 10314  & 298  & 100 & 80 & \cite{bergerFILESpopc50chol}$^*$ & \cite{ferreira13} \\
                               &   POPC &50 & 78 & 5782   & 298  & 100 & 80 & \cite{bergerFILESpopc60chol}$^*$ & \cite{ferreira13} \\
CHARMM36\cite{klauda10,lim12}\todoi{Hubert Santuz, these files are already in Gitbub, however to get doi citation it should be put to Zenodo}    & POPC   & 128&  0 & 5120  & 303  & 150 & 100 & ? & SI  \\
 \todoi{Hubert Santuz, these files are already in Gitbub, however to get doi citation it should be put to Zenodo}                                & POPC   & 100  & 24  &  4960   & 303 & 200 & 100 & ? & SI \\
  \todoi{Hubert Santuz, these files are already in Gitbub, however to get doi citation it should be put to Zenodo}                               & POPC   & 80  & 80  &  4496    & 303 & 200 & 100 & ? & SI \\
MacRog\cite{maciejewski14}     & POPC   & 128 & 0  & 6400   & 310 & 400 & 200 & \cite{macrogCHOLfiles}$^*$ & SI \\ 
                          & POPC   & 114  & 14 & 6400  & 310  & 400 & 200 & \cite{macrogCHOLfiles}$^*$ & SI    \\
                          & POPC   & 72   & 56 & 6400  & 310  & 400 & 200 & \cite{macrogCHOLfiles}$^*$ & SI    \\
                             & POPC   & 64  & 64 & 6400  & 310  & 400 & 200 & \cite{macrogCHOLfiles}$^*$ & SI    \\
                             & POPC   & 56   & 72 & 6400  & 310  & 400 & 200 & \cite{macrogCHOLfiles}$^*$ & SI    \\
\end{tabular}
\end{table*}

%The initial structures for the simulations in low hydrated conditions are made by removing the
%water molecules from the fully hydrated system to achieve the targeted hydration conditions.

%For the initial structure of CHARMM36-~\cite{klauda10,lim12} and MacRog-~\cite{maciejewski14,kulig14} simulations with cholesterol,
%the required amount of lipid molecules were replaced with cholesterol

\section{Results and Discussion}

\subsection{Full hydration: Experimental order parameters for glycerol backbone and headgroup}\label{experiments}
The specific deuteration of $\alpha$-, $\beta$- and g$_3$- segments of dipalmitoylphosphatidylcholine (DPPC) has been successful, 
allowing the order parameter measurements for these segments by $^2$H--NMR~\cite{gally75,brown77,brown78,akutsu81}.
In addition, the order parameters for all glycerol backbone and choline headgroup segments in egg yolk lecithin~\cite{hong95a},
1,2-dimyristoyl-sn-glycero-3-phosphocholine (DMPC)~\cite{hong95b,gross97,dvinskikh05a}, 
1,2-dioleoyl-sn-glycero-3-phosphocholine (DOPC)~\cite{warschawski05} and POPC~\cite{warschawski05,ferreira13}
have been measured with several different implementations of $^{13}$C NMR experiments.
The experimental absolute values of glycerol backbone and choline order parameters from various publications are shown in Fig.~\ref{HGorderparameters}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{comparison3.pdf}
%\todo{DONE}\\
%The data in this figure should be updated to be consistent with the systems listed in Table~\ref{systems}. At least, Tjornhammar et al and CHARMM-UA values are missing. Some results in the figure are outdated due to longer simulations reported etc.} \\
\todo{Markus Miettinen suggested that we should consider making one more figure where only experimental data would be shown and that would be discussed in Section~\ref{experiments}} \\
\todo{There has been some discussion in the blog that maybe we should include the signs and stereospecific labeling here as well, and that maybe this figure should be clarified. My view is that for the main conclusion, i.e. CHARMM36, GAFFlipid and MacRog are the best here, this format is good enough. However, if someone wants to improve this and/or collect all the data with the signs and/or stereospecific labeling it would be even better. There is also very interesting discussion about interactive figure.}
  \caption{\label{HGorderparameters}
  Order parameteres from simulations listed in Table~\ref{systems} and experiments for glycerol and choline groups.
  The experimental values were taken from the following publications: DMPC 303K from \cite{gross97}, DMPC 314K from \cite{dvinskikh05a}, DPPC 310K from \cite{gally75}, 
  DPPC 323K from \cite{akutsu81}, and POPC 298K \cite{ferreira13}.
  The vertical bars shown for some of the computational values are not error bars, but demonstrate that for 
  these systems we had two data sets; the ends of the bars mark the values from the two sets, and the dot marks their measurement-time-weighted average. 
  The H{\"o}gberg et al. force field in Table~\ref{systems} is abbreviated in this figure as HoNiLy, Tj{\"o}rnhammar et al. as TjoEdh, CHARMM36-UA as C36UA,
  Rabinovic et al. as RabLyu and Ulmschneider as UlmUlm.
} 
\end{figure}

In general there is a good agreement between the order parameters measured with different experimental NMR techniques: Almost all the 
reported values are inside variation of $\pm$0.02 (which is also the error estimate given by Gross et al.~\cite{gross97}) 
for all fully hydrated PC bilayer, regardless of the variation in their acyl chain composition and the temperature.
Exception are the somewhat lower order parameters sometimes reported having been measured with $^{13}$C--NMR~\cite{hong95a,hong95b,warschawski05}.
These experiments have not seen in Fig.~\ref{HGorderparameters} as the reported error bars are either relatively large~\cite{hong95a,hong95b}, 
or the spectral resolution is quite low and the numerical lineshape simulations have not been used in the analysis~\cite{warschawski05}.
Therefore it is highly likely that these reported lower order parameters are due to lower experimental 
accuracy and that we exclude these values from our discussion. 
Motivated by the high experimental repeatability, we have highlighted in 
Fig.~\ref{HGorderparameters} subjective sweet spots (light blue areas), within which we expect the calculated absolute 
values of order parameters of a well-performing force field to fall.

%In addition to the phosphatidylcholine lipids, similar values of the glycerol backbone order parameters have been measured
%for phophatidylethanolamine (PE) and phopshatidylglycerol (PG) in E. Coli extract~\cite{gally81},
%indicating that the glycerol backbone structure is similar independent of the headgroup chemistry and lipid environment.
%Further, choline order parameters measured from mouse fibroblast L-M cell are similar to the ones in model
%membranes~\cite{scherer87}.

In addition to the numerical values, an important feature of the glycerol backbone is the 
inequality of order parameters for the two hydrogens attached to the same carbon in g$_1$ and g$_3$ segments,
while the hydrogens in choline $\alpha$ and $\beta$ segments give equal values.
Note that in this work we call the phenomena of inequal order parameters for hydrogens attached to the same carbon as ''forking'' to avoid 
confusion with dipolar and quadrupolar splitting in NMR terminology. Forking is also observed experimentally for the C$_2$ carbon in the sn-2
chain of all phosholipids, and it is known to arise from differently sampled orientations of the two C-H bonds, not from two 
different populations of lipid conformations~\cite{engel81}. The forking in glycerol backbone g$_3$ segment is small ($\approx$ 0.02) 
and some experiments only report the larger or the average value~\cite{akutsu81,ferreira13}. 
In contrast, forking is significant for the glycerol backbone g$_1$ segment, whose lower order parameter is close to zero and the
larger one has absolute values around 0.13-0.15. Forking was studied in detail by Gally et al.~\cite{gally81}, who used E. Coli to 
stereospecifically deuterate the different hydrogens attached to the g$_1$ or g$_3$ groups in PE lipids, and measured the order parameters from the lipid 
extract. This experiment gave the lower order parameter when deuterium was in the S position of g$_1$ or R position for g$_3$.
Since the glycerol backbone order parameters are very similar irrespective of the headgroup chemistry (PC,PE and PG) or lipid 
environment~\cite{gally81}, it is reasonable to assume that the stereospecifity measured for the PE lipids
holds also for the PC lipids.

In Fig.~\ref{HGorderparameters} we have shown the absolute values of order parameters as these are accessible
with both $^2$H NMR and $^{13}$C NMR techniques. However, $^{13}$C NMR techniques allow also the measurement of 
the sign of the order parameter~\cite{hong95a,hong95b,gross97}. The measured sign is negative for almost all the carbons 
discussed in this work, only $\alpha$ is positive~\cite{hong95a,hong95b,gross97}.

Combining the experimental information of the sign~\cite{hong95a,hong95b,gross97} and the stereospecifity 
measurements~\cite{gally81} with the absolute value measurements from various techniques~\cite{gally75,akutsu81,gross97,dvinskikh05a,ferreira13}
having high quantitative accuracy,
the most detailed experimentally available order parameter information for the glycerol backbone and choline segments of POPC is obtained.
These data are shown in Fig.~\ref{HGorderparameters2}.
\begin{figure*}[]
  \centering
  \includegraphics[width=8.6cm]{HGorderparameters5.eps} \\
  \todo{Markus Miettinen suggested that we should consider ``spreading'' the values similarly to the figure~\ref{HGorderparameters}} \\
  \caption{\label{HGorderparameters2}
  Order parameteres from simulations with Berger-POPC-07, CHARMM36, GAFFlipid and Macrog force fields together with experimental values for POPC glycerol and choline groups.
  The magnitudes for experimental order parameters are taken from Ferreira et al.~\cite{ferreira13}, the signs are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
  and Gross et al.~\cite{gross97}, and the R/S labeling is based in the measurements by Gally et al.~\cite{gally81}.
} 
\end{figure*}

\subsection{Full hydration: Comparison between simulation models and experiments}

The order parameters of the glycerol backbone and headgroup calculated from different force fields for various lipids has been 
previously compared to experiments~\cite{shinoda97,hogberg08,castro08,klauda10,kapla12,dickson12,poger12,ferreira13,chowdhary13,maciejewski14}. 
The general conclusion from these works seems to be that the CHARMM based~\cite{hogberg08,klauda10}, GAFFlipid~\cite{dickson12} and
MacRog~\cite{maciejewski14} force fields perform better for the glycerol backbone and headgroup structures than the Gromos based models~\cite{castro08,kapla12,poger12,ferreira13}.
However, none of the studies exploits the full potential of the available experimental data discussed in previous section; the quantitative accuracy, known signs and stereospecific labeling of
the experimental order parameters.

To get a general idea about the quality of the glycerol backbone and choline headgroup structures in different models, we calculated the absolute
values of the order parameters for these parts from fourteen different lipid models (Table~\ref{systems}) and 
plotted the results together with experimental values in Fig.~\ref{HGorderparameters}.
Two criteria were used to judge the quality of the model: {\bf forkings} there must not be significant forking in the $\alpha$ and $\beta$ carbons,
there must be only moderate forking in the g$_3$ carbon and there must be significant forking in the g$_1$ carbon, {\bf absolute values}
should be preferably inside to the subjective sweet spots determined from experiments (blue shaded regions in Fig.~\ref{HGorderparameters}).
None of the studied force fields fulfills these criteria, however, three force fields are closer than others: CHARMM36~\cite{klauda10}, MacRog~\cite{maciejewski14} and GAFFlipid~\cite{dickson12}.
The results for each force field in respect to the above criteria are summarized in Table~\ref{??} \todo{This kind of table should be done or not?}.

The top three models (CHARMM36, MacRog and GAFFlipid) together with the most used lipid model (Berger based model) 
were subjected to a more careful comparison including the signs and the sterospecific labeling in Fig.~\ref{HGorderparameters2}.
The essential additional information given by this comparison is that the sign of the $\beta$ carbon order parameter is correct only in CHARMM36 model.

\subsection{Full hydration: Atomistic resolution structures in different models}

The results in the previous section revealed significant differences of the glycerol backbone and choline headgroup
order parameters between different molecular dynamics simulation models.
However, it is not straightforward to conclude which kind of structural differences (if any)
between the models the results indicate, because the mapping from the order parameters to the 
structure is not unique. In this section we demonstrate that 1) the differences in order parameters
indicate significantly different structural sampling strongly correlating with the dihedral angles of the related bonds,
and that 2) the comparison between experimental and simulated order parameters can be used to exclude
nonrealistic structural samping in molecular dynamics simulations. The demonstration is done for 
the dihedral angles defined by the g$_3$-g$_2$-g$_1$-O(sn-1) segments in the glycerol backbone and 
the N-$\beta$-$\alpha$-O segments in the headgroup. These dihedrals were chosen for demonstration, because 
significant differences between the models are observed around these segments in Fig.~\ref{HGorderparameters2}.
We note that performing a similar comparison through all the dihedrals in all the 14 models would probably give highly useful
information to improve the accuracy of simulations, however this is beyond the scope of the current report.

The dihedral angle distributions for the  g$_3$-g$_2$-g$_1$-O(sn-1) dihedral calculated from different models are
shown in Fig.~\ref{dihDISTS}. The distribution is qualitatively different for the Berger-POPC-07 model, showing a maximum in 
the gaughe$^+$-conformation (60$^o$) compared to all the other models showing a maximum in the trans-conformation (180$^o$).
The distributions in all the other models have the same general features, the main difference being that the
fraction of configurations in gaughe$^-$-conformation (-60$^o$) is zero for the MacRog, detectable for the CHARMM36 and
equally large to the gaughe$^+$ fraction in GAFFlipid. From the results we conclude that most likely the wrongly sampled
dihedral angle for the g$_2$-g$_1$ bond explains the significant discrepancy to the experimental order parameters
for the g$_1$ segment in the Berger-POPC-07 model (Fig.~\ref{HGorderparameters2}). 
The result that models preferring the trans conformation for this dihedral gives more realistic order parameters
is in agreement with previous crystal structure and $^1$H NMR studies~\cite{hauser80,hauser81,hauser81b,hauser88,pascher92,marsh06}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{g1-g2_Cdihs2.eps}
  \caption{\label{dihDISTS}
    Dihedral angle distributions for g$_3$-g$_2$-g$_1$-O(sn-1) dihedral from different models (POPC bilayer in full hydration).
      } 
\end{figure}

The dihedral angle distribution for the  N-$\beta$-$\alpha$-O dihedral calculated from the same four models is 
shown in Fig.~\ref{dihDISTS2}. Also for this dihedral there are significant differences in the gauche-trans fractions.
The gaughe conformations are dominant in the CHARMM36, in MacRog there are only trans conformations present,
and in the Berger-POPC-07 and GAFFlipid gaughe and trans conformation has equal probability. 
On the other hand comparison of $\alpha$ and $\beta$ order parameters in Fig.~\ref{HGorderparameters2}
reveals that the CHARMM36 is closest to the experimental results being the only model having correct
sign (negative) for the $\beta$ order parameter. This result is again in agreement with previous 
crystal structure, $^1$H NMR and Raman spectroscopy studies~\cite{hauser80,hauser81,hauser81b,akutsu81b} which suggest that
this dihedral has only gaughe conformation in the absense of ions.

%Interestingly, the probability of the gaughe conformations correlates with the order parameter difference between the $\beta$ and $\alpha$ segments:
%the larger the gaughe fraction the larger the order parameter difference. This suggestion together with the results in 
%Fig.~\ref{HGorderparameters2} would indicate that the correct gaughe-trans fraction for N-$\beta$-$\alpha$-O dihedral is 
%larger than in GAFFlipid but smaller than in CHARMM36. This happens to be quite close to the gaughe-trans fraction in
%the Berger model. However, there is significant forking and numerical values are off from the experiments in the Berger
%model suggesting that it has some other inaccuracies.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{a-bDIHS2.eps}
  \caption{\label{dihDISTS2}
    Dihedral angle distributions for N-$\beta$-$\alpha$-O dihedral from different models (POPC bilayer in full hydration).
  } 
\end{figure}

The used examples show that the glycerol backbone and headgroup order parameters reflect the atomistic resolution structure
and that the comparison with experiments allows the assesment of the quality of the suggested structure. We were able to pinpoint
specific problems in the structures in different models and suggest potential improvement strategies.
If the improved atomistic resolution
molecular dynamics simulation model would reproduce the order parameters and other experimental observables (like chemical shift anisotropy)
with experimental accuracy, it would give an interpretation for the atomistic resolution structure of the glycerol backbone and 
choline~\cite{seelig77b,skarjune79,jacobs80,davis83,akutsu91,hong95b,semchyschyn04}. The research along these lines is left, however,
for future studies.

\subsection{Response to dehydration and cholesterol content}
In addition to pure phosphatidylcholine bilayers at full hydration, the choline headgroup order parameters
have been measured under various different conditions~\cite{gally75,brown77,brown78,akutsu81,altenbach84,scherer89,bechinger91,ulrich94,dvinskikh05b,castro08,kapla12,ferreira13}.
Also the order parameters for the glycerol backbone have been measured with $^{13}$C NMR in dehydrated conditions~\cite{dvinskikh05b}, and as a function 
of anesthetics~\cite{castro08} and glycolipids~\cite{kapla12} for DMPC, and as a function of cholesterol 
concentration for POPC~\cite{ferreira13}. Due to the high resolution in the NMR (especially $^2$H NMR) experiments,
even very small order parameter changes resulting from the varying conditions can be measured (see
http://nmrlipids.blogspot.fi/2014/02/accuracy-of-order-parameter-measurements.html
for more discussion.
\todo{This is temporarily linked to the blog. Currently I think that the posts (with comments included) to which we want to cite are put into 
the Figshare in pdf format as they are in when the publication is submitted. Figshare allows commenting to continue the discussion.}
However, as already discussed above, it is not simple to deduce 
the structural changes from order parameter changes~\cite{akutsu91,semchyschyn04}. Consequently, comparison of the order parameters
between simulations and experiments in different conditions can be used measure the quality of the force field 
in different situations, and, if the quality is good, to potentially interpret the structural changes in experiments.
Here we exemplify such comparison for a lipid bilayer under low hydration levels and mixed with cholesterol. 
The interaction between ions and phosphatidylcholine bilayer is discussed in a separate work~\cite{ionpaper}.

\subsubsection{Phospholipid bilayer with low hydration level}
The experimental order parameters available in the literature~\cite{dvinskikh05b,ulrich94,bechinger91} 
for the glycerol backbone and choline as a function of hydration level are shown in Fig.~\ref{ordPhydr}. 
The independently reported values for choline segments are in good agreement with each other (despite of 
slight differences in temperature and acyl chain composition),
showing increase for both segments with decreasing hydration level. It should be noted that only 
absolute values were measured in the original experiments~\cite{dvinskikh05b,ulrich94,bechinger91}, but
we have included the signs measured separately~\cite{hong95a,hong95b,gross97}. 
Consequently, the $\beta$ order parameter with negative sign actually increases with dehydration 
since the absolute value decreases~\cite{dvinskikh05b,ulrich94,bechinger91}.
Slight decrease for the glycerol backbone g$_3$- and g$_2$- order parameters were observed with dehydration, 
while g$_1$ remained practically unchanged~\cite{dvinskikh05b}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{OrderParameterDEHYDpresentation.eps}
\todo{Several people think that this plot should be clarified.}
  \caption{\label{ordPhydr}
    The effect of dehydration on glycerol and choline order parameters in experiments.
    The magnitudes of order parameters are measured for DMPC ($^{13}$C NMR) at 314~K~\cite{dvinskikh05b}, 
    for POPC ($^2$H NMR) at 296~K~\cite{bechinger91} and for DOPC ($^2$H NMR) at 303~K~\cite{ulrich94}. 
    The signs are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
    and Gross et al.~\cite{gross97}.
  }
\end{figure}

Lipid bilayer dehydration has been studied also with molecular dynamics simulations~\cite{mashl01,pertsin05,pertsin07,eun09,eun10,schneck12},
typically motivated by the  discussion about the origin of the ``hydration repulsion''~\cite{israelachvili,israelachvili96,sparr11}.
However, the used simulation models are not typically compared to the experimental choline and glycerol backbone
order parameters (except by Mashl et al.~\cite{mashl01}).
In Fig.~\ref{ordPhydr} the glycerol backbone and choline order parameters as a function of hydration level are shown 
for the CHARMM36, MacRog and GAFFlipid models (having the most realistic atomistic resolution structures) together with the Berger based model 
(which is the most used lipid model). The choline order parameters increase with dehydration in all simulation
models, in qualitative agreement with experiments. 
The measured decrease in both g$_3$ and g$_2$ order parameters with dehydration is reproduced only in CHARMM36.

The qualitative agreement with experiments in all simulation models for the $\alpha$ and $\beta$ order parameters  
as a function of hydration indicates that the structural response of the choline headgroup to dehydration is somewhat realistic
despite the unrealistic structures at full hydration. 
The most likely explanation is that the choline group
orients more parallel to the membrane plane with dehydration due to restricted interlamellar space. 
Indeed, the P-N angle vector angle with membrane normal as a function of dehydration shows an increase for
all models as a function of dehydration in Fig.~\ref{PNangle}.
However, the qualitative agreement in the lipid response to dehydration does not guarantee the correct 
free energy landscape if the simulation model has incorrect structure. The influence of this issue to 
dehydration energetics studied with simulations~\cite{eun09,schneck12} is left for future studies.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{PNangles.eps}

  \caption{\label{PNangle}
    The angle between membrane normal and P-N vector in choline segment as function of
    hydration level calculated from different simulations.
  }
\end{figure}

The response of the glycerol backbone to dehydration seems to be more subtle than of the choline headgroup 
as CHARMM36 is the only model that reproduces the decrease in g$_2$ and g$_3$ segments.

\subsubsection{Phospholipid bilayer mixed with cholesterol}
Phospholipid--cholesterol interactions have been widely studied with theoretical~\cite{huang99,zhu07,rog09,alwarawrah12} and
experimental methods~\cite{brown78,marsh10,ferreira13,marsh13}, since cholesterol is abundant in biological membranes and
it has been suggested to be an important player, for example, in domain formation~\cite{simons04,somerharju09}.
It is widely agreed that cholesterol orders lipid acyl tails thus decreasing the area per molecule (condensing effect),
however, the influence of cholesterol on the lipid headgroup and glycerol backbone are sill debated~\cite{huang99,simons04,somerharju09}.
For example, it has been suggested that the surrounding lipids shield cholesterol from interactions with water by 
reorienting their headgroups (``umbrella model'')~\cite{huang99} or that cholesterol acts as a spacer for the headgroups thus increasing 
their entropy and dynamics (``superlattice model'')~\cite{somerharju09}. 
\todo{It has been suggested that we should use direct quotations from original papers here for clarity.}
Both of these suggestions have been supported
by molecular dynamics simulations~\cite{zhu07,alwarawrah12}, and other simulations suggest specific
interactions between the glycerol backbone and cholesterol~\cite{rog09}, however the glycerol backbone and choline headgroup behaviour
as a function of cholesterol content is not compared to experiments in these studies.

The choline headgroup and glycerol backbone order parameters for POPC measured by $^{13}$C NMR~\cite{ferreira13} and DPPC choline order parameters 
measured by $^{2}$H NMR~\cite{brown78} are shown in Fig.~\ref{ordPchol} as a function of cholesterol content.
The agreement between different experimental results is again very good, showing only very modest changes in 
the choline order parameters as a function of cholesterol content. It should be noted, however, that very small
changes are measurable with high resolution $^{2}$H NMR experiments
and cholesterol causes a measurable increase in the $\beta$ order parameter and a forking in the $\alpha$ order
parameter~\cite{brown78}, but these effects are so small that they are barely visible in the scale used in Fig.~\ref{ordPchol}.
Further, the effects of cholesterol on the glycerol backbone order parameters for POPC from $^{13}$C NMR experiment~\cite{ferreira13} 
is in good agreement with the results for the phosphatidylethanolamine (PE) measured by $^{2}$H NMR~\cite{ghosh82}.
These results further support the idea that the glycerol backbone structural behaviour is independent of the
headgroup composition~\cite{gally81} and that the headgroup stucture is independent of the acyl chain region content unless
charges are present~\cite{scherer87}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{OrderParameterCHOLexpVSsim.eps}
  \caption{\label{ordPchol}
    The effect of cholesterol content on the glycerol backbone and choline order parameters in experiments~\cite{brown78,ferreira13} and simulations
    with the Berger-POPC-07/H\"oltje-CHOL-13, CHARMM36 and MacRog force fields. The signs in the experimental values are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
    and Gross et al.~\cite{gross97}.  The most order parameters from Berger-POPC-07/H\"oltje-CHOL-13 model for g$_1$ are beoynd the y-axis scale.}
\end{figure}

In addition to the experimental data, the previously published simulation results from the Berger-POPC-07/H\"oltje-CHOL-13 model~\cite{ferreira13},
and our results from CHARMM36 and MacRog force fields  
are shown in Fig.~\ref{ordPchol}. As already pointed out previosly, the Berger-POPC-07/H\"oltje-CHOL-13 model
seriously overestimates the effect of cholesterol on the phospholipid glycerol backbone and choline segments~\cite{ferreira13}.
In contrast, the responses of both CHARMM36 and MacRog are in better agreement
with experiments, however CHARMM36 seems to better reproduce the experimentally observed modest changes in the glycerol backbone segments 
g$_2$ and g$_3$ with high concetrations of cholesterol. Thus we have calculated the glycerol backbone dihedral angle distributions
as a function of cholesterol in CHARMM36 (shown in Supplementary material) to resolve the cholesterol induced structural changes. The only detectable change due to the
addition of cholesterol is the small decrease of gaughe- and increase of gaughe+ probability of g3-g2-g1-O(sn-1) dihedral.

It should be noted that the CHARMM36 force field parameters (dihedral potentials) for the glycerol backbone have been tuned 
the dihedral potentials to reproduce the correct order parameters at fully hydrated conditions~\cite{klauda10}. 
This procedure contains a risk of overfitting, which would manifest itself as wrong responses to changing conditions. 
According to our results, tuning seems not to lead to overfitting problems in the case of dehydration or lipid-cholesterol mixtures.

\section{Conclusions}
The atomistic resolution structures sampled by the glycerol backbone and choline headgroup
in phoshatidylcholine bilayers are not known despite of vast amount of accurate experimental
data. Atomistic resolution molecular dynamics simulation model which would reproduce the
experimental data would automatically resolve the structures giving an interpretation of experimental results.
In this work we have collected and reviewed the experimental C-H bond vector order
parameters available in literature. These experimental parameters are then compared to
different atomistic resolution simulation models for fully hydrated bilayer, dehydrated bilayer and
lipid bilayer containing cholesterol. Our results have led to the following conclusions:

- The C-H bond order parameters measured with different NMR techniques are in good agreement
with each others. By combining the experimental results from various sources we concluded
that the order parameters for each C-H bond are known with quantitative accuracy of $\pm$0.02.

- None of the tested models (14 different models) produces the order parameters with the experimental
accuracy for fully hydrated phoshatidylcholine lipid bilayer. However, the CHARMM36, GAFFlipid and MacRog 
models are relatively close. The structures of these models together with the most used lipid model (Berger based) 
were subjected to more careful studies. The results revealed that the current models are not accurate
enough to resolve the atomistic resolution structures sampled by glycerol backbone and choline headgroup.  
However, the correlation between dihedral angle distributions and order parameter differences was found, 
suggesting that careful adjustment of dihedral potentials would potentially lead to the model with correct
structure.

- Independent of the accuracy for fully hydrated lipid bilayer, all the models reproduced the choline response
to the dehydration. This can be explained by the change in the P-N vector tilting more parallel to the membrane
which leads to the increase of order parameters despite of the initial configuration. It should be however noted
that the correct qualitative response do not necessarily indicate correct energetics.

- The response of glycerol backbone and choline headgroup to the cholesterol content is described more
realistically in CHARMM36 and MacRog models than in the Berger based model.

In general, we conclude that the atomistic resolution classical molecular dynamics simulations 
is extremely convenient tool to give structural interpretation for the high resolution NMR data~\cite{ferreira14}. 
However, in the case of phoshatidylcholine glycerol backbone and choline headgroup there is some
further model development required.

This work has been, and continues to be, progressed and discussed through the blog: nmrlipids.blogspot.fi. 
Everyone is invited to join the discussion and make contributions through the blog. 
The manuscript will be eventually submitted to an appropriate scientific journal. 
Everyone who has contributed to the work through the blog will be offered 
coauthorship. For more details see: nmrlipids.blogspot.fi.

{\bf Acknowledgements: }
OHSO acknowledges Tiago Ferreira and Paavo Kinnunen for useful discussions, the Emil Aaltonen foundation for financial support, Aalto Science-IT project and CSC-IT Center for Science for computational resources.

\bibliographystyle{apsrev}
\bibliography{refs}

\newpage

\appendix
\begin{center}
{\bf SUPPLEMENTARY INFORMATION}
\end{center}
\subsection{Simulation details} 
%\todo{When I was adding the simulation details to the manuscript, I realized that the section was getting very long and
%it would do the paper more difficult to read. I suggest that we do this in the following way: \\
%- For the simulation data which are taken directly from previous publications, we only cite the original publication
%in the tables where the systems are described. This would be the case, for example, for Berger POPC/chol, DMPC, Rabinovich et al. and Hogberg et al. \\
%- For the simulations which are reproduced using mdp file already delivered by original authors we can cite that, thus we do not have to copy the 
%simulation detail section from the original publication. This is the case, for example, for Tjornhammar et al. \\
%- When simulations are done specifically for this project and there is something non-trivial we
%put the details in the supplementary material. This is the case, for example, for GAFFlipid, Lipid14 and poger et al. \\
%- In all cases all simulation files, also trajectory, would be ideally shared in Zenodo.
%}
\subsubsection{Berger based models}
For the berger based models we use here the following naming convention: 
Berger - \{{\it molecule name}\} - \{{\it year when model published first time}\} \{{\it citation}\}.
The reason is that there are several different molecular topologies which are using the non-bonded parameters originally
developed by Berger et al.~\cite{berger97}. Thus the common factor in the berger based models are the non-bonded parameters,
while the molecule specific parameters might somewhat vary. However, the majority of the molecular level topologies are 
relying (especially for the glycerol backbone and headgroup) on the parameters originally introduced by Marrink et al.~\cite{marrink98}.
This is the case for all the Berger based simulations discussed in this work.

POPC simulations at full hydration in 298 K and simulations as a function of cholesterol are the same as previous publications~\cite{ferreira13,ferreira15}.
In these simulation the POPC parameters introduced by Ollila et al~\cite{ollila07a} are used, which are using Berger non-bonded parameters~\cite{berger97}
and molecular topology is from Tieleman et al.~\cite{tieleman99} with improved double bond dihedrals by Bachar et al.~\cite{bachar04}. 
Thus they are called Berger-POPC-07~\cite{ollila07a}. The cholesterol model is based on the parameters by H\"oltje et al.~\cite{holtje01} with the
exeption that the atom types were changed from CH2/CH3 to LP2/LP3 to avoid overcondensation of the bilayer as suggested in ref.~\cite{tieleman06}.
Since this modification was introduced by Ferreira et al.~\cite{ferreira13}, we call the used cholesterol model as H\"oltje-CHOL-13~\cite{ferreira13}.

For the POPC at 323~K and POPC in low hydration the same force field parameters are used.
For DPPC the implementation of Berger parameters~\cite{berger97} by Peter Tieleman et al. are used~\cite{marrink98}.
For all of these simulations the timestep of 2~fs was used with leap-frog integrator. Covalent bond lengths were constrained with LINCS algorithm~\cite{hess97,hess07}. 
Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with real space cut-off at 1.0~nm was used 
for electrostatics. Plain cut-off was used for the Lennard-Jones interactions with a 1.0~nm cut-off.
The neighbour lists were updated every 5th step with cut-off at 1.0~nm. Temperature was coupled separately
for lipids and water to 298~K with the velocity-rescale method~\cite{bussi07} with coupling constant 0.1~ps.
Pressure was semi-isotropically coupled to the atmospheric pressure with the Berendsen method~\cite{berendsen84}.

%The results for the POPC/cholesterol systems with the Berger lipid~\cite{berger97} and modified H\"oltje
%cholesterol model~\cite{holtje01} were taken directly from Ref.~\cite{ferreira13}.

\subsubsection{CHARMM36}

{\it DPPC}
\todo{Markus Miettinen made the files.}

Timestep of 1~fs was used with leap-frog integrator. Covalent bonds with hydrogens were constrained with LINCS algorithm~\cite{hess97,hess07}. 
Coordinates were written every 5~ps. PME~\cite{darden93,essman95} with real space cut-off at 1.4~nm was used 
for electrostatics. Lennard-Jones interactions were switched to zero between 0.8~nm and 1.2~nm.
The neighbour lists were updated every 5th step with cut-off 1.4~nm. Temperature was coupled separately
for lipids and water to 303~K with the velocity-rescale method~\cite{bussi07} with coupling constant 0.2~ps.
Pressure was semi-isotropically coupled to the atmospheric pressure with the Berendsen method~\cite{berendsen84}.

{\it POPC}
The starting structures for the pure POPC and DOPC simulations was taken from the Slipids~\cite{jambeck12b} website (http://people.su.se/$\sim$jjm/Stockholm\_Lipids/Downloads.html).
The starting structures for mixed POPC/Cholesterol simulations were constructed with the ChARMM-GUI website~\cite{jo08}. 
They contained 100 POPC/24 Cholesterol molecules and 80 POPC/80 Cholesterol molecules for
the simulations of 20\% Cholesterol and 50\% Cholesterol respectively. The TIP3P water model~\cite{jorgensen83} was used to
solvate the system.
The publicly available CHARMM36 forcefield parameters (http://www.gromacs.org/@api/deki/files/184/=charmm36.ff\\\_4.5.4\_ref.tgz) 
by Piggot et al. \cite{piggot12} were used. Cholesterol parameters came
from Lim et al. \cite{lim12} and were converted into Gromacs files by using the PyTopol tool (https://github.com/resal81/PyTopol).  
Single point energy calculation was done to assess the conversion. 
%(At the time I've made the simulations, the charmm36 ff for cholesterol in gromacs format was not available at
%http://mackerell.umaryland.edu/charmm_ff.shtml#gromacs).
Simulations were performed for 200ns and the last 100ns was used for the calculations. Timestep of 2fs was
used with leap-frog integrator. All bond lengths were constrained with LINCS~\cite{hess97,hess07}. Temperature was maintened at
303 K with the velocity-rescale method~\cite{bussi07} and a time constant of 0.2 ps. Pressure was maintained semiisotropically
at 1 bar using the Parrinello-Rahman algorithm~\cite{parrinello81} with a time constant of 1.0 ps. The neighbour list
was updated every 10th step with a cut-off of 1.2 nm. Lennard-Jones interactions were switched to zero
between 0.8 nm and 1.2 nm. PME~\cite{darden93,essman95} with real space cut-off at 1.2nm was used for electrostatics.

\subsubsection{MacRog}
The lipid force field parameters were obtained from the developers and they correspond to the published DPPC parameters~\cite{maciejewski14} with the inclusion of the 
double bond parameters. This inclusion of unsaturated lipid tails will be published in the near future. 
\todo{You have recent paper where you use POPC/chol simulations with this model: Kulig et al. BBA 1848 (2015) 422-432 http://dx.doi.org/10.1016/j.bbamem.2014.10.032.
would this be a correct refence here?}
A bilayer with 288 POPC lipids was hydrated with 12600 TIP3P water~\cite{jorgensen83} molecules ($\sim$44/lipid) and simulated for 100~ns with a time step of 2~fs. Data was saved 
every 10~ps and the first 20~ns of the trajectory was discarded from the analysis. 
All bond lengths were constrained with LINCS~\cite{hess97,hess07}. The temperatures of the lipids and the solvent were separately coupled to the Nos\'{e}--Hoover thermostat~\cite{nose84,hoover85} 
with a target temperature of 310~K and a time constant of 0.4~ps. Semi-isotropical pressure coupling to 1~bar was obtained with the Parrinello--Rahman 
barostat~\cite{parrinello81} with a time constant of 1~ps. PME~\cite{darden93,essman95} was employed to calculate the long-range electrostatic interactions. Lennard-Jones interactions were cut off 
at 1~nm and the dispersion correction was applied to both energy and pressure. A neighbour list with a radius of 1~nm was updated every step.

Identical parameters were employed for both full hydration and for the dehydration simulations. The dehydration simulations were also run for 100~ns 
with data saved every 10~ps.

The initial structures for the simulations with 10, 40, 50 and 60 mol\% of cholesterol were obtained by replacing 14, 56, 64 or 72 POPC molecules 
with cholesterol molecules in the initial structure containing 128 POPC molecules. These systems were simulated for 400 ns and the first 200 ns was 
discarded from analysis. Data was saved every 100 ps.

\subsubsection{GAFFLipid}
The initial structure in Lipidbook \cite{domanski10} had different glycerol backbone isomers in different leaflets. 
To generate the initial structure we took the structure delivered by Slipid developers~\cite{jambeck12b}. Also this structure
had one lipid with different glycerol bakcbone isomer. This lipid and one lipid from opposite leaflet were removed
after the system was equilibrated.

The force field parameters were generated using files obtained from the Lipidbook website (http://lipidbook.bioch.ox.ac.uk/package/show/id/150.html)~\cite{domanski10}. 
The conversion to Gromacs compatible formats was performed using the acpype tool~\cite{silva12}. The accuracy of the conversion was checked by calculating 
the total energy of a single POPC lipid molecule using the sander program which is part of the AmberTools14 package~\cite{ferrer13} and Gromacs 4.6.5. 
A difference of 0.002 kcal/mol was obtained between the two programs.

Timestep of 2~fs was used in Langevin dynamics with zero friction term and collision frequency of 1.0~ps$^{-1}$. 
Covalent bonds with hydrogens were constrained with LINCS algorithm~\cite{hess97,hess07}.
Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with real space cut-off at 1.0~nm was used 
for electrostatics. Plain cut-off with 1~nm was used for Lennard-Jones interactions. 
The neighbour lists were updated every 5th step with cut-off 1.0~nm. 
Pressure was semi-isotropically coupled to the 1~bar pressure with the Berendsen method~\cite{berendsen84}.
\subsubsection{Lipid14}
%The used simulation files and part of the trajectroy are available~\cite{lipid14files}.
The initial structure was taken directly from the Lipidbook~\cite{domanski10}.
The Amber compatible force field parameters were generated using the tleap program which is integrated in the AmberTools14 package~\cite{ferrer13}. 
A workflow similar to the one used previously for the conversion and validation of the GAFFLipid parameters was followed here. 
As before, a negligible energy difference of 0.003 kcal/mol was obtained between the two programs.

Timestep of 2~fs was used in Langevin dynamics with zero friction term and collision frequency of 1.0~ps$^{-1}$. 
Covalent bonds with hydrogens were constrained with LINCS algorithm~\cite{hess97,hess07}.
Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with real space cut-off at 1.0~nm was used 
for electrostatics. Plain cut-off with 1~nm was used for Lennart-Jones interactions. Dispersion correction
was used for energy and pressure. The neighbour lists were updated every 5th step with cut-off 1.0~nm. 
Pressure was semi-isotropically coupled to the 1~bar pressure with the Berendsen method~\cite{berendsen84}.

\subsubsection{Poger et al.}
The Poger lipids are derived from GROMOS G53A6~\cite{poger10} and were initially coined 53A6-L (L for lipids), and are now part of GROMOS G54A7~\cite{poger12}. 
They work with the SPC water model~\cite{berendsen81}. The initial hydrated bilayer structure of 128 DPPC/5841 water molecules as well as force field parameters were downloaded 
from David Poger's web site (http://compbio.chemistry.uq.edu.au/~david/) on April 2012. 
We noticed that the same files downloaded in October 2013 appear to lack two dihedral angles in the choline headgroup (only one dihedral of type gd\_29 allowing 
the rotation of the 3 choline methyls) compared to the April 2012 version (3 dihedrals of type gd\_29 for the 3 choline methyls). This should not affect the 
bilayer structure and only change the kinetics of the choline methyls rotation, however the October 2013 version has not been tested.

MD Simulations (two repetitions with independent initial velocities) were run for 100 ns using a 2 fs time step and the analysis 
was performed on the last 50 ns. Coordinates were saved every 50 ps for analysis. All bond lengths were constrained with the LINCS algorithm~\cite{hess97,hess07}. Temperature was kept 
at 323K with the v-rescale~\cite{bussi07} thermostat with a time constant of 0.1 ps (DPPC and water coupled separetly). Pressure was maintained semi-isotropically at 1 bar using 
the Parrinello-Rahman barostat~\cite{parrinello81} using a 4 ps time constant and a compressibility of 4.5e-5 bar$^{-1}$. For non-bonded interactions, two conditions were tested:
i) A 0.8-1.4 nm twin-range cutoff with the neighbor list updated every 5 steps for both electrostatics and Lennard-Jones. For the former the generalized reaction 
field (RF) with a dielectric permitivity of 62 was used beyond the 1.4 nm cutoff~\cite{tironi95}. This is the original setup that Poger et al.~\cite{poger10} used.
ii) PME~\cite{darden93,essman95} electrostatics with a real space cutoff of 1.0 nm, a Fourier spacing of 0.12 nm and an interpolation order of 4, LJ computed with a 1.0-1.4 nm twin-range cutoff, 
neighbor list updated every 5 steps. Note that Poger and Mark tested the effect of PME vs RF in the ref.~\cite{poger12}, but used a 1.0 nm cutoff with PME and 1.4 nm with RF for LJ 
interactions. Since 0.8-1.4 nm twin-range cutoff for LJ interations is used in the parametrization of GROMOS force field we decided to use that
also in the simulations with PME.
%Since there is no dispersion correction in GROMOS (and since dispersion correction should not be used anyway for systems with interfaces), we decided to stick 
%to the same LJ cutoff as that with the RF scheme i), that is 1.4 nm.

Since Poger lipids come from GROMOS force field, it is important to note that GROMOS uses the RF scheme for computing electrostatics (this is the method used for the 
force field parameterization). Using setup i) based on RF, we were able to reproduce the results (i.e. area per lipid~0.63 nm$^2$) from the original work only with 
GROMACS version $<=$ 4.0.* (the original authors~\cite{poger10} used GROMACS version 3.3.3). On going to versions $>=$ 4.5.*, the area per lipid dropped below 0.58 nm$^2$. 
The GROMACS developers were contacted and a redmine issue opened (http://redmine.gromacs.org/issues/1400). The difference comes from the new Trotter decomposition 
introduced in version 4.5. A fix has been introduced in version 4.6.6 that allows a recovery of ~0.615 nm$^2$. The results in terms of area per lipid using the different 
GROMACS versions are here~\cite{fuchs14}.
% http://www.dsimb.inserm.fr/~fuchs/project\_Samuli/Poger\_DPPC/tests\_gmx\_versions/)[Samuli, you may want to add this Figure+Legend+Comment 
%o a stable link (Figshare?), let me know]. 
%Of note, some tests (Miguel Machuqueiro, personnal communication) with the same RF scheme i) on simulations with proteins 
%(thus with explicit hydrogens) showed that this setup is still unstable and eventually leads to crashes with versions 4.6.6 and 4.6.7 (and so does version 4.5.*).
Thus we decided to use only the PME setup ii) for computing the order parameter since it gives stable results whatever the GROMACS version. We obtained an area per 
lipid of ~0.615 nm$^2$, below 0.648 nm$^2$ found by the original authors with their PME setup (see~\cite{poger12}). We explained that by the fact we used 
a 1.4 nm for the LJ cutoff and they used 1.0 nm.

\subsubsection{Slipid}
%The used simulation files and part of the trajectroy are available~\cite{slipidsFILES}.

Initial coordinates for a hydrated DPPC bilayer (30 waters/lipid) at 323K were taken directly
from http://people.su.se/\~jjm/Stockholm\_Lipids/Downloads.html  The Slipids force field~\cite{jambeck12} was used for the the all atom description of DPPC, and
water was described with the TIP3P water model~\cite{jorgensen83}. Simulations were performed within the NPT ensemble using the GROMACS 4.6.1 simulation
package~\cite{hess08}. The nose--hoover rescaling thermostat~\cite{nose84,hoover85} was used with reference temperature of
323 K and a relaxation time constant of 0.5 ps. Water and lipids were coupled separately to
the heat bath. Pressure was kept constant at 1.013 bar using a semi--isotropic Parinello--Rahman
barostat~\cite{parrinello81} with a time constant of 10.0 ps. Equations of motion were
integrated with the leapfrog algorithm using a timestep of 2 fs. Long range
electrostatic interactions were calculated using the PME method~\cite{darden93,essman95}, with a fourth order
smoothing spline. A real space cutoff of 1.0 nm was employed with grid spacing of 0.12 in the reciprocal space.
Lennard--Jones potentials were cutoff at 1.4 nm, with dispersion correction applied both to
energy and pressure. All covalent bonds in lipids were constrained using the LINCS algorithm~\cite{hess97}, 
whereas water molecules were constrained using SETTLE~\cite{miyamoto92}. Twinrange cutoffs,
1.0 nm and 1.6 nm, were used for the neighborlists with the longrange neighbor list updated every
10 steps. This simulation protocol corresponds to the protocol used in Ref~\cite{jambeck13}. 
%The simulations were run for XXX ns with the last XXX ns used for calculating the order parameters. 
\subsubsection{Kukol}
A bilayer patch with 512 POPC lipids was constructed and hydrated with $\sim$40 SPC water molecules per lipid. 
The force field parameters were obtained from Lipidbook \cite{domanski10}.
This bilayer was simulated with a 2~fs time step for a total of 50~ns and coordinates were saved every 100~ps. 
All bonds were constrained with LINCS~\cite{hess97,hess07}. PME~\cite{darden93,essman95} was employed for the long-range electrostatics. Lennard-Jones interactions 
were cut off at 1.4~nm. A neighbour list with a radius of 0.8~nm was updated every 5~steps. The constant temperature of 298~K 
was maintained with the Berendsen thermostat \cite{berendsen84} with a time constant of 0.1~ps. The Berendsen barostat \cite{berendsen84} 
was employed for semi-isotropical pressure coupling at 1~bar.

%Additionally, the reaction field electrostatics with a twin-range cut-off scheme [cite GROMOS] was employed yet the results 
%were consistent with those obtained with PME. (This paragraph can be discarded.)

\subsubsection{Chiu et al.}
The force field parameters and the initial configuration were available through the Lipidbook~\cite{domanski10}.
Timestep of 2~fs was used with leap-frog integrator. Covalent bond lengths were constrained with LINCS algorithm~\cite{hess97,hess07}. 
Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with real space cut-off at 1.0~nm was used 
for electrostatics. Twin range cut-off was used for the Lennardt-Jones interactions with short and long cut-offs at 1.0~nm and 1.6~nm, respectively.
The neighbour lists were updated every 5th step with cut-off at 1.0~nm. Temperature was coupled separately
for lipids and water to 298~K with the velocity-rescale method~\cite{bussi07} with coupling constant 0.2~ps.
Pressure was semi-isotropically coupled to the atmospheric pressure with the Parrinello-Rahman method~\cite{parrinello81}.
%\subsubsection{H\"ogberg et al. and Rabonovich et al.}
%Simulations done in \cite{hogberg08} and \cite{rabinovich14} are used.

%Time step 2fs, Ewald summation,
%cut-off 14 Å, update of neighbour list each 10 steps,
%Nose-Hoover thermostat,  Parrinello-Rahman barostat,
%semi-anizotropic pressure coupling,
%long-range isotropic pressure correction
\subsubsection{Ulmschneider}
The initial structure containing 128 POPC molecules with 3328 TIP3P water~\cite{jorgensen83} molecules (26 per lipid) was downloaded from Lipidbook \cite{domanski10} 
together with the topologies. This bilayer was simulated for 100~ns with a time step of 2~fs and the data was saved every 10~ps. The bonds involving hydrogen atoms were 
constrained with LINCS~\cite{hess97,hess07}. The temperature was kept at 298~K with the Berendsen thermostat~\cite{berendsen84}. The pressure was semi-isotropically coupled to the Berendsen 
barostat~\cite{berendsen84} with a time constant of 1~ps and a target pressure of 1~bar. PME~\cite{darden93,essman95} was employed for long range electrostatics and a cut-off of 1~ns was employed for 
the Lennard-Jones interactions. A neighbour list with a radius of 1~nm was updated every 10~steps.

Additionally, the simulations were repeated with the dispersion correction applied to pressure and temperature. Even though the area per lipid decrease
d slightly, the head group order parameters were only slightly affected.

\subsubsection{Tj\"ornhammar et al.}
%The used simulation files and part of the trajectroy are available~\cite{tjornhammarfiles}.

The gel phase structure delivered by Tj\"ornhammar  and edholm~\cite{tjornhammar14} was ran a 5ns at 70 degrees in order to destroy the 
ordered gel configuration. This was followed by 200ns simulation at 50 degrees. The last 100ns of this simulation was used for analysis. 
The same mdp file as in the paper's~\cite{tjornhammar14} SI was used except for the temperatures.

\subsubsection{CHARMM36-UA}
A hydrated bilayer consisting of 128 DLPC lipids and 3840 water molecules is modeled by the force field of Lee and co-workers~\cite{lee14}, 
which is a combination of the all-atom CHARMM36 force-field~\cite{klauda10} and the united-atom Berger model~\cite{berger97}. 
The nonbonding interactions are calculated using an atom-based switching function with inner and outer cutoffs of 0.8 and 1.2 nm~\cite{lee14}. 
Long range electrostatic interactions are implemented using the particle-particle particle-mesh solver with a relative accuracy of 0.0001. The system 
is first equilibrated for 30 ns in the NP$\gamma$T ensemble (Nose-Hoover~\cite{nose84,hoover85} style thermostat and barostat with anisotropic pressure coupling) 
at 323 K and 1 bar with timestep of 1 fs, the next 20 ns of dynamics are taken for calculation of configurational averages. 
\todo{The simulation times given currently contradict with the Table~\ref{systems}.}
Simulations were 
carried out by using the LAMMPS package~\cite{plimpton95}. %( http://doi.org/10.1006/jcph.1995.1039 )., the input files are available ( http://dx.doi.org/10.5281/zenodo.13821 ).

%\subsubsection{OPLS-AA}
%\todo{Joona Tynkkynen. We have this only with 150mM of NaCl delivered by Joona Tynkkynen. Options are to remove these results from this publicatio, run the simulations without ions (if not yet available), 
%  or include the results with ions. In my understanding this is a proto version of MacRog so it could be left out as well. However, for historical reasons and 
%to understand literature it might be useful to include also these results. COMMENT: Matti Javanainen suggested that we would remove these results. If there are no objections and the
%data without ions will not be delivered, we will remove this.}

\subsection{Dihedral angle distributions as a function of cholesterol in CHARMM36}
\begin{figure*}[]
  \centering
  \includegraphics[width=17.2cm]{dihsCHOLcharmm.pdf}
  \caption{\label{dihsCHOLcharmm}
    The effect of cholesterol content on the glycerol backbone and choline dihedral angles in CHARMM36 model.}
\end{figure*}

%\newpage

\onecolumngrid

 \listoftodos

\end{document}

### 2015-01-20 HGmodel_draft3.tex
new version

@@
 would automatically resolve the structures giving an interpretation of experimental results, if the model
-would reproduce the experimental data. In this work we compare the C-H bond vector order
+reproduced the experimental data. In this work we compare the C--H bond vector order
 parameters for glycerol backbone and choline headgroup between 14 different atomistic resolution
-models and experiments in fully hydrated lipid bilayer. The current models are not accurately enough to resolve the structure.
+models and experiments in a fully hydrated lipid bilayer. The current models are not accurately enough to resolve the structure.
 However, closer inspection of three best performing models (CHARMM36, GAFFlipid and MacRog) suggest
@@
 would resolve the structure. Despite of the inaccuracy in the fully hydrated
-structures, the response to the dehydration, i.e. P-N vector tilting more parallel to membrane normal, 
+structures, the response to dehydration, i.e. the tilting of the P--N vector more parallel to membrane normal, 
 is qualitatively correct in all models. The CHARMM36 and MacRog models describe the interactions
-between lipids and cholesterol better than Berger/H\"oltje model.
-This work has been, and continues to be, progressed and discussed through the blog: nmrlipids.blogspot.fi. 
+between lipids and cholesterol better than the Berger/H\"oltje model.
+This work has been, and continues to be, progressed and discussed through the blog: \texttt{nmrlipids.blogspot.fi}. 
 Everyone is invited to join the discussion and make contributions through the blog. 
@@
 Everyone who has contributed to the work through the blog will be offered 
-coauthorship. For more details see: nmrlipids.blogspot.fi.
+coauthorship. For more details see: \texttt{nmrlipids.blogspot.fi}.
 
@@
 
-Phospholipids containing different polar headgroups and different acyl chains are essential building blocks of 
+Phospholipids containing various polar headgroups and acyl chains are essential building blocks of 
 biological membranes. Lamellar phospholipid bilayer structures have been widely studied with various experimental 
-and theoretical techniques as a simple model for the biological membranes~\cite{lipowsky95,tieleman97,klauda08,edholm08,tieleman10,piggot12,rabinovich13,marsh13}. 
-Phospholipid molecules are composed of hydrophobic acyl chains and hydrophilic headgroup, which are connected by glycerol backbone,
-see Fig.~\ref{POPCstructure} for the structure of 1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
-The behaviour of the acyl chains in a bilayer is relatively well understood~\cite{Israelachvili80,lipowsky95,tieleman97,klauda08,edholm08,tieleman10,marsh13}. 
-The structures sampled by the glycerol backbone and choline in liquid bilayer state, however, are not fully 
+and theoretical techniques as a simple model for cellular membranes~\cite{lipowsky95,tieleman97,klauda08,edholm08,tieleman10,piggot12,rabinovich13,marsh13}. 
+Phospholipid molecules are composed of hydrophobic acyl chains which are connected by a glycerol backbone to a hydrophilic headgroup.
+See Fig.~\ref{POPCstructure} for the structure of 1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
+The behaviour of the acyl chains in a lipid bilayer is relatively well understood~\cite{Israelachvili80,lipowsky95,tieleman97,klauda08,edholm08,tieleman10,marsh13}. 
+The conformations sampled by the glycerol backbone and choline in a fluid bilayer are, however, not fully 
 resolved since even the most accurate scattering and Nuclear Magnetic Resonance (NMR)
-techniques give only a set of values that the structure has to fulfill, but
-there is no unique way to derive the actual structure from these parameters~\cite{seelig77b,skarjune79,Israelachvili80,jacobs80,davis83,strenk85,akutsu91,hong95b,hong96,semchyschyn04}.
+techniques give only a set of values that the structure has to fulfil, but
+there is no unique way to derive the actual structure from them~\cite{seelig77b,skarjune79,Israelachvili80,jacobs80,davis83,strenk85,akutsu91,hong95b,hong96,semchyschyn04}.
 Some structural details have been extracted from crystal structure, $^1$H NMR studies and Raman spectroscopy~\cite{hauser80,hauser81,hauser81b,akutsu81b,pascher92,hauser88,marsh06}
-but general consensus about the structures sampled in the liquid state have not been reached~\cite{seelig77b,skarjune79,Israelachvili80,jacobs80,davis83,strenk85,hauser88,akutsu91,hong95b,hong96,semchyschyn04,marsh06}.
+but general consensus about the structures sampled in the fluid state have not been reached~\cite{seelig77b,skarjune79,Israelachvili80,jacobs80,davis83,strenk85,hauser88,akutsu91,hong95b,hong96,semchyschyn04,marsh06}.
 On the other hand, the glycerol backbone structures are similar for various biologically
@@
 real cells (mouse fibroblast L-M cell)~\cite{scherer87}.
-Thus, the resolvation of phosphatidylcholine glycerol backbone and choline structures would be 
-useful for understanding wide range of different biological membranes.
+Thus, the solution of phosphatidylcholine glycerol backbone and choline structures would be 
+useful for understanding a wide range of different biological membranes.
 
-Classical atomistic resolution molecular dynamics simulations have been widely used to study  
+Classical atomistic molecular dynamics simulations have been widely used to study  
 lipid bilayers~\cite{tieleman97,klauda08,edholm08,tieleman10,piggot12,rabinovich13}. As these models provide an atomistic
-resolution description of the whole lipid molecule, they have potential to resolve the glycerol backbone and 
-headgroup structures. In particular, the experimental C-H bond order parameters for the glycerol backbone 
+resolution description of the whole lipid molecule, they have potential to solve the glycerol backbone and 
+headgroup structures. In particular, the experimental C--H bond order parameters for the glycerol backbone 
 (g$_1$, g$_2$ and g$_3$) and choline ($\alpha$ and $\beta$) segments (see Fig.~\ref{POPCstructure} for definitions) are among the main parameters used in
-attempts to derive the structures from experimental data~\cite{seelig77b,skarjune79,jacobs80,davis83,akutsu91,hong95b,semchyschyn04}.
+attempts to derive lipid structures from experimental data~\cite{seelig77b,skarjune79,jacobs80,davis83,akutsu91,hong95b,semchyschyn04}.
 These parameters are also routinely compared between experiments and simulations for the acyl chains~\cite{tieleman97,klauda08,edholm08,tieleman10,piggot12}.
 Thus, the structures sampled in a simulation model that reproduces these and other experimental parameters, automatically
-give an interpretation of the experiments, in other words they can be considered as reasonable atomistic resolution descriptions of
-the behaviour of lipid molecules in a bilayer.
+give an interpretation of the experiments. In other words they can be considered as reasonable atomistic resolution descriptions of
+the behavior of lipid molecules in a bilayer.
 
@@
 in some studies~\cite{shinoda97,hogberg08,castro08,klauda10,kapla12,dickson12,poger12,ferreira13,chowdhary13,maciejewski14}, 
-however much less frequently than for acyl tail chains~\cite{tieleman97,klauda08,edholm08,tieleman10,piggot12}.
+however much less frequently than for the acyl  chains~\cite{tieleman97,klauda08,edholm08,tieleman10,piggot12}.
 The main reason is probably that the existing experimental data for the glycerol backbone
@@
 In addition to the order parameters, also dihedral angles for glycerol backbone and headgroup estimated from experiments have been sometimes used to 
-assess the quality of simulation model~\cite{robinson94,essex94,kothekar96,hyvonen97,shinoda97,duong99}.
+assess the quality of a simulation model~\cite{robinson94,essex94,kothekar96,hyvonen97,shinoda97,duong99}.
 
@@
 order parameters we can distinguish the more reasonable structures from the less reasonable ones. However, none of the current models 
-is accurate enough to resolve the atomistic resolution stuctures.
+is accurate enough to properly resolve the atomistic resolution stuctures.
 
-In addition to the fully--hydrated single--component lipid bilayers, the glycerol backbone and choline order parameters
+In addition to fully hydrated single component lipid bilayers, the glycerol backbone and choline order parameters
 have been measured under a large number of different conditions. For example, as a function of hydration level~\cite{bechinger91,ulrich94,dvinskikh05b}, cholesterol content~\cite{brown78,ferreira13}
@@
   \centering
-  \includegraphics[width=8.6cm]{POPCstructure.eps}
+%  \includegraphics[width=8.6cm]{POPCstructure.eps}
 
@@
 
-This work has been done as an open collaboration by using nmrlipids.blogspot.fi as an communication platform.
+This work has been done as an open collaboration by using \texttt{nmrlipids.blogspot.fi} as a communication platform.
 The approach is inspired by the Polymath project~\cite{gowers09}, however there are some essential differences. The manuscript pointing out problems in the glycerol backbone and
@@
 All the contributions were done publicly through the blog and the blog contributors were offered coauthorship. The final author list
-is based on the self-assesment of the authors scientific contribution to the project.
+is based on the self-assesment of the authors' scientific contribution to the project.
 
-Large portion of the used simulation files and scripts are shared through the GitHub organization https://github.com/NMRlipids
-and Zenodo community https://zenodo.org/collection/user-nmrlipids.
+Large portion of the used simulation files and scripts are shared through the GitHub organization \texttt{https://github.com/NMRlipids}
+and the Zenodo community \texttt{https://zenodo.org/collection/user-nmrlipids}.
 
@@
 Maybe we should change the name of this section II.B. to: "Analysis and terminology of order parameters from experiments and simulations" or similar, and also move the "forking" explanation here.}
-The order parameter of a hydrocarbon C-H vector is defined as $S_{{\rm CH}}=\frac{3}{2}\langle \cos^2 \theta-1 \rangle$, where
-the average is an ensemble average over the sampled conformations, and $\theta$ is the angle between the C-H bond and the membrane normal.
+The order parameter of a hydrocarbon C--H vector is defined as $S_{{\rm CH}}=\frac{3}{2}\langle \cos^2 \theta-1 \rangle$, where
+the angle brackets denote an ensemble average over the sampled conformations, and $\theta$ is the angle between the C--H bond and the membrane normal.
 The order parameters can be measured by detecting quadrupolar splitting with $^2$H NMR~\cite{seelig77c} or by detecting dipolar 
@@
 $|S_{{\rm CH}}|=\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c} d_\mathrm{CH}$, where
-values between 20.2-22.7 kHz are used for $\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c}$,
+values between 20.2--22.7 kHz are used for $\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c}$,
 depending on the original authors~\cite{hong95a,gross97,dvinskikh05a,ferreira13}.
@@
 with each other (see Results and Discussion), indicating very high quantitative accuracy of the measurements.
-For more detailed discussion see http://nmrlipids.blogspot.fi/2014/02/accuracy-of-order-parameter-measurements.html
+For a more detailed discussion see \texttt{http://nmrlipids.blogspot.fi/2014/02/accuracy-of-order-parameter-measurements.html}
 \todo{This is temporarily linked to the blog. Currently I think that the posts (with comments included) to which we want to cite are put into 
@@
 For the united atom models the hydrogen locations were generated 
-post-simulationally using the positions of the heavy atoms in the simulation trajectories.
-The statistical error estimates were calculated for the best performing simulation
+in the trajectories post-simulationally using the positions of the heavy atoms and the known hydrocarbon geometries.
+The statistical errors were estimated for the best performing simulation
 models by calculating the error of the mean for the average over individual lipids in the system.
@@
 slow compared to the typical simulation timescales~\cite{vogel12}. On the other hand, another recent
-study shows that the slowest rotational correlation functions of C-H bond (g$_1$) reaches plateau (S$_{CH}^2$)
-after 200ns in the Berger-POPC-07~\cite{ollila07a} model, and that the dynamics of this segment is significantly too slow in simulations
-compared to the experiments~\cite{ferreira15}. In practise, less than 200~ns data set is enough for the order parameter
+study shows that the slowest rotational correlation functions of a C--H bond (g$_1$) reaches a plateau (S$_{CH}^2$)
+after 200~ns in the Berger-POPC-07~\cite{ollila07a} model, and that the dynamics of this segment is significantly too slow in simulations
+compared to the experiments~\cite{ferreira15}. In practise, less than 200~ns of simulation data is enough for the order parameter
 calculation due to the average over different lipid molecules. In conclusion, if the sampling with typical simulation times
-is not enough for the convergense of the order parameters, then the simulation models has significantly too slow dynamics.
+is not enough for the convergence of the order parameters, then the simulation models has significantly too slow dynamics.
 
 \subsection{Simulated systems}
-All simulations are ran with a standard setup for planar lipid bilayer in zero tension and constant temperature
-with periodic boundary conditions in all directions by using Gromacs software package~\cite{hess08} 
-(version numbers 4.5-4.6) or LAMMPS~\cite{plimpton95}.
-The number of molecules, temperatures and the length of simulations for all the simulated systems 
+All simulations are ran with a standard setup for a planar lipid bilayer in zero tension and constant temperature
+with periodic boundary conditions in all directions by using the GROMACS software package~\cite{hess08} 
+(version numbers 4.5.X--4.6.X) or LAMMPS~\cite{plimpton95}.
+The number of molecules, simulation temperatures and the length of simulations of all the simulated systems 
 are listed in Tables~\ref{systems},~\ref{systemsDEHYD} and~\ref{systemsCHOL}. Full simulation
-details are given in the Supplementary Information (SI) or in the original publications if the
-data is used previously. For some systems also the simulation related files and trajectories are publicly
-available. The references pointing to simulation details and files are also listed in Tables~\ref{systems},~\ref{systemsDEHYD} and~\ref{systemsCHOL}.
+details are given in the Supplementary Information (SI) or in the original publications in case the
+data is used previously therein. Additionally, the files related to the simulations and the resulting trajectories are publicly
+available for some systems. The references pointing to simulation details and files are also listed in Tables~\ref{systems},~\ref{systemsDEHYD} and~\ref{systemsCHOL}.
 
@@
 \centering
-\caption{Simulated single component lipid bilayers with full hydration. The simulation file data sets marked with $^*$ include also part of the trajectory.
+\caption{Simulated single component fully hydrated lipid bilayers. The simulation file data sets marked with $^*$ also include a part of the trajectory.
   If simulation data from previously published work has been directly used, the original publication is cited for simulation details. For other systems the simulation details
-  are given in Supplementary Information. 
+  are given in the Supplementary Information. 
 }\label{systems}
@@
 \centering
-\caption{Simulated lipid bilayers with cholesterol. The simulation file data sets marked with $^*$ include also part of the trajectory.
+\caption{Simulated lipid bilayers containing cholesterol. The simulation file data sets marked with $^*$ include also part of the trajectory.
 }\label{systemsCHOL}
@@
 
-\subsection{Full hydration: Experimental order parameters for glycerol backbone and headgroup}\label{experiments}
+\subsection{Full hydration: Experimental order parameters for the glycerol backbone and headgroup}\label{experiments}
 The specific deuteration of $\alpha$-, $\beta$- and g$_3$- segments of dipalmitoylphosphatidylcholine (DPPC) has been successful, 
@@
   \centering
-  \includegraphics[width=8.6cm]{comparison3.pdf}
+%  \includegraphics[width=8.6cm]{comparison3.pdf}
 %\todo{DONE}\\
@@
   Order parameteres from simulations listed in Table~\ref{systems} and experiments for glycerol and choline groups.
-  The experimental values were taken from the following publications: DMPC 303K from \cite{gross97}, DMPC 314K from \cite{dvinskikh05a}, DPPC 310K from \cite{gally75}, 
-  DPPC 323K from \cite{akutsu81}, and POPC 298K \cite{ferreira13}.
+  The experimental values were taken from the following publications: DMPC 303~K from \cite{gross97}, DMPC 314~K from \cite{dvinskikh05a}, DPPC 310K from \cite{gally75}, 
+  DPPC 323~K from \cite{akutsu81}, and POPC 298~K \cite{ferreira13}.
   The vertical bars shown for some of the computational values are not error bars, but demonstrate that for 
@@
   The H{\"o}gberg et al. force field in Table~\ref{systems} is abbreviated in this figure as HoNiLy, Tj{\"o}rnhammar et al. as TjoEdh, CHARMM36-UA as C36UA,
-  Rabinovic et al. as RabLyu and Ulmschneider as UlmUlm.
+  Rabinovic et al. as RabLyu and Ulmschneider and Ulmschneider as UlmUlm.
 } 
@@
 In general there is a good agreement between the order parameters measured with different experimental NMR techniques: Almost all the 
-reported values are inside variation of $\pm$0.02 (which is also the error estimate given by Gross et al.~\cite{gross97}) 
-for all fully hydrated PC bilayer, regardless of the variation in their acyl chain composition and the temperature.
-Exception are the somewhat lower order parameters sometimes reported having been measured with $^{13}$C--NMR~\cite{hong95a,hong95b,warschawski05}.
-These experiments have not seen in Fig.~\ref{HGorderparameters} as the reported error bars are either relatively large~\cite{hong95a,hong95b}, 
+reported values are within a variation of $\pm$0.02 (which is also the error estimate given by Gross et al.~\cite{gross97}) 
+for all fully hydrated PC bilayer, regardless of the variation in their acyl chain composition and temperature.
+Exceptions are the somewhat lower order parameters sometimes reported from measurements using $^{13}$C--NMR~\cite{hong95a,hong95b,warschawski05}.
+These experiments are not shown in Fig.~\ref{HGorderparameters} as the reported error bars are either relatively large~\cite{hong95a,hong95b}, 
 or the spectral resolution is quite low and the numerical lineshape simulations have not been used in the analysis~\cite{warschawski05}.
-Therefore it is highly likely that these reported lower order parameters are due to lower experimental 
-accuracy and that we exclude these values from our discussion. 
+Due to this end, it is highly likely that these reported lower order parameters are due to lower experimental 
+accuracy and therefore we exclude them from our discussion. 
 Motivated by the high experimental repeatability, we have highlighted in 
-Fig.~\ref{HGorderparameters} subjective sweet spots (light blue areas), within which we expect the calculated absolute 
-values of order parameters of a well-performing force field to fall.
+Fig.~\ref{HGorderparameters} the subjective sweet spots (light blue areas), within which we expect the calculated absolute 
+values of the order parameters of a well-performing force field to fall.
 
@@
 while the hydrogens in choline $\alpha$ and $\beta$ segments give equal values.
-Note that in this work we call the phenomena of inequal order parameters for hydrogens attached to the same carbon as ''forking'' to avoid 
-confusion with dipolar and quadrupolar splitting in NMR terminology. Forking is also observed experimentally for the C$_2$ carbon in the sn-2
-chain of all phosholipids, and it is known to arise from differently sampled orientations of the two C-H bonds, not from two 
-different populations of lipid conformations~\cite{engel81}. The forking in glycerol backbone g$_3$ segment is small ($\approx$ 0.02) 
-and some experiments only report the larger or the average value~\cite{akutsu81,ferreira13}. 
+Note that in this work we call the phenomena of unequal order parameters for hydrogens attached to the same carbon as ``forking'' to avoid 
+confusion with dipolar and quadrupolar splitting in NMR terminology. Forking is also observed experimentally for the C$_2$ carbon in the \textit{sn}-2
+chain of all phosholipids, and it is known to arise from differently sampled orientations of the two C--H bonds, not from two 
+separate populations of lipid conformations~\cite{engel81}. The forking in glycerol backbone g$_3$ segment is small ($\approx$ 0.02) 
+and some experiments only report the larger value or the average value~\cite{akutsu81,ferreira13}. 
 In contrast, forking is significant for the glycerol backbone g$_1$ segment, whose lower order parameter is close to zero and the
-larger one has absolute values around 0.13-0.15. Forking was studied in detail by Gally et al.~\cite{gally81}, who used E. Coli to 
+larger one has an absolute value of approximately 0.13--0.15. Forking was studied in detail by Gally et al.~\cite{gally81}, who used E. Coli to 
 stereospecifically deuterate the different hydrogens attached to the g$_1$ or g$_3$ groups in PE lipids, and measured the order parameters from the lipid 
@@
 the sign of the order parameter~\cite{hong95a,hong95b,gross97}. The measured sign is negative for almost all the carbons 
-discussed in this work, only $\alpha$ is positive~\cite{hong95a,hong95b,gross97}.
+discussed in this work as only $\alpha$ is positive~\cite{hong95a,hong95b,gross97}.
 
-Combining the experimental information of the sign~\cite{hong95a,hong95b,gross97} and the stereospecifity 
-measurements~\cite{gally81} with the absolute value measurements from various techniques~\cite{gally75,akutsu81,gross97,dvinskikh05a,ferreira13}
-having high quantitative accuracy,
+By combining the experimental information of the sign~\cite{hong95a,hong95b,gross97} and the stereospecific
+measurements~\cite{gally81} with the accurate absolute value measurements from various techniques~\cite{gally75,akutsu81,gross97,dvinskikh05a,ferreira13},
 the most detailed experimentally available order parameter information for the glycerol backbone and choline segments of POPC is obtained.
@@
   \centering
-  \includegraphics[width=8.6cm]{HGorderparameters5.eps} \\
+%  \includegraphics[width=8.6cm]{HGorderparameters5.eps} \\
   \todo{Markus Miettinen suggested that we should consider ``spreading'' the values similarly to the figure~\ref{HGorderparameters}} \\
   \caption{\label{HGorderparameters2}
-  Order parameteres from simulations with Berger-POPC-07, CHARMM36, GAFFlipid and Macrog force fields together with experimental values for POPC glycerol and choline groups.
+  Order parameteres from simulations with Berger-POPC-07, CHARMM36, GAFFlipid and MacRog force fields together with experimental values for POPC glycerol and choline groups.
   The magnitudes for experimental order parameters are taken from Ferreira et al.~\cite{ferreira13}, the signs are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
@@
 
-The order parameters of the glycerol backbone and headgroup calculated from different force fields for various lipids has been 
+The order parameters of the glycerol backbone and headgroup calculated from different force fields for various lipids have been 
 previously compared to experiments~\cite{shinoda97,hogberg08,castro08,klauda10,kapla12,dickson12,poger12,ferreira13,chowdhary13,maciejewski14}. 
-The general conclusion from these works seems to be that the CHARMM based~\cite{hogberg08,klauda10}, GAFFlipid~\cite{dickson12} and
-MacRog~\cite{maciejewski14} force fields perform better for the glycerol backbone and headgroup structures than the Gromos based models~\cite{castro08,kapla12,poger12,ferreira13}.
-However, none of the studies exploits the full potential of the available experimental data discussed in previous section; the quantitative accuracy, known signs and stereospecific labeling of
+The general conclusion from these studies seems to be that the CHARMM based~\cite{hogberg08,klauda10}, GAFFlipid~\cite{dickson12} and
+MacRog~\cite{maciejewski14} force fields perform better for the glycerol backbone and headgroup structures than the GROMOS based models~\cite{castro08,kapla12,poger12,ferreira13}.
+However, none of the studies exploits the full potential of the available experimental data discussed in previous section, i.e. the quantitative accuracy, known signs and stereospecific labeling of
 the experimental order parameters.
 
-To get a general idea about the quality of the glycerol backbone and choline headgroup structures in different models, we calculated the absolute
+To get a general idea of the quality of the glycerol backbone and choline headgroup structures in different models, we calculated the absolute
 values of the order parameters for these parts from fourteen different lipid models (Table~\ref{systems}) and 
 plotted the results together with experimental values in Fig.~\ref{HGorderparameters}.
-Two criteria were used to judge the quality of the model: {\bf forkings} there must not be significant forking in the $\alpha$ and $\beta$ carbons,
+Two criteria were used to judge the quality of the model: there must not be significant  {\bf forking} in the $\alpha$ and $\beta$ carbons,
 there must be only moderate forking in the g$_3$ carbon and there must be significant forking in the g$_1$ carbon, {\bf absolute values}
 should be preferably inside to the subjective sweet spots determined from experiments (blue shaded regions in Fig.~\ref{HGorderparameters}).
-None of the studied force fields fulfills these criteria, however, three force fields are closer than others: CHARMM36~\cite{klauda10}, MacRog~\cite{maciejewski14} and GAFFlipid~\cite{dickson12}.
+None of the studied force fields fulfils these criteria.
+However, three force fields perform better than others, namely CHARMM36~\cite{klauda10}, MacRog~\cite{maciejewski14} and GAFFlipid~\cite{dickson12}.
 The results for each force field in respect to the above criteria are summarized in Table~\ref{??} \todo{This kind of table should be done or not?}.
@@
 nonrealistic structural samping in molecular dynamics simulations. The demonstration is done for 
-the dihedral angles defined by the g$_3$-g$_2$-g$_1$-O(sn-1) segments in the glycerol backbone and 
+the dihedral angles defined by the g$_3$-g$_2$-g$_1$-O(\textit{sn}-1) segments in the glycerol backbone and 
 the N-$\beta$-$\alpha$-O segments in the headgroup. These dihedrals were chosen for demonstration, because 
@@
 We note that performing a similar comparison through all the dihedrals in all the 14 models would probably give highly useful
-information to improve the accuracy of simulations, however this is beyond the scope of the current report.
+information on how to improve the accuracy of the models yet this is beyond the scope of the current report.
 
-The dihedral angle distributions for the  g$_3$-g$_2$-g$_1$-O(sn-1) dihedral calculated from different models are
+The dihedral angle distributions for the  g$_3$-g$_2$-g$_1$-O(\textit{sn}-1) dihedral calculated from different models are
 shown in Fig.~\ref{dihDISTS}. The distribution is qualitatively different for the Berger-POPC-07 model, showing a maximum in 
-the gaughe$^+$-conformation (60$^o$) compared to all the other models showing a maximum in the trans-conformation (180$^o$).
+the gauche$^+$-conformation (60$^o$) compared to all the other models showing a maximum in the anti-conformation (180$^o$).
 The distributions in all the other models have the same general features, the main difference being that the
-fraction of configurations in gaughe$^-$-conformation (-60$^o$) is zero for the MacRog, detectable for the CHARMM36 and
-equally large to the gaughe$^+$ fraction in GAFFlipid. From the results we conclude that most likely the wrongly sampled
+fraction of configurations in the gauche$^-$-conformation (-60$^o$) is zero for the MacRog, detectable for the CHARMM36 and
+equally large to the gauche$^+$ fraction in GAFFlipid. From the results we conclude that most likely the wrongly sampled
 dihedral angle for the g$_2$-g$_1$ bond explains the significant discrepancy to the experimental order parameters
 for the g$_1$ segment in the Berger-POPC-07 model (Fig.~\ref{HGorderparameters2}). 
-The result that models preferring the trans conformation for this dihedral gives more realistic order parameters
-is in agreement with previous crystal structure and $^1$H NMR studies~\cite{hauser80,hauser81,hauser81b,hauser88,pascher92,marsh06}.
+In conclusion, models preferring the anti conformation for this dihedral give more realistic order parameters and
+this in agreement with previous crystal structure and $^1$H NMR studies~\cite{hauser80,hauser81,hauser81b,hauser88,pascher92,marsh06}.
 \begin{figure}[]
   \centering
-  \includegraphics[width=8.6cm]{g1-g2_Cdihs2.eps}
+%  \includegraphics[width=8.6cm]{g1-g2_Cdihs2.eps}
   \caption{\label{dihDISTS}
-    Dihedral angle distributions for g$_3$-g$_2$-g$_1$-O(sn-1) dihedral from different models (POPC bilayer in full hydration).
+    Dihedral angle distributions for g$_3$-g$_2$-g$_1$-O(\textit{sn}-1) dihedral from different models (POPC bilayer in full hydration).
       } 
@@
 The dihedral angle distribution for the  N-$\beta$-$\alpha$-O dihedral calculated from the same four models is 
-shown in Fig.~\ref{dihDISTS2}. Also for this dihedral there are significant differences in the gauche-trans fractions.
-The gaughe conformations are dominant in the CHARMM36, in MacRog there are only trans conformations present,
-and in the Berger-POPC-07 and GAFFlipid gaughe and trans conformation has equal probability. 
-On the other hand comparison of $\alpha$ and $\beta$ order parameters in Fig.~\ref{HGorderparameters2}
-reveals that the CHARMM36 is closest to the experimental results being the only model having correct
+shown in Fig.~\ref{dihDISTS2}. Also for this dihedral there are significant differences in the gauche--anti fractions.
+The gauche conformations are dominant in the CHARMM36, in MacRog there are only anti conformations present,
+and in the Berger-POPC-07 and GAFFlipid gauche and anti conformations have equal probabilities. 
+On the other hand, comparison of $\alpha$ and $\beta$ order parameters in Fig.~\ref{HGorderparameters2}
+reveals that for these carbons the CHARMM36 is closest to the experimental results and it is also the only model that has the correct
 sign (negative) for the $\beta$ order parameter. This result is again in agreement with previous 
 crystal structure, $^1$H NMR and Raman spectroscopy studies~\cite{hauser80,hauser81,hauser81b,akutsu81b} which suggest that
-this dihedral has only gaughe conformation in the absense of ions.
+this dihedral is in the gauche conformation in the absence of ions.
 
@@
   \centering
-  \includegraphics[width=8.6cm]{a-bDIHS2.eps}
+%  \includegraphics[width=8.6cm]{a-bDIHS2.eps}
   \caption{\label{dihDISTS2}
@@
 The used examples show that the glycerol backbone and headgroup order parameters reflect the atomistic resolution structure
-and that the comparison with experiments allows the assesment of the quality of the suggested structure. We were able to pinpoint
+and that the comparison with experiments allows the assessment of the quality of the suggested structure. We were able to pinpoint
 specific problems in the structures in different models and suggest potential improvement strategies.
-If the improved atomistic resolution
-molecular dynamics simulation model would reproduce the order parameters and other experimental observables (like chemical shift anisotropy)
+If the improved atomistic
+molecular dynamics simulation model reproduced the order parameters and other experimental observables (like chemical shift anisotropy)
 with experimental accuracy, it would give an interpretation for the atomistic resolution structure of the glycerol backbone and 
@@
 even very small order parameter changes resulting from the varying conditions can be measured (see
-http://nmrlipids.blogspot.fi/2014/02/accuracy-of-order-parameter-measurements.html
+\texttt{http://nmrlipids.blogspot.fi/2014/02/accuracy-of-order-parameter-measurements.html}
 for more discussion.
@@
 the structural changes from order parameter changes~\cite{akutsu91,semchyschyn04}. Consequently, comparison of the order parameters
-between simulations and experiments in different conditions can be used measure the quality of the force field 
+between simulations and experiments in different conditions can be used to assess the quality of the force field 
 in different situations, and, if the quality is good, to potentially interpret the structural changes in experiments.
-Here we exemplify such comparison for a lipid bilayer under low hydration levels and mixed with cholesterol. 
-The interaction between ions and phosphatidylcholine bilayer is discussed in a separate work~\cite{ionpaper}.
+Here we exemplify such comparison for a lipid bilayer under low hydration levels and when varying amounts of cholesterol is included in the bilayer. 
+The interaction between ions and a phosphatidylcholine bilayer will be discussed in a separate study~\cite{ionpaper}.
 
@@
 for the glycerol backbone and choline as a function of hydration level are shown in Fig.~\ref{ordPhydr}. 
-The independently reported values for choline segments are in good agreement with each other (despite of 
+The independently reported values for choline segments are in good agreement with each other (despite 
 slight differences in temperature and acyl chain composition),
@@
 absolute values were measured in the original experiments~\cite{dvinskikh05b,ulrich94,bechinger91}, but
-we have included the signs measured separately~\cite{hong95a,hong95b,gross97}. 
+we have included the signs measured in separate studies~\cite{hong95a,hong95b,gross97}. 
 Consequently, the $\beta$ order parameter with negative sign actually increases with dehydration 
@@
   \centering
-  \includegraphics[width=8.6cm]{OrderParameterDEHYDpresentation.eps}
+%  \includegraphics[width=8.6cm]{OrderParameterDEHYDpresentation.eps}
 \todo{Several people think that this plot should be clarified.}
@@
 order parameters (except by Mashl et al.~\cite{mashl01}).
-In Fig.~\ref{ordPhydr} the glycerol backbone and choline order parameters as a function of hydration level are shown 
-for the CHARMM36, MacRog and GAFFlipid models (having the most realistic atomistic resolution structures) together with the Berger based model 
-(which is the most used lipid model). The choline order parameters increase with dehydration in all simulation
+In Fig.~\ref{ordPhydr} the glycerol backbone and choline order parameters are shown as a function of hydration level for the CHARMM36, MacRog and GAFFlipid models (having the most realistic atomistic resolution structures) together with the Berger based model
+(which is the most used lipid model). The choline order parameters increase with dehydration in all
 models, in qualitative agreement with experiments. 
@@
 orients more parallel to the membrane plane with dehydration due to restricted interlamellar space. 
-Indeed, the P-N angle vector angle with membrane normal as a function of dehydration shows an increase for
+Indeed, the P--N (phoshate phosphorus to choline nitrogen) vector angle with membrane normal shows an increase for
 all models as a function of dehydration in Fig.~\ref{PNangle}.
 However, the qualitative agreement in the lipid response to dehydration does not guarantee the correct 
-free energy landscape if the simulation model has incorrect structure. The influence of this issue to 
+free energy landscape if the simulation model has an incorrect structure. The influence of this issue to 
 dehydration energetics studied with simulations~\cite{eun09,schneck12} is left for future studies.
@@
   \centering
-  \includegraphics[width=8.6cm]{PNangles.eps}
+%  \includegraphics[width=8.6cm]{PNangles.eps}
 
   \caption{\label{PNangle}
-    The angle between membrane normal and P-N vector in choline segment as function of
+    The angle between membrane normal and P--N vector as function of
     hydration level calculated from different simulations.
@@
 
-The response of the glycerol backbone to dehydration seems to be more subtle than of the choline headgroup 
+The response of the glycerol backbone to dehydration seems to be more subtle than that of the choline headgroup 
 as CHARMM36 is the only model that reproduces the decrease in g$_2$ and g$_3$ segments.
 
-\subsubsection{Phospholipid bilayer mixed with cholesterol}
+\subsubsection{Cholesterol-containing phospholipid bilayer}
 Phospholipid--cholesterol interactions have been widely studied with theoretical~\cite{huang99,zhu07,rog09,alwarawrah12} and
-experimental methods~\cite{brown78,marsh10,ferreira13,marsh13}, since cholesterol is abundant in biological membranes and
-it has been suggested to be an important player, for example, in domain formation~\cite{simons04,somerharju09}.
+experimental methods~\cite{brown78,marsh10,ferreira13,marsh13}, since cholesterol is abundant in biological membranes.
+It has been suggested to be an important player, for example, in domain formation~\cite{simons04,somerharju09}.
 It is widely agreed that cholesterol orders lipid acyl tails thus decreasing the area per molecule (condensing effect),
@@
 and cholesterol causes a measurable increase in the $\beta$ order parameter and a forking in the $\alpha$ order
-parameter~\cite{brown78}, but these effects are so small that they are barely visible in the scale used in Fig.~\ref{ordPchol}.
+parameter~\cite{brown78}. These effects are, however, so small that they are barely visible in the scale used in Fig.~\ref{ordPchol}.
 Further, the effects of cholesterol on the glycerol backbone order parameters for POPC from $^{13}$C NMR experiment~\cite{ferreira13} 
-is in good agreement with the results for the phosphatidylethanolamine (PE) measured by $^{2}$H NMR~\cite{ghosh82}.
+are in good agreement with the results for the phosphatidylethanolamine (PE) measured by $^{2}$H NMR~\cite{ghosh82}.
 These results further support the idea that the glycerol backbone structural behaviour is independent of the
@@
   \centering
-  \includegraphics[width=8.6cm]{OrderParameterCHOLexpVSsim.eps}
+%  \includegraphics[width=8.6cm]{OrderParameterCHOLexpVSsim.eps}
   \caption{\label{ordPchol}
@@
     with the Berger-POPC-07/H\"oltje-CHOL-13, CHARMM36 and MacRog force fields. The signs in the experimental values are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
-    and Gross et al.~\cite{gross97}.  The most order parameters from Berger-POPC-07/H\"oltje-CHOL-13 model for g$_1$ are beoynd the y-axis scale.}
+    and Gross et al.~\cite{gross97}.  Most order parameters from Berger-POPC-07/H\"oltje-CHOL-13 model for g$_1$ are beoynd the y-axis scale.}
 \end{figure}
@@
 In contrast, the responses of both CHARMM36 and MacRog are in better agreement
-with experiments, however CHARMM36 seems to better reproduce the experimentally observed modest changes in the glycerol backbone segments 
-g$_2$ and g$_3$ with high concetrations of cholesterol. Thus we have calculated the glycerol backbone dihedral angle distributions
+with experiments. However, CHARMM36 seems to better reproduce the experimentally observed modest changes in the glycerol backbone segments 
+g$_2$ and g$_3$ with high concentrations of cholesterol. Thus we have calculated the glycerol backbone dihedral angle distributions
 as a function of cholesterol in CHARMM36 (shown in Supplementary material) to resolve the cholesterol induced structural changes. The only detectable change due to the
-addition of cholesterol is the small decrease of gaughe- and increase of gaughe+ probability of g3-g2-g1-O(sn-1) dihedral.
+addition of cholesterol is the small decrease of gauche- and increase of gaughe+ probability of g3-g2-g1-O(\textit{sn}-1) dihedral.
 
-It should be noted that the CHARMM36 force field parameters (dihedral potentials) for the glycerol backbone have been tuned 
-the dihedral potentials to reproduce the correct order parameters at fully hydrated conditions~\cite{klauda10}. 
+It should be noted that the CHARMM36 force field parameters (dihedral potentials) for the glycerol backbone have been tuned to reproduce the correct order parameters at fully hydrated conditions~\cite{klauda10}.
 This procedure contains a risk of overfitting, which would manifest itself as wrong responses to changing conditions. 
-According to our results, tuning seems not to lead to overfitting problems in the case of dehydration or lipid-cholesterol mixtures.
+According to our results, this tuning does not seem to lead to overfitting problems in the case of dehydration or lipid--cholesterol mixtures.
 
@@
 in phoshatidylcholine bilayers are not known despite of vast amount of accurate experimental
-data. Atomistic resolution molecular dynamics simulation model which would reproduce the
+data. Atomistic resolution molecular dynamics simulation model which reproduced the
 experimental data would automatically resolve the structures giving an interpretation of experimental results.
-In this work we have collected and reviewed the experimental C-H bond vector order
-parameters available in literature. These experimental parameters are then compared to
-different atomistic resolution simulation models for fully hydrated bilayer, dehydrated bilayer and
-lipid bilayer containing cholesterol. Our results have led to the following conclusions:
+In this work we have collected and reviewed the experimental C--H bond vector order
+parameters available in the literature. These experimental parameters are then compared to
+different atomistic resolution simulation models for a fully hydrated bilayer, bilayers dehydrated to different extents, and
+lipid bilayers containing various amounts of cholesterol. Our results have led to the following conclusions:
 
-- The C-H bond order parameters measured with different NMR techniques are in good agreement
+- The C--H bond order parameters measured with different NMR techniques are in good agreement
 with each others. By combining the experimental results from various sources we concluded
-that the order parameters for each C-H bond are known with quantitative accuracy of $\pm$0.02.
+that the order parameters for each C--H bond are known with a quantitative accuracy of $\pm$0.02.
 
 - None of the tested models (14 different models) produces the order parameters with the experimental
-accuracy for fully hydrated phoshatidylcholine lipid bilayer. However, the CHARMM36, GAFFlipid and MacRog 
+accuracy for a fully hydrated phoshatidylcholine lipid bilayer. However, the CHARMM36, GAFFlipid and MacRog 
 models are relatively close. The structures of these models together with the most used lipid model (Berger based) 
@@
 However, the correlation between dihedral angle distributions and order parameter differences was found, 
-suggesting that careful adjustment of dihedral potentials would potentially lead to the model with correct
+suggesting that careful adjustment of dihedral potentials would potentially lead to the model with a correct
 structure.
 
-- Independent of the accuracy for fully hydrated lipid bilayer, all the models reproduced the choline response
-to the dehydration. This can be explained by the change in the P-N vector tilting more parallel to the membrane
-which leads to the increase of order parameters despite of the initial configuration. It should be however noted
-that the correct qualitative response do not necessarily indicate correct energetics.
+- Independent of the accuracy for a fully hydrated lipid bilayer, all the models reproduced the choline response
+to dehydration. This can be explained by the change in the P--N vector tilting more parallel to the membrane,
+which leads to the increase of order parameters despite of the initial configuration. It should be noted, however,
+that the correct qualitative response does not necessarily indicate correct energetics.
 
@@
 is extremely convenient tool to give structural interpretation for the high resolution NMR data~\cite{ferreira14}. 
-However, in the case of phoshatidylcholine glycerol backbone and choline headgroup there is some
-further model development required.
+However, in the case of phoshatidylcholine glycerol backbone and choline headgroup further model development is still required.
 
-This work has been, and continues to be, progressed and discussed through the blog: nmrlipids.blogspot.fi. 
+This work has been, and continues to be, progressed and discussed through a blog at \texttt{nmrlipids.blogspot.fi}. 
 Everyone is invited to join the discussion and make contributions through the blog. 
@@
 Everyone who has contributed to the work through the blog will be offered 
-coauthorship. For more details see: nmrlipids.blogspot.fi.
+coauthorship. For more details, see \texttt{nmrlipids.blogspot.fi}.
 
@@
 \subsubsection{Berger based models}
-For the berger based models we use here the following naming convention: 
+For the Berger based models we use here the following naming convention: 
 Berger - \{{\it molecule name}\} - \{{\it year when model published first time}\} \{{\it citation}\}.
 The reason is that there are several different molecular topologies which are using the non-bonded parameters originally
-developed by Berger et al.~\cite{berger97}. Thus the common factor in the berger based models are the non-bonded parameters,
+developed by Berger et al.~\cite{berger97}. Thus the common factor in the Berger based models are the non-bonded parameters,
 while the molecule specific parameters might somewhat vary. However, the majority of the molecular level topologies are 
@@
 
-POPC simulations at full hydration in 298 K and simulations as a function of cholesterol are the same as previous publications~\cite{ferreira13,ferreira15}.
-In these simulation the POPC parameters introduced by Ollila et al~\cite{ollila07a} are used, which are using Berger non-bonded parameters~\cite{berger97}
-and molecular topology is from Tieleman et al.~\cite{tieleman99} with improved double bond dihedrals by Bachar et al.~\cite{bachar04}. 
+POPC simulations at full hydration at 298~K and simulations studying the effect of cholesterol are the same as in previous publications~\cite{ferreira13,ferreira15}.
+In these simulation the POPC parameters introduced by Ollila et al~\cite{ollila07a} are used, which are using the non-bonded parameters of Berger~\cite{berger97}
+and a molecular topology from Tieleman et al.~\cite{tieleman99} with improved double bond dihedrals by Bachar et al.~\cite{bachar04}. 
 Thus they are called Berger-POPC-07~\cite{ollila07a}. The cholesterol model is based on the parameters by H\"oltje et al.~\cite{holtje01} with the
-exeption that the atom types were changed from CH2/CH3 to LP2/LP3 to avoid overcondensation of the bilayer as suggested in ref.~\cite{tieleman06}.
+exception that the atom types were changed from CH2/CH3 to LP2/LP3 to avoid overcondensation of the bilayer as suggested in ref.~\cite{tieleman06}.
 Since this modification was introduced by Ferreira et al.~\cite{ferreira13}, we call the used cholesterol model as H\"oltje-CHOL-13~\cite{ferreira13}.
@@
 For DPPC the implementation of Berger parameters~\cite{berger97} by Peter Tieleman et al. are used~\cite{marrink98}.
-For all of these simulations the timestep of 2~fs was used with leap-frog integrator. Covalent bond lengths were constrained with LINCS algorithm~\cite{hess97,hess07}. 
-Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with real space cut-off at 1.0~nm was used 
+For all of these simulations a timestep of 2~fs was used with a leapfrog integrator. Covalent bond lengths were constrained with the LINCS algorithm~\cite{hess97,hess07}. 
+Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with real space cut-off of 1.0~nm was used 
 for electrostatics. Plain cut-off was used for the Lennard-Jones interactions with a 1.0~nm cut-off.
-The neighbour lists were updated every 5th step with cut-off at 1.0~nm. Temperature was coupled separately
-for lipids and water to 298~K with the velocity-rescale method~\cite{bussi07} with coupling constant 0.1~ps.
+The neighbor lists with cut-off of 1.0~nm were updated every 5 steps. Temperature was coupled separately
+for lipids and water to 298~K using the velocity-rescale method~\cite{bussi07} with coupling constant 0.1~ps.
 Pressure was semi-isotropically coupled to the atmospheric pressure with the Berendsen method~\cite{berendsen84}.
@@
 
-Timestep of 1~fs was used with leap-frog integrator. Covalent bonds with hydrogens were constrained with LINCS algorithm~\cite{hess97,hess07}. 
-Coordinates were written every 5~ps. PME~\cite{darden93,essman95} with real space cut-off at 1.4~nm was used 
+Timestep of 1~fs was used with the leapfrog integrator. Covalent bonds with hydrogens were constrained with LINCS algorithm~\cite{hess97,hess07}. 
+Coordinates were written every 5~ps. PME~\cite{darden93,essman95} with real space cut-off of 1.4~nm was used 
 for electrostatics. Lennard-Jones interactions were switched to zero between 0.8~nm and 1.2~nm.
-The neighbour lists were updated every 5th step with cut-off 1.4~nm. Temperature was coupled separately
-for lipids and water to 303~K with the velocity-rescale method~\cite{bussi07} with coupling constant 0.2~ps.
+The neighbour lists with a cut-off of 1.4~nm were updated every 5 steps. Temperature was coupled separately
+for lipids and water to 303~K using the velocity-rescale method~\cite{bussi07} with coupling constant of 0.2~ps.
 Pressure was semi-isotropically coupled to the atmospheric pressure with the Berendsen method~\cite{berendsen84}.
@@
 The starting structures for the pure POPC and DOPC simulations was taken from the Slipids~\cite{jambeck12b} website (http://people.su.se/$\sim$jjm/Stockholm\_Lipids/Downloads.html).
-The starting structures for mixed POPC/Cholesterol simulations were constructed with the ChARMM-GUI website~\cite{jo08}. 
-They contained 100 POPC/24 Cholesterol molecules and 80 POPC/80 Cholesterol molecules for
-the simulations of 20\% Cholesterol and 50\% Cholesterol respectively. The TIP3P water model~\cite{jorgensen83} was used to
+The starting structures for mixed POPC/Cholesterol simulations were constructed with the CHARMM-GUI website~\cite{jo08}. 
+They contained 100 POPC/24 cholesterol molecules and 80 POPC/80 cholesterol molecules for
+the simulations of 20\% cholesterol and 50\% cholesterol respectively. The TIP3P water model~\cite{jorgensen83} was used to
 solvate the system.
-The publicly available CHARMM36 forcefield parameters (http://www.gromacs.org/@api/deki/files/184/=charmm36.ff\\\_4.5.4\_ref.tgz) 
+The publicly available CHARMM36 forcefield parameters (\texttt{http://www.gromacs.org/@api/deki/files/184/=charmm36.ff\\\_4.5.4\_ref.tgz}) 
 by Piggot et al. \cite{piggot12} were used. Cholesterol parameters came
-from Lim et al. \cite{lim12} and were converted into Gromacs files by using the PyTopol tool (https://github.com/resal81/PyTopol).  
+from Lim et al. \cite{lim12} and were converted into GROMACS format with the PyTopol tool (\texttt{https://github.com/resal81/PyTopol}).  
 Single point energy calculation was done to assess the conversion. 
@@
 %http://mackerell.umaryland.edu/charmm_ff.shtml#gromacs).
-Simulations were performed for 200ns and the last 100ns was used for the calculations. Timestep of 2fs was
-used with leap-frog integrator. All bond lengths were constrained with LINCS~\cite{hess97,hess07}. Temperature was maintened at
-303 K with the velocity-rescale method~\cite{bussi07} and a time constant of 0.2 ps. Pressure was maintained semiisotropically
-at 1 bar using the Parrinello-Rahman algorithm~\cite{parrinello81} with a time constant of 1.0 ps. The neighbour list
-was updated every 10th step with a cut-off of 1.2 nm. Lennard-Jones interactions were switched to zero
-between 0.8 nm and 1.2 nm. PME~\cite{darden93,essman95} with real space cut-off at 1.2nm was used for electrostatics.
+Simulations were performed for 200~ns and the last 100~ns was used for the calculations. Timestep of 2~fs was
+used with leapfrog integrator. All bond lengths were constrained with LINCS~\cite{hess97,hess07}. Temperature was maintened at
+303~K with the velocity-rescale method~\cite{bussi07} and a time constant of 0.2~ps. Pressure was maintained semiisotropically
+at 1~bar using the Parrinello--Rahman algorithm~\cite{parrinello81} with a time constant of 1.0~ps. The neighbour list with a cut-off of 1.2~nm
+was updated every 10 steps. Lennard-Jones interactions were switched to zero
+between 0.8~nm and 1.2~nm. PME~\cite{darden93,essman95} with real space cut-off of 1.2~nm was used for electrostatics.
 
@@
 The initial structure in Lipidbook \cite{domanski10} had different glycerol backbone isomers in different leaflets. 
-To generate the initial structure we took the structure delivered by Slipid developers~\cite{jambeck12b}. Also this structure
-had one lipid with different glycerol bakcbone isomer. This lipid and one lipid from opposite leaflet were removed
+To generate the initial structure we took the structure delivered by Slipids developers~\cite{jambeck12b}. Also this structure
+had one lipid with different glycerol backbone isomer. This lipid and one lipid from opposite leaflet were removed
 after the system was equilibrated.
 
-The force field parameters were generated using files obtained from the Lipidbook website (http://lipidbook.bioch.ox.ac.uk/package/show/id/150.html)~\cite{domanski10}. 
-The conversion to Gromacs compatible formats was performed using the acpype tool~\cite{silva12}. The accuracy of the conversion was checked by calculating 
-the total energy of a single POPC lipid molecule using the sander program which is part of the AmberTools14 package~\cite{ferrer13} and Gromacs 4.6.5. 
+The force field parameters were generated using files obtained from the Lipidbook website (\texttt{http://lipidbook.bioch.ox.ac.uk/package/show/id/150.html})~\cite{domanski10}. 
+The conversion to GROMACS compatible formats was performed using the acpype tool~\cite{silva12}. The accuracy of the conversion was checked by calculating 
+the total energy of a single POPC lipid molecule using the sander program which is part of the AmberTools14 package~\cite{ferrer13} and version 4.6.5 of GROMACS. 
 A difference of 0.002 kcal/mol was obtained between the two programs.
@@
 Timestep of 2~fs was used in Langevin dynamics with zero friction term and collision frequency of 1.0~ps$^{-1}$. 
-Covalent bonds with hydrogens were constrained with LINCS algorithm~\cite{hess97,hess07}.
-Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with real space cut-off at 1.0~nm was used 
+Covalent bonds with hydrogens were constrained with the LINCS algorithm~\cite{hess97,hess07}.
+Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with a real space cut-off at 1.0~nm was used 
 for electrostatics. Plain cut-off with 1~nm was used for Lennard-Jones interactions. 
-The neighbour lists were updated every 5th step with cut-off 1.0~nm. 
-Pressure was semi-isotropically coupled to the 1~bar pressure with the Berendsen method~\cite{berendsen84}.
+The neighbour lists with a cut-off of 1.0~nm were updated every 5 steps. 
+Pressure was semi-isotropically coupled to a pressure of 1~bar with the Berendsen method~\cite{berendsen84}.
+
 \subsubsection{Lipid14}
@@
 Covalent bonds with hydrogens were constrained with LINCS algorithm~\cite{hess97,hess07}.
-Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with real space cut-off at 1.0~nm was used 
-for electrostatics. Plain cut-off with 1~nm was used for Lennart-Jones interactions. Dispersion correction
-was used for energy and pressure. The neighbour lists were updated every 5th step with cut-off 1.0~nm. 
-Pressure was semi-isotropically coupled to the 1~bar pressure with the Berendsen method~\cite{berendsen84}.
+Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with real space cut-off of 1.0~nm was used 
+for electrostatics. Plain cut-off with 1~nm was used for Lennard-Jones interactions. Dispersion correction
+was applied for both energy and pressure. The neighbor lists with a cut-off of 1.0~nm were updated every 5 steps. 
+Pressure was semi-isotropically coupled to a pressure of 1~bar with the Berendsen method~\cite{berendsen84}.
 
 \subsubsection{Poger et al.}
-The Poger lipids are derived from GROMOS G53A6~\cite{poger10} and were initially coined 53A6-L (L for lipids), and are now part of GROMOS G54A7~\cite{poger12}. 
-They work with the SPC water model~\cite{berendsen81}. The initial hydrated bilayer structure of 128 DPPC/5841 water molecules as well as force field parameters were downloaded 
-from David Poger's web site (http://compbio.chemistry.uq.edu.au/~david/) on April 2012. 
+The Poger lipids are derived from GROMOS G53A6~\cite{poger10} and were initially coined 53A6-L (L for lipids).
+They are now part of GROMOS G54A7~\cite{poger12} and parametrized to work with the SPC water model~\cite{berendsen81}.
+The initial hydrated bilayer structure of 128 DPPC and 5841 water molecules as well as force field parameters were downloaded
+from David Poger's web site (\texttt{http://compbio.chemistry.uq.edu.au/~david/}) on April 2012. 
 We noticed that the same files downloaded in October 2013 appear to lack two dihedral angles in the choline headgroup (only one dihedral of type gd\_29 allowing 
 the rotation of the 3 choline methyls) compared to the April 2012 version (3 dihedrals of type gd\_29 for the 3 choline methyls). This should not affect the 
-bilayer structure and only change the kinetics of the choline methyls rotation, however the October 2013 version has not been tested.
+bilayer structure and only change the kinetics of the choline methyls rotation. However the October 2013 version has not been tested in this study.
 
-MD Simulations (two repetitions with independent initial velocities) were run for 100 ns using a 2 fs time step and the analysis 
-was performed on the last 50 ns. Coordinates were saved every 50 ps for analysis. All bond lengths were constrained with the LINCS algorithm~\cite{hess97,hess07}. Temperature was kept 
-at 323K with the v-rescale~\cite{bussi07} thermostat with a time constant of 0.1 ps (DPPC and water coupled separetly). Pressure was maintained semi-isotropically at 1 bar using 
-the Parrinello-Rahman barostat~\cite{parrinello81} using a 4 ps time constant and a compressibility of 4.5e-5 bar$^{-1}$. For non-bonded interactions, two conditions were tested:
-i) A 0.8-1.4 nm twin-range cutoff with the neighbor list updated every 5 steps for both electrostatics and Lennard-Jones. For the former the generalized reaction 
-field (RF) with a dielectric permitivity of 62 was used beyond the 1.4 nm cutoff~\cite{tironi95}. This is the original setup that Poger et al.~\cite{poger10} used.
-ii) PME~\cite{darden93,essman95} electrostatics with a real space cutoff of 1.0 nm, a Fourier spacing of 0.12 nm and an interpolation order of 4, LJ computed with a 1.0-1.4 nm twin-range cutoff, 
-neighbor list updated every 5 steps. Note that Poger and Mark tested the effect of PME vs RF in the ref.~\cite{poger12}, but used a 1.0 nm cutoff with PME and 1.4 nm with RF for LJ 
-interactions. Since 0.8-1.4 nm twin-range cutoff for LJ interations is used in the parametrization of GROMOS force field we decided to use that
+MD Simulations (two repetitions with independent initial velocities) were run for 100~ns using a 2~fs time step and the analysis 
+was performed on the last 50~ns. Coordinates were saved every 50~ps for analysis. All bond lengths were constrained with the LINCS algorithm~\cite{hess97,hess07}. Temperature was kept 
+at 323~K employing the velocity-rescale~\cite{bussi07} thermostat with a time constant of 0.1~ps (DPPC and water coupled separetly). Pressure was maintained semi-isotropically at 1~bar using 
+the Parrinello--Rahman barostat~\cite{parrinello81} using a 4~ps time constant and a compressibility of 4.5e-5~bar$^{-1}$. For non-bonded interactions, two conditions were tested:
+i) A 0.8--1.4 nm twin-range cut-off with the neighbor list updated every 5 steps for both electrostatics and Lennard-Jones (LJ) interactions. For the former the generalized reaction 
+field (RF) with a dielectric permittivity of 62 was used beyond the 1.4~nm cut-off~\cite{tironi95}. This is the original setup that Poger et al.~\cite{poger10} used.
+ii) PME~\cite{darden93,essman95} electrostatics with a real space cut-off of 1.0~nm, a Fourier spacing of 0.12~nm and an interpolation order of 4, LJ interactions computed with a 1.0--1.4~nm twin-range cut-off,
+neighbor list updated every 5 steps. Note that Poger and Mark tested the effect of PME vs RF in ref.~\cite{poger12}, but used a 1.0~nm cut-off with PME and 1.4~nm with RF for LJ
+interactions. Since 0.8--1.4~nm twin-range cut-off for LJ interations is used in the parametrization of GROMOS force field, we decided to use that
 also in the simulations with PME.
@@
 
-Since Poger lipids come from GROMOS force field, it is important to note that GROMOS uses the RF scheme for computing electrostatics (this is the method used for the 
-force field parameterization). Using setup i) based on RF, we were able to reproduce the results (i.e. area per lipid~0.63 nm$^2$) from the original work only with 
-GROMACS version $<=$ 4.0.* (the original authors~\cite{poger10} used GROMACS version 3.3.3). On going to versions $>=$ 4.5.*, the area per lipid dropped below 0.58 nm$^2$. 
-The GROMACS developers were contacted and a redmine issue opened (http://redmine.gromacs.org/issues/1400). The difference comes from the new Trotter decomposition 
-introduced in version 4.5. A fix has been introduced in version 4.6.6 that allows a recovery of ~0.615 nm$^2$. The results in terms of area per lipid using the different 
-GROMACS versions are here~\cite{fuchs14}.
+Since Poger lipids come from the GROMOS force field, it is important to note that GROMOS uses the RF scheme for computing electrostatics (this is the method used for the 
+force field parameterization). Using setup i) based on RF, we were able to reproduce the results (i.e. area per lipid value of 0.63~nm$^2$) from the original work only with 
+GROMACS versions 4.0.X and earlier (the original authors~\cite{poger10} used GROMACS version 3.3.3). When switching to versions 4.5.X and above, the area per lipid dropped to below 0.58~nm$^2$. 
+The GROMACS developers were contacted and a redmine issue opened (\texttt{http://redmine.gromacs.org/issues/1400}). The difference comes from the new Trotter decomposition 
+introduced in versions 4.5.X. A fix has been introduced in version 4.6.6 that allows a recovery of an area per lipid value of 0.615~nm$^2$. The results in terms of area per lipid using the different 
+GROMACS versions are available at~\cite{fuchs14}.
 % http://www.dsimb.inserm.fr/~fuchs/project\_Samuli/Poger\_DPPC/tests\_gmx\_versions/)[Samuli, you may want to add this Figure+Legend+Comment 
@@
 %(thus with explicit hydrogens) showed that this setup is still unstable and eventually leads to crashes with versions 4.6.6 and 4.6.7 (and so does version 4.5.*).
-Thus we decided to use only the PME setup ii) for computing the order parameter since it gives stable results whatever the GROMACS version. We obtained an area per 
-lipid of ~0.615 nm$^2$, below 0.648 nm$^2$ found by the original authors with their PME setup (see~\cite{poger12}). We explained that by the fact we used 
-a 1.4 nm for the LJ cutoff and they used 1.0 nm.
+Thus we decided to use only the PME setup ii) for computing the order parameter since it gives stable results regardless of the GROMACS version. We obtained an area per 
+lipid of 0.615~nm$^2$, below 0.648 nm$^2$ found by the original authors with their PME setup (see~\cite{poger12}). We explained that by the fact that we used 
+a 1.4~nm for the LJ cut-off whereas a value of 1.0~nm was used in the original publication.
 
-\subsubsection{Slipid}
+\subsubsection{Slipids}
 %The used simulation files and part of the trajectroy are available~\cite{slipidsFILES}.
 
-Initial coordinates for a hydrated DPPC bilayer (30 waters/lipid) at 323K were taken directly
-from http://people.su.se/\~jjm/Stockholm\_Lipids/Downloads.html  The Slipids force field~\cite{jambeck12} was used for the the all atom description of DPPC, and
-water was described with the TIP3P water model~\cite{jorgensen83}. Simulations were performed within the NPT ensemble using the GROMACS 4.6.1 simulation
-package~\cite{hess08}. The nose--hoover rescaling thermostat~\cite{nose84,hoover85} was used with reference temperature of
-323 K and a relaxation time constant of 0.5 ps. Water and lipids were coupled separately to
-the heat bath. Pressure was kept constant at 1.013 bar using a semi--isotropic Parinello--Rahman
-barostat~\cite{parrinello81} with a time constant of 10.0 ps. Equations of motion were
-integrated with the leapfrog algorithm using a timestep of 2 fs. Long range
+Initial coordinates for a hydrated DPPC (at 323~K) and POPC (at 310~K) bilayers (30 and 40 waters/lipid, respectively) were taken directly from the Slipids home page \texttt{http://people.su.se/\~jjm/Stockholm\_Lipids/Downloads.html}.
+The Slipids force field~\cite{jambeck12,XXX} ADD 2ND SLIPIDS PAPER!
+was used for the the all atom descriptions of DPPC and POPC, and
+water was described with the TIP3P water model~\cite{jorgensen83}. Simulations were performed within the NPT ensemble using the GROMACS 4.6.X simulation
+package~\cite{hess08}. The Nos\'{e}--Hoover thermostat~\cite{nose84,hoover85} was used with reference temperatures of
+323~K (DPPC) and 310~K (POPC) and a relaxation time constant of 0.5~ps. Water and lipids were coupled separately to
+the heat bath. Pressure was kept constant at 1.013~bar using a semi--isotropic Parrinello--Rahman
+barostat~\cite{parrinello81} with a time constant of 10.0~ps. Equations of motion were
+integrated with the leapfrog algorithm using a timestep of 2~fs. Long range
 electrostatic interactions were calculated using the PME method~\cite{darden93,essman95}, with a fourth order
-smoothing spline. A real space cutoff of 1.0 nm was employed with grid spacing of 0.12 in the reciprocal space.
-Lennard--Jones potentials were cutoff at 1.4 nm, with dispersion correction applied both to
-energy and pressure. All covalent bonds in lipids were constrained using the LINCS algorithm~\cite{hess97}, 
-whereas water molecules were constrained using SETTLE~\cite{miyamoto92}. Twinrange cutoffs,
-1.0 nm and 1.6 nm, were used for the neighborlists with the longrange neighbor list updated every
+smoothing spline. A real space cut-off of 1.0~nm was employed with grid spacing of 0.12~nm in the reciprocal space.
+Lennard-Jones potentials were cut off at 1.4~nm, with a dispersion correction applied to both energy and pressure.
+All covalent bonds in lipids were constrained using the LINCS algorithm~\cite{hess97},
+whereas water molecules were constrained using SETTLE~\cite{miyamoto92}. Twin-range cutoffs,
+1.0~nm and 1.6~nm, were used for the neighbor lists with the longrange neighbor list updated every
 10 steps. This simulation protocol corresponds to the protocol used in Ref~\cite{jambeck13}. 
 %The simulations were run for XXX ns with the last XXX ns used for calculating the order parameters.
+
 \subsubsection{Kukol}
@@
 The force field parameters and the initial configuration were available through the Lipidbook~\cite{domanski10}.
-Timestep of 2~fs was used with leap-frog integrator. Covalent bond lengths were constrained with LINCS algorithm~\cite{hess97,hess07}. 
-Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with real space cut-off at 1.0~nm was used 
-for electrostatics. Twin range cut-off was used for the Lennardt-Jones interactions with short and long cut-offs at 1.0~nm and 1.6~nm, respectively.
-The neighbour lists were updated every 5th step with cut-off at 1.0~nm. Temperature was coupled separately
-for lipids and water to 298~K with the velocity-rescale method~\cite{bussi07} with coupling constant 0.2~ps.
-Pressure was semi-isotropically coupled to the atmospheric pressure with the Parrinello-Rahman method~\cite{parrinello81}.
+Timestep of 2~fs was used with leapfrog integrator. Covalent bond lengths were constrained with LINCS algorithm~\cite{hess97,hess07}. 
+Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with real space cut-off of 1.0~nm was used 
+for electrostatics. Twin range cut-off was used for the Lennard-Jones interactions with short and long cut-offs of 1.0~nm and 1.6~nm, respectively.
+The neighbour lists with a cut-off of 1.0~nm were updated every 5 steps. Temperature was coupled separately
+for lipids and water to 298~K with the velocity-rescale method~\cite{bussi07} with a coupling constant 0.2~ps.
+Pressure was semi-isotropically coupled to the atmospheric pressure with the Parrinello--Rahman method~\cite{parrinello81}.
 %\subsubsection{H\"ogberg et al. and Rabonovich et al.}
@@
 Additionally, the simulations were repeated with the dispersion correction applied to pressure and temperature. Even though the area per lipid decrease
-d slightly, the head group order parameters were only slightly affected.
+d slightly, the headgroup order parameters were only slightly affected.
 
@@
 
-The gel phase structure delivered by Tj\"ornhammar  and edholm~\cite{tjornhammar14} was ran a 5ns at 70 degrees in order to destroy the 
-ordered gel configuration. This was followed by 200ns simulation at 50 degrees. The last 100ns of this simulation was used for analysis. 
-The same mdp file as in the paper's~\cite{tjornhammar14} SI was used except for the temperatures.
+The gel phase DPPC bilayer structure delivered by Tj\"ornhammar  and Edholm~\cite{tjornhammar14} was ran for 5~ns at 343~K in order to destroy the 
+ordered gel configuration. This was followed by a 200~ns simulation at 323~K, i.e. in the fluid phase. The last 100~ns of this simulation was used for analysis. 
+The same mdp file as in the Supplementary Information section of the original paper ~\cite{tjornhammar14} was used except for the simulation temperature.
 
 \subsubsection{CHARMM36-UA}
-A hydrated bilayer consisting of 128 DLPC lipids and 3840 water molecules is modeled by the force field of Lee and co-workers~\cite{lee14}, 
-which is a combination of the all-atom CHARMM36 force-field~\cite{klauda10} and the united-atom Berger model~\cite{berger97}. 
-The nonbonding interactions are calculated using an atom-based switching function with inner and outer cutoffs of 0.8 and 1.2 nm~\cite{lee14}. 
-Long range electrostatic interactions are implemented using the particle-particle particle-mesh solver with a relative accuracy of 0.0001. The system 
-is first equilibrated for 30 ns in the NP$\gamma$T ensemble (Nose-Hoover~\cite{nose84,hoover85} style thermostat and barostat with anisotropic pressure coupling) 
-at 323 K and 1 bar with timestep of 1 fs, the next 20 ns of dynamics are taken for calculation of configurational averages. 
+A hydrated bilayer consisting of 128 DLPC lipids and 3840 water molecules is modeled by the force field of Lee and co-workers~\cite{lee14}.
+This force field is a combination of the all-atom CHARMM36 force-field~\cite{klauda10} and the united-atom Berger model~\cite{berger97}. 
+The non-bonded interactions are calculated using an atom-based switching function with short and long cut-offs of 0.8 and 1.2~nm~\cite{lee14}. 
+Long range electrostatic interactions are implemented using the particle-particle particle-mesh solver with a relative accuracy of $10^{-4}$. The system 
+is first equilibrated for 30~ns in the NP$\gamma$T ensemble (Nos\'{e}--Hoover~\cite{nose84,hoover85} style thermostat and barostat with anisotropic pressure coupling) 
+at 323~K and 1~bar with timestep of 1~fs. The following 20~ns of dynamics are taken for calculation of configurational averages. 
 \todo{The simulation times given currently contradict with the Table~\ref{systems}.}
@@
   \centering
-  \includegraphics[width=17.2cm]{dihsCHOLcharmm.pdf}
+%  \includegraphics[width=17.2cm]{dihsCHOLcharmm.pdf}
   \caption{\label{dihsCHOLcharmm}

### 2015-01-20 HGmodel_draft3.tex
new version

@@
   \centering
-%  \includegraphics[width=8.6cm]{POPCstructure.eps}
+  \includegraphics[width=8.6cm]{POPCstructure.eps}
 
@@
 Berger-POPC-07~\cite{ollila07a}          &   POPC & 128 & 7290  & 298  & 270 & 240 & \cite{bergerFILESpopc}$^*$ & \cite{ferreira15} \\
-\sout {Berger-POPC-07~\cite{ollila07a}}\todoi{Not included in the Fig. 2, to be removed}          &   POPC & 128 & 7290  & 323  & 56 & 56  & ? & SI \\
+\sout {Berger-POPC-07~\cite{ollila07a}}\todoi{Not included in the Fig. 2, to be removed?}          &   POPC & 128 & 7290  & 323  & 56 & 56  & ? & SI \\
 Berger-DPPC-98~\cite{marrink98}          &   DPPC & 72 & 2864  & 323  & 140 & 100  & \cite{bergerDPPCfiles} & SI \\
@@
 CHARMM36\cite{klauda10}       & POPC   & 72  &  2242 & 303 & 30 & 20  & \cite{charmm36filesSHORT}$^*$ & SI \\
-CHARMM36\cite{klauda10}\todoi{Hubert Santuz, these files are already in Gitbub, however to get doi citation it should be put to Zenodo}        & POPC   & 128 &  5120    & 303 & 150 & 100  & ? & SI \\
-CHARMM36\cite{klauda10}\todoi{Samuli, put to Zenodo}        & DPPC   & 72  &  2189 & 323 & 30 & 25  & ? & SI \\
-CHARMM36\cite{klauda10}\todoi{Not included in the Fig. 2. Should we remove or include?}        & DOPC   & 128 &  5120    & 303 & 150 & 100  & ? & SI \\
+CHARMM36\cite{klauda10}      & POPC   & 128 &  5120    & 303 & 150 & 100  & ? \todoi{Permanent link in progress by Hubert Santuz}   & SI \\
+CHARMM36\cite{klauda10}       & DPPC   & 72  &  2189 & 323 & 30 & 25  & ?\todoi{Samuli, put to Zenodo}  & SI \\
+CHARMM36\cite{klauda10}\todoi{Not included in the Fig.
+2.
+Should we remove or include?}        & DOPC   & 128 &  5120    & 303 & 150 & 100  & ?\todoi{Permanent link in progress by Hubert Santuz}  & SI \\
 MacRog\cite{maciejewski14}  & POPC & 288  & 12600 & 310 & 100 & 80  & \cite{macrogFILES}$^*$ & SI  \\
@@
 Lipid14\cite{dickson14}         & POPC  & 72 & 2234 & 303 & 100 & 50  & \cite{lipid14files}$^*$ & SI \\
-Poger\cite{poger10}             & DPPC  & 128 & 5841 & 323 & 2$\times$100 & 2$\times$50 & \cite{pogerFILES} & SI \\
+Poger\cite{poger10}             & DPPC  & 128 & 5841 & 323 & 2$\times$100 & 2$\times$50 & \cite{pogerFILES}\todoi{Permanent link in progress by Patrick Fuchs} & SI \\
 Slipid\cite{jambeck12}          & DPPC & 128 & 3840 & 323 & 150 & 100 & \cite{slipidsFILES}$^*$ & SI \\
@@
   \centering
-%  \includegraphics[width=8.6cm]{comparison3.pdf}
+  \includegraphics[width=8.6cm]{comparison3.pdf}
 %\todo{DONE}\\
@@
   \centering
-%  \includegraphics[width=8.6cm]{HGorderparameters5.eps} \\
+  \includegraphics[width=8.6cm]{HGorderparameters5.eps} \\
   \todo{Markus Miettinen suggested that we should consider ``spreading'' the values similarly to the figure~\ref{HGorderparameters}} \\
@@
   \centering
-%  \includegraphics[width=8.6cm]{g1-g2_Cdihs2.eps}
+  \includegraphics[width=8.6cm]{g1-g2_Cdihs2.eps}
   \caption{\label{dihDISTS}
@@
   \centering
-%  \includegraphics[width=8.6cm]{a-bDIHS2.eps}
+  \includegraphics[width=8.6cm]{a-bDIHS2.eps}
   \caption{\label{dihDISTS2}
@@
   \centering
-%  \includegraphics[width=8.6cm]{OrderParameterDEHYDpresentation.eps}
+  \includegraphics[width=8.6cm]{OrderParameterDEHYDpresentation.eps}
 \todo{Several people think that this plot should be clarified.}
@@
   \centering
-%  \includegraphics[width=8.6cm]{PNangles.eps}
+  \includegraphics[width=8.6cm]{PNangles.eps}
 
@@
   \centering
-%  \includegraphics[width=8.6cm]{OrderParameterCHOLexpVSsim.eps}
+  \includegraphics[width=8.6cm]{OrderParameterCHOLexpVSsim.eps}
   \caption{\label{ordPchol}
@@
 introduced in versions 4.5.X. A fix has been introduced in version 4.6.6 that allows a recovery of an area per lipid value of 0.615~nm$^2$. The results in terms of area per lipid using the different 
-GROMACS versions are available at~\cite{fuchs14}.
+GROMACS versions are available at~\cite{fuchs14}\todoi{Permanent link in progress by Patrick Fuchs}.
 % http://www.dsimb.inserm.fr/~fuchs/project\_Samuli/Poger\_DPPC/tests\_gmx\_versions/)[Samuli, you may want to add this Figure+Legend+Comment 
@@
 at 323~K and 1~bar with timestep of 1~fs. The following 20~ns of dynamics are taken for calculation of configurational averages. 
-\todo{The simulation times given currently contradict with the Table~\ref{systems}.}
-Simulations were 
-carried out by using the LAMMPS package~\cite{plimpton95}. %( http://doi.org/10.1006/jcph.1995.1039 )., the input files are available ( http://dx.doi.org/10.5281/zenodo.13821 ).
+Simulations were carried out by using the LAMMPS package~\cite{plimpton95}. %( http://doi.org/10.1006/jcph.1995.1039 )., the input files are available ( http://dx.doi.org/10.5281/zenodo.13821 ).
 
@@
   \centering
-%  \includegraphics[width=17.2cm]{dihsCHOLcharmm.pdf}
+  \includegraphics[width=17.2cm]{dihsCHOLcharmm.pdf}
   \caption{\label{dihsCHOLcharmm}
@@
 \end{figure*}
+
+\subsection{General issues to think of about the manuscript}
+\todo{Should we emphasize more the excellent collection of simulation trajectories collected to Zenodo?}
+\todo{Maybe we should mention that CHARMM is significanlty more computationally expensive than other all-atom models.
+This seems evident for a person who has ran CHARMM and other all-atom models with the same computer. However, there
+may be some people who have never used any other model than CHARMM. In this case it may be difficult to realize.
+Would anyone have a good bencmark about this? I have assumed that the reason for the slowness are the extra Lennart-Jones
+interactionsi in water, however, I have never looked into it. Would someone have some hard facts on this?}
 

### 2015-01-20 HGmodel_draft3.tex
new version

@@
 %\singlespacing
+
+\onecolumngrid
+{\color{red}
+Matti Javanainen has made the text modifications to the manuscript in excellent way.
+He modified directly the manuscript tex file available in the GitHub and then made a pull request to add a new version of the manuscript into the GitHub repository.
+In addition, he used latexdiff which can be generated to show the pdf with changes.
+
+After Matti had added his version I looked the changes through and agreed. Then I changed his version to the up-to-date version and made some minor modifications.
+
+All the mentioned files can be found from the GitHub repository
+https://github.com/NMRLipids/nmrlipids.blogspot.fi
+
+up-to date manuscript
+HGmodel\_draft3.tex
+
+Version Matti uploaded
+HGmodel\_draft3\_matti.tex
+
+pdf showing the changes made by matti:
+HGmodel\_draft3\_matti\_latexdiff.pdf
+
+I think that this is a very good and convenient way to work collaboratively with the manuscript.
+If you want you can also make a pull request directly using the up-to-date manuscript file name (HGmodel\_draft3.tex).
+Then I will accept the pull request, or we will do some discussion before acception.
+Using latexdiff is very good, but the history will also be saved automatically to the GitHub.
+
+If you find this difficult (as it may be if you are not familiar with GitHub) you can also send your comments and modifications in some other way.
+However, keep in mind that it is very important to clearly report somehow what you have done.
+This is to ease my work to handle the input from large amount of contributors.
+}
 

### 2015-01-23 HGmodelMANUSCRIPT/HGmodel_draft3.tex
add HGmodel manuscript

--- 2015-01-20 HGmodel_draft3.tex
@@
 {\color{red}
-Matti Javanainen has made the text modifications to the manuscript in excellent way.
-He modified directly the manuscript tex file available in the GitHub and then made a pull request to add a new version of the manuscript into the GitHub repository.
+Matti Javanainen has made the text modifications to the manuscript in excellent way. He modified directly the manuscript 
+tex file available in the GitHub and then made a pull request to add a new version of the manuscript into the GitHub repository. 
 In addition, he used latexdiff which can be generated to show the pdf with changes.
 
-After Matti had added his version I looked the changes through and agreed. Then I changed his version to the up-to-date version and made some minor modifications.
+After Matti had added his version I looked the changes through and agreed. Then I changed his version to the up-to-date version 
+and made some minor modifications.
 
@@
 
-I think that this is a very good and convenient way to work collaboratively with the manuscript.
-If you want you can also make a pull request directly using the up-to-date manuscript file name (HGmodel\_draft3.tex).
-Then I will accept the pull request, or we will do some discussion before acception.
-Using latexdiff is very good, but the history will also be saved automatically to the GitHub.
+I think that this is a very good and convenient way to work collaboratively with the manuscript. If you want you can also make 
+a pull request directly using the up-to-date manuscript file name (HGmodel\_draft3.tex). Then I will accept the pull request, 
+or we will do some discussion before acception. Using latexdiff is very good, but the history will also be saved automatically 
+to the GitHub.
 
-If you find this difficult (as it may be if you are not familiar with GitHub) you can also send your comments and modifications in some other way.
-However, keep in mind that it is very important to clearly report somehow what you have done.
+If you find this difficult (as it may be if you are not familiar with GitHub) you can also send your comments and modifications 
+in some other way. However, keep in mind that it is very important to clearly report somehow what you have done. 
 This is to ease my work to handle the input from large amount of contributors.
@@
   \centering
-  \includegraphics[width=8.6cm]{comparison3.pdf}
+  \includegraphics[width=8.6cm]{../DATAreportediINblog/comparisonSorted.pdf}
 %\todo{DONE}\\
@@
 \todo{Markus Miettinen suggested that we should consider making one more figure where only experimental data would be shown and that would be discussed in Section~\ref{experiments}} \\
-\todo{There has been some discussion in the blog that maybe we should include the signs and stereospecific labeling here as well, and that maybe this figure should be clarified.
-My view is that for the main conclusion, i.e.
-CHARMM36, GAFFlipid and MacRog are the best here, this format is good enough.
-However, if someone wants to improve this and/or collect all the data with the signs and/or stereospecific labeling it would be even better.
-There is also very interesting discussion about interactive figure.}
+\todo{Markus has now made a new figure with signs included.
+I think that this is very good.
+I also think that we should try to get the interactive figure behind permalink (i.e.
+Zenodo or figshare) and cite that in the paper.}
   \caption{\label{HGorderparameters}

### 2015-01-23 HGmodelMANUSCRIPT/HGmodel_draft.tex
add HGmodel manuscript

% \documentclass[rmp,aps,floatfix,authordate1-4,preprint]{revtex4}
% \documentclass[prb,aps,floatfix,authordate1-4,preprint]{revtex4}

\documentclass[pre,aps,floatfix,authordate1-4,twocolumn]{revtex4-1}
%\documentclass[pre,aps,floatfix,authordate1-4]{revtex4-1}

%\documentclass[aps,prl,preprint,superscriptaddress]{revtex4}

%\documentclass[aps,prl,preprint,groupedaddress]{revtex4}

\usepackage{rotating} 
\usepackage{times}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[obeyFinal]{easy-todo}
%\usepackage{scrtime}
\usepackage[multiple]{footmisc}
%\usepackage[sort&compress]{natbib}
%\bibpunct{(}{)}{,}{n}{}{}
%\renewcommand{\bibnumfmt}[1]{#1.}

\begin{document}
%\singlespacing

\title{Towards atomistic resolution structure of phosphatidylcholine glycerol backbone and choline headgroup at different ambient conditions}

%\author{Alexandru Botan}
%
%\author{Andrea Catte}
%\author{Fernando Favela}
\author{Patrick Fuchs}
\thanks{The authors are listed in alphaphetical order.}
\thanks{The author list is not completed.}
\thanks{Institut Jacques Monod, CNRS, Université Paris Diderot, Sorbonne Paris Cité, Paris, France}
%\thanks{Université Paris Diderot, Sorbonne Paris Cité, Paris, France}
%\author{Matti Javanainen}
%\thanks{Tampere University of Technology, Tampere, Finland}
%\author{Waldemar Kulig}
\author{Antti Lamberg}
\thanks{Department of Chemical Engineering, Kyoto University, Kyoto, Japan}
%\author{Alexander Lyubartsev}
\author{Markus S. Miettinen}
\thanks{Fachbereich Physik, Freie Universität Berlin, Berlin, Germany}
%\author{Luca Monticelli}
%\thanks{5IBCP, CNRS UMR 5086, Lyon, France}
%\author{Jukka M\"a\"att\"a}
%\thanks{Aalto University, Espoo, Finland}
%\setcounter{page}{1}
%\author{Vasily Oganesyan}
\author{O. H. Samuli Ollila} 
\thanks{{\bf Author to whom correspondence may be addressed. E-mail: samuli.ollila@aalto.fi.}}
\thanks{Helsinki Biophysics and Biomembrane Group, Department of Biomedical Engineering and Computational Science, Aalto University, Espoo, Finland}
%\author{Marius Retegan}
%\thanks{Max Planck Institute for Chemical Energy Conversion, Mülheim an der Ruhr, Germany}
%\author{Hubert Santuz}
%\thanks{INSERM, U1134, DSIMB; Institut National de la Transfusion Sanguine (INTS); Laboratoire d'Excellence GR-Ex, Paris, France}
%\thanks{Université Paris Diderot, Sorbonne Paris Cité, Paris, France}
%\author{Joona Tynkkynen}
%\author{Mark Wilson}
%\author{Alexander Vogel}

\begin{abstract}
Phospholipids are essential building blocks of biological membranes.
Despite of vast amount of accurate experimental data the atomistic resolution structures sampled by the glycerol backbone and choline headgroup
in phoshatidylcholine bilayers are not known. Atomistic resolution molecular dynamics simulation model 
would automatically resolve the structures giving an interpretation of experimental results, if the model
would reproduce the experimental data. In this work we compare the C-H bond vector order
parameters for glycerol backbone and choline headgroup between 12 different atomistic resolution
models and experiments in fully hydrated lipid bilayer. The current models are not accurately enough to resolve the structure.
However, closer inspection of three best performing models (CHARMM36, GAFFlipid and MacRog) suggest
that improvements in the sampled dihedral angle distributions would potentilly lead to the model which
would resolve the structure. Despite of the inaccuracy in the fully hydrated
structures, the response to the dehydration, i.e. P-N vector tilting more parallel to membrane normal, 
is qualitatively correct in all models. The CHARMM36 and MacRog models describe the interactions
between lipids and cholesterol better than Berger/H\"oltje model.
This work has been, and continues to be, progressed and discussed through the blog: nmrlipids.blogspot.fi. 
Everyone is invited to join the discussion and make contributions through the blog. 
The manuscript will be eventually submitted to an appropriate scientific journal. 
Everyone who has contributed to the work through the blog will be offered 
coauthorship. For more details see: nmrlipids.blogspot.fi.

\end{abstract}

\maketitle

~\vspace{0.3cm}\\
{\it \bf}

\section{Introduction}

Phospholipids containing different polar headgroups and different acyl chains are essential building blocks of 
biological membranes. Lamellar phospholipid bilayer structures have been widely studied with various experimental 
and theoretical techniques as a simple model for the biological membranes~\cite{lipowsky95,tieleman97,klauda08,edholm08,tieleman10,piggot12,rabinovich13,marsh13}. 
Phospholipid molecules are composed of hydrophobic acyl chains and hydrophilic headgroup, which are connected by glycerol backbone,
see Fig.~\ref{POPCstructure} for the structure of 1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
The behaviour of the acyl chains in a bilayer is relatively well understood~\cite{Israelachvili80,lipowsky95,tieleman97,klauda08,edholm08,tieleman10,marsh13}. 
The structures sampled by the glycerol backbone and choline, however, are not fully 
resolved since even the most accurate scattering and Nuclear Magnetic Resonance (NMR)
techniques give only a set of values that the structure has to fulfill, but
there is no unique way to derive the actual structure from these parameters~\cite{seelig77b,skarjune79,Israelachvili80,jacobs80,davis83,akutsu91,hong95b,semchyschyn04}.
Several approaches have been used to resolve structural details, but general consensus has not been reached~\cite{seelig77b,skarjune79,Israelachvili80,jacobs80,davis83,akutsu91,hong95b,semchyschyn04}. 
On the other hand, the glycerol backbone structures are similar for various biologically
relevant lipid species (phosphatidylcholine (PC), phosphatidylethanolamine (PE) and phosphatidylglycerol (PG)) 
in various environments~\cite{gally81} and the headgroup choline structures are similar in model membranes and
real cells (mouse fibroblast L-M cell)~\cite{scherer87}.
Thus, the resolvation of phosphatidylcholine glycerol backbone and choline structures would be 
useful for understanding wide range of different biological membranes.

Classical atomistic resolution molecular dynamics simulations have been widely used to study  
lipid bilayers~\cite{tieleman97,klauda08,edholm08,tieleman10,piggot12,rabinovich13}. As these models provide an atomistic
resolution description of the whole lipid molecule, they have potential to resolve the glycerol backbone and 
headgroup structures. In particular, the experimental C-H bond order parameters for the glycerol backbone 
(g$_1$, g$_2$ and g$_3$) and choline ($\alpha$ and $\beta$) segments (see Fig.~\ref{POPCstructure} for definitions) are among the main parameters used in
attempts to derive the structures from experimental data~\cite{seelig77b,skarjune79,jacobs80,davis83,akutsu91,hong95b,semchyschyn04}.
These parameters are also routinely compared between experiments and simulations for the acyl chains~\cite{tieleman97,klauda08,edholm08,tieleman10,piggot12}.
Thus, the structures sampled in a simulation model that reproduces these and other experimental parameters, automatically
give an interpretation of the experiments, in other words they can be considered as reasonable atomistic resolution descriptions of
the behaviour of lipid molecules in a bilayer.

The glycerol backbone and choline headgroup order parameters have been compared between simulations and experiments
in some studies~\cite{hogberg08,castro08,klauda10,kapla12,dickson12,poger12,ferreira13,chowdhary13,maciejewski14}, 
however much less frequently than for acyl tail chains~\cite{tieleman97,klauda08,edholm08,tieleman10,piggot12}.
The main reason is probably that the existing experimental data for the glycerol backbone
and choline headgroups is scattered over many publications and published in a format that is difficult to understand without some NMR expertise.

In this work we first review the most relevant experimental data for the glycerol backbone and choline headgroup order parameters
in a phosphatidylcholine lipid bilayer. Then the available atomistic resolution lipid models are carefully compared to the 
experimental data. The comparison reveals that the CHARMM36~\cite{klauda10}, GAFFlipid~\cite{dickson12} and MacRog models~\cite{maciejewski14}
have the most realistic glycerol backbone and choline structures. We also compare the glycerol backbone and choline 
structures between the most often used (Berger) lipid model~\cite{berger97} and the best performing models, to demonstrate that by using the 
order parameters we can distinguish the more reasonable structures from the less reasonable ones. However, none of the current models 
is accurate enough to resolve the atomistic resolution stuctures.

In addition to the fully--hydrated single--component lipid bilayers, the glycerol backbone and choline order parameters
have been measured under a large number of different conditions. For example, as a function of hydration level~\cite{bechinger91,ulrich94,dvinskikh05b}, cholesterol content~\cite{brown78,ferreira13}
ion concentration~\cite{brown77,akutsu81,altenbach84,roux90,roux91}, temperature~\cite{gally75}, charged lipid content~\cite{roux90,roux91}, charged surfactant content~\cite{scherer89}, 
drug molecule concentration~\cite{kelusky84,castro08}, and protein content~\cite{roux89,kuchinka89} (listing only the publications most relevant for this work and the pioneering studies).
Awareness of the existence of this type of data allows the comparison of structural responses to varying conditions between simulations and experiments,
which can be used to validate the simulation models and to interpret the original experiments. 
In this publication we demonstrate the power of this approach for understanding the behaviour of a bilayer as a function of hydration level and cholesterol concentration.
Choline headgroup order parameters as function of ion concentration, and their relation to the ion binding affinity, are discussed elsewhere~\cite{ionpaper}.

  \begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{POPCstructure.eps}

  \caption{\label{POPCstructure}
    Chemical structure of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).}

\end{figure}

\section{methods}

\subsection{Open collaboration}
\todo{How much should we write about this?}

This work has been done, and is continuing, as an open collaboration through nmrlipids.blogspot.fi.

\subsection{Analysis}
The order parameter of a hydrocarbon C-H vector is defined as $S_{{\rm CH}}=\frac{3}{2}\langle \cos^2 \theta-1 \rangle$, where
the average is an ensemble average over the sampled conformations, and $\theta$ is the angle between the C-H bond and the membrane normal.
The order parameters can be measured by detecting quadrupolar splitting with $^2$H NMR~\cite{seelig77c} or by detecting dipolar 
splitting with $^{13}$C NMR~\cite{hong95a,gross97,dvinskikh05a,ferreira13}. The measurements are based on
different physical interactions and also the connection between order parameters and quadrupolar or dipolar splitting
are different. The order parameters from the measured quadrupolar splitting $\Delta \nu_Q$ ($^2$H NMR) are calculated using 
the equation $|S_{{\rm CD}}|=\frac{4}{3} \frac{e^2qQ}{h} \Delta \nu_Q$, where the value for the static quadrupole
splitting constant is estimated from various experiments to be 170~kHz leading to a numerical relation $|S_{{\rm CD}}|=0.00784 \times \Delta \nu_Q$~\cite{seelig77c}. 
The order parameters from the measured dipolar splitting $d_\mathrm{CH}$ ($^{13}$C NMR) are calculated using equation
$|S_{{\rm CH}}|=\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c} d_\mathrm{CH}$, where
values between 20.2-22.7 kHz are used for $\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c}$,
depending on the original authors~\cite{hong95a,gross97,dvinskikh05a,ferreira13}.
It is important to note that the order parameters measured with different techniques based on different physical interactions are in good agreement
with each other (see Results and Discussion), indicating very high quantitative accuracy of the measurements.
For more detailed discussion see the Supplementary material \todo{The idea is to publish all the blog content as supplementary material or through
Figshare or Zenodo}.

The order parameters from simulations were calculated directly using the definition.
%$S_{{\rm CH}}=\frac{3}{2} \langle \cos^2 \theta-1 \rangle$. 
For the united atom models the hydrogen locations were generated 
post-simulationally using the positions of the heavy atoms in the simulation trajectories.
The statistical error estimates were calculated for the best performing simulation
models by calculating the error of the mean for the average over individual lipids in the system.

\subsection{Simulated systems}
All simulations are ran with a standard setup for planar lipid bilayer in zero tension
with periodic boundary conditions with Gromacs software package (version numbers ?-? \todo{Data contributors: please deliver the version number you have used}).
The number of molecules, temperatures and the length of simulations for all the fully 
hydrated lipid bilayer simulations are listed in Table~\ref{systems}. Full simulation
details for all simulations are given the next section.

\begin{table*}[htb]
\centering
\caption{Simulated pure lipid bilayers with full hydration.
}\label{systems}
\begin{tabular}{c c c c c c c}
%\hline
Force field & lipid  & \footnote{The number of lipid molecules}N$_{\rm l}$   &  \footnote{The number of water molecules}N$_{\rm w}$ & \footnote{Simulation temperature}T (K)  & \footnote{The total simulation time}t$_{{\rm sim}}$(ns) \todoi{Do we need this or would it be enough to write into the text that "All the simulations were sufficinetly equlibrated before the analysis"?} & \footnote{Time frames used in the analysis}t$_{{\rm anal}}$ (ns)\\
\hline
Berger\cite{berger97}          &   POPC & 128 & 7290  & 298  & 270 & 240  \\
Berger\cite{berger97}          &   POPC & 128 & 7290  & 323  & 56 & 56  \\
Berger\cite{berger97}\todoi{Jukka Maatta, please check the information and deliver missing numbers}          &   DPPC & 72 & ?  & 323  & ? & 100  \\
Berger\cite{berger97}          &   DMPC & 128 & 5097  & 323  & ? & 100  \\
CHARMM36\cite{klauda10}        & POPC   & 72  &  2242 & 303 & 30 & 20  \\
CHARMM36\cite{klauda10}\todoi{Hubert Santuz, please check the information and deliver missing numbers}        & POPC   & 128 &  ?    & 303 & 150 & 50  \\
CHARMM36\cite{klauda10}        & DPPC   & 72  &  2189 & 323 & 30 & 25  \\
CHARMM36\cite{klauda10}\todoi{Hubert Santuz, please check the information and deliver missing numbers}        & DOPC   & 128 &  ?    & 303 & 150 & 50  \\
MacRog\cite{maciejewski14}\todoi{Matti Javanainen, please check the information and deliver missing numbers}  & POPC & 288  & ? & 310 & 100 & 80  \\
GAFFlipid\cite{dickson12}       & POPC & 126  & 3948  & 303 & 137 & 32  \\
Lipid14\cite{dickson14}         & POPC  & 72 & 2234 & 303 & 100 & 50  \\
Poger\cite{poger10}\todoi{Patrick Fuchs, please check the information and deliver missing numbers}             & POPC  &  &  & & &  \\
Slipid\cite{jambeck12}\todoi{Jukka Maatta, please check the information and deliver missing numbers}          & DPPC & ? & ? & ? & ? & 30?  \\
Kukol\cite{kukol09} \todoi{Matti Javanainen, please check the information and deliver missing numbers}         & ?   & ? & ? & ? & ? & ?  \\
Chiu et al.\cite{chiu09}     & POPC  & 128 & 3552  & 298 & 56 & 50 \\
H\"ogberg et al.\cite{hogberg08}\todoi{Alexander Lyubartsev, please check the information and deliver missing numbers}  & POPC   &  ? & ?  & 303 & ? & ?  \\
H\"ogberg et al.\cite{hogberg08}\todoi{Alexander Lyubartsev, please check the information and deliver missing numbers}  & DMPC   &  ? & ?  & 303 & ? & ?  \\
Ulmschneider\cite{Ulmschneider09}\todoi{Matti Javanainen, please check the information and deliver missing numbers}    & POPC  & 128 & ? & 310 & 100 & 50 \\
Tj\"ornhammar et al.\cite{tjornhammar14}\todoi{Matti Javanainen, please check the information and deliver missing numbers}    & DPPC  & 144 & ? & ? & 200 & 100 \\
OPLS-AA\cite{rog09b}\todoi{We have this only with 150mM of NaCl delivered by Joona Tynkkynen. Options are to remove these results from this publicatio, run the simulations without ions (if not yet available), 
  or include the results with ions. In my understanding this is a proto version of MacRog so it could be left out as well. However, for historical reasons and 
to understand literature it might be useful to include also these results.}         &   &  &  & & &  \\
CHARMM36-UA~\cite{henin08,lee14}\todo{Alexandru Botan}     & DLPC   & 128  & ?  & 323 & ? & 20? \\
\end{tabular}
\end{table*}

The initial structures for the simulations in low hydrated conditions are made by removing the
water molecules from the fully hydrated system to achieve the targeted hydration conditions.

The results for the POPC/cholesterol systems with the Berger lipid~\cite{berger97} and H\"oltje
cholesterol model~\cite{holtje01} were taken directly from Ref.~\cite{ferreira13}.
For the initial structure of CHARMM36-~\cite{klauda10,lim12} and MacRog-~\cite{maciejewski14,kulig14} simulations with cholesterol,
the required amount of lipid molecules were replaced with cholesterol \todo{Contributors: Is this correct?}.

\subsection{Simulation details} 
\todo{Markus Miettinen suggested that maybe these should be put as a table?} \\
\todo{I think that the mdp files should be preferably shared in addition of writing the details here.}
\subsubsection{Berger}

The Berger force field was used for the POPC~\cite{berger97}, with the dihedral potential next to the double bond 
taken from~\cite{bachar04}. The simulations are identical to previous publications~\cite{ollila07a,ferreira13,ferreira14b}.
Timestep of 2~fs was used with leap-frog integrator. Covalent bond lengths were constrained with LINCS algorithm~\cite{hess97,hess07}. 
Coordinates were written every 10~ps. PME with real space cut-off at 1.0~nm was used 
for electrostatics. Plain cut-off was used for the Lennard-Jones interactions with a 1.0~nm cut-off.
The neighbour lists were updated every 5th step with cut-off at 1.0~nm. Temperature was coupled separately
for lipids and water to 298~K with the velocity-rescale method~\cite{bussi07} with coupling constant 0.1~ps$^{-1}$.
Pressure was semi-isotropically coupled to the atmospheric pressure with the Berendsen method~\cite{berendsen84}.

\subsubsection{CHARMM36}
\todo{Markus Miettinen made the files.}

Timestep of 1~fs was used with leap-frog integrator. Covalent bonds with hydrogens were constrained with LINCS algorithm~\cite{hess97,hess07}. 
Coordinates were written every 5~ps. PME with real space cut-off at 1.4~nm was used 
for electrostatics. Lennard-Jones interactions were switched to zero between 0.8~nm and 1.2~nm.
The neighbour lists were updated every 5th step with cut-off 1.4~nm. Temperature was coupled separately
for lipids and water to 303~K with the velocity-rescale method~\cite{bussi07} with coupling constant 0.2~ps$^{-1}$.
Pressure was semi-isotropically coupled to the atmospheric pressure with the Berendsen method~\cite{berendsen84}.

\todo{Hubert Santuz made the simulations with cholesterol.}

\subsubsection{MacRog}
\todo{Matti Javanainen}
\subsubsection{GAFFLipid}
\todo{Marius Retegan made the files}
The initial structure in Lipidbook had different glycerol backbone isomers in different leaflets. 
To generate the initial structure we took the structure delivered by Slipid developers. Also this strucutre
had on lipid with different glycerol bakcbone isomer. This lipid and one lipid from opposite leaflet were removed
after the system was equilibrated.

Timestep of 2~fs was used in Langevin dynamics with zero friction term and collision frequency of 1.0~ps$^{-1}$. 
Covalent bonds with hydrogens were constrained with LINCS algorithm~\cite{hess97,hess07}.
Coordinates were written every 10~ps. PME with real space cut-off at 1.0~nm was used 
for electrostatics. Plain cut-off with 1~nm was used for Lennardt-Jones interactions. 
The neighbour lists were updated every 5th step with cut-off 1.0~nm. 
Pressure was semi-isotropically coupled to the 1~bar pressure with the Berendsen method~\cite{berendsen84}.
\subsubsection{Lipid14}
\todo{Marius Retegan made the files}

Timestep of 2~fs was used in Langevin dynamics with zero friction term and collision frequency of 1.0~ps$^{-1}$. 
Covalent bonds with hydrogens were constrained with LINCS algorithm~\cite{hess97,hess07}.
Coordinates were written every 10~ps. PME with real space cut-off at 1.0~nm was used 
for electrostatics. Plain cut-off with 1~nm was used for Lennardt-Jones interactions. Dispersion correction
was used for energy and pressure. The neighbour lists were updated every 5th step with cut-off 1.0~nm. 
Pressure was semi-isotropically coupled to the 1~bar pressure with the Berendsen method~\cite{berendsen84}.

\subsubsection{Poger et al.}
\todo{Patrick Fuchs, maybe we should shortly mention the Gromacs version issue here?}
\subsubsection{Slipid}
\todo{Matti Javanainen}
\subsubsection{Kukol}
\todo{Matti Javanainen}
\subsubsection{Chiu et al.}
The force field parameters and the initial configuration were downloaded from: \todo{The page I used does not exist anymore}.
Timestep of 2~fs was used with leap-frog integrator. Covalent bond lengths were constrained with LINCS algorithm~\cite{hess97,hess07}. 
Coordinates were written every 10~ps. PME with real space cut-off at 1.0~nm was used 
for electrostatics. Twin range cut-off was used for the Lennardt-Jones interactions with short and long cut-offs at 1.0~nm and 1.6~nm, respectively.
The neighbour lists were updated every 5th step with cut-off at 1.0~nm. Temperature was coupled separately
for lipids and water to 298~K with the velocity-rescale method~\cite{bussi07} with coupling constant 0.2~ps$^{-1}$.
Pressure was semi-isotropically coupled to the atmospheric pressure with the Parrinello-Rahman method~\cite{parrinello81}.
\subsubsection{H\"ogberg et al.}
\todo{Alexander Luybartsev}
\subsubsection{Ulmschneider}
\todo{Matti Javanainen}
\subsubsection{OPLS-AA}
\todo{Joona Tynkkynen. We have this only with 150mM of NaCl delivered by Joona Tynkkynen. Options are to remove these results from this publicatio, run the simulations without ions (if not yet available), 
  or include the results with ions. In my understanding this is a proto version of MacRog so it could be left out as well. However, for historical reasons and 
to understand literature it might be useful to include also these results.}

\section{Results and Discussion}

\subsection{Full hydration: Experimental order parameters for glycerol backbone and headgroup}\label{experiments}
The specific deuteration of $\alpha$-, $\beta$- and g$_3$- segments of dipalmitoylphosphatidylcholine (DPPC) has been successful, 
allowing the order parameter measurements for these segments by $^2$H--NMR~\cite{gally75,brown77,brown78,akutsu81}.
In addition, the order parameters for all glycerol backbone and choline headgroup segments in egg yolk lecithin~\cite{hong95a},
1,2-dimyristoyl-sn-glycero-3-phosphocholine (DMPC)~\cite{hong95b,gross97,dvinskikh05a}, 
1,2-dioleoyl-sn-glycero-3-phosphocholine (DOPC)~\cite{warschawski05} and POPC~\cite{warschawski05,ferreira13}
have been measured with several different implementations of $^{13}$C NMR experiments.
The experimental absolute values of glycerol backbone and choline order parameters from various publications are shown in Fig.~\ref{HGorderparameters}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{comparison12.eps}
\todo{Figure should be updated. At least, the recent results from the model by Tjornhammar and Edholm are missing. Also CHARMM-UA results.} \\
\todo{Markus Miettinen suggested that we should consider making one more figure where only experimental data would be shown and that would be discussed in Section~\ref{experiments}}
  \caption{\label{HGorderparameters}
  Order parameteres from simulations and experiments for glycerol and choline groups.
  The experimental values were taken from the following publications: DMPC 303K from \cite{gross97}, DMPC 314K from \cite{dvinskikh05a}, DPPC 310K from \cite{gally75}, 
  DPPC 323K from \cite{akutsu81}, and POPC 298K \cite{ferreira13}.
  The vertical bars shown for some of the computational values are not error bars, but demonstrate that for 
  these systems we had two data sets; the ends of the bars mark the values from the two sets, and the dot marks their measurement-time-weighted average. 
} 
\end{figure}

In general there is a good agreement between the order parameters measured with different experimental NMR techniques: Almost all the 
reported values are inside variation of $\pm$0.02 (which is also the error estimate given by Gross et al.~\cite{gross97}) 
for all fully hydrated PC bilayer, regardless of the variation in their acyl chain composition and the temperature.
Exception are the somewhat lower order parameters sometimes reported having been measured with $^{13}$C--NMR~\cite{hong95a,hong95b,warschawski05}.
These experiments have not seen in Fig.~\ref{HGorderparameters} as the reported error bars are either relatively large~\cite{hong95a,hong95b}, 
or the spectral resolution is quite low and the numerical lineshape simulations have not been used in the analysis~\cite{warschawski05}.
Therefore it is highly likely that these reported lower order parameters are due to lower experimental 
accuracy and that we exclude these values from our discussion. 
Motivated by the high experimental repeatability, we have highlighted in 
Fig.~\ref{HGorderparameters} subjective sweet spots (light blue areas), within which we expect the calculated absolute 
values of order parameters of a well-performing force field to fall.

In addition to the phosphatidylcholine lipids, similar values of the glycerol backbone order parameters have been measured
for phophatidylethanolamine (PE) and phopshatidylglycerol (PG) in E. Coli extract~\cite{gally81},
indicating that the glycerol backbone structure is similar independent of the headgroup chemistry and lipid environment.
Further, choline order parameters measured from mouse fibroblast L-M cell are similar to the ones in model
membranes~\cite{scherer87}.

In addition to the numerical values, an important feature of the glycerol backbone is the 
inequality of order parameters for the two hydrogens attached to the same carbon in g$_1$ and g$_3$ segments,
while the hydrogens in choline $\alpha$ and $\beta$ segments give equal values.
Note that in this work we call the phenomena of inequal order parameters for hydrogens attached to the same carbon as ''forking'' to avoid 
confusion with dipolar and quadrupolar splitting in NMR terminology. Forking is also observed experimentally for the C$_2$ carbon in the sn-2
chain of all phosholipids, and it is known to arise from differently sampled orientations of the two C-H bonds, not from two 
different populations of lipid conformations~\cite{engel81}. The forking in glycerol backbone g$_3$ segment is small ($\approx$ 0.02) 
and some experiments only report the larger or the average value~\cite{akutsu81,ferreira13}. 
In contrast, forking is significant for the glycerol backbone g$_1$ segment, whose lower order parameter is close to zero and the
larger one has absolute values around 0.13-0.15. Forking was studied in detail by Gally et al.~\cite{gally81}, who used E. Coli to 
stereospecifically deuterate the different hydrogens attached to the g$_1$ or g$_3$ groups in PE lipids, and measured the order parameters from the lipid 
extract. This experiment gave the lower order parameter when deuterium was in the S position of g$_1$ or R position for g$_3$.
Since the glycerol backbone order parameters are very similar irrespective of the headgroup chemistry (PC,PE and PG) or lipid 
environment~\cite{gally81}, it is reasonable to assume that the stereospecifity measured for the PE lipids
holds also for the PC lipids.

In Fig.~\ref{HGorderparameters} we have shown the absolute values of order parameters as these are accessible
with both $^2$H NMR and $^{13}$C NMR techniques. However, $^{13}$C NMR techniques allow also the measurement of 
the sign of the order parameter~\cite{hong95a,hong95b,gross97}. The measured sign is negative for almost all the carbons 
discussed in this work, only $\alpha$ is positive~\cite{hong95a,hong95b,gross97}.

Combining the experimental information of the sign~\cite{hong95a,hong95b,gross97} and the stereospecifity 
measurements~\cite{gally81} with the absolute value measurements from various techniques~\cite{gally75,akutsu81,gross97,dvinskikh05a,ferreira13}
having high quantitative accuracy,
the most detailed experimentally available order parameter information for the glycerol backbone and choline segments of POPC is obtained.
These data are shown in Fig.~\ref{HGorderparameters2}.
\begin{figure*}[]
  \centering
  \includegraphics[width=8.6cm]{HGorderparameters5.eps} \\
  \todo{Markus Miettinen suggested that we should consider ``spreading'' the values similarly to the figure~\ref{HGorderparameters}} \\
  \todo{I think that we should include the error bars from simulations in this figure. I also think that for CHARMM36 in this figure we should use the 
    results delivered by Hubert Santuz \\ 
    (https://github.com/NMRLipids/nmrlipids.blogspot.fi/blob/master/DATAreportediINblog/POPC/CHARMM36-303K\_blogged-15-01-14.dat) \\
    which are from longer simulations than mine \\
    (https://github.com/NMRLipids/nmrlipids.blogspot.fi/blob/master/DATAreportediINblog/POPC/CHARMM36-303K\_blogged-20-11-13.dat). \\
  For this we would need error bars from simulations by Hubert.}
  \caption{\label{HGorderparameters2}
  Order parameteres from simulations with Berger, CHARMM36, GAFFlipid and Macrog force fields together with experimental values for POPC glycerol and choline groups.
  The magnitudes for experimental order parameters are taken from Ferreira et al.~\cite{ferreira13}, the signs are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
  and Gross et al.~\cite{gross97}, and the R/S labeling is based in the measurements by Gally et al.~\cite{gally81}.
} 
\end{figure*}

\subsection{Full hydration: Comparison between simulation models and experiments}

The order parameters of the glycerol backbone and headgroup calculated from different force fields for various lipids has been 
previously compared to experiments~\cite{hogberg08,castro08,klauda10,kapla12,dickson12,poger12,ferreira13,chowdhary13,maciejewski14}. 
The general conclusion from these works seems to be that the CHARMM based~\cite{hogberg08,klauda10}, GAFFlipid~\cite{dickson12} and
MacRog~\cite{maciejewski14} force fields perform better for the glycerol backbone and headgroup structures than the Gromos based models~\cite{castro08,kapla12,poger12,ferreira13}.
However, none of the studies exploits the full potential of the available experimental data discussed in previous section; the quantitative accuracy, known signs and stereospecific labeling of
the experimental order parameters.

To get a general idea about the quality of the glycerol backbone and choline headgroup structures in different models, we calculated the absolute
values of the order parameters for these parts from twelve \todo{This number may change} different lipid models (Table~\ref{systems}) and 
plotted the results together with experimental values in Fig.~\ref{HGorderparameters}.
Two criteria were used to judge the quality of the model: {\bf forkings} there must not be significant forking in the $\alpha$ and $\beta$ carbons,
there must be only moderate forking in the g$_3$ carbon and there must be significant forking in the g$_1$ carbon, {\bf absolute values}
should be preferably inside to the subjective sweet spots determined from experiments (blue shaded regions in Fig.~\ref{HGorderparameters}).
None of the studied force fields fulfills these criteria, however, three force fields are closer than others: CHARMM36~\cite{klauda10}, MacRog~\cite{maciejewski14} and GAFFlipid~\cite{dickson12}.
The results for each force field in respect to the above criteria are summarized in Table~\ref{??} \todo{This kind of table should be done or not?}.

The top three models (CHARMM36, MacRog and GAFFlipid) together with the most used lipid model (Berger model) 
were subjected to a more careful comparison including the signs and the sterospecific labeling in Fig.~\ref{HGorderparameters2}.
The essential additional information given by this comparison is that the sign of the $\beta$ carbon order parameter is correct only in CHARMM36 model.

\subsection{Full hydration: Atomistic resolution structures in different models}

The results in the previous section revealed significant differences of the glycerol backbone and choline headgroup
order parameters between different molecular dynamics simulation models.
However, it is not straightforward to conclude which kind of structural differences (if any)
between the models the results indicate, because the mapping from the order parameters to the 
structure is not unique. In this section we demonstrate that 1) the differences in order parameters
indicate significantly different structural sampling strongly correlating with the dihedral angles of the related bonds,
and that 2) the comparison between experimental and simulated order parameters can be used to exclude
nonrealistic structural samping in molecular dynamics simulations. The demonstration is done for 
the dihedral angles defined by the g$_3$-g$_2$-g$_1$-O(sn-1) segments in the glycerol backbone and 
the N-$\beta$-$\alpha$-O segments in the headgroup. These dihedrals were chosen for demonstration, because 
significant differences between the models are observed around these segments in Fig.~\ref{HGorderparameters2}.
We note that performing a similar comparison through all the dihedrals in all the 12 \todo{this number may change} models would probably give highly useful
information to improve the accuracy of simulations, however this is beyond the scope of the current report.

The dihedral angle distributions for the  g$_3$-g$_2$-g$_1$-O(sn-1) dihedral calculated from different models are
shown in Fig.~\ref{dihDISTS}. The distribution is qualitatively different for the Berger model, showing a maximum in 
the gaughe$^+$-conformation (60$^o$) compared to all the other models showing a maximum in the trans-conformation (180$^o$).
The distributions in all the other models have the same general features, the main difference being that the
fraction of configurations in gaughe$^-$-conformation (-60$^o$) is zero for the MacRog, detectable for the CHARMM36 and
equally large to the gaughe$^+$ fraction in GAFFlipid. From the results we conclude that most likely the wrongly sampled
dihedral angle for the g$_2$-g$_1$ bond explains the significant discrepancy to the experimental order parameters
for the g$_1$ segment in the Berger model (Fig.~\ref{HGorderparameters2}).
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{g1-g2_Cdihs.eps}
  \caption{\label{dihDISTS}
    Dihedral angle distributions for g$_3$-g$_2$-g$_1$-O(sn-1) dihedral from different models (POPC bilayer in full hydration).
  } 
\end{figure}

The dihedral angle distribution for the  N-$\beta$-$\alpha$-O dihedral calculated from the same four models is 
shown in Fig.~\ref{dihDISTS2}. Also for this dihedral there are significant differences in the gauche-trans fractions.
The gaughe conformations are dominant in the CHARMM36 while in MacRog there are only trans conformations present.
Interestingly, the probability of the gaughe conformations correlates with the order parameter difference between the $\beta$ and $\alpha$ segments:
the larger the gaughe fraction the larger the order parameter difference. This suggestion together with the results in 
Fig.~\ref{HGorderparameters2} would indicate that the correct gaughe-trans fraction for N-$\beta$-$\alpha$-O dihedral is 
larger than in GAFFlipid but smaller than in CHARMM36. This happens to be quite close to the gaughe-trans fraction in
the Berger model. However, there is significant forking and numerical values are off from the experiments in the Berger
model suggesting that it has some other inaccuracies.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{a-bDIHS.eps}
  \caption{\label{dihDISTS2}
    Dihedral angle distributions for N-$\beta$-$\alpha$-O dihedral from different models (POPC bilayer in full hydration).
  } 
\end{figure}

The used examples show that the glycerol backbone order parameters reflect the atomistic resolution structure
and that the comparison with experiments allows the assesment of the quality of the suggested structure. We were able to pinpoint
specific problems in the structures in different models and suggest potential improvement strategies.
If the improved atomistic resolution
molecular dynamics simulation model would reproduce the order parameters and other experimental observables (like chemical shift anisotropy)
with experimental accuracy, it would give an interpretation for the atomistic resolution structure of the glycerol backbone and 
choline~\cite{seelig77b,skarjune79,jacobs80,davis83,akutsu91,hong95b,semchyschyn04}. The research along these lines is left, however,
for future studies.

\subsection{Response to dehydration and cholesterol content}
In addition to pure phosphatidylcholine bilayers at full hydration, the choline headgroup order parameters
have been measured under various different conditions~\cite{gally75,brown77,brown78,akutsu81,altenbach84,scherer89,bechinger91,ulrich94,dvinskikh05b,castro08,kapla12,ferreira13}.
Also the order parameters for the glycerol backbone have been measured with $^{13}$C NMR in dehydrated conditions~\cite{dvinskikh05b}, and as a function 
of anesthetics~\cite{castro08} and glycolipids~\cite{kapla12} for DMPC, and as a function of cholesterol 
concentration for POPC~\cite{ferreira13}. Due to the high resolution in the NMR (especially $^2$H NMR) experiments,
even very small order parameter changes resulting from the varying conditions can be measured (see supplementary
material for more detailed discussion \todo{The idea is to publish all the blog content as supplementary material or through
Figshare or Zenodo}.). However, as already discussed above, it is not simple to deduce 
the structural changes from order parameter changes~\cite{akutsu91,semchyschyn04}. Consequently, comparison of the order parameters
between simulations and experiments in different conditions can be used measure the quality of the force field 
in different situations, and, if the quality is good, to potentially interpret the structural changes in experiments.
Here we exemplify such comparison for a lipid bilayer under low hydration levels and mixed with cholesterol. 
The interaction between ions and phosphatidylcholine bilayer is discussed in a separate work~\cite{ionpaper}.

\subsubsection{Phospholipid bilayer with low hydration level}
The experimental order parameters available in the literature~\cite{dvinskikh05b,ulrich94,bechinger91} 
for the glycerol backbone and choline as a function of hydration level are shown in Fig.~\ref{ordPhydr}. 
The independently reported values for choline segments are in good agreement with each other (despite of 
slight differences in temperature and acyl chain composition),
showing increase for both segments with decreasing hydration level. It should be noted that only 
absolute values were measured in the original experiments~\cite{dvinskikh05b,ulrich94,bechinger91}, but
we have included the signs measured separately~\cite{hong95a,hong95b,gross97}. 
Consequently, the $\beta$ order parameter with negative sign actually increases with dehydration 
since the absolute value decreases~\cite{dvinskikh05b,ulrich94,bechinger91}.
Slight decrease for the glycerol backbone g$_3$- and g$_2$- order parameters were observed with dehydration, 
while g$_1$ remained practically unchanged~\cite{dvinskikh05b}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{OrderParameterDEHYDpresentation.eps}
\todo{Markus Miettinen suggested that this plot should be clarified.}
  \caption{\label{ordPhydr}
    The effect of dehydration on glycerol and choline order parameters in experiments.
    The magnitudes of order parameters are measured for DMPC ($^{13}$C NMR) at 314~K~\cite{dvinskikh05b}, 
    for POPC ($^2$H NMR) at 296~K~\cite{bechinger91} and for DOPC ($^2$H NMR) at 303~K~\cite{ulrich94}. 
    The signs are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
    and Gross et al.~\cite{gross97}.
  }
\end{figure}

Lipid bilayer dehydration has been studied also with molecular dynamics simulations~\cite{mashl01,pertsin05,pertsin07,eun09,eun10,schneck12},
typically motivated by the  discussion about the origin of the ``hydration repulsion''~\cite{israelachvili,israelachvili96,sparr11}.
However, the used simulation models are not typically compared to the experimental choline and glycerol backbone
order parameters (except by Mashl et al.~\cite{mashl01}).
In Fig.~\ref{ordPhydr} the glycerol backbone and choline order parameters as a function of hydration level are shown 
for the CHARMM36, MacRog and GAFFlipid models (having the most realistic atomistic resolution structures) together with the Berger model 
(which is the most used lipid model). The choline order parameters increase with dehydration in all simulation
models, in qualitative agreement with experiments. 
The measured decrease in both g$_3$ and g$_2$ order parameters with dehydration is reproduced only in CHARMM36.

The qualitative agreement with experiments in all simulation models for the $\alpha$ and $\beta$ order parameters  
as a function of hydration indicates that the structural response of the choline headgroup to dehydration is somewhat realistic
despite the unrealistic structures at full hydration. 
The most likely explanation is that the choline group
orients more parallel to the membrane plane with dehydration due to restricted interlamellar space. 
Indeed, the P-N angle vector angle with membrane normal as a function of dehydration shows an increase for
all models as a function of dehydration in Fig.~\ref{PNangle}.
However, the qualitative agreement in the lipid response to dehydration does not guarantee the correct 
free energy landscape if the simulation model has incorrect structure. The influence of this issue to 
dehydration energetics studied with simulations~\cite{eun09,schneck12} is left for future studies.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{PNangles.eps}

  \caption{\label{PNangle}
    The angle between membrane normal and P-N vector in choline segment as function of
    hydration level calculated from different simulations.
  }
\end{figure}

The response of the glycerol backbone to dehydration seems to be more subtle than of the choline headgroup 
as CHARMM36 is the only model that reproduces the decrease in g$_2$ and g$_3$ segments.

\subsubsection{Phospholipid bilayer mixed with cholesterol}
Phospholipid--cholesterol interactions have been widely studied with theoretical~\cite{huang99,zhu07,rog09,alwarawrah12} and
experimental methods~\cite{brown78,marsh10,ferreira13,marsh13}, since cholesterol is abundant in biological membranes and
it has been suggested to be an important player, for example, in domain formation~\cite{simons04,somerharju09}.
It is widely agreed that cholesterol orders lipid acyl tails thus decreasing the area per molecule (condensing effect),
however, the influence of cholesterol on the lipid headgroup and glycerol backbone are sill debated~\cite{huang99,simons04,somerharju09}.
For example, it has been suggested that the surrounding lipids shield cholesterol from interactions with water by 
tilting their headgroups (``umbrella model'')~\cite{huang99} or that cholesterol acts as a spacer for the headgroups thus increasing 
their entropy and dynamics (``superlattice model'')~\cite{somerharju09}. Both of these suggestions have been supported
by molecular dynamics simulations~\cite{zhu07,alwarawrah12}, and other simulations suggest specific
interactions between the glycerol backbone and cholesterol~\cite{rog09}, however the glycerol backbone and choline headgroup behaviour
as a function of cholesterol content is not compared to experiments in these studies.

The choline headgroup and glycerol backbone order parameters for POPC measured by $^{13}$C NMR~\cite{ferreira13} and DPPC choline order parameters 
measured by $^{2}$H NMR~\cite{brown78} are shown in Fig.~\ref{ordPchol} as a function of cholesterol content.
The agreement between different experimental results is again very good, showing only very modest changes in 
the choline order parameters as a function of cholesterol content. It should be noted, however, that very small
changes are measurable with high resolution $^{2}$H NMR experiments
and cholesterol causes a measurable increase in the $\beta$ order parameter and a forking in the $\alpha$ order
parameter~\cite{brown78}, but these effects are so small that they are barely visible in the scale used in Fig.~\ref{ordPchol}.
Further, the effects of cholesterol on the glycerol backbone order parameters for POPC from $^{13}$C NMR experiment~\cite{ferreira13} 
is in good agreement with the results for the phosphatidylethanolamine (PE) measured by $^{2}$H NMR~\cite{ghosh82}.
These results further support the idea that the glycerol backbone structural behaviour is independent of the
headgroup composition~\cite{gally81} and that the headgroup stucture is independent of the acyl chain region content unless
charges are present~\cite{scherer87}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{OrderParameterCHOLexpVSsim.eps}
  \caption{\label{ordPchol}
    The effect of cholesterol content on the glycerol backbone and choline order parameters in experiments~\cite{brown78,ferreira13} and simulations
    with the Berger/H\"oltje, CHARMM36 and MacRog force fields. The signs in the experimental values are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
    and Gross et al.~\cite{gross97}.  The most order parameters from Berger/Höltje model for g$_1$ are beoynd the y-axis scale.}
\end{figure}

In addition to the experimental data, the previously published simulation results from the Berger/H\"oltje model~\cite{ferreira13},
and our results from CHARMM36 and MacRog force fields  
are shown in Fig.~\ref{ordPchol}. As already pointed out previosly, the Berger/H\"oltje model
seriously overestimates the effect of cholesterol on the phospholipid glycerol backbone and choline segments~\cite{ferreira13}.
In contrast, the responses of CHARMM36 and MacRog are in better agreement
with experiments. However, in the case of MacRog the possibility of discrepancy with 50 mol-\% of cholesterol
and above cannot be excluded since there is data only up to 40 mol-\% 
\todo{It would be nice to have data also for 50 mol-\% or above with MacRog to get rid of this sentence}. 
CHARMM36 seems to even reproduce the modest changes observed experimentally in glycerol backbone segments 
g$_2$ and g$_3$ with high concetrations of cholesterol \todo{It would be nice to look from CHARMM36 simulations
what happens to the g3 segment (the slight increase of the order parameters as a function of cholesterol)
and see if we could explain it.}.

It should be noted that the CHARMM36 force field parameters (dihedral potentials) for the glycerol backbone have been tuned 
the dihedral potentials to reproduce the correct order parameters at fully hydrated conditions~\cite{klauda10}. 
This procedure contains a risk of overfitting, which would manifest itself as wrong responses to changing conditions. 
According to our results, tuning seems not to lead to overfitting problems in the case of dehydration or lipid-cholesterol mixtures.

\section{Conclusions}
The atomistic resolution structures sampled by the glycerol backbone and choline headgroup
in phoshatidylcholine bilayers are not known despite of vast amount of accurate experimental
data. Atomistic resolution molecular dynamics simulation model which would reproduce the
experimental data would automatically resolve the structures giving an interpretation of experimental results.
In this work we have collected and reviewed the experimental C-H bond vector order
parameters available in literature. These experimental parameters are then compared to
different atomistic resolution simulation models for fully hydrated bilayer, dehydrated bilayer and
lipid bilayer containing cholesterol. Our results have led to the following conclusions:

- The C-H bond order parameters measured with different NMR techniques are in good agreement
with each others. By combining the experimental results from various sources we concluded
that the order parameters for each C-H bond are known with quantitative accuracy of $\pm$0.02.

- None of the tested models (12 different models  \todo{this number may change}) produces the order parameters with the experimental
accuracy for fully hydrated phoshatidylcholine lipid bilayer. However, the CHARMM36, GAFFlipid and MacRog 
models are relatively close. The structures of these models together with the most used lipid model (Berger) 
were subjected to more careful studies. The results revealed that the current models are not accurate
enough to resolve the atomistic resolution structures sampled by glycerol backbone and choline headgroup.  
However, the correlation between dihedral angle distributions and order parameter differences was found, 
suggesting that careful adjustment of dihedral potentials would potentially lead to the model with correct
structure.

- Independent of the accuracy for fully hydrated lipid bilayer, all the models reproduced the choline response
to the dehydration. This can be explained by the change in the P-N vector tilting more parallel to the membrane
which leads to the increase of order parameters despite of the initial configuration. It should be however noted
that the correct qualitative response do not necessarily indicate correct energetics.

- The response of glycerol backbone and choline headgroup to the cholesterol content is described more
realistically in CHARMM36 and MacRog models than in the Berger model.

In general, we conclude that the atomistic resolution classical molecular dynamics simulations 
is extremely convenient tool to give structural interpretation for the high resolution NMR data~\cite{ferreira14}. 
However, in the case of phoshatidylcholine glycerol backbone and choline headgroup there is some
further model development required.

This work has been, and continues to be, progressed and discussed through the blog: nmrlipids.blogspot.fi. 
Everyone is invited to join the discussion and make contributions through the blog. 
The manuscript will be eventually submitted to an appropriate scientific journal. 
Everyone who has contributed to the work through the blog will be offered 
coauthorship. For more details see: nmrlipids.blogspot.fi.

{\bf Acknowledgements: }
OHSO acknowledges Tiago Ferreira and Paavo Kinnunen for useful discussions, the Emil Aaltonen foundation for financial support, Aalto Science-IT project and CSC-IT Center for Science for computational resources.

\bibliographystyle{apsrev}
\bibliography{refs}

\newpage

\onecolumngrid

 \listoftodos

\end{document}

### 2015-01-28 HGmodelMANUSCRIPT/HGmodel_draft3.tex
Added table ranking the force fields at full hydration.

@@
 However, three force fields perform better than others, namely CHARMM36~\cite{klauda10}, MacRog~\cite{maciejewski14} and GAFFlipid~\cite{dickson12}.
-The results for each force field in respect to the above criteria are summarized in Table~\ref{??} \todo{This kind of table should be done or not?}.
+The results for each force field in respect to the above criteria are summarized in Figure~\ref{FullHydrationComparisonTable}.
+\todo{This kind of table should be done or not? DONE, see Fig.~\ref{FullHydrationComparisonTable}}.
+\begin{figure}[]
+  \centering
+  \includegraphics[width=8.6cm]{../DATAreportediINblog/comparisonTable.pdf}
+  \caption{\label{FullHydrationComparisonTable}
+  Ranking of force fields based on data of Fig.~\ref{HGorderparameters}.
+  } 
+\end{figure}
 

### 2015-01-31 HGmodelMANUSCRIPT/HGmodel_draft3.tex
Updated the dehydration plot.
Added modified cholestol figure, acknowledgements of MSM, and fixed typos in addresses.

@@
 \author{Patrick Fuchs}
-\thanks{Institut Jacques Monod, CNRS, Université Paris Diderot, Sorbonne Paris Cité, Paris, France}
+\thanks{Institut Jacques Monod, CNRS, Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
 %\thanks{Université Paris Diderot, Sorbonne Paris Cité, Paris, France}
@@
 \author{Markus S. Miettinen}
-\thanks{Fachbereich Physik, Freie Universität Berlin, Berlin, Germany}
+\thanks{Fachbereich Physik, Freie Universit\"at Berlin, Berlin, Germany}
 \author{Luca Monticelli}
@@
 \author{Marius Retegan}
-\thanks{Max Planck Institute for Chemical Energy Conversion, Mülheim an der Ruhr, Germany}
+\thanks{Max Planck Institute for Chemical Energy Conversion, M\"ulheim an der Ruhr, Germany}
 \author{Hubert Santuz}
 \thanks{INSERM, U1134, DSIMB; Institut National de la Transfusion Sanguine (INTS); Laboratoire d'Excellence GR-Ex, Paris, France}
-\thanks{Université Paris Diderot, Sorbonne Paris Cité, Paris, France}
+\thanks{Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
 \author{Joona Tynkkynen}
@@
 This work has been done as an open collaboration by using \texttt{nmrlipids.blogspot.fi} as a communication platform.
-The approach is inspired by the Polymath project~\cite{gowers09}, however there are some essential differences. The manuscript pointing out problems in the glycerol backbone and
+The approach is inspired by the Polymath project~\cite{gowers09}, however there are some essential differences. 
+The manuscript pointing out problems in the glycerol backbone and
 headgroup structure in the most used molecular dynamics simulation model for lipid bilayers (Berger based models) was used as a starting point~\cite{ollila13}.
@@
 http://nmrlipids.blogspot.fi/2014/04/on-signs-of-order-parameters.html somewhere.
-Maybe we should change the name of this section II.B. to: "Analysis and terminology of order parameters from experiments and simulations" or similar, and also move the "forking" explanation here.}
+Maybe we should change the name of this section II.B. to: "Analysis and terminology of order parameters from experiments 
+and simulations" or similar, and also move the "forking" explanation here.}
 The order parameter of a hydrocarbon C--H vector is defined as $S_{{\rm CH}}=\frac{3}{2}\langle \cos^2 \theta-1 \rangle$, where
@@
   \centering
-  \includegraphics[width=8.6cm]{OrderParameterDEHYDpresentation.eps}
-\todo{Several people think that this plot should be clarified.}
+  \includegraphics[width=8.6cm]{../DATAreportediINblog/dehydration.pdf}
+\todo{Several people think that this plot should be clarified. ------> DONE.}
   \caption{\label{ordPhydr}
@@
   \centering
-  \includegraphics[width=8.6cm]{OrderParameterCHOLexpVSsim.eps}
+  \includegraphics[width=8.6cm]{../DATAreportediINblog/cholesterolization.pdf}
   \caption{\label{ordPchol}
@@
 
+\pagebreak
 \section{Conclusions}
@@
 
+MSM acknowledges  financial support from the Volkswagen Foundation (86110).
 \bibliographystyle{apsrev}

### 2015-02-05 HGmodelMANUSCRIPT/HGmodel_draft3.tex
Added modified PN-vector-vs-hydration plot
updating information from blog to the manuscript

@@
 between lipids and cholesterol better than the Berger/H\"oltje model.
-This work has been, and continues to be, progressed and discussed through the blog: \texttt{nmrlipids.blogspot.fi}. 
+This work has been, and continues to be, progressed and discussed through the blog: \url{nmrlipids.blogspot.fi}. 
 Everyone is invited to join the discussion and make contributions through the blog. 
@@
 Everyone who has contributed to the work through the blog will be offered 
-coauthorship. For more details see: \texttt{nmrlipids.blogspot.fi}.
+coauthorship. For more details see: \url{nmrlipids.blogspot.fi}.
 
@@
 
-This work has been done as an open collaboration by using \texttt{nmrlipids.blogspot.fi} as a communication platform.
+This work has been done as an open collaboration by using \url{nmrlipids.blogspot.fi} as a communication platform.
 The approach is inspired by the Polymath project~\cite{gowers09}, however there are some essential differences. 
@@
 
-Large portion of the used simulation files and scripts are shared through the GitHub organization \texttt{https://github.com/NMRlipids}
-and the Zenodo community \texttt{https://zenodo.org/collection/user-nmrlipids}.
+Large portion of the used simulation files and scripts are shared through the GitHub organization \url{https://github.com/NMRlipids}
+and the Zenodo community \url{https://zenodo.org/collection/user-nmrlipids}.
 
@@
 with each other (see Results and Discussion), indicating very high quantitative accuracy of the measurements.
-For a more detailed discussion see \texttt{http://nmrlipids.blogspot.fi/2014/02/accuracy-of-order-parameter-measurements.html}
+For a more detailed discussion see \url{http://nmrlipids.blogspot.fi/2014/02/accuracy-of-order-parameter-measurements.html}
 \todo{This is temporarily linked to the blog. Currently I think that the posts (with comments included) to which we want to cite are put into 
@@
 Berger-POPC-07~\cite{ollila07a}          &   POPC & 128 & 7290  & 298  & 270 & 240 & \cite{bergerFILESpopc}$^*$ & \cite{ferreira15} \\
-\sout {Berger-POPC-07~\cite{ollila07a}}\todoi{Not included in the Fig. 2, to be removed?}          &   POPC & 128 & 7290  & 323  & 56 & 56  & ? & SI \\
+%\sout {Berger-POPC-07~\cite{ollila07a}}\todoi{Not included in the Fig. 2, to be removed?}          &   POPC & 128 & 7290  & 323  & 56 & 56  & ? & SI \\
+\todoi{DONE} 
 Berger-DPPC-98~\cite{marrink98}          &   DPPC & 72 & 2864  & 323  & 140 & 100  & \cite{bergerDPPCfiles} & SI \\
@@
 CHARMM36\cite{klauda10}       & POPC   & 72  &  2242 & 303 & 30 & 20  & \cite{charmm36filesSHORT}$^*$ & SI \\
-CHARMM36\cite{klauda10}      & POPC   & 128 &  5120    & 303 & 150 & 100  & ? \todoi{Permanent link in progress by Hubert Santuz}   & SI \\
+CHARMM36\cite{klauda10}      & POPC   & 128 &  5120    & 303 & 150 & 100  & \cite{charmm36files}$^*$ \todoi{DONE}   & SI \\
 CHARMM36\cite{klauda10}       & DPPC   & 72  &  2189 & 323 & 30 & 25  & ?\todoi{Samuli, put to Zenodo}  & SI \\
-CHARMM36\cite{klauda10}\todoi{Not included in the Fig.
+%CHARMM36\cite{klauda10}\todoi{Not included in the Fig.
 2.
 Should we remove or include?}        & DOPC   & 128 &  5120    & 303 & 150 & 100  & ?\todoi{Permanent link in progress by Hubert Santuz}  & SI \\
+\todoi{DONE} \todoi{Permanent link for DOPC CHARMM36 simulation by Hubert Santuz is not required now. However, it would be probably useful in the future.}
 MacRog\cite{maciejewski14}  & POPC & 288  & 12600 & 310 & 100 & 80  & \cite{macrogFILES}$^*$ & SI  \\
 GAFFlipid\cite{dickson12}       & POPC & 126  & 3948  & 303 & 137 & 32  & \cite{GAFFlipidFILES}$^*$ & SI \\
+GAFFlipid\cite{dickson12}       & DPPC & 72  & 2197  & 323 & 90 & 50  & \cite{GAFFlipidFILESdppc}$^*$ & SI \\
 Lipid14\cite{dickson14}         & POPC  & 72 & 2234 & 303 & 100 & 50  & \cite{lipid14files}$^*$ & SI \\
-Poger\cite{poger10}             & DPPC  & 128 & 5841 & 323 & 2$\times$100 & 2$\times$50 & \cite{pogerFILES}\todoi{Permanent link in progress by Patrick Fuchs} & SI \\
+Poger\cite{poger10}             & DPPC  & 128 & 5841 & 323 & 2$\times$100 & 2$\times$50 & \cite{pogerFILESpme1,pogerFILESpme2}$^*$\todoi{DONE} & SI \\
 Slipid\cite{jambeck12}          & DPPC & 128 & 3840 & 323 & 150 & 100 & \cite{slipidsFILES}$^*$ & SI \\
+Slipid\cite{jambeck12}          & POPC & 128 & 5120 & 303 & 200 & 150 & \cite{slipidsFILESpopc}$^*$ & SI \\
 Kukol\cite{kukol09}          & POPC   & 512 & 20564 & 298 & 50 & 30  & \cite{kukolFILES}$^*$ & SI \\
 Chiu et al.\cite{chiu09} \todoi{Samuli, put to Zenodo}     & POPC  & 128 & 3552  & 298 & 56 & 50  & ? & SI \\
-Rabinovich et al.\cite{rabinovich14}\todoi{Alexander Luybartsev, let us know if you share the files.}  & POPC   &  128 & 3840  & 303 & 100 & 80  & ? & \cite{rabinovich14}  \\
-H\"ogberg et al.\cite{hogberg08}\todoi{Alexander Luybartsev, let us know if you share the files.}  & DMPC   &  98 & 3840  & 303 & 75 & 50 & ? & \cite{hogberg08} \\
-Ulmschneider\cite{Ulmschneider09}    & POPC  & 128 & 3328 & 310 & 100 & 50 & \cite{ulmschneiderFILES}$^*$ & SI \\
+H\"ogberg08\cite{rabinovich14}\todoi{Alexander Luybartsev, let us know if you share the files.}  & POPC   &  128 & 3840  & 303 & 100 & 80  & ? & \cite{rabinovich14}  \\
+H\"ogberg08\cite{hogberg08}\todoi{Alexander Luybartsev, let us know if you share the files.}  & DMPC   &  98 & 3840  & 303 & 75 & 50 & ? & \cite{hogberg08} \\
+Ulmschneiders\cite{Ulmschneider09}    & POPC  & 128 & 3328 & 310 & 100 & 50 & \cite{ulmschneiderFILES}$^*$ & SI \\
 Tj\"ornhammar et al.\cite{tjornhammar14}   & DPPC  & 144 & 7056 & 323 & 200 & 100 & \cite{tjornhammarfiles}$^*$ & \cite{tjornhammar14} \\
@@
                                &   POPC &50 & 78 & 5782   & 298  & 100 & 80 & \cite{bergerFILESpopc60chol}$^*$ & \cite{ferreira13} \\
-CHARMM36\cite{klauda10,lim12}\todoi{Hubert Santuz, these files are already in Gitbub, however to get doi citation it should be put to Zenodo}    & POPC   & 128&  0 & 5120  & 303  & 150 & 100 & ?
-& SI  \\
- \todoi{Hubert Santuz, these files are already in Gitbub, however to get doi citation it should be put to Zenodo}                                & POPC   & 100  & 24  &  4960   & 303 & 200 & 100 & ?
-& SI \\
-  \todoi{Hubert Santuz, these files are already in Gitbub, however to get doi citation it should be put to Zenodo}                               & POPC   & 80  & 80  &  4496    & 303 & 200 & 100 & ?
-& SI \\
+CHARMM36\cite{klauda10,lim12}\todoi{DONE}    & POPC   & 128&  0 & 5120  & 303  & 150 & 100 & \cite{charmm36files}$^*$  & SI  \\
+ \todoi{DONE}                                & POPC   & 100  & 24  &  4960   & 303 & 200 & 100 & \cite{charmm36files20perCHOL}$^*$ & SI \\
+  \todoi{DONE}                               & POPC   & 80  & 80  &  4496    & 303 & 200 & 100 & \cite{charmm36files50perCHOL}$^*$ & SI \\
 MacRog\cite{maciejewski14}     & POPC   & 128 & 0  & 6400   & 310 & 400 & 200 & \cite{macrogCHOLfiles}$^*$ & SI \\ 
@@
   \includegraphics[width=8.6cm]{../DATAreportediINblog/comparisonSorted.pdf}
-%\todo{DONE}\\
-%The data in this figure should be updated to be consistent with the systems listed in Table~\ref{systems}.
-At least, Tjornhammar et al and CHARMM-UA values are missing.
-Some results in the figure are outdated due to longer simulations reported etc.} \\
 \todo{Markus Miettinen suggested that we should consider making one more figure where only experimental data would be shown and that would be discussed in Section~\ref{experiments}} \\
-\todo{Markus has now made a new figure with signs included.
-I think that this is very good.
-I also think that we should try to get the interactive figure behind permalink (i.e.
-Zenodo or figshare) and cite that in the paper.}
+\todo{I think that this figure is now very good. There is also the interactive version by Hubert Santuz now in https://plot.ly/~HubertSantuz/72/lipid-force-field-comparison/
+we should figure out which is the most practical way to put that behind permalink once it finalized (Zenodo, figshare or something else?) and then put a citation in the paper.}
   \caption{\label{HGorderparameters}
@@
 The results for each force field in respect to the above criteria are summarized in Figure~\ref{FullHydrationComparisonTable}.
-\todo{This kind of table should be done or not? DONE, see Fig.~\ref{FullHydrationComparisonTable}}.
+\todo{The table is now done, see Fig.~\ref{FullHydrationComparisonTable}. The text should be updated accordingly.}.
 \begin{figure}[]
@@
 even very small order parameter changes resulting from the varying conditions can be measured (see
-\texttt{http://nmrlipids.blogspot.fi/2014/02/accuracy-of-order-parameter-measurements.html}
+\url{http://nmrlipids.blogspot.fi/2014/02/accuracy-of-order-parameter-measurements.html}
 for more discussion.
@@
   \includegraphics[width=8.6cm]{../DATAreportediINblog/dehydration.pdf}
-\todo{Several people think that this plot should be clarified. ------> DONE.}
+\todo{Figure has been modified to elucidate changes. The text should be updated accordingly. The same applies to the cholesterol figure.}
   \caption{\label{ordPhydr}
@@
   \centering
-  \includegraphics[width=8.6cm]{PNangles.eps}
+  \includegraphics[width=8.6cm]{../DATAreportediINblog/dehydrationPN.pdf}
 
@@
 
-This work has been, and continues to be, progressed and discussed through a blog at \texttt{nmrlipids.blogspot.fi}. 
+This work has been, and continues to be, progressed and discussed through a blog at \url{nmrlipids.blogspot.fi}. 
 Everyone is invited to join the discussion and make contributions through the blog. 
@@
 Everyone who has contributed to the work through the blog will be offered 
-coauthorship. For more details, see \texttt{nmrlipids.blogspot.fi}.
+coauthorship. For more details, see \url{nmrlipids.blogspot.fi}.
 
@@
 solvate the system.
-The publicly available CHARMM36 forcefield parameters (\texttt{http://www.gromacs.org/@api/deki/files/184/=charmm36.ff\\\_4.5.4\_ref.tgz}) 
+The publicly available CHARMM36 forcefield parameters (\url{http://www.gromacs.org/@api/deki/files/184/=charmm36.ff\\\_4.5.4\_ref.tgz}) 
 by Piggot et al. \cite{piggot12} were used. Cholesterol parameters came
-from Lim et al. \cite{lim12} and were converted into GROMACS format with the PyTopol tool (\texttt{https://github.com/resal81/PyTopol}).  
+from Lim et al. \cite{lim12} and were converted into GROMACS format with the PyTopol tool (\url{https://github.com/resal81/PyTopol}).  
 Single point energy calculation was done to assess the conversion. 
@@
 
-The force field parameters were generated using files obtained from the Lipidbook website (\texttt{http://lipidbook.bioch.ox.ac.uk/package/show/id/150.html})~\cite{domanski10}. 
+The force field parameters were generated using files obtained from the Lipidbook website (\url{http://lipidbook.bioch.ox.ac.uk/package/show/id/150.html})~\cite{domanski10}. 
 The conversion to GROMACS compatible formats was performed using the acpype tool~\cite{silva12}. The accuracy of the conversion was checked by calculating 
@@
 The initial hydrated bilayer structure of 128 DPPC and 5841 water molecules as well as force field parameters were downloaded
-from David Poger's web site (\texttt{http://compbio.chemistry.uq.edu.au/~david/}) on April 2012. 
+from David Poger's web site (\url{http://compbio.chemistry.uq.edu.au/~david/}) on April 2012. 
 We noticed that the same files downloaded in October 2013 appear to lack two dihedral angles in the choline headgroup (only one dihedral of type gd\_29 allowing 
@@
 the Parrinello--Rahman barostat~\cite{parrinello81} using a 4~ps time constant and a compressibility of 4.5e-5~bar$^{-1}$. For non-bonded interactions, two conditions were tested:
-i) A 0.8--1.4 nm twin-range cut-off with the neighbor list updated every 5 steps for both electrostatics and Lennard-Jones (LJ) interactions. For the former the generalized reaction 
+i) A 0.8--1.4 nm twin-range cut-off with the neighbor list updated every 5 steps for both electrostatics and Lennard-Jones (LJ) interactions 
+(simulation files available at~\cite{pogerFILESrf1,pogerFILESrf2}). For the former the generalized reaction 
 field (RF) with a dielectric permittivity of 62 was used beyond the 1.4~nm cut-off~\cite{tironi95}. This is the original setup that Poger et al.~\cite{poger10} used.
 ii) PME~\cite{darden93,essman95} electrostatics with a real space cut-off of 1.0~nm, a Fourier spacing of 0.12~nm and an interpolation order of 4, LJ interactions computed with a 1.0--1.4~nm twin-range cut-off,
-neighbor list updated every 5 steps. Note that Poger and Mark tested the effect of PME vs RF in ref.~\cite{poger12}, but used a 1.0~nm cut-off with PME and 1.4~nm with RF for LJ
+neighbor list updated every 5 steps (simulation files available at~\cite{pogerFILESpme1,pogerFILESpme2}). 
+Note that Poger and Mark tested the effect of PME vs RF in ref.~\cite{poger12}, but used a 1.0~nm cut-off with PME and 1.4~nm with RF for LJ
 interactions. Since 0.8--1.4~nm twin-range cut-off for LJ interations is used in the parametrization of GROMOS force field, we decided to use that
@@
 GROMACS versions 4.0.X and earlier (the original authors~\cite{poger10} used GROMACS version 3.3.3). When switching to versions 4.5.X and above, the area per lipid dropped to below 0.58~nm$^2$. 
-The GROMACS developers were contacted and a redmine issue opened (\texttt{http://redmine.gromacs.org/issues/1400}). The difference comes from the new Trotter decomposition 
+The GROMACS developers were contacted and a redmine issue opened (\url{http://redmine.gromacs.org/issues/1400}). The difference comes from the new Trotter decomposition 
 introduced in versions 4.5.X. A fix has been introduced in version 4.6.6 that allows a recovery of an area per lipid value of 0.615~nm$^2$. The results in terms of area per lipid using the different 
-GROMACS versions are available at~\cite{fuchs14}\todoi{Permanent link in progress by Patrick Fuchs}.
+GROMACS versions are available at~\cite{pogerFILESrf2}\todoi{DONE}.
 % http://www.dsimb.inserm.fr/~fuchs/project\_Samuli/Poger\_DPPC/tests\_gmx\_versions/)[Samuli, you may want to add this Figure+Legend+Comment 
@@
 
-Initial coordinates for a hydrated DPPC (at 323~K) and POPC (at 310~K) bilayers (30 and 40 waters/lipid, respectively) were taken directly from the Slipids home page \texttt{http://people.su.se/\~jjm/Stockholm\_Lipids/Downloads.html}.
+Initial coordinates for a hydrated DPPC (at 323~K) and POPC (at 310~K) bilayers (30 and 40 waters/lipid, respectively) were taken directly from the Slipids home page \url{http://people.su.se/\~jjm/Stockholm\_Lipids/Downloads.html}.
 The Slipids force field~\cite{jambeck12,XXX} ADD 2ND SLIPIDS PAPER!
@@
 \subsection{General issues to think of about the manuscript}
-\todo{Should we emphasize more the excellent collection of simulation trajectories collected to Zenodo?}
+\todo{Should we emphasize more the excellent collection of simulation trajectories collected to Zenodo?} \\
 \todo{Maybe we should mention that CHARMM is significanlty more computationally expensive than other all-atom models.

### 2015-02-06 LIPIDionINTERACT/LIPIDionINTERACT.tex
publish first draft of lipid-ion interaction manuscript

% \documentclass[rmp,aps,floatfix,authordate1-4,preprint]{revtex4}
% \documentclass[prb,aps,floatfix,authordate1-4,preprint]{revtex4}

\documentclass[pre,aps,floatfix,authordate1-4,twocolumn]{revtex4-1}
%\documentclass[pre,aps,floatfix,authordate1-4]{revtex4-1}

%\documentclass[aps,prl,preprint,superscriptaddress]{revtex4}

%\documentclass[aps,prl,preprint,groupedaddress]{revtex4}

\usepackage{rotating} 
\usepackage{times}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{caption}
\usepackage{subcaption}
%\usepackage[sort&compress]{natbib}
%\bibpunct{(}{)}{,}{n}{}{}
%\renewcommand{\bibnumfmt}[1]{#1.}
\usepackage[obeyFinal]{easy-todo}

\begin{document}
%\singlespacing

\title{Binding of cations to phospholipid bilayers}

%\author{Patrick Fuchs}
%\author{Matti Javanainen}
%\author{Antti Lamberg}
%\author{Alexander Lyubartsev}
%\author{Markus S. Miettinen}
%\author{Luca Monticelli}
%\author{Jukka Määttä}
%\setcounter{page}{1}
\author{O. H. Samuli Ollila}
\footnote{Author to whom 
correspondence may be addressed. E-mail: samuli.ollila@aalto.fi.\\
}
\affiliation{
  Helsinki Biophysics and Biomembrane Group,
Department of Biomedical Engineering and Computational Science, Aalto University, Espoo, Finland
}
%\author{Marius Retegan}
%\author{Hubert Santuz}
%\author{Joona Tynkkynen}
%\author{Alexander Vogel}

\begin{abstract}
%We compare the order parameters predicted for the hydrocarbon segments in lipid bilayer headgroup region
%by the Berger molecular dynamics simulation model to those measured by Nuclear Magnetic Resonance (NMR) experiments.
%We first show results for a fully hydrated POPC bilayer, and then focus on changes of the order parameters as
%a function of hydration level, NaCl and CaCl$_2$ concentrations, and cholesterol content. The experimental headgroup order 
%parameters are never reproduced.
%This indicates that under all of these conditions the used model is unable to correctly reproduce the headgroup structure. 
%Consequently, many of the conclusions drawn over the years from this model might be erroneous.
%This manuscript has not been submitted to any journal, instead its contents are discussed at 
%nmrlipids.blogspot.fi.
Despite of vast amount of experimental and theoretical studies the binding affinity of cations, especially biologically relevant Na$^+$ and Ca$^{2+}$ ions,
into a phosholipid bilayer is not agreed in the literature. Here we directly compare the measured choline headgroup order parameters
to the simulations with different models in the presence of different cations. We conclude that the least complicated explanation
for the experimental and theoretical observations is that the Na$^+$ ions do not penetrate into a phosphatidylcholine lipid
bilayer in mM concentrations, in contrast to Ca$^{2+}$. Further, the binding affinity of Na$^+$ is overestimated in almost all molecular dynamics simulation models.
However, the electrometer concept connecting the choline order parameter changes to the penetrating charges is valid also in simulations.
This work has been, and continues to be, progressed and discussed through the blog: nmrlipids.blogspot.fi.
Everyone is invited to join the discussion and make contributions through the blog. The manuscript will be
eventually submitted to an appropriate scientific journal. Everyone who has contributed to the work through the
blog will be offered coauthorship. For more details see: nmrlipids.blogspot.fi.
\end{abstract}

\maketitle

~\vspace{0.3cm}\\
{\it \bf}

\section{Introduction}

The cation interactions with phospholipid membranes are present in large amount
of physiological processes, nerve cell signalling being the prime example.
Thus, the interactions between different cations and phospholipid bilayers have been very widely studied experimentally
and theoretically. It is practically agreed that the relative binding affinity of different
ions follows the Hofmeister series~\cite{eisenberg79,cevc90,tocanne90,binder02,celma07,leontidis09,vacha09a,klasczyk10,harb13}, however the quantitative binding affinities of different
ions are not agreed in the literature. The extensive reviews about work done prior 1990~\cite{cevc90,tocanne90}
concluded that monovalent cations (lithium being an exception) interact only weakly with phopholipid bilayers, 
while the interactions are significant for multivalent ions. This conclusion has been supported by
further studies where the bilayer poperties have been intact with increasing monovalent salt 
concentration~\cite{binder02,pabst07,filippov09}. On the other hand, the weak interactions of monovalent ions, especially
sodium, has been also questioned in several more recent studies suggesting stronger 
binding~\cite{bockmann03,bockmann04,vacha09a,manyes05,manyes06,fukuma07,leontidis09,ferber11,morata12,klasczyk10,harb13}. In this work we concentrate on the 
interactions between phospholipid bilayer and two widely studied and biologically relevant cations, monovalent sodium and divalent calcium.

The experimental studies supporting the weak interactions between sodium ions and phospholipid bilayer
have shown that NaCl with mM concentrations has a negligible effect on 
choline headgroup order parameters~\cite{akutsu81}, area per molecule~\cite{pabst07}, dipole potential~\cite{clarke99}
and lipid lateral diffusion~\cite{filippov09} while these properties were significantly affected by the presense
of CaCl$_2$ or other multivalent ions. In addition, water sorption isotherm for POPC/NaCl system
was essentially similar to NaCl in pure water indicating only weak interaction between ion and lipid~\cite{binder02}.
Only minor changes in POPC infrared spectra were observed in the presense of NaCl compared to the significant 
changes in the presense of CaCl$_2$ and other multivalent ions, and it was again concluded that the Na$^+$-lipid interactions are weak~\cite{binder02}.

%More specifically, the conclusion that monovalent ions (except lithium)
%have very mild effects on phase transition temperature, electrophoretic 
%mobility, infrared spectra, area per molecule, bending rigidity, deuterium order parameters
%and lateral diffusion of lipids has been supported also by several more recent studies~\cite{akutsu81,cevc90,binder02,petrache06,pabst07,filippov09}.
%The exception in these studies was again lithium, for which larger effects were observed~\cite{cevc90}.
%Multivalent ions, on the other hand, typically have a larger effect on the above 
%quantities~\cite{akutsu81,cevc90,binder02,pabst07}.

In contrast, experimentally observed decrease in fluorescent probe rotational and translational dynamics
in lipid bilayer due to addition of NaCl has suggested significant ion binding also with mM 
concentrations~\cite{bockmann03,vacha09a,harb13}. However, the reduced lateral diffusion is not observed
in NMR experiments without probes suggesting that it would arise from interactions between probe and ion,
not between lipid and ion~\cite{filippov09}. NaCl binding to phospholipids leading to the change of the strength of 
bilayer against rupture and area per lipid reduction has been also suggested from Atomistic Force Microscopy (AFM)~\cite{manyes05,manyes06,fukuma07,ferber11,morata12}.
Interpretation of calorimetric measurements has been controversial: Previously the small effect of
monovalent ions (except Lithium)  on phase transition temperature compared to multivalent ions was interpreted such that 
only multivalent ions and Lithium specifically bind to phosholipid bilayer~\cite{cevc90}, however more recently the
small changes in calorimetric experiments have been interpreted to indicate also Na$^+$ binding~\cite{bockmann03,klasczyk10}.
In electrophoresis measurements of phosphatidylcholine vesicles the NaCl can increase the originally negative zeta potential 
close to zero, however positive zeta potential can be typically reached only with multivalent ions or Lithium
ions~\cite{eisenberg79,tatulian87,manyes05,manyes06,klasczyk10}. The lack of significant positive electrophoretic mobility in the presence
of NaCl has been rekognized to contradict with suggested strong binding of Na$^+$ and the contradiction
has been explained by the effect of cl$^-$ ions to the electrophoretic mobility~\cite{berkowitz06,knecht13}.

%However, the origin of negative mobility of vesicles in the absence of ions is not known and 
%electrophoretic mobility depends also on the location of slip plane, not only on directly bound ions~\cite{??},
%thus the definite conclusions about Na-lipid interactions are difficult to draw from these experiments.

%in some atomistic simulations both, Na$^+$ and Ca$^{2+}$ ions have been observed to bind  
%lipid bilayers and change their properties significantly~\cite{pandit03,bockmann03,bockmann04}. Support for these results
%has been found from calorimetry, AFM experiments and diffusion measurements with 
%probes~\cite{bockmann03,vacha09a,garciamanyes05,fukuma07,klasczyk10}.
%The conclusions from these experiments, however, often contradict with the conclusions from 
%other calorimetric experiments and spectroscopy~\cite{akutsu81,cevc90,binder02},
%and it has been shown that the results from probe diffusion measurements arise from the
%behaviour of the probe, not lipids~\cite{filippov09}.

%Especially the Na$^+$ binding has been widely discussed in recent molecular dynamics simulation
%literature. 
In atomistic resolution molecular dynamics simulations all the generally used models seems
to predict binding of Na${^+}$ ions into a phoshatidylcholine lipid bilayer, 
but the strength of binding depends on the force field~\cite{bockmann03,bockmann04,sachs04,berkowitz06,cordomi09,valley11,berkowitz12}. 
%Generally the CHARMM based force fields predict less Na$^+$ binding and changes in bilayer
%properties than the GROMOS based force fields~\cite{sachs04,valley11}.
The reduction of lipid lateral diffusion due to Na binding in simulations is in agreement with 
fluorescent probe measurements~\cite{bockmann03,vacha09a,harb13}, but not with the NMR experiments~\cite{filippov09}.
The AFM experiments are interpereted to agree with the area per lipid reduction due to Na binding
in simulations~\cite{manyes05,manyes06,fukuma07,ferber11,morata12}, however compared to the scattering experiments the are reduction 
is observed with too low concentrations in simulations compared to experiments~\cite{pabst07}.
The simulations has been suggested to predict larger positive electrophoretic mobility than experimentally
measured with NaCl, which has been explained by choride ion behaviour~\cite{berkowitz06,knecht13}.

%It has been concluded that the simulation results are in agreement with 
%Comparisons have been done
%between simulations and elecrophoresis, calorimetric and 
%diffusion experiments~\cite{bockmann03,vacha09a,klasczyk11,knecht13,knecht13b}.
%However, direct comparison of atomistic simulations to these measurements is nontrivial,
%because the measured changes can arise from many different origins. For example, the
%electrophoretic mobility depends on the exact location of the shear plane, the changes in 
%water properties in the presense of ions might affect the results, and in diffusion
%experiments the ions directly affect the probe diffusion without actually interfering with
%the bilayer~\cite{filippov09}.

It is agreed in the literature that the Ca$^{2+}$ ions do penetrate in the phosphatidylcholine bilayer and
signifincantyl affect membrane properties already with mM concentrations of CaCl$_2$, however the strength of the binding is not agreed~\cite{tatulian87,altenbach84,bockmann04}.

One of the experimental methods used to measure the ion binding is the so called
''electrometer concept'' where the changes of the C-H bond order parameters for the 
choline $\alpha$ and $\beta$ segments (see Fig.~\ref{POPCstructure} for definitions and 1-palmitoyl-2-oleoylphosphatidylcholine (POPC) 
chemical structure) are used to determine the amount of penetrating charges into a bilayer~\cite{akutsu81,altenbach84,seelig87,scherer89}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{POPCstructure.eps}

  \caption{\label{POPCstructure}
    Chemical structure of POPC.}

\end{figure}
The concept is based on the findings that the absolute value of the $\beta$ order parameter
increases and the $\alpha$ order parameter decreases with inceasing amount of penetrating 
charges~\cite{akutsu81,altenbach84,seelig87,scherer89}. 
From the simulation point of view the electrometer concept is particularly interesting since one
can directly compare the actual measurable quantity, i.e. order parameters, to the simulations.
In this work we compare the ion induced changes in phosphatidylcholine headgroup order parameters
between different simulation models to resolve the above discussed controversies in the literature.

%In this work we concentrate on the effect of Na$^+$ and Ca$^{2+}$ since those are widely discussed 
%ions in the simulation literature~\cite{pandit03,bockmann03,bockmann04,??}, and the experimental data is available~\cite{akutsu81,altenbach84}.

In our recent work we showed that the experimental order parameters for choline and glycerol backbone were not
quantitatively reproduced by any available lipid model in fully hydrated conditions, however, the response of the choline order parameters
to the dehydration were qualitatively correct in all models and the response to the cholesterol  concentration
was correct for the CHARMM36 model~\cite{botan15}. In this work, we show that the qualitative response of order parameters
to penetrating cations is qualitatively correct but the partition of ions is too strong in all the studied models.

\section{Results and Discussion}
The effect of NaCl on the choline and glycerol bakcbone order parameters from experiments by Akutsu et al.~\cite{akutsu81}
and various simulations are shown in Fig.~\ref{ordPnacl}. 
The simulated systems are described in the table~\ref{IONsystems} and further details are given in Supplementary Information. 
The experimental results are shown for DPPC bilayer due to more extensive data set~\cite{akutsu81}, 
however the measurements for POPC gave essentially the same results~\cite{altenbach84}.

\begin{figure}[]
  \centering
  \includegraphics[width=8cm]{OrderParameterIONSnaclSIGN.eps}
  \caption{\label{ordPnacl}
    Order parameters as a function of NaCl concentration from simulations 
     with the Berger, CHARMM36, MacRog and Orange force fields compared to the experiments
     by Akutsu et al.~\cite{akutsu81}. The signs are assumed to be the same as measured by Hong et al.~\cite{hong95a}, Hong et al.~\cite{hong95a} and Gross et al.~\cite{gross97}. 
     The effect of ions to the g2 and g3 were not measured, thus only the values without ions are shown. The straight line between the results with and without ions is 
     plotted to guide the eye. 
   }
\end{figure}

\begin{table*}[htb]
\centering
\caption{Simulated lipid bilayers with ions.
}\label{IONsystems}
\begin{tabular}{c c c c c c c c c c c c}
%\hline
Force field & lipid & [Ion] mM & \footnote{The number of lipid molecules}N$_{\rm l}$   &  \footnote{The number of water molecules}N$_{\rm w}$   & \footnote{The number of Na$^+$ molecules}N$_{\rm Na}$  & \footnote{The number of Ca$^{2+}$ molecules}N$_{\rm Ca}$   &  \footnote{The number of Cl molecules}N$_{\rm Cl}$ & \footnote{Simulation temperature}T (K)  & \footnote{The total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time frames used in the analysis}t$_{{\rm anal}}$ (ns) & Files\\
\hline
Berger-POPC-07\cite{ollila07a}   &   POPC & 0          & 128 & 7290 & 0  & 0  & 0 & 298  & 270 & 240 & \cite{bergerFILESpopc}  \\
Berger-POPC-07\cite{ollila07a}, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}   &   POPC & 190 (NaCl) & 128 & 7202 & 44  & 0  & 44 &298  & 110 & 50 &?\todoi{Samuli put to Zenodo} \\
Berger-POPC-07\cite{ollila07a}, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}   &   POPC & 190 (CaCl$_2$) & 128 & 7157 & 0 & 44  & 88 &298 & 110 & 50 &?\todoi{Samuli put to Zenodo}  \\
CHARMM36\cite{klauda10}   & POPC & 0           & 72 & 2242 & 0  & 0 & 0 & 303  & 30 & 20 & \cite{charmm36filesSHORT} \\
CHARMM36\cite{klauda10}, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}   & POPC & 140 (NaCl)  & 72 & 2085 & 13  & 0 & 13 & 303  & 80 & 60 &?\todoi{Samuli put to Zenodo}\\
CHARMM36\cite{klauda10}, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}   & POPC & 280 (NaCl)  & 72 & 2085 & 26  & 0 & 26 & 303  & 80 & 60 &?\todoi{Samuli put to Zenodo} \\
CHARMM36\cite{klauda10}, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}  & POPC & 570 (NaCl)  & 72 & 2085 & 52  & 0 & 52 & 303  & 80 & 60 &?\todoi{Samuli put to Zenodo} \\
CHARMM36\cite{klauda10}, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}   & POPC & 85  (NaCl)  & 72 & 2085 & 78  & 0 & 78 & 303  & 73 & 60 &?\todoi{Samuli put to Zenodo} \\
MacRog\cite{maciejewski14}\todoi{Is this the same as in the first manuscript for full hydration?}  & POPC & 0 & ? & ? & 0 & 0 & 0 & ? & ?&?  &?  \\
MacRog\cite{maciejewski14}\todoi{Matti Javanainen, please check the numbers and let us know if you share some files through Zenodo} , ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}  & POPC & ? (NaCl) & ? & ? & ? & ? & ? & ? & ?&?  &?  \\
MacRog\cite{maciejewski14}\todoi{Matti Javanainen, please check the numbers and let us know if you share some files through Zenodo} , ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}        & POPC &  ? (NaCl) & ? & ? & ? & ? & ? & ? & ?&?  &?  \\
MacRog\cite{maciejewski14}\todoi{Matti Javanainen, please check the numbers and let us know if you share some files through Zenodo} , ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}        & POPC &   ? (NaCl) & ? & ? & ? & ? & ? & ? & ?&?  &?  \\
MacRog\cite{maciejewski14}\todoi{Matti Javanainen, please check the numbers and let us know if you share some files through Zenodo} , ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}        & POPC &   ? (NaCl) & ? & ? & ? & ? & ? & ? & ?&?  &?  \\
Orange, ionFF~\cite{??}\todoi{Samuli check the numbers}   &   POPC & 0 & ? & ? & 0 & 0  & 0 & 298 & ? & ? &?\todoi{Jukka Määttä and Luca Monticelli, please let us know if we can share some files. This is unpublished model.}  \\
Orange, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}   &   POPC & 250 (NaCl) & 72 & 2802 & 26 & 0  & 26 & 298 & 120 & 60 &?\todoi{Jukka Määttä and Luca Monticelli, please let us know if we can share some files through Zenodo. This is unpublished model.}  \\
Orange, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}   &   POPC & 240 (CaCl$_2$)  & 72 & 2802 & 0 & 26  & 52 & 298 & 120 & 60 &? \todoi{Jukka Määttä and Luca Monticelli, please let us know if we can share some files. This is unpublished model.} \\
Slipid\cite{jambeck12}   &   DPPC & 0 & 128 &3840 & 0 & 0  & 0 & 323 & 150 & 100 &~\cite{slipidsFILES}  \\
Slipid\cite{jambeck12}, ionFF~\cite{??}\todoi{Andrea Catte, please check the numbers and let us know if you share some files through Zenodo}    &   DPPC & 150 (NaCl) & ? & ? & ? & ?  & ? & 323 & ? & ? &?  \\
\end{tabular}
\end{table*}

The most important observation is that the experimental order parameters for the headgroup and glycerol bakcbone g$_3$ segment 
are practically unhanged even though 1M NaCl concentration was added to the system. Thus, the presense of mM concentrations
of NaCl do not affect the structure of these parts of lipids. On the other hand, the presence of multivalent ions or 
charged amphiphiles changes the $\alpha$ and $\beta$ order parameters in a systematic way suggesting that the amount
and sign of charge penetrated into a bilayer can be measured from the changes of these parameters~\cite{akutsu81,altenbach84,seelig87,scherer89}.
This is known as the ''electrometer concept''. Thus, the most straightforward explanation for the experimental results shown
in Fig.~\ref{ordPnacl} is that the Na$^+$ and Cl$^-$ ions do not essentially penetrate into a phospholipid bilayer below 1M concentrations.

The response of order parameters to the NaCl concentration in simulations depends on the used model in Fig.~\ref{ordPnacl}:
The addition of NaCl leads to a significant changes for choline and $g_2$ and g$_3$ segments in glycerol bakcbone
in Berger and MacRog models, while only moderate changes are seen in CHARMM36 and Orange force fields.
The experimental data for glycerol backbone as a function of NaCl concentration is available only for the
g$_3$ segment. The changes in g$_1$ and g$_2$ are unlikely since all the other order parameters are unaffected
and $^1$H NMR data indicates that multivalent cations binding to phospholipid interact mainly with choline group
leaving glycerol backbone conformations intact~\cite{hauser76,hauser78}.

%Akutsu et al.~\cite{akutsu81} is shown in Fig.~\ref{ordPnacl} together with the results from 
%various simulations. The experiments were later on done also for POPC bilayer~\cite{altenbach84}, and the  
%results were essentially identical. The more extensive experimental data set DPPC bilayer is used here.

%The result is in line with the large number of experiments with different techniques where
%NaCl did not affect PC lipid bilayer properties with these concentrations~\cite{cevc90,binder02,petrache06,pabst07,filippov09}.

The core idea of the electrometer concept is that the decrease (increase) of $\beta$- and $\alpha$-carbon order parameters
is related to cation (anion) penetration into a lipid bilayer. 
It should be noted that in the original experiments the absolute value of $\beta$ order parameter was increasing with the presense of cations,
however it was shown later on that the $\beta$ order parameter is negative~\cite{hong95a,hong95b,gross97}, 
thus the value is actually increasing (becoming less negative).
The order parameter changes for  $\beta$- and $\alpha$-segments 
are plotted in Fig.~\ref{ordPions} in the presense of NaCl and CaCl$_2$ from experiments and different simulations. 
The clear decrease seen with CaCl$_2$ compared to the negligible effect of NaCl. Interpreted in terms of
electrometer concept, the result indicates that the Ca$^{2+}$ ions signifincantly penetrate into PC bilayer in contrast to 
Na$^+$~\cite{akutsu81,altenbach84}.

%The experimental results~\cite{akutsu81}, plotted in Fig.~\ref{ordPions} shows a distinct difference
%in behaviour as a function of ions: the choline order parameters were unaffected by the addition of NaCl, 
%whereas a systematic decrease was observed by the addition of CaCl$_2$. Following the electrometer
%concept this means that the binding constant for divalent Ca$^{2+}$ is higher than for monovalent Na$^+$
%which is in line with the large amount of experimental data with different techniques~\cite{cevc90,binder02,petrache06,pabst07,filippov09}.
\begin{figure}[]
  \centering
  \includegraphics[width=8cm]{OrderParameterIONSchanges.eps}
  \caption{\label{ordPions}
    The change of order parameters for $\beta$ and $\alpha$ segments as a function of NaCl (left column) and CaCl$_2$ (right column) concentrations from simulations 
    and experiments by Akutsu et al.~\cite{akutsu81}. The signs of the order parameters without ions are taken from~\cite{hong95a,hong95b,gross97} and it is reasonable
    to assume that the sign do not change with concentrations repsented here~\cite{altenbach84}. Note that the $\beta$ order parameter with negative sign decreases here with cation
    penetration in contrast to measured absolute value increase in the origical studues~\cite{akutsu81,altenbach84}.
    Regarding the actual electrometer concept this not relevant, however. 
    The experimental accuracy of order parameter changes is much higher than the accuracy of quantitative values due to high resolution of $^2$H NMR spectroscopy~\cite{??}. 
   }
\todo{I think that we have to run Orange simulation with higher NaCl concentration. I will do it.} \\
\todo{I think that we have to run CHARMM36 with CaCl$_2$. If anyone happen to have the data and is willing to share,
let us know. If not, then I will do it.} \\
\todo{ I think that it would be very interesting to test the modified CHARMM ions: Venable et al. dx.doi.org/10.1021/jp401512z, J. Phys. Chem. B 2013, 117, 10183−10192.
According to the paper, these parameters improve the ion binding to the charged lipid bilayers. Even though I think that testing this would be highly relevant, I do not have
time to do it now.}
\end{figure}

The differences in NaCl behaviour between different models already seen in Fig.~\ref{ordPnacl} are seen more clearly in Fig.~\ref{ordPions} showing the changes.
\todo{The discussion about simulation results with CaCl$_2$ to be written once we have all the results.}

%It should be noted that when the electrometer concept was developed and original experiments done in 1980s, 
%the signs of the order parameters were unknown. When the signs were measured in 1990s it turned
%out that the $\beta$ order parameter was negative while the $\alpha$ order parameter was 
%positive~\cite{hong95a,hong95b,gross97}. This means that actually the both $\beta$ and $\alpha$
%order parameters are decreasing with penetrating charge (absolute values are increasing and decreasing, respectively), as seen in Figs.~\ref{ordPions}.

%is based on measuring the changes in the
%$\beta$- and $\alpha$-carbon order parameters as a function of ion concetration~\cite{akutsu81,seelig87,scherer89,semchyschyn04}.
%Thus the changes in these order parameters as function of NaCl and CaCl$_2$ concentrations in experiments and different simulations
%are plotted in Fig.~\ref{ordPions}.

%The measured choline order parameters for DPPC as function of NaCl and CaCl$_2$~\cite{akutsu81} 
%are plotted in Fig.~\ref{ordPions}. 
%Also for the CaCl$_2$ experiments with POPC gives essentially similar results~\cite{altenbach84}.

%Also the results with Orange force field in Fig.~\ref{ordPions} shows a distinct difference between different salts:
%CaCl$_2$ significantly reduces the $\beta$- and $\alpha$-order parameters while significantly smaller or negligible effect is seen with NaCl. 
%Note, however, that in the Orange force field addition of Ca$^{2+}$ leads 
%to changes in order parameters that are larger than in experiments.

To study the correlation between order parameter changes and ion partitioning into a bilayer, 
suggested by the electrometer concept~\cite{akutsu81,altenbach84,seelig87,scherer89}, the density distributions as a function of membrane normal 
are shown in Fig.~\ref{IONdensCOMP} for lipids and cations from different simulation models. 
%to see if larger ion induced effects
%in choline order parameters correlates with ion binding also in simulation models.
\begin{figure}[]
  \centering
  \includegraphics[width=8cm]{IONdensities.eps}
  \caption{\label{IONdensCOMP}
    Mass density profiles for lipids and ions from simulations with different force fields. 
    The lipid densities are divided with 100 to make them visible with the used y-axis scale.
  }
\todo{This figure is in progress, however, the main point can be seen. \\
- I think that we should use number densities instead of mass densities to make the difference between Na and Ca distributions more clear. 
Now the difference comparison is little bit distracted due to the different mass of the atoms. \\
- The lipid density from MacRog simulation is missing. Maybe Matti Javanainen could provide this?
If the data will be uploaded to Zenodo, it also trivial to calculate it with script. \\
- It seems that relative integrated (total) ion density in different models do not correspond the reported relative concentrations.
For example, total Na density seems bigger in Berger than in MacRog even though there is only minot difference in concentration, and into the wrong direction. \\
- I think that CHARMM36 simulation with CaCl$_2$ should be done \\
- Figure editing will be finished (centering of the graphs, labels, etc.) will be done once the above issues have been sorted out.
 }
\end{figure}
It is evident from the figure that the Na$^+$ density peaks at the lipid bilayer interface are higher (indicating stronger ion binding) for the models
experiencing also larger changes in order parameters with NaCl concentration shown in Fig.~\ref{ordPnacl}. 
The both Na$^+$ binding affinity and order parameter changes are smallest, thus closest to the experiments, for the Orange model, and then incease 
in order CHARMM36, MacRog, and Berger having the highest binding affinity and order parameter changes.
%Simultaneously, it is evident from Figs.~\ref{ordPnacl} and~\ref{ordPions} that the order parameter decrease due to addition
%of NaCl is more pronounced in Berger and MacRog models compared to the CHARMM36 and Orange. 
Thus, the correlation between decreasing choline order parameters
and amount of penetrating cations (i.e. the electrometer concept) is also seen in simulations, independently of the quantitative quality of the used simulation model.

The most straightforward explanation for our results is that Na$^+$ ions do not practically penetrate into a PC lipid bilayer
with mM concentrations, thus the presense of NaCl does not affect the bilayer properties as observed in various experiments~\cite{akutsu81,altenbach84,clarke99,binder02,pabst07,filippov09}.
Consequently, the Na$^+$ penetration and concominant changes in order parameters, area per molecule and lateral diffusion 
seen in almost all simulation models would be artefact due to overestimated attraction between ions and lipid bilayer.
Even though this would also explain the absense of positive zeta potential in electrophoresis experiments~\cite{eisenberg79,tatulian87,manyes05,manyes06,klasczyk10},  
the presented data do not rule out the suggested possiblity of equal binding of Na$^+$ and Cl$^-$ ions~\cite{knecht13},
however, this equal binding should happen in such a way that the bilayer properties are unaffected.
Negligible binding of Na$^+$ with mM concentrations suggested here differs from the conclusions made from 
measurements of fluorescent probe dynamics~\cite{bockmann03,vacha09a,harb13}, membrane hardiness with AFM~\cite{manyes05,manyes06,fukuma07,ferber11,morata12} and calorimetry~\cite{bockmann03,klasczyk10}.
However, the fluorescent measurement results may arise from direct interactions between probe and ions, as already suggested by Filippov et al.~\cite{filippov09}. 
Further, the calorimetric results have been also interpreted to support negligible binding~\cite{cevc90}, and AFM result is relatively indirect, thus
there may be alternative explanations as well.

The effect of Ca$^{2+}$ on the $\alpha$ and $\beta$ order parameters is overestimated in all the simulation models, however,
qualitatively the change is correct, i.e. order parameters are decreasing.
The overestimation of the effect could be due to too strong partitioning or too sensitive headgroup model to the presense of ions.
\todo{The discussion should be finished when we have CHARMM results with CaCl$_2$.} \\
\todo{The $^1$H NMR experiments suggest that the N-$\beta$-$\alpha$-O dihedral is only in gaughe conformation in the absense of ions, but
in the presense of multivalent ions there would be also anti conformations present [Hauser et al. BBA {\bf 508}, 450 (1978), Hauser et al. Chem. Phys. Lipids {\bf 29}, 103 (1981)].
I think that we should check if this happens in simulations.}

%In conclusion, our results show that the Na$^+$ association to PC bilayer is significantly too strong in the Berger and MacRog models,
%while Orange and CHARMM36 force fields predict more realistic binding affinity. Since some modest changes are also seen in 
%Orange and CHARMM36 results, more careful simulation studies would be needed to judge if the Na$^+$ association is slightly 
%too strong also in these  models. The stonger Na$^+$ ion association in GROMOS based force fields compared to CHARMM based force 
%has been already reported in the literature~\cite{berkowitz06,cordomi09,valley11,berkowitz12,sachs04,valley11},
%however the novelty of this work in this respect is that we make a direct comparison
%to the experiments which can be used as measure for the model quality.

The origin for suggested inaccuracies in lipid-ion interactions in simulation models us unknown. 
In principle, the incorrect choline structure~\cite{botan15}, lack of polarizability~\cite{leontyev11} or
the used water model could cause such results. The effect of changes in lipid and ion model ot the ion
partition is discussed in supplementary material.
\todo{In the Orange simulation only lipid model is changed, respect to Berger, and Jukka tested the effect
of 0.7 charge scaling for Na ion (suggested by Leontyev et al.~\cite{leontyev11} to compensate to the lack 
of electronic polarizability in the model). I think we should discuss these things in supplementary material.
Even though we cannot be fully conclusive, there is some essential information also in these results.}

%From the technical point of view it is interesting to note that the same ion model is used in the Berger and Orange simulation, only difference 
%being in the lipid model, indicating that the partitioning can be strongly reduced just by modifying the lipid
%model parameters.

%In situations where a positive charge penetrates into a bilayer (e.g. cationic surfactant~\cite{scherer89} or multivalent ion~\cite{akutsu81}) 
%a decrease in choline order parameters is observed in experiments, as also seen in CaCl$_2$ results in Fig.~\ref{ordPions}.
%Also in all simulations analyzed in this work where cation is penetrating into a bilayer, the choline order parameters
%decrease as seen in Fig.~\ref{ordPions}. However, in the case of CaCl$_2$ the decrease in too pronounced in 
%simulations which may arise from too strong partitioning of Ca$^{2+}$ or overestimated influence of charge on lipid conformation.

\section{Conclusions}
We have compared phosholipid bilayer interactions with NaCl and CaCl$_2$ between different molecular dynamics simulation
models and $^2$H NMR experiments. The comparison led to the following conclusions

- The most straighforward explanation for the various experimental obsevations is that there is no Na$^+$ ion binding
into the phospholipid bilayer with mM concentrations, in constrast to Ca$^{2+}$ which specifically binds.

- The Na$^+$ is overestimated in almost all molecular dynamics simulation models, however, from the publicly available
models the CHARMM36 has the most realistic description.

- The electrometer concept suggesting connection between $\alpha$ and $\beta$ order parameter decrease and
cation partitioning~\cite{akutsu81,altenbach84,seelig87,scherer89} works also in simulations, despite of inaccuracies in actual atomistic resolution structures.

%Our discussion gives further support to the idea that choline orientation and ion partitioning can
%be experimentally measured by using the choline group order parameters as suggested in the series of
%publications by Seelig and co workers.
%However, we have updated the details of the relation by including the information about 
%the sings.

- \todo{Final conclusions about the structural response to be written once we have all the results}

%In conclusion, the penetrating induced decrease in order parameters can be seen in all models
%independently of the detailed molecular structure in the model, corresponding to the dehydration case
%where increase was seen in all models. The most likely explanation is that, despite of the structural model, the choline group
%orients more parallel to the membrane normal with penetrating cations. In contrast to the dehydration results,
%we are not aware of experimental data for the glycerol order parameters as a function of penetrating ions. 
%Furher, our results show that the Na$^+$ partitioning into a PC bilayer with Berger model is a
%simulation artefact, thus the research lines based these results should be taken with caution.

%This work has been done, and will be progressed, as an open collaboration through nmrlipids.blogspot.fi.

%Similar significant structural changes in the headgroup region are seen upon the addition
%of sodium and calcium ions in the Berger model. This contradicts with the experiments where
%sodium does not change the headgroup structure and the changes observed for calcium are different from those seen in simulations.
%These results indicate that the changes in the headgroup structure due to a permeating charge are not correct
%in the Berger model and, in particular, that sodium ions penetrating the headgroup region is erroneous.

This work has been, and will be further, progressed and discussed through the blog: nmrlipids.blogspot.fi. 
Everyone is invited to join the discussion and make contributions through the blog. 
The manuscript will be eventually submitted to an appropriate scientific journal. 
Everyone who has contributed to the work through the blog will be offered 
coauthorship. For more details see: nmrlipids.blogspot.fi.

{\bf Acknowledgements: }
OHSO acknowledges Tiago Ferreira for very useful discussions, the Emil Aaltonen foundation for financial support, Aalto Science-IT project and CSC-IT Center for Science for computational resources.

\bibliographystyle{apsrev}
\bibliography{refs}

\newpage

\appendix
\begin{center}
{\bf SUPPLEMENTARY INFORMATION}
\end{center}

\section{methods}

\subsection{Simulated systems}
All simulations are ran with a standard setup for planar lipid bilayer in zero tension
with periodic boundary conditions with Gromacs software package (version numbers 4.5-X-4.6.X).
%The number of molecules, temperatures and the length of simulations for all the fully 
%hydrated lipid bilayer simulations are listed in Table~\ref{systems}. Full simulation
%details for all simulations are given the next section.

%For intitial configuration of the simulations with NaCl and CaCl$_2$, the required number of randomly placed water molecules were replaced by 
%ions in fully hydrated simulations. 
%The molar concentration was calculated as ??. 
%The number of different molecules, temperatures and 
%the length of simulations for all the ion containing simulations are show in Table~\ref{IONsystems}.

%the POPC bilayer was solvated with 7202 water molecules, 44 Na$^+$ and Cl$^-$ ions,
%corresponding roughly to [NaCl] $\approx$ 200~mM. For simulations with CaCl$_2$, the POPC bilayer was solvated with 
%7157 water molecules, 44 Ca$^{2+}$ and 88 Cl$^-$ ions, corresponding roughly to [CaCl$_2$] $\approx$ 200~mM. 
%The order parameters were calculated from the last 50~ns of a trajectory
%totalling 110~ns.

\subsection{Simulation details}
\subsubsection{Berger}
The simulation without ions is the same as in~\cite{botan15}. The starting structures for simulations with ions is
made by replacing water molecules with appropriate amount of ions under study.
\todo{Samuli, finalize and check the methods.}

The Berger force field was used for the POPC~\cite{berger97}, with the dihedral potential next to the double bond 
taken from~\cite{bachar04}. The simulations are identical to previous publications~\cite{ollila07a,ferreira13,ferreira15}.
Timestep of 2~fs was used with leap-frog integrator. Covalent bond lengths were constrained with LINCS algorithm~\cite{hess97,hess07}. 
Coordinated were written every 10~ps. PME with real space cut-off 1.0~nm was used 
for electrostatics. Plain cut-off was used for the Lennart-Jones interactions with a 1.0~nm cut-off.
The neighbour list was updated every 5th step with cut-off 1.0~nm. Temperature was coupled separately
for lipids and water to 298~K with the velocity-rescale method~\cite{bussi07} with coupling constant 0.1~ps$^{-1}$.
Pressure was semi-isotropically coupled to the athmospheric pressure with the Berendsen method~\cite{berendsen84}.

\subsubsection{CHARMM36}
The simulation without ions is the same as in~\cite{botan15}. The starting structures for simulations with ions is
made by replacing water molecules with appropriate amount of ions under study.
\todo{Samuli, finalize and check the methods.}

Timestep of 1~fs was used with leap-frog integrator. Covalent bonds with hydrogens were constrained with LINCS algorithm~\cite{hess97,hess07}. 
Coordinated were written every 5~ps. PME with real space cut-off 1.4~nm was used 
for electrostatics. Lennart-Jones interactions were switched to zero between 0.8~nm and 1.2~nm.
The neighbour list was updated every 5th step with cut-off 1.4~nm. Temperature was coupled separately
for lipids and water to 303~K with the velocity-rescale method~\cite{bussi07} with coupling constant 0.2~ps.
Pressure was semi-isotropically coupled to the athmospheric pressure with the Berendsen method~\cite{berendsen84}.

\subsubsection{MacRog}
\todo{Matti Javanainen, please deliver the details.}

\subsubsection{Orange}
\todo{Jukka Maatta and Luca Monticelli, please deliver as much details as you can.}

\subsubsection{Slipids}
The simulation without ions is the same as in~\cite{botan15}.
\todo{Andrea Catte, please deliver the details.}

\subsection{analysis}
The order parameters were calculated from simulation trajectories directly applying the equation
$S_{{\rm CH}}=\frac{3}{2} \langle \cos^2 \theta-1 \rangle$. For united atom models the hydrogen locations
were regenerated for each molecule in each frame after the simulation trajectory was created.
??The statistical error estimate for each order parameter calculated from simulation was roughly
0.01, which is much smaller than the differences discussed in this work.??

%The order parameters calculated from previously published quadrupolar splitting data were calculated using 
%the equation $|S_{{\rm CD}}|=0.00784 \times \Delta \nu_Q$, where $\Delta \nu_Q$ is the quadrupolar splitting measured
%by $^2$H NMR (for more details see~\cite{seelig77c} and supplementary information).
%The order parameters measured with $^{13}$C NMR were calculated from dipolar splitting by using equation
%$|S_{{\rm CH}}|=\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c} d_\mathrm{CH}$, where
%$d_\mathrm{CH}$ is the measured dipolar splitting and
%values between 20.2-22.7 kHz were used for $\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c}$,
%depending on the original authors~\cite{hong95a,gross97,dvinskikh05,ferreira13} (for more discussion see supplementary material).

%The results for choline order parameters as a function of CaCl$_2$ from simulations and experiments
%in Fig.~\ref{ordPions} shows a significant decrease in both cases, however the decrease is more pronounced
%in simulations, indicating too high Ca$^{2+}$ partitioning of too strong effect to the order parameters
%in simulations. Also the decrease in order parameters in simulations with penetrating Na$^+$ ions (MacRog and Berger)
%is in qualitative agreement with experiments where penetrating positive ions are studied, Ca$^{2+}$ by Akutsu et al~\cite{akutsu81} also
%shown in Fig.~\ref{ordPions} and anionic surfactants by Scherer et al.~\cite{scherer89}.

%By comparing Figs~\ref{ordPnacl},~\ref{ordPions} and ~\ref{IONdensCOMP} it becomes clear that Na$^+$ binding is stronger (Fig. 4) in the force fields in which the 
%Na$^+$-induced changes in the order parameters are larger (Figs.~\ref{ordPnacl} and~\ref{ordPions}), i.e., in Berger and MacRog. In CHARMM36 and Orange, in which 
%the Na$^+$-induced changes in order parameters are smaller, also the partitioning of Na$^+$ is significantly less. It seems that the correlation between ion binding and order parameter 
%changes that was suggested from the NMR experiments~\cite{akutsu81,seelig87} can be reproduced in MD simulations.

%In simulations the ion induced changes in the order parameters shown in Figs.~\ref{ordPnacl} and ~\ref{ordPions} depends
%on the used model: NaCl induces a dramatic decrease of the $\beta$- and $\alpha$-carbon order parameters in Berger and MacRog models
%while in CHARMM36 and Orange force fields only small or nonexistent changes are induced.
%The order parameter results are in with the ion distribution results in Fig.~\ref{IONdensCOMP}
%showing that signifanct fraction of Na$^+$ ions is located in the bilayer region
%for the MacRog and Berger models, while the ion penetration is much less for the
%Orange and CHARMM36.

%The order parameters calculated from simulations with the Berger and MacRog force fields 
%for $\beta$ and $\alpha$ were clearly decreased by the presence of NaCl. Also order parameters
%for g$_3$ and g$_2$ were affected. The decrease of  $\beta$ and $\alpha$ order parameters
%is the signature of penetrating positive charges in the experiments, and indeed, the ion density
%distrubutions show that for both force fields a significant amount of Na$^+$ is located
%in the bilayer region. Slight decrease in  $\beta$ and $\alpha$ order parameters with 
%NaCl is also observed in CHARMM, and even less in the Orange. This observation is also
%in line with the density distributions of ions which show that CHARMM has some bound ions, Orange has less,
%and both have much less compared to the Berger and MacRog models.

%The simulation results
%are in line with previous works~\cite{??}. As a conclusion, the results strongly indicate that the
%partitioning of Na$^+$ is strongly overestimated in the Berger and MacRog models, and also significantly
%overestimated in the CHARMM model. There seems to be weak partitioning present also in the Orange model,
%however to conclude if this is too strong or not more data should be generated. This is not done since the

\onecolumngrid
\listoftodos

\end{document}

### 2015-02-06 LIPIDionINTERACT/calcCHANGES.sh

@@ -1,0 +1,12 @@
+cat BERGERnaclCONS.dat | awk '{if(NR==1){a=$2;b=$3;c=$4;d=$5;e=$6;f=$7;g=$8;h=$9;}print $1" " $2-a" "$3-b" "$4-c" "$5-d" "$6-e" "$7-f" "$8-g" "$9-h}' > BERGERnaclCONSchange.dat
+cat CHARMMnaclCONS.dat | awk '{if(NR==1){a=$2;b=$3;c=$4;d=$5;e=$6;f=$7;g=$8;h=$9;}print $1" " $2-a" "$3-b" "$4-c" "$5-d" "$6-e" "$7-f" "$8-g" "$9-h}' > CHARMMnaclCONSchange.dat
+cat ORANGEnaclCONS.dat | awk '{if(NR==1){a=$2;b=$3;c=$4;d=$5;e=$6;f=$7;g=$8;h=$9;}print $1" " $2-a" "$3-b" "$4-c" "$5-d" "$6-e" "$7-f" "$8-g" "$9-h}' > ORANGEnaclCONSchange.dat
+cat MACROGnaclCONS.dat | awk '{if(NR==1){a=$2;b=$3;c=$4;d=$5;e=$6;f=$7;g=$8;h=$9;}print $1" " $2-a" "$3-b" "$4-c" "$5-d" "$6-e" "$7-f" "$8-g" "$9-h}' > MACROGnaclCONSchange.dat
+cat SLIPIDSnaclCONSdppc.dat | awk '{if(NR==1){a=$2;b=$3;c=$4;d=$5;e=$6;f=$7;g=$8;h=$9;}print $1" " $2-a" "$3-b" "$4-c" "$5-d" "$6-e" "$7-f" "$8-g" "$9-h}' > SLIPIDSnaclCONSchangeDPPC.dat
+cat SLIPIDSnaclCONSpopc.dat | awk '{if(NR==1){a=$2;b=$3;c=$4;d=$5;e=$6;f=$7;g=$8;h=$9;}print $1" " $2-a" "$3-b" "$4-c" "$5-d" "$6-e" "$7-f" "$8-g" "$9-h}' > SLIPIDSnaclCONSchangePOPC.dat
+
[…]

### 2015-02-09 LIPIDionINTERACT/LIPIDionINTERACT.tex
update lipid-ion interaction manuscript

@@
 
-%\author{Patrick Fuchs}
-%\author{Matti Javanainen}
-%\author{Antti Lamberg}
-%\author{Alexander Lyubartsev}
-%\author{Markus S. Miettinen}
-%\author{Luca Monticelli}
-%\author{Jukka Määttä}
+\author{Andrea Catte}
+\thanks{University of East Anglia, Norwich, United Kingdom}
+\author{Matti Javanainen}
+\thanks{Tampere University of Technology, Tampere, Finland}
+\author{Markus S. Miettinen}
+\thanks{Fachbereich Physik, Freie Universit\"at Berlin, Berlin, Germany}
 %\setcounter{page}{1}
 \author{O. H. Samuli Ollila} 
-\footnote{Author to whom 
-correspondence may be addressed. E-mail: samuli.ollila@aalto.fi.\\
-}
-\affiliation{
-  Helsinki Biophysics and Biomembrane Group,
-Department of Biomedical Engineering and Computational Science, Aalto University, Espoo, Finland
-}
-%\author{Marius Retegan}
-%\author{Hubert Santuz}
-%\author{Joona Tynkkynen}
-%\author{Alexander Vogel}
+\thanks{{\bf Author to whom correspondence may be addressed. E-mail: samuli.ollila@aalto.fi.}}
+\thanks{Helsinki Biophysics and Biomembrane Group, Department of Biomedical Engineering and Computational Science, Aalto University, Espoo, Finland}
+%\footnote{Author to whom 
+%correspondence may be addressed. E-mail: samuli.ollila@aalto.fi.\\
+%}
+%\affiliation{
+%  Helsinki Biophysics and Biomembrane Group,
+%Department of Biomedical Engineering and Computational Science, Aalto University, Espoo, Finland
+%}
 
@@
 Slipid\cite{jambeck12}   &   DPPC & 0 & 128 &3840 & 0 & 0  & 0 & 323 & 150 & 100 &~\cite{slipidsFILES}  \\
-Slipid\cite{jambeck12}, ionFF~\cite{??}\todoi{Andrea Catte, please check the numbers and let us know if you share some files through Zenodo}    &   DPPC & 150 (NaCl) & ?
-& ?
-& ?
-& ?
-& ?
-& 323 & ?
-& ?
-&?
-\\
+Slipid\cite{jambeck12}, ionFF~\cite{??}\todoi{Andrea Catte, please let us know if you share some files through Zenodo}    &   DPPC & 150 (NaCl) & 600 & 18000 & 49 & 0  & 49 & 323 & 100 & 40 &?  \\
 \end{tabular}
@@
 The simulation without ions is the same as in~\cite{botan15}.
-\todo{Andrea Catte, please deliver the details.}
+\todo{DONE}
+
+For the simulations with ions, the starting DPPC lipid bilayer, which was built with the online CHARMM-GUI
+(http://www.charmm-gui.org/), contained 600 lipids, 30 water molecules/lipid, Na+ and Cl- ions (150 mM NaCl). 
+The TIP3p water model was used to solvate the system. AA MD simulations of DPPC lipid bilayers were performed 
+at ten different temperatures (283, 298, 303, 308, 312, 313, 314, 318, 323, and 333 K) using the GROMACS software 
+package version 4.5.5 and the Stockholm lipids (Slipids) force field parameters for phospholipids. After energy 
+minimization and a short equilibration run of 50 ps (time step 1fs), 100 ns production runs were performed using 
+a time step of 2 fs with leap-frog integrator. All covalent bonds were constrained with the LINCS
+algorithm. Coordinates were written every 100 ps. PME with real space cut-off at 1.0 nm was used for Coulomb 
+interactions. Lennard-Jones (LJ) interactions were switched to zero between 1.0 nm and 1.4 nm. The neighbour 
+lists were updated every 10th step with a cut-off of 1.6 nm. Temperature was coupled separately for upper and 
+bottom leaflets of the lipid bilayer, and for water to one of the temperatures reported above with the Nosè-Hoover 
+thermostat using a time constant of 0.5 ps. Pressure was semi-isotropically coupled to the atmospheric pressure 
+with the Parrinello-Rahman barostat using a time constant of 10 ps.
+The last 40 ns of each simulation was employed for the analysis of DPPC choline and glycerol backbone order parameters.
 

### 2015-02-12 LIPIDionINTERACT/LIPIDionINTERACT.tex
update ion-lipid interactions manuscript

@@
 \author{Andrea Catte}
+\thanks{The authors are listed in alphaphetical order.}
+\thanks{The author list is not completed.}
 \thanks{University of East Anglia, Norwich, United Kingdom}
@@
 %\setcounter{page}{1}
+\author{Vasily S. Oganesyan}
+\thanks{University of East Anglia, Norwich, United Kingdom}
 \author{O. H. Samuli Ollila} 
@@
 Berger-POPC-07\cite{ollila07a}, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}   &   POPC & 190 (CaCl$_2$) & 128 & 7157 & 0 & 44  & 88 &298 & 110 & 50 &?\todoi{Samuli put to Zenodo}  \\
+\hline
 CHARMM36\cite{klauda10}   & POPC & 0           & 72 & 2242 & 0  & 0 & 0 & 303  & 30 & 20 & \cite{charmm36filesSHORT} \\
@@
 CHARMM36\cite{klauda10}, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}   & POPC & 85  (NaCl)  & 72 & 2085 & 78  & 0 & 78 & 303  & 73 & 60 &?\todoi{Samuli put to Zenodo} \\
-MacRog\cite{maciejewski14}\todoi{Is this the same as in the first manuscript for full hydration?}  & POPC & 0 & ? & ? & 0 & 0 & 0 & ? & ?&?  &?  \\
-MacRog\cite{maciejewski14}\todoi{Matti Javanainen, please check the numbers and let us know if you share some files through Zenodo} , ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}  & POPC & ?
-(NaCl) & ?
-& ?
-& ?
-& ?
-& ?
-& ?
-& ?&?
-&?
-\\
-MacRog\cite{maciejewski14}\todoi{Matti Javanainen, please check the numbers and let us know if you share some files through Zenodo} , ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}        & POPC &  ?
-(NaCl) & ?
-& ?
-& ?
-& ?
-& ?
-& ?
-& ?&?
-&?
-\\
-MacRog\cite{maciejewski14}\todoi{Matti Javanainen, please check the numbers and let us know if you share some files through Zenodo} , ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}        & POPC &   ?
-(NaCl) & ?
-& ?
-& ?
-& ?
-& ?
-& ?
-& ?&?
-&?
-\\
-MacRog\cite{maciejewski14}\todoi{Matti Javanainen, please check the numbers and let us know if you share some files through Zenodo} , ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}        & POPC &   ?
-(NaCl) & ?
-& ?
-& ?
-& ?
-& ?
-& ?
-& ?&?
-&?
-\\
+\hline
+MacRog\cite{maciejewski14}\todoi{DONE}  & POPC & 0 & 288 & 14400 & 0 & 0 & 0 & 310 & 90&40  &~\cite{macrogdehydFILES}  \\
+MacRog\cite{maciejewski14}\todoi{DONE} , ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}  & POPC & 100 (NaCl) & 288 & 14554 & 27 & 0 & 27 & 310 & 90&50  & \cite{macrogIONfiles} \\
+MacRog\cite{maciejewski14}\todoi{DONE} , ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}        & POPC &  200 (NaCl) & 288 & 14500 & 54 & 0 & 54 & 310 & 90&50  &\cite{macrogIONfiles}  \\
+MacRog\cite{maciejewski14}\todoi{DONE} , ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}        & POPC &   300 (NaCl) & 288 & 14446 & 81 & 0 & 81 & 310 & 90&50  & \cite{macrogIONfiles} \\
+MacRog\cite{maciejewski14}\todoi{DONE} , ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}        & POPC &   400 (NaCl) & 288 & 14392 & 108 & 0 & 108 & 310 & 90& 50  & \cite{macrogIONfiles}  \\
+\hline
 Orange, ionFF~\cite{??}\todoi{Samuli check the numbers}   &   POPC & 0 & ?
@@
 This is unpublished model.} \\
+\hline
 Slipid\cite{jambeck12}   &   DPPC & 0 & 128 &3840 & 0 & 0  & 0 & 323 & 150 & 100 &~\cite{slipidsFILES}  \\
@@
 \subsubsection{MacRog}
-\todo{Matti Javanainen, please deliver the details.}
+\todo{DONE}
+The simulation parameters are identical to those employed in our earlier study~\cite{botan15} for the full 
+hydration and dehydration simulations. The initial structures with varying amounts of NaCl were constructed from an 
+extensively hydrated bilayer by replacing water molecules with ions using the gromacs tool genion. Even at the highest 
+considered salt concentration, the amount of water molecules per lipid after this replacement process was still greater than 50.
 

### 2015-02-16 HGmodelMANUSCRIPT/HGmodel_draft3.tex
updating information from blog to the manuscript
updates to HGmodel draft

@@
 %
-\author{Andrea Catte}
-\thanks{University of East Anglia, Norwich, United Kingdom}
+%\author{Andrea Catte}
+%\thanks{University of East Anglia, Norwich, United Kingdom}
 \author{Fernando Favela}
@@
 %\setcounter{page}{1}
-\author{Vasily S. Oganesyan}
-\thanks{University of East Anglia, Norwich, United Kingdom}
+%\author{Vasily S. Oganesyan}
+%\thanks{University of East Anglia, Norwich, United Kingdom}
 \author{O. H. Samuli Ollila} 
@@
 \thanks{Max Planck Institute for Chemical Energy Conversion, M\"ulheim an der Ruhr, Germany}
+\author{Tomasz Rog}
+\thanks{Tampere University of Technology, Tampere, Finland}
 \author{Hubert Santuz}
@@
 Phospholipids are essential building blocks of biological membranes.
-Despite of vast amount of accurate experimental data the atomistic resolution structures sampled by the glycerol backbone and choline headgroup
-in phoshatidylcholine bilayers are not known. Atomistic resolution molecular dynamics simulation model 
+Despite of vast amount of accurate experimental data, the atomistic resolution structures sampled by the glycerol backbone and choline headgroup
+in phoshatidylcholine bilayers are not known. Atomistic resolution molecular dynamics simulations model 
 would automatically resolve the structures giving an interpretation of experimental results, if the model
-reproduced the experimental data. In this work we compare the C--H bond vector order
-parameters for glycerol backbone and choline headgroup between 14 different atomistic resolution
-models and experiments in a fully hydrated lipid bilayer. The current models are not accurately enough to resolve the structure.
+reproduced the experimental data. In the present work we simulate phosphatidylcholine (PC) lipid bilayers with 13 different
+atomistic models, and we compare simulations with experiments (in fully hydrated conditions) in terms of C--H bond vector order
+parameters for the glycerol backbone and choline headgroups. None of the current models is not accurately enough to resolve the structure with experimental accuracy.
 However, closer inspection of three best performing models (CHARMM36, GAFFlipid and MacRog) suggest
@@
 structures between the most often used (Berger based) lipid model~\cite{berger97} 
-\todo{Based on the results and discussions with Olle Edholm and Peter Tieleman in the blog, it seems that 
-the glycerol backbone dihedrals are not the same in the model used in the publication by Berger et al.~\cite{berger97}
-and in the model shared in the webpage of Peter Tieleman (http://wcm.ucalgary.ca/tieleman/downloads).
-In this work (as it is quite often done in other works as well) we currently refer to the model in the webpage as ''Berger'' model.
-This is misleading. We have to call it something else. One suggestion could be to call it ''BergerM'', because it was firt time
-used in Marrink et al. Biophys. J. 74,931-943 (1998). Any other suggestions?\\
-COMMENT: Patrick Fuchs suggested ''Berger-T''. This is better. \\
-COMMENT2: Current trial, based in blog discussion, is to use Berger-MOLECULE-YEAR [citation].
+\todo{If there are no objections, we will leave the naming convention of different Berger versions as it is now.
+%Based on the results and discussions with Olle Edholm and Peter Tieleman in the blog, it seems that 
+%the glycerol backbone dihedrals are not the same in the model used in the publication by Berger et al.~\cite{berger97}
+%and in the model shared in the webpage of Peter Tieleman (http://wcm.ucalgary.ca/tieleman/downloads).
+%In this work (as it is quite often done in other works as well) we currently refer to the model in the webpage as ''Berger'' model.
+%This is misleading. We have to call it something else. One suggestion could be to call it ''BergerM'', because it was firt time
+%used in Marrink et al. Biophys. J. 74,931-943 (1998). Any other suggestions?\\
+%COMMENT: Patrick Fuchs suggested ''Berger-T''. This is better. \\
+%COMMENT2: Current trial, based in blog discussion, is to use Berger-MOLECULE-YEAR [citation].
 }
@@
 
-\subsection{Analysis}
-\todo{ We should mention the sign issue and refer to the blog post:
-http://nmrlipids.blogspot.fi/2014/04/on-signs-of-order-parameters.html somewhere.
-Maybe we should change the name of this section II.B. to: "Analysis and terminology of order parameters from experiments 
-and simulations" or similar, and also move the "forking" explanation here.}
-The order parameter of a hydrocarbon C--H vector is defined as $S_{{\rm CH}}=\frac{3}{2}\langle \cos^2 \theta-1 \rangle$, where
-the angle brackets denote an ensemble average over the sampled conformations, and $\theta$ is the angle between the C--H bond and the membrane normal.
-The order parameters can be measured by detecting quadrupolar splitting with $^2$H NMR~\cite{seelig77c} or by detecting dipolar 
+\subsection{Order parameters from experiments}
+\todo{DONE
+%We should mention the sign issue and refer to the blog post:
+%http://nmrlipids.blogspot.fi/2014/04/on-signs-of-order-parameters.html somewhere.
+%Maybe we should change the name of this section II.B. to: "Analysis and terminology of order parameters from experiments 
+%and simulations" or similar, and also move the "forking" explanation here.
+}
+The order parameter of a hydrocarbon C--H vector is defined as 
+\begin{equation}\label{orderP}
+S_{{\rm CH}}=\frac{3}{2}\langle \cos^2 \theta-1 \rangle,
+\end{equation} 
+where the angle brackets denote an ensemble average over the sampled conformations, and $\theta$ is the angle between the C--H bond and the membrane normal.
+The absolute values of order parameters can be measured by detecting quadrupolar splitting with $^2$H NMR~\cite{seelig77c} or by detecting dipolar 
 splitting with $^{13}$C NMR~\cite{hong95a,gross97,dvinskikh05a,ferreira13}. The measurements are based on
 different physical interactions and also the connection between order parameters and quadrupolar or dipolar splitting
-are different. The order parameters from the measured quadrupolar splitting $\Delta \nu_Q$ ($^2$H NMR) are calculated using 
+are different. The absolute values of order parameters from the measured quadrupolar splitting $\Delta \nu_Q$ ($^2$H NMR) are calculated using 
 the equation $|S_{{\rm CD}}|=\frac{4}{3} \frac{e^2qQ}{h} \Delta \nu_Q$, where the value for the static quadrupole
 splitting constant is estimated from various experiments to be 170~kHz leading to a numerical relation $|S_{{\rm CD}}|=0.00784 \times \Delta \nu_Q$~\cite{seelig77c}. 
-The order parameters from the measured dipolar splitting $d_\mathrm{CH}$ ($^{13}$C NMR) are calculated using equation
+The absolute values of order parameters from the measured dipolar splitting $d_\mathrm{CH}$ ($^{13}$C NMR) are calculated using equation
 $|S_{{\rm CH}}|=\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c} d_\mathrm{CH}$, where
@@
 \todo{This is temporarily linked to the blog. Currently I think that the posts (with comments included) to which we want to cite are put into 
-the Figshare in pdf format as they are in when the publication is submitted. Figshare allows commenting to continue the discussion.}.
-
-The order parameters from simulations were calculated directly using the definition.
+the Figshare in pdf format as they are in when the publication is submitted. Figshare allows commenting to continue the discussion.
+COMMENT: There is also the option to put the most relevant blog posts as a supplementary material. The advantage of this would
+be that the posts would be peer reviewed as well.}.
+
+{\bf The absolute values of order parameters are accessible with both $^2$H NMR and $^{13}$C NMR techniques. 
+However, only $^{13}$C NMR techniques allow also the measurement of the sign of the order parameter~\cite{hong95a,hong95b,gross97}. 
+The measured sign is negative for almost all the carbons discussed in this work as only $\alpha$ is positive~\cite{hong95a,hong95b,gross97}. 
+For more detailed discussion about the sign measurements of the order parameters see \url{http://nmrlipids.blogspot.fi/2014/04/on-signs-of-order-parameters.html}
+
+For most CH$_2$ segments in fluid phosholipid bilayer the order parameters are equal for both hydrogens attached to the same carbon.
+However, in some cases (e.g. g$_1$, g$_3$ and  C$_2$ carbon in the \textit{sn}-2 chain) the order parameters are not equal and this 
+can be observed with both $^2$H NMR and $^{13}$C NMR techniques. In the present work we call the phenomena of unequal order parameters 
+for hydrogens attached to the same carbon as {\it forking} to avoid confusion with dipolar and quadrupolar splitting in NMR terminology. 
+Forking has been studied in detail with $^2$H NMR techniques by deuterating the R or S position in CH$_2$ segment, and
+the studies show that the forking arises from differently sampled orientations of the two C--H bonds, not from two 
+separate populations of lipid conformations~\cite{engel81,gally81}.
+}
+
+\subsection{Order parameters from simulations}
+The order parameters from simulations were calculated directly using the definition as in Eq.~\ref{orderP}.
 %$S_{{\rm CH}}=\frac{3}{2} \langle \cos^2 \theta-1 \rangle$. 
-For the united atom models the hydrogen locations were generated 
+For the united atom models the hydrogen positions were generated 
 in the trajectories post-simulationally using the positions of the heavy atoms and the known hydrocarbon geometries.
-The statistical errors were estimated for the best performing simulation
-models by calculating the error of the mean for the average over individual lipids in the system.
+{\bf For the statistical error estimates, the time average of order parameters were first calculated separately
+for each lipid in the system. Then it was assumed that different lipids are statistically independent 
+entities (which should be the case in fluid phase) and the error of the mean for the average over individual 
+lipids in the system was calculated and used as error bar for the order parameters.}
 
@@
 
-\begin{table*}[htb]
-\centering
-\caption{Simulated single component fully hydrated lipid bilayers. The simulation file data sets marked with $^*$ also include a part of the trajectory.
+\begin{table*}[]
+\centering
+\caption{Simulated single component fully hydrated lipid bilayer systems. The simulation file data sets marked with $^*$ also include a part of the trajectory.
   If simulation data from previously published work has been directly used, the original publication is cited for simulation details. For other systems the simulation details
   are given in the Supplementary Information.
+The abbreviations are 1-palmitoyl-2-oleoylphosphatidylcholine (POPC), dipalmitoylphosphatidylcholine (DPPC), 1,2-dimyristoyl-sn-glycero-3-phosphocholine (DMPC),
+1,2-dioleoyl-sn-glycero-3-phosphocholine (DOPC) and dilauroylphosphatidylcholine (DLPC).
 }\label{systems}
@@
 Berger-DMPC-04~\cite{gurtovenko04}          &   DMPC & 128 & 5097  & 323  & 130 & 100  & \cite{dmpcFILES} & \cite{miettinen09} \\
-CHARMM36\cite{klauda10}       & POPC   & 72  &  2242 & 303 & 30 & 20  & \cite{charmm36filesSHORT}$^*$ & SI \\
-CHARMM36\cite{klauda10}      & POPC   & 128 &  5120    & 303 & 150 & 100  & \cite{charmm36files}$^*$ \todoi{DONE}   & SI \\
-CHARMM36\cite{klauda10}       & DPPC   & 72  &  2189 & 323 & 30 & 25  & ?\todoi{Samuli, put to Zenodo}  & SI \\
+CHARMM36~\cite{klauda10}       & POPC   & 72  &  2242 & 303 & 30 & 20  & \cite{charmm36filesSHORT}$^*$ & SI \\
+CHARMM36~\cite{klauda10}      & POPC   & 128 &  5120    & 303 & 150 & 100  & \cite{charmm36files}$^*$ \todoi{DONE}   & SI \\
+CHARMM36~\cite{klauda10}       & DPPC   & 72  &  2189 & 323 & 30 & 25  & ?\todoi{Samuli, put to Zenodo}  & SI \\
 %CHARMM36\cite{klauda10}\todoi{Not included in the Fig.
@@
 \todoi{DONE} \todoi{Permanent link for DOPC CHARMM36 simulation by Hubert Santuz is not required now. However, it would be probably useful in the future.}
-MacRog\cite{maciejewski14}  & POPC & 288  & 12600 & 310 & 100 & 80  & \cite{macrogFILES}$^*$ & SI  \\
-GAFFlipid\cite{dickson12}       & POPC & 126  & 3948  & 303 & 137 & 32  & \cite{GAFFlipidFILES}$^*$ & SI \\
-GAFFlipid\cite{dickson12}       & DPPC & 72  & 2197  & 323 & 90 & 50  & \cite{GAFFlipidFILESdppc}$^*$ & SI \\
-Lipid14\cite{dickson14}         & POPC  & 72 & 2234 & 303 & 100 & 50  & \cite{lipid14files}$^*$ & SI \\
-Poger\cite{poger10}             & DPPC  & 128 & 5841 & 323 & 2$\times$100 & 2$\times$50 & \cite{pogerFILESpme1,pogerFILESpme2}$^*$\todoi{DONE} & SI \\
-Slipid\cite{jambeck12}          & DPPC & 128 & 3840 & 323 & 150 & 100 & \cite{slipidsFILES}$^*$ & SI \\
-Slipid\cite{jambeck12}          & POPC & 128 & 5120 & 303 & 200 & 150 & \cite{slipidsFILESpopc}$^*$ & SI \\
-Kukol\cite{kukol09}          & POPC   & 512 & 20564 & 298 & 50 & 30  & \cite{kukolFILES}$^*$ & SI \\
-Chiu et al.\cite{chiu09} \todoi{Samuli, put to Zenodo}     & POPC  & 128 & 3552  & 298 & 56 & 50  & ? & SI \\
-H\"ogberg08\cite{rabinovich14}\todoi{Alexander Luybartsev, let us know if you share the files.}  & POPC   &  128 & 3840  & 303 & 100 & 80  & ? & \cite{rabinovich14}  \\
-H\"ogberg08\cite{hogberg08}\todoi{Alexander Luybartsev, let us know if you share the files.}  & DMPC   &  98 & 3840  & 303 & 75 & 50 & ? & \cite{hogberg08} \\
-Ulmschneiders\cite{Ulmschneider09}    & POPC  & 128 & 3328 & 310 & 100 & 50 & \cite{ulmschneiderFILES}$^*$ & SI \\
-Tj\"ornhammar et al.\cite{tjornhammar14}   & DPPC  & 144 & 7056 & 323 & 200 & 100 & \cite{tjornhammarfiles}$^*$ & \cite{tjornhammar14} \\
+MacRog~\cite{maciejewski14}  & POPC & 288  & 12600 & 310 & 100 & 80  & \cite{macrogFILES}$^*$ & SI  \\
+GAFFlipid~\cite{dickson12}       & POPC & 126  & 3948  & 303 & 137 & 32  & \cite{GAFFlipidFILES}$^*$ & SI \\
+GAFFlipid~\cite{dickson12}       & DPPC & 72  & 2197  & 323 & 90 & 50  & \cite{GAFFlipidFILESdppc}$^*$ & SI \\
+Lipid14 \cite{dickson14}         & POPC  & 72 & 2234 & 303 & 100 & 50  & \cite{lipid14files}$^*$ & SI \\
+Poger \cite{poger10}             & DPPC  & 128 & 5841 & 323 & 2$\times$100 & 2$\times$50 & \cite{pogerFILESpme1,pogerFILESpme2}$^*$\todoi{DONE} & SI \\
+Slipid \cite{jambeck12}          & DPPC & 128 & 3840 & 323 & 150 & 100 & \cite{slipidsFILES}$^*$ & SI \\
+Slipid \cite{jambeck12}          & POPC & 128 & 5120 & 303 & 200 & 150 & \cite{slipidsFILESpopc}$^*$ & SI \\
+Kukol \cite{kukol09}          & POPC   & 512 & 20564 & 298 & 50 & 30  & \cite{kukolFILES}$^*$ & SI \\
+Chiu \cite{chiu09} \todoi{Samuli, put to Zenodo}     & POPC  & 128 & 3552  & 298 & 56 & 50  & ? & SI \\
+H\"ogberg08 \cite{rabinovich14}\todoi{Alexander Luybartsev, let us know if you share the files.}  & POPC   &  128 & 3840  & 303 & 100 & 80  & ? & \cite{rabinovich14}  \\
+H\"ogberg08 \cite{hogberg08}\todoi{Alexander Luybartsev, let us know if you share the files.}  & DMPC   &  98 & 3840  & 303 & 75 & 50 & ? & \cite{hogberg08} \\
+Ulmschneiders \cite{Ulmschneider09}    & POPC  & 128 & 3328 & 310 & 100 & 50 & \cite{ulmschneiderFILES}$^*$ & SI \\
+Tj\"ornhammar14 \cite{tjornhammar14}   & DPPC  & 144 & 7056 & 323 & 200 & 100 & \cite{tjornhammarfiles}$^*$ & \cite{tjornhammar14} \\
 %\sout{OPLS-AA}\cite{rog09b}\todoi{We have this only with 150mM of NaCl delivered by Joona Tynkkynen.
@@
 \subsection{Full hydration: Experimental order parameters for the glycerol backbone and headgroup}\label{experiments}
-The specific deuteration of $\alpha$-, $\beta$- and g$_3$- segments of dipalmitoylphosphatidylcholine (DPPC) has been successful, 
-allowing the order parameter measurements for these segments by $^2$H--NMR~\cite{gally75,brown77,brown78,akutsu81}.
-In addition, the order parameters for all glycerol backbone and choline headgroup segments in egg yolk lecithin~\cite{hong95a},
-1,2-dimyristoyl-sn-glycero-3-phosphocholine (DMPC)~\cite{hong95b,gross97,dvinskikh05a}, 
-1,2-dioleoyl-sn-glycero-3-phosphocholine (DOPC)~\cite{warschawski05} and POPC~\cite{warschawski05,ferreira13}
-have been measured with several different implementations of $^{13}$C NMR experiments.
-The experimental absolute values of glycerol backbone and choline order parameters from various publications are shown in Fig.~\ref{HGorderparameters}.
+The specific deuteration of $\alpha$-, $\beta$- and g$_3$- segments of DPPC has been successful, 
+allowing the absolute value order parameter measurements for these segments by $^2$H--NMR~\cite{gally75,brown77,brown78,akutsu81}.
+In addition, the absolute values of order parameters for all glycerol backbone and choline headgroup segments in egg yolk lecithin~\cite{hong95a},
+DMPC~\cite{hong95b,gross97,dvinskikh05a}, DOPC~\cite{warschawski05} and POPC~\cite{warschawski05,ferreira13}
+have been measured with several different implementations of $^{13}$C NMR experiments. In addition, the signs of order parameters in some systems
+are measured with $^{13}$C NMR tehcniques~\cite{hong95a,hong95b,gross97}. The experimental absolute values of glycerol backbone 
+and choline order parameters from various publications are shown in Fig.~\ref{HGorderparameters}.
 \begin{figure}[]
@@
   \includegraphics[width=8.6cm]{../DATAreportediINblog/comparisonSorted.pdf}
-\todo{Markus Miettinen suggested that we should consider making one more figure where only experimental data would be shown and that would be discussed in Section~\ref{experiments}} \\
+\todo{Markus Miettinen suggested that we should consider making one more figure where only experimental data would be shown and that would be discussed in Section~\ref{experiments}. \\
+COMMENT: I do not think that we need such figure anymore since this figure is more clear now.} \\
 \todo{I think that this figure is now very good. There is also the interactive version by Hubert Santuz now in https://plot.ly/~HubertSantuz/72/lipid-force-field-comparison/
@@
 while the hydrogens in choline $\alpha$ and $\beta$ segments give equal values.
-Note that in this work we call the phenomena of unequal order parameters for hydrogens attached to the same carbon as ``forking'' to avoid 
-confusion with dipolar and quadrupolar splitting in NMR terminology. Forking is also observed experimentally for the C$_2$ carbon in the \textit{sn}-2
-chain of all phosholipids, and it is known to arise from differently sampled orientations of the two C--H bonds, not from two 
-separate populations of lipid conformations~\cite{engel81}. The forking in glycerol backbone g$_3$ segment is small ($\approx$ 0.02) 
+The forking in glycerol backbone g$_3$ segment is small ($\approx$ 0.02) 
 and some experiments only report the larger value or the average value~\cite{akutsu81,ferreira13}. 
@@
 holds also for the PC lipids.
-
-In Fig.~\ref{HGorderparameters} we have shown the absolute values of order parameters as these are accessible
-with both $^2$H NMR and $^{13}$C NMR techniques. However, $^{13}$C NMR techniques allow also the measurement of 
-the sign of the order parameter~\cite{hong95a,hong95b,gross97}. The measured sign is negative for almost all the carbons 
-discussed in this work as only $\alpha$ is positive~\cite{hong95a,hong95b,gross97}.
 

### 2015-02-16 LIPIDionINTERACT/LIPIDionINTERACT.tex
update lipid-ion manuscript

@@
    }
-\todo{I think that we have to run Orange simulation with higher NaCl concentration. I will do it.} \\
+\todo{The ion concentrations for, at least, Berger, CHARMM36 and Orange has to be recalculated, see http://nmrlipids.blogspot.com/2015/02/the-first-draft-of-ion-lipid.html?showComment=1423771449743\#c5884666173819273918 \\
+ I think that we have to run Orange simulation with higher NaCl concentration. I will do it.} \\
 \todo{I think that we have to run CHARMM36 with CaCl$_2$. If anyone happen to have the data and is willing to share,

### 2015-02-17 HGmodelMANUSCRIPT/HGmodel_draft3.tex
update HGmodel manuscript

@@
 
-\subsection{Order parameters from experiments}
+\subsection{Order parameters from experiments}\label{expORDp}
 \todo{DONE
@@
 Slipid \cite{jambeck12}          & DPPC & 128 & 3840 & 323 & 150 & 100 & \cite{slipidsFILES}$^*$ & SI \\
-Slipid \cite{jambeck12}          & POPC & 128 & 5120 & 303 & 200 & 150 & \cite{slipidsFILESpopc}$^*$ & SI \\
+Slipid \cite{jambeck12b}          & POPC & 128 & 5120 & 303 & 200 & 150 & \cite{slipidsFILESpopc}$^*$ & SI \\
 Kukol \cite{kukol09}          & POPC   & 512 & 20564 & 298 & 50 & 30  & \cite{kukolFILES}$^*$ & SI \\
@@
 \subsection{Full hydration: Experimental order parameters for the glycerol backbone and headgroup}\label{experiments}
-The specific deuteration of $\alpha$-, $\beta$- and g$_3$- segments of DPPC has been successful, 
+{\bf The specific deuteration of $\alpha$-, $\beta$- and g$_3$- segments of DPPC has been successful, 
 allowing the absolute value order parameter measurements for these segments by $^2$H--NMR~\cite{gally75,brown77,brown78,akutsu81}.
@@
 have been measured with several different implementations of $^{13}$C NMR experiments. In addition, the signs of order parameters in some systems
-are measured with $^{13}$C NMR tehcniques~\cite{hong95a,hong95b,gross97}. The experimental absolute values of glycerol backbone 
-and choline order parameters from various publications are shown in Fig.~\ref{HGorderparameters}.
+are measured with $^{13}$C NMR tehcniques~\cite{hong95a,hong95b,gross97}. The experimental values of glycerol backbone 
+and choline order parameters from various publications~\cite{gally75,akutsu81,gross97,dvinskikh05a,ferreira13} with the signs measured in~\cite{hong95a,hong95b,gross97} are shown in Fig.~\ref{HGorderparameters}.
+}
 \begin{figure}[]
@@
 
-In addition to the numerical values, an important feature of the glycerol backbone is the 
-inequality of order parameters for the two hydrogens attached to the same carbon in g$_1$ and g$_3$ segments,
-while the hydrogens in choline $\alpha$ and $\beta$ segments give equal values.
+{\bf In addition to the numerical values, an important feature of the glycerol backbone is the 
+forking (see section~\ref{expORDp}) of the order parameters in g$_1$ and g$_3$ segments, in contrast to the choline segments $\alpha$ and $\beta$. }
 The forking in glycerol backbone g$_3$ segment is small ($\approx$ 0.02) 
@@
 
-By combining the experimental information of the sign~\cite{hong95a,hong95b,gross97} and the stereospecific
-measurements~\cite{gally81} with the accurate absolute value measurements from various techniques~\cite{gally75,akutsu81,gross97,dvinskikh05a,ferreira13},
-the most detailed experimentally available order parameter information for the glycerol backbone and choline segments of POPC is obtained.
-These data are shown in Fig.~\ref{HGorderparameters2}.
+{\bf By combining the absolute value measurements~\cite{ferreira13}, the sign measurements~\cite{hong95a,hong95b,gross97} and the stereospecific
+measurements~\cite{gally81} the most detailed experimentally available order parameter information for the glycerol backbone and choline 
+segments of POPC bilayer is collected and the data are shown in Fig.~\ref{HGorderparameters2}. }
 \begin{figure*}[]
@@
 
-To get a general idea of the quality of the glycerol backbone and choline headgroup structures in different models, we calculated the absolute
-values of the order parameters for these parts from fourteen different lipid models (Table~\ref{systems}) and 
+To get a general idea of the quality of the glycerol backbone and choline headgroup structures in different models, we calculated 
+the order parameters for these parts from thirteen different lipid models (Table~\ref{systems}) and 
 plotted the results together with experimental values in Fig.~\ref{HGorderparameters}.
 Two criteria were used to judge the quality of the model: there must not be significant  {\bf forking} in the $\alpha$ and $\beta$ carbons,
-there must be only moderate forking in the g$_3$ carbon and there must be significant forking in the g$_1$ carbon, {\bf absolute values}
+there must be only moderate forking in the g$_3$ carbon and there must be significant forking in the g$_1$ carbon, the {\bf magnitude}
 should be preferably inside to the subjective sweet spots determined from experiments (blue shaded regions in Fig.~\ref{HGorderparameters}).
-None of the studied force fields fulfils these criteria.
-However, three force fields perform better than others, namely CHARMM36~\cite{klauda10}, MacRog~\cite{maciejewski14} and GAFFlipid~\cite{dickson12}.
 The results for each force field in respect to the above criteria are summarized in Figure~\ref{FullHydrationComparisonTable}.
-\todo{The table is now done, see Fig.~\ref{FullHydrationComparisonTable}. The text should be updated accordingly.}.
+\todo{DONE}.
 \begin{figure}[]
@@
   Ranking of force fields based on data of Fig.~\ref{HGorderparameters}. 
+  Markus, please describe the definition of M, F and $\Sigma$ here.
   } 
@@
 
-The top three models (CHARMM36, MacRog and GAFFlipid) together with the most used lipid model (Berger based model) 
-were subjected to a more careful comparison including the signs and the sterospecific labeling in Fig.~\ref{HGorderparameters2}.
-The essential additional information given by this comparison is that the sign of the $\beta$ carbon order parameter is correct only in CHARMM36 model.
+{\bf
+None of the studied force fields fulfils these criteria completely, however CHARMM36 is pretty close. 
+This is not surprising since the dihedral potentials in this model are tuned to reproduce these parameters better against experiments~\cite{klauda10}.
+The next models in the list are CHARMM36-UA~\cite{henin08,lee14} and H\"ogberg08~\cite{hogberg08} which is also not surprising since
+these models are using CHARMM bonded potentials for glycerol backbone and choline. The fourth and the fifth models in the list, MacRog~\cite{maciejewski14} and
+GAFFlipid~\cite{dickson12}, have independently determined dihedral potentials. All the models based on Gromos potentials and Slipid perform less well.
+In the present work we subject the CHARMM36, MacRog, GAFFlipid and Berger-POPC-07 to a more careful comparison including the sterospecific labeling  
+(Fig.~\ref{HGorderparameters2}), and atomistic level structure and responses to the dehydration and cholesterol content in the following sections.
+These models are selected for more detailed studies since they are the best representatives of different dihedral potential parametrization techniques 
+(CHARMM36, MacRog, GAFFlipid), and the Berger based models are the most used lipid model in the literature.
+}
+%The top three models (CHARMM36, MacRog and GAFFlipid) together with the most used lipid model (Berger based model) 
+%were subjected to a more careful comparison including the signs and the sterospecific labeling in Fig.~\ref{HGorderparameters2}.
+%The essential additional information given by this comparison is that the sign of the $\beta$ carbon order parameter is correct only in CHARMM36 model.
 
@@
 significant differences between the models are observed around these segments in Fig.~\ref{HGorderparameters2}.
-We note that performing a similar comparison through all the dihedrals in all the 14 models would probably give highly useful
+We note that performing a similar comparison through all the dihedrals in all the 13 models would probably give highly useful
 information on how to improve the accuracy of the models yet this is beyond the scope of the current report.
@@
 In conclusion, models preferring the anti conformation for this dihedral give more realistic order parameters and
-this in agreement with previous crystal structure and $^1$H NMR studies~\cite{hauser80,hauser81,hauser81b,hauser88,pascher92,marsh06}.
+this is in agreement with previous crystal structure and $^1$H NMR studies~\cite{hauser80,hauser81,hauser81b,hauser88,pascher92,marsh06}.
 \begin{figure}[]
@@
 \url{http://nmrlipids.blogspot.fi/2014/02/accuracy-of-order-parameter-measurements.html}
-for more discussion.
+for more discussion.)
 \todo{This is temporarily linked to the blog. Currently I think that the posts (with comments included) to which we want to cite are put into 
@@
 
-- None of the tested models (14 different models) produces the order parameters with the experimental
+- None of the tested models (13 different models) produces the order parameters with the experimental
 accuracy for a fully hydrated phoshatidylcholine lipid bilayer. However, the CHARMM36, GAFFlipid and MacRog 
@@
 
+{\bf It should be noted that the area per molecule with these settings for the GAFFlipid model was 61.6~\AA$^2$,
+while the original publication reported 63.9~\AA$^2$~\cite{dickson12}. However, the same parameters and Amber to Gromacs
+conversion reproduced the area per molecule from original publication for the lipid14 model (see next section).}
+
 \subsubsection{Lipid14}
@@
 
+{\bf The area per molecule with these settings was 65.4~\AA$^2$ which is in agreement with the value reported in the original publciation 65.6$\pm$0.5~\AA$^2$~\cite{dickson14}.}
+
 \subsubsection{Poger et al.}
-The Poger lipids are derived from GROMOS G53A6~\cite{poger10} and were initially coined 53A6-L (L for lipids).
-They are now part of GROMOS G54A7~\cite{poger12} and parametrized to work with the SPC water model~\cite{berendsen81}.
-The initial hydrated bilayer structure of 128 DPPC and 5841 water molecules as well as force field parameters were downloaded
+The Poger lipids are derived from GROMOS G53A6~\cite{poger10} and were initially coined 53A6-L (L for lipids). They are now part of GROMOS G54A7~\cite{poger12} 
+and parametrized to work with the SPC water model~\cite{berendsen81}. The initial hydrated bilayer structure of 128 DPPC and 5841 water molecules as well as force field parameters were downloaded 
 from David Poger's web site (\url{http://compbio.chemistry.uq.edu.au/~david/}) on April 2012. 
@@
 
-Initial coordinates for a hydrated DPPC (at 323~K) and POPC (at 310~K) bilayers (30 and 40 waters/lipid, respectively) were taken directly from the Slipids home page \url{http://people.su.se/\~jjm/Stockholm\_Lipids/Downloads.html}.
-The Slipids force field~\cite{jambeck12,XXX} ADD 2ND SLIPIDS PAPER!
-was used for the the all atom descriptions of DPPC and POPC, and
+Initial coordinates for a hydrated DPPC (at 323~K) and POPC (at 310~K) bilayers (30 and 40 waters/lipid, respectively) were taken directly from the 
+Slipids home page \url{http://people.su.se/\~jjm/Stockholm\_Lipids/Downloads.html}.
+The Slipids force field~\cite{jambeck12,jambeck12b} was used for the the all atom descriptions of DPPC and POPC, and
 water was described with the TIP3P water model~\cite{jorgensen83}. Simulations were performed within the NPT ensemble using the GROMACS 4.6.X simulation

### 2015-02-18 LIPIDionINTERACT/LIPIDionINTERACT.tex
fix the ion concetrations

--- 2015-02-12 LIPIDionINTERACT/LIPIDionINTERACT.tex
@@
 \caption{Simulated lipid bilayers with ions.
-}\label{IONsystems}
+The ion concentrations are the concentration of ions in buffer to solute the lipid bilayers and calculated as [ion]=(N$_{\rm ion} \times$[water])/N$_{\rm w}$,
+where [water]=55.5M. These correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.}\label{IONsystems}
 \begin{tabular}{c c c c c c c c c c c c}
@@
 Berger-POPC-07\cite{ollila07a}   &   POPC & 0          & 128 & 7290 & 0  & 0  & 0 & 298  & 270 & 240 & \cite{bergerFILESpopc}  \\
-Berger-POPC-07\cite{ollila07a}, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}   &   POPC & 190 (NaCl) & 128 & 7202 & 44  & 0  & 44 &298  & 110 & 50 &?\todoi{Samuli put to Zenodo} \\
-Berger-POPC-07\cite{ollila07a}, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}   &   POPC & 190 (CaCl$_2$) & 128 & 7157 & 0 & 44  & 88 &298 & 110 & 50 &?\todoi{Samuli put to Zenodo}  \\
+Berger-POPC-07\cite{ollila07a}, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}   &   POPC & 340 (NaCl) & 128 & 7202 & 44  & 0  & 44 &298  & 110 & 50 &?\todoi{Samuli put to Zenodo} \\
+Berger-POPC-07\cite{ollila07a}, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}   &   POPC & 340 (CaCl$_2$) & 128 & 7157 & 0 & 44  & 88 &298 & 110 & 50 &?\todoi{Samuli put to Zenodo}  \\
 \hline
 CHARMM36\cite{klauda10}   & POPC & 0           & 72 & 2242 & 0  & 0 & 0 & 303  & 30 & 20 & \cite{charmm36filesSHORT} \\
-CHARMM36\cite{klauda10}, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}   & POPC & 140 (NaCl)  & 72 & 2085 & 13  & 0 & 13 & 303  & 80 & 60 &?\todoi{Samuli put to Zenodo}\\
-CHARMM36\cite{klauda10}, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}   & POPC & 280 (NaCl)  & 72 & 2085 & 26  & 0 & 26 & 303  & 80 & 60 &?\todoi{Samuli put to Zenodo} \\
-CHARMM36\cite{klauda10}, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}  & POPC & 570 (NaCl)  & 72 & 2085 & 52  & 0 & 52 & 303  & 80 & 60 &?\todoi{Samuli put to Zenodo} \\
-CHARMM36\cite{klauda10}, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}   & POPC & 85  (NaCl)  & 72 & 2085 & 78  & 0 & 78 & 303  & 73 & 60 &?\todoi{Samuli put to Zenodo} \\
+CHARMM36\cite{klauda10}, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}   & POPC & 350 (NaCl)  & 72 & 2085 & 13  & 0 & 13 & 303  & 80 & 60 &?\todoi{Samuli put to Zenodo}\\
+CHARMM36\cite{klauda10}, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}   & POPC & 690 (NaCl)  & 72 & 2085 & 26  & 0 & 26 & 303  & 80 & 60 &?\todoi{Samuli put to Zenodo} \\
+CHARMM36\cite{klauda10}, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}  & POPC & 1380 (NaCl)  & 72 & 2085 & 52  & 0 & 52 & 303  & 80 & 60 &?\todoi{Samuli put to Zenodo} \\
+CHARMM36\cite{klauda10}, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}   & POPC & 2080  (NaCl)  & 72 & 2085 & 78  & 0 & 78 & 303  & 73 & 60 &?\todoi{Samuli put to Zenodo} \\
 \hline
@@
 MacRog\cite{maciejewski14}\todoi{DONE} , ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}  & POPC & 100 (NaCl) & 288 & 14554 & 27 & 0 & 27 & 310 & 90&50  & \cite{macrogIONfiles} \\
-MacRog\cite{maciejewski14}\todoi{DONE} , ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}        & POPC &  200 (NaCl) & 288 & 14500 & 54 & 0 & 54 & 310 & 90&50  &\cite{macrogIONfiles}  \\
-MacRog\cite{maciejewski14}\todoi{DONE} , ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}        & POPC &   300 (NaCl) & 288 & 14446 & 81 & 0 & 81 & 310 & 90&50  & \cite{macrogIONfiles} \\
-MacRog\cite{maciejewski14}\todoi{DONE} , ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}        & POPC &   400 (NaCl) & 288 & 14392 & 108 & 0 & 108 & 310 & 90& 50  & \cite{macrogIONfiles}  \\
+MacRog\cite{maciejewski14}\todoi{DONE} , ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}        & POPC &  210 (NaCl) & 288 & 14500 & 54 & 0 & 54 & 310 & 90&50  &\cite{macrogIONfiles}  \\
+MacRog\cite{maciejewski14}\todoi{DONE} , ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}        & POPC &   310 (NaCl) & 288 & 14446 & 81 & 0 & 81 & 310 & 90&50  & \cite{macrogIONfiles} \\
+MacRog\cite{maciejewski14}\todoi{DONE} , ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}        & POPC &   420 (NaCl) & 288 & 14392 & 108 & 0 & 108 & 310 & 90& 50  & \cite{macrogIONfiles}  \\
 \hline
@@
 This is unpublished model.}  \\
-Orange, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}   &   POPC & 250 (NaCl) & 72 & 2802 & 26 & 0  & 26 & 298 & 120 & 60 &?\todoi{Jukka Määttä and Luca Monticelli, please let us know if we can share some files through Zenodo.
+Orange, ionFF~\cite{??}\ &   POPC & 140 (NaCl) & 72 & 2866 & 7 & 0  & 7 & 298 & running & ? &?  \\
+Orange, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}   &   POPC & 510 (NaCl) & 72 & 2802 & 26 & 0  & 26 & 298 & 120 & 60 &?\todoi{Jukka Määttä and Luca Monticelli, please let us know if we can share some files through Zenodo.
 This is unpublished model.}  \\
-Orange, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}   &   POPC & 240 (CaCl$_2$)  & 72 & 2802 & 0 & 26  & 52 & 298 & 120 & 60 &?
+Orange, ionFF~\cite{??}  &   POPC & 1000 (NaCl) & 72 & 2780 & 50 & 0  & 50 & 298 & running & ? &? \\
+Orange, ionFF~\cite{??}\todoi{Appropriate reference for the ion model?}   &   POPC & 510 (CaCl$_2$)  & 72 & 2802 & 0 & 26  & 52 & 298 & 120 & 60 &?
 \todoi{Jukka Määttä and Luca Monticelli, please let us know if we can share some files.
@@
    }
-\todo{I think that we have to run Orange simulation with higher NaCl concentration. I will do it.} \\
+\todo{I have already Orange simulations running with 1M and 140mM NaCl concentrations. I think I should still run CHARMM36 with 1M and maybe with close to 150mM. 
+Some of the simulations had accidentally over 1M concentration and those are not useful for this plot.} \\
 \todo{I think that we have to run CHARMM36 with CaCl$_2$. If anyone happen to have the data and is willing to share,
@@
 If the data will be uploaded to Zenodo, it also trivial to calculate it with script. \\
-- It seems that relative integrated (total) ion density in different models do not correspond the reported relative concentrations.
-For example, total Na density seems bigger in Berger than in MacRog even though there is only minot difference in concentration, and into the wrong direction. \\
+- CONCENTRATION CALCULATION ISSUE FIXED, SEE http://nmrlipids.blogspot.com/2015/02/the-first-draft-of-ion-lipid.html?showComment=1423750867868\#c3435589938183508901. \\
+The systems are here now with quite different concentrations so comparison is not straigthforward. One option would be to run simulations of all
+systems with 150mM which would be close to physiological concentrations, and show the density distributions from those here. \\
 - I think that CHARMM36 simulation with CaCl$_2$ should be done \\

### 2015-02-20 HGmodelMANUSCRIPT/HGmodel_draft3.tex
updates to HGmodel manuscript

@@
 between lipids and cholesterol better than the Berger/H\"oltje model.
-This work has been, and continues to be, progressed and discussed through the blog: \url{nmrlipids.blogspot.fi}. 
-Everyone is invited to join the discussion and make contributions through the blog. 
-The manuscript will be eventually submitted to an appropriate scientific journal. 
-Everyone who has contributed to the work through the blog will be offered 
-coauthorship. For more details see: \url{nmrlipids.blogspot.fi}.
+{\bf This work has been done as an open collaboration by using the \url{nmrlipids.blogspot.fi} as an
+communication platform. All the scientific contributions have been done through the blog and
+are publicly available. In addition, almost all simulation trajectories and files are made available
+in the Zenodo community \url{https://zenodo.org/collection/user-nmrlipids} which has became the most extensive
+publicly available collection of molecular dynamics simulation trajectories of lipid bilayers.
+}
+%This work has been, and continues to be, progressed and discussed through the blog: \url{nmrlipids.blogspot.fi}. 
+%Everyone is invited to join the discussion and make contributions through the blog. 
+%The manuscript will be eventually submitted to an appropriate scientific journal. 
+%Everyone who has contributed to the work through the blog will be offered 
+%coauthorship. For more details see: \url{nmrlipids.blogspot.fi}.
 
@@
 The approach is inspired by the Polymath project~\cite{gowers09}, however there are some essential differences. 
-The manuscript pointing out problems in the glycerol backbone and
-headgroup structure in the most used molecular dynamics simulation model for lipid bilayers (Berger based models) was used as a starting point~\cite{ollila13}.
-After publishing the initial manuscript, the blog was opened for further contributions and discussion from anyone interested. 
-All the contributions were done publicly through the blog and the blog contributors were offered coauthorship. The final author list
-is based on the self-assesment of the authors' scientific contribution to the project.
-
-Large portion of the used simulation files and scripts are shared through the GitHub organization \url{https://github.com/NMRlipids}
-and the Zenodo community \url{https://zenodo.org/collection/user-nmrlipids}.
+{\bf The present approach was started by publishing a manuscript~\cite{ollila13} discussing the glycerol backbone and choline structures 
+in a Berger based model (the most used molecular dynamics simulation model for lipid bilayers).
+After publishing the initial manuscript, the open invitation for further contributions and discussion through the blog was presented.
+All the scientific contributions are done publicly through the blog and all contributors were offered coauthorship. 
+The acception of authorship is based on the self-assesment of the authors' scientific contribution to the project.
+The contributions from each author are summarized in ?? (the location depends on the journal probably).
+
+Almost all simulation data, including running parameters for reprocution and trajectories for further analysis, are collected
+to the nmrlipids community in Zenodo file repository \url{https://zenodo.org/collection/user-nmrlipids}.
+Thus, besides the main topic of this manuscript we present the most extensive available collection of simulations trajectories
+for lipid bilayers, which opens up numerous possiblities for different analyses with much less effort than previously required.
+Furher information, e.g. scripts, figures and manuscript text files, are available in the GitHub organization \url{https://github.com/NMRlipids}.
+}
 
@@
 CHARMM36~\cite{klauda10}      & POPC   & 128 &  5120    & 303 & 150 & 100  & \cite{charmm36files}$^*$ \todoi{DONE}   & SI \\
-CHARMM36~\cite{klauda10}       & DPPC   & 72  &  2189 & 323 & 30 & 25  & ?\todoi{Samuli, put to Zenodo}  & SI \\
+CHARMM36~\cite{klauda10}       & DPPC   & 72  &  2189 & 323 & 30 & 25  & \cite{charmmFILESdppc}$^*$\todoi{DONE}  & SI \\
 %CHARMM36\cite{klauda10}\todoi{Not included in the Fig.
@@
 Kukol \cite{kukol09}          & POPC   & 512 & 20564 & 298 & 50 & 30  & \cite{kukolFILES}$^*$ & SI \\
-Chiu \cite{chiu09} \todoi{Samuli, put to Zenodo}     & POPC  & 128 & 3552  & 298 & 56 & 50  & ? & SI \\
+Chiu \cite{chiu09} \todoi{DONE}     & POPC  & 128 & 3552  & 298 & 56 & 50  & \cite{chiuFILES}$^*$ & SI \\
 H\"ogberg08 \cite{rabinovich14}\todoi{Alexander Luybartsev, let us know if you share the files.}  & POPC   &  128 & 3840  & 303 & 100 & 80  & ? & \cite{rabinovich14}  \\
@@
 
-In general, we conclude that the atomistic resolution classical molecular dynamics simulations 
-is extremely convenient tool to give structural interpretation for the high resolution NMR data~\cite{ferreira14}. 
-However, in the case of phoshatidylcholine glycerol backbone and choline headgroup further model development is still required.
-
-This work has been, and continues to be, progressed and discussed through a blog at \url{nmrlipids.blogspot.fi}. 
-Everyone is invited to join the discussion and make contributions through the blog. 
-The manuscript will be eventually submitted to an appropriate scientific journal. 
-Everyone who has contributed to the work through the blog will be offered 
-coauthorship. For more details, see \url{nmrlipids.blogspot.fi}.
+- {\bf Besides the main conclusions, we have created the most extensive
+publicly available collection of molecular dynamics simulation trajectories of lipid bilayers
+into the Zenodo community \url{https://zenodo.org/collection/user-nmrlipids}. 
+This collection opens up numerous possiblities for different analyses with
+much less effort than previously required.
+
+In general, we conclude that to fully utilize the potential of atomistic resolution classical molecular dynamics simulations
+in the structural interpretation of the high resolution NMR data~\cite{ferreira14} for lipid bialyers one has to  
+improve the phoshatidylcholine glycerol backbone and choline headgroup parameters.
+
+This work has been done as an open collaboration by using the \url{nmrlipids.blogspot.fi} as an communication
+platform. All the scientific contributions have been done through the blog and are publicly
+available. }
+
+%This work has been, and continues to be, progressed and discussed through a blog at \url{nmrlipids.blogspot.fi}. 
+%Everyone is invited to join the discussion and make contributions through the blog. 
+%The manuscript will be eventually submitted to an appropriate scientific journal. 
+%Everyone who has contributed to the work through the blog will be offered 
+%coauthorship. For more details, see \url{nmrlipids.blogspot.fi}.
 
@@
 \subsection{General issues to think of about the manuscript}
-\todo{Should we emphasize more the excellent collection of simulation trajectories collected to Zenodo?} \\
+\todo{DONE}\\
 \todo{Maybe we should mention that CHARMM is significanlty more computationally expensive than other all-atom models.

### 2015-02-24 LIPIDionINTERACT/LIPIDionINTERACT.tex
update experimental data in lipid-ion interaction manuscript

@@
      with the Berger, CHARMM36, MacRog and Orange force fields compared to the experiments
-     by Akutsu et al.~\cite{akutsu81}. The signs are assumed to be the same as measured by Hong et al.~\cite{hong95a}, Hong et al.~\cite{hong95a} and Gross et al.~\cite{gross97}. 
-     The effect of ions to the g2 and g3 were not measured, thus only the values without ions are shown. The straight line between the results with and without ions is 
+     by Akutsu et al.~\cite{akutsu81} and Altenbach et al.~\cite{altenbach84}.
+The signs are assumed to be the same as measured by Hong et al.~\cite{hong95a}, Hong et al.~\cite{hong95a} and Gross et al.~\cite{gross97}.
+     The experimental data for the effect of ions to the glycerol backbone is not found, thus only the values without ions are shown. 
+     The straight line between the results with and without ions is plotted to guide the eye. 
+   }
+\end{figure}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=8cm]{OrderParameterIONSCaCl.eps}
+  \caption{\label{ordPcacl}
+    Order parameters as a function of CaCl concentration from simulations 
+     with the Berger 
+%CHARMM36, MacRog 
+and Orange force fields compared to the experiments by Akutsu et al.~\cite{akutsu81} and Altenbach et al.~\cite{altenbach84}. 
+The signs are assumed to be the same as measured by Hong et al.~\cite{hong95a}, Hong et al.~\cite{hong95a} and Gross et al.~\cite{gross97}. 
+     The effect of ions to the g1 and g2 were not measured, thus only the values without ions are shown. The straight line between the results with and without ions is 
      plotted to guide the eye. 

### 2015-02-24 LIPIDionINTERACT/calcCHANGES.sh

@@ -8,5 +8,10 @@ cat SLIPIDSnaclCONSpopc.dat | awk '{if(NR==1){a=$2;b=$3;c=$4;d=$5;e=$6;f=$7;g=$8
 cat BERGERcaclCONS.dat | awk '{if(NR==1){a=$2;b=$3;c=$4;d=$5;e=$6;f=$7;g=$8;h=$9;}print $1" " $2-a" "$3-b" "$4-c" "$5-d" "$6-e" "$7-f" "$8-g" "$9-h}' > BERGERcaclCONSchange.dat
 cat ORANGEcaclCONS.dat | awk '{if(NR==1){a=$2;b=$3;c=$4;d=$5;e=$6;f=$7;g=$8;h=$9;}print $1" " $2-a" "$3-b" "$4-c" "$5-d" "$6-e" "$7-f" "$8-g" "$9-h}' > ORANGEcaclCONSchange.dat
 
-cat EXPnaclCONS.dat | awk '{if(NR==1){a=$2;b=$3}print $1" " $2-a" "$3-b}' > EXPnaclCONSchange.dat
-cat EXPcaclCONS.dat | awk '{if(NR==1){a=$2;b=$3}print $1" " $2-a" "$3-b}' > EXPcaclCONSchange.dat
+#cat EXPnaclCONS.dat | awk '{if(NR==1){a=$2;b=$3}print $1" " $2-a" "$3-b}' > EXPnaclCONSchange.dat
+#cat EXPcaclCONS.dat | awk '{if(NR==1){a=$2;b=$3}print $1" " $2-a" "$3-b}' > EXPcaclCONSchange.dat
+cat ../DATAreportediINblog/DPPC/NaCl/EXP-Akutsu1981.dat | awk '{if($0!~"#") print $0}' | awk '{if(NR==1){a=$2;b=$3}print $1" " $2-a" "$3-b}' > EXP-Akutsu1981-DPPC-NaClchange.dat
+cat ../DATAreportediINblog/DPPC/CaCl/EXP-Akutsu1981-T332K.dat | awk '{if($0!~"#") print $0}' | awk '{if(NR==1){a=$2;b=$3}print $1" " $2-a" "$3-b}' > EXP-Akutsu1981-DPPC-CaClchange-T332.dat
[…]

### 2015-03-12 HGmodelMANUSCRIPT/HGmodel_draft3.tex
Fixed typo in equation (1)
Added definitions of M, F, Sigma and color scheme to FF ranking Fig.

@@
 \begin{equation}\label{orderP}
-S_{{\rm CH}}=\frac{3}{2}\langle \cos^2 \theta-1 \rangle,
+S_{{\rm CH}}=\frac{1}{2}\langle 3 \cos^2 \theta-1 \rangle,
 \end{equation} 
@@
   \caption{\label{FullHydrationComparisonTable}
-  Ranking of force fields based on data of Fig.~\ref{HGorderparameters}. 
-  Markus, please describe the definition of M, F and $\Sigma$ here.
+  Rough ranking of force fields based on data of Fig.~\ref{HGorderparameters}.
+  "M" indicates a magnitude problem,
+  "F" a forking problem.
+  Letter size shows the level (0-4) of severity;
+  the $\Sigma$-column shows the sum of these, i.e., the "total severity".
+  Color scheme:
+  "within experimental error" (dark green),
+  "almost within experimental error" (light green),
+  "clear deviation from experiments" (light red), and
+  "major deviation from experiments" (dark red).
   } 

### 2015-03-17 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
add HGmodel manuscript in ACS format

--- 2015-03-12 HGmodelMANUSCRIPT/HGmodel_draft3.tex
@@
-% \documentclass[rmp,aps,floatfix,authordate1-4,preprint]{revtex4}
-% \documentclass[prb,aps,floatfix,authordate1-4,preprint]{revtex4}
-
-\documentclass[pre,aps,floatfix,authordate1-4,twocolumn]{revtex4-1}
-%\documentclass[pre,aps,floatfix,authordate1-4]{revtex4-1}
-
-%\documentclass[aps,prl,preprint,superscriptaddress]{revtex4}
-
-%\documentclass[aps,prl,preprint,groupedaddress]{revtex4}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% This is a (brief) model paper using the achemso class
+%% The document class accepts keyval options, which should include
+%% the target journal and optionally the manuscript type.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass[journal=jacsat,manuscript=article]{achemso}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Place any additional packages needed here.  Only include packages
+%% which are essential, to avoid problems later. Do NOT use any
+%% packages which require e-TeX (for example etoolbox): the e-TeX
+%% extensions are not currently available on the ACS conversion
+%% servers.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
+\usepackage[obeyFinal]{easy-todo}
 \usepackage{rotating}
-\usepackage{times}
-\usepackage{graphicx}
-\usepackage{setspace}
-\usepackage{amsmath}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage[obeyFinal]{easy-todo}
-%\usepackage{scrtime}
-\usepackage[multiple]{footmisc}
-%\usepackage[sort&compress]{natbib}
-%\bibpunct{(}{)}{,}{n}{}{}
-%\renewcommand{\bibnumfmt}[1]{#1.}
-\usepackage[normalem]{ulem}
-
-\begin{document}
-%\singlespacing
-
-\onecolumngrid
-{\color{red}
-Matti Javanainen has made the text modifications to the manuscript in excellent way. He modified directly the manuscript 
-tex file available in the GitHub and then made a pull request to add a new version of the manuscript into the GitHub repository. 
-In addition, he used latexdiff which can be generated to show the pdf with changes.
-
-After Matti had added his version I looked the changes through and agreed. Then I changed his version to the up-to-date version 
-and made some minor modifications.
-
-All the mentioned files can be found from the GitHub repository
-https://github.com/NMRLipids/nmrlipids.blogspot.fi
-
-up-to date manuscript
-HGmodel\_draft3.tex
-
-Version Matti uploaded
-HGmodel\_draft3\_matti.tex
-
-pdf showing the changes made by matti:
-HGmodel\_draft3\_matti\_latexdiff.pdf
-
-I think that this is a very good and convenient way to work collaboratively with the manuscript. If you want you can also make 
-a pull request directly using the up-to-date manuscript file name (HGmodel\_draft3.tex). Then I will accept the pull request, 
-or we will do some discussion before acception. Using latexdiff is very good, but the history will also be saved automatically 
-to the GitHub.
-
-If you find this difficult (as it may be if you are not familiar with GitHub) you can also send your comments and modifications 
-in some other way. However, keep in mind that it is very important to clearly report somehow what you have done. 
-This is to ease my work to handle the input from large amount of contributors.
-}
-
-\title{Towards atomistic resolution structure of phosphatidylcholine glycerol backbone and choline headgroup at different ambient conditions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% If issues arise when submitting your manuscript, you may want to
+%% un-comment the next line.  This provides information on the
+%% version of every file you have used.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%\listfiles
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Place any additional macros here.  Please use \newcommand* where
+%% possible, and avoid layout-changing macros (which are not used
+%% when typesetting).
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Meta-data block
+%% ---------------
+%% Each author should be given as a separate \author command.
+%%
+%% Corresponding authors should have an e-mail given after the author
+%% name as an \email command. Phone and fax numbers can be given
+%% using \phone and \fax, respectively; this information is optional.
+%%
+%% The affiliation of authors is given after the authors; each
+%% \affiliation command applies to all preceding authors not already
+%% assigned an affiliation.
+%%
+%% The affiliation takes an option argument for the short name.  This
+%% will typically be something like "University of Somewhere".
+%%
+%% The \altaffiliation macro should be used for new address, etc.
+%% On the other hand, \alsoaffiliation is used on a per author basis
+%% when authors are associated with multiple institutions.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\author{Andrew N. Other}
+%\altaffiliation{A shared footnote}
+%\author{Fred T. Secondauthor}
+%\altaffiliation{Current address: Some other place, Othert\"own,
+%Germany}
+%\author{I. Ken Groupleader}
+%\altaffiliation{A shared footnote}
+%\email{i.k.groupleader@unknown.uu}
+%\phone{+123 (0)123 4445556}
+%\fax{+123 (0)123 4445557}
+%\affiliation[Unknown University]
+%{Department of Chemistry, Unknown University, Unknown Town}
+%\alsoaffiliation[Second University]
+%{Department of Chemistry, Second University, Nearby Town}
+%\author{Susanne K. Laborator}
+%\email{s.k.laborator@bigpharma.co}
+%\affiliation[BigPharma]
+%{Lead Discovery, BigPharma, Big Town, USA}
+%\author{Kay T. Finally}
+%\affiliation[Unknown University]
+%{Department of Chemistry, Unknown University, Unknown Town}
+%\alsoaffiliation[Second University]
+%{Department of Chemistry, Second University, Nearby Town}
 
 \author{Alexandru Botan}
-\thanks{The authors are listed in alphaphetical order.}
-\thanks{The author list is not completed.}
-\thanks{Institut Lumi\`ere Mati\`ere, UMR5306 Universit\'e Lyon 1-CNRS, Universit\'e de Lyon 69622 Villeurbanne, France}
+%\affiliation{The authors are listed in alphaphetical order.}
+%\affiliation{The author list is not completed.}
+\affiliation[Lyon CNRS]{Institut Lumi\`ere Mati\`ere, UMR5306 Universit\'e Lyon 1-CNRS, Universit\'e de Lyon 69622 Villeurbanne, France}
 %
 %\author{Andrea Catte}
-%\thanks{University of East Anglia, Norwich, United Kingdom}
+%\affiliation{University of East Anglia, Norwich, United Kingdom}
 \author{Fernando Favela}
+\affiliation[Mexico]{Mexico}
 \author{Patrick Fuchs}
-\thanks{Institut Jacques Monod, CNRS, Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
-%\thanks{Université Paris Diderot, Sorbonne Paris Cité, Paris, France}
+\affiliation[CNRS paris]{Institut Jacques Monod, CNRS, Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
+\alsoaffiliation[Diderot University]
+{Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
 \author{Matti Javanainen}
-\thanks{Tampere University of Technology, Tampere, Finland}
+\affiliation[Tampere University of Technology]
+{Tampere University of Technology}
 \author{Waldemar Kulig}
-\thanks{Tampere University of Technology, Tampere, Finland}
+\affiliation[Tampere University of Technology]
+{Tampere University of Technology, Tampere, Finland}
 \author{Antti Lamberg}
-\thanks{Department of Chemical Engineering, Kyoto University, Kyoto, Japan}
-%\author{Alexander Lyubartsev}
+\affiliation[Kyoto University]
+{Department of Chemical Engineering, Kyoto University, Kyoto, Japan}
 \author{Markus S. Miettinen}
-\thanks{Fachbereich Physik, Freie Universit\"at Berlin, Berlin, Germany}
+\affiliation[Freie Universitat Berlin] 
+{Fachbereich Physik, Freie Universitat Berlin, Berlin, Germany}
 \author{Luca Monticelli}
-\thanks{5IBCP, CNRS UMR 5086, Lyon, France}
-\author{Jukka M\"a\"att\"a}
-\thanks{Aalto University, Espoo, Finland}
-%\setcounter{page}{1}
-%\author{Vasily S. Oganesyan}
-%\thanks{University of East Anglia, Norwich, United Kingdom}
+\affiliation[UMR] 
+{5IBCP, CNRS UMR 5086, Lyon, France}
+\author{Jukka Maatta}
+\affiliation[Aalto University]
+{Aalto University, Espoo, Finland}
 \author{O. H. Samuli Ollila} 
-\thanks{{\bf Author to whom correspondence may be addressed. E-mail: samuli.ollila@aalto.fi.}}
-\thanks{Helsinki Biophysics and Biomembrane Group, Department of Biomedical Engineering and Computational Science, Aalto University, Espoo, Finland}
+\email{samuli.ollila@aalto.fi.}
+\affiliation[Aalto University]
+{Aalto University, Espoo, Finland}
 \author{Marius Retegan}
-\thanks{Max Planck Institute for Chemical Energy Conversion, M\"ulheim an der Ruhr, Germany}
+\affiliation[Max Planck]
+{Max Planck Institute for Chemical Energy Conversion, Mulheim an der Ruhr, Germany}
 \author{Tomasz Rog}
-\thanks{Tampere University of Technology, Tampere, Finland}
+\affiliation[Tampere University of Technology]
+{Tampere University of Technology, Tampere, Finland}
 \author{Hubert Santuz}
-\thanks{INSERM, U1134, DSIMB; Institut National de la Transfusion Sanguine (INTS); Laboratoire d'Excellence GR-Ex, Paris, France}
-\thanks{Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
+\affiliation[INSERM]
+{INSERM, U1134, DSIMB; Institut National de la Transfusion Sanguine (INTS); Laboratoire d'Excellence GR-Ex, Paris, France}
+\alsoaffiliation[Diderot]
+{Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
 \author{Joona Tynkkynen}
-\thanks{Tampere University of Technology, Tampere, Finland}
-%\author{Mark Wilson}
-%\author{Alexander Vogel}
-
+\affiliation[Tampere University of Technology]
+{Tampere University of Technology, Tampere, Finland}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% The document title should be given as usual. Some journals require
+%% a running title from the author: this should be supplied as an
+%% optional argument to \title.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\title[An \textsf{achemso} demo]
+  {Towards atomistic resolution structure of phosphatidylcholine glycerol backbone and choline headgroup at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Some journals require a list of abbreviations or keywords to be
+%% supplied. These should be set up here, and will be printed after
+%% the title and author information, if needed.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\abbreviations{IR,NMR,UV}
+\keywords{American Chemical Society, \LaTeX}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% The manuscript does not need to include \maketitle, which is
+%% executed automatically.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{document}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% The "tocentry" environment can be used to create an entry for the
+%% graphical table of contents. It is given here as some journals
+%% require that it is printed as part of the abstract page. It will
+%% be automatically moved as appropriate.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{tocentry}
+
+Some journals require a graphical entry for the Table of Contents.
+This should be laid out ``print ready'' so that the sizing of the
+text is correct.
+
+Inside the \texttt{tocentry} environment, the font used is Helvetica
+8\,pt, as required by \emph{Journal of the American Chemical
+Society}.
+
+The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
+permitted for  \emph{Journal of the American Chemical Society}
+graphical table of content entries. The box will not resize if the
+content is too big: instead it will overflow the edge of the box.
+
+This box and the associated title will always be printed on a
+separate page at the end of the document.
+
+\end{tocentry}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% The abstract environment will automatically gobble the contents
+%% if an abstract is not used by the target journal.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{abstract}
@@
 would automatically resolve the structures giving an interpretation of experimental results, if the model
-reproduced the experimental data. In the present work we simulate phosphatidylcholine (PC) lipid bilayers with 13 different
+reproduced the experimental data. 
+In the present work we simulate phosphatidylcholine (PC) lipid bilayers with 13 different
 atomistic models, and we compare simulations with experiments (in fully hydrated conditions) in terms of C--H bond vector order
-parameters for the glycerol backbone and choline headgroups. None of the current models is not accurately enough to resolve the structure with experimental accuracy.
-However, closer inspection of three best performing models (CHARMM36, GAFFlipid and MacRog) suggest
-that improvements in the sampled dihedral angle distributions would potentilly lead to the model which
-would resolve the structure. Despite of the inaccuracy in the fully hydrated
-structures, the response to dehydration, i.e. the tilting of the P--N vector more parallel to membrane normal, 
-is qualitatively correct in all models. The CHARMM36 and MacRog models describe the interactions
-between lipids and cholesterol better than the Berger/H\"oltje model.
-{\bf This work has been done as an open collaboration by using the \url{nmrlipids.blogspot.fi} as an
+parameters for the glycerol backbone and choline headgroups. 
+We show that this comparison can be used to judge the atomistic resolution structural accuracy of the models,
+thus allowing the usage of molecular dynamics simulations to interpret the structures of biomolecules in 
+biologically relevant conditions from NMR experiments. Further, we review previous experimental
+data for dehydrated lipid bilayers and cholesterol containing bilayers, and interpretate it with simulations.
+The interpretation reveals that the increased choline order parameters indicate P--N vector tilting more parallel to membrane,
+and that cholesterol induces only minor changes to the glycerol backbone structure.
+%None of the current models is not accurately enough to resolve the structure with experimental accuracy.
+%However, closer inspection of three best performing models (CHARMM36, GAFFlipid and MacRog) suggest
+%that improvements in the sampled dihedral angle distributions would potentilly lead to the model which
+%would resolve the structure. Despite of the inaccuracy in the fully hydrated
+%structures, the response to dehydration, i.e. the tilting of the P--N vector more parallel to membrane normal, 
+%is qualitatively correct in all models. The CHARMM36 and MacRog models describe the interactions
+%between lipids and cholesterol better than the Berger/H\"oltje model.
+This work has been done as an open collaboration by using the \url{nmrlipids.blogspot.fi} as an
 communication platform. All the scientific contributions have been done through the blog and
@@
 publicly available collection of molecular dynamics simulation trajectories of lipid bilayers.
-}
-%This work has been, and continues to be, progressed and discussed through the blog: \url{nmrlipids.blogspot.fi}. 
-%Everyone is invited to join the discussion and make contributions through the blog. 
-%The manuscript will be eventually submitted to an appropriate scientific journal. 
-%Everyone who has contributed to the work through the blog will be offered 
-%coauthorship. For more details see: \url{nmrlipids.blogspot.fi}.
-
 \end{abstract}
 
-\maketitle
-
-~\vspace{0.3cm}\\
-{\it \bf}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Start the main part of the manuscript here.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 but general consensus about the structures sampled in the fluid state have not been reached~\cite{seelig77b,skarjune79,Israelachvili80,jacobs80,davis83,strenk85,hauser88,akutsu91,hong95b,hong96,semchyschyn04,marsh06}.
-On the other hand, the glycerol backbone structures are similar for various biologically
+On the other hand, the structural parameters for the glycerol backbone are similar for various biologically
 relevant lipid species (phosphatidylcholine (PC), phosphatidylethanolamine (PE) and phosphatidylglycerol (PG)) 
-in various environments~\cite{gally81} and the headgroup choline structures are similar in model membranes and
+in various environments~\cite{gally81} and the structural parameters for the headgroup choline structures are similar in model membranes and
 real cells (mouse fibroblast L-M cell)~\cite{scherer87}.
@@
 structures between the most often used (Berger based) lipid model~\cite{berger97} 
-\todo{If there are no objections, we will leave the naming convention of different Berger versions as it is now.
-%Based on the results and discussions with Olle Edholm and Peter Tieleman in the blog, it seems that 
-%the glycerol backbone dihedrals are not the same in the model used in the publication by Berger et al.~\cite{berger97}
-%and in the model shared in the webpage of Peter Tieleman (http://wcm.ucalgary.ca/tieleman/downloads).
-%In this work (as it is quite often done in other works as well) we currently refer to the model in the webpage as ''Berger'' model.
-%This is misleading. We have to call it something else. One suggestion could be to call it ''BergerM'', because it was firt time
-%used in Marrink et al. Biophys. J. 74,931-943 (1998). Any other suggestions?\\
-%COMMENT: Patrick Fuchs suggested ''Berger-T''. This is better. \\
-%COMMENT2: Current trial, based in blog discussion, is to use Berger-MOLECULE-YEAR [citation].
-}
+%\todo{If there are no objections, we will leave the naming convention of different Berger versions as it is now.}
 and the best performing models, to demonstrate that by using the 
@@
 \subsection{Open collaboration}
-\todo{Should we write more about this?}
-
-This work has been done as an open collaboration by using \url{nmrlipids.blogspot.fi} as a communication platform.
+%\todo{Should we write more about this?}
+
+This work has been done as an open collaboration by using \url{nmrlipids.blogspot.fi}
+\todo{Current idea is that we will assign doi for the blog by using the help from  winnower administrators:
+https://thewinnower.com/posts/archiving-and-aggregating-alternative-scholarly-content-dois-for-blogs.
+Once done, that doi will be cited here and other places where blog content needs to be cited.} 
+as a communication platform.
 The approach is inspired by the Polymath project~\cite{gowers09}, however there are some essential differences. 
-{\bf The present approach was started by publishing a manuscript~\cite{ollila13} discussing the glycerol backbone and choline structures 
+The present approach was started by publishing a manuscript~\cite{ollila13} discussing the glycerol backbone and choline structures 
 in a Berger based model (the most used molecular dynamics simulation model for lipid bilayers).
 After publishing the initial manuscript, the open invitation for further contributions and discussion through the blog was presented.
-All the scientific contributions are done publicly through the blog and all contributors were offered coauthorship. 
+All the scientific contributions are done publicly through the blog and all contributors were offered coauthorship
+according to the guidelines defined in the beginning of the project\todo{Citation to the On Credits in the blog here}. 
 The acception of authorship is based on the self-assesment of the authors' scientific contribution to the project.
-The contributions from each author are summarized in ?? (the location depends on the journal probably).
+The contributions from each author are summarized in the Supplementary Information. %(the location depends on the journal probably).
 
@@
 for lipid bilayers, which opens up numerous possiblities for different analyses with much less effort than previously required.
-Furher information, e.g. scripts, figures and manuscript text files, are available in the GitHub organization \url{https://github.com/NMRlipids}.
-}
+Furher information, e.g. scripts, figures and manuscript text files, are available in the GitHub 
+organization \url{https://github.com/NMRlipids} \todo{Doi will be assigned to the GitHub repository before submission}.
 
 \subsection{Order parameters from experiments}\label{expORDp}
-\todo{DONE
-%We should mention the sign issue and refer to the blog post:
-%http://nmrlipids.blogspot.fi/2014/04/on-signs-of-order-parameters.html somewhere.
-%Maybe we should change the name of this section II.B. to: "Analysis and terminology of order parameters from experiments 
-%and simulations" or similar, and also move the "forking" explanation here.
-}
 The order parameter of a hydrocarbon C--H vector is defined as 
@@
 For a more detailed discussion see \url{http://nmrlipids.blogspot.fi/2014/02/accuracy-of-order-parameter-measurements.html}
-\todo{This is temporarily linked to the blog. Currently I think that the posts (with comments included) to which we want to cite are put into 
-the Figshare in pdf format as they are in when the publication is submitted. Figshare allows commenting to continue the discussion.
-COMMENT: There is also the option to put the most relevant blog posts as a supplementary material. The advantage of this would
-be that the posts would be peer reviewed as well.}.
-
-{\bf The absolute values of order parameters are accessible with both $^2$H NMR and $^{13}$C NMR techniques. 
+\todo{Citation to the blog will be added when available.}
+
+The absolute values of order parameters are accessible with both $^2$H NMR and $^{13}$C NMR techniques. 
 However, only $^{13}$C NMR techniques allow also the measurement of the sign of the order parameter~\cite{hong95a,hong95b,gross97}. 
@@
 For more detailed discussion about the sign measurements of the order parameters see \url{http://nmrlipids.blogspot.fi/2014/04/on-signs-of-order-parameters.html}
+\todo{Citation to the blog will be added when available.}
 
@@
 separate populations of lipid conformations~\cite{engel81,gally81}.
-}
 
@@
 in the trajectories post-simulationally using the positions of the heavy atoms and the known hydrocarbon geometries.
-{\bf For the statistical error estimates, the time average of order parameters were first calculated separately
+For the statistical error estimates, the time average of order parameters were first calculated separately
 for each lipid in the system. Then it was assumed that different lipids are statistically independent 
 entities (which should be the case in fluid phase) and the error of the mean for the average over individual 
-lipids in the system was calculated and used as error bar for the order parameters.}
+lipids in the system was calculated and used as error bar for the order parameters.
 
@@
 data is used previously therein. Additionally, the files related to the simulations and the resulting trajectories are publicly
-available for some systems. The references pointing to simulation details and files are also listed in Tables~\ref{systems},~\ref{systemsDEHYD} and~\ref{systemsCHOL}.
-
-\begin{table*}[]
+available for almost all systems in the Zenodo collection \url{https://zenodo.org/collection/user-nmrlipids}. 
+The references pointing to simulation details and files are also listed in Tables~\ref{systems},~\ref{systemsDEHYD} and~\ref{systemsCHOL}.
+
+\begin{table}[]
 \centering
@@
 1,2-dioleoyl-sn-glycero-3-phosphocholine (DOPC) and dilauroylphosphatidylcholine (DLPC).
+$^a$ The number of lipid molecules
+$^b$ The number of water molecules
+$^c$ Simulation temperature
+$^d$ The total simulation time
+$^e$ Time frames used in the analysis
+$^f$ Reference link for the downloadable simulation files
+$^g$ Reference for the full simulation details
 }\label{systems}
@@
 %\hline
-Force field & lipid  & \footnote{The number of lipid molecules}N$_{\rm l}$   &  \footnote{The number of water molecules}N$_{\rm w}$ & \footnote{Simulation temperature}T (K)  & \footnote{The total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time frames used in the analysis}t$_{{\rm anal}}$ (ns) & \footnote{Reference link for the downloadable simulation files}Files  &  \footnote{Reference for the full simulation details} Details\\
+Force field & lipid  & $^a$N$_{\rm l}$   &  $^b$N$_{\rm w}$ &  $^c$T (K)  &  $^d$t$_{{\rm sim}}$(ns) &  $^e$t$_{{\rm anal}}$ (ns) &  $^f$Files  &  $^g$Details\\
 \hline
-Berger-POPC-07~\cite{ollila07a}          &   POPC & 128 & 7290  & 298  & 270 & 240 & \cite{bergerFILESpopc}$^*$ & \cite{ferreira15} \\
+Berger-POPC-07~\cite{ollila07a}          &   POPC & 128 & 7290  & 298  & 270 & 240 & [\citenum{bergerFILESpopc}]$^*$ & [\citenum{ferreira15}] \\
 %\sout {Berger-POPC-07~\cite{ollila07a}}\todoi{Not included in the Fig. 2, to be removed?}          &   POPC & 128 & 7290  & 323  & 56 & 56  & ? & SI \\
-\todoi{DONE} 
-Berger-DPPC-98~\cite{marrink98}          &   DPPC & 72 & 2864  & 323  & 140 & 100  & \cite{bergerDPPCfiles} & SI \\
-Berger-DMPC-04~\cite{gurtovenko04}          &   DMPC & 128 & 5097  & 323  & 130 & 100  & \cite{dmpcFILES} & \cite{miettinen09} \\
-CHARMM36~\cite{klauda10}       & POPC   & 72  &  2242 & 303 & 30 & 20  & \cite{charmm36filesSHORT}$^*$ & SI \\
-CHARMM36~\cite{klauda10}      & POPC   & 128 &  5120    & 303 & 150 & 100  & \cite{charmm36files}$^*$ \todoi{DONE}   & SI \\
-CHARMM36~\cite{klauda10}       & DPPC   & 72  &  2189 & 323 & 30 & 25  & \cite{charmmFILESdppc}$^*$\todoi{DONE}  & SI \\
+Berger-DPPC-98~\cite{marrink98}          &   DPPC & 72 & 2864  & 323  & 140 & 100  & [\citenum{bergerDPPCfiles}] & SI \\
+Berger-DMPC-04~\cite{gurtovenko04}          &   DMPC & 128 & 5097  & 323  & 130 & 100  & [\citenum{dmpcFILES}] & [\citenum{miettinen09}] \\
+CHARMM36~\cite{klauda10}       & POPC   & 72  &  2242 & 303 & 30 & 20  & [\citenum{charmm36filesSHORT}]$^*$ & SI \\
+CHARMM36~\cite{klauda10}      & POPC   & 128 &  5120    & 303 & 150 & 100  & [\citenum{charmm36files}]$^*$   & SI \\
+CHARMM36~\cite{klauda10}       & DPPC   & 72  &  2189 & 323 & 30 & 25  & [\citenum{charmmFILESdppc}]$^*$  & SI \\
 %CHARMM36\cite{klauda10}\todoi{Not included in the Fig.
@@
 Should we remove or include?}        & DOPC   & 128 &  5120    & 303 & 150 & 100  & ?\todoi{Permanent link in progress by Hubert Santuz}  & SI \\
-\todoi{DONE} \todoi{Permanent link for DOPC CHARMM36 simulation by Hubert Santuz is not required now. However, it would be probably useful in the future.}
-MacRog~\cite{maciejewski14}  & POPC & 288  & 12600 & 310 & 100 & 80  & \cite{macrogFILES}$^*$ & SI  \\
-GAFFlipid~\cite{dickson12}       & POPC & 126  & 3948  & 303 & 137 & 32  & \cite{GAFFlipidFILES}$^*$ & SI \\
-GAFFlipid~\cite{dickson12}       & DPPC & 72  & 2197  & 323 & 90 & 50  & \cite{GAFFlipidFILESdppc}$^*$ & SI \\
-Lipid14 \cite{dickson14}         & POPC  & 72 & 2234 & 303 & 100 & 50  & \cite{lipid14files}$^*$ & SI \\
-Poger \cite{poger10}             & DPPC  & 128 & 5841 & 323 & 2$\times$100 & 2$\times$50 & \cite{pogerFILESpme1,pogerFILESpme2}$^*$\todoi{DONE} & SI \\
-Slipid \cite{jambeck12}          & DPPC & 128 & 3840 & 323 & 150 & 100 & \cite{slipidsFILES}$^*$ & SI \\
-Slipid \cite{jambeck12b}          & POPC & 128 & 5120 & 303 & 200 & 150 & \cite{slipidsFILESpopc}$^*$ & SI \\
-Kukol \cite{kukol09}          & POPC   & 512 & 20564 & 298 & 50 & 30  & \cite{kukolFILES}$^*$ & SI \\
-Chiu \cite{chiu09} \todoi{DONE}     & POPC  & 128 & 3552  & 298 & 56 & 50  & \cite{chiuFILES}$^*$ & SI \\
-H\"ogberg08 \cite{rabinovich14}\todoi{Alexander Luybartsev, let us know if you share the files.}  & POPC   &  128 & 3840  & 303 & 100 & 80  & ? & \cite{rabinovich14}  \\
-H\"ogberg08 \cite{hogberg08}\todoi{Alexander Luybartsev, let us know if you share the files.}  & DMPC   &  98 & 3840  & 303 & 75 & 50 & ? & \cite{hogberg08} \\
-Ulmschneiders \cite{Ulmschneider09}    & POPC  & 128 & 3328 & 310 & 100 & 50 & \cite{ulmschneiderFILES}$^*$ & SI \\
-Tj\"ornhammar14 \cite{tjornhammar14}   & DPPC  & 144 & 7056 & 323 & 200 & 100 & \cite{tjornhammarfiles}$^*$ & \cite{tjornhammar14} \\
-%\sout{OPLS-AA}\cite{rog09b}\todoi{We have this only with 150mM of NaCl delivered by Joona Tynkkynen.
-Options are to remove these results from this publicatio, run the simulations without ions (if not yet available),
-%  or include the results with ions. In my understanding this is a proto version of MacRog so it could be left out as well. However, for historical reasons and 
-%to understand literature it might be useful to include also these results. COMMENT: Matti Javanainen suggested that we would remove these results. If there are no objections and the
-%data without ions will not be delivered, we will remove this.}         &   &  &  & & & & & \\
-CHARMM36-UA~\cite{henin08,lee14}     & DLPC   & 128  & 3840  & 323 & 30 & 20 & \cite{charmmUAfiles} & SI \\
+%\todoi{Permanent link for DOPC CHARMM36 simulation by Hubert Santuz is not required now. However, it would be probably useful in the future.}
+MacRog~\cite{maciejewski14}  & POPC & 288  & 12600 & 310 & 100 & 80  & [\citenum{macrogFILES}]$^*$ & SI  \\
+GAFFlipid~\cite{dickson12}       & POPC & 126  & 3948  & 303 & 137 & 32  & [\citenum{GAFFlipidFILES}]$^*$ & SI \\
+GAFFlipid~\cite{dickson12}       & DPPC & 72  & 2197  & 323 & 90 & 50  & [\citenum{GAFFlipidFILESdppc}]$^*$ & SI \\
+Lipid14 \cite{dickson14}         & POPC  & 72 & 2234 & 303 & 100 & 50  & [\citenum{lipid14files}]$^*$ & SI \\
+Poger \cite{poger10}             & DPPC  & 128 & 5841 & 323 & 2$\times$100 & 2$\times$50 & [\citenum{pogerFILESpme1,pogerFILESpme2}]$^*$ & SI \\
+Slipid \cite{jambeck12}          & DPPC & 128 & 3840 & 323 & 150 & 100 & [\citenum{slipidsFILES}]$^*$ & SI \\
+Slipid \cite{jambeck12b}          & POPC & 128 & 5120 & 303 & 200 & 150 & [\citenum{slipidsFILESpopc}]$^*$ & SI \\
+Kukol \cite{kukol09}          & POPC   & 512 & 20564 & 298 & 50 & 30  & [\citenum{kukolFILES}]$^*$ & SI \\
+Chiu \cite{chiu09}      & POPC  & 128 & 3552  & 298 & 56 & 50  & [\citenum{chiuFILES}]$^*$ & SI \\
+H\"ogberg08 \cite{rabinovich14}  & POPC   &  128 & 3840  & 303 & 100 & 80  & - & [\citenum{rabinovich14}]  \\
+H\"ogberg08 \cite{hogberg08}  & DMPC   &  98 & 3840  & 303 & 75 & 50 & - & [\citenum{hogberg08}] \\
+Ulmschneiders \cite{Ulmschneider09}    & POPC  & 128 & 3328 & 310 & 100 & 50 & [\citenum{ulmschneiderFILES}]$^*$ & SI \\
+Tj\"ornhammar14 \cite{tjornhammar14}   & DPPC  & 144 & 7056 & 323 & 200 & 100 & [\citenum{tjornhammarfiles}]$^*$ & [\citenum{tjornhammar14}] \\
+CHARMM36-UA~\cite{henin08,lee14}     & DLPC   & 128  & 3840  & 323 & 30 & 20 & [\citenum{charmmUAfiles}] & SI \\
 \end{tabular}
-\end{table*}
-
-\begin{table*}[htb]
+\end{table}
+
+\begin{table}[htb]
 \centering
 \caption{Simulated single component lipid bilayers with varying hydration levels. The simulation file data sets marked with $^*$ include also part of the trajectory.
+$^a$ Water/lipid molar ratio
+$^b$ The number of lipid molecules
+$^c$ The number of water molecules
+$^d$ Simulation temperature
+$^e$ The total simulation time
+$^f$ Time frames used in the analysis
+$^g$ Reference link for the downloadable simulation files
+$^h$ Reference for the full simulation details
 }\label{systemsDEHYD}
@@
 %\hline
-Force field & lipid & \footnote{Water/lipid molar ratio} n (w/l)   & \footnote{The number of lipid molecules}N$_{\rm l}$   &  \footnote{The number of water molecules}N$_{\rm w}$ & \footnote{Simulation temperature}T (K)  & \footnote{The total simulation time}t$_{{\rm sim}}$(ns)  & \footnote{Time frames used in the analysis}t$_{{\rm anal}}$ (ns)& \footnote{Reference link for the downloadable simulation files}Files  &  \footnote{Reference for the full simulation details} Details\\
+Force field & lipid & $^a$n (w/l)   & $^b$N$_{\rm l}$   &  $^c$N$_{\rm w}$ & $^d$T (K)  & $^e$t$_{{\rm sim}}$(ns)  & $^f$t$_{{\rm anal}}$ (ns)& $^g$Files  &  $^h$Details\\
 \hline
-Berger-POPC-07~\cite{ollila07a}          &   POPC & 57  &128 & 7290  & 298  & 270 & 240 & \cite{bergerFILESpopc}$^*$ & SI \\
-                                        &   POPC & 7  &128 & 896   & 298  & 60 & 50 & \cite{bergerDEHYDfiles}$^*$ & SI \\
-CHARMM36\cite{klauda10}              & POPC   & 31 & 72  &  2242 & 303 & 30 & 20 & \cite{charmm36filesSHORT}$^*$ & SI \\
-                               & POPC   & 15 & 72 &  1080  & 303 & 59 & 40 & \cite{charmm36files15wPERl}$^*$ & SI \\
-                            & POPC   & 7  & 72  &  504  & 303 & 60 & 20 & \cite{charmm36files7wPERl}$^*$ & SI \\
-MacRog\cite{maciejewski14}     & POPC   & 50 & 288  & 14400 & 310 & 90 & 40 & \cite{macrogdehydFILES}$^*$ & SI \\
-                               & POPC   & 25 & 288  & 7200 & 310 & 100 & 50 & \cite{macrogdehydFILES}$^*$ & SI \\
-                                & POPC   & 20 & 288  & 5760 & 310 & 100 & 50 & \cite{macrogdehydFILES}$^*$ & SI \\
-                                & POPC   & 15 & 288  & 4320 & 310 & 100 & 50 & \cite{macrogdehydFILES}$^*$ & SI \\
-                                & POPC   & 10 & 288  & 2880 & 310 & 100 & 50 & \cite{macrogdehydFILES}$^*$ & SI \\
-                                & POPC   & 5 & 288   & 1440 & 310 & 100 & 50 & \cite{macrogdehydFILES}$^*$ & SI \\
-GAFFlipid\cite{dickson12}      & POPC   & 31& 126  & 3948  & 303 & 137 & 32 & \cite{GAFFlipidFILES}$^*$ & SI  \\
-                               & POPC   & 7& 126  & 896   & 303 & 130 & 40 & \cite{gaffDEHYDfiles}$^*$ & SI \\
+Berger-POPC-07~\cite{ollila07a}          &   POPC & 57  &128 & 7290  & 298  & 270 & 240 & [\citenum{bergerFILESpopc}]$^*$ & SI \\
+                                        &   POPC & 7  &128 & 896   & 298  & 60 & 50 & [\citenum{bergerDEHYDfiles}]$^*$ & SI \\
+CHARMM36\cite{klauda10}              & POPC   & 31 & 72  &  2242 & 303 & 30 & 20 & [\citenum{charmm36filesSHORT}]$^*$ & SI \\
+                               & POPC   & 15 & 72 &  1080  & 303 & 59 & 40 & [\citenum{charmm36files15wPERl}]$^*$ & SI \\
+                            & POPC   & 7  & 72  &  504  & 303 & 60 & 20 & [\citenum{charmm36files7wPERl}]$^*$ & SI \\
+MacRog\cite{maciejewski14}     & POPC   & 50 & 288  & 14400 & 310 & 90 & 40 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
+                               & POPC   & 25 & 288  & 7200 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
+                                & POPC   & 20 & 288  & 5760 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
+                                & POPC   & 15 & 288  & 4320 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
+                                & POPC   & 10 & 288  & 2880 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
+                                & POPC   & 5 & 288   & 1440 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
+GAFFlipid\cite{dickson12}      & POPC   & 31& 126  & 3948  & 303 & 137 & 32 & [\citenum{GAFFlipidFILES}]$^*$ & SI  \\
+                               & POPC   & 7& 126  & 896   & 303 & 130 & 40 & [\citenum{gaffDEHYDfiles}]$^*$ & SI \\
 \end{tabular}
-\end{table*}
-
-\begin{table*}[htb]
+\end{table}
+
+\begin{sidewaystable}[]
 \centering
 \caption{Simulated lipid bilayers containing cholesterol. The simulation file data sets marked with $^*$ include also part of the trajectory.
+$^a$ The number of lipid molecules
+$^b$ The number of cholesterol molecules
+$^c$ Cholesterol concentration (mol\%)
+$^d$ The number of water molecules
+$^e$ Simulation temperature
+$^f$ The total simulation time
+$^g$ Time frames used in the analysis
+$^h$ Reference link for the downloadable simulation files
+$^i$ Reference for the full simulation details
 }\label{systemsCHOL}
-\begin{tabular}{c c c c c c c c c c}
+\begin{tabular}{c c c c c c c c c c c}
 %\hline
-Force field & lipid  & \footnote{The number of lipid molecules}N$_{\rm l}$  & \footnote{number of cholesterol molecules}N$_{\rm chol}$   &  \footnote{The number of water molecules}N$_{\rm w}$ & \footnote{Simulation temperature}T (K)  & \footnote{The total simulation time}t$_{{\rm sim}}$(ns)  & \footnote{Time frames used in the analysis}t$_{{\rm anal}}$ (ns)\\
+Force field & lipid   & $^a$N$_{\rm l}$ & $^b$N$_{\rm chol}$ &$^c$C$_{\rm CHOL}$  &  $^d$N$_{\rm w}$ & $^e$T (K)  & $^f$t$_{{\rm sim}}$(ns)  & $^g$t$_{{\rm anal}}$ (ns)& $^h$Files  &  $^i$Details\\
 \hline
-Berger-POPC-07~\cite{ollila07a}/H\"oltje-CHOL-13~\cite{holtje01,ferreira13}          &   POPC &128 & 0 & 7290  & 298  & 270 & 240 & \cite{bergerFILESpopc}$^*$ & \cite{ferreira15} \\
-                               &   POPC &120 & 8 & 7290   & 298  & 100 & 80 & \cite{bergerFILESpopc7chol}$^*$ & \cite{ferreira13} \\
-                               &   POPC &110 & 18 & 8481  & 298  & 100 & 80 & \cite{bergerFILESpopc15chol}$^*$ & \cite{ferreira13}  \\
-                               &   POPC &84 & 44 & 6794   & 298  & 100 & 80 & \cite{bergerFILESpopc34chol}$^*$ & \cite{ferreira13} \\
-                               &   POPC &64 & 64 & 10314  & 298  & 100 & 80 & \cite{bergerFILESpopc50chol}$^*$ & \cite{ferreira13} \\
-                               &   POPC &50 & 78 & 5782   & 298  & 100 & 80 & \cite{bergerFILESpopc60chol}$^*$ & \cite{ferreira13} \\
-CHARMM36\cite{klauda10,lim12}\todoi{DONE}    & POPC   & 128&  0 & 5120  & 303  & 150 & 100 & \cite{charmm36files}$^*$  & SI  \\
- \todoi{DONE}                                & POPC   & 100  & 24  &  4960   & 303 & 200 & 100 & \cite{charmm36files20perCHOL}$^*$ & SI \\
-  \todoi{DONE}                               & POPC   & 80  & 80  &  4496    & 303 & 200 & 100 & \cite{charmm36files50perCHOL}$^*$ & SI \\
-MacRog\cite{maciejewski14}     & POPC   & 128 & 0  & 6400   & 310 & 400 & 200 & \cite{macrogCHOLfiles}$^*$ & SI \\ 
-                          & POPC   & 114  & 14 & 6400  & 310  & 400 & 200 & \cite{macrogCHOLfiles}$^*$ & SI    \\
-                          & POPC   & 72   & 56 & 6400  & 310  & 400 & 200 & \cite{macrogCHOLfiles}$^*$ & SI    \\
-                             & POPC   & 64  & 64 & 6400  & 310  & 400 & 200 & \cite{macrogCHOLfiles}$^*$ & SI    \\
-                             & POPC   & 56   & 72 & 6400  & 310  & 400 & 200 & \cite{macrogCHOLfiles}$^*$ & SI    \\
+Berger-POPC-07~\cite{ollila07a}&   POPC &128 & 0 &0\% & 7290  & 298  & 270 & 240 & [\citenum{bergerFILESpopc}]$^*$ & [\citenum{ferreira15}] \\
+/H\"oltje-CHOL-13~\cite{holtje01,ferreira13}   &    & &  &   &   &  &  &  &  \\
+                               &   POPC &120 & 8 & 6\% &7290   & 298  & 100 & 80 & [\citenum{bergerFILESpopc7chol}]$^*$ & [\citenum{ferreira13}] \\
+                               &   POPC &110 & 18& 14\% & 8481  & 298  & 100 & 80 & [\citenum{bergerFILESpopc15chol}]$^*$ & [\citenum{ferreira13}]  \\
+                               &   POPC &84 & 44 & 34\%  & 6794   & 298  & 100 & 80 & [\citenum{bergerFILESpopc34chol}]$^*$ & [\citenum{ferreira13}] \\
+                               &   POPC &64 & 64 & 50\% & 10314  & 298  & 100 & 80 & [\citenum{bergerFILESpopc50chol}]$^*$ & [\citenum{ferreira13}] \\
+                               &   POPC &50 & 78 & 61\% & 5782   & 298  & 100 & 80 & [\citenum{bergerFILESpopc60chol}]$^*$ & [\citenum{ferreira13}] \\
+CHARMM36\cite{klauda10,lim12}    & POPC   & 128& 0& 0\% & 5120  & 303  & 150 & 100 & [\citenum{charmm36files}]$^*$  & SI  \\
+                                & POPC   & 100 & 24 & 19\%  &  4960   & 303 & 200 & 100 & [\citenum{charmm36files20perCHOL}]$^*$ & SI \\
+                               & POPC   & 80 & 80 &50\%  &  4496    & 303 & 200 & 100 & [\citenum{charmm36files50perCHOL}]$^*$ & SI \\
+MacRog\cite{maciejewski14}     & POPC   & 128 & 0 & 0\% & 6400  & 310 & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI \\ 
+                          & POPC   & 114  & 14 & 11\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
+                          & POPC   & 72   & 56 &  44\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
+                             & POPC   & 64  & 64 & 50\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
+                             & POPC   & 56   & 72 & 56\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
 \end{tabular}
-\end{table*}
-
-%The initial structures for the simulations in low hydrated conditions are made by removing the
-%water molecules from the fully hydrated system to achieve the targeted hydration conditions.
-
-%For the initial structure of CHARMM36-~\cite{klauda10,lim12} and MacRog-~\cite{maciejewski14,kulig14} simulations with cholesterol,
-%the required amount of lipid molecules were replaced with cholesterol
+\end{sidewaystable}
 
@@
 \subsection{Full hydration: Experimental order parameters for the glycerol backbone and headgroup}\label{experiments}
-{\bf The specific deuteration of $\alpha$-, $\beta$- and g$_3$- segments of DPPC has been successful, 
+The specific deuteration of $\alpha$-, $\beta$- and g$_3$- segments of DPPC has been successful, 
 allowing the absolute value order parameter measurements for these segments by $^2$H--NMR~\cite{gally75,brown77,brown78,akutsu81}.
@@
 have been measured with several different implementations of $^{13}$C NMR experiments. In addition, the signs of order parameters in some systems
-are measured with $^{13}$C NMR tehcniques~\cite{hong95a,hong95b,gross97}. The experimental values of glycerol backbone 
+are measured with $^{13}$C NMR techniques~\cite{hong95a,hong95b,gross97}. The experimental values of glycerol backbone 
 and choline order parameters from various publications~\cite{gally75,akutsu81,gross97,dvinskikh05a,ferreira13} with the signs measured in~\cite{hong95a,hong95b,gross97} are shown in Fig.~\ref{HGorderparameters}.
-}
 \begin{figure}[]
-  \centering
+%  \centering
   \includegraphics[width=8.6cm]{../DATAreportediINblog/comparisonSorted.pdf}
-\todo{Markus Miettinen suggested that we should consider making one more figure where only experimental data would be shown and that would be discussed in Section~\ref{experiments}. \\
-COMMENT: I do not think that we need such figure anymore since this figure is more clear now.} \\
-\todo{I think that this figure is now very good. There is also the interactive version by Hubert Santuz now in https://plot.ly/~HubertSantuz/72/lipid-force-field-comparison/
+%\todo{Markus Miettinen suggested that we should consider making one more figure where only experimental data would be shown and that would be discussed in Section~\ref{experiments}. \\
+%COMMENT: I do not think that we need such figure anymore since this figure is more clear now.} \\
+\newline
+\todo{There is also the interactive version by Hubert Santuz now in https://plot.ly/~HubertSantuz/72/lipid-force-field-comparison/
 we should figure out which is the most practical way to put that behind permalink once it finalized (Zenodo, figshare or something else?) and then put a citation in the paper.}
@@
   Order parameteres from simulations listed in Table~\ref{systems} and experiments for glycerol and choline groups.
-  The experimental values were taken from the following publications: DMPC 303~K from \cite{gross97}, DMPC 314~K from \cite{dvinskikh05a}, DPPC 310K from \cite{gally75}, 
-  DPPC 323~K from \cite{akutsu81}, and POPC 298~K \cite{ferreira13}.
+The experimental values were taken from the following publications:
+ DMPC 303~K from \cite{gross97},
+ DMPC 314~K from \cite{dvinskikh05a},
+ DPPC 322~K from \cite{gally75},
+ DPPC 323~K from \cite{akutsu81},
+ POPC 296~K from \cite{bechinger91}, and
+ POPC 300~K from \cite{ferreira13}.
   The vertical bars shown for some of the computational values are not error bars, but demonstrate that for 
-  these systems we had two data sets; the ends of the bars mark the values from the two sets, and the dot marks their measurement-time-weighted average. 
-  The H{\"o}gberg et al. force field in Table~\ref{systems} is abbreviated in this figure as HoNiLy, Tj{\"o}rnhammar et al. as TjoEdh, CHARMM36-UA as C36UA,
-  Rabinovic et al. as RabLyu and Ulmschneider and Ulmschneider as UlmUlm.
+these systems we had at least two data sets (see Table~\ref{systems});
+the ends of the bars mark the extreme values from the sets, and the dot marks their measurement-time-weighted average. 
 } 
@@
 accuracy and therefore we exclude them from our discussion. 
+For more details, see \url{http://nmrlipids.blogspot.fi/2014/02/accuracy-of-order-parameter-measurements.html}
+\todo{Citation to the blog will be added when available.}
 Motivated by the high experimental repeatability, we have highlighted in 
@@
 
-{\bf In addition to the numerical values, an important feature of the glycerol backbone is the 
-forking (see section~\ref{expORDp}) of the order parameters in g$_1$ and g$_3$ segments, in contrast to the choline segments $\alpha$ and $\beta$. }
+In addition to the numerical values, an important feature of the glycerol backbone is the 
+forking (see section~\ref{expORDp}) of the order parameters in g$_1$ and g$_3$ segments, in contrast to the choline segments $\alpha$ and $\beta$. 
 The forking in glycerol backbone g$_3$ segment is small ($\approx$ 0.02) 
@@
 
-{\bf By combining the absolute value measurements~\cite{ferreira13}, the sign measurements~\cite{hong95a,hong95b,gross97} and the stereospecific
-measurements~\cite{gally81} the most detailed experimentally available order parameter information for the glycerol backbone and choline 
-segments of POPC bilayer is collected and the data are shown in Fig.~\ref{HGorderparameters2}. }
+The most detailed experimentally available order parameter information for the glycerol backbone and choline 
+segments of POPC bilayer is collected by taking the absolute values from~\cite{ferreira13}, the signs from~\cite{hong95a,hong95b,gross97} 
+and the stereospecific labeling from~\cite{gally81}, and shown in Fig.~\ref{HGorderparameters2}.
 \begin{figure*}[]
@@
   \includegraphics[width=8.6cm]{HGorderparameters5.eps} \\
-  \todo{Markus Miettinen suggested that we should consider ``spreading'' the values similarly to the figure~\ref{HGorderparameters}} \\
+  \todo{This figure should be updated similar to Fig.~\ref{HGorderparameters}} \\
   \caption{\label{HGorderparameters2}
@@
 The results for each force field in respect to the above criteria are summarized in Figure~\ref{FullHydrationComparisonTable}.
-\todo{DONE}.
 \begin{figure}[]
@@
   \includegraphics[width=8.6cm]{../DATAreportediINblog/comparisonTable.pdf}
+  \newline
   \caption{\label{FullHydrationComparisonTable}
@@
 
-{\bf
 None of the studied force fields fulfils these criteria completely, however CHARMM36 is pretty close. 
@@
 (CHARMM36, MacRog, GAFFlipid), and the Berger based models are the most used lipid model in the literature.
-}
-%The top three models (CHARMM36, MacRog and GAFFlipid) together with the most used lipid model (Berger based model) 
-%were subjected to a more careful comparison including the signs and the sterospecific labeling in Fig.~\ref{HGorderparameters2}.
-%The essential additional information given by this comparison is that the sign of the $\beta$ carbon order parameter is correct only in CHARMM36 model.
 
@@
 for more discussion.)
-\todo{This is temporarily linked to the blog. Currently I think that the posts (with comments included) to which we want to cite are put into 
-the Figshare in pdf format as they are in when the publication is submitted. Figshare allows commenting to continue the discussion.}
+\todo{Citation to the blog will be added when available.}
 However, as already discussed above, it is not simple to deduce 
@@
 since the absolute value decreases~\cite{dvinskikh05b,ulrich94,bechinger91}.
-Slight decrease for the glycerol backbone g$_3$- and g$_2$- order parameters were observed with dehydration, 
-while g$_1$ remained practically unchanged~\cite{dvinskikh05b}.
+Slight decrease for the glycerol backbone order parameters were observed with dehydration~\cite{dvinskikh05b}.
 \begin{figure}[]
@@
   \includegraphics[width=8.6cm]{../DATAreportediINblog/dehydration.pdf}
-\todo{Figure has been modified to elucidate changes. The text should be updated accordingly. The same applies to the cholesterol figure.}
+  \todo{It should be checked that we have the up-to-date GAFF results in this plot.} \\
   \caption{\label{ordPhydr}
@@
     The signs are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
-    and Gross et al.~\cite{gross97}.
+    and Gross et al.~\cite{gross97}. To elucidate changes as a function of hydration level, the 
+    simulation results are shifted to match better with experiments in the full hydration.
+    The magnitude of shifting is written in the y-axis panel.
   }
@@
 order parameters (except by Mashl et al.~\cite{mashl01}).
-In Fig.~\ref{ordPhydr} the glycerol backbone and choline order parameters are shown as a function of hydration level for the CHARMM36, MacRog and GAFFlipid models (having the most realistic atomistic resolution structures) together with the Berger based model
-(which is the most used lipid model). The choline order parameters increase with dehydration in all
+In Fig.~\ref{ordPhydr} the glycerol backbone and choline order parameters are shown as a function of hydration level for the CHARMM36, 
+MacRog and GAFFlipid models (having the most realistic atomistic resolution structures) together with the Berger based model 
+(which is the most used lipid model). 
+To elucidate changes the simulation results are shifted to match better with experiments with full hydration.
+The magnitude of shifting is written in the y-axis label.
+The choline order parameters increase with dehydration in all
 models, in qualitative agreement with experiments. 
 The measured decrease in both g$_3$ and g$_2$ order parameters with dehydration is reproduced only in CHARMM36.
+\todo{This sentence must be rewritten after the up-to-date results for GAFFlipid are in the figure.}
 
@@
 all models as a function of dehydration in Fig.~\ref{PNangle}.
-However, the qualitative agreement in the lipid response to dehydration does not guarantee the correct 
-free energy landscape if the simulation model has an incorrect structure. The influence of this issue to 
-dehydration energetics studied with simulations~\cite{eun09,schneck12} is left for future studies.
+
+It should be noted, however, that free energy landscape is not correct in the models which
+are not able to reproduce the experimental order parameters. Thus the energetical response
+to the dehydration may not be correct even though the order parameter response is qualitatively correct.
+This issue may have some influence on the dehydration energetic calculations made with the Berger model~\cite{eun09,schneck12}.
 \begin{figure}[]
@@
 as CHARMM36 is the only model that reproduces the decrease in g$_2$ and g$_3$ segments.
+\todo{This sentence must be rewritten after the up-to-date results for GAFFlipid are in the figure.}
 
@@
 their entropy and dynamics (``superlattice model'')~\cite{somerharju09}. 
-\todo{It has been suggested that we should use direct quotations from original papers here for clarity.}
+%\todo{It has been suggested that we should use direct quotations from original papers here for clarity.}
 Both of these suggestions have been supported
 by molecular dynamics simulations~\cite{zhu07,alwarawrah12}, and other simulations suggest specific
-interactions between the glycerol backbone and cholesterol~\cite{rog09}, however the glycerol backbone and choline headgroup behaviour
-as a function of cholesterol content is not compared to experiments in these studies.
+interactions between the glycerol backbone and cholesterol~\cite{rog09}. However,
+in the above mentioned studies, the glycerol backbone and choline headgroup behaviour
+as a function of cholesterol content has not been compared to experiments.
 
@@
   \includegraphics[width=8.6cm]{../DATAreportediINblog/cholesterolization.pdf}
+  \todo{The data delivered by Fernando Favela should be added here.} \\
   \caption{\label{ordPchol}
@@
     with the Berger-POPC-07/H\"oltje-CHOL-13, CHARMM36 and MacRog force fields. The signs in the experimental values are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
-    and Gross et al.~\cite{gross97}.  Most order parameters from Berger-POPC-07/H\"oltje-CHOL-13 model for g$_1$ are beoynd the y-axis scale.}
+    and Gross et al.~\cite{gross97}.  Most order parameters from Berger-POPC-07/H\"oltje-CHOL-13 model for g$_1$ are beoynd the y-axis scale.
+    To elucidate changes as a function of the cholesterol concentration, the 
+    simulation results are shifted to match better without cholesterol.   
+    The magnitude of shifting is written in the y-axis panel.
+}
 \end{figure}
@@
 In addition to the experimental data, the previously published simulation results from the Berger-POPC-07/H\"oltje-CHOL-13 model~\cite{ferreira13},
-and our results from CHARMM36 and MacRog force fields  
-are shown in Fig.~\ref{ordPchol}. As already pointed out previosly, the Berger-POPC-07/H\"oltje-CHOL-13 model
+and our results from CHARMM36 and MacRog force fields are shown in Fig.~\ref{ordPchol}. 
+To elucidate changes the simulation results are shifted to match better with experiments without cholesterol.
+The magnitude of shifting is written in the y-axis label.
+As already pointed out previosly, the Berger-POPC-07/H\"oltje-CHOL-13 model
 seriously overestimates the effect of cholesterol on the phospholipid glycerol backbone and choline segments~\cite{ferreira13}.
-In contrast, the responses of both CHARMM36 and MacRog are in better agreement
-with experiments. However, CHARMM36 seems to better reproduce the experimentally observed modest changes in the glycerol backbone segments 
-g$_2$ and g$_3$ with high concentrations of cholesterol. Thus we have calculated the glycerol backbone dihedral angle distributions
+In contrast, the responses of both CHARMM36 and MacRog are in better agreement with experiments. 
+%However, CHARMM36 seems to better reproduce the experimentally observed modest changes in the glycerol backbone segments 
+%g$_2$ and g$_3$ with high concentrations of cholesterol. 
+Thus we have calculated the glycerol backbone dihedral angle distributions
 as a function of cholesterol in CHARMM36 (shown in Supplementary material) to resolve the cholesterol induced structural changes. The only detectable change due to the
@@
 
-- None of the tested models (13 different models) produces the order parameters with the experimental
-accuracy for a fully hydrated phoshatidylcholine lipid bilayer. However, the CHARMM36, GAFFlipid and MacRog 
-models are relatively close. The structures of these models together with the most used lipid model (Berger based) 
-were subjected to more careful studies. The results revealed that the current models are not accurate
-enough to resolve the atomistic resolution structures sampled by glycerol backbone and choline headgroup.  
-However, the correlation between dihedral angle distributions and order parameter differences was found, 
-suggesting that careful adjustment of dihedral potentials would potentially lead to the model with a correct
-structure.
-
-- Independent of the accuracy for a fully hydrated lipid bilayer, all the models reproduced the choline response
-to dehydration. This can be explained by the change in the P--N vector tilting more parallel to the membrane,
-which leads to the increase of order parameters despite of the initial configuration. It should be noted, however,
-that the correct qualitative response does not necessarily indicate correct energetics.
-
-- The response of glycerol backbone and choline headgroup to the cholesterol content is described more
-realistically in CHARMM36 and MacRog models than in the Berger based model.
-
-- {\bf Besides the main conclusions, we have created the most extensive
+- Comparison of order parameters between experiments and different atomistic resolution models 
+together with the structural analysis shows that these parameters can be used
+to judge the structural accuracy of the model. Thus the combination of atomistic resolution 
+molecular dynamics simulations and NMR experiments can be used to resolve the atomistic resolution
+structures of biomolecules in biologically relevant conditions. 
+This approach can be extended from lipids also, for example, to membrane proteins.
+
+%None of the tested models (13 different models) produces the order parameters with the experimental
+%accuracy for a fully hydrated phoshatidylcholine lipid bilayer. However, the CHARMM36, GAFFlipid and MacRog 
+%models are relatively close. The structures of these models together with the most used lipid model (Berger based) 
+%were subjected to more careful studies. The results revealed that the current models are not accurate
+%enough to resolve the atomistic resolution structures sampled by glycerol backbone and choline headgroup.  
+%However, the correlation between dihedral angle distributions and order parameter differences was found, 
+%suggesting that careful adjustment of dihedral potentials would potentially lead to the model with a correct
+%structure.
+
+- The review of previous experimental results revealed that the 
+choline order parameters are increasing when the bilayer is dehydrated. Our simulations show that
+this can be explained by P--N vector tilting more parallel to the membrane. 
+These results strongly supports and complements the idea that the charge induced choline tilting and 
+partition can be measured by using the order parameter changes~\cite{ionpaper,scherer89}.
+
+%Independent of the accuracy for a fully hydrated lipid bilayer, all the models reproduced the choline response
+%to dehydration. This can be explained by the change in the P--N vector tilting more parallel to the membrane,
+%which leads to the increase of order parameters despite of the initial configuration. It should be noted, however,
+%that the correct qualitative response does not necessarily indicate correct energetics.
+
+- Only modest changes of glycerol backbone and choline order parameters are observed experimentally with increasing cholesterol concentration.
+When interpreted by using the simulations with the most realistic response to the cholesterol concentration,
+this observation can be explained such that the cholesterol induces only minor changes in the g$_3$-g$_2$-g$_1$-O(sn-1)
+dihedral in the glycerol backbone.
+
+%- The response of glycerol backbone and choline headgroup to the cholesterol content is described more
+%realistically in CHARMM36 and MacRog models than in the Berger based model.
+
+- Besides the main conclusions, we have created the most extensive
 publicly available collection of molecular dynamics simulation trajectories of lipid bilayers
@@
 In general, we conclude that to fully utilize the potential of atomistic resolution classical molecular dynamics simulations
-in the structural interpretation of the high resolution NMR data~\cite{ferreira14} for lipid bialyers one has to  
+in the structural interpretation of the high resolution NMR data~\cite{ferreira14} for lipid bilayers one has to  
 improve the phoshatidylcholine glycerol backbone and choline headgroup parameters.
@@
 platform. All the scientific contributions have been done through the blog and are publicly
-available. }
+available. \todo{Citation to the blog will be added when available.}
 
@@
 
-{\bf Acknowledgements: }
-OHSO acknowledges Tiago Ferreira and Paavo Kinnunen for useful discussions, the Emil Aaltonen foundation for financial support, Aalto Science-IT project and CSC-IT Center for Science for computational resources.
+%\listoftodos
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% The "Acknowledgement" section can be given in all manuscript
+%% classes.  This should be given within the "acknowledgement"
+%% environment, which will make the correct section or running title.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{acknowledgement}
+
+\todo{All authors, please add your acknowledgements.}
+
+OHSO acknowledges Tiago Ferreira and Paavo Kinnunen for useful discussions, 
+the Emil Aaltonen foundation for financial support, 
+Aalto Science-IT project and CSC-IT Center for Science for computational resources.
 
 MSM acknowledges  financial support from the Volkswagen Foundation (86110).
-\bibliographystyle{apsrev}
+\end{acknowledgement}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% The same is true for Supporting Information, which should use the
+%% suppinfo environment.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{suppinfo}
+
+Simulation details, one figure and author contributions.
+
+%This will usually read something like: ``Experimental procedures and
+%characterization data for all new compounds. The class will
+%automatically add a sentence pointing to the information on-line:
+
+\end{suppinfo}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% The appropriate \bibliography command should be placed here.
+%% Notice that the class file automatically sets \bibliographystyle
+%% and also names the section correctly.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \bibliography{refs}
@@
 \newpage
-
-\appendix
 \begin{center}
@@
 \subsection{Simulation details} 
-%\todo{When I was adding the simulation details to the manuscript, I realized that the section was getting very long and
-%it would do the paper more difficult to read. I suggest that we do this in the following way: \\
-%- For the simulation data which are taken directly from previous publications, we only cite the original publication
-%in the tables where the systems are described. This would be the case, for example, for Berger POPC/chol, DMPC, Rabinovich et al. and Hogberg et al. \\
-%- For the simulations which are reproduced using mdp file already delivered by original authors we can cite that, thus we do not have to copy the 
-%simulation detail section from the original publication. This is the case, for example, for Tjornhammar et al. \\
-%- When simulations are done specifically for this project and there is something non-trivial we
-%put the details in the supplementary material. This is the case, for example, for GAFFlipid, Lipid14 and poger et al. \\
-%- In all cases all simulation files, also trajectory, would be ideally shared in Zenodo.
-%}
 \subsubsection{Berger based models}
@@
 
-%The results for the POPC/cholesterol systems with the Berger lipid~\cite{berger97} and modified H\"oltje
-%cholesterol model~\cite{holtje01} were taken directly from Ref.~\cite{ferreira13}.
-
 \subsubsection{CHARMM36}
@@
 {\it DPPC}
-\todo{Markus Miettinen made the files.}
+\todo{Markus Miettinen, please describe how you made the files.}
 
@@
 by Piggot et al. \cite{piggot12} were used. Cholesterol parameters came
-from Lim et al. \cite{lim12} and were converted into GROMACS format with the PyTopol tool (\url{https://github.com/resal81/PyTopol}).  
+from Lim et al. \cite{lim12} and were converted into GROMACS format with the PyTopol tool~\cite{salari15}.  
 Single point energy calculation was done to assess the conversion. 
@@
 The lipid force field parameters were obtained from the developers and they correspond to the published DPPC parameters~\cite{maciejewski14} with the inclusion of the 
-double bond parameters. This inclusion of unsaturated lipid tails will be published in the near future. 
-\todo{You have recent paper where you use POPC/chol simulations with this model: Kulig et al. BBA 1848 (2015) 422-432 http://dx.doi.org/10.1016/j.bbamem.2014.10.032.
-would this be a correct refence here?}
+double bond parameters~\cite{kulig15}. 
+%This inclusion of unsaturated lipid tails will be published in the near future. 
+%\todo{You have recent paper where you use POPC/chol simulations with this model: Kulig et al. BBA 1848 (2015) 422-432 http://dx.doi.org/10.1016/j.bbamem.2014.10.032.
+%would this be a correct refence here?}
 A bilayer with 288 POPC lipids was hydrated with 12600 TIP3P water~\cite{jorgensen83} molecules ($\sim$44/lipid) and simulated for 100~ns with a time step of 2~fs. Data was saved 
@@
 
-{\bf It should be noted that the area per molecule with these settings for the GAFFlipid model was 61.6~\AA$^2$,
+It should be noted that the area per molecule with these settings for the GAFFlipid model was 61.6~\AA$^2$,
 while the original publication reported 63.9~\AA$^2$~\cite{dickson12}. However, the same parameters and Amber to Gromacs
-conversion reproduced the area per molecule from original publication for the lipid14 model (see next section).}
+conversion reproduced the area per molecule from original publication for the lipid14 model (see next section).
 
 \subsubsection{Lipid14}
-%The used simulation files and part of the trajectroy are available~\cite{lipid14files}.
 The initial structure was taken directly from the Lipidbook~\cite{domanski10}.
@@
 
-{\bf The area per molecule with these settings was 65.4~\AA$^2$ which is in agreement with the value reported in the original publciation 65.6$\pm$0.5~\AA$^2$~\cite{dickson14}.}
+The area per molecule with these settings was 65.4~\AA$^2$ which is in agreement with the value reported in the original publication 65.6$\pm$0.5~\AA$^2$~\cite{dickson14}.
 
@@
 also in the simulations with PME.
-%Since there is no dispersion correction in GROMOS (and since dispersion correction should not be used anyway for systems with interfaces), we decided to stick 
-%to the same LJ cutoff as that with the RF scheme i), that is 1.4 nm.
 
@@
 introduced in versions 4.5.X. A fix has been introduced in version 4.6.6 that allows a recovery of an area per lipid value of 0.615~nm$^2$. The results in terms of area per lipid using the different 
-GROMACS versions are available at~\cite{pogerFILESrf2}\todoi{DONE}.
-% http://www.dsimb.inserm.fr/~fuchs/project\_Samuli/Poger\_DPPC/tests\_gmx\_versions/)[Samuli, you may want to add this Figure+Legend+Comment 
-%o a stable link (Figshare?), let me know]. 
-%Of note, some tests (Miguel Machuqueiro, personnal communication) with the same RF scheme i) on simulations with proteins 
-%(thus with explicit hydrogens) showed that this setup is still unstable and eventually leads to crashes with versions 4.6.6 and 4.6.7 (and so does version 4.5.*).
+GROMACS versions are available at~\cite{pogerFILESrf2}.
 Thus we decided to use only the PME setup ii) for computing the order parameter since it gives stable results regardless of the GROMACS version. We obtained an area per 
@@
 \subsubsection{Slipids}
-%The used simulation files and part of the trajectroy are available~\cite{slipidsFILES}.
-
 Initial coordinates for a hydrated DPPC (at 323~K) and POPC (at 310~K) bilayers (30 and 40 waters/lipid, respectively) were taken directly from the 
@@
 10 steps. This simulation protocol corresponds to the protocol used in Ref~\cite{jambeck13}.
-%The simulations were run for XXX ns with the last XXX ns used for calculating the order parameters.
 
@@
 
-%Additionally, the reaction field electrostatics with a twin-range cut-off scheme [cite GROMOS] was employed yet the results 
-%were consistent with those obtained with PME. (This paragraph can be discarded.)
-
 \subsubsection{Chiu et al.}
@@
 Pressure was semi-isotropically coupled to the atmospheric pressure with the Parrinello--Rahman method~\cite{parrinello81}.
-%\subsubsection{H\"ogberg et al. and Rabonovich et al.}
-%Simulations done in \cite{hogberg08} and \cite{rabinovich14} are used.
-
-%Time step 2fs, Ewald summation,
-%cut-off 14 Å, update of neighbour list each 10 steps,
-%Nose-Hoover thermostat,  Parrinello-Rahman barostat,
-%semi-anizotropic pressure coupling,
-%long-range isotropic pressure correction
+
 \subsubsection{Ulmschneider}
@@
 \subsubsection{Tj\"ornhammar et al.}
-%The used simulation files and part of the trajectroy are available~\cite{tjornhammarfiles}.
-
 The gel phase DPPC bilayer structure delivered by Tj\"ornhammar  and Edholm~\cite{tjornhammar14} was ran for 5~ns at 343~K in order to destroy the 
@@
 
-%\subsubsection{OPLS-AA}
-%\todo{Joona Tynkkynen.
-We have this only with 150mM of NaCl delivered by Joona Tynkkynen.
-Options are to remove these results from this publicatio, run the simulations without ions (if not yet available),
-%  or include the results with ions. In my understanding this is a proto version of MacRog so it could be left out as well. However, for historical reasons and 
-%to understand literature it might be useful to include also these results. COMMENT: Matti Javanainen suggested that we would remove these results. If there are no objections and the
-%data without ions will not be delivered, we will remove this.}
-
 \subsection{Dihedral angle distributions as a function of cholesterol in CHARMM36}
@@
 
-\subsection{General issues to think of about the manuscript}
-\todo{DONE}\\
-\todo{Maybe we should mention that CHARMM is significanlty more computationally expensive than other all-atom models.
-This seems evident for a person who has ran CHARMM and other all-atom models with the same computer. However, there
-may be some people who have never used any other model than CHARMM. In this case it may be difficult to realize.
-Would anyone have a good bencmark about this? I have assumed that the reason for the slowness are the extra Lennart-Jones
-interactionsi in water, however, I have never looked into it. Would someone have some hard facts on this?}
+\subsection{Author Contributions}
+\todo{Please write a text summarizing your contribution} \\
+{\it Alexandru Botan} \\
+{\it Fernando Favela}  \\
+{\it Matti Javanainen} \\
+{\it Waldemar Kulig}  \\
+{\it Antti Lamberg}  \\
+{\it Markus S. Miettinen}  \\
+{\it Luca Monticelli}  \\
+{\it Jukka Maatta}  \\
+{\it O. H. Samuli Ollila} Designed and managed the work. Ran and analyzed several simulations. Wrote the manuscript.  \\
+{\it Marius Retegan}  \\
+{\it Tomasz Rog}  \\
+{\it Hubert Santuz}  \\
+{\it Joona Tynkkynen}  \\
+
+%\subsection{General issues to think of about the manuscript}
+%\todo{Maybe we should mention that CHARMM is significanlty more computationally expensive than other all-atom models.
+%This seems evident for a person who has ran CHARMM and other all-atom models with the same computer. However, there
+%may be some people who have never used any other model than CHARMM. In this case it may be difficult to realize.
+%Would anyone have a good bencmark about this? I have assumed that the reason for the slowness are the extra Lennart-Jones
+%interactionsi in water, however, I have never looked into it. Would someone have some hard facts on this?}
 
 %\newpage
-
-\onecolumngrid
-
- \listoftodos
-
 \end{document}

### 2015-03-17 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
update HGmodel manuscript in ACS format

@@
 %\affiliation{University of East Anglia, Norwich, United Kingdom}
-\author{Fernando Favela}
-\affiliation[Mexico]{Mexico}
+\author{Fernando Favela-Rosales}
+\affiliation[Mexico]{Departamento de F{\'i}sica - CINVESTAV}
 \author{Patrick Fuchs}
@@
 {\it Fernando Favela}  \\
+{\it Patrick Fuchs} \\
 {\it Matti Javanainen} \\

### 2015-03-18 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
update HGmodel manuscript in ACS format

@@
 \affiliation[Tampere University of Technology]
-{Tampere University of Technology, Tampere, Finland}
+{Department of Physics, Tampere University of Technology, Tampere, Finland}
 \author{Antti Lamberg}
@@
 the Emil Aaltonen foundation for financial support, 
-Aalto Science-IT project and CSC-IT Center for Science for computational resources.
+Aalto Science -- IT project and CSC -- IT Center for Science for computational resources.
 
 MSM acknowledges financial support from the Volkswagen Foundation (86110).
+
+WK acknowledges CSC -- IT Centre for Science (Espoo, Finland) is acknowledged for excellent computational resources (project number tty3995).
+
 \end{acknowledgement}
@@
 {\it Matti Javanainen} \\
-{\it Waldemar Kulig}  \\
+{\it Waldemar Kulig} prepared the MD simulations with cholesterol for the MacRog FF \\
 {\it Antti Lamberg}  \\

### 2015-03-18 HGmodelMANUSCRIPT/coverLETTER.tex
add cover letter

\documentclass[12pt]{letter}
\usepackage{graphicx,baskervald,microtype}
\usepackage{hyperref,amsmath,xcolor}
\usepackage{pgfplots,marginnote}
\usepackage[top=0.5in,bottom=0.5in,left=1in,right=1in]{geometry}
%\newgeometry{margin=2.5cm}
\newcommand{\omamargin}[1]{\marginnote{\textbf{#1}}[7pt]}
\begin{document}
\reversemarginpar
\pagestyle{empty}
\noindent O. H. Samuli Ollila \\
\noindent Helsinki Biophysics and Biomembrane Group\\
\noindent Department of Biomedical Engineering and Computational Science\\
\noindent P.O. Box 12200, FI-00076 Aalto University, Espoo, Finland\\
\noindent samuli.ollila@aalto.fi, Tel. +358503746963, Fax. +358 9 470 23182 \\

%THESE ARE COPYPASTED FROM THE JACS AUTHOR GUIDELINES AND I HAVE TRIED TO ANWER THESE
%
%Articles of high scientific quality, originality, significance, and conceptual novelty that
%are of interest to the wide and diverse contemporary readership
%of JACS will be given priority for publication.
%
%it is required that (a) the level of theory and methodology employed must be adequate for the
%problem at hand, and (b) the manuscript must provide significant chemical insights or have
%substantial predictive value.

To Whom It May Concern,

Please find attached to this message a manuscript that we would like to submit for publication 
in \textit{Journal of the american chemical society} (JACS): 
`Towards atomistic resolution structure of phosphatidylcholine glycerol backbone and choline headgroup at different ambient conditions', 
by A. Botan et al. In the manuscript we combine extensive amount of experimental Nuclear Magnetic Resonance (NMR)
and molecular dynamics simulation data to understand the atomistic resolution structure of biologically abundant phospholipid molecules.

We believe that the manuscript is interesting for the wide and diverse contemporary audience of
JACS since the used combination of experiments and simulations is shown to be successfull in
resolving the atomistic resolution structure of biomolecules in the various biologically relevant conditions.
The approach is demonstrated for lipids, however, the extension to, e.g. membrane proteins is straightforward. 
The understanding of the atomistic resolution structures of molecules in biologically relevant conditions 
significantly advances understanding about fundamental chemistry and also supports the development of, 
for example, new biomaterials or drug delivery systems.

The work has been conducted by using a novel open collaboration concept, and all the scientific contributions
are done publicly through the nmrlipids.blogspot.fi. Thus all the scientific content related to this work
(including raw data, discussions, manuscript drafts, etc.)
has been publicly available during the whole course of the project through the blog. This material and discussion possibility
will be kept available also after the publication. This approach has significantly increased and will increase the scientific
quality and impact of the work. In addition, we have shared all the simulation raw data in such a format
that it can be reanalyzed easily also for other purposes. In addition to the scientic content,
we believe that the success of this approach in the field of chemistry is highly interesting 
for the JACS audience.

The usage of molecular dynamics simulations and open collaboration are adequate for the
problem at hand since the simulations are currently the most straightforward method to 
construct the atomistic resolution structure of molecules to interpret the NMR experiments. On the other hand,
enormous amount of simulation work was required for this work, and the only practical
solution to collect this data is the used open collaboration approach.
As result, we have produced original chemical insight for the atomistic resolution 
structures of phospholipids in various biologically relevant conditions.

We hope that you can consider this manuscript to be published as an article in the Journal of the american chemical society. \\

Sincerely yours,

On behalf of the authors,

O. H. Samuli Ollila

\end{document}

### 2015-03-19 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Contributions & Acknowledgements of Matti Javanainen and Joona Tynkkynen

@@
 \affiliation[Tampere University of Technology]
-{Tampere University of Technology}
+{Department of Physics, Tampere University of Technology, Tampere, Finland}
 \author{Waldemar Kulig}
@@
 \affiliation[Tampere University of Technology]
-{Tampere University of Technology, Tampere, Finland}
+{Department of Physics, Tampere University of Technology, Tampere, Finland}
 \author{Hubert Santuz}
@@
 \affiliation[Tampere University of Technology]
-{Tampere University of Technology, Tampere, Finland}
+{Department of Physics, Tampere University of Technology, Tampere, Finland}
 
@@
 Poger \cite{poger10}             & DPPC  & 128 & 5841 & 323 & 2$\times$100 & 2$\times$50 & [\citenum{pogerFILESpme1,pogerFILESpme2}]$^*$ & SI \\
-Slipid \cite{jambeck12}          & DPPC & 128 & 3840 & 323 & 150 & 100 & [\citenum{slipidsFILES}]$^*$ & SI \\
-Slipid \cite{jambeck12b}          & POPC & 128 & 5120 & 303 & 200 & 150 & [\citenum{slipidsFILESpopc}]$^*$ & SI \\
+Slipids \cite{jambeck12}          & DPPC & 128 & 3840 & 323 & 150 & 100 & [\citenum{slipidsFILES}]$^*$ & SI \\
+Slipids \cite{jambeck12b}          & POPC & 128 & 5120 & 303 & 200 & 150 & [\citenum{slipidsFILESpopc}]$^*$ & SI \\
 Kukol \cite{kukol09}          & POPC   & 512 & 20564 & 298 & 50 & 30  & [\citenum{kukolFILES}]$^*$ & SI \\
@@
 these models are using CHARMM bonded potentials for glycerol backbone and choline. The fourth and the fifth models in the list, MacRog~\cite{maciejewski14} and
-GAFFlipid~\cite{dickson12}, have independently determined dihedral potentials. All the models based on Gromos potentials and Slipid perform less well.
+GAFFlipid~\cite{dickson12}, have independently determined dihedral potentials. All the models based on Gromos potentials and Slipids perform less well.
 In the present work we subject the CHARMM36, MacRog, GAFFlipid and Berger-POPC-07 to a more careful comparison including the sterospecific labeling  
@@
 
+MJ and JT acknowledge and CSC - IT Center for Science for computational resources (project number tty3979).
+MJ also acknowledges the Finnish Doctoral Programme in Computational Sciences (FICS) for funding.
+
 \end{acknowledgement}
@@
 {\it Patrick Fuchs} \\
-{\it Matti Javanainen} \\
+{\it Matti Javanainen} prepared and performed simulations with multiple lipid models and analyzed the results. Supervised the work of JT.\\
 {\it Waldemar Kulig} prepared the MD simulations with cholesterol for the MacRog FF \\
@@
 {\it Hubert Santuz}  \\
-{\it Joona Tynkkynen}  \\
+{\it Joona Tynkkynen} prepared and performed the dehydration simulations with the MacRog FF.\\
 

### 2015-03-20 FFactor/FFcomp.sh
move form factor files to new repository

@@ -1,138 +1,0 @@
-#GAFFlipids
-rm popcRUN4.tpr
-wget  https://zenodo.org/record/12654/files/popcRUN4.tpr
-wget  https://zenodo.org/record/12652/files/popcRUN4.trr.gz
-wget https://raw.githubusercontent.com/NMRLipids/nmrlipids.blogspot.fi/master/POPCgaff/electronsGAFF.dat
-gunzip popcRUN4.trr.gz
-cp popcRUN4.tpr ANALtpr.tpr
-echo non-Water System | trjconv -f popcRUN4.trr -s ANALtpr.tpr -fit translation -o ANALtraj.xtc
-cp electronsGAFF.dat electrons.dat
-echo 0 | g_density -f ANALtraj.xtc -s ANALtpr.tpr -ei electrons.dat -dens electron -o electronDENSITY.xvg -xvg none -sl 100
-sh FFstructCALC.sh
-mv tmp4.dat electronDENSITYgaff.dat
-mv FFactor.dat FFactorGAFF.dat
-#LIPID14
-wget https://zenodo.org/record/12767/files/popcRUN2.trr
-rm popcRUN2.tpr
-wget https://zenodo.org/record/12767/files/popcRUN2.tpr
-wget https://raw.githubusercontent.com/NMRLipids/nmrlipids.blogspot.fi/master/POPClipid14/electronsLIPID14.dat
-cp popcRUN2.tpr ANALtpr.tpr
-echo non-Water System | trjconv -f popcRUN2.trr -s ANALtpr.tpr -fit translation -o ANALtraj.xtc
-cp electronsLIPID14.dat electrons.dat
-echo 0 | g_density -f ANALtraj.xtc -s ANALtpr.tpr -ei electrons.dat -dens electron -o electronDENSITY.xvg -xvg none -sl 100
-sh FFstructCALC.sh
-mv tmp4.dat electronDENSITYlipid14.dat
-mv FFactor.dat FFactorLIPID14.dat
-#BERGER
-#POPC
-wget https://zenodo.org/record/13279/files/popc0-25ns.trr
-wget https://zenodo.org/record/13279/files/popc25-50ns.trr
-rm popc407.tpr
-wget https://zenodo.org/record/13279/files/popc407.tpr
-wget https://raw.githubusercontent.com/NMRLipids/nmrlipids.blogspot.fi/master/POPCberger/electronsBERGER.dat
[…]

### 2015-03-20 FFactor/FFstructCALC.sh

@@ -1,10 +1,0 @@
-transz=$(cat electronDENSITY.xvg | awk 'BEGIN{min=1000;}{if($2<min){min=$2; minx=$1;}}END{print minx}')
-boxz=$(tail -n 1 electronDENSITY.xvg | awk '{print $1}')
-slice=$(cat electronDENSITY.xvg | awk '{if(NR==2) print $1}')
-cat electronDENSITY.xvg | awk -v transz=$transz '{print $1-transz" "$2}' > tmp.dat
-cat tmp.dat | awk -v boxz=$boxz -v slice=$slice '{if($1<(-1)*boxz/2){print $1+boxz+slice" "$2}}' > tmp1.dat
-cat tmp.dat | awk -v boxz=$boxz '{if($1>(-1)*boxz/2 && $1<boxz/2){print $1" "$2}}' > tmp2.dat
-cat tmp.dat | awk -v boxz=$boxz '{if($1>boxz/2){print $1-boxz" "$2}}' > tmp3.dat
-cat tmp3.dat tmp2.dat tmp1.dat > tmp4.dat
-bulkDENS=$(tail -n 1 tmp4.dat | awk '{print $2}')
-cat tmp4.dat | awk -v slice=$slice -v bulkDENS=$bulkDENS 'BEGIN{scale=0.01;}{for(q=0;q<1000;q=q+1){F[q]=F[q]+($2-bulkDENS)*cos(scale*q*$1)*slice;}}END{for(q=0;q<1000;q=q+1){print 0.1*q*scale" "0.01*sqrt(F[q]*F[q])}}' > FFactor.dat

### 2015-03-22 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
add Acknowledgements
Contributions of Marius Retegan

@@
 MJ also acknowledges the Finnish Doctoral Programme in Computational Sciences (FICS) for funding.
-
+JT, MJ, TR and WK acknowledge the funding from the Academy of Finland (Centre of Excellence program) 
+and the European Research Council (Advanced Grant project CROWDED-PRO-LIPIDS).
 \end{acknowledgement}
@@
 {\it O. H. Samuli Ollila} Designed and managed the work. Ran and analyzed several simulations. Wrote the manuscript.  \\
-{\it Marius Retegan}  \\
+{\it Marius Retegan} Prepared and validated the GROMACS compatible parameter files for GAFFlipid and Lipid14 force fields.\\
 {\it Tomasz Rog}  \\

### 2015-03-24 HGmodelMANUSCRIPT/OLDversions/HGmodel_draft3.tex
move old versions to own folder

--- 2015-03-12 HGmodelMANUSCRIPT/HGmodel_draft3.tex
@@
   \includegraphics[width=8.6cm]{../DATAreportediINblog/comparisonSorted.pdf}
-\todo{Markus Miettinen suggested that we should consider making one more figure where only experimental data would be shown and that would be discussed in Section~\ref{experiments}. \\
+\todo{Markus Miettinen suggested that we should consider making one more figure where only experimental data would be shown and that would be discussed in Section~\ref{experiments}.
 COMMENT: I do not think that we need such figure anymore since this figure is more clear now.} \\
@@
   Order parameteres from simulations listed in Table~\ref{systems} and experiments for glycerol and choline groups.
-  The experimental values were taken from the following publications: DMPC 303~K from \cite{gross97}, DMPC 314~K from \cite{dvinskikh05a}, DPPC 310K from \cite{gally75}, 
-  DPPC 323~K from \cite{akutsu81}, and POPC 298~K \cite{ferreira13}.
+  The experimental values were taken from the following publications:
+  DMPC 303~K from \cite{gross97},
+  DMPC 314~K from \cite{dvinskikh05a},
+  DPPC 322~K from \cite{gally75}, 
+  DPPC 323~K from \cite{akutsu81},
+  POPC 296~K from \cite{bechinger91}, and
+  POPC 300~K from \cite{ferreira13}.
   The vertical bars shown for some of the computational values are not error bars, but demonstrate that for 
-  these systems we had two data sets; the ends of the bars mark the values from the two sets, and the dot marks their measurement-time-weighted average. 
-  The H{\"o}gberg et al. force field in Table~\ref{systems} is abbreviated in this figure as HoNiLy, Tj{\"o}rnhammar et al. as TjoEdh, CHARMM36-UA as C36UA,
-  Rabinovic et al. as RabLyu and Ulmschneider and Ulmschneider as UlmUlm.
+  these systems we had at least two data sets (see Table~\ref{systems});
+  the ends of the bars mark the extreme values from the sets, and the dot marks their measurement-time-weighted average. 
 } 

### 2015-03-24 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
update Fuchs information & fix affiliations and citations

@@
 \affiliation[Mexico]{Departamento de F{\'i}sica - CINVESTAV}
-\author{Patrick Fuchs}
-\affiliation[CNRS paris]{Institut Jacques Monod, CNRS, Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
-\alsoaffiliation[Diderot University]
-{Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
+\author{Patrick F. J. Fuchs}
+\affiliation[CNRS Paris]{Institut Jacques Monod, CNRS, Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
+%\alsoaffiliation[Diderot University]
+%{Institut Jacques Monod, CNRS UMR 7592, Universit{\'e} Paris Diderot, Sorbonne Paris Cit{\'e}, Paris, France}
 \author{Matti Javanainen}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliography{refs}
 
@@
 {\it Fernando Favela}  \\
-{\it Patrick Fuchs} \\
+{\it Patrick F. J. Fuchs} Ran and analyzed the Poger simulations. Provided scientific information which significantly advanced the project (signs and forking of order parameters). \\
 {\it Matti Javanainen} prepared and performed simulations with multiple lipid models and analyzed the results. Supervised the work of JT.\\
@@
 %\newpage
+
+\bibliography{refs}
+
 \end{document}

### 2015-03-25 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Update HGmodel_ACStemplate.tex
Clarified the criteria used to shift the data in dehydration.pdf and cholesterolization.pdf in the captions.

@@
 {5IBCP, CNRS UMR 5086, Lyon, France}
-\author{Jukka Maatta}
+\author{Jukka M{\"a}{\"a}tt{\"a}}
 \affiliation[Aalto University]
@@
     The signs are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
-    and Gross et al.~\cite{gross97}. To elucidate changes as a function of hydration level, the 
-    simulation results are shifted to match better with experiments in the full hydration.
-    The magnitude of shifting is written in the y-axis panel.
+    and Gross et al.~\cite{gross97}.
+    Note that to elucidate the relative change as a function of hydration level,
+    the simulation results are shifted such that the (smaller) $S_\mathrm{CH}$
+    matches (within $\pm$0.01) the experimental value at full hydration;
+    the shift magnitudes for each of the force fields are listed ($S_\mathrm{CH}$+shift) in the y-label.
   }
@@
     and Gross et al.~\cite{gross97}.  Most order parameters from Berger-POPC-07/H\"oltje-CHOL-13 model for g$_1$ are beoynd the y-axis scale.
-    To elucidate changes as a function of the cholesterol concentration, the 
-    simulation results are shifted to match better without cholesterol.   
-    The magnitude of shifting is written in the y-axis panel.
+    Note that to elucidate the relative change as a function of cholesterol content,
+    the simulation results are shifted such that the (smaller) $S_\mathrm{CH}$
+    matches (within $\pm$0.01) the experimental value without cholesterol;
+    the shift magnitudes for each of the force fields are listed ($S_\mathrm{CH}$+shift) in the y-label.
 }
@@
 WK acknowledges CSC -- IT Centre for Science (Espoo, Finland) is acknowledged for excellent computational resources (project number tty3995).
+
+JM acknowledges CSC -- IT Center for Science for computational resources.
 
@@
 {\it Luca Monticelli}  \\
-{\it Jukka Maatta}  \\
+{\it Jukka M{\"a}{\"a}tt{\"a}}  prepared and performed simulations with Berger and Slipids models and analyzed the results.\\
 {\it O. H. Samuli Ollila} Designed and managed the work. Ran and analyzed several simulations. Wrote the manuscript.  \\

### 2015-03-26 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Move figures in the Fig folder & update tex file
missing Favela data in table 3

@@
   \centering
-  \includegraphics[width=8.6cm]{POPCstructure.eps}
+  \includegraphics[width=8.6cm]{../Fig/POPCstructure.pdf}
 
@@
 CHARMM36\cite{klauda10,lim12}   & POPC   & 128& 0& 0\% & 5120  & 303  & 150 & 100 & [\citenum{charmm36files}]$^*$  & SI  \\
+                                & POPC   & 512& 0& 0\% & ?  & 298  & 170 & 100 & ?  & ?  \\
+                                & POPC   & ?& ?& 10\% & ?  & 298  & 170 & 100 & ?  & ?  \\
+                                & POPC   & ?& ?& 15\% & ?  & 298  & 170 & 100 & ?  & ?  \\
                                 & POPC   & 100 & 24 & 19\%  &  4960   & 303 & 200 & 100 & [\citenum{charmm36files20perCHOL}]$^*$ & SI \\
+                                & POPC   & ?& ?& 20\% & ?  & 298  & 170 & 100 & ?  & ?  \\
+                                & POPC   & ?& ?& 25\% & ?  & 298  & 170 & 100 & ?  & ?  \\
+                                & POPC   & ?& ?& 35\% & ?  & 298  & 170 & 100 & ?  & ?  \\
+                                & POPC   & 256& 256& 50\% & ?  & 298  & 170 & 100 & ?  & ?  \\
                                & POPC   & 80 & 80 &50\%  &  4496    & 303 & 200 & 100 & [\citenum{charmm36files50perCHOL}]$^*$ & SI \\
@@
   \centering
-  \includegraphics[width=8.6cm]{HGorderparameters5.eps} \\
+  \includegraphics[width=8.6cm]{../Fig/HGorderparameters5.pdf} \\
   \todo{This figure should be updated similar to Fig.~\ref{HGorderparameters}} \\
@@
   \centering
-  \includegraphics[width=8.6cm]{g1-g2_Cdihs2.eps}
+  \includegraphics[width=8.6cm]{../Fig/g1-g2_Cdihs2.pdf}
   \caption{\label{dihDISTS}
@@
   \centering
-  \includegraphics[width=8.6cm]{a-bDIHS2.eps}
+  \includegraphics[width=8.6cm]{../Fig/a-bDIHS2.pdf}
   \caption{\label{dihDISTS2}
@@
   \centering
-  \includegraphics[width=17.2cm]{dihsCHOLcharmm.pdf}
+  \includegraphics[width=17.2cm]{../Fig/dihsCHOLcharmm.pdf}
   \caption{\label{dihsCHOLcharmm}

### 2015-03-26 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
add new CHARMM36-UA information by Claire Loison

@@
 {Department of Chemical Engineering, Kyoto University, Kyoto, Japan}
+\author{Claire Loison}
+\affiliation[Lyon CNRS]{Institut Lumi\`ere Mati\`ere, UMR5306 Universit\'e Lyon 1-CNRS, Universit\'e de Lyon 69622 Villeurbanne, France}
 \author{Markus S. Miettinen}
@@
 Tj\"ornhammar14 \cite{tjornhammar14}   & DPPC  & 144 & 7056 & 323 & 200 & 100 & [\citenum{tjornhammarfiles}]$^*$ & [\citenum{tjornhammar14}] \\
-CHARMM36-UA~\cite{henin08,lee14}     & DLPC   & 128  & 3840  & 323 & 30 & 20 & [\citenum{charmmUAfiles}] & SI \\
+CHARMM36-UA~\cite{lee14}     & DPPC   & 72  & 2189  & 323 & 80 & 50 & ? & SI \\
+CHARMMXX-UA~\cite{henin08,lee14}     & DLPC   & 128  & 3840  & 323 & 30 & 20 & [\citenum{charmmUAfiles}] & SI \\
 \end{tabular}
@@
 and the European Research Council (Advanced Grant project CROWDED-PRO-LIPIDS).
+
+CL and AB acknowledges financial support from the French National Research Agency
+(ANR: “Biolubrication by phospholipid membranes” Bioub2012) and
+computing time allocation from P{\^o}le Scientifique de Mod{\'e}lisation Num{\'e}rique from the ENS Lyon (PSMN),
+and Centre Informatique National de l'Enseignement Sup{\'e}rieur (CINES, Montpelier, France)
+(Project c2015096850).
+
 \end{acknowledgement}
@@
 \todo{Please write a text summarizing your contribution} \\
-{\it Alexandru Botan} \\
+{\it Alexandru Botan} provided simulation results for charmm36-UA\\
 {\it Fernando Favela}  \\
@@
 {\it Antti Lamberg}  \\
+{\it Claire Loison} provided simulation results for charmm36-UA \\
 {\it Markus S. Miettinen}  \\

### 2015-03-27 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Update Contributions of Hubert Santuz

@@
 \affiliation[INSERM]
-{INSERM, U1134, DSIMB; Institut National de la Transfusion Sanguine (INTS); Laboratoire d'Excellence GR-Ex, Paris, France}
+{INSERM, UMR\_S 1134, DSIMB, Paris, France}
 \alsoaffiliation[Diderot]
-{Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, Paris, France}
+{Universit\'e Paris Diderot, Sorbonne Paris Cit\'e, UMR\_S 1134, Paris, France}
+\alsoaffiliation[INTS]
+{Institut National de la Transfusion Sanguine (INTS), Paris, France}
+\alsoaffiliation[Labex]
+{Laboratoire d'Excellence GR-Ex, Paris, France}
 \author{Joona Tynkkynen}
@@
 
+HS acknowleges Catherine Etchebest and St\'ephane T\'eletch\'ea for useful discussions and continued support, GENCI-CINES for the HPC resources (Grant 2014-c2014077209).
+
 \end{acknowledgement}
@@
 {\it Tomasz Rog}  \\
-{\it Hubert Santuz}  \\
+{\it Hubert Santuz} prepared and performed the cholesterol simulations with CHARMM36 and analyzed the results. \\
 {\it Joona Tynkkynen} prepared and performed the dehydration simulations with the MacRog FF.\\

### 2015-03-27 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Minor Corrections + Update pdf

@@
 
-HS acknowleges Catherine Etchebest and St\'ephane T\'eletch\'ea for useful discussions and continued support, GENCI-CINES for the HPC resources (Grant 2014-c2014077209).
+HS acknowleges Catherine Etchebest and St\'ephane T\'eletch\'ea for useful discussions and continued support, the HPC resources granted from GENCI-CINES (Grant 2014-c2014077209) and computer facilities provided by R\'egion Ile de France and INTS (SESAME 2009 project).
 

### 2015-03-30 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Update path of dehydrationPN.pdf
add Favela infromation

@@
 \author{Fernando Favela-Rosales}
-\affiliation[Mexico]{Departamento de F{\'i}sica - CINVESTAV}
+\affiliation[Mexico]{Departamento de F\'isica, Centro de Investigaci\'on y de Estudios Avanzados del IPN, Apartado Postal 14-740, 07000 M\'exico D.F., M\'exico}
 \author{Patrick F. J. Fuchs}
@@
 
+Fernando Favela acknowledges CONACYT and DGAPA UNAM IG100513
+for financial support, Cluster H\'ibrido de Superc\'omputo Xiuhcoatl - CINVESTAV and Miztli - UNAM for computational resources.
+
 \end{acknowledgement}
@@
 {\it Alexandru Botan} provided simulation results for charmm36-UA\\
-{\it Fernando Favela}  \\
+{\it Fernando Favela} prepared, performed and analyzed simulations
+with popc and cholesterol for the CHARMM36 All-Atom FF. \\
 {\it Patrick F. J. Fuchs} Ran and analyzed the Poger simulations. Provided scientific information which significantly advanced the project (signs and forking of order parameters). \\

### 2015-03-30 POPCberger/PNvectorANGLE.sh
add PN vector calculation script for Berger

@@ -1,0 +1,3 @@
+grep "  P" popc1-5prot.gro |awk 'BEGIN{FIELDWIDTHS = "5 5 5 5 8 8 8"}{print $5" "$6" "$7}' > Pdat.dat
+grep "  N" popc1-5prot.gro |awk 'BEGIN{FIELDWIDTHS = "5 5 5 5 8 8 8"}{print $5" "$6" "$7}' > Ndat.dat
+paste Pdat.dat Ndat.dat | awk '{x=$1-$4; y=$2-$5; z=$3-$6; angle=angle+atan2(sqrt(x*x+y*y),sqrt(z*z));sum=sum+1;}END{print 180*angle/(sum*3.14)}'

### 2015-03-31 DLPCberger/dehydration/12waterPERpc/downloadFILES.sh
Add Berger DLPC dehydration data & update others

@@ -1,0 +1,7 @@
+wget https://zenodo.org/record/16293/files/dlpc.itp
+wget https://zenodo.org/record/16293/files/init2_TI5.gro
+wget https://zenodo.org/record/16293/files/lipid.itp
+wget https://zenodo.org/record/16293/files/system.top
+wget https://zenodo.org/record/16293/files/TI.mdp
+wget https://zenodo.org/record/16293/files/topol.tpr
+wget https://zenodo.org/record/16293/files/traj_dil.xtc

### 2015-03-31 DLPCberger/dehydration/16waterPERpc/MASTER.sh

@@ -1,0 +1,3 @@
+#sh downloadFILES.sh
+#sh ../make402tpr.sh
+./calcORDPwithERROR.sh

### 2015-03-31 DLPCberger/dehydration/16waterPERpc/calcORDP2.sh

@@ -1,0 +1,9 @@
+awk -v Cname="   C5" -v Hname="  H51" -f /home/ollilas1/Dropbox/codes/gro_OP.awk runPROT.gro
+awk -v Cname="   C5" -v Hname="  H52" -f /home/ollilas1/Dropbox/codes/gro_OP.awk runPROT.gro
+awk -v Cname="   C6" -v Hname="  H61" -f /home/ollilas1/Dropbox/codes/gro_OP.awk runPROT.gro
+awk -v Cname="   C6" -v Hname="  H62" -f /home/ollilas1/Dropbox/codes/gro_OP.awk runPROT.gro
+awk -v Cname="  C12" -v Hname=" H121" -f /home/ollilas1/Dropbox/codes/gro_OP.awk runPROT.gro
+awk -v Cname="  C12" -v Hname=" H122" -f /home/ollilas1/Dropbox/codes/gro_OP.awk runPROT.gro
+awk -v Cname="  C13" -v Hname=" H131" -f /home/ollilas1/Dropbox/codes/gro_OP.awk runPROT.gro
+awk -v Cname="  C32" -v Hname=" H321" -f /home/ollilas1/Dropbox/codes/gro_OP.awk runPROT.gro
+awk -v Cname="  C32" -v Hname=" H322" -f /home/ollilas1/Dropbox/codes/gro_OP.awk runPROT.gro

### 2015-03-31 DLPCberger/dehydration/16waterPERpc/calcORDPwithERROR.sh

@@ -1,0 +1,58 @@
+#!/bin/bash
+Nlipids=72
+tprname402=topol402.tpr
+tprname=topol.tpr
+trajname=traj_dil.xtc
+starttime=0
+#echo 'q
+#' | make_ndx -f $tprname
+#echo 0 | trjconv -f $trajname -s $tprname -o trjtmpINBOX.xtc -n index.ndx -pbc res -b $starttime
+#echo 'keep 2
+#splitres 0
+#q
+#' | make_ndx -f $tprname
+for((  j = 1 ;  j <= $Nlipids;  j=j+1  ))
+do
+    tmp=1000
+    i=$(($j+$tmp))
+    rm runPROT.gro
+    echo "$j" | /home/ollilas1/gromacs/gromacs402/bin/protonate -f trjtmpINBOX.xtc -s $tprname402 -o runPROT.gro -n index.ndx
+    sh calcORDP2.sh > OrderParameters_"$j".dat
+done
+echo '0
+0
+0
+0
+0
+0
+0
+0
+0' > OrderParametersSUM.dat
+echo '0
+0
+0
+0
+0
+0
+0
+0
+0' > OrderParametersSDsum.dat
+for((  j = 1 ;  j <= $Nlipids;  j=j+1  ))
+do
+    tmp=1000
+    i=$(($j+$tmp))
+    paste OrderParametersSUM.dat OrderParameters_"$j".dat | awk '{print $1+$2}' > tmp.dat
+    mv tmp.dat OrderParametersSUM.dat
+    sum=$(($sum+1))
+done
+cat OrderParametersSUM.dat | awk -v sum=$sum '{print $1/sum}' > OrderParametersAV.dat
+sum=0
+for((  j = 1 ;  j <= $Nlipids;  j=j+1  ))
+do
+    tmp=1000
+    i=$(($j+$tmp))
+    paste OrderParametersAV.dat OrderParameters_"$j".dat OrderParametersSDsum.dat | awk '{print $3+($1-$2)*($1-$2)}' > tmp.dat
+    mv tmp.dat OrderParametersSDsum.dat
+    sum=$(($sum+1))
+done
+paste OrderParametersAV.dat OrderParametersSDsum.dat | awk -v sum=$sum '{print $1" "sqrt($2/sum)/sqrt(sum)}'

### 2015-03-31 DLPCberger/dehydration/16waterPERpc/downloadFILES.sh

@@ -1,0 +1,7 @@
+wget https://zenodo.org/record/16292/files/dlpc.itp
+wget https://zenodo.org/record/16292/files/init2_TI5.gro
+wget https://zenodo.org/record/16292/files/lipid.itp
+wget https://zenodo.org/record/16292/files/system.top
+wget https://zenodo.org/record/16292/files/TI.mdp
+wget https://zenodo.org/record/16292/files/topol.tpr
+wget https://zenodo.org/record/16292/files/traj_dil.xtc

### 2015-03-31 DLPCberger/dehydration/20waterPERpc/calcORDPwithERROR.sh

@@ -1,0 +1,58 @@
+#!/bin/bash
+Nlipids=72
+tprname402=topol402.tpr
+tprname=topol.tpr
+trajname=traj_dil.xtc
+starttime=0
+echo 'q
+' | make_ndx -f $tprname
+echo 0 | trjconv -f $trajname -s $tprname -o trjtmpINBOX.xtc -n index.ndx -pbc res -b $starttime
+echo 'keep 2
+splitres 0
+q
+' | make_ndx -f $tprname
+for((  j = 1 ;  j <= $Nlipids;  j=j+1  ))
+do
+    tmp=1000
+    i=$(($j+$tmp))
+    rm runPROT.gro
+    echo "$j" | /home/ollilas1/gromacs/gromacs402/bin/protonate -f trjtmpINBOX.xtc -s $tprname402 -o runPROT.gro -n index.ndx
+    sh calcORDP2.sh > OrderParameters_"$j".dat
+done
+echo '0
+0
+0
+0
+0
+0
+0
+0
+0' > OrderParametersSUM.dat
+echo '0
+0
+0
+0
+0
+0
+0
+0
+0' > OrderParametersSDsum.dat
+for((  j = 1 ;  j <= $Nlipids;  j=j+1  ))
+do
+    tmp=1000
+    i=$(($j+$tmp))
+    paste OrderParametersSUM.dat OrderParameters_"$j".dat | awk '{print $1+$2}' > tmp.dat
+    mv tmp.dat OrderParametersSUM.dat
+    sum=$(($sum+1))
+done
+cat OrderParametersSUM.dat | awk -v sum=$sum '{print $1/sum}' > OrderParametersAV.dat
+sum=0
+for((  j = 1 ;  j <= $Nlipids;  j=j+1  ))
+do
+    tmp=1000
+    i=$(($j+$tmp))
+    paste OrderParametersAV.dat OrderParameters_"$j".dat OrderParametersSDsum.dat | awk '{print $3+($1-$2)*($1-$2)}' > tmp.dat
+    mv tmp.dat OrderParametersSDsum.dat
+    sum=$(($sum+1))
+done
+paste OrderParametersAV.dat OrderParametersSDsum.dat | awk -v sum=$sum '{print $1" "sqrt($2/sum)/sqrt(sum)}'

### 2015-03-31 DLPCberger/dehydration/20waterPERpc/downloadFILES.sh

@@ -1,0 +1,7 @@
+wget https://zenodo.org/record/16291/files/dlpc.itp
+wget https://zenodo.org/record/16291/files/init2_TI5.gro
+wget https://zenodo.org/record/16291/files/lipid.itp
+wget https://zenodo.org/record/16291/files/system.top
+wget https://zenodo.org/record/16291/files/TI.mdp
+wget https://zenodo.org/record/16291/files/topol.tpr
+wget https://zenodo.org/record/16291/files/traj_dil.xtc

### 2015-03-31 DLPCberger/dehydration/24waterPERpc/downloadFILES.sh

@@ -1,0 +1,7 @@
+wget https://zenodo.org/record/16289/files/dlpc.itp
+wget https://zenodo.org/record/16289/files/init2_TI5.gro
+wget https://zenodo.org/record/16289/files/lipid.itp
+wget https://zenodo.org/record/16289/files/system.top
+wget https://zenodo.org/record/16289/files/TI.mdp
+wget https://zenodo.org/record/16289/files/topol.tpr
+wget https://zenodo.org/record/16289/files/traj_dil.xtc

### 2015-03-31 DLPCberger/dehydration/28waterPERpc/MASTER.sh

@@ -1,0 +1,3 @@
+sh downloadFILES.sh
+sh ../make402tpr.sh
+./calcORDPwithERROR.sh

### 2015-03-31 DLPCberger/dehydration/28waterPERpc/downloadFILES.sh

@@ -1,0 +1,7 @@
+wget https://zenodo.org/record/16287/files/dlpc.itp
+wget https://zenodo.org/record/16287/files/init2_TI5.gro
+wget https://zenodo.org/record/16287/files/lipid.itp
+wget https://zenodo.org/record/16287/files/system.top
+wget https://zenodo.org/record/16287/files/TI.mdp
+wget https://zenodo.org/record/16287/files/topol.tpr
+wget https://zenodo.org/record/16287/files/traj_dil.xtc

### 2015-03-31 DLPCberger/dehydration/4waterPERpc/calcHGop.sh

@@ -1,0 +1,7 @@
+wget https://zenodo.org/record/16295/files/dlpc.itp
+wget https://zenodo.org/record/16295/files/init2_TI5.gro
+wget https://zenodo.org/record/16295/files/lipid.itp
+wget https://zenodo.org/record/16295/files/system.top
+wget https://zenodo.org/record/16295/files/TI.mdp
+wget https://zenodo.org/record/16295/files/topol.tpr
+wget https://zenodo.org/record/16295/files/traj_dil.xtc

### 2015-03-31 DLPCberger/dehydration/4waterPERpc/calcORDPwithERROR.sh

@@ -1,0 +1,57 @@
+Nlipids=72
+tprname402=topol402.tpr
+tprname=topol.tpr
+trajname=traj_dil.xtc
+starttime=0
+echo 'q
+' | make_ndx -f $tprname
+echo 0 | trjconv -f $trajname -s $tprname -o trjtmpINBOX.xtc -n index.ndx -pbc res -b $starttime
+echo 'keep 2
+splitres 0
+q
+' | make_ndx -f $tprname
+for((  j = 1 ;  j <= $Nlipids;  j=j+1  ))
+do
+    tmp=1000
+    i=$(($j+$tmp))
+    rm runPROT.gro
+    echo "$j" | /home/ollilas1/gromacs/gromacs402/bin/protonate -f trjtmpINBOX.xtc -s $tprname402 -o runPROT.gro -n index.ndx
+    sh calcORDP2.sh > OrderParameters_"$j".dat
+done
+echo '0
+0
+0
+0
+0
+0
+0
+0
+0' > OrderParametersSUM.dat
+echo '0
+0
+0
+0
+0
+0
+0
+0
+0' > OrderParametersSDsum.dat
+for((  j = 1 ;  j <= $Nlipids;  j=j+1  ))
+do
+    tmp=1000
+    i=$(($j+$tmp))
+    paste OrderParametersSUM.dat OrderParameters_"$j".dat | awk '{print $1+$2}' > tmp.dat
+    mv tmp.dat OrderParametersSUM.dat
+    sum=$(($sum+1))
+done
+cat OrderParametersSUM.dat | awk -v sum=$sum '{print $1/sum}' > OrderParametersAV.dat
+sum=0
+for((  j = 1 ;  j <= $Nlipids;  j=j+1  ))
+do
+    tmp=1000
+    i=$(($j+$tmp))
+    paste OrderParametersAV.dat OrderParameters_"$j".dat OrderParametersSDsum.dat | awk '{print $3+($1-$2)*($1-$2)}' > tmp.dat
+    mv tmp.dat OrderParametersSDsum.dat
+    sum=$(($sum+1))
+done
+paste OrderParametersAV.dat OrderParametersSDsum.dat | awk -v sum=$sum '{print $1" "sqrt($2/sum)/sqrt(sum)}'

### 2015-03-31 DLPCberger/dehydration/4waterPERpc/make402tpr.sh

@@ -1,0 +1,1 @@
+/home/ollilas1/gromacs/gromacs402/bin/protonate -f TI.mdp -c init2_TI5.gro -p system.top -o topol402.tpr -v

### 2015-03-31 DLPCberger/dehydration/8waterPERpc/MASTER.sh

@@ -1,0 +1,3 @@
+sh downloadFILES.sh
+sh make402tpr.sh
+../calcORDPwithERROR.sh

### 2015-03-31 DLPCberger/dehydration/8waterPERpc/downloadFILES.sh

@@ -1,0 +1,7 @@
+wget https://zenodo.org/record/16294/files/dlpc.itp
+wget https://zenodo.org/record/16294/files/init2_TI5.gro
+wget https://zenodo.org/record/16294/files/lipid.itp
+wget https://zenodo.org/record/16294/files/system.top
+wget https://zenodo.org/record/16294/files/TI.mdp
+wget https://zenodo.org/record/16294/files/topol.tpr
+wget https://zenodo.org/record/16294/files/traj_dil.xtc

### 2015-03-31 DLPCberger/dehydration/make402tpr.sh

@@ -1,0 +1,1 @@
+/home/ollilas1/gromacs/gromacs402/bin/grompp -f TI.mdp -c init2_TI5.gro -p system.top -o topol402.tpr -v

### 2015-03-31 DLPCberger/dehydration/makePLOTfile.sh

@@ -1,0 +1,25 @@
+echo '#
+#Dehydration results for DLPC from Berger force field. Simulation data delivered by Matej Kanduc:
+#
+#http://dx.doi.org/10.5281/zenodo.16295
+#http://dx.doi.org/10.5281/zenodo.16294
+#http://dx.doi.org/10.5281/zenodo.16293
+#http://dx.doi.org/10.5281/zenodo.16292
+#http://dx.doi.org/10.5281/zenodo.16291
+#http://dx.doi.org/10.5281/zenodo.16289
+#http://dx.doi.org/10.5281/zenodo.16287
+#
+#Analysis done with scripts
+#*waterPERpc/downloadFILES.sh
+#make402tpr.sh
+#*waterPERpc/calcORDPwithERROR.sh
+
+'
+for((j=1;j<10;j=j+1))
+do
+for i in 4 8 12 16 20 24 28
+do
+cat "$i"waterPERpc/OrderParametersAV.dat | awk -v j=$j -v i=$i '{if(NR==j){print i" "$1}}' 
+done
+echo " "
+done

### 2015-03-31 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
update hydration results in the manuscript

@@
                                         &   POPC & 7  &128 & 896   & 298  & 60 & 50 & [\citenum{bergerDEHYDfiles}]$^*$ & SI \\
-CHARMM36\cite{klauda10}              & POPC   & 31 & 72  &  2242 & 303 & 30 & 20 & [\citenum{charmm36filesSHORT}]$^*$ & SI \\
+CHARMM36\cite{klauda10}          & POPC   & 40 & 128 &  5120   & 303 & 150 & 100  & [\citenum{charmm36files}]$^*$   & SI \\
+                                 & POPC   & 31  & 72  &  2242 & 303 & 30 & 20 & [\citenum{charmm36filesSHORT}]$^*$ & SI \\
                                & POPC   & 15 & 72 &  1080  & 303 & 59 & 40 & [\citenum{charmm36files15wPERl}]$^*$ & SI \\
@@
   \centering
-  \includegraphics[width=8.6cm]{../DATAreportediINblog/dehydration.pdf}
-  \todo{It should be checked that we have the up-to-date GAFF results in this plot.} \\
+  \includegraphics[width=8.6cm]{../DATAreportediINblog/dehydration.pdf} \\
+  \todo{DONE} \\
   \caption{\label{ordPhydr}
@@
   \centering
-  \includegraphics[width=8.6cm]{../DATAreportediINblog/dehydrationPN.pdf}
+  \includegraphics[width=8.6cm]{../Fig/dehydrationPN.pdf}
 

### 2015-04-02 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
update dehydration data & add DLPC berger results

@@
                                         &   POPC & 7  &128 & 896   & 298  & 60 & 50 & [\citenum{bergerDEHYDfiles}]$^*$ & SI \\
+Berger-DLPC-?~\cite{??}          &   DLPC & 28  &72 & 2016  & 300  & ? & 80 & [\citenum{bergerFILESdlpc28}]$^*$ & SI \\
+                                 &   DLPC & 24  &72 & 1728  & 300  & ? & 80 & [\citenum{bergerFILESdlpc24}]$^*$ & SI \\
+                                 &   DLPC & 20  &72 & 1440  & 300  & ? & 80 & [\citenum{bergerFILESdlpc20}]$^*$ & SI \\
+                                 &   DLPC & 16  &72 & 1152  & 300  & ? & 80 & [\citenum{bergerFILESdlpc16}]$^*$ & SI \\
+                                 &   DLPC & 12  &72 & 864  & 300  & ? & 80 & [\citenum{bergerFILESdlpc12}]$^*$ & SI \\
+                                 &   DLPC & 8  &72 & 576  & 300  & ? & 80 & [\citenum{bergerFILESdlpc8}]$^*$ & SI \\
+                                 &   DLPC & 4  &72 & 288  & 300  & ? & 80 & [\citenum{bergerFILESdlpc4}]$^*$ & SI \\
 CHARMM36\cite{klauda10}          & POPC   & 40 & 128 &  5120   & 303 & 150 & 100  & [\citenum{charmm36files}]$^*$   & SI \\

### 2015-04-02 DLPCberger/dehydration/20waterPERpc/calcORDPchains.sh

@@ -1,0 +1,30 @@
+#!/bin/bash
+tprname=topol.tpr
+tprname402=topol402.tpr
+trajname=traj_dil.xtc
+trajINBOX=trjtmpINBOX.xtc
+trajgroname=trajPROT.gro
+sn1outname=OrderParamSN1.dat
+sn2outname=OrderParamSN2.dat
+mappingFILE=/wrk/ollilas1/HGmodel/NMRlipids/NmrLipidsCholXray/MAPPING/mappingDLPCberger.txt
+groOPpath=../../../scripts/gro_OP.awk
+echo System | trjconv -f $trajname -s $tprname -o $trajINBOX -pbc res -b 0
+echo DPP | /home/ollilas1/gromacs/gromacs402/bin/protonate -f $trajINBOX -s $tprname402 -o $trajgroname
+for((  j = 3 ;  j <= 13;  j=j+1  ))
+do
+Cname=$(grep M_G1C"$j"_M $mappingFILE | awk '{printf "%5s\n",$2}')
+H1name=$(grep M_G1C"$j"H1_M $mappingFILE | awk '{printf "%5s\n",$2}')
+H2name=$(grep M_G1C"$j"H2_M $mappingFILE | awk '{printf "%5s\n",$2}')
+H1op=$(awk -v Cname="$Cname" -v Hname="$H1name" -f $groOPpath $trajgroname)
+H2op=$(awk -v Cname="$Cname" -v Hname="$H2name" -f $groOPpath $trajgroname)
+echo $j $H1op $H2op >> $sn1outname
+done
+for((  j = 3 ;  j <= 13;  j=j+1  ))
+do
+Cname=$(grep M_G2C"$j"_M $mappingFILE | awk '{printf "%5s\n",$2}')
+H1name=$(grep M_G2C"$j"H1_M $mappingFILE | awk '{printf "%5s\n",$2}')
+H2name=$(grep M_G2C"$j"H2_M $mappingFILE | awk '{printf "%5s\n",$2}')
+H1op=$(awk -v Cname="$Cname" -v Hname="$H1name" -f $groOPpath $trajgroname)
+H2op=$(awk -v Cname="$Cname" -v Hname="$H2name" -f $groOPpath $trajgroname)
+echo $j $H1op $H2op >> $sn2outname
+done

### 2015-04-07 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
update dehydration discussion after adding Berger DLPC results

@@
 data for dehydrated lipid bilayers and cholesterol containing bilayers, and interpretate it with simulations.
-The interpretation reveals that the increased choline order parameters indicate P--N vector tilting more parallel to membrane,
+The interpretation suggest that the increased choline order parameters indicate P--N vector tilting more parallel to membrane,
 and that cholesterol induces only minor changes to the glycerol backbone structure.
@@
 The magnitude of shifting is written in the y-axis label.
-The choline order parameters increase with dehydration in all
-models, in qualitative agreement with experiments. 
-The measured decrease in both g$_3$ and g$_2$ order parameters with dehydration is reproduced only in CHARMM36.
-\todo{This sentence must be rewritten after the up-to-date results for GAFFlipid are in the figure.}
+The choline order parameter increase with dehydration is seen in all
+models despite of some additional fluctuations. Thus the choline order parameter response to dehydration can be
+intepreted to be in qualitative agreement with experiments.
+The situation is significantly more complicated for the glycerol backbone segments 
+and none of the models can be interpreted to qualitatively agree with experiments
+for all the segmets.
+\todo{DONE}
 
@@
 all models as a function of dehydration in Fig.~\ref{PNangle}.
-
-It should be noted, however, that free energy landscape is not correct in the models which
+However, the amount of increase depends on the model. Especially the DLPC simulations with Berger model
+predict significantly stronger P--N vector tilt compared with the other models. The Berger model
+has also generally larger P--N vector angles and its choline order parameters are more off from
+experiments than other models. Thus the relatively modest tilting with dehydration
+predicted by MacRog, CHARMM36 and GAFFlipid is probably more realistic.
+
+It should be also noted that the free energy landscape is not correct in the models which
 are not able to reproduce the experimental order parameters. Thus the energetical response
@@
 The response of the glycerol backbone to dehydration seems to be more subtle than that of the choline headgroup 
-as CHARMM36 is the only model that reproduces the decrease in g$_2$ and g$_3$ segments.
-\todo{This sentence must be rewritten after the up-to-date results for GAFFlipid are in the figure.}
+and none of the models reproduced the experimental results with the accuracy allowing
+the structural interpretation.
+\todo{DONE}
 
@@
 - The review of previous experimental results revealed that the 
-choline order parameters are increasing when the bilayer is dehydrated. Our simulations show that
+choline order parameters are increasing when the bilayer is dehydrated. Our simulations suggest that
 this can be explained by P--N vector tilting more parallel to the membrane. 

### 2015-04-07 DATAreportediINblog/DPPC/cholesterol/QStoOP.sh
add experimental DPPC order parameter data as a function of cholesterol

@@ -1,0 +1,3 @@
+cat brownDATAalpha1.dat | awk '{print 100*$1/2" "$2*0.00784}' > brownOPalpha1.dat
+cat brownDATAalpha2.dat | awk '{print 100*$1/2" "$2*0.00784}' > brownOPalpha2.dat
+cat brownDATAbeta.dat | awk '{print 100*$1/2" "(-1)*$2*0.00784}' > brownOPbeta.dat

### 2015-04-08 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
add Favela data into cholesterol figure

@@
   \centering
-  \includegraphics[width=8.6cm]{../DATAreportediINblog/cholesterolization.pdf}
-  \todo{The data delivered by Fernando Favela should be added here.} \\
+  \includegraphics[width=8.6cm]{../DATAreportediINblog/cholesterolization.pdf} \\
+  \todo{DONE} \\
   \caption{\label{ordPchol}
@@
   \caption{\label{dihsCHOLcharmm}
-    The effect of cholesterol content on the glycerol backbone and choline dihedral angles in CHARMM36 model.}
+    The effect of cholesterol content on the POPC glycerol backbone and choline dihedral angles in CHARMM36 model (T=303~K).}
 \end{figure*}

### 2015-04-08 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
update authorlist

@@
 {Department of Physics, Tampere University of Technology, Tampere, Finland}
+\author{Matej Kanduc}
+\affiliation[Freie Universitat Berlin] 
+{Fachbereich Physik, Freie Universitat Berlin, Berlin, Germany}
 \author{Waldemar Kulig}
@@
 {\it Matti Javanainen} prepared and performed simulations with multiple lipid models and analyzed the results. Supervised the work of JT.\\
+{\it Matej Kanduc} \\
 {\it Waldemar Kulig} prepared the MD simulations with cholesterol for the MacRog FF \\

### 2015-04-09 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
update DLPC berger results for analysis between 20-80ns

@@
                                         &   POPC & 7  &128 & 896   & 298  & 60 & 50 & [\citenum{bergerDEHYDfiles}]$^*$ & SI \\
-Berger-DLPC-?~\cite{??}          &   DLPC & 28  &72 & 2016  & 300  & ? & 80 & [\citenum{bergerFILESdlpc28}]$^*$ & SI \\
-                                 &   DLPC & 24  &72 & 1728  & 300  & ? & 80 & [\citenum{bergerFILESdlpc24}]$^*$ & SI \\
-                                 &   DLPC & 20  &72 & 1440  & 300  & ? & 80 & [\citenum{bergerFILESdlpc20}]$^*$ & SI \\
-                                 &   DLPC & 16  &72 & 1152  & 300  & ? & 80 & [\citenum{bergerFILESdlpc16}]$^*$ & SI \\
-                                 &   DLPC & 12  &72 & 864  & 300  & ? & 80 & [\citenum{bergerFILESdlpc12}]$^*$ & SI \\
-                                 &   DLPC & 8  &72 & 576  & 300  & ? & 80 & [\citenum{bergerFILESdlpc8}]$^*$ & SI \\
-                                 &   DLPC & 4  &72 & 288  & 300  & ? & 80 & [\citenum{bergerFILESdlpc4}]$^*$ & SI \\
+Berger-DLPC-?~\cite{??}          &   DLPC & 28  &72 & 2016  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc28}]$^*$ & [\citenum{kanduc13}] \\
+                                 &   DLPC & 24  &72 & 1728  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc24}]$^*$ & [\citenum{kanduc13}] \\
+                                 &   DLPC & 20  &72 & 1440  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc20}]$^*$ & [\citenum{kanduc13}] \\
+                                 &   DLPC & 16  &72 & 1152  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc16}]$^*$ & [\citenum{kanduc13}] \\
+                                 &   DLPC & 12  &72 & 864  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc12}]$^*$ & [\citenum{kanduc13}] \\
+                                 &   DLPC & 8  &72 & 576  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc8}]$^*$ & [\citenum{kanduc13}] \\
+                                 &   DLPC & 4  &72 & 288  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc4}]$^*$ & [\citenum{kanduc13}] \\
 CHARMM36\cite{klauda10}          & POPC   & 40 & 128 &  5120   & 303 & 150 & 100  & [\citenum{charmm36files}]$^*$   & SI \\

### 2015-04-09 DLPCberger/dehydration/calcORDPwithERROR.sh

@@ -1,0 +1,58 @@
+#!/bin/bash
+Nlipids=72
+tprname402=topol402.tpr
+tprname=topol.tpr
+trajname=traj_dil.xtc
+starttime=20000
+echo 'q
+' | make_ndx -f $tprname
+echo 0 | trjconv -f $trajname -s $tprname -o trjtmpINBOX.xtc -n index.ndx -pbc res -b $starttime
+echo 'keep 2
+splitres 0
+q
+' | make_ndx -f $tprname
+for((  j = 1 ;  j <= $Nlipids;  j=j+1  ))
+do
+    tmp=1000
+    i=$(($j+$tmp))
+    rm runPROT.gro
+    echo "$j" | /home/ollilas1/gromacs/gromacs402/bin/protonate -f trjtmpINBOX.xtc -s $tprname402 -o runPROT.gro -n index.ndx
+    sh calcORDP2.sh > OrderParameters_"$j".dat
+done
+echo '0
+0
+0
+0
+0
+0
+0
+0
+0' > OrderParametersSUM.dat
+echo '0
+0
+0
+0
+0
+0
+0
+0
+0' > OrderParametersSDsum.dat
+for((  j = 1 ;  j <= $Nlipids;  j=j+1  ))
+do
+    tmp=1000
+    i=$(($j+$tmp))
+    paste OrderParametersSUM.dat OrderParameters_"$j".dat | awk '{print $1+$2}' > tmp.dat
+    mv tmp.dat OrderParametersSUM.dat
+    sum=$(($sum+1))
+done
+cat OrderParametersSUM.dat | awk -v sum=$sum '{print $1/sum}' > OrderParametersAV.dat
+sum=0
+for((  j = 1 ;  j <= $Nlipids;  j=j+1  ))
+do
+    tmp=1000
+    i=$(($j+$tmp))
+    paste OrderParametersAV.dat OrderParameters_"$j".dat OrderParametersSDsum.dat | awk '{print $3+($1-$2)*($1-$2)}' > tmp.dat
+    mv tmp.dat OrderParametersSDsum.dat
+    sum=$(($sum+1))
+done
+paste OrderParametersAV.dat OrderParametersSDsum.dat | awk -v sum=$sum '{print $1" "sqrt($2/sum)/sqrt(sum)}' >> OrderParametersTMP.dat

### 2015-04-09 DLPCberger/dehydration/makePLOTfile.sh

@@ -9,6 +9,8 @@ echo '#
 #http://dx.doi.org/10.5281/zenodo.16289
 #http://dx.doi.org/10.5281/zenodo.16287
 #
+#20-80ns of the trajectories is used in the analysis
+#
 #Analysis done with scripts
 #*waterPERpc/downloadFILES.sh
 #make402tpr.sh

### 2015-04-13 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
update information from Kanduc

--- 2015-04-08 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
@@
 {Department of Physics, Tampere University of Technology, Tampere, Finland}
+\author{Matej Kandu\v{c}}
+\affiliation[Freie Universit\"{a}t Berlin] 
+{Fachbereich Physik, Freie Universitat Berlin, Berlin, Germany}
 \author{Waldemar Kulig}
@@
 \author{Markus S. Miettinen}
-\affiliation[Freie Universitat Berlin] 
+\affiliation[Freie Universit\"{a}t Berlin] 
 {Fachbereich Physik, Freie Universitat Berlin, Berlin, Germany}
@@
                                         &   POPC & 7  &128 & 896   & 298  & 60 & 50 & [\citenum{bergerDEHYDfiles}]$^*$ & SI \\
-Berger-DLPC-?~\cite{??}          &   DLPC & 28  &72 & 2016  & 300  & ? & 80 & [\citenum{bergerFILESdlpc28}]$^*$ & SI \\
-                                 &   DLPC & 24  &72 & 1728  & 300  & ? & 80 & [\citenum{bergerFILESdlpc24}]$^*$ & SI \\
-                                 &   DLPC & 20  &72 & 1440  & 300  & ? & 80 & [\citenum{bergerFILESdlpc20}]$^*$ & SI \\
-                                 &   DLPC & 16  &72 & 1152  & 300  & ? & 80 & [\citenum{bergerFILESdlpc16}]$^*$ & SI \\
-                                 &   DLPC & 12  &72 & 864  & 300  & ? & 80 & [\citenum{bergerFILESdlpc12}]$^*$ & SI \\
-                                 &   DLPC & 8  &72 & 576  & 300  & ? & 80 & [\citenum{bergerFILESdlpc8}]$^*$ & SI \\
-                                 &   DLPC & 4  &72 & 288  & 300  & ? & 80 & [\citenum{bergerFILESdlpc4}]$^*$ & SI \\
+Berger-DLPC-13~\cite{kanduc13}          &   DLPC & 28  &72 & 2016  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc28}]$^*$ & [\citenum{kanduc13}] \\
+                                 &   DLPC & 24  &72 & 1728  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc24}]$^*$ & [\citenum{kanduc13}] \\
+                                 &   DLPC & 20  &72 & 1440  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc20}]$^*$ & [\citenum{kanduc13}] \\
+                                 &   DLPC & 16  &72 & 1152  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc16}]$^*$ & [\citenum{kanduc13}] \\
+                                 &   DLPC & 12  &72 & 864  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc12}]$^*$ & [\citenum{kanduc13}] \\
+                                 &   DLPC & 8  &72 & 576  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc8}]$^*$ & [\citenum{kanduc13}] \\
+                                 &   DLPC & 4  &72 & 288  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc4}]$^*$ & [\citenum{kanduc13}] \\
 CHARMM36\cite{klauda10}          & POPC   & 40 & 128 &  5120   & 303 & 150 & 100  & [\citenum{charmm36files}]$^*$   & SI \\
@@
 {\it Matti Javanainen} prepared and performed simulations with multiple lipid models and analyzed the results. Supervised the work of JT.\\
+{\it Matej Kanduc} provided simulation results for the Berger DLPC model \\
 {\it Waldemar Kulig} prepared the MD simulations with cholesterol for the MacRog FF \\

### 2015-04-14 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
updating manuscript based on comments by Ferreira

@@
 The absolute values of order parameters can be measured by detecting quadrupolar splitting with $^2$H NMR~\cite{seelig77c} or by detecting dipolar 
-splitting with $^{13}$C NMR~\cite{hong95a,gross97,dvinskikh05a,ferreira13}. The measurements are based on
+splitting with $^1$H-$^{13}$C NMR~\cite{hong95a,gross97,dvinskikh05a,ferreira13}. The measurements are based on
 different physical interactions and also the connection between order parameters and quadrupolar or dipolar splitting
@@
 splitting constant is estimated from various experiments to be 170~kHz leading to a numerical relation $|S_{{\rm CD}}|=0.00784 \times \Delta \nu_Q$~\cite{seelig77c}. 
-The absolute values of order parameters from the measured dipolar splitting $d_\mathrm{CH}$ ($^{13}$C NMR) are calculated using equation
+The absolute values of order parameters from the effective dipolar coupling $d_\mathrm{CH}$ ($^1$H-$^{13}$C NMR) are calculated using equation
 $|S_{{\rm CH}}|=\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c} d_\mathrm{CH}$, where
@@
 depending on the original authors~\cite{hong95a,gross97,dvinskikh05a,ferreira13}.
+The effective dipolar coupling $d_\mathrm{CH}$ is related to the measured dipolar splitting $\Delta \nu_{CH}$ 
+through scaling factor which depends on the pulse sequence used in the $^1$H-$^{13}$C NMR experiment~\cite{hong95a,gross97,dvinskikh05a,ferreira13}.
 It is important to note that the order parameters measured with different techniques based on different physical interactions are in good agreement
@@
 
-The absolute values of order parameters are accessible with both $^2$H NMR and $^{13}$C NMR techniques. 
-However, only $^{13}$C NMR techniques allow also the measurement of the sign of the order parameter~\cite{hong95a,hong95b,gross97}. 
+The absolute values of order parameters are accessible with both $^2$H NMR and $^1$H-$^{13}$C NMR techniques. 
+However, only $^1$H-$^{13}$C NMR techniques allow also the measurement of the sign of the order parameter~\cite{hong95a,hong95b,gross97}. 
 The measured sign is negative for almost all the carbons discussed in this work as only $\alpha$ is positive~\cite{hong95a,hong95b,gross97}. 
@@
 However, in some cases (e.g. g$_1$, g$_3$ and  C$_2$ carbon in the \textit{sn}-2 chain) the order parameters are not equal and this 
-can be observed with both $^2$H NMR and $^{13}$C NMR techniques. In the present work we call the phenomena of unequal order parameters 
+can be observed with both $^2$H NMR and $^1$H-$^{13}$C NMR techniques. In the present work we call the phenomena of unequal order parameters 
 for hydrogens attached to the same carbon as {\it forking} to avoid confusion with dipolar and quadrupolar splitting in NMR terminology. 
@@
 The specific deuteration of $\alpha$-, $\beta$- and g$_3$- segments of DPPC has been successful, 
-allowing the absolute value order parameter measurements for these segments by $^2$H--NMR~\cite{gally75,brown77,brown78,akutsu81}.
+allowing the absolute value order parameter measurements for these segments by $^2$H NMR~\cite{gally75,brown77,brown78,akutsu81}.
 In addition, the absolute values of order parameters for all glycerol backbone and choline headgroup segments in egg yolk lecithin~\cite{hong95a},
 DMPC~\cite{hong95b,gross97,dvinskikh05a}, DOPC~\cite{warschawski05} and POPC~\cite{warschawski05,ferreira13}
-have been measured with several different implementations of $^{13}$C NMR experiments. In addition, the signs of order parameters in some systems
-are measured with $^{13}$C NMR techniques~\cite{hong95a,hong95b,gross97}. The experimental values of glycerol backbone 
+have been measured with several different implementations of $^1$H-$^{13}$C NMR experiments. In addition, the signs of order parameters in some systems
+are measured with $^1$H-$^{13}$C NMR techniques~\cite{hong95a,hong95b,gross97}. The experimental values of glycerol backbone 
 and choline order parameters from various publications~\cite{gally75,akutsu81,gross97,dvinskikh05a,ferreira13} with the signs measured in~\cite{hong95a,hong95b,gross97} are shown in Fig.~\ref{HGorderparameters}.
@@
 for all fully hydrated PC bilayer, regardless of the variation in their acyl chain composition and temperature.
-Exceptions are the somewhat lower order parameters sometimes reported from measurements using $^{13}$C--NMR~\cite{hong95a,hong95b,warschawski05}.
+Exceptions are the somewhat lower order parameters sometimes reported from measurements using $^1$H-$^{13}$C NMR~\cite{hong95a,hong95b,warschawski05}.
 These experiments are not shown in Fig.~\ref{HGorderparameters} as the reported error bars are either relatively large~\cite{hong95a,hong95b}, 
@@
 have been measured under various different conditions~\cite{gally75,brown77,brown78,akutsu81,altenbach84,scherer89,bechinger91,ulrich94,dvinskikh05b,castro08,kapla12,ferreira13}.
-Also the order parameters for the glycerol backbone have been measured with $^{13}$C NMR in dehydrated conditions~\cite{dvinskikh05b}, and as a function 
+Also the order parameters for the glycerol backbone have been measured with $^1$H-$^{13}$C NMR in dehydrated conditions~\cite{dvinskikh05b}, and as a function 
 of anesthetics~\cite{castro08} and glycolipids~\cite{kapla12} for DMPC, and as a function of cholesterol 
@@
     The effect of dehydration on glycerol and choline order parameters in experiments.
-    The magnitudes of order parameters are measured for DMPC ($^{13}$C NMR) at 314~K~\cite{dvinskikh05b}, 
+    The magnitudes of order parameters are measured for DMPC ($^1$H-$^{13}$C NMR) at 314~K~\cite{dvinskikh05b}, 
     for POPC ($^2$H NMR) at 296~K~\cite{bechinger91} and for DOPC ($^2$H NMR) at 303~K~\cite{ulrich94}. 
@@
 
-The choline headgroup and glycerol backbone order parameters for POPC measured by $^{13}$C NMR~\cite{ferreira13} and DPPC choline order parameters 
+The choline headgroup and glycerol backbone order parameters for POPC measured by $^1$H-$^{13}$C NMR~\cite{ferreira13} and DPPC choline order parameters 
 measured by $^{2}$H NMR~\cite{brown78} are shown in Fig.~\ref{ordPchol} as a function of cholesterol content.
@@
 parameter~\cite{brown78}. These effects are, however, so small that they are barely visible in the scale used in Fig.~\ref{ordPchol}.
-Further, the effects of cholesterol on the glycerol backbone order parameters for POPC from $^{13}$C NMR experiment~\cite{ferreira13} 
+Further, the effects of cholesterol on the glycerol backbone order parameters for POPC from $^1$H-$^{13}$C NMR experiment~\cite{ferreira13} 
 are in good agreement with the results for the phosphatidylethanolamine (PE) measured by $^{2}$H NMR~\cite{ghosh82}.

### 2015-04-14 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
add Hogberg et al. simulation file citations

@@
 with periodic boundary conditions in all directions by using the GROMACS software package~\cite{hess08} 
-(version numbers 4.5.X--4.6.X) or LAMMPS~\cite{plimpton95}.
+(version numbers 4.5.X--4.6.X), LAMMPS~\cite{plimpton95} or MDynaMix~\cite{Lyubartsev00}.
 The number of molecules, simulation temperatures and the length of simulations of all the simulated systems 
@@
 Chiu \cite{chiu09}      & POPC  & 128 & 3552  & 298 & 56 & 50  & [\citenum{chiuFILES}]$^*$ & SI \\
-H\"ogberg08 \cite{rabinovich14}  & POPC   &  128 & 3840  & 303 & 100 & 80  & - & [\citenum{rabinovich14}]  \\
-H\"ogberg08 \cite{hogberg08}  & DMPC   &  98 & 3840  & 303 & 75 & 50 & - & [\citenum{hogberg08}] \\
+H\"ogberg08 \cite{rabinovich14}  & POPC   &  128 & 3840  & 303 & 100 & 80  &  [\citenum{hogbergPOPCfiles}]$^*$ & [\citenum{rabinovich14}]  \\
+H\"ogberg08 \cite{hogberg08}  & DMPC   &  98 & 3840  & 303 & 75 & 50 &  [\citenum{hogbergDMPCfiles}]$^*$ & [\citenum{hogberg08}] \\
 Ulmschneiders \cite{Ulmschneider09}    & POPC  & 128 & 3328 & 310 & 100 & 50 & [\citenum{ulmschneiderFILES}]$^*$ & SI \\
@@
 {\it Marius Retegan} Prepared and validated the GROMACS compatible parameter files for GAFFlipid and Lipid14 force fields.\\
-{\it Tomasz Rog}  \\
+{\it Tomasz Rog} Provided the topologies for the MacRog model and also the full hydration simulation data with this model. \\
 {\it Hubert Santuz} prepared and performed the cholesterol simulations with CHARMM36 and analyzed the results. \\

### 2015-04-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
update information from Favela

@@
 CHARMM36\cite{klauda10,lim12}   & POPC   & 128& 0& 0\% & 5120  & 303  & 150 & 100 & [\citenum{charmm36files}]$^*$  & SI  \\
-                                & POPC   & 512& 0& 0\% & ?  & 298  & 170 & 100 & ?  & ?  \\
-                                & POPC   & ?& ?& 10\% & ?  & 298  & 170 & 100 & ?  & ?  \\
-                                & POPC   & ?& ?& 15\% & ?  & 298  & 170 & 100 & ?  & ?  \\
+                                & POPC   & 512& 0& 0\% & 23943  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
+                                & POPC   & 460& 52& 10\% & 23569  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
+                                & POPC   & 436& 76& 15\% & 23331  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
                                 & POPC   & 100 & 24 & 19\%  &  4960   & 303 & 200 & 100 & [\citenum{charmm36files20perCHOL}]$^*$ & SI \\
-                                & POPC   & ?& ?& 20\% & ?  & 298  & 170 & 100 & ?  & ?  \\
-                                & POPC   & ?& ?& 25\% & ?  & 298  & 170 & 100 & ?  & ?  \\
-                                & POPC   & ?& ?& 35\% & ?  & 298  & 170 & 100 & ?  & ?  \\
-                                & POPC   & 256& 256& 50\% & ?  & 298  & 170 & 100 & ?  & ?  \\
+                                & POPC   & 410& 102& 20\% & 20972  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
+                                & POPC   & 384& 128& 25\% & 22327  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
+                                & POPC   & 332& 180& 35\% & 21340  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
+                                & POPC   & 256& 256& 50\% & 20334  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
                                & POPC   & 80 & 80 &50\%  &  4496    & 303 & 200 & 100 & [\citenum{charmm36files50perCHOL}]$^*$ & SI \\
@@
 
-{\it DPPC}
+{\it DPPC and POPC with 72 lipids}
 \todo{Markus Miettinen, please describe how you made the files.}
@@
 
-{\it POPC}
-The starting structures for the pure POPC and DOPC simulations was taken from the Slipids~\cite{jambeck12b} website (http://people.su.se/$\sim$jjm/Stockholm\_Lipids/Downloads.html).
+{\it POPC with 128 lipids}
+The starting structures for the pure POPC simulations was taken from the Slipids~\cite{jambeck12b} website (http://people.su.se/$\sim$jjm/Stockholm\_Lipids/Downloads.html).
 The starting structures for mixed POPC/Cholesterol simulations were constructed with the CHARMM-GUI website~\cite{jo08}. 
@@
 between 0.8~nm and 1.2~nm. PME~\cite{darden93,essman95} with real space cut-off of 1.2~nm was used for electrostatics.
+
+{\it POPC with 512 lipids}
+POPC The starting structures for simulations were constructed with
+the CHARMM-GUI website~\cite{jo08}. 512 POPC lipids (256 per leafleat) were 
+used for the initial 0\% concentration and they were subsequently 
+substituted by cholesterol molecules to fulfil the desired concentration.
+The TIP3P~\cite{jorgensen83} water model was used to solvate the system. 
+The publicly available port of the CHARMM36 forcefield parameters (\url{mackerell.umaryland.edu/CHARMM_ff_params.html}) 
+was used for both POPC and cholesterol. 
+Simulations were performed for 170 ns and the last 100 ns was used for the calculations.
+Timestep of 1 fs was used with the leapfrog integrator. Covalent bonds with hydrogen 
+were constrained with LINCS algorithm. Coordinates were written every 10 ps.
+PME with real space cut-off of 1.2 nm was used for electrostatics. Lennard-Jones 
+interactions were switched to zero between 0.8 nm and 1.2 nm. The neighbour lists with a
+cut-off of 1.2 nm were updated every 5 steps. Temperature was coupled separately for cholesterol, lipids
+and water to 298 K using the velocity-rescale method with coupling constant of 0.2 ps.
+Pressure was semi-isotropically coupled to the atmospheric pressure with the Berendsen
+method.
 

### 2015-04-16 LIPIDionINTERACT/calcCHANGES.sh
Moving lipid-ion interaction manuscript files to the new repo

@@ -1,17 +1,0 @@
-cat BERGERnaclCONS.dat | awk '{if(NR==1){a=$2;b=$3;c=$4;d=$5;e=$6;f=$7;g=$8;h=$9;}print $1" " $2-a" "$3-b" "$4-c" "$5-d" "$6-e" "$7-f" "$8-g" "$9-h}' > BERGERnaclCONSchange.dat
-cat CHARMMnaclCONS.dat | awk '{if(NR==1){a=$2;b=$3;c=$4;d=$5;e=$6;f=$7;g=$8;h=$9;}print $1" " $2-a" "$3-b" "$4-c" "$5-d" "$6-e" "$7-f" "$8-g" "$9-h}' > CHARMMnaclCONSchange.dat
-cat ORANGEnaclCONS.dat | awk '{if(NR==1){a=$2;b=$3;c=$4;d=$5;e=$6;f=$7;g=$8;h=$9;}print $1" " $2-a" "$3-b" "$4-c" "$5-d" "$6-e" "$7-f" "$8-g" "$9-h}' > ORANGEnaclCONSchange.dat
-cat MACROGnaclCONS.dat | awk '{if(NR==1){a=$2;b=$3;c=$4;d=$5;e=$6;f=$7;g=$8;h=$9;}print $1" " $2-a" "$3-b" "$4-c" "$5-d" "$6-e" "$7-f" "$8-g" "$9-h}' > MACROGnaclCONSchange.dat
-cat SLIPIDSnaclCONSdppc.dat | awk '{if(NR==1){a=$2;b=$3;c=$4;d=$5;e=$6;f=$7;g=$8;h=$9;}print $1" " $2-a" "$3-b" "$4-c" "$5-d" "$6-e" "$7-f" "$8-g" "$9-h}' > SLIPIDSnaclCONSchangeDPPC.dat
-cat SLIPIDSnaclCONSpopc.dat | awk '{if(NR==1){a=$2;b=$3;c=$4;d=$5;e=$6;f=$7;g=$8;h=$9;}print $1" " $2-a" "$3-b" "$4-c" "$5-d" "$6-e" "$7-f" "$8-g" "$9-h}' > SLIPIDSnaclCONSchangePOPC.dat
-
[…]

### 2015-04-17 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
update authors

@@
 \affiliation[Lyon CNRS]{Institut Lumi\`ere Mati\`ere, UMR5306 Universit\'e Lyon 1-CNRS, Universit\'e de Lyon 69622 Villeurbanne, France}
+\author{Alexander Lyubartsev}
+\affiliation[Stockholm University]
+{Division of Physical Chemistry, Department of Materials and Environmental Chemistry, Stockholm University, S-106 91 Stockholm, SWEDEN}
 \author{Markus S. Miettinen}
@@
 \author{Luca Monticelli}
-\affiliation[UMR] 
-{5IBCP, CNRS UMR 5086, Lyon, France}
+\affiliation[IBCP] 
+{Institut de Biologie et Chimie des Prot\'eines (IBCP), CNRS UMR 5086, Lyon, France}
 \author{Jukka M{\"a}{\"a}tt{\"a}}
@@
 
+Luca Monticelli acknowledges funding proved by the Institut national de la sant\'e et de la recherche m\'edicale (INSERM).
+
 \end{acknowledgement}
@@
 {\it Claire Loison} provided simulation results for charmm36-UA \\
+{\it Alexander Lyubartsev} \\
 {\it Markus S. Miettinen}  \\
-{\it Luca Monticelli}  \\
+{\it Luca Monticelli} Critical discussions in all phases of the project. Collaboration with Jukka  M{\"a}{\"a}tt{\"a}. \\
 {\it Jukka M{\"a}{\"a}tt{\"a}}  prepared and performed simulations with Berger and Slipids models and analyzed the results.\\

### 2015-04-18 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Added description of how CHARMM36 72 lipid POPC and DPPC files were created.

--- 2015-04-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
@@
 
-{\it DPPC and POPC with 72 lipids}
-\todo{Markus Miettinen, please describe how you made the files.}
-
+{\it DPPC and POPC with 72 lipids.}
+%
+The starting structures in PDB format were downloaded from the
+NIH/NHLBI Laboratory of Computational Biology Membrane Biophysics Section
+website (\url{http://www.lobos.nih.gov/mbs/coords.shtml}),
+which refers to these as the final structures (for DPPC after 40~ns, POPC after 35~ns)
+of the NPT lipid bilayer trajectories used in the original CHARMM36 publication~\cite{klauda10}.
+The TIP3P~\cite{jorgensen83} water model was used to solvate the system.
+%
+The publicly available CHARMM36 parameters in Gromacs format
+(September 2013 update: {\tt charmm36\_gmx\_format\_sep13.tgz})
+from the MacKerell Lab website
+(\url{http://mackerell.umaryland.edu/CHARMM_ff_params.html}) were used.
 Timestep of 1~fs was used with the leapfrog integrator. Covalent bonds with hydrogens were constrained with LINCS algorithm~\cite{hess97,hess07}. 
@@
 
-{\it POPC with 128 lipids}
-The starting structures for the pure POPC simulations was taken from the Slipids~\cite{jambeck12b} website (http://people.su.se/$\sim$jjm/Stockholm\_Lipids/Downloads.html).
+{\it POPC with 128 lipids.}
+%
+The starting structures for the pure POPC simulations was taken from the Slipids~\cite{jambeck12b} website
+(\url{http://people.su.se/$\sim$jjm/Stockholm\_Lipids/Downloads.html}).
 The starting structures for mixed POPC/Cholesterol simulations were constructed with the CHARMM-GUI website~\cite{jo08}. 
@@
 solvate the system.
-The publicly available CHARMM36 forcefield parameters (\url{http://www.gromacs.org/@api/deki/files/184/=charmm36.ff\\\_4.5.4\_ref.tgz}) 
+The publicly available CHARMM36 forcefield parameters
+(\url{http://www.gromacs.org/@api/deki/files/184/=charmm36.ff\\\_4.5.4\_ref.tgz}) 
 by Piggot et al. \cite{piggot12} were used. Cholesterol parameters came
@@
 
-{\it POPC with 512 lipids}
+{\it POPC with 512 lipids.}
+%
 POPC The starting structures for simulations were constructed with

### 2015-04-20 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
update MS

--- 2015-04-17 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
@@
 {\it POPC with 128 lipids}
-The starting structures for the pure POPC simulations was taken from the Slipids~\cite{jambeck12b} website (http://people.su.se/$\sim$jjm/Stockholm\_Lipids/Downloads.html).
+The starting structures for the pure POPC simulations was taken from the Slipids~\cite{jambeck12b} website (\url{http://mmkluster.fos.su.se/slipids/Downloads.html}).
 The starting structures for mixed POPC/Cholesterol simulations were constructed with the CHARMM-GUI website~\cite{jo08}. 
@@
 Initial coordinates for a hydrated DPPC (at 323~K) and POPC (at 310~K) bilayers (30 and 40 waters/lipid, respectively) were taken directly from the 
-Slipids home page \url{http://people.su.se/\~jjm/Stockholm\_Lipids/Downloads.html}.
-The Slipids force field~\cite{jambeck12,jambeck12b} was used for the the all atom descriptions of DPPC and POPC, and
+Slipids home page \url{http://mmkluster.fos.su.se/slipids/Downloads.html}.  The Slipids force field~\cite{jambeck12,jambeck12b} was used for the the all atom descriptions of DPPC and POPC, and
 water was described with the TIP3P water model~\cite{jorgensen83}. Simulations were performed within the NPT ensemble using the GROMACS 4.6.X simulation

### 2015-04-21 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex

@@
 for lipid bilayers, which opens up numerous possiblities for different analyses with much less effort than previously required.
-Furher information, e.g. scripts, figures and manuscript text files, are available in the GitHub 
+Further information, e.g. scripts, figures and manuscript text files, are available in the GitHub 
 organization \url{https://github.com/NMRlipids} \todo{Doi will be assigned to the GitHub repository before submission}.
@@
 Tj\"ornhammar14 \cite{tjornhammar14}   & DPPC  & 144 & 7056 & 323 & 200 & 100 & [\citenum{tjornhammarfiles}]$^*$ & [\citenum{tjornhammar14}] \\
-CHARMM36-UA~\cite{lee14}     & DPPC   & 72  & 2189  & 323 & 80 & 50 & ? & SI \\
-CHARMMXX-UA~\cite{henin08,lee14}     & DLPC   & 128  & 3840  & 323 & 30 & 20 & [\citenum{charmmUAfiles}] & SI \\
+Lee-CHARMM36-UA~\cite{henin08,lee14}     & DPPC   & 72  & 2189  & 323 & 70 & 50 & ? & SI \\
+Botan-CHARMM36-UA   & DLPC   & 128  & 3840  & 323 & 30 & 20 & [\citenum{charmmUAfiles}] & SI \\
 \end{tabular}
@@
 
-\subsubsection{CHARMM36-UA}
-A hydrated bilayer consisting of 128 DLPC lipids and 3840 water molecules is modeled by the force field of Lee and co-workers~\cite{lee14}.
-This force field is a combination of the all-atom CHARMM36 force-field~\cite{klauda10} and the united-atom Berger model~\cite{berger97}. 
-The non-bonded interactions are calculated using an atom-based switching function with short and long cut-offs of 0.8 and 1.2~nm~\cite{lee14}. 
+\subsubsection{Lee-CHARMM36-UA}
+A hydrated bilayer consisting of 72 DLPC lipids and 2189 water molecules
+ is constructed using the model by Lee et al. ~\cite{lee14}. This model describes 
+ the important all-atom CHARMM36 character of the lipid headgroup but 
+reduces the details of the lipid chains into a united-atom model.
+The initial  equilibrated structure was downloaded from the web page of J. Klauda, 
+Department of Chemical and Biomolecular Engineering University of Maryland.
+ The parameter files are taken from the Supplementary Material of Ref. \citenum{lee14}.  
+This bilayer was equilibrated for 20~ns and the production run was 50~ns long,  with data was saved every 20~ps. 
+ The equations of motion were integrated using the multiple time step Verlet r-RESPA algorithm with
+a time step of 2 fs, and a calculation of the electrostatic forces only every two timesteps.
+ Covalent bonds between heavy and hydrogen atoms were constrained 
+using SHAKE/RATTLE algorithm~\cite{miyamoto92}.
+
+The temperature was kept at 323~K with a Langevin thermostat
+with a damping coefficient of 5~ps.  The modified NAMD version of the
+Nose-Hoover barostat with Langevin dynamics (piston period of
+0.1 ps and piston decay time of 0.05 ps)  was used semi-isotropically 
+to reach the averaged target pressure of 1~bar and an averaged zero surface tension.
+PME~\cite{darden93,essman95} was employed for long range electrostatics. 
+A cut-off of 1.2~nm was employed for  the Lennard-Jones interactions, with 
+a force-based switching function  for distances beyond 1~nm.
+A neighbour list with a radius of 1.4~nm was updated every 10~timesteps.
+
+NAMD was developed by the Theoretical and Computational Biophysics Group in the
+ Beckman Institute for Advanced Science and Technology at the University
+ of Illinois at Urbana-Champaign\cite{namd}.
+
+\subsubsection{Botan-CHARMM36-UA}
+This model is similar to the Lee-CHARMM36-UA model with a few  differences.
+A hydrated bilayer consisting of 128 DLPC lipids and 3840 water
+molecules is described by a model derived from the CHARMM27-UA model by H\'enin et al.~\cite{henin08}.
+
+The distribution into all-atom  (AA) and united-atom (UA) parts within the lipids is the same
+as in the original CHARMM27-UA model by H\'enin et al.~\cite{henin08}.  
+This distribution differs from the one by Lee-C36-UA because  in Botan's model, C22 and C32  a
+nd their hydrogens are merged into  united atoms, whereas in Lee's model their hydrogens are described explicitely.
+The  united-atom Berger model ~\cite{berger97} was used for the tails and the UA-AA interactions as in Ref. ~\citenum{henin08}. 
+The difference between this model and the original oneby H\'enin~\cite{henin08} is the replacement of the 
+AA parameters for the heads by  the parameters of the all-atom
+ CHARMM36 force-field~\cite{klauda10}. Contrarily to the model by Lee et al.~\cite{lee14}, no reparametrization was done.
+
+The non-bonded interactions are calculated using an atom-based switching function with short and long cut-offs of 0.8 and 1.2~nm. 
 Long range electrostatic interactions are implemented using the particle-particle particle-mesh solver with a relative accuracy of $10^{-4}$. The system 

### 2015-04-22 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
small updates related to the CHARMM-UA

@@
 with periodic boundary conditions in all directions by using the GROMACS software package~\cite{hess08} 
-(version numbers 4.5.X--4.6.X), LAMMPS~\cite{plimpton95} or MDynaMix~\cite{Lyubartsev00}.
+(version numbers 4.5.X--4.6.X), LAMMPS~\cite{plimpton95}, MDynaMix~\cite{Lyubartsev00} or NAMD~\cite{phillips05}.
 The number of molecules, simulation temperatures and the length of simulations of all the simulated systems 
@@
 Tj\"ornhammar14 \cite{tjornhammar14}   & DPPC  & 144 & 7056 & 323 & 200 & 100 & [\citenum{tjornhammarfiles}]$^*$ & [\citenum{tjornhammar14}] \\
-Lee-CHARMM36-UA~\cite{henin08,lee14}     & DPPC   & 72  & 2189  & 323 & 70 & 50 & ? & SI \\
-Botan-CHARMM36-UA   & DLPC   & 128  & 3840  & 323 & 30 & 20 & [\citenum{charmmUAfiles}] & SI \\
+Lee-CHARMM36-UA~\cite{lee14}     & DPPC   & 72  & 2189  & 323 & 70 & 50 & ? & SI \\
+Botan-CHARMM36-UA~\cite{henin08}   & DLPC   & 128  & 3840  & 323 & 30 & 20 & [\citenum{charmmUAfiles}] & SI \\
 \end{tabular}
@@
  Beckman Institute for Advanced Science and Technology at the University
- of Illinois at Urbana-Champaign\cite{namd}.
+ of Illinois at Urbana-Champaign\cite{phillips05}.
 
@@
 as in the original CHARMM27-UA model by H\'enin et al.~\cite{henin08}.  
-This distribution differs from the one by Lee-C36-UA because  in Botan's model, C22 and C32  a
-nd their hydrogens are merged into  united atoms, whereas in Lee's model their hydrogens are described explicitely.
+This distribution differs from the one by Lee-C36-UA because  in Botan's model, C22 and C32 and 
+their hydrogens are merged into  united atoms, whereas in Lee's model their hydrogens are described explicitely.
 The  united-atom Berger model ~\cite{berger97} was used for the tails and the UA-AA interactions as in Ref. ~\citenum{henin08}. 
-The difference between this model and the original oneby H\'enin~\cite{henin08} is the replacement of the 
+The difference between this model and the original one by H\'enin~\cite{henin08} is the replacement of the 
 AA parameters for the heads by  the parameters of the all-atom

### 2015-04-22 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex

@@
 as in the original CHARMM27-UA model by H\'enin et al.~\cite{henin08}.  
-This distribution differs from the one by Lee-C36-UA because  in Botan's model, C22 and C32 and 
+This distribution differs from the one by Lee-CHARMM36-UA model solely
+for the first methylene groups of the two acyl chains (C22 and C32 in CHARMM36 topology): 
 their hydrogens are merged into  united atoms, whereas in Lee's model their hydrogens are described explicitely.

### 2015-04-22 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
updates to manuscript

@@
 %\sout {Berger-POPC-07~\cite{ollila07a}}\todoi{Not included in the Fig. 2, to be removed?}          &   POPC & 128 & 7290  & 323  & 56 & 56  & ? & SI \\
-Berger-DPPC-98~\cite{marrink98}          &   DPPC & 72 & 2864  & 323  & 140 & 100  & [\citenum{bergerDPPCfiles}] & SI \\
-Berger-DMPC-04~\cite{gurtovenko04}          &   DMPC & 128 & 5097  & 323  & 130 & 100  & [\citenum{dmpcFILES}] & [\citenum{miettinen09}] \\
+Berger-DPPC-98~\cite{marrink98}          &   DPPC & 72 & 2864  & 323  & 140 & 100  & [\citenum{bergerDPPCfiles}]$^*$ & SI \\
+Berger-DMPC-04~\cite{gurtovenko04}          &   DMPC & 128 & 5097  & 323  & 130 & 100  & [\citenum{dmpcFILES}]$^*$ & [\citenum{miettinen09}] \\
 CHARMM36~\cite{klauda10}       & POPC   & 72  &  2242 & 303 & 30 & 20  & [\citenum{charmm36filesSHORT}]$^*$ & SI \\
@@
 Tj\"ornhammar14 \cite{tjornhammar14}   & DPPC  & 144 & 7056 & 323 & 200 & 100 & [\citenum{tjornhammarfiles}]$^*$ & [\citenum{tjornhammar14}] \\
-Lee-CHARMM36-UA~\cite{lee14}     & DPPC   & 72  & 2189  & 323 & 70 & 50 & ? & SI \\
+Lee-CHARMM36-UA~\cite{lee14}     & DPPC   & 72  & 2189  & 323 & 70 & 50 & [\citenum{charmmUAfilesLEE}]$^*$ & SI \\
 Botan-CHARMM36-UA~\cite{henin08}   & DLPC   & 128  & 3840  & 323 & 30 & 20 & [\citenum{charmmUAfiles}] & SI \\
@@
 {\it Waldemar Kulig} prepared the MD simulations with cholesterol for the MacRog FF \\
-{\it Antti Lamberg}  \\
-{\it Claire Loison} provided simulation results for charmm36-UA \\
-{\it Alexander Lyubartsev} \\
+{\it Antti Lamberg} Prepared and performed simulations with Berger and its variants to show the importance of the signs and stereospecific labeling of the order parameters. \\
+{\it Claire Loison} provided simulation results for charmm36-UA. \\
+{\it Alexander Lyubartsev} provided simulation results for Högberg08 force field.\\
 {\it Markus S. Miettinen}  \\

### 2015-04-27 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Add citations to the webarchive pages of the blog

@@
 
-This work has been done as an open collaboration by using \url{nmrlipids.blogspot.fi}
-\todo{Current idea is that we will assign doi for the blog by using the help from  winnower administrators:
-https://thewinnower.com/posts/archiving-and-aggregating-alternative-scholarly-content-dois-for-blogs.
-Once done, that doi will be cited here and other places where blog content needs to be cited.} 
+This work has been done as an open collaboration by using \url{nmrlipids.blogspot.fi}~\cite{nmrlipids}
+\todo{DONE}
 as a communication platform.
@@
 All the scientific contributions are done publicly through the blog and all contributors were offered coauthorship
-according to the guidelines defined in the beginning of the project\todo{Citation to the On Credits in the blog here}. 
+according to the guidelines defined in the beginning of the project~\cite{creditsPOST}\todo{DONE}. 
 The acception of authorship is based on the self-assesment of the authors' scientific contribution to the project.
@@
 with each other (see Results and Discussion), indicating very high quantitative accuracy of the measurements.
-For a more detailed discussion see \url{http://nmrlipids.blogspot.fi/2014/02/accuracy-of-order-parameter-measurements.html}
-\todo{Citation to the blog will be added when available.}
+For a more detailed discussion see Ref.~\citenum{accuracyPOST}. \todo{DONE}
 
@@
 The measured sign is negative for almost all the carbons discussed in this work as only $\alpha$ is positive~\cite{hong95a,hong95b,gross97}. 
-For more detailed discussion about the sign measurements of the order parameters see \url{http://nmrlipids.blogspot.fi/2014/04/on-signs-of-order-parameters.html}
-\todo{Citation to the blog will be added when available.}
+For more detailed discussion about the sign measurements of the order parameters see Ref.~\citenum{signpost}. \todo{DONE}
 
@@
 accuracy and therefore we exclude them from our discussion. 
-For more details, see \url{http://nmrlipids.blogspot.fi/2014/02/accuracy-of-order-parameter-measurements.html}
-\todo{Citation to the blog will be added when available.}
+For more details, see Ref.~\citenum{accuracyPOST}. \todo{DONE}
 Motivated by the high experimental repeatability, we have highlighted in 
@@
 concentration for POPC~\cite{ferreira13}. Due to the high resolution in the NMR (especially $^2$H NMR) experiments,
-even very small order parameter changes resulting from the varying conditions can be measured (see
-\url{http://nmrlipids.blogspot.fi/2014/02/accuracy-of-order-parameter-measurements.html}
-for more discussion.)
-\todo{Citation to the blog will be added when available.}
+even very small order parameter changes resulting from the varying conditions can be measured 
+(see Ref.~\citenum{accuracyPOST} for more discussion.) \todo{DONE}
 However, as already discussed above, it is not simple to deduce 
@@
 platform. All the scientific contributions have been done through the blog and are publicly
-available. \todo{Citation to the blog will be added when available.}
+available~\cite{nmrlipids}. \todo{DONE}
 
@@
 
-\todo{All authors, please add your acknowledgements.}
+\todo{DONE}
 
@@
 \subsection{Author Contributions}
-\todo{Please write a text summarizing your contribution} \\
 {\it Alexandru Botan} provided simulation results for charmm36-UA\\
@@
 {\it Alexander Lyubartsev} provided simulation results for Högberg08 force field.\\
-{\it Markus S. Miettinen}  \\
+{\it Markus S. Miettinen} \todo{Please write a text summarizing your contribution} \\
 {\it Luca Monticelli} Critical discussions in all phases of the project. Collaboration with Jukka  M{\"a}{\"a}tt{\"a}. \\

### 2015-05-05 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Edited Table 1 + Included the modified Figure 3

@@
 
-\begin{table}[]
+\begin{sidewaystable}[]
 \centering
-\caption{Simulated single component fully hydrated lipid bilayer systems. The simulation file data sets marked with $^*$ also include a part of the trajectory.
-  If simulation data from previously published work has been directly used, the original publication is cited for simulation details. For other systems the simulation details
-  are given in the Supplementary Information.
-The abbreviations are 1-palmitoyl-2-oleoylphosphatidylcholine (POPC), dipalmitoylphosphatidylcholine (DPPC), 1,2-dimyristoyl-sn-glycero-3-phosphocholine (DMPC),
-1,2-dioleoyl-sn-glycero-3-phosphocholine (DOPC) and dilauroylphosphatidylcholine (DLPC).
-$^a$ The number of lipid molecules
-$^b$ The number of water molecules
-$^c$ Simulation temperature
-$^d$ The total simulation time
-$^e$ Time frames used in the analysis
-$^f$ Reference link for the downloadable simulation files
-$^g$ Reference for the full simulation details
+\caption{Fully hydrated single component lipid bilayer systems simulated for Fig.~\ref{HGorderparameters}:
+  1,2-dimyristoyl-sn-glycero-3-phosphocholine (DMPC),
+  %1,2-dioleoyl-sn-glycero-3-phosphocholine (DOPC),
+  dilauroylphosphatidylcholine (DLPC),
+  dipalmitoylphosphatidylcholine (DPPC), and
+  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
+The {\bf bolded} systems were used also for Fig.~\ref{HGorderparameters2}.
+$^a$ Number of lipid molecules.
+$^b$ Number of water molecules.
+$^c$ Temperature.
+$^d$ Total simulation time.
+$^e$ Time used for analysis.
+$^f$ Reference link for the downloadable simulation files;
+     the data sets marked with $^*$ also include a part of the trajectory.
+$^g$ Reference for the full simulation details;
+     the original publication is cited if simulation data from previously published work has been directly used,
+     for other systems the simulation details are given in the Supplementary Information.
 }\label{systems}
-\begin{tabular}{c c c c c c c c c}
+\begin{tabular}{l l r r r r r c c c}
 %\hline
-Force field & lipid  & $^a$N$_{\rm l}$   &  $^b$N$_{\rm w}$ &  $^c$T (K)  &  $^d$t$_{{\rm sim}}$(ns) &  $^e$t$_{{\rm anal}}$ (ns) &  $^f$Files  &  $^g$Details\\
+Force field & lipid & $^a$N$_{\rm l}$   &  $^b$N$_{\rm w}$ &  $^c$T (K)  &  $^d$t$_{{\rm sim}}$(ns) &  $^e$t$_{{\rm anal}}$ (ns) &  $^f$Files  &  $^g$Details & Notes\\
 \hline
-Berger-POPC-07~\cite{ollila07a}          &   POPC & 128 & 7290  & 298  & 270 & 240 & [\citenum{bergerFILESpopc}]$^*$ & [\citenum{ferreira15}] \\
+Berger-DMPC-04~\cite{gurtovenko04} & DMPC & 128 & 5097 & 323 & 130 & 100 & [\citenum{dmpcFILES}]$^*$       & [\citenum{miettinen09}] \\
+Berger-DPPC-98~\cite{marrink98}    & DPPC &  72 & 2864 & 323 & 140 & 100 & [\citenum{bergerDPPCfiles}]$^*$ & SI \\
+{\bf
+Berger-POPC-07}~\cite{ollila07a}   & POPC & 128 & 7290 & 298 & 270 & 240 & [\citenum{bergerFILESpopc}]$^*$ & [\citenum{ferreira15}] \\
 %\sout {Berger-POPC-07~\cite{ollila07a}}\todoi{Not included in the Fig. 2, to be removed?}          &   POPC & 128 & 7290  & 323  & 56 & 56  & ? & SI \\
-Berger-DPPC-98~\cite{marrink98}          &   DPPC & 72 & 2864  & 323  & 140 & 100  & [\citenum{bergerDPPCfiles}]$^*$ & SI \\
-Berger-DMPC-04~\cite{gurtovenko04}          &   DMPC & 128 & 5097  & 323  & 130 & 100  & [\citenum{dmpcFILES}]$^*$ & [\citenum{miettinen09}] \\
+CHARMM36~\cite{klauda10}  & DPPC &  72 & 2189 & 323 &  30 &  25 & [\citenum{charmmFILESdppc}]$^*$    & SI \\
+CHARMM36~\cite{klauda10}  & DPPC &  72 & 2189 & 323 &  ? &  130? & [?]$^*$    & ? & Magnitudes from Fig.~S4 of Klauda et al.~\cite{klauda10}, signs matched to our simulations. \\
 CHARMM36~\cite{klauda10}  & POPC &  72 & 2242 & 303 &  30 &  20 & [\citenum{charmm36filesSHORT}]$^*$ & SI \\
-CHARMM36~\cite{klauda10}      & POPC   & 128 &  5120    & 303 & 150 & 100  & [\citenum{charmm36files}]$^*$   & SI \\
-CHARMM36~\cite{klauda10}       & DPPC   & 72  &  2189 & 323 & 30 & 25  & [\citenum{charmmFILESdppc}]$^*$  & SI \\
+{\bf
+CHARMM36}~\cite{klauda10} & POPC & 128 & 5120 & 303 & 200 & 100 & [\citenum{charmm36files}]$^*$      & SI \\
 %CHARMM36\cite{klauda10}\todoi{Not included in the Fig.
@@
 MacRog~\cite{maciejewski14}  & POPC & 288 & 12600 & 310 & 100 &  80 & [\citenum{macrogFILES}]$^*$      & SI  \\
-GAFFlipid~\cite{dickson12}       & POPC & 126  & 3948  & 303 & 137 & 32  & [\citenum{GAFFlipidFILES}]$^*$ & SI \\
+{\bf
+MacRog}~\cite{maciejewski14} & POPC & 128 &  6400 & 310 & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$  & SI  \\
+MacRog~\cite{maciejewski14}  & POPC & 288 & 14400 & 310 &  90 &  40 & [\citenum{macrogdehydFILES}]$^*$ & SI  \\
 GAFFlipid~\cite{dickson12}   & DPPC &  72 & 2197 & 323 &  90 & 50 & [\citenum{GAFFlipidFILESdppc}]$^*$ & SI \\
+GAFFlipid~\cite{dickson12}   & DPPC &  ? & ? & 323 & ? &  250? & [?]$^*$    & ? & Magnitudes from Fig.~9 of Dickson et al.~\cite{dickson12}, signs matched to our simulations. \\
+{\bf
+GAFFlipid}~\cite{dickson12}  & POPC & 126 & 3948 & 303 & 137 & 32 & [\citenum{GAFFlipidFILES}]$^*$ & SI \\
 Lipid14 \cite{dickson14}         & POPC  & 72 & 2234 & 303 & 100 & 50  & [\citenum{lipid14files}]$^*$ & SI \\
@@
 Chiu \cite{chiu09}      & POPC  & 128 & 3552  & 298 & 56 & 50  & [\citenum{chiuFILES}]$^*$ & SI \\
+H\"ogberg08 \cite{hogberg08}    & DMPC &  98 & 3840 & 303 &  75 & 50 &  [\citenum{hogbergDMPCfiles}]$^*$ & [\citenum{hogberg08}] \\
 H\"ogberg08 \cite{rabinovich14} & POPC & 128 & 3840 & 303 & 100 & 80 &  [\citenum{hogbergPOPCfiles}]$^*$ & [\citenum{rabinovich14}]  \\
-H\"ogberg08 \cite{hogberg08}  & DMPC   &  98 & 3840  & 303 & 75 & 50 &  [\citenum{hogbergDMPCfiles}]$^*$ & [\citenum{hogberg08}] \\
 Ulmschneiders \cite{Ulmschneider09}    & POPC  & 128 & 3328 & 310 & 100 & 50 & [\citenum{ulmschneiderFILES}]$^*$ & SI \\
 Tj\"ornhammar14 \cite{tjornhammar14}   & DPPC  & 144 & 7056 & 323 & 200 & 100 & [\citenum{tjornhammarfiles}]$^*$ & [\citenum{tjornhammar14}] \\
+Botan-CHARMM36-UA~\cite{henin08} & DLPC & 128 & 3840 & 323 & 30 & 20 & [\citenum{charmmUAfiles}]        & SI \\
 Lee-CHARMM36-UA~\cite{lee14}     & DPPC &  72 & 2189 & 323 & 70 & 50 & [\citenum{charmmUAfilesLEE}]$^*$ & SI \\
-Botan-CHARMM36-UA~\cite{henin08}   & DLPC   & 128  & 3840  & 323 & 30 & 20 & [\citenum{charmmUAfiles}] & SI \\
 \end{tabular}
-\end{table}
+\end{sidewaystable}
 
@@
   \centering
-  \includegraphics[width=8.6cm]{../Fig/HGorderparameters5.pdf} \\
-  \todo{This figure should be updated similar to Fig.~\ref{HGorderparameters}} \\
+  \includegraphics[width=8.6cm]{../DATAreportediINblog/stereospecificOPs.pdf} \\
+  \todo{This figure should be updated similar to Fig.~\ref{HGorderparameters}. DONE.} \\
   \caption{\label{HGorderparameters2}
-  Order parameteres from simulations with Berger-POPC-07, CHARMM36, GAFFlipid and MacRog force fields together with experimental values for POPC glycerol and choline groups.
-  The magnitudes for experimental order parameters are taken from Ferreira et al.~\cite{ferreira13}, the signs are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
-  and Gross et al.~\cite{gross97}, and the R/S labeling is based in the measurements by Gally et al.~\cite{gally81}.
+  Order parameteres for POPC glycerol and choline groups
+  from simulations with Berger-POPC-07, MacRog, GAFFlipid, and CHARMM36 force fields
+  (the {\bf bolded} systems in Table~\ref{systems})
+  together with experimental values.
+  The error bars of simulation data are standard errors of mean (see Methods for details).
+  The magnitudes for experimental order parameters are taken from Ferreira et al.~\cite{ferreira13},
+  the signs are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
+  and Gross et al.~\cite{gross97}, and the R/S labeling is based on the measurements by Gally et al.~\cite{gally81}.
 }

### 2015-05-07 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
updates in manuscript

@@
 This work has been done as an open collaboration by using \url{nmrlipids.blogspot.fi}~\cite{nmrlipids}
-\todo{DONE}
 as a communication platform.
@@
 All the scientific contributions are done publicly through the blog and all contributors were offered coauthorship
-according to the guidelines defined in the beginning of the project~\cite{creditsPOST}\todo{DONE}. 
+according to the guidelines defined in the beginning of the project~\cite{creditsPOST}.
 The acception of authorship is based on the self-assesment of the authors' scientific contribution to the project.
@@
 Further information, e.g. scripts, figures and manuscript text files, are available in the GitHub 
-organization \url{https://github.com/NMRlipids} \todo{Doi will be assigned to the GitHub repository before submission}.
+organization \url{https://github.com/NMRlipids} \todo{I will not assign a doi yet. We can wait for the journal decision.}.
 
@@
 with each other (see Results and Discussion), indicating very high quantitative accuracy of the measurements.
-For a more detailed discussion see Ref.~\citenum{accuracyPOST}. \todo{DONE}
+For a more detailed discussion see Ref.~\citenum{accuracyPOST}.
 
@@
 The measured sign is negative for almost all the carbons discussed in this work as only $\alpha$ is positive~\cite{hong95a,hong95b,gross97}. 
-For more detailed discussion about the sign measurements of the order parameters see Ref.~\citenum{signpost}. \todo{DONE}
+For more detailed discussion about the sign measurements of the order parameters see Ref.~\citenum{signpost}.
 
@@
      for other systems the simulation details are given in the Supplementary Information.
+$^h$ Magnitudes from Fig.~S4 of Klauda et al.~\cite{klauda10}, signs matched to our simulations.
+$^i$ Magnitudes from Fig.~9 of Dickson et al.~\cite{dickson12}, signs matched to our simulations.
 }\label{systems}
-\begin{tabular}{l l r r r r r c c c}
+\begin{tabular}{l l r r r r r c c}
 %\hline
-Force field & lipid & $^a$N$_{\rm l}$   &  $^b$N$_{\rm w}$ &  $^c$T (K)  &  $^d$t$_{{\rm sim}}$(ns) &  $^e$t$_{{\rm anal}}$ (ns) &  $^f$Files  &  $^g$Details & Notes\\
+Force field & lipid & $^a$N$_{\rm l}$   &  $^b$N$_{\rm w}$ &  $^c$T (K)  &  $^d$t$_{{\rm sim}}$(ns) &  $^e$t$_{{\rm anal}}$ (ns) &  $^f$Files  &  $^g$Details\\
 \hline
@@
 Berger-POPC-07}~\cite{ollila07a}   & POPC & 128 & 7290 & 298 & 270 & 240 & [\citenum{bergerFILESpopc}]$^*$ & [\citenum{ferreira15}] \\
-%\sout {Berger-POPC-07~\cite{ollila07a}}\todoi{Not included in the Fig. 2, to be removed?}          &   POPC & 128 & 7290  & 323  & 56 & 56  & ? & SI \\
 CHARMM36~\cite{klauda10}  & DPPC &  72 & 2189 & 323 &  30 &  25 & [\citenum{charmmFILESdppc}]$^*$    & SI \\
-CHARMM36~\cite{klauda10}  & DPPC &  72 & 2189 & 323 &  ? &  130? & [?]$^*$    & ? & Magnitudes from Fig.~S4 of Klauda et al.~\cite{klauda10}, signs matched to our simulations. \\
+CHARMM36~\cite{klauda10}  & DPPC &  72 & 2189 & 323 &  -  & 130 & -   & [\citenum{klauda10}]$^h$  \\
 CHARMM36~\cite{klauda10}  & POPC &  72 & 2242 & 303 &  30 &  20 & [\citenum{charmm36filesSHORT}]$^*$ & SI \\
@@
 GAFFlipid~\cite{dickson12}   & DPPC &  72 & 2197 & 323 &  90 & 50 & [\citenum{GAFFlipidFILESdppc}]$^*$ & SI \\
-GAFFlipid~\cite{dickson12}   & DPPC &  ? & ? & 323 & ? &  250? & [?]$^*$    & ? & Magnitudes from Fig.~9 of Dickson et al.~\cite{dickson12}, signs matched to our simulations. \\
+GAFFlipid~\cite{dickson12}   & DPPC &  72 & 2167 & 323 & 250 &  250 & -   &  [\citenum{dickson12}]$^i$   \\
 {\bf
@@
   \includegraphics[width=8.6cm]{../DATAreportediINblog/comparisonSorted.pdf}
-%\todo{Markus Miettinen suggested that we should consider making one more figure where only experimental data would be shown and that would be discussed in Section~\ref{experiments}. \\
-%COMMENT: I do not think that we need such figure anymore since this figure is more clear now.} \\
 \newline
-\todo{There is also the interactive version by Hubert Santuz now in https://plot.ly/~HubertSantuz/72/lipid-force-field-comparison/
-we should figure out which is the most practical way to put that behind permalink once it finalized (Zenodo, figshare or something else?) and then put a citation in the paper.}
+%\todo{There is also the interactive version by Hubert Santuz now in https://plot.ly/~HubertSantuz/72/lipid-force-field-comparison/
+%we should figure out which is the most practical way to put that behind permalink once it finalized (Zenodo, figshare or something else?) and then put a citation in the paper.}
   \caption{\label{HGorderparameters}
@@
 the ends of the bars mark the extreme values from the sets, and the dot marks their measurement-time-weighted average. 
+The interactive version of this figure is available at  https://plot.ly/$\sim$HubertSantuz/72/lipid-force-field-comparison/.
 } 
@@
 accuracy and therefore we exclude them from our discussion. 
-For more details, see Ref.~\citenum{accuracyPOST}. \todo{DONE}
+For more details, see Ref.~\citenum{accuracyPOST}. 
 Motivated by the high experimental repeatability, we have highlighted in 
@@
   \includegraphics[width=8.6cm]{../DATAreportediINblog/stereospecificOPs.pdf} \\
-  \todo{This figure should be updated similar to Fig.~\ref{HGorderparameters}. DONE.} \\
   \caption{\label{HGorderparameters2}
@@
 even very small order parameter changes resulting from the varying conditions can be measured 
-(see Ref.~\citenum{accuracyPOST} for more discussion.) \todo{DONE}
+(see Ref.~\citenum{accuracyPOST} for more discussion.) 
 However, as already discussed above, it is not simple to deduce 
@@
   \includegraphics[width=8.6cm]{../DATAreportediINblog/dehydration.pdf} \\
-  \todo{DONE} \\
     \caption{\label{ordPhydr}
@@
 for all the segmets.
-\todo{DONE}
 
@@
 the structural interpretation.
-\todo{DONE}
 
@@
   \includegraphics[width=8.6cm]{../DATAreportediINblog/cholesterolization.pdf} \\
-  \todo{DONE} \\
   \caption{\label{ordPchol}
@@
 platform. All the scientific contributions have been done through the blog and are publicly
-available~\cite{nmrlipids}. \todo{DONE}
+available~\cite{nmrlipids}.
 
@@
 
-\todo{DONE}
-
-OHSO acknowledges Tiago Ferreira and Paavo Kinnunen for useful discussions, 
-the Emil Aaltonen foundation for financial support, 
-Aalto Science -- IT project and CSC -- IT Center for Science for computational resources.
-
-MSM acknowledges financial support from the Volkswagen Foundation (86110).
-
-WK acknowledges CSC -- IT Centre for Science (Espoo, Finland) is acknowledged for excellent computational resources (project number tty3995).
-
-JM acknowledges CSC -- IT Center for Science for computational resources.
-
+AB and CL acknowledges financial support from the French National Research Agency
+(ANR: “Biolubrication by phospholipid membranes” Bioub2012) and
+computing time allocation from P{\^o}le Scientifique de Mod{\'e}lisation Num{\'e}rique from the ENS Lyon (PSMN),
+and Centre Informatique National de l'Enseignement Sup{\'e}rieur (CINES, Montpelier, France)
+(Project c2015096850).
+FFR acknowledges CONACYT and DGAPA UNAM IG100513
+for financial support, Cluster H\'ibrido de Superc\'omputo Xiuhcoatl - CINVESTAV and Miztli - UNAM for computational resources. 
 MJ and JT acknowledge and CSC - IT Center for Science for computational resources (project number tty3979).
@@
 and the European Research Council (Advanced Grant project CROWDED-PRO-LIPIDS).
-
-CL and AB acknowledges financial support from the French National Research Agency
-(ANR: “Biolubrication by phospholipid membranes” Bioub2012) and
-computing time allocation from P{\^o}le Scientifique de Mod{\'e}lisation Num{\'e}rique from the ENS Lyon (PSMN),
-and Centre Informatique National de l'Enseignement Sup{\'e}rieur (CINES, Montpelier, France)
-(Project c2015096850).
-
-HS acknowleges Catherine Etchebest and St\'ephane T\'eletch\'ea for useful discussions and continued support, the HPC resources granted from GENCI-CINES (Grant 2014-c2014077209) and computer facilities provided by R\'egion Ile de France and INTS (SESAME 2009 project).
-
-Fernando Favela acknowledges CONACYT and DGAPA UNAM IG100513
-for financial support, Cluster H\'ibrido de Superc\'omputo Xiuhcoatl - CINVESTAV and Miztli - UNAM for computational resources.
-
-Luca Monticelli acknowledges funding proved by the Institut national de la sant\'e et de la recherche m\'edicale (INSERM).
+WK acknowledges CSC -- IT Centre for Science (Espoo, Finland) is acknowledged for excellent computational resources (project number tty3995).
+MSM acknowledges financial support from the Volkswagen Foundation (86110).
+LM acknowledges funding proved by the Institut national de la sant\'e et de la recherche m\'edicale (INSERM).
+JM acknowledges CSC -- IT Center for Science for computational resources.
+OHSO acknowledges Tiago Ferreira and Paavo Kinnunen for useful discussions, 
+the Emil Aaltonen foundation for financial support, 
+Aalto Science -- IT project and CSC -- IT Center for Science for computational resources. 
+HS acknowleges Catherine Etchebest and St\'ephane T\'eletch\'ea for useful discussions and continued support, the HPC 
+resources granted from GENCI-CINES (Grant 2014-c2014077209) and computer facilities provided by R\'egion Ile de France and INTS (SESAME 2009 project).
 
@@
 
-%\subsection{General issues to think of about the manuscript}
-%\todo{Maybe we should mention that CHARMM is significanlty more computationally expensive than other all-atom models.
-%This seems evident for a person who has ran CHARMM and other all-atom models with the same computer. However, there
-%may be some people who have never used any other model than CHARMM. In this case it may be difficult to realize.
-%Would anyone have a good bencmark about this? I have assumed that the reason for the slowness are the extra Lennart-Jones
-%interactionsi in water, however, I have never looked into it. Would someone have some hard facts on this?}
-
 %\newpage

### 2015-05-07 HGmodelMANUSCRIPT/coverLETTER.tex
Update coverLETTER.tex

@@
 by A. Botan et al. In the manuscript we combine extensive amount of experimental Nuclear Magnetic Resonance (NMR)
-and molecular dynamics simulation data to understand the atomistic resolution structure of biologically abundant phospholipid molecules.
+and molecular dynamics simulation data to understand the atomistic resolution structure of phospholipid molecules and their assemblies.
 
@@
 JACS since the used combination of experiments and simulations is shown to be successfull in
-resolving the atomistic resolution structure of biomolecules in the various biologically relevant conditions.
+resolving the atomistic resolution structure of biomolecules in various biologically relevant conditions.
 The approach is demonstrated for lipids, however, the extension to, e.g. membrane proteins is straightforward. 
-The understanding of the atomistic resolution structures of molecules in biologically relevant conditions 
-significantly advances understanding about fundamental chemistry and also supports the development of, 
-for example, new biomaterials or drug delivery systems.
+Such an atomistic resolution of  biomolecular structure
+significantly advances the understanding in fundamental chemistry, molecular biology, and also supports the development of applications, 
+for example new biomaterials or drug delivery systems.
 
@@
 The usage of molecular dynamics simulations and open collaboration are adequate for the
-problem at hand since the simulations are currently the most straightforward method to 
-construct the atomistic resolution structure of molecules to interpret the NMR experiments. On the other hand,
-enormous amount of simulation work was required for this work, and the only practical
-solution to collect this data is the used open collaboration approach.
+problem at hand. Indeed, such simulations are currently the most straightforward method to 
+construct atomistic resolution structures to interpret NMR experiments, but
+enormous amount of simulation work was required for this work. The only practical
+solution to collect this data is the open collaboration.
 As result, we have produced original chemical insight for the atomistic resolution 

### 2015-05-07 HGmodelMANUSCRIPT/coverLETTER.tex
update cover letter

@@
 by A. Botan et al. In the manuscript we combine extensive amount of experimental Nuclear Magnetic Resonance (NMR)
-and molecular dynamics simulation data to understand the atomistic resolution structure of phospholipid molecules and their assemblies.
+and molecular dynamics simulation data to understand the atomistic resolution structure of biologically abundant 
+phospholipid molecules and their assemblies.
 

### 2015-05-07 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Mention data management in Hubert Santuz contribution
typo fix

@@
 and the European Research Council (Advanced Grant project CROWDED-PRO-LIPIDS).
-WK acknowledges CSC -- IT Centre for Science (Espoo, Finland) is acknowledged for excellent computational resources (project number tty3995).
+WK acknowledges CSC -- IT Centre for Science (Espoo, Finland) for excellent computational resources (project number tty3995).
 MSM acknowledges financial support from the Volkswagen Foundation (86110).
@@
 {\it Tomasz Rog} Provided the topologies for the MacRog model and also the full hydration simulation data with this model. \\
-{\it Hubert Santuz} prepared and performed the cholesterol simulations with CHARMM36 and analyzed the results. \\
+{\it Hubert Santuz} prepared and performed the cholesterol simulations with CHARMM36 and analyzed the results. 
+Made significant contribution in the data management of the project.\\
 {\it Joona Tynkkynen} prepared and performed the dehydration simulations with the MacRog FF.\\

### 2015-05-11 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Fixed simulation times of Berger-DPPC-98 in Table 1 to reported (60/30ns): http://nmrlipids.blogspot.com/2014/11/new-version-of-manuscript.html?showComment=1421231975768#c3582830273528800239
udpate CHARMM36 liteture information in table

@@
 Berger-DMPC-04~\cite{gurtovenko04} & DMPC & 128 & 5097 & 323 & 130 & 100 & [\citenum{dmpcFILES}]$^*$       & [\citenum{miettinen09}] \\
-Berger-DPPC-98~\cite{marrink98}    & DPPC &  72 & 2864 & 323 & 140 & 100 & [\citenum{bergerDPPCfiles}]$^*$ & SI \\
+Berger-DPPC-98~\cite{marrink98}    & DPPC &  72 & 2864 & 323 & 60 & 30 & [\citenum{bergerDPPCfiles}]$^*$ & SI \\
 {\bf
@@
 CHARMM36~\cite{klauda10}  & DPPC &  72 & 2189 & 323 &  30 &  25 & [\citenum{charmmFILESdppc}]$^*$    & SI \\
-CHARMM36~\cite{klauda10}  & DPPC &  72 & 2189 & 323 &  -  & 130 & -   & [\citenum{klauda10}]$^h$  \\
+CHARMM36~\cite{klauda10}  & DPPC &  72 & 2189 & 323 &  130  & - & -   & [\citenum{klauda10}]$^h$  \\
 CHARMM36~\cite{klauda10}  & POPC &  72 & 2242 & 303 &  30 &  20 & [\citenum{charmm36filesSHORT}]$^*$ & SI \\
@@
 %\todoi{Permanent link for DOPC CHARMM36 simulation by Hubert Santuz is not required now. However, it would be probably useful in the future.}
-MacRog~\cite{maciejewski14}  & POPC & 288 & 12600 & 310 & 100 &  80 & [\citenum{macrogFILES}]$^*$      & SI  \\
+MacRog~\cite{kulig15b}  & POPC & 288 & 12600 & 310 & 100 &  80 & [\citenum{macrogFILES}]$^*$      & SI  \\
 {\bf
-MacRog}~\cite{maciejewski14} & POPC & 128 &  6400 & 310 & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$  & SI  \\
-MacRog~\cite{maciejewski14}  & POPC & 288 & 14400 & 310 &  90 &  40 & [\citenum{macrogdehydFILES}]$^*$ & SI  \\
+MacRog}~\cite{kulig15b} & POPC & 128 &  6400 & 310 & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$  & SI  \\
+MacRog~\cite{kulig15b}  & POPC & 288 & 14400 & 310 &  90 &  40 & [\citenum{macrogdehydFILES}]$^*$ & SI  \\
 GAFFlipid~\cite{dickson12}   & DPPC &  72 & 2197 & 323 &  90 & 50 & [\citenum{GAFFlipidFILESdppc}]$^*$ & SI \\
@@
                             & POPC   & 7  & 72  &  504  & 303 & 60 & 20 & [\citenum{charmm36files7wPERl}]$^*$ & SI \\
-MacRog\cite{maciejewski14}     & POPC   & 50 & 288  & 14400 & 310 & 90 & 40 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
+MacRog\cite{kulig15b}     & POPC   & 50 & 288  & 14400 & 310 & 90 & 40 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 25 & 288  & 7200 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
@@
                                & POPC   & 80 & 80 &50\%  &  4496    & 303 & 200 & 100 & [\citenum{charmm36files50perCHOL}]$^*$ & SI \\
-MacRog\cite{maciejewski14}     & POPC   & 128 & 0 & 0\% & 6400  & 310 & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI \\ 
+MacRog\cite{kulig15b}     & POPC   & 128 & 0 & 0\% & 6400  & 310 & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI \\ 
                           & POPC   & 114  & 14 & 11\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
@@
 \subsubsection{MacRog}
-The lipid force field parameters were obtained from the developers and they correspond to the published DPPC parameters~\cite{maciejewski14} with the inclusion of the 
-double bond parameters~\cite{kulig15}. 
+The lipid force field parameters were obtained from the developers and the parameters are presented in Refs.~\citenum{maciejewski14,kulig15b}. 
+%correspond to the published DPPC parameters~\cite{maciejewski14} with the inclusion of the double bond parameters~\cite{kulig15}. 
 %This inclusion of unsaturated lipid tails will be published in the near future. 

### 2015-05-11 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Two new versions of the Cholesterization plot. Which one to use?

@@
   \centering
-  \includegraphics[width=8.6cm]{../DATAreportediINblog/cholesterolization.pdf} \\
+  \includegraphics[width=7.3cm]{../DATAreportediINblog/cholesterolization_lp.pdf}
+    \includegraphics[width=7.3cm]{../DATAreportediINblog/cholesterolization_smooth.pdf}
   \caption{\label{ordPchol}
+	\todo{Markus prepared two versions of this plot. The one on left has all the data points joint by lines;
+	the one on right shows smooth trend lines (cubic polynomials) to ease qualitative comparison.
+	Samuli likes the one on left, Markus the one on right. Do others have opinions?}
     The effect of cholesterol content on the glycerol backbone and choline order parameters in experiments~\cite{brown78,ferreira13} and simulations

### 2015-05-11 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
update Fig. 9 suggestion into the manuscript

--- 2015-05-11 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
@@
   \centering
-  \includegraphics[width=8.6cm]{../DATAreportediINblog/cholesterolization.pdf} \\
+  \includegraphics[width=7.3cm]{../DATAreportediINblog/cholesterolization_lp.pdf}
+    \includegraphics[width=7.3cm]{../DATAreportediINblog/cholesterolization_smooth.pdf}
   \caption{\label{ordPchol}
@@
 }
+	\todo{Markus prepared two versions of this plot. The one on left has all the data points joint by lines;
+	the one on right shows smooth trend lines (cubic polynomials) to ease qualitative comparison.
+	Samuli likes the one on left, Markus the one on right. Do others have opinions?}
 \end{figure}

### 2015-05-12 POPClipid14/makeELECTRONSdat.sh
Cleaning repository

@@ -1,3 +1,0 @@
-grep " 1POPC" endSTRUCTURE.gro | awk '{print $2}' > tmp.dat
-cat tmp.dat | awk '{if($0~"C") print $1" =6"; if($0~"H") print $1" =1"; if($0~"O") print $1" =6"; if($0~"P") print $1" =15"; if($0~"N") print $1" =7";
-}' > electrons.dat

### 2015-05-13 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Organize stereospecific Figure
Gnuplot script (gpl) and pdf output have been move the Fig folder.
A Makefile entry has been made to reproduce the figure
The tex source has been update with the new location
Two versions of dehydration plot.
Streamlined title.

@@
 \title[An \textsf{achemso} demo]
-  {Towards atomistic resolution structure of phosphatidylcholine glycerol backbone and choline headgroup at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}}
+  {Towards atomistic resolution structure of phosphatidylcholine headgroup and glycerol backbone at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}}
 
@@
   \centering
-  \includegraphics[width=8.6cm]{../DATAreportediINblog/stereospecificOPs.pdf} \\
+  \includegraphics[width=8.6cm]{../Fig/stereospecificOPs.pdf} \\
   \caption{\label{HGorderparameters2}
@@
   \centering
-  \includegraphics[width=8.6cm]{../DATAreportediINblog/dehydration.pdf} \\
+  \includegraphics[width=7.3cm]{../DATAreportediINblog/dehydration_lp.pdf}
+    \includegraphics[width=7.3cm]{../DATAreportediINblog/dehydration_smooth.pdf}
     \caption{\label{ordPhydr}
@@
   }
+  	\todo{Markus prepared two versions of this plot. The one on left has all the data points joint by lines;
+	the one on right shows smooth trend lines (cubic polynomials) to ease qualitative comparison.
+	Samuli likes the one on left, Markus the one on right. Do others have opinions?}
 \end{figure}

### 2015-05-13 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Fix issue https://github.com/NMRLipids/nmrlipids.blogspot.fi/issues/41 and increase font also in Fig. 5
Small change to test PR.

@@
   \centering
-  \includegraphics[width=8.6cm]{../Fig/g1-g2_Cdihs2.pdf}
+  \includegraphics[width=8.6cm]{../Fig/g1-g2_Cdihs2.eps}
   \caption{\label{dihDISTS}
@@
   \centering
-  \includegraphics[width=8.6cm]{../Fig/a-bDIHS2.pdf}
+  \includegraphics[width=8.6cm]{../Fig/a-bDIHS2.eps}
   \caption{\label{dihDISTS2}
@@
 Indeed, the P--N (phoshate phosphorus to choline nitrogen) vector angle with membrane normal shows an increase for
-all models as a function of dehydration in Fig.~\ref{PNangle}.
-However, the amount of increase depends on the model. Especially the DLPC simulations with Berger model
+all models as a function of dehydration (Fig.~\ref{PNangle}). However,
+the amount of increase depends on the model. Especially the DLPC simulations with Berger model
 predict significantly stronger P--N vector tilt compared with the other models. The Berger model

### 2015-05-13 HGmodelMANUSCRIPT/coverLETTER.tex

--- 2015-05-07 HGmodelMANUSCRIPT/coverLETTER.tex
@@
 
-To Whom It May Concern,
+Dear Editor,
 
-Please find attached to this message a manuscript that we would like to submit for publication 
-in \textit{Journal of the american chemical society} (JACS): 
-`Towards atomistic resolution structure of phosphatidylcholine glycerol backbone and choline headgroup at different ambient conditions', 
-by A. Botan et al. In the manuscript we combine extensive amount of experimental Nuclear Magnetic Resonance (NMR)
-and molecular dynamics simulation data to understand the atomistic resolution structure of phospholipid molecules and their assemblies.
+Please find attached to this message a manuscript entitled `Towards atomistic resolution structure of phosphatidylcholine 
+glycerol backbone and choline headgroup at different ambient conditions' by A. Botan et al., which we would like to submit 
+for consideration of publication in \textit{Journal of the American Chemical Society} (JACS). In the manuscript a large amount of
+existing experimental Nuclear Magnetic Resonance (NMR) data is reviewed and combined with an exceptionally comprehensive 
+collection of generated molecular dynamics (MD) simulation data in order to understand the atomistic resolution structure 
+of biologically abundant phospholipid molecules and their assemblies.
 
-We believe that the manuscript is interesting for the wide and diverse contemporary audience of
-JACS since the used combination of experiments and simulations is shown to be successfull in
-resolving the atomistic resolution structure of biomolecules in various biologically relevant conditions.
-The approach is demonstrated for lipids, however, the extension to, e.g. membrane proteins is straightforward. 
-Such an atomistic resolution of  biomolecular structure
-significantly advances the understanding in fundamental chemistry, molecular biology, and also supports the development of applications, 
+We believe that the manuscript is interesting to the wide and diverse contemporary audience of
+JACS as it demonstrates how the combination of experimental NMR data with MD simulations is shown 
+to be successful in resolving the atomistic resolution structure of biomolecules in various biologically 
+relevant conditions. The approach is demonstrated for lipids, however, the extension to, e.g. membrane 
+proteins is straightforward. Such an atomistic resolution of biomolecular structure significantly advances 
+the understanding in fundamental chemistry, molecular biology, and also supports the development of applications, 
 for example new biomaterials or drug delivery systems.
 
+In addition to its main message, the manuscript also provides an extensive comparison on the quality of the head group 
+structures of numerous widely used lipid models in varying biologically relevant conditions. The behavior
+of the lipid head group region plays a key role in multiple processes, such as partitioning of drugs or 
+nanoparticles into membranes and interactions of peripheral proteins with membranes, both of which are 
+currently intensively studied. Therefore the manuscript provides crucial information for a large group
+of computational scientists who are struggling to select the best possible lipid model for the problem at hand.
+
 The work has been conducted by using a novel open collaboration concept, and all the scientific contributions
-are done publicly through the nmrlipids.blogspot.fi. Thus all the scientific content related to this work
-(including raw data, discussions, manuscript drafts, etc.)
-has been publicly available during the whole course of the project through the blog. This material and discussion possibility
-will be kept available also after the publication. This approach has significantly increased and will increase the scientific
-quality and impact of the work. In addition, we have shared all the simulation raw data in such a format
-that it can be reanalyzed easily also for other purposes. In addition to the scientic content,
-we believe that the success of this approach in the field of chemistry is highly interesting 
-for the JACS audience.
+are made publicly through the blog-based project web page located at \texttt{nmrlipids.blogspot.fi}. 
+Thus all the scientific content related to this work (including raw data, discussions, manuscript drafts, etc.)
+has been publicly available during the whole course of the project through this blog. This material and the possibility to 
+discuss it on the blog page will remain available also after the publication. This approach has significantly 
+increased and will further increase the scientific quality and impact of the work. In addition, we have shared 
+all the simulation raw data in a format that allows easy reanalysis and therefore reuse of the data for other purposes. 
+In addition to the scientic content, we believe that the success of this fresh approach in the field of chemistry 
+is highly interesting to the audience of JACS.
 
@@
 enormous amount of simulation work was required for this work. The only practical
-solution to collect this data is the open collaboration.
+solution to collect this data was the open collaboration concept.
 As result, we have produced original chemical insight for the atomistic resolution 
@@
 
-We hope that you can consider this manuscript to be published as an article in the Journal of the american chemical society. \\
+We hope that you can consider this manuscript to be published as an article in the Journal of the American Chemical Society. \\
 

### 2015-05-13 HGmodelMANUSCRIPT/coverLETTER.tex
Update coverLETTER.tex

@@
-\documentclass[12pt]{letter}
+\documentclass[11pt]{letter}
 \usepackage{graphicx,baskervald,microtype}
@@
 currently intensively studied. Therefore the manuscript provides crucial information for a large group
-of computational scientists who are struggling to select the best possible lipid model for the problem at hand.
+of computational scientists who are struggling to select the best possible lipid model for their research.
 

### 2015-05-13 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Changes after fast read-through.

@@
 for the glycerol backbone and choline as a function of hydration level are shown in Fig.~\ref{ordPhydr}. 
-The independently reported values for choline segments are in good agreement with each other (despite 
+The independently reported values for the choline segments are in good agreement with each other (despite 
 slight differences in temperature and acyl chain composition),
-showing increase for both segments with decreasing hydration level. It should be noted that only 
+and both ($\beta$ and $\alpha$) show increase with decreasing hydration level. It should be noted that only 
 absolute values were measured in the original experiments~\cite{dvinskikh05b,ulrich94,bechinger91}, but
@@
 since the absolute value decreases~\cite{dvinskikh05b,ulrich94,bechinger91}.
-Slight decrease for the glycerol backbone order parameters were observed with dehydration~\cite{dvinskikh05b}.
+Slight decrease for the glycerol backbone order parameters have been observed with dehydration~\cite{dvinskikh05b}.
 \begin{figure}[]
@@
 typically motivated by the  discussion about the origin of the ``hydration repulsion''~\cite{israelachvili,israelachvili96,sparr11}.
-However, the used simulation models are not typically compared to the experimental choline and glycerol backbone
+However, the used simulation models are typically not compared to the experimental choline and glycerol backbone
 order parameters (except by Mashl et al.~\cite{mashl01}).
@@
 The magnitude of shifting is written in the y-axis label.
-The choline order parameter increase with dehydration is seen in all
+The increase of the choline order parameters with dehydration is seen in all
 models despite of some additional fluctuations. Thus the choline order parameter response to dehydration can be
@@
 has also generally larger P--N vector angles and its choline order parameters are more off from
-experiments than other models. Thus the relatively modest tilting with dehydration
+experiments than the other three models. Thus the relatively modest tilting with dehydration
 predicted by MacRog, CHARMM36 and GAFFlipid is probably more realistic.
@@
   \caption{\label{PNangle}
-    The angle between membrane normal and P--N vector as function of
+    The average angle between membrane normal and P--N vector as function of
     hydration level calculated from different simulations.

### 2015-05-14 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex

@@
 \subsubsection{Phospholipid bilayer with low hydration level}
-The experimental order parameters available in the literature~\cite{dvinskikh05b,ulrich94,bechinger91} 
-for the glycerol backbone and choline as a function of hydration level are shown in Fig.~\ref{ordPhydr}. 
-The independently reported values for the choline segments are in good agreement with each other (despite 
-slight differences in temperature and acyl chain composition),
-and both ($\beta$ and $\alpha$) show increase with decreasing hydration level. It should be noted that only 
-absolute values were measured in the original experiments~\cite{dvinskikh05b,ulrich94,bechinger91}, but
-we have included the signs measured in separate studies~\cite{hong95a,hong95b,gross97}. 
-Consequently, the $\beta$ order parameter with negative sign actually increases with dehydration 
-since the absolute value decreases~\cite{dvinskikh05b,ulrich94,bechinger91}.
-Slight decrease for the glycerol backbone order parameters have been observed with dehydration~\cite{dvinskikh05b}.
+Fig.~\ref{ordPhydr} shows the published~\cite{dvinskikh05b,ulrich94,bechinger91} experimental order parameters
+for the glycerol backbone and choline as a function of hydration level.
+Despite slight differences in temperature and acyl chain composition,
+the three independently reported data sets for the choline ($\beta$ and $\alpha$) segments agree well with each other:
+Both order parameters increase with decreasing hydration level.
+The glycerol backbone order parameters (g$_{3}$, g$_{2}$, g$_{1}$), in contrast,
+have been observed~\cite{dvinskikh05b} to slightly decrease with dehydration.
+Note that the original experiments~\cite{dvinskikh05b,ulrich94,bechinger91} measured only 
+absolute values, but here
+we included the signs measured in separate studies~\cite{hong95a,hong95b,gross97}. 
+Consequently, the negative $\beta$ order parameter actually increases with dehydration 
+as its absolute value decreases~\cite{dvinskikh05b,ulrich94,bechinger91}.
 \begin{figure}[]
@@
 Lipid bilayer dehydration has been studied also with molecular dynamics simulations~\cite{mashl01,pertsin05,pertsin07,eun09,eun10,schneck12},
-typically motivated by the  discussion about the origin of the ``hydration repulsion''~\cite{israelachvili,israelachvili96,sparr11}.
-However, the used simulation models are typically not compared to the experimental choline and glycerol backbone
-order parameters (except by Mashl et al.~\cite{mashl01}).
-In Fig.~\ref{ordPhydr} the glycerol backbone and choline order parameters are shown as a function of hydration level for the CHARMM36, 
-MacRog and GAFFlipid models (having the most realistic atomistic resolution structures) together with the Berger based model 
-(which is the most used lipid model). 
-To elucidate changes the simulation results are shifted to match better with experiments with full hydration.
-The magnitude of shifting is written in the y-axis label.
-The increase of the choline order parameters with dehydration is seen in all
-models despite of some additional fluctuations. Thus the choline order parameter response to dehydration can be
-intepreted to be in qualitative agreement with experiments.
-The situation is significantly more complicated for the glycerol backbone segments 
-and none of the models can be interpreted to qualitatively agree with experiments
-for all the segmets.
-
-The qualitative agreement with experiments in all simulation models for the $\alpha$ and $\beta$ order parameters  
-as a function of hydration indicates that the structural response of the choline headgroup to dehydration is somewhat realistic
-despite the unrealistic structures at full hydration. 
-The most likely explanation is that the choline group
-orients more parallel to the membrane plane with dehydration due to restricted interlamellar space. 
-Indeed, the P--N (phoshate phosphorus to choline nitrogen) vector angle with membrane normal shows an increase for
-all models as a function of dehydration (Fig.~\ref{PNangle}). However,
+typically motivated by the  discussion concerning the origin of the ``hydration repulsion''~\cite{israelachvili,israelachvili96,sparr11}.
+Only one~\cite{mashl01} of these studies, however, compared their simulation model to the experimental choline and glycerol backbone
+order parameters.
+Fig.~\ref{ordPhydr} shows these order parameters as a function of hydration level for the CHARMM36, 
+MacRog and GAFFlipid models (having the most realistic atomistic resolution structures) and a Berger-based model 
+(which is the most used lipid model);
+note that the simulation results have been vertically shifted to ease the comparison with experimental response to dehydration.
+Despite of some fluctuations, the increase of the choline ($\beta$ and $\alpha$) order parameters is seen in all four
+models. The response of the choline order parameters to dehydration can, therefore, be
+interpreted to qualitatively agree with experiments.
+The situation is significantly more complicated for the glycerol backbone: 
+None of the four models produced the experimentally seen trends
+in all the (g$_{3}$, g$_{2}$, g$_{1}$) segments.
+
+The qualitative agreement of the $\alpha$ and $\beta$ order parameters with experiments in all four simulation models (Fig.~\ref{ordPhydr}) indicates that,
+despite the unrealistic structures at full hydration (Figs.~\ref{HGorderparameters} and~\ref{FullHydrationComparisonTable}),
+the structural response of the choline headgroup to dehydration is somewhat realistic.
+A likely explanation is that as the interlamellar space shrinks with dehydration, the whole choline group
+orients more parallel to the membrane.
+Indeed, upon dehydration the angle between P--N (phoshate phosphorus to choline nitrogen) vector and membrane normal increases for
+all the four models (Fig.~\ref{PNangle}). However,
 the amount of increase depends on the model. Especially the DLPC simulations with Berger model
-predict significantly stronger P--N vector tilt compared with the other models. The Berger model
+predict significantly stronger P--N vector tilt than the other models. The Berger model
 has also generally larger P--N vector angles and its choline order parameters are more off from
-experiments than the other three models. Thus the relatively modest tilting with dehydration
+experiments than the other three models (Fig.~\ref{HGorderparameters2}). Thus the relatively modest tilting with dehydration
 predicted by MacRog, CHARMM36 and GAFFlipid is probably more realistic.
 
-It should be also noted that the free energy landscape is not correct in the models which
-are not able to reproduce the experimental order parameters. Thus the energetical response
-to the dehydration may not be correct even though the order parameter response is qualitatively correct.
-This issue may have some influence on the dehydration energetic calculations made with the Berger model~\cite{eun09,schneck12}.
+It must be stressed that in the models
+incapable of reproducing the experimental order parameters the free energy landscape is not correct. Thus
+even though the order parameter response to dehydration is qualitatively correct,
+the energetic response is likely to be incorrect.
+This may have some influence on dehydration energetic calculations made using the Berger model~\cite{eun09,schneck12}.
 \begin{figure}[]
@@
 
-The response of the glycerol backbone to dehydration seems to be more subtle than that of the choline headgroup 
-and none of the models reproduced the experimental results with the accuracy allowing
-the structural interpretation.
+The response of the glycerol backbone seems to be more subtle than that of the choline headgroup;
+none of the four models reproduced the experimental trends upon dehydration with enough accuracy to invite a
+structural interpretation.
 
@@
 It is widely agreed that cholesterol orders lipid acyl tails thus decreasing the area per molecule (condensing effect),
-however, the influence of cholesterol on the lipid headgroup and glycerol backbone are sill debated~\cite{huang99,simons04,somerharju09}.
+however, the influence of cholesterol on the lipid headgroup and glycerol backbone are still debated~\cite{huang99,simons04,somerharju09}.
 For example, it has been suggested that the surrounding lipids shield cholesterol from interactions with water by 
@@
 and cholesterol causes a measurable increase in the $\beta$ order parameter and a forking in the $\alpha$ order
-parameter~\cite{brown78}. These effects are, however, so small that they are barely visible in the scale used in Fig.~\ref{ordPchol}.
+parameter~\cite{brown78}. These effects are, however, so small that they are barely visible on the scale used in Fig.~\ref{ordPchol}.
 Further, the effects of cholesterol on the glycerol backbone order parameters for POPC from $^1$H-$^{13}$C NMR experiment~\cite{ferreira13} 
@@
 These results further support the idea that the glycerol backbone structural behaviour is independent of the
-headgroup composition~\cite{gally81} and that the headgroup stucture is independent of the acyl chain region content unless
+headgroup composition~\cite{gally81} and that the headgroup structure is largely independent of the acyl chain region content unless
 charges are present~\cite{scherer87}.
@@
     and Gross et al.~\cite{gross97}.  Most order parameters from Berger-POPC-07/H\"oltje-CHOL-13 model for g$_1$ are beoynd the y-axis scale.
-    Note that to elucidate the relative change as a function of cholesterol content,
+    In order to elucidate the relative change as a function of cholesterol content,
     the simulation results are shifted such that the (smaller) $S_\mathrm{CH}$
@@
 and our results from CHARMM36 and MacRog force fields are shown in Fig.~\ref{ordPchol}. 
-To elucidate changes the simulation results are shifted to match better with experiments without cholesterol.
+To elucidate changes, the simulation results are shifted to match better with experiments without cholesterol.
 The magnitude of shifting is written in the y-axis label.
-As already pointed out previosly, the Berger-POPC-07/H\"oltje-CHOL-13 model
-seriously overestimates the effect of cholesterol on the phospholipid glycerol backbone and choline segments~\cite{ferreira13}.
-In contrast, the responses of both CHARMM36 and MacRog are in better agreement with experiments. 
+As already pointed out previously, the Berger-POPC-07/H\"oltje-CHOL-13 model
+seriously exaggerates the effect of cholesterol on the phospholipid glycerol backbone and choline segments~\cite{ferreira13}.
+In comprasion, the choline responses of both CHARMM36 and MacRog are in better qualitative agreement with experiments,
+and also the glycerol backbone responses appear less drastic.
 %However, CHARMM36 seems to better reproduce the experimentally observed modest changes in the glycerol backbone segments 
@@
 Thus we have calculated the glycerol backbone dihedral angle distributions
-as a function of cholesterol in CHARMM36 (shown in Supplementary material) to resolve the cholesterol induced structural changes. The only detectable change due to the
-addition of cholesterol is the small decrease of gauche- and increase of gaughe+ probability of g3-g2-g1-O(\textit{sn}-1) dihedral.
+as a function of cholesterol in CHARMM36 (shown in Supplementary material) to resolve the cholesterol-induced structural changes. The only detectable change due to the
+addition of cholesterol is the small decrease of gauche(-) and increase of gaughe(+) probability of the g$_3$-g$_2$-g$_1$-O(\textit{sn}-1) dihedral.
+In conclusion, the results suggest that the significant effect of cholesterol on the lipid structures observed in simulations~\cite{zhu07,rog09,alwarawrah12} are overestimated by the model used in these studies, and the cholesterol only induces very small structural changes in the glycerol backbone.
 

### 2015-05-14 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Careful reading of the cholesterol part.

@@
 \subsubsection{Cholesterol-containing phospholipid bilayer}
-Phospholipid--cholesterol interactions have been widely studied with theoretical~\cite{huang99,zhu07,rog09,alwarawrah12} and
-experimental methods~\cite{brown78,marsh10,ferreira13,marsh13}, since cholesterol is abundant in biological membranes.
-It has been suggested to be an important player, for example, in domain formation~\cite{simons04,somerharju09}.
-It is widely agreed that cholesterol orders lipid acyl tails thus decreasing the area per molecule (condensing effect),
-however, the influence of cholesterol on the lipid headgroup and glycerol backbone are still debated~\cite{huang99,simons04,somerharju09}.
-For example, it has been suggested that the surrounding lipids shield cholesterol from interactions with water by 
-reorienting their headgroups (``umbrella model'')~\cite{huang99} or that cholesterol acts as a spacer for the headgroups thus increasing 
+As cholesterol is abundant in biological membranes and
+has been suggested to be an important player, for example, in domain formation~\cite{simons04,somerharju09},
+phospholipid--cholesterol interactions have been extensively studied with theoretical~\cite{huang99,zhu07,rog09,alwarawrah12} and
+experimental~\cite{brown78,marsh10,ferreira13,marsh13} methods.
+It is widely agreed that cholesterol orders lipid acyl tails and thus decreases the area per molecule (condensing effect),
+but its influence on the lipid headgroup and glycerol backbone remains debated~\cite{huang99,simons04,somerharju09}.
+It has been suggested, for example, that the surrounding phospholipids shield cholesterol from exposure to water by 
+reorienting their headgroups (``umbrella model'')~\cite{huang99} or that cholesterol acts as a spacer between the headgroups to increase 
 their entropy and dynamics (``superlattice model'')~\cite{somerharju09}. 
 %\todo{It has been suggested that we should use direct quotations from original papers here for clarity.}
-Both of these suggestions have been supported
-by molecular dynamics simulations~\cite{zhu07,alwarawrah12}, and other simulations suggest specific
-interactions between the glycerol backbone and cholesterol~\cite{rog09}. However,
-in the above mentioned studies, the glycerol backbone and choline headgroup behaviour
-as a function of cholesterol content has not been compared to experiments.
-
-The choline headgroup and glycerol backbone order parameters for POPC measured by $^1$H-$^{13}$C NMR~\cite{ferreira13} and DPPC choline order parameters 
-measured by $^{2}$H NMR~\cite{brown78} are shown in Fig.~\ref{ordPchol} as a function of cholesterol content.
-The agreement between different experimental results is again very good, showing only very modest changes in 
-the choline order parameters as a function of cholesterol content. It should be noted, however, that very small
-changes are measurable with high resolution $^{2}$H NMR experiments
-and cholesterol causes a measurable increase in the $\beta$ order parameter and a forking in the $\alpha$ order
-parameter~\cite{brown78}. These effects are, however, so small that they are barely visible on the scale used in Fig.~\ref{ordPchol}.
-Further, the effects of cholesterol on the glycerol backbone order parameters for POPC from $^1$H-$^{13}$C NMR experiment~\cite{ferreira13} 
-are in good agreement with the results for the phosphatidylethanolamine (PE) measured by $^{2}$H NMR~\cite{ghosh82}.
-These results further support the idea that the glycerol backbone structural behaviour is independent of the
+%Both of these suggestions 
+Molecular dynamics simulations have supported both
+the umbrella~\cite{alwarawrah12} as well as the superlattice~\cite{zhu07} model,
+in addition to suggesting specific interactions of cholesterol with the glycerol backbone~\cite{rog09}.
+In these studies~\cite{zhu07,rog09,alwarawrah12}
+the responses of the glycerol backbone and choline headgroup 
+to increasing cholesterol content were not, however, compared to experiments.
+
+Fig.~\ref{ordPchol} shows the responses of the choline headgroup ($\beta$ and $\alpha$)
+order parameters of POPC (measured by $^1$H-$^{13}$C NMR~\cite{ferreira13}) and DPPC
+($^{2}$H NMR~\cite{brown78}) to increasing cholesterol content.
+Again, the two independent data sets agree very well:
+Only very modest ($\Delta S_\mathrm{CH}<0.03$) changes occur in  the choline order parameters as cholesterol content
+increases from 0 to 60\%.
+The extreme sensitivity of the
+high resolution $^{2}$H NMR experiments
+is beautifully demonstrated by the measurable~\cite{brown78}
+(but barely visible on the scale used in Fig.~\ref{ordPchol})
+cholesterol-induced forking of the $\alpha$ order parameter.
+
+We note that the modest ($\Delta S_\mathrm{CH}<0.02$ for g$_1$; $<0.04$ for g$_2$, g$_3$; see Fig.~\ref{ordPchol})
+effects of cholesterol
+on the glycerol backbone order parameters of POPC
+measured by $^1$H-$^{13}$C NMR~\cite{ferreira13}
+agree well with the results for phosphatidylethanolamine (PE) measured by $^{2}$H NMR~\cite{ghosh82}.
+This further supports the ideas that the glycerol backbone structural behaviour is independent of the
 headgroup composition~\cite{gally81} and that the headgroup structure is largely independent of the acyl chain region content unless
@@
 
-In addition to the experimental data, the previously published simulation results from the Berger-POPC-07/H\"oltje-CHOL-13 model~\cite{ferreira13},
-and our results from CHARMM36 and MacRog force fields are shown in Fig.~\ref{ordPchol}. 
-To elucidate changes, the simulation results are shifted to match better with experiments without cholesterol.
-The magnitude of shifting is written in the y-axis label.
-As already pointed out previously, the Berger-POPC-07/H\"oltje-CHOL-13 model
-seriously exaggerates the effect of cholesterol on the phospholipid glycerol backbone and choline segments~\cite{ferreira13}.
-In comprasion, the choline responses of both CHARMM36 and MacRog are in better qualitative agreement with experiments,
-and also the glycerol backbone responses appear less drastic.
-%However, CHARMM36 seems to better reproduce the experimentally observed modest changes in the glycerol backbone segments 
-%g$_2$ and g$_3$ with high concentrations of cholesterol. 
-Thus we have calculated the glycerol backbone dihedral angle distributions
-as a function of cholesterol in CHARMM36 (shown in Supplementary material) to resolve the cholesterol-induced structural changes. The only detectable change due to the
-addition of cholesterol is the small decrease of gauche(-) and increase of gaughe(+) probability of the g$_3$-g$_2$-g$_1$-O(\textit{sn}-1) dihedral.
-In conclusion, the results suggest that the significant effect of cholesterol on the lipid structures observed in simulations~\cite{zhu07,rog09,alwarawrah12} are overestimated by the model used in these studies, and the cholesterol only induces very small structural changes in the glycerol backbone.
-
-It should be noted that the CHARMM36 force field parameters (dihedral potentials) for the glycerol backbone have been tuned to reproduce the correct order parameters at fully hydrated conditions~\cite{klauda10}.
-This procedure contains a risk of overfitting, which would manifest itself as wrong responses to changing conditions. 
-According to our results, this tuning does not seem to lead to overfitting problems in the case of dehydration or lipid--cholesterol mixtures.
+In addition to the experimental data, Fig.~\ref{ordPchol} shows
+our results for the CHARMM36 and MacRog force fields
+and the previously published~\cite{ferreira13}
+Berger-POPC-07/H\"oltje-CHOL-13 results.
+Note that the simulation data are shifted vertically to ease comparison with experimental responses.
+As previously pointed out~\cite{ferreira13}, the Berger-based model
+seriously exaggerates the effect of cholesterol on the phospholipid glycerol backbone and choline headgroup.
+In comparison, the choline responses of CHARMM36 and MacRog are in better qualitative agreement with experiments, and the glycerol backbone responses, 
+although still somewhat larger than what is seen experimentally,
+much less drastic.
+Therefore,
+to resolve the nature of cholesterol-induced structural changes,
+we calculated from CHARMM36 the glycerol backbone dihedral angle distributions
+at various cholesterol contents (Supplementary material): The only detectable change
+is the small decrease of gauche(-) and increase of gaughe(+) probability of the g$_3$-g$_2$-g$_1$-O(\textit{sn}-1) dihedral.
+In conclusion, our results suggest that the significant effects of cholesterol on lipid conformations observed in simulations~\cite{zhu07,rog09,alwarawrah12} are overestimated by the computational models used;
+cholesterol only induces very small structural changes in the glycerol backbone.
+
+Finally, it is important to note that the CHARMM36 force field parameters (glycerol backbone dihedral potentials)
+have been tuned to reproduce the experimental order parameters at full hydration~\cite{klauda10}. 
+This approach introduces a risk of overfitting, which would manifest itself as wrong responses to changing conditions. 
+Interestingly, according to our results, tuning did not lead to overfitting problems as far as dehydration or cholesterol content are considered.
 

### 2015-05-14 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Added edits to Conclusions too.

@@
 in phoshatidylcholine bilayers are not known despite of vast amount of accurate experimental
-data. Atomistic resolution molecular dynamics simulation model which reproduced the
-experimental data would automatically resolve the structures giving an interpretation of experimental results.
+data. An atomistic resolution molecular dynamics simulation model that would reproduce the
+experimental data would automatically resolve the structures, thus giving an unprecedently detailed interpretation of the experimental data.
 In this work we have collected and reviewed the experimental C--H bond vector order
-parameters available in the literature. These experimental parameters are then compared to
-different atomistic resolution simulation models for a fully hydrated bilayer, bilayers dehydrated to different extents, and
-lipid bilayers containing various amounts of cholesterol. Our results have led to the following conclusions:
-
-- The C--H bond order parameters measured with different NMR techniques are in good agreement
-with each others. By combining the experimental results from various sources we concluded
+parameters available in the literature. These accurate experimental data are then compared to 13
+different atomistic resolution simulation models for a fully hydrated lipid bilayer system, followed by bilayers dehydrated to different extents, and
+finally bilayers containing various amounts of cholesterol. We are led to the following four main conclusions:
+
+(1) The C--H bond order parameters measured with different NMR techniques are consistent.
+By combining the experimental results from various sources we concluded
 that the order parameters for each C--H bond are known with a quantitative accuracy of $\pm$0.02.
 
-- Comparison of order parameters between experiments and different atomistic resolution models 
-together with the structural analysis shows that these parameters can be used
-to judge the structural accuracy of the model. Thus the combination of atomistic resolution 
+(2) Comparison of order parameters between experiments and different atomistic resolution models 
+together with structural analysis showed that the order parameters can be used
+to judge the structural accuracy of a model. Thus the combination of atomistic resolution 
 molecular dynamics simulations and NMR experiments can be used to resolve the atomistic resolution
 structures of biomolecules in biologically relevant conditions. 
-This approach can be extended from lipids also, for example, to membrane proteins.
+This approach can be extended from lipids to, for example, membrane proteins.
 
@@
 
-- The review of previous experimental results revealed that the 
-choline order parameters are increasing when the bilayer is dehydrated. Our simulations suggest that
-this can be explained by P--N vector tilting more parallel to the membrane. 
-These results strongly supports and complements the idea that the charge induced choline tilting and 
-partition can be measured by using the order parameter changes~\cite{ionpaper,scherer89}.
+(3) The review of previous experimental results revealed that when a bilayer is dehydrated the 
+choline order parameters increase. Our simulations suggested that
+this can be explained by the P--N vector tilting more parallel to the membrane. 
+This strongly supports and complements the idea that charge-induced choline tilting
+can be measured using order parameter changes~\cite{ionpaper,scherer89}.
 
@@
 
-- Only modest changes of glycerol backbone and choline order parameters are observed experimentally with increasing cholesterol concentration.
-When interpreted by using the simulations with the most realistic response to the cholesterol concentration,
-this observation can be explained such that the cholesterol induces only minor changes in the g$_3$-g$_2$-g$_1$-O(sn-1)
-dihedral in the glycerol backbone.
+(4) Only modest changes of glycerol backbone and choline order parameters are observed experimentally with increasing cholesterol content.
+When interpreted using the computational lipid model that we found to have the most realistic response to cholesterol,
+this observation means that cholesterol induces only minor changes in (the g$_3$-g$_2$-g$_1$-O(sn-1)
+dihedral of) the glycerol backbone, in other words, there is no major conformational change of the lipid.
 
@@
 
-- Besides the main conclusions, we have created the most extensive
+(+) Besides these four main conclusions, we note that we have created the most extensive
 publicly available collection of molecular dynamics simulation trajectories of lipid bilayers
-into the Zenodo community \url{https://zenodo.org/collection/user-nmrlipids}. 
-This collection opens up numerous possiblities for different analyses with
-much less effort than previously required.
-
-In general, we conclude that to fully utilize the potential of atomistic resolution classical molecular dynamics simulations
-in the structural interpretation of the high resolution NMR data~\cite{ferreira14} for lipid bilayers one has to  
-improve the phoshatidylcholine glycerol backbone and choline headgroup parameters.
-
-This work has been done as an open collaboration by using the \url{nmrlipids.blogspot.fi} as an communication
-platform. All the scientific contributions have been done through the blog and are publicly
-available~\cite{nmrlipids}.
+(\url{https://zenodo.org/collection/user-nmrlipids}). The mere existence of
+this collection opens up numerous possibilities for unforeseen analyses, such as data mining,
+and rapid testing of ideas with
+much less computational effort than previously required.
+
+In general, we conclude that in order to fully utilize the potential of atomistic-resolution classical molecular dynamics simulations
+in the structural interpretation of high resolution NMR data~\cite{ferreira14} for lipid bilayers, one must  
+improve the phoshatidylcholine glycerol backbone and choline headgroup parameters of the existing force fields.
+
+This work has been done as a fully open collaboration, using \url{nmrlipids.blogspot.fi} as the communication
+platform. All the scientific contributions have been communicated publicly through this blog~\cite{nmrlipids}.
 

### 2015-05-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
small updates to the text

@@
 seriously exaggerates the effect of cholesterol on the phospholipid glycerol backbone and choline headgroup.
-In comparison, the choline responses of CHARMM36 and MacRog are in better qualitative agreement with experiments, and the glycerol backbone responses, 
-although still somewhat larger than what is seen experimentally,
-much less drastic.
-Therefore,
-to resolve the nature of cholesterol-induced structural changes,
+In comparison, the choline and glycerol backbone responses of CHARMM36 and MacRog are in better qualitative 
+agreement with experiments. 
+Therefore, to resolve the nature of cholesterol-induced structural changes,
 we calculated from CHARMM36 the glycerol backbone dihedral angle distributions
@@
 this observation means that cholesterol induces only minor changes in (the g$_3$-g$_2$-g$_1$-O(sn-1)
-dihedral of) the glycerol backbone, in other words, there is no major conformational change of the lipid.
+dihedral of the glycerol backbone, in other words, there is no major conformational change of the lipid.
 

### 2015-05-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
The abstract.

--- 2015-05-14 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
@@
 Phospholipids are essential building blocks of biological membranes.
-Despite of vast amount of accurate experimental data, the atomistic resolution structures sampled by the glycerol backbone and choline headgroup
-in phoshatidylcholine bilayers are not known. Atomistic resolution molecular dynamics simulations model 
-would automatically resolve the structures giving an interpretation of experimental results, if the model
-reproduced the experimental data. 
-In the present work we simulate phosphatidylcholine (PC) lipid bilayers with 13 different
-atomistic models, and we compare simulations with experiments (in fully hydrated conditions) in terms of C--H bond vector order
-parameters for the glycerol backbone and choline headgroups. 
-We show that this comparison can be used to judge the atomistic resolution structural accuracy of the models,
-thus allowing the usage of molecular dynamics simulations to interpret the structures of biomolecules in 
-biologically relevant conditions from NMR experiments. Further, we review previous experimental
-data for dehydrated lipid bilayers and cholesterol containing bilayers, and interpretate it with simulations.
-The interpretation suggest that the increased choline order parameters indicate P--N vector tilting more parallel to membrane,
-and that cholesterol induces only minor changes to the glycerol backbone structure.
+Despite of vast amount of very accurate experimental data, the atomistic resolution structures sampled by the glycerol backbone and choline headgroup
+in phoshatidylcholine bilayers are not known. Atomistic resolution molecular dynamics simulations 
+have the potential to resolve the structures, and to give an arrestingly intuitive interpretation of the experimental data---but only if the simulations
+reproduce the data within experimental accuracy.
+In the present work, we simulated phosphatidylcholine (PC) lipid bilayers with 13 different
+atomistic models, and compared simulations with NMR experiments in terms of the highly structurally sensitive C--H bond vector order
+parameters. Focusing on the glycerol backbone and choline headgroups, 
+we showed that the order parameter comparison can be used to judge the atomistic resolution structural accuracy of the models.
+Accurate models, in turn, allow molecular dynamics simulations to be used as an interpretation tool that translates these NMR data
+into a dynamic three dimensional representation of biomolecules in biologically relevant conditions.
+In addition to lipid bilayers in fully hydrated conditions, we reviewed previous experimental
+data for dehydrated bilayers and cholesterol-containing bilayers, and interpreted them with simulations.
+Although none of the existing models reached experimental accuracy, by critically comparing them we were able to distill relevant chemical information:
+(1) increase of choline order parameters indicates the P--N vector tilting more parallel to the membrane,
+and (2) cholesterol induces only minor changes to the PC (glycerol backbone) structure.
 %None of the current models is not accurately enough to resolve the structure with experimental accuracy.
@@
 %between lipids and cholesterol better than the Berger/H\"oltje model.
-This work has been done as an open collaboration by using the \url{nmrlipids.blogspot.fi} as an
-communication platform. All the scientific contributions have been done through the blog and
-are publicly available. In addition, almost all simulation trajectories and files are made available
-in the Zenodo community \url{https://zenodo.org/collection/user-nmrlipids} which has became the most extensive
+This work has been done as a fully open collaboration, using \url{nmrlipids.blogspot.fi} as a
+communication platform; all the scientific contributions were made publicly on this blog.
+During the open research process, the repository holding our simulation trajectories and files
+(\url{https://zenodo.org/collection/user-nmrlipids}) has become the most extensive
 publicly available collection of molecular dynamics simulation trajectories of lipid bilayers.

### 2015-05-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Introduction.

--- 2015-05-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
@@
 and theoretical techniques as a simple model for cellular membranes~\cite{lipowsky95,tieleman97,klauda08,edholm08,tieleman10,piggot12,rabinovich13,marsh13}. 
-Phospholipid molecules are composed of hydrophobic acyl chains which are connected by a glycerol backbone to a hydrophilic headgroup.
-See Fig.~\ref{POPCstructure} for the structure of 1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
+Phospholipid molecules are composed of hydrophobic acyl chains connected by a glycerol backbone to a hydrophilic headgroup;
+see Fig.~\ref{POPCstructure} for the structure of 1-palmitoyl-2-oleoyl\-phosphatidyl\-choline (POPC).
 The behaviour of the acyl chains in a lipid bilayer is relatively well understood~\cite{Israelachvili80,lipowsky95,tieleman97,klauda08,edholm08,tieleman10,marsh13}. 
 The conformations sampled by the glycerol backbone and choline in a fluid bilayer are, however, not fully 
-resolved since even the most accurate scattering and Nuclear Magnetic Resonance (NMR)
-techniques give only a set of values that the structure has to fulfil, but
+resolved as even the most accurate scattering and Nuclear Magnetic Resonance (NMR)
+techniques give only a set of values that the structure has to fulfill, but
 there is no unique way to derive the actual structure from them~\cite{seelig77b,skarjune79,Israelachvili80,jacobs80,davis83,strenk85,akutsu91,hong95b,hong96,semchyschyn04}.
-Some structural details have been extracted from crystal structure, $^1$H NMR studies and Raman spectroscopy~\cite{hauser80,hauser81,hauser81b,akutsu81b,pascher92,hauser88,marsh06}
-but general consensus about the structures sampled in the fluid state have not been reached~\cite{seelig77b,skarjune79,Israelachvili80,jacobs80,davis83,strenk85,hauser88,akutsu91,hong95b,hong96,semchyschyn04,marsh06}.
-On the other hand, the structural parameters for the glycerol backbone are similar for various biologically
+Some structural details have been extracted from crystal structures, $^1$H NMR studies, and Raman spectroscopy~\cite{hauser80,hauser81,hauser81b,akutsu81b,pascher92,hauser88,marsh06},
+but general consensus concerning the structures sampled in the fluid state has not been reached~\cite{seelig77b,skarjune79,Israelachvili80,jacobs80,davis83,strenk85,hauser88,akutsu91,hong95b,hong96,semchyschyn04,marsh06}.
+Importantly, the structural parameters for the glycerol backbone are similar for various biologically
 relevant lipid species (phosphatidylcholine (PC), phosphatidylethanolamine (PE) and phosphatidylglycerol (PG)) 
-in various environments~\cite{gally81} and the structural parameters for the headgroup choline structures are similar in model membranes and
+in various environments~\cite{gally81}, and the structural parameters for the choline headgroup are similar in model membranes and
 real cells (mouse fibroblast L-M cell)~\cite{scherer87}.
-Thus, the solution of phosphatidylcholine glycerol backbone and choline structures would be 
+Thus, resolving the PC-lipid glycerol and choline structures would be 
 useful for understanding a wide range of different biological membranes.
@@
 lipid bilayers~\cite{tieleman97,klauda08,edholm08,tieleman10,piggot12,rabinovich13}. As these models provide an atomistic
-resolution description of the whole lipid molecule, they have potential to solve the glycerol backbone and 
-headgroup structures. In particular, the experimental C--H bond order parameters for the glycerol backbone 
-(g$_1$, g$_2$ and g$_3$) and choline ($\alpha$ and $\beta$) segments (see Fig.~\ref{POPCstructure} for definitions) are among the main parameters used in
+resolution description of the whole lipid molecule, they have the potential to solve the glycerol backbone and 
+headgroup structures. The experimental C--H bond order parameters
+(routinely compared between experiments and simulations for the acyl chains~\cite{tieleman97,klauda08,edholm08,tieleman10,piggot12})
+are also known for the glycerol backbone 
+(g$_1$, g$_2$, and g$_3$) and choline ($\alpha$ and $\beta$) segments (see Fig.~\ref{POPCstructure} for definitions) and are among the main parameters used in
 attempts to derive lipid structures from experimental data~\cite{seelig77b,skarjune79,jacobs80,davis83,akutsu91,hong95b,semchyschyn04}.
-These parameters are also routinely compared between experiments and simulations for the acyl chains~\cite{tieleman97,klauda08,edholm08,tieleman10,piggot12}.
-Thus, the structures sampled in a simulation model that reproduces these and other experimental parameters, automatically
-give an interpretation of the experiments. In other words they can be considered as reasonable atomistic resolution descriptions of
+Notably, the structures sampled in a simulation that reproduces these %and other experimental
+parameters will automatically comprise an interpretation of the experiments. In other words, such simulations can be considered as an accurate atomistic resolution description of
 the behavior of lipid molecules in a bilayer.
 
-The glycerol backbone and choline headgroup order parameters have been compared between simulations and experiments
-in some studies~\cite{shinoda97,hogberg08,castro08,klauda10,kapla12,dickson12,poger12,ferreira13,chowdhary13,maciejewski14}, 
-however much less frequently than for the acyl  chains~\cite{tieleman97,klauda08,edholm08,tieleman10,piggot12}.
-The main reason is probably that the existing experimental data for the glycerol backbone
-and choline headgroups is scattered over many publications and published in a format that is difficult to understand without some NMR expertise. 
-In addition to the order parameters, also dihedral angles for glycerol backbone and headgroup estimated from experiments have been sometimes used to 
+Only a few studies~\cite{shinoda97,hogberg08,castro08,klauda10,kapla12,dickson12,poger12,ferreira13,chowdhary13,maciejewski14}
+have compared
+the glycerol backbone and choline headgroup order parameters between simulations and experiments. 
+The main reason probably is that the existing experimental data for the glycerol backbone
+and choline headgroups are scattered over many publications and published in a format that is difficult to understand without some NMR expertise. 
+In addition to the order parameters, dihedral angles for the glycerol backbone and headgroup estimated from experiments have sometimes been used to 
 assess the quality of a simulation model~\cite{robinson94,essex94,kothekar96,hyvonen97,shinoda97,duong99}.
 
-In this work we first review the most relevant experimental data for the glycerol backbone and choline headgroup order parameters
+In this work, we first review the most relevant experimental data for the glycerol backbone and choline headgroup order parameters
 in a phosphatidylcholine lipid bilayer. Then the available atomistic resolution lipid models are carefully compared to the 
-experimental data. The comparison reveals that the CHARMM36~\cite{klauda10}, GAFFlipid~\cite{dickson12} and MacRog models~\cite{maciejewski14}
+experimental data. The comparison reveals that the CHARMM36~\cite{klauda10}, GAFFlipid~\cite{dickson12}, and MacRog~\cite{maciejewski14} models
 have the most realistic glycerol backbone and choline structures. We also compare the glycerol backbone and choline 
-structures between the most often used (Berger based) lipid model~\cite{berger97} 
+structures between the most often used (Berger-based) lipid model~\cite{berger97} 
 %\todo{If there are no objections, we will leave the naming convention of different Berger versions as it is now.}
@@
 order parameters we can distinguish the more reasonable structures from the less reasonable ones. However, none of the current models 
-is accurate enough to properly resolve the atomistic resolution stuctures.
+is accurate enough to properly resolve the atomistic resolution structures.
 
 In addition to fully hydrated single component lipid bilayers, the glycerol backbone and choline order parameters
-have been measured under a large number of different conditions. For example, as a function of hydration level~\cite{bechinger91,ulrich94,dvinskikh05b}, cholesterol content~\cite{brown78,ferreira13}
+have been measured under a large number of changing conditions: hydration level~\cite{bechinger91,ulrich94,dvinskikh05b}, cholesterol content~\cite{brown78,ferreira13},
 ion concentration~\cite{brown77,akutsu81,altenbach84,roux90,roux91}, temperature~\cite{gally75}, charged lipid content~\cite{roux90,roux91}, charged surfactant content~\cite{scherer89}, 
 drug molecule concentration~\cite{browning82,kelusky84,castro08}, and protein content~\cite{roux89,kuchinka89} (listing only the publications most relevant for this work and the pioneering studies).
-Awareness of the existence of this type of data allows the comparison of structural responses to varying conditions between simulations and experiments,
-which can be used to validate the simulation models and to interpret the original experiments. 
-In this publication we demonstrate the power of this approach for understanding the behaviour of a bilayer as a function of hydration level and cholesterol concentration.
+Existence of these data allows the comparison of structural responses to varying conditions between simulations and experiments,
+in other words, validation of the simulation models and interpretation of the original experiments. 
+Here we demonstrate the power of this approach in understanding the behaviour of a bilayer as a function of hydration level and cholesterol content.
 Choline headgroup order parameters as function of ion concentration, and their relation to the ion binding affinity, are discussed elsewhere~\cite{ionpaper}.

### 2015-05-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Methods: Open collaboration.

@@
 
-\section{methods}
+\section{Methods}
 
@@
 
-This work has been done as an open collaboration by using \url{nmrlipids.blogspot.fi}~\cite{nmrlipids}
+This work has been done as a fully open collaboration, using the \url{nmrlipids.blogspot.fi} blog~\cite{nmrlipids}
 as a communication platform.
-The approach is inspired by the Polymath project~\cite{gowers09}, however there are some essential differences. 
-The present approach was started by publishing a manuscript~\cite{ollila13} discussing the glycerol backbone and choline structures 
-in a Berger based model (the most used molecular dynamics simulation model for lipid bilayers).
-After publishing the initial manuscript, the open invitation for further contributions and discussion through the blog was presented.
-All the scientific contributions are done publicly through the blog and all contributors were offered coauthorship
-according to the guidelines defined in the beginning of the project~\cite{creditsPOST}.
-The acception of authorship is based on the self-assesment of the authors' scientific contribution to the project.
-The contributions from each author are summarized in the Supplementary Information. %(the location depends on the journal probably).
-
-Almost all simulation data, including running parameters for reprocution and trajectories for further analysis, are collected
-to the nmrlipids community in Zenodo file repository \url{https://zenodo.org/collection/user-nmrlipids}.
-Thus, besides the main topic of this manuscript we present the most extensive available collection of simulations trajectories
-for lipid bilayers, which opens up numerous possiblities for different analyses with much less effort than previously required.
-Further information, e.g. scripts, figures and manuscript text files, are available in the GitHub 
-organization \url{https://github.com/NMRlipids} \todo{I will not assign a doi yet. We can wait for the journal decision.}.
+Our approach is inspired by the Polymath project~\cite{gowers09}, however there are some essential differences. 
+We started by publishing a manuscript~\cite{ollila13} discussing the glycerol backbone and choline structures 
+in a Berger-based model (the most used molecular dynamics simulation model for lipid bilayers).
+Simultaneously,  we presented an open invitation for further contributions and discussion on the blog.
+All the scientific contributions were made publicly through the blog. Every contributor was offered coauthorship
+according to the guidelines defined in the beginning of the project~\cite{creditsPOST};
+the acceptance of the offer was based on authors' self-assesment of their scientific contribution.
+These contributions are summarized in the Supplementary Information. %(the location depends on the journal probably).
+
+Almost all simulation data, including input files for reproduction and trajectories for further analysis, are collected
+on our CERN-hosted Zenodo file repository (\url{https://zenodo.org/collection/user-nmrlipids}).
+Thus, in addition to the main topic of this manuscript, we present the most extensive publicly available collection of simulation trajectories
+for lipid bilayers, opening up numerous possiblities for different analyses with much less effort than previously required.
+Further information, such as scripts, figures, and manuscript text files, are available through our GitHub 
+repository (\url{https://github.com/NMRlipids}). \todo{I will not assign a doi yet. We can wait for the journal decision.}.
 

### 2015-05-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Methods: Order parameters from experiments.

--- 2015-05-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
@@
 splitting with $^1$H-$^{13}$C NMR~\cite{hong95a,gross97,dvinskikh05a,ferreira13}. The measurements are based on
-different physical interactions and also the connection between order parameters and quadrupolar or dipolar splitting
+different physical interactions, and also the connections between order parameter and quadrupolar or dipolar splitting
 are different. The absolute values of order parameters from the measured quadrupolar splitting $\Delta \nu_Q$ ($^2$H NMR) are calculated using 
@@
 splitting constant is estimated from various experiments to be 170~kHz leading to a numerical relation $|S_{{\rm CD}}|=0.00784 \times \Delta \nu_Q$~\cite{seelig77c}. 
-The absolute values of order parameters from the effective dipolar coupling $d_\mathrm{CH}$ ($^1$H-$^{13}$C NMR) are calculated using equation
+The absolute values of order parameters from the effective dipolar coupling $d_\mathrm{CH}$ ($^1$H-$^{13}$C NMR) are calculated using the equation
 $|S_{{\rm CH}}|=\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c} d_\mathrm{CH}$, where
-values between 20.2--22.7 kHz are used for $\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c}$,
+values between 20.2--22.7~kHz are used for $\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c}$,
 depending on the original authors~\cite{hong95a,gross97,dvinskikh05a,ferreira13}.
-The effective dipolar coupling $d_\mathrm{CH}$ is related to the measured dipolar splitting $\Delta \nu_{CH}$ 
-through scaling factor which depends on the pulse sequence used in the $^1$H-$^{13}$C NMR experiment~\cite{hong95a,gross97,dvinskikh05a,ferreira13}.
+The effective dipolar coupling $d_\mathrm{CH}$ is related to the measured dipolar splitting $\Delta \nu_\mathrm{CH}$ 
+through a scaling factor that depends on the pulse sequence used in the $^1$H-$^{13}$C NMR experiment~\cite{hong95a,gross97,dvinskikh05a,ferreira13}.
 It is important to note that the order parameters measured with different techniques based on different physical interactions are in good agreement
 with each other (see Results and Discussion), indicating very high quantitative accuracy of the measurements.
-For a more detailed discussion see Ref.~\citenum{accuracyPOST}.
+For a more detailed discussion, see Ref.~\citenum{accuracyPOST}.
 
@@
 However, only $^1$H-$^{13}$C NMR techniques allow also the measurement of the sign of the order parameter~\cite{hong95a,hong95b,gross97}. 
-The measured sign is negative for almost all the carbons discussed in this work as only $\alpha$ is positive~\cite{hong95a,hong95b,gross97}. 
-For more detailed discussion about the sign measurements of the order parameters see Ref.~\citenum{signpost}.
-
-For most CH$_2$ segments in fluid phosholipid bilayer the order parameters are equal for both hydrogens attached to the same carbon.
-However, in some cases (e.g. g$_1$, g$_3$ and  C$_2$ carbon in the \textit{sn}-2 chain) the order parameters are not equal and this 
-can be observed with both $^2$H NMR and $^1$H-$^{13}$C NMR techniques. In the present work we call the phenomena of unequal order parameters 
-for hydrogens attached to the same carbon as {\it forking} to avoid confusion with dipolar and quadrupolar splitting in NMR terminology. 
+The measured sign is negative for almost all the carbons discussed in this work, only the choline $\alpha$ is positive~\cite{hong95a,hong95b,gross97}. 
+For more detailed discussion about the sign measurements of the order parameters, see Ref.~\citenum{signpost}.
+
+For most CH$_2$ segments in a fluid phospholipid bilayer, the order parameters of both hydrogens are equal.
+However, in some cases (e.g., g$_1$, g$_3$, and  C$_2$ carbon in the \textit{sn}-2 chain) the two order parameters are not equal;
+this  can be observed with both $^2$H NMR and $^1$H-$^{13}$C NMR techniques. In the present work,
+to avoid confusion with the dipolar and quadrupolar splittings in NMR terminology,
+we call the phenomenon of unequal order parameters for hydrogens attached to the same carbon {\it forking}. 
 Forking has been studied in detail with $^2$H NMR techniques by deuterating the R or S position in CH$_2$ segment, and
-the studies show that the forking arises from differently sampled orientations of the two C--H bonds, not from two 
+the studies show that it arises from differently sampled orientations of the two C--H bonds, not from two 
 separate populations of lipid conformations~\cite{engel81,gally81}.

### 2015-05-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Comments by Monticelli added and contributions updated

--- 2015-05-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
@@
 However, only $^1$H-$^{13}$C NMR techniques allow also the measurement of the sign of the order parameter~\cite{hong95a,hong95b,gross97}. 
-The measured sign is negative for almost all the carbons discussed in this work as only $\alpha$ is positive~\cite{hong95a,hong95b,gross97}. 
-For more detailed discussion about the sign measurements of the order parameters see Ref.~\citenum{signpost}.
+The measured sign is negative for almost all the carbons discussed in this work, except for $\alpha$ which is positive~\cite{hong95a,hong95b,gross97}. 
+For more detailed discussion about the determination of the sign of the order parameters see Ref.~\citenum{signpost}.
 
@@
 For more details, see Ref.~\citenum{accuracyPOST}. 
-Motivated by the high experimental repeatability, we have highlighted in 
+Motivated by the high experimental reproducibility, we have highlighted in 
 Fig.~\ref{HGorderparameters} the subjective sweet spots (light blue areas), within which we expect the calculated absolute 
@@
 between simulations and experiments in different conditions can be used to assess the quality of the force field 
-in different situations, and, if the quality is good, to potentially interpret the structural changes in experiments.
+in different situations, and, if the quality is good, to interpret the structural changes in experiments.
 Here we exemplify such comparison for a lipid bilayer under low hydration levels and when varying amounts of cholesterol is included in the bilayer. 
@@
 {\it Alexander Lyubartsev} provided simulation results for Högberg08 force field.\\
-{\it Markus S. Miettinen} \todo{Please write a text summarizing your contribution} \\
+{\it Markus S. Miettinen} Co-designed the project with OHSO and supported in the work management. Provided trajectories for Berger DMPC; prepared parameter files for 72 lipids \\
 {\it Luca Monticelli} Critical discussions in all phases of the project. Collaboration with Jukka  M{\"a}{\"a}tt{\"a}. \\
 {\it Jukka M{\"a}{\"a}tt{\"a}}  prepared and performed simulations with Berger and Slipids models and analyzed the results.\\
-{\it O. H. Samuli Ollila} Designed and managed the work. Ran and analyzed several simulations. Wrote the manuscript.  \\
+{\it O. H. Samuli Ollila}Co-designed the project with MSM and managed the work. Ran and analyzed several simulations. Wrote the manuscript. \\
 {\it Marius Retegan} Prepared and validated the GROMACS compatible parameter files for GAFFlipid and Lipid14 force fields.\\

### 2015-05-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Figs. 7 and 9 chosen based on the voting https://github.com/NMRLipids/nmrlipids.blogspot.fi/issues/38

@@
   \includegraphics[width=7.3cm]{../DATAreportediINblog/dehydration_lp.pdf}
-    \includegraphics[width=7.3cm]{../DATAreportediINblog/dehydration_smooth.pdf}
+%    \includegraphics[width=7.3cm]{../DATAreportediINblog/dehydration_smooth.pdf}
     \caption{\label{ordPhydr}
@@
   }
-  	\todo{Markus prepared two versions of this plot. The one on left has all the data points joint by lines;
-	the one on right shows smooth trend lines (cubic polynomials) to ease qualitative comparison.
-	Samuli likes the one on left, Markus the one on right. Do others have opinions?}
+%  	\todo{Markus prepared two versions of this plot. The one on left has all the data points joint by lines;
+%	the one on right shows smooth trend lines (cubic polynomials) to ease qualitative comparison.
+%	Samuli likes the one on left, Markus the one on right. Do others have opinions?}
 \end{figure}
@@
   \includegraphics[width=7.3cm]{../DATAreportediINblog/cholesterolization_lp.pdf}
-    \includegraphics[width=7.3cm]{../DATAreportediINblog/cholesterolization_smooth.pdf}
+%    \includegraphics[width=7.3cm]{../DATAreportediINblog/cholesterolization_smooth.pdf}
   \caption{\label{ordPchol}
@@
 }
-	\todo{Markus prepared two versions of this plot. The one on left has all the data points joint by lines;
-	the one on right shows smooth trend lines (cubic polynomials) to ease qualitative comparison.
-	Samuli likes the one on left, Markus the one on right. Do others have opinions?}
+%	\todo{Markus prepared two versions of this plot. The one on left has all the data points joint by lines;
+%	the one on right shows smooth trend lines (cubic polynomials) to ease qualitative comparison.
+%	Samuli likes the one on left, Markus the one on right. Do others have opinions?}
 \end{figure}

### 2015-05-15 HGmodelMANUSCRIPT/HGmodel_ACSsupplementary.tex
Separate main text and supplementary

--- 2015-05-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
@@
 \title[An \textsf{achemso} demo]
-  {Towards atomistic resolution structure of phosphatidylcholine headgroup and glycerol backbone at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}}
+  {SUPPLEMENTARY INFORMATION: Towards atomistic resolution structure of phosphatidylcholine headgroup and glycerol backbone at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\abbreviations{IR,NMR,UV}
-\keywords{American Chemical Society, \LaTeX}
+%\abbreviations{IR,NMR,UV}
+%\keywords{American Chemical Society, \LaTeX}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{tocentry}
+%\begin{tocentry}
 
-Some journals require a graphical entry for the Table of Contents.
-This should be laid out ``print ready'' so that the sizing of the
-text is correct.
+%Some journals require a graphical entry for the Table of Contents.
+%This should be laid out ``print ready'' so that the sizing of the
+%text is correct.
 
-Inside the \texttt{tocentry} environment, the font used is Helvetica
-8\,pt, as required by \emph{Journal of the American Chemical
-Society}.
+%Inside the \texttt{tocentry} environment, the font used is Helvetica
+%8\,pt, as required by \emph{Journal of the American Chemical
+%Society}.
 
-The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
-permitted for  \emph{Journal of the American Chemical Society}
-graphical table of content entries. The box will not resize if the
-content is too big: instead it will overflow the edge of the box.
+%The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
+%permitted for  \emph{Journal of the American Chemical Society}
+%graphical table of content entries. The box will not resize if the
+%content is too big: instead it will overflow the edge of the box.
 
-This box and the associated title will always be printed on a
-separate page at the end of the document.
+%This box and the associated title will always be printed on a
+%separate page at the end of the document.
 
-\end{tocentry}
+%\end{tocentry}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{abstract}
-Phospholipids are essential building blocks of biological membranes.
-Despite of vast amount of very accurate experimental data, the atomistic resolution structures sampled by the glycerol backbone and choline headgroup
-in phoshatidylcholine bilayers are not known. Atomistic resolution molecular dynamics simulations 
-have the potential to resolve the structures, and to give an arrestingly intuitive interpretation of the experimental data---but only if the simulations
-reproduce the data within experimental accuracy.
-In the present work, we simulated phosphatidylcholine (PC) lipid bilayers with 13 different
-atomistic models, and compared simulations with NMR experiments in terms of the highly structurally sensitive C--H bond vector order
-parameters. Focusing on the glycerol backbone and choline headgroups, 
-we showed that the order parameter comparison can be used to judge the atomistic resolution structural accuracy of the models.
-Accurate models, in turn, allow molecular dynamics simulations to be used as an interpretation tool that translates these NMR data
-into a dynamic three dimensional representation of biomolecules in biologically relevant conditions.
-In addition to lipid bilayers in fully hydrated conditions, we reviewed previous experimental
-data for dehydrated bilayers and cholesterol-containing bilayers, and interpreted them with simulations.
-Although none of the existing models reached experimental accuracy, by critically comparing them we were able to distill relevant chemical information:
-(1) increase of choline order parameters indicates the P--N vector tilting more parallel to the membrane,
-and (2) cholesterol induces only minor changes to the PC (glycerol backbone) structure.
-%None of the current models is not accurately enough to resolve the structure with experimental accuracy.
-%However, closer inspection of three best performing models (CHARMM36, GAFFlipid and MacRog) suggest
-%that improvements in the sampled dihedral angle distributions would potentilly lead to the model which
-%would resolve the structure. Despite of the inaccuracy in the fully hydrated
-%structures, the response to dehydration, i.e. the tilting of the P--N vector more parallel to membrane normal, 
-%is qualitatively correct in all models. The CHARMM36 and MacRog models describe the interactions
-%between lipids and cholesterol better than the Berger/H\"oltje model.
-This work has been done as a fully open collaboration, using \url{nmrlipids.blogspot.fi} as a
-communication platform; all the scientific contributions were made publicly on this blog.
-During the open research process, the repository holding our simulation trajectories and files
-(\url{https://zenodo.org/collection/user-nmrlipids}) has become the most extensive
-publicly available collection of molecular dynamics simulation trajectories of lipid bilayers.
-\end{abstract}
 
@@
 
-\section{Introduction}
-
-Phospholipids containing various polar headgroups and acyl chains are essential building blocks of 
-biological membranes. Lamellar phospholipid bilayer structures have been widely studied with various experimental 
-and theoretical techniques as a simple model for cellular membranes~\cite{lipowsky95,tieleman97,klauda08,edholm08,tieleman10,piggot12,rabinovich13,marsh13}. 
-Phospholipid molecules are composed of hydrophobic acyl chains connected by a glycerol backbone to a hydrophilic headgroup;
-see Fig.~\ref{POPCstructure} for the structure of 1-palmitoyl-2-oleoyl\-phosphatidyl\-choline (POPC).
-The behaviour of the acyl chains in a lipid bilayer is relatively well understood~\cite{Israelachvili80,lipowsky95,tieleman97,klauda08,edholm08,tieleman10,marsh13}. 
-The conformations sampled by the glycerol backbone and choline in a fluid bilayer are, however, not fully 
-resolved as even the most accurate scattering and Nuclear Magnetic Resonance (NMR)
-techniques give only a set of values that the structure has to fulfill, but
-there is no unique way to derive the actual structure from them~\cite{seelig77b,skarjune79,Israelachvili80,jacobs80,davis83,strenk85,akutsu91,hong95b,hong96,semchyschyn04}.
-Some structural details have been extracted from crystal structures, $^1$H NMR studies, and Raman spectroscopy~\cite{hauser80,hauser81,hauser81b,akutsu81b,pascher92,hauser88,marsh06},
-but general consensus concerning the structures sampled in the fluid state has not been reached~\cite{seelig77b,skarjune79,Israelachvili80,jacobs80,davis83,strenk85,hauser88,akutsu91,hong95b,hong96,semchyschyn04,marsh06}.
-Importantly, the structural parameters for the glycerol backbone are similar for various biologically
-relevant lipid species (phosphatidylcholine (PC), phosphatidylethanolamine (PE) and phosphatidylglycerol (PG)) 
-in various environments~\cite{gally81}, and the structural parameters for the choline headgroup are similar in model membranes and
-real cells (mouse fibroblast L-M cell)~\cite{scherer87}.
-Thus, resolving the PC-lipid glycerol and choline structures would be 
-useful for understanding a wide range of different biological membranes.
-
-Classical atomistic molecular dynamics simulations have been widely used to study  
-lipid bilayers~\cite{tieleman97,klauda08,edholm08,tieleman10,piggot12,rabinovich13}. As these models provide an atomistic
-resolution description of the whole lipid molecule, they have the potential to solve the glycerol backbone and 
-headgroup structures. The experimental C--H bond order parameters
-(routinely compared between experiments and simulations for the acyl chains~\cite{tieleman97,klauda08,edholm08,tieleman10,piggot12})
-are also known for the glycerol backbone 
-(g$_1$, g$_2$, and g$_3$) and choline ($\alpha$ and $\beta$) segments (see Fig.~\ref{POPCstructure} for definitions) and are among the main parameters used in
-attempts to derive lipid structures from experimental data~\cite{seelig77b,skarjune79,jacobs80,davis83,akutsu91,hong95b,semchyschyn04}.
-Notably, the structures sampled in a simulation that reproduces these %and other experimental
-parameters will automatically comprise an interpretation of the experiments. In other words, such simulations can be considered as an accurate atomistic resolution description of
-the behavior of lipid molecules in a bilayer.
-
-Only a few studies~\cite{shinoda97,hogberg08,castro08,klauda10,kapla12,dickson12,poger12,ferreira13,chowdhary13,maciejewski14}
-have compared
-the glycerol backbone and choline headgroup order parameters between simulations and experiments. 
-The main reason probably is that the existing experimental data for the glycerol backbone
-and choline headgroups are scattered over many publications and published in a format that is difficult to understand without some NMR expertise. 
-In addition to the order parameters, dihedral angles for the glycerol backbone and headgroup estimated from experiments have sometimes been used to 
-assess the quality of a simulation model~\cite{robinson94,essex94,kothekar96,hyvonen97,shinoda97,duong99}.
-
-In this work, we first review the most relevant experimental data for the glycerol backbone and choline headgroup order parameters
-in a phosphatidylcholine lipid bilayer. Then the available atomistic resolution lipid models are carefully compared to the 
-experimental data. The comparison reveals that the CHARMM36~\cite{klauda10}, GAFFlipid~\cite{dickson12}, and MacRog~\cite{maciejewski14} models
-have the most realistic glycerol backbone and choline structures. We also compare the glycerol backbone and choline 
-structures between the most often used (Berger-based) lipid model~\cite{berger97} 
-%\todo{If there are no objections, we will leave the naming convention of different Berger versions as it is now.}
-and the best performing models, to demonstrate that by using the 
-order parameters we can distinguish the more reasonable structures from the less reasonable ones. However, none of the current models 
-is accurate enough to properly resolve the atomistic resolution structures.
-
-In addition to fully hydrated single component lipid bilayers, the glycerol backbone and choline order parameters
-have been measured under a large number of changing conditions: hydration level~\cite{bechinger91,ulrich94,dvinskikh05b}, cholesterol content~\cite{brown78,ferreira13},
-ion concentration~\cite{brown77,akutsu81,altenbach84,roux90,roux91}, temperature~\cite{gally75}, charged lipid content~\cite{roux90,roux91}, charged surfactant content~\cite{scherer89}, 
-drug molecule concentration~\cite{browning82,kelusky84,castro08}, and protein content~\cite{roux89,kuchinka89} (listing only the publications most relevant for this work and the pioneering studies).
-Existence of these data allows the comparison of structural responses to varying conditions between simulations and experiments,
-in other words, validation of the simulation models and interpretation of the original experiments. 
-Here we demonstrate the power of this approach in understanding the behaviour of a bilayer as a function of hydration level and cholesterol content.
-Choline headgroup order parameters as function of ion concentration, and their relation to the ion binding affinity, are discussed elsewhere~\cite{ionpaper}.
-
-  \begin{figure}[]
-  \centering
-  \includegraphics[width=8.6cm]{../Fig/POPCstructure.pdf}
-
-  \caption{\label{POPCstructure}
-    Chemical structure of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).}
-
-\end{figure}
-
-\section{Methods}
-
-\subsection{Open collaboration}
-%\todo{Should we write more about this?}
-
-This work has been done as a fully open collaboration, using the \url{nmrlipids.blogspot.fi} blog~\cite{nmrlipids}
-as a communication platform.
-Our approach is inspired by the Polymath project~\cite{gowers09}, however there are some essential differences. 
-We started by publishing a manuscript~\cite{ollila13} discussing the glycerol backbone and choline structures 
-in a Berger-based model (the most used molecular dynamics simulation model for lipid bilayers).
-Simultaneously,  we presented an open invitation for further contributions and discussion on the blog.
-All the scientific contributions were made publicly through the blog. Every contributor was offered coauthorship
-according to the guidelines defined in the beginning of the project~\cite{creditsPOST};
-the acceptance of the offer was based on authors' self-assesment of their scientific contribution.
-These contributions are summarized in the Supplementary Information. %(the location depends on the journal probably).
-
-Almost all simulation data, including input files for reproduction and trajectories for further analysis, are collected
-on our CERN-hosted Zenodo file repository (\url{https://zenodo.org/collection/user-nmrlipids}).
-Thus, in addition to the main topic of this manuscript, we present the most extensive publicly available collection of simulation trajectories
-for lipid bilayers, opening up numerous possiblities for different analyses with much less effort than previously required.
-Further information, such as scripts, figures, and manuscript text files, are available through our GitHub 
-repository (\url{https://github.com/NMRlipids}). \todo{I will not assign a doi yet. We can wait for the journal decision.}.
-
-\subsection{Order parameters from experiments}\label{expORDp}
-The order parameter of a hydrocarbon C--H vector is defined as 
-\begin{equation}\label{orderP}
-S_{{\rm CH}}=\frac{1}{2}\langle 3 \cos^2 \theta-1 \rangle,
-\end{equation} 
-where the angle brackets denote an ensemble average over the sampled conformations, and $\theta$ is the angle between the C--H bond and the membrane normal.
-The absolute values of order parameters can be measured by detecting quadrupolar splitting with $^2$H NMR~\cite{seelig77c} or by detecting dipolar 
-splitting with $^1$H-$^{13}$C NMR~\cite{hong95a,gross97,dvinskikh05a,ferreira13}. The measurements are based on
-different physical interactions and also the connection between order parameters and quadrupolar or dipolar splitting
-are different. The absolute values of order parameters from the measured quadrupolar splitting $\Delta \nu_Q$ ($^2$H NMR) are calculated using 
-the equation $|S_{{\rm CD}}|=\frac{4}{3} \frac{e^2qQ}{h} \Delta \nu_Q$, where the value for the static quadrupole
-splitting constant is estimated from various experiments to be 170~kHz leading to a numerical relation $|S_{{\rm CD}}|=0.00784 \times \Delta \nu_Q$~\cite{seelig77c}. 
-The absolute values of order parameters from the effective dipolar coupling $d_\mathrm{CH}$ ($^1$H-$^{13}$C NMR) are calculated using equation
-$|S_{{\rm CH}}|=\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c} d_\mathrm{CH}$, where
-values between 20.2--22.7 kHz are used for $\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c}$,
-depending on the original authors~\cite{hong95a,gross97,dvinskikh05a,ferreira13}.
-The effective dipolar coupling $d_\mathrm{CH}$ is related to the measured dipolar splitting $\Delta \nu_{CH}$ 
-through scaling factor which depends on the pulse sequence used in the $^1$H-$^{13}$C NMR experiment~\cite{hong95a,gross97,dvinskikh05a,ferreira13}.
-It is important to note that the order parameters measured with different techniques based on different physical interactions are in good agreement
-with each other (see Results and Discussion), indicating very high quantitative accuracy of the measurements.
-For a more detailed discussion see Ref.~\citenum{accuracyPOST}.
-
-The absolute values of order parameters are accessible with both $^2$H NMR and $^1$H-$^{13}$C NMR techniques. 
-However, only $^1$H-$^{13}$C NMR techniques allow also the measurement of the sign of the order parameter~\cite{hong95a,hong95b,gross97}. 
-The measured sign is negative for almost all the carbons discussed in this work, except for $\alpha$ which is positive~\cite{hong95a,hong95b,gross97}. 
-For more detailed discussion about the determination of the sign of the order parameters see Ref.~\citenum{signpost}.
-
-For most CH$_2$ segments in fluid phosholipid bilayer the order parameters are equal for both hydrogens attached to the same carbon.
-However, in some cases (e.g. g$_1$, g$_3$ and  C$_2$ carbon in the \textit{sn}-2 chain) the order parameters are not equal and this 
-can be observed with both $^2$H NMR and $^1$H-$^{13}$C NMR techniques. In the present work we call the phenomena of unequal order parameters 
-for hydrogens attached to the same carbon as {\it forking} to avoid confusion with dipolar and quadrupolar splitting in NMR terminology. 
-Forking has been studied in detail with $^2$H NMR techniques by deuterating the R or S position in CH$_2$ segment, and
-the studies show that the forking arises from differently sampled orientations of the two C--H bonds, not from two 
-separate populations of lipid conformations~\cite{engel81,gally81}.
-
-\subsection{Order parameters from simulations}
-The order parameters from simulations were calculated directly using the definition as in Eq.~\ref{orderP}.
-%$S_{{\rm CH}}=\frac{3}{2} \langle \cos^2 \theta-1 \rangle$. 
-For the united atom models the hydrogen positions were generated 
-in the trajectories post-simulationally using the positions of the heavy atoms and the known hydrocarbon geometries.
-For the statistical error estimates, the time average of order parameters were first calculated separately
-for each lipid in the system. Then it was assumed that different lipids are statistically independent 
-entities (which should be the case in fluid phase) and the error of the mean for the average over individual 
-lipids in the system was calculated and used as error bar for the order parameters.
-
-It has been recently pointed out that the sampling of individual dihedral angles might be very
-slow compared to the typical simulation timescales~\cite{vogel12}. On the other hand, another recent
-study shows that the slowest rotational correlation functions of a C--H bond (g$_1$) reaches a plateau (S$_{CH}^2$)
-after 200~ns in the Berger-POPC-07~\cite{ollila07a} model, and that the dynamics of this segment is significantly too slow in simulations
-compared to the experiments~\cite{ferreira15}. In practise, less than 200~ns of simulation data is enough for the order parameter
-calculation due to the average over different lipid molecules. In conclusion, if the sampling with typical simulation times
-is not enough for the convergence of the order parameters, then the simulation models has significantly too slow dynamics.
-
-\subsection{Simulated systems}
-All simulations are ran with a standard setup for a planar lipid bilayer in zero tension and constant temperature
-with periodic boundary conditions in all directions by using the GROMACS software package~\cite{hess08} 
-(version numbers 4.5.X--4.6.X), LAMMPS~\cite{plimpton95}, MDynaMix~\cite{Lyubartsev00} or NAMD~\cite{phillips05}.
-The number of molecules, simulation temperatures and the length of simulations of all the simulated systems 
-are listed in Tables~\ref{systems},~\ref{systemsDEHYD} and~\ref{systemsCHOL}. Full simulation
-details are given in the Supplementary Information (SI) or in the original publications in case the
-data is used previously therein. Additionally, the files related to the simulations and the resulting trajectories are publicly
-available for almost all systems in the Zenodo collection \url{https://zenodo.org/collection/user-nmrlipids}. 
-The references pointing to simulation details and files are also listed in Tables~\ref{systems},~\ref{systemsDEHYD} and~\ref{systemsCHOL}.
-
-\begin{sidewaystable}[]
-\centering
-\caption{Fully hydrated single component lipid bilayer systems simulated for Fig.~\ref{HGorderparameters}:
-  1,2-dimyristoyl-sn-glycero-3-phosphocholine (DMPC),
-  %1,2-dioleoyl-sn-glycero-3-phosphocholine (DOPC),
-  dilauroylphosphatidylcholine (DLPC),
-  dipalmitoylphosphatidylcholine (DPPC), and
-  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).
-The {\bf bolded} systems were used also for Fig.~\ref{HGorderparameters2}.
-$^a$ Number of lipid molecules.
-$^b$ Number of water molecules.
-$^c$ Temperature.
-$^d$ Total simulation time.
-$^e$ Time used for analysis.
-$^f$ Reference link for the downloadable simulation files;
-     the data sets marked with $^*$ also include a part of the trajectory.
-$^g$ Reference for the full simulation details;
-     the original publication is cited if simulation data from previously published work has been directly used,
-     for other systems the simulation details are given in the Supplementary Information.
-$^h$ Magnitudes from Fig.~S4 of Klauda et al.~\cite{klauda10}, signs matched to our simulations.
-$^i$ Magnitudes from Fig.~9 of Dickson et al.~\cite{dickson12}, signs matched to our simulations.
-}\label{systems}
-\begin{tabular}{l l r r r r r c c}
-%\hline
-Force field & lipid & $^a$N$_{\rm l}$   &  $^b$N$_{\rm w}$ &  $^c$T (K)  &  $^d$t$_{{\rm sim}}$(ns) &  $^e$t$_{{\rm anal}}$ (ns) &  $^f$Files  &  $^g$Details\\
-\hline
-Berger-DMPC-04~\cite{gurtovenko04} & DMPC & 128 & 5097 & 323 & 130 & 100 & [\citenum{dmpcFILES}]$^*$       & [\citenum{miettinen09}] \\
-Berger-DPPC-98~\cite{marrink98}    & DPPC &  72 & 2864 & 323 & 60 & 30 & [\citenum{bergerDPPCfiles}]$^*$ & SI \\
-{\bf
-Berger-POPC-07}~\cite{ollila07a}   & POPC & 128 & 7290 & 298 & 270 & 240 & [\citenum{bergerFILESpopc}]$^*$ & [\citenum{ferreira15}] \\
-CHARMM36~\cite{klauda10}  & DPPC &  72 & 2189 & 323 &  30 &  25 & [\citenum{charmmFILESdppc}]$^*$    & SI \\
-CHARMM36~\cite{klauda10}  & DPPC &  72 & 2189 & 323 &  130  & - & -   & [\citenum{klauda10}]$^h$  \\
-CHARMM36~\cite{klauda10}  & POPC &  72 & 2242 & 303 &  30 &  20 & [\citenum{charmm36filesSHORT}]$^*$ & SI \\
-{\bf
-CHARMM36}~\cite{klauda10} & POPC & 128 & 5120 & 303 & 200 & 100 & [\citenum{charmm36files}]$^*$      & SI \\
-%CHARMM36\cite{klauda10}\todoi{Not included in the Fig.
-2.
-Should we remove or include?}        & DOPC   & 128 &  5120    & 303 & 150 & 100  & ?\todoi{Permanent link in progress by Hubert Santuz}  & SI \\
-%\todoi{Permanent link for DOPC CHARMM36 simulation by Hubert Santuz is not required now. However, it would be probably useful in the future.}
-MacRog~\cite{kulig15b}  & POPC & 288 & 12600 & 310 & 100 &  80 & [\citenum{macrogFILES}]$^*$      & SI  \\
-{\bf
-MacRog}~\cite{kulig15b} & POPC & 128 &  6400 & 310 & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$  & SI  \\
-MacRog~\cite{kulig15b}  & POPC & 288 & 14400 & 310 &  90 &  40 & [\citenum{macrogdehydFILES}]$^*$ & SI  \\
-GAFFlipid~\cite{dickson12}   & DPPC &  72 & 2197 & 323 &  90 & 50 & [\citenum{GAFFlipidFILESdppc}]$^*$ & SI \\
-GAFFlipid~\cite{dickson12}   & DPPC &  72 & 2167 & 323 & 250 &  250 & -   &  [\citenum{dickson12}]$^i$   \\
-{\bf
-GAFFlipid}~\cite{dickson12}  & POPC & 126 & 3948 & 303 & 137 & 32 & [\citenum{GAFFlipidFILES}]$^*$ & SI \\
-Lipid14 \cite{dickson14}         & POPC  & 72 & 2234 & 303 & 100 & 50  & [\citenum{lipid14files}]$^*$ & SI \\
-Poger \cite{poger10}             & DPPC  & 128 & 5841 & 323 & 2$\times$100 & 2$\times$50 & [\citenum{pogerFILESpme1,pogerFILESpme2}]$^*$ & SI \\
-Slipids \cite{jambeck12}          & DPPC & 128 & 3840 & 323 & 150 & 100 & [\citenum{slipidsFILES}]$^*$ & SI \\
-Slipids \cite{jambeck12b}          & POPC & 128 & 5120 & 303 & 200 & 150 & [\citenum{slipidsFILESpopc}]$^*$ & SI \\
-Kukol \cite{kukol09}          & POPC   & 512 & 20564 & 298 & 50 & 30  & [\citenum{kukolFILES}]$^*$ & SI \\
-Chiu \cite{chiu09}      & POPC  & 128 & 3552  & 298 & 56 & 50  & [\citenum{chiuFILES}]$^*$ & SI \\
-H\"ogberg08 \cite{hogberg08}    & DMPC &  98 & 3840 & 303 &  75 & 50 &  [\citenum{hogbergDMPCfiles}]$^*$ & [\citenum{hogberg08}] \\
-H\"ogberg08 \cite{rabinovich14} & POPC & 128 & 3840 & 303 & 100 & 80 &  [\citenum{hogbergPOPCfiles}]$^*$ & [\citenum{rabinovich14}]  \\
-Ulmschneiders \cite{Ulmschneider09}    & POPC  & 128 & 3328 & 310 & 100 & 50 & [\citenum{ulmschneiderFILES}]$^*$ & SI \\
-Tj\"ornhammar14 \cite{tjornhammar14}   & DPPC  & 144 & 7056 & 323 & 200 & 100 & [\citenum{tjornhammarfiles}]$^*$ & [\citenum{tjornhammar14}] \\
-Botan-CHARMM36-UA~\cite{henin08} & DLPC & 128 & 3840 & 323 & 30 & 20 & [\citenum{charmmUAfiles}]        & SI \\
-Lee-CHARMM36-UA~\cite{lee14}     & DPPC &  72 & 2189 & 323 & 70 & 50 & [\citenum{charmmUAfilesLEE}]$^*$ & SI \\
-\end{tabular}
-\end{sidewaystable}
-
-\begin{table}[htb]
-\centering
-\caption{Simulated single component lipid bilayers with varying hydration levels. The simulation file data sets marked with $^*$ include also part of the trajectory.
-$^a$ Water/lipid molar ratio
-$^b$ The number of lipid molecules
-$^c$ The number of water molecules
-$^d$ Simulation temperature
-$^e$ The total simulation time
-$^f$ Time frames used in the analysis
-$^g$ Reference link for the downloadable simulation files
-$^h$ Reference for the full simulation details
-}\label{systemsDEHYD}
-\begin{tabular}{c c c c c c c c c c}
-%\hline
-Force field & lipid & $^a$n (w/l)   & $^b$N$_{\rm l}$   &  $^c$N$_{\rm w}$ & $^d$T (K)  & $^e$t$_{{\rm sim}}$(ns)  & $^f$t$_{{\rm anal}}$ (ns)& $^g$Files  &  $^h$Details\\
-\hline
-Berger-POPC-07~\cite{ollila07a}          &   POPC & 57  &128 & 7290  & 298  & 270 & 240 & [\citenum{bergerFILESpopc}]$^*$ & SI \\
-                                        &   POPC & 7  &128 & 896   & 298  & 60 & 50 & [\citenum{bergerDEHYDfiles}]$^*$ & SI \\
-Berger-DLPC-13~\cite{kanduc13}          &   DLPC & 28  &72 & 2016  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc28}]$^*$ & [\citenum{kanduc13}] \\
-                                 &   DLPC & 24  &72 & 1728  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc24}]$^*$ & [\citenum{kanduc13}] \\
-                                 &   DLPC & 20  &72 & 1440  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc20}]$^*$ & [\citenum{kanduc13}] \\
-                                 &   DLPC & 16  &72 & 1152  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc16}]$^*$ & [\citenum{kanduc13}] \\
-                                 &   DLPC & 12  &72 & 864  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc12}]$^*$ & [\citenum{kanduc13}] \\
-                                 &   DLPC & 8  &72 & 576  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc8}]$^*$ & [\citenum{kanduc13}] \\
-                                 &   DLPC & 4  &72 & 288  & 300  & 80 & 60 & [\citenum{bergerFILESdlpc4}]$^*$ & [\citenum{kanduc13}] \\
-CHARMM36\cite{klauda10}          & POPC   & 40 & 128 &  5120   & 303 & 150 & 100  & [\citenum{charmm36files}]$^*$   & SI \\
-                                 & POPC   & 31  & 72  &  2242 & 303 & 30 & 20 & [\citenum{charmm36filesSHORT}]$^*$ & SI \\
-                               & POPC   & 15 & 72 &  1080  & 303 & 59 & 40 & [\citenum{charmm36files15wPERl}]$^*$ & SI \\
-                            & POPC   & 7  & 72  &  504  & 303 & 60 & 20 & [\citenum{charmm36files7wPERl}]$^*$ & SI \\
-MacRog\cite{kulig15b}     & POPC   & 50 & 288  & 14400 & 310 & 90 & 40 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
-                               & POPC   & 25 & 288  & 7200 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
-                                & POPC   & 20 & 288  & 5760 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
-                                & POPC   & 15 & 288  & 4320 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
-                                & POPC   & 10 & 288  & 2880 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
-                                & POPC   & 5 & 288   & 1440 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
-GAFFlipid\cite{dickson12}      & POPC   & 31& 126  & 3948  & 303 & 137 & 32 & [\citenum{GAFFlipidFILES}]$^*$ & SI  \\
-                               & POPC   & 7& 126  & 896   & 303 & 130 & 40 & [\citenum{gaffDEHYDfiles}]$^*$ & SI \\
-\end{tabular}
-\end{table}
-
-\begin{sidewaystable}[]
-\centering
-\caption{Simulated lipid bilayers containing cholesterol. The simulation file data sets marked with $^*$ include also part of the trajectory.
-$^a$ The number of lipid molecules
-$^b$ The number of cholesterol molecules
-$^c$ Cholesterol concentration (mol\%)
-$^d$ The number of water molecules
-$^e$ Simulation temperature
-$^f$ The total simulation time
-$^g$ Time frames used in the analysis
-$^h$ Reference link for the downloadable simulation files
-$^i$ Reference for the full simulation details
-}\label{systemsCHOL}
-\begin{tabular}{c c c c c c c c c c c}
-%\hline
-Force field & lipid   & $^a$N$_{\rm l}$ & $^b$N$_{\rm chol}$ &$^c$C$_{\rm CHOL}$  &  $^d$N$_{\rm w}$ & $^e$T (K)  & $^f$t$_{{\rm sim}}$(ns)  & $^g$t$_{{\rm anal}}$ (ns)& $^h$Files  &  $^i$Details\\
-\hline
-Berger-POPC-07~\cite{ollila07a}&   POPC &128 & 0 &0\% & 7290  & 298  & 270 & 240 & [\citenum{bergerFILESpopc}]$^*$ & [\citenum{ferreira15}] \\
-/H\"oltje-CHOL-13~\cite{holtje01,ferreira13}   &    & &  &   &   &  &  &  &  \\
-                               &   POPC &120 & 8 & 6\% &7290   & 298  & 100 & 80 & [\citenum{bergerFILESpopc7chol}]$^*$ & [\citenum{ferreira13}] \\
-                               &   POPC &110 & 18& 14\% & 8481  & 298  & 100 & 80 & [\citenum{bergerFILESpopc15chol}]$^*$ & [\citenum{ferreira13}]  \\
-                               &   POPC &84 & 44 & 34\%  & 6794   & 298  & 100 & 80 & [\citenum{bergerFILESpopc34chol}]$^*$ & [\citenum{ferreira13}] \\
-                               &   POPC &64 & 64 & 50\% & 10314  & 298  & 100 & 80 & [\citenum{bergerFILESpopc50chol}]$^*$ & [\citenum{ferreira13}] \\
-                               &   POPC &50 & 78 & 61\% & 5782   & 298  & 100 & 80 & [\citenum{bergerFILESpopc60chol}]$^*$ & [\citenum{ferreira13}] \\
-CHARMM36\cite{klauda10,lim12}   & POPC   & 128& 0& 0\% & 5120  & 303  & 150 & 100 & [\citenum{charmm36files}]$^*$  & SI  \\
-                                & POPC   & 512& 0& 0\% & 23943  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
-                                & POPC   & 460& 52& 10\% & 23569  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
-                                & POPC   & 436& 76& 15\% & 23331  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
-                                & POPC   & 100 & 24 & 19\%  &  4960   & 303 & 200 & 100 & [\citenum{charmm36files20perCHOL}]$^*$ & SI \\
-                                & POPC   & 410& 102& 20\% & 20972  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
-                                & POPC   & 384& 128& 25\% & 22327  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
-                                & POPC   & 332& 180& 35\% & 21340  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
-                                & POPC   & 256& 256& 50\% & 20334  & 298  & 170 & 100 & [\citenum{charmmPOPC512files}]$^*$  & SI  \\
-                               & POPC   & 80 & 80 &50\%  &  4496    & 303 & 200 & 100 & [\citenum{charmm36files50perCHOL}]$^*$ & SI \\
-MacRog\cite{kulig15b}     & POPC   & 128 & 0 & 0\% & 6400  & 310 & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI \\ 
-                          & POPC   & 114  & 14 & 11\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
-                          & POPC   & 72   & 56 &  44\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
-                             & POPC   & 64  & 64 & 50\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
-                             & POPC   & 56   & 72 & 56\% & 6400  & 310  & 400 & 200 & [\citenum{macrogCHOLfiles}]$^*$ & SI    \\
-\end{tabular}
-\end{sidewaystable}
-
-\section{Results and Discussion}
-
-\subsection{Full hydration: Experimental order parameters for the glycerol backbone and headgroup}\label{experiments}
-The specific deuteration of $\alpha$-, $\beta$- and g$_3$- segments of DPPC has been successful, 
-allowing the absolute value order parameter measurements for these segments by $^2$H NMR~\cite{gally75,brown77,brown78,akutsu81}.
-In addition, the absolute values of order parameters for all glycerol backbone and choline headgroup segments in egg yolk lecithin~\cite{hong95a},
-DMPC~\cite{hong95b,gross97,dvinskikh05a}, DOPC~\cite{warschawski05} and POPC~\cite{warschawski05,ferreira13}
-have been measured with several different implementations of $^1$H-$^{13}$C NMR experiments. In addition, the signs of order parameters in some systems
-are measured with $^1$H-$^{13}$C NMR techniques~\cite{hong95a,hong95b,gross97}. The experimental values of glycerol backbone 
-and choline order parameters from various publications~\cite{gally75,akutsu81,gross97,dvinskikh05a,ferreira13} with the signs measured in~\cite{hong95a,hong95b,gross97} are shown in Fig.~\ref{HGorderparameters}.
-\begin{figure}[]
-%  \centering
-  \includegraphics[width=8.6cm]{../DATAreportediINblog/comparisonSorted.pdf}
-\newline
-%\todo{There is also the interactive version by Hubert Santuz now in https://plot.ly/~HubertSantuz/72/lipid-force-field-comparison/
-%we should figure out which is the most practical way to put that behind permalink once it finalized (Zenodo, figshare or something else?) and then put a citation in the paper.}
-  \caption{\label{HGorderparameters}
-  Order parameteres from simulations listed in Table~\ref{systems} and experiments for glycerol and choline groups.
-The experimental values were taken from the following publications:
- DMPC 303~K from \cite{gross97},
- DMPC 314~K from \cite{dvinskikh05a},
- DPPC 322~K from \cite{gally75},
- DPPC 323~K from \cite{akutsu81},
- POPC 296~K from \cite{bechinger91}, and
- POPC 300~K from \cite{ferreira13}.
-  The vertical bars shown for some of the computational values are not error bars, but demonstrate that for 
-these systems we had at least two data sets (see Table~\ref{systems});
-the ends of the bars mark the extreme values from the sets, and the dot marks their measurement-time-weighted average. 
-The interactive version of this figure is available at  https://plot.ly/$\sim$HubertSantuz/72/lipid-force-field-comparison/.
-} 
-\end{figure}
-
-In general there is a good agreement between the order parameters measured with different experimental NMR techniques: Almost all the 
-reported values are within a variation of $\pm$0.02 (which is also the error estimate given by Gross et al.~\cite{gross97}) 
-for all fully hydrated PC bilayer, regardless of the variation in their acyl chain composition and temperature.
-Exceptions are the somewhat lower order parameters sometimes reported from measurements using $^1$H-$^{13}$C NMR~\cite{hong95a,hong95b,warschawski05}.
-These experiments are not shown in Fig.~\ref{HGorderparameters} as the reported error bars are either relatively large~\cite{hong95a,hong95b}, 
-or the spectral resolution is quite low and the numerical lineshape simulations have not been used in the analysis~\cite{warschawski05}.
-Due to this end, it is highly likely that these reported lower order parameters are due to lower experimental 
-accuracy and therefore we exclude them from our discussion. 
-For more details, see Ref.~\citenum{accuracyPOST}. 
-Motivated by the high experimental reproducibility, we have highlighted in 
-Fig.~\ref{HGorderparameters} the subjective sweet spots (light blue areas), within which we expect the calculated absolute 
-values of the order parameters of a well-performing force field to fall.
-
-%In addition to the phosphatidylcholine lipids, similar values of the glycerol backbone order parameters have been measured
-%for phophatidylethanolamine (PE) and phopshatidylglycerol (PG) in E. Coli extract~\cite{gally81},
-%indicating that the glycerol backbone structure is similar independent of the headgroup chemistry and lipid environment.
-%Further, choline order parameters measured from mouse fibroblast L-M cell are similar to the ones in model
-%membranes~\cite{scherer87}.
-
-In addition to the numerical values, an important feature of the glycerol backbone is the 
-forking (see section~\ref{expORDp}) of the order parameters in g$_1$ and g$_3$ segments, in contrast to the choline segments $\alpha$ and $\beta$. 
-The forking in glycerol backbone g$_3$ segment is small ($\approx$ 0.02) 
-and some experiments only report the larger value or the average value~\cite{akutsu81,ferreira13}. 
-In contrast, forking is significant for the glycerol backbone g$_1$ segment, whose lower order parameter is close to zero and the
-larger one has an absolute value of approximately 0.13--0.15. Forking was studied in detail by Gally et al.~\cite{gally81}, who used E. Coli to 
-stereospecifically deuterate the different hydrogens attached to the g$_1$ or g$_3$ groups in PE lipids, and measured the order parameters from the lipid 
-extract. This experiment gave the lower order parameter when deuterium was in the S position of g$_1$ or R position for g$_3$.
-Since the glycerol backbone order parameters are very similar irrespective of the headgroup chemistry (PC,PE and PG) or lipid 
-environment~\cite{gally81}, it is reasonable to assume that the stereospecifity measured for the PE lipids
-holds also for the PC lipids.
-
-The most detailed experimentally available order parameter information for the glycerol backbone and choline 
-segments of POPC bilayer is collected by taking the absolute values from~\cite{ferreira13}, the signs from~\cite{hong95a,hong95b,gross97} 
-and the stereospecific labeling from~\cite{gally81}, and shown in Fig.~\ref{HGorderparameters2}.
-\begin{figure*}[]
-  \centering
-  \includegraphics[width=8.6cm]{../Fig/stereospecificOPs.pdf} \\
-  \caption{\label{HGorderparameters2}
-  Order parameteres for POPC glycerol and choline groups
-  from simulations with Berger-POPC-07, MacRog, GAFFlipid, and CHARMM36 force fields
-  (the {\bf bolded} systems in Table~\ref{systems})
-  together with experimental values.
-  The error bars of simulation data are standard errors of mean (see Methods for details).
-  The magnitudes for experimental order parameters are taken from Ferreira et al.~\cite{ferreira13},
-  the signs are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
-  and Gross et al.~\cite{gross97}, and the R/S labeling is based on the measurements by Gally et al.~\cite{gally81}.
-}
-\end{figure*}
-
-\subsection{Full hydration: Comparison between simulation models and experiments}
-
-The order parameters of the glycerol backbone and headgroup calculated from different force fields for various lipids have been 
-previously compared to experiments~\cite{shinoda97,hogberg08,castro08,klauda10,kapla12,dickson12,poger12,ferreira13,chowdhary13,maciejewski14}. 
-The general conclusion from these studies seems to be that the CHARMM based~\cite{hogberg08,klauda10}, GAFFlipid~\cite{dickson12} and
-MacRog~\cite{maciejewski14} force fields perform better for the glycerol backbone and headgroup structures than the GROMOS based models~\cite{castro08,kapla12,poger12,ferreira13}.
-However, none of the studies exploits the full potential of the available experimental data discussed in previous section, i.e. the quantitative accuracy, known signs and stereospecific labeling of
-the experimental order parameters.
-
-To get a general idea of the quality of the glycerol backbone and choline headgroup structures in different models, we calculated 
-the order parameters for these parts from thirteen different lipid models (Table~\ref{systems}) and 
-plotted the results together with experimental values in Fig.~\ref{HGorderparameters}.
-Two criteria were used to judge the quality of the model: there must not be significant  {\bf forking} in the $\alpha$ and $\beta$ carbons,
-there must be only moderate forking in the g$_3$ carbon and there must be significant forking in the g$_1$ carbon, the {\bf magnitude}
-should be preferably inside to the subjective sweet spots determined from experiments (blue shaded regions in Fig.~\ref{HGorderparameters}).
-The results for each force field in respect to the above criteria are summarized in Figure~\ref{FullHydrationComparisonTable}.
-\begin{figure}[]
-  \centering
-  \includegraphics[width=8.6cm]{../DATAreportediINblog/comparisonTable.pdf}
-  \newline
-  \caption{\label{FullHydrationComparisonTable}
-Rough ranking of force fields based on data of Fig.~\ref{HGorderparameters}.
-"M" indicates a magnitude problem,
-"F" a forking problem.
-Letter size shows the level (0-4) of severity;
-the $\Sigma$-column shows the sum of these, i.e., the "total severity".
-Color scheme:
-"within experimental error" (dark green),
-"almost within experimental error" (light green),
-"clear deviation from experiments" (light red), and
-"major deviation from experiments" (dark red).
-  } 
-\end{figure}
-
-None of the studied force fields fulfils these criteria completely, however CHARMM36 is pretty close. 
-This is not surprising since the dihedral potentials in this model are tuned to reproduce these parameters better against experiments~\cite{klauda10}.
-The next models in the list are CHARMM36-UA~\cite{henin08,lee14} and H\"ogberg08~\cite{hogberg08} which is also not surprising since
-these models are using CHARMM bonded potentials for glycerol backbone and choline. The fourth and the fifth models in the list, MacRog~\cite{maciejewski14} and
-GAFFlipid~\cite{dickson12}, have independently determined dihedral potentials. All the models based on Gromos potentials and Slipids perform less well.
-In the present work we subject the CHARMM36, MacRog, GAFFlipid and Berger-POPC-07 to a more careful comparison including the sterospecific labeling  
-(Fig.~\ref{HGorderparameters2}), and atomistic level structure and responses to the dehydration and cholesterol content in the following sections.
-These models are selected for more detailed studies since they are the best representatives of different dihedral potential parametrization techniques 
-(CHARMM36, MacRog, GAFFlipid), and the Berger based models are the most used lipid model in the literature.
-
-\subsection{Full hydration: Atomistic resolution structures in different models}
-
-The results in the previous section revealed significant differences of the glycerol backbone and choline headgroup
-order parameters between different molecular dynamics simulation models.
-However, it is not straightforward to conclude which kind of structural differences (if any)
-between the models the results indicate, because the mapping from the order parameters to the 
-structure is not unique. In this section we demonstrate that 1) the differences in order parameters
-indicate significantly different structural sampling strongly correlating with the dihedral angles of the related bonds,
-and that 2) the comparison between experimental and simulated order parameters can be used to exclude
-nonrealistic structural samping in molecular dynamics simulations. The demonstration is done for 
-the dihedral angles defined by the g$_3$-g$_2$-g$_1$-O(\textit{sn}-1) segments in the glycerol backbone and 
-the N-$\beta$-$\alpha$-O segments in the headgroup. These dihedrals were chosen for demonstration, because 
-significant differences between the models are observed around these segments in Fig.~\ref{HGorderparameters2}.
-We note that performing a similar comparison through all the dihedrals in all the 13 models would probably give highly useful
-information on how to improve the accuracy of the models yet this is beyond the scope of the current report.
-
-The dihedral angle distributions for the  g$_3$-g$_2$-g$_1$-O(\textit{sn}-1) dihedral calculated from different models are
-shown in Fig.~\ref{dihDISTS}. The distribution is qualitatively different for the Berger-POPC-07 model, showing a maximum in 
-the gauche$^+$-conformation (60$^o$) compared to all the other models showing a maximum in the anti-conformation (180$^o$).
-The distributions in all the other models have the same general features, the main difference being that the
-fraction of configurations in the gauche$^-$-conformation (-60$^o$) is zero for the MacRog, detectable for the CHARMM36 and
-equally large to the gauche$^+$ fraction in GAFFlipid. From the results we conclude that most likely the wrongly sampled
-dihedral angle for the g$_2$-g$_1$ bond explains the significant discrepancy to the experimental order parameters
-for the g$_1$ segment in the Berger-POPC-07 model (Fig.~\ref{HGorderparameters2}). 
-In conclusion, models preferring the anti conformation for this dihedral give more realistic order parameters and
-this is in agreement with previous crystal structure and $^1$H NMR studies~\cite{hauser80,hauser81,hauser81b,hauser88,pascher92,marsh06}.
-\begin{figure}[]
-  \centering
-  \includegraphics[width=8.6cm]{../Fig/g1-g2_Cdihs2.eps}
-  \caption{\label{dihDISTS}
-    Dihedral angle distributions for g$_3$-g$_2$-g$_1$-O(\textit{sn}-1) dihedral from different models (POPC bilayer in full hydration).
-      } 
-\end{figure}
-
-The dihedral angle distribution for the  N-$\beta$-$\alpha$-O dihedral calculated from the same four models is 
-shown in Fig.~\ref{dihDISTS2}. Also for this dihedral there are significant differences in the gauche--anti fractions.
-The gauche conformations are dominant in the CHARMM36, in MacRog there are only anti conformations present,
-and in the Berger-POPC-07 and GAFFlipid gauche and anti conformations have equal probabilities. 
-On the other hand, comparison of $\alpha$ and $\beta$ order parameters in Fig.~\ref{HGorderparameters2}
-reveals that for these carbons the CHARMM36 is closest to the experimental results and it is also the only model that has the correct
-sign (negative) for the $\beta$ order parameter. This result is again in agreement with previous 
-crystal structure, $^1$H NMR and Raman spectroscopy studies~\cite{hauser80,hauser81,hauser81b,akutsu81b} which suggest that
-this dihedral is in the gauche conformation in the absence of ions.
-
-%Interestingly, the probability of the gaughe conformations correlates with the order parameter difference between the $\beta$ and $\alpha$ segments:
-%the larger the gaughe fraction the larger the order parameter difference. This suggestion together with the results in 
-%Fig.~\ref{HGorderparameters2} would indicate that the correct gaughe-trans fraction for N-$\beta$-$\alpha$-O dihedral is 
-%larger than in GAFFlipid but smaller than in CHARMM36. This happens to be quite close to the gaughe-trans fraction in
-%the Berger model. However, there is significant forking and numerical values are off from the experiments in the Berger
-%model suggesting that it has some other inaccuracies.
-\begin{figure}[]
-  \centering
-  \includegraphics[width=8.6cm]{../Fig/a-bDIHS2.eps}
-  \caption{\label{dihDISTS2}
-    Dihedral angle distributions for N-$\beta$-$\alpha$-O dihedral from different models (POPC bilayer in full hydration).
-  } 
-\end{figure}
-
-The used examples show that the glycerol backbone and headgroup order parameters reflect the atomistic resolution structure
-and that the comparison with experiments allows the assessment of the quality of the suggested structure. We were able to pinpoint
-specific problems in the structures in different models and suggest potential improvement strategies.
-If the improved atomistic
-molecular dynamics simulation model reproduced the order parameters and other experimental observables (like chemical shift anisotropy)
-with experimental accuracy, it would give an interpretation for the atomistic resolution structure of the glycerol backbone and 
-choline~\cite{seelig77b,skarjune79,jacobs80,davis83,akutsu91,hong95b,semchyschyn04}. The research along these lines is left, however,
-for future studies.
-
-\subsection{Response to dehydration and cholesterol content}
-In addition to pure phosphatidylcholine bilayers at full hydration, the choline headgroup order parameters
-have been measured under various different conditions~\cite{gally75,brown77,brown78,akutsu81,altenbach84,scherer89,bechinger91,ulrich94,dvinskikh05b,castro08,kapla12,ferreira13}.
-Also the order parameters for the glycerol backbone have been measured with $^1$H-$^{13}$C NMR in dehydrated conditions~\cite{dvinskikh05b}, and as a function 
-of anesthetics~\cite{castro08} and glycolipids~\cite{kapla12} for DMPC, and as a function of cholesterol 
-concentration for POPC~\cite{ferreira13}. Due to the high resolution in the NMR (especially $^2$H NMR) experiments,
-even very small order parameter changes resulting from the varying conditions can be measured 
-(see Ref.~\citenum{accuracyPOST} for more discussion.) 
-However, as already discussed above, it is not simple to deduce 
-the structural changes from order parameter changes~\cite{akutsu91,semchyschyn04}. Consequently, comparison of the order parameters
-between simulations and experiments in different conditions can be used to assess the quality of the force field 
-in different situations, and, if the quality is good, to interpret the structural changes in experiments.
-Here we exemplify such comparison for a lipid bilayer under low hydration levels and when varying amounts of cholesterol is included in the bilayer. 
-The interaction between ions and a phosphatidylcholine bilayer will be discussed in a separate study~\cite{ionpaper}.
-
-\subsubsection{Phospholipid bilayer with low hydration level}
-Fig.~\ref{ordPhydr} shows the published~\cite{dvinskikh05b,ulrich94,bechinger91} experimental order parameters
-for the glycerol backbone and choline as a function of hydration level.
-Despite slight differences in temperature and acyl chain composition,
-the three independently reported data sets for the choline ($\beta$ and $\alpha$) segments agree well with each other:
-Both order parameters increase with decreasing hydration level.
-The glycerol backbone order parameters (g$_{3}$, g$_{2}$, g$_{1}$), in contrast,
-have been observed~\cite{dvinskikh05b} to slightly decrease with dehydration.
-Note that the original experiments~\cite{dvinskikh05b,ulrich94,bechinger91} measured only 
-absolute values, but here
-we included the signs measured in separate studies~\cite{hong95a,hong95b,gross97}. 
-Consequently, the negative $\beta$ order parameter actually increases with dehydration 
-as its absolute value decreases~\cite{dvinskikh05b,ulrich94,bechinger91}.
-\begin{figure}[]
-  \centering
-  \includegraphics[width=7.3cm]{../DATAreportediINblog/dehydration_lp.pdf}
-    \includegraphics[width=7.3cm]{../DATAreportediINblog/dehydration_smooth.pdf}
-    \caption{\label{ordPhydr}
-    The effect of dehydration on glycerol and choline order parameters in experiments.
-    The magnitudes of order parameters are measured for DMPC ($^1$H-$^{13}$C NMR) at 314~K~\cite{dvinskikh05b}, 
-    for POPC ($^2$H NMR) at 296~K~\cite{bechinger91} and for DOPC ($^2$H NMR) at 303~K~\cite{ulrich94}. 
-    The signs are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
-    and Gross et al.~\cite{gross97}.
-    Note that to elucidate the relative change as a function of hydration level,
-    the simulation results are shifted such that the (smaller) $S_\mathrm{CH}$
-    matches (within $\pm$0.01) the experimental value at full hydration;
-    the shift magnitudes for each of the force fields are listed ($S_\mathrm{CH}$+shift) in the y-label.
-  }
-  	\todo{Markus prepared two versions of this plot. The one on left has all the data points joint by lines;
-	the one on right shows smooth trend lines (cubic polynomials) to ease qualitative comparison.
-	Samuli likes the one on left, Markus the one on right. Do others have opinions?}
-\end{figure}
-
-Lipid bilayer dehydration has been studied also with molecular dynamics simulations~\cite{mashl01,pertsin05,pertsin07,eun09,eun10,schneck12},
-typically motivated by the  discussion concerning the origin of the ``hydration repulsion''~\cite{israelachvili,israelachvili96,sparr11}.
-Only one~\cite{mashl01} of these studies, however, compared their simulation model to the experimental choline and glycerol backbone
-order parameters.
-Fig.~\ref{ordPhydr} shows these order parameters as a function of hydration level for the CHARMM36, 
-MacRog and GAFFlipid models (having the most realistic atomistic resolution structures) and a Berger-based model 
-(which is the most used lipid model);
-note that the simulation results have been vertically shifted to ease the comparison with experimental response to dehydration.
-Despite of some fluctuations, the increase of the choline ($\beta$ and $\alpha$) order parameters is seen in all four
-models. The response of the choline order parameters to dehydration can, therefore, be
-interpreted to qualitatively agree with experiments.
-The situation is significantly more complicated for the glycerol backbone: 
-None of the four models produced the experimentally seen trends
-in all the (g$_{3}$, g$_{2}$, g$_{1}$) segments.
-
-The qualitative agreement of the $\alpha$ and $\beta$ order parameters with experiments in all four simulation models (Fig.~\ref{ordPhydr}) indicates that,
-despite the unrealistic structures at full hydration (Figs.~\ref{HGorderparameters} and~\ref{FullHydrationComparisonTable}),
-the structural response of the choline headgroup to dehydration is somewhat realistic.
-A likely explanation is that as the interlamellar space shrinks with dehydration, the whole choline group
-orients more parallel to the membrane.
-Indeed, upon dehydration the angle between P--N (phoshate phosphorus to choline nitrogen) vector and membrane normal increases for
-all the four models (Fig.~\ref{PNangle}). However,
-the amount of increase depends on the model. Especially the DLPC simulations with Berger model
-predict significantly stronger P--N vector tilt than the other models. The Berger model
-has also generally larger P--N vector angles and its choline order parameters are more off from
-experiments than the other three models (Fig.~\ref{HGorderparameters2}). Thus the relatively modest tilting with dehydration
-predicted by MacRog, CHARMM36 and GAFFlipid is probably more realistic.
-
-It must be stressed that in the models
-incapable of reproducing the experimental order parameters the free energy landscape is not correct. Thus
-even though the order parameter response to dehydration is qualitatively correct,
-the energetic response is likely to be incorrect.
-This may have some influence on dehydration energetic calculations made using the Berger model~\cite{eun09,schneck12}.
-\begin{figure}[]
-  \centering
-  \includegraphics[width=8.6cm]{../Fig/dehydrationPN.pdf}
-
-  \caption{\label{PNangle}
-    The average angle between membrane normal and P--N vector as function of
-    hydration level calculated from different simulations.
-  }
-\end{figure}
-
-The response of the glycerol backbone seems to be more subtle than that of the choline headgroup;
-none of the four models reproduced the experimental trends upon dehydration with enough accuracy to invite a
-structural interpretation.
-
-\subsubsection{Cholesterol-containing phospholipid bilayer}
-As cholesterol is abundant in biological membranes and
-has been suggested to be an important player, for example, in domain formation~\cite{simons04,somerharju09},
-phospholipid--cholesterol interactions have been extensively studied with theoretical~\cite{huang99,zhu07,rog09,alwarawrah12} and
-experimental~\cite{brown78,marsh10,ferreira13,marsh13} methods.
-It is widely agreed that cholesterol orders lipid acyl tails and thus decreases the area per molecule (condensing effect),
-but its influence on the lipid headgroup and glycerol backbone remains debated~\cite{huang99,simons04,somerharju09}.
-It has been suggested, for example, that the surrounding phospholipids shield cholesterol from exposure to water by 
-reorienting their headgroups (``umbrella model'')~\cite{huang99} or that cholesterol acts as a spacer between the headgroups to increase 
-their entropy and dynamics (``superlattice model'')~\cite{somerharju09}. 
-%\todo{It has been suggested that we should use direct quotations from original papers here for clarity.}
-%Both of these suggestions 
-Molecular dynamics simulations have supported both
-the umbrella~\cite{alwarawrah12} as well as the superlattice~\cite{zhu07} model,
-in addition to suggesting specific interactions of cholesterol with the glycerol backbone~\cite{rog09}.
-In these studies~\cite{zhu07,rog09,alwarawrah12}
-the responses of the glycerol backbone and choline headgroup 
-to increasing cholesterol content were not, however, compared to experiments.
-
-Fig.~\ref{ordPchol} shows the responses of the choline headgroup ($\beta$ and $\alpha$)
-order parameters of POPC (measured by $^1$H-$^{13}$C NMR~\cite{ferreira13}) and DPPC
-($^{2}$H NMR~\cite{brown78}) to increasing cholesterol content.
-Again, the two independent data sets agree very well:
-Only very modest ($\Delta S_\mathrm{CH}<0.03$) changes occur in  the choline order parameters as cholesterol content
-increases from 0 to 60\%.
-The extreme sensitivity of the
-high resolution $^{2}$H NMR experiments
-is beautifully demonstrated by the measurable~\cite{brown78}
-(but barely visible on the scale used in Fig.~\ref{ordPchol})
-cholesterol-induced forking of the $\alpha$ order parameter.
-
-We note that the modest ($\Delta S_\mathrm{CH}<0.02$ for g$_1$; $<0.04$ for g$_2$, g$_3$; see Fig.~\ref{ordPchol})
-effects of cholesterol
-on the glycerol backbone order parameters of POPC
-measured by $^1$H-$^{13}$C NMR~\cite{ferreira13}
-agree well with the results for phosphatidylethanolamine (PE) measured by $^{2}$H NMR~\cite{ghosh82}.
-This further supports the ideas that the glycerol backbone structural behaviour is independent of the
-headgroup composition~\cite{gally81} and that the headgroup structure is largely independent of the acyl chain region content unless
-charges are present~\cite{scherer87}.
-\begin{figure}[]
-  \centering
-  \includegraphics[width=7.3cm]{../DATAreportediINblog/cholesterolization_lp.pdf}
-    \includegraphics[width=7.3cm]{../DATAreportediINblog/cholesterolization_smooth.pdf}
-  \caption{\label{ordPchol}
-    The effect of cholesterol content on the glycerol backbone and choline order parameters in experiments~\cite{brown78,ferreira13} and simulations
-    with the Berger-POPC-07/H\"oltje-CHOL-13, CHARMM36 and MacRog force fields. The signs in the experimental values are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
-    and Gross et al.~\cite{gross97}.  Most order parameters from Berger-POPC-07/H\"oltje-CHOL-13 model for g$_1$ are beoynd the y-axis scale.
-    In order to elucidate the relative change as a function of cholesterol content,
-    the simulation results are shifted such that the (smaller) $S_\mathrm{CH}$
-    matches (within $\pm$0.01) the experimental value without cholesterol;
-    the shift magnitudes for each of the force fields are listed ($S_\mathrm{CH}$+shift) in the y-label.
-}
-	\todo{Markus prepared two versions of this plot. The one on left has all the data points joint by lines;
-	the one on right shows smooth trend lines (cubic polynomials) to ease qualitative comparison.
-	Samuli likes the one on left, Markus the one on right. Do others have opinions?}
-\end{figure}
-
-In addition to the experimental data, Fig.~\ref{ordPchol} shows
-our results for the CHARMM36 and MacRog force fields
-and the previously published~\cite{ferreira13}
-Berger-POPC-07/H\"oltje-CHOL-13 results.
-Note that the simulation data are shifted vertically to ease comparison with experimental responses.
-As previously pointed out~\cite{ferreira13}, the Berger-based model
-seriously exaggerates the effect of cholesterol on the phospholipid glycerol backbone and choline headgroup.
-In comparison, the choline and glycerol backbone responses of CHARMM36 and MacRog are in better qualitative 
-agreement with experiments. 
-Therefore, to resolve the nature of cholesterol-induced structural changes,
-we calculated from CHARMM36 the glycerol backbone dihedral angle distributions
-at various cholesterol contents (Supplementary material): The only detectable change
-is the small decrease of gauche(-) and increase of gaughe(+) probability of the g$_3$-g$_2$-g$_1$-O(\textit{sn}-1) dihedral.
-In conclusion, our results suggest that the significant effects of cholesterol on lipid conformations observed in simulations~\cite{zhu07,rog09,alwarawrah12} are overestimated by the computational models used;
-cholesterol only induces very small structural changes in the glycerol backbone.
-
-Finally, it is important to note that the CHARMM36 force field parameters (glycerol backbone dihedral potentials)
-have been tuned to reproduce the experimental order parameters at full hydration~\cite{klauda10}. 
-This approach introduces a risk of overfitting, which would manifest itself as wrong responses to changing conditions. 
-Interestingly, according to our results, tuning did not lead to overfitting problems as far as dehydration or cholesterol content are considered.
-
-\pagebreak
-\section{Conclusions}
-The atomistic resolution structures sampled by the glycerol backbone and choline headgroup
-in phoshatidylcholine bilayers are not known despite of vast amount of accurate experimental
-data. An atomistic resolution molecular dynamics simulation model that would reproduce the
-experimental data would automatically resolve the structures, thus giving an unprecedently detailed interpretation of the experimental data.
-In this work we have collected and reviewed the experimental C--H bond vector order
-parameters available in the literature. These accurate experimental data are then compared to 13
-different atomistic resolution simulation models for a fully hydrated lipid bilayer system, followed by bilayers dehydrated to different extents, and
-finally bilayers containing various amounts of cholesterol. We are led to the following four main conclusions:
-
-(1) The C--H bond order parameters measured with different NMR techniques are consistent.
-By combining the experimental results from various sources we concluded
-that the order parameters for each C--H bond are known with a quantitative accuracy of $\pm$0.02.
-
-(2) Comparison of order parameters between experiments and different atomistic resolution models 
-together with structural analysis showed that the order parameters can be used
-to judge the structural accuracy of a model. Thus the combination of atomistic resolution 
-molecular dynamics simulations and NMR experiments can be used to resolve the atomistic resolution
-structures of biomolecules in biologically relevant conditions. 
-This approach can be extended from lipids to, for example, membrane proteins.
-
-%None of the tested models (13 different models) produces the order parameters with the experimental
-%accuracy for a fully hydrated phoshatidylcholine lipid bilayer. However, the CHARMM36, GAFFlipid and MacRog 
-%models are relatively close. The structures of these models together with the most used lipid model (Berger based) 
-%were subjected to more careful studies. The results revealed that the current models are not accurate
-%enough to resolve the atomistic resolution structures sampled by glycerol backbone and choline headgroup.  
-%However, the correlation between dihedral angle distributions and order parameter differences was found, 
-%suggesting that careful adjustment of dihedral potentials would potentially lead to the model with a correct
-%structure.
-
-(3) The review of previous experimental results revealed that when a bilayer is dehydrated the 
-choline order parameters increase. Our simulations suggested that
-this can be explained by the P--N vector tilting more parallel to the membrane. 
-This strongly supports and complements the idea that charge-induced choline tilting
-can be measured using order parameter changes~\cite{ionpaper,scherer89}.
-
-%Independent of the accuracy for a fully hydrated lipid bilayer, all the models reproduced the choline response
-%to dehydration. This can be explained by the change in the P--N vector tilting more parallel to the membrane,
-%which leads to the increase of order parameters despite of the initial configuration. It should be noted, however,
-%that the correct qualitative response does not necessarily indicate correct energetics.
-
-(4) Only modest changes of glycerol backbone and choline order parameters are observed experimentally with increasing cholesterol content.
-When interpreted using the computational lipid model that we found to have the most realistic response to cholesterol,
-this observation means that cholesterol induces only minor changes in (the g$_3$-g$_2$-g$_1$-O(sn-1)
-dihedral of the glycerol backbone, in other words, there is no major conformational change of the lipid.
-
-%- The response of glycerol backbone and choline headgroup to the cholesterol content is described more
-%realistically in CHARMM36 and MacRog models than in the Berger based model.
-
-(+) Besides these four main conclusions, we note that we have created the most extensive
-publicly available collection of molecular dynamics simulation trajectories of lipid bilayers
-(\url{https://zenodo.org/collection/user-nmrlipids}). The mere existence of
-this collection opens up numerous possibilities for unforeseen analyses, such as data mining,
-and rapid testing of ideas with
-much less computational effort than previously required.
-
-In general, we conclude that in order to fully utilize the potential of atomistic-resolution classical molecular dynamics simulations
-in the structural interpretation of high resolution NMR data~\cite{ferreira14} for lipid bilayers, one must  
-improve the phoshatidylcholine glycerol backbone and choline headgroup parameters of the existing force fields.
-
-This work has been done as a fully open collaboration, using \url{nmrlipids.blogspot.fi} as the communication
-platform. All the scientific contributions have been communicated publicly through this blog~\cite{nmrlipids}.
-
-%This work has been, and continues to be, progressed and discussed through a blog at \url{nmrlipids.blogspot.fi}. 
-%Everyone is invited to join the discussion and make contributions through the blog. 
-%The manuscript will be eventually submitted to an appropriate scientific journal. 
-%Everyone who has contributed to the work through the blog will be offered 
-%coauthorship. For more details, see \url{nmrlipids.blogspot.fi}.
-
-%\listoftodos
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The "Acknowledgement" section can be given in all manuscript
-%% classes.  This should be given within the "acknowledgement"
-%% environment, which will make the correct section or running title.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{acknowledgement}
-
-AB and CL acknowledges financial support from the French National Research Agency
-(ANR: “Biolubrication by phospholipid membranes” Bioub2012) and
-computing time allocation from P{\^o}le Scientifique de Mod{\'e}lisation Num{\'e}rique from the ENS Lyon (PSMN),
-and Centre Informatique National de l'Enseignement Sup{\'e}rieur (CINES, Montpelier, France)
-(Project c2015096850).
-FFR acknowledges CONACYT and DGAPA UNAM IG100513
-for financial support, Cluster H\'ibrido de Superc\'omputo Xiuhcoatl - CINVESTAV and Miztli - UNAM for computational resources. 
-MJ and JT acknowledge and CSC - IT Center for Science for computational resources (project number tty3979).
-MJ also acknowledges the Finnish Doctoral Programme in Computational Sciences (FICS) for funding.
-JT, MJ, TR and WK acknowledge the funding from the Academy of Finland (Centre of Excellence program) 
-and the European Research Council (Advanced Grant project CROWDED-PRO-LIPIDS).
-WK acknowledges CSC -- IT Centre for Science (Espoo, Finland) for excellent computational resources (project number tty3995).
-MSM acknowledges financial support from the Volkswagen Foundation (86110).
-LM acknowledges funding proved by the Institut national de la sant\'e et de la recherche m\'edicale (INSERM).
-JM acknowledges CSC -- IT Center for Science for computational resources.
-OHSO acknowledges Tiago Ferreira and Paavo Kinnunen for useful discussions, 
-the Emil Aaltonen foundation for financial support, 
-Aalto Science -- IT project and CSC -- IT Center for Science for computational resources. 
-HS acknowleges Catherine Etchebest and St\'ephane T\'eletch\'ea for useful discussions and continued support, the HPC 
-resources granted from GENCI-CINES (Grant 2014-c2014077209) and computer facilities provided by R\'egion Ile de France and INTS (SESAME 2009 project).
-
-\end{acknowledgement}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The same is true for Supporting Information, which should use the
-%% suppinfo environment.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{suppinfo}
-
-Simulation details, one figure and author contributions.
-
-%This will usually read something like: ``Experimental procedures and
-%characterization data for all new compounds. The class will
-%automatically add a sentence pointing to the information on-line:
-
-\end{suppinfo}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The appropriate \bibliography command should be placed here.
-%% Notice that the class file automatically sets \bibliographystyle
-%% and also names the section correctly.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newpage
-\begin{center}
-{\bf SUPPLEMENTARY INFORMATION}
-\end{center}
-\subsection{Simulation details} 
-\subsubsection{Berger based models}
+%\begin{center}
+%{\bf SUPPLEMENTARY INFORMATION}
+%\end{center}
+\section{Simulation details} 
+\subsection{Berger based models}
 For the Berger based models we use here the following naming convention: 
@@
 
-\subsubsection{CHARMM36}
+\subsection{CHARMM36}
 
@@
 
-\subsubsection{MacRog}
+\subsection{MacRog}
 The lipid force field parameters were obtained from the developers and the parameters are presented in Refs.~\citenum{maciejewski14,kulig15b}. 
@@
 
-\subsubsection{GAFFLipid}
+\subsection{GAFFLipid}
 The initial structure in Lipidbook \cite{domanski10} had different glycerol backbone isomers in different leaflets. 
@@
 
-\subsubsection{Lipid14}
+\subsection{Lipid14}
 The initial structure was taken directly from the Lipidbook~\cite{domanski10}.
@@
 
-\subsubsection{Poger et al.}
+\subsection{Poger et al.}
 The Poger lipids are derived from GROMOS G53A6~\cite{poger10} and were initially coined 53A6-L (L for lipids). They are now part of GROMOS G54A7~\cite{poger12} 
@@
 
-\subsubsection{Slipids}
+\subsection{Slipids}
 Initial coordinates for a hydrated DPPC (at 323~K) and POPC (at 310~K) bilayers (30 and 40 waters/lipid, respectively) were taken directly from the 
@@
 
-\subsubsection{Kukol}
+\subsection{Kukol}
 A bilayer patch with 512 POPC lipids was constructed and hydrated with $\sim$40 SPC water molecules per lipid. 
@@
 
-\subsubsection{Chiu et al.}
+\subsection{Chiu et al.}
 The force field parameters and the initial configuration were available through the Lipidbook~\cite{domanski10}.
@@
 
-\subsubsection{Ulmschneider}
+\subsection{Ulmschneider}
 The initial structure containing 128 POPC molecules with 3328 TIP3P water~\cite{jorgensen83} molecules (26 per lipid) was downloaded from Lipidbook \cite{domanski10} 
@@
 
-\subsubsection{Tj\"ornhammar et al.}
+\subsection{Tj\"ornhammar et al.}
 The gel phase DPPC bilayer structure delivered by Tj\"ornhammar  and Edholm~\cite{tjornhammar14} was ran for 5~ns at 343~K in order to destroy the 
@@
 
-\subsubsection{Lee-CHARMM36-UA}
+\subsection{Lee-CHARMM36-UA}
 A hydrated bilayer consisting of 72 DLPC lipids and 2189 water molecules
@@
 
-\subsubsection{Botan-CHARMM36-UA}
+\subsection{Botan-CHARMM36-UA}
 This model is similar to the Lee-CHARMM36-UA model with a few  differences.
@@
 
-\subsection{Dihedral angle distributions as a function of cholesterol in CHARMM36}
+\section{Dihedral angle distributions as a function of cholesterol in CHARMM36}
 \begin{figure*}[]
@@
 
-\subsection{Author Contributions}
+\section{Author Contributions}
 {\it Alexandru Botan} provided simulation results for charmm36-UA\\
@@
 {\it Jukka M{\"a}{\"a}tt{\"a}}  prepared and performed simulations with Berger and Slipids models and analyzed the results.\\
-{\it O. H. Samuli Ollila}Co-designed the project with MSM and managed the work. Ran and analyzed several simulations. Wrote the manuscript. \\
+{\it O. H. Samuli Ollila} Co-designed the project with MSM and managed the work. Ran and analyzed several simulations. Wrote the manuscript. \\
 {\it Marius Retegan} Prepared and validated the GROMACS compatible parameter files for GAFFlipid and Lipid14 force fields.\\

### 2015-05-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Separate main text and supplementary

@@
 
+We acknowledge all the discussion participants at nmrlipids.blogspot.fi.
 AB and CL acknowledges financial support from the French National Research Agency
@@
 
-\newpage
-\begin{center}
-{\bf SUPPLEMENTARY INFORMATION}
-\end{center}
-\subsection{Simulation details} 
-\subsubsection{Berger based models}
-For the Berger based models we use here the following naming convention: 
-Berger - \{{\it molecule name}\} - \{{\it year when model published first time}\} \{{\it citation}\}.
-The reason is that there are several different molecular topologies which are using the non-bonded parameters originally
-developed by Berger et al.~\cite{berger97}. Thus the common factor in the Berger based models are the non-bonded parameters,
-while the molecule specific parameters might somewhat vary. However, the majority of the molecular level topologies are 
-relying (especially for the glycerol backbone and headgroup) on the parameters originally introduced by Marrink et al.~\cite{marrink98}.
-This is the case for all the Berger based simulations discussed in this work.
-
-POPC simulations at full hydration at 298~K and simulations studying the effect of cholesterol are the same as in previous publications~\cite{ferreira13,ferreira15}.
-In these simulation the POPC parameters introduced by Ollila et al~\cite{ollila07a} are used, which are using the non-bonded parameters of Berger~\cite{berger97}
-and a molecular topology from Tieleman et al.~\cite{tieleman99} with improved double bond dihedrals by Bachar et al.~\cite{bachar04}. 
-Thus they are called Berger-POPC-07~\cite{ollila07a}. The cholesterol model is based on the parameters by H\"oltje et al.~\cite{holtje01} with the
-exception that the atom types were changed from CH2/CH3 to LP2/LP3 to avoid overcondensation of the bilayer as suggested in ref.~\cite{tieleman06}.
-Since this modification was introduced by Ferreira et al.~\cite{ferreira13}, we call the used cholesterol model as H\"oltje-CHOL-13~\cite{ferreira13}.
-
-For the POPC at 323~K and POPC in low hydration the same force field parameters are used.
-For DPPC the implementation of Berger parameters~\cite{berger97} by Peter Tieleman et al. are used~\cite{marrink98}.
-For all of these simulations a timestep of 2~fs was used with a leapfrog integrator. Covalent bond lengths were constrained with the LINCS algorithm~\cite{hess97,hess07}. 
-Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with real space cut-off of 1.0~nm was used 
-for electrostatics. Plain cut-off was used for the Lennard-Jones interactions with a 1.0~nm cut-off.
-The neighbor lists with cut-off of 1.0~nm were updated every 5 steps. Temperature was coupled separately
-for lipids and water to 298~K using the velocity-rescale method~\cite{bussi07} with coupling constant 0.1~ps.
-Pressure was semi-isotropically coupled to the atmospheric pressure with the Berendsen method~\cite{berendsen84}.
-
-\subsubsection{CHARMM36}
-
-{\it DPPC and POPC with 72 lipids.}
-%
-The starting structures in PDB format were downloaded from the
-NIH/NHLBI Laboratory of Computational Biology Membrane Biophysics Section
-website (\url{http://www.lobos.nih.gov/mbs/coords.shtml}),
-which refers to these as the final structures (for DPPC after 40~ns, POPC after 35~ns)
-of the NPT lipid bilayer trajectories used in the original CHARMM36 publication~\cite{klauda10}.
-The TIP3P~\cite{jorgensen83} water model was used to solvate the system.
-%
-The publicly available CHARMM36 parameters in Gromacs format
-(September 2013 update: {\tt charmm36\_gmx\_format\_sep13.tgz})
-from the MacKerell Lab website
-(\url{http://mackerell.umaryland.edu/CHARMM_ff_params.html}) were used.
-Timestep of 1~fs was used with the leapfrog integrator. Covalent bonds with hydrogens were constrained with LINCS algorithm~\cite{hess97,hess07}. 
-Coordinates were written every 5~ps. PME~\cite{darden93,essman95} with real space cut-off of 1.4~nm was used 
-for electrostatics. Lennard-Jones interactions were switched to zero between 0.8~nm and 1.2~nm.
-The neighbour lists with a cut-off of 1.4~nm were updated every 5 steps. Temperature was coupled separately
-for lipids and water to 303~K using the velocity-rescale method~\cite{bussi07} with coupling constant of 0.2~ps.
-Pressure was semi-isotropically coupled to the atmospheric pressure with the Berendsen method~\cite{berendsen84}.
-
-{\it POPC with 128 lipids}
-The starting structures for the pure POPC simulations was taken from the Slipids~\cite{jambeck12b} website (\url{http://mmkluster.fos.su.se/slipids/Downloads.html}).
-The starting structures for mixed POPC/Cholesterol simulations were constructed with the CHARMM-GUI website~\cite{jo08}. 
-They contained 100 POPC/24 cholesterol molecules and 80 POPC/80 cholesterol molecules for
-the simulations of 20\% cholesterol and 50\% cholesterol respectively. The TIP3P water model~\cite{jorgensen83} was used to
-solvate the system.
-The publicly available CHARMM36 forcefield parameters
-(\url{http://www.gromacs.org/@api/deki/files/184/=charmm36.ff\\\_4.5.4\_ref.tgz}) 
-by Piggot et al. \cite{piggot12} were used. Cholesterol parameters came
-from Lim et al. \cite{lim12} and were converted into GROMACS format with the PyTopol tool~\cite{salari15}.  
-Single point energy calculation was done to assess the conversion. 
-%(At the time I've made the simulations, the charmm36 ff for cholesterol in gromacs format was not available at
-%http://mackerell.umaryland.edu/charmm_ff.shtml#gromacs).
-Simulations were performed for 200~ns and the last 100~ns was used for the calculations. Timestep of 2~fs was
-used with leapfrog integrator. All bond lengths were constrained with LINCS~\cite{hess97,hess07}. Temperature was maintened at
-303~K with the velocity-rescale method~\cite{bussi07} and a time constant of 0.2~ps. Pressure was maintained semiisotropically
-at 1~bar using the Parrinello--Rahman algorithm~\cite{parrinello81} with a time constant of 1.0~ps. The neighbour list with a cut-off of 1.2~nm
-was updated every 10 steps. Lennard-Jones interactions were switched to zero
-between 0.8~nm and 1.2~nm. PME~\cite{darden93,essman95} with real space cut-off of 1.2~nm was used for electrostatics.
-
-{\it POPC with 512 lipids.}
-%
-POPC The starting structures for simulations were constructed with
-the CHARMM-GUI website~\cite{jo08}. 512 POPC lipids (256 per leafleat) were 
-used for the initial 0\% concentration and they were subsequently 
-substituted by cholesterol molecules to fulfil the desired concentration.
-The TIP3P~\cite{jorgensen83} water model was used to solvate the system. 
-The publicly available port of the CHARMM36 forcefield parameters (\url{mackerell.umaryland.edu/CHARMM_ff_params.html}) 
-was used for both POPC and cholesterol. 
-Simulations were performed for 170 ns and the last 100 ns was used for the calculations.
-Timestep of 1 fs was used with the leapfrog integrator. Covalent bonds with hydrogen 
-were constrained with LINCS algorithm. Coordinates were written every 10 ps.
-PME with real space cut-off of 1.2 nm was used for electrostatics. Lennard-Jones 
-interactions were switched to zero between 0.8 nm and 1.2 nm. The neighbour lists with a
-cut-off of 1.2 nm were updated every 5 steps. Temperature was coupled separately for cholesterol, lipids
-and water to 298 K using the velocity-rescale method with coupling constant of 0.2 ps.
-Pressure was semi-isotropically coupled to the atmospheric pressure with the Berendsen
-method.
-
-\subsubsection{MacRog}
-The lipid force field parameters were obtained from the developers and the parameters are presented in Refs.~\citenum{maciejewski14,kulig15b}. 
-%correspond to the published DPPC parameters~\cite{maciejewski14} with the inclusion of the double bond parameters~\cite{kulig15}. 
-%This inclusion of unsaturated lipid tails will be published in the near future. 
-%\todo{You have recent paper where you use POPC/chol simulations with this model: Kulig et al. BBA 1848 (2015) 422-432 http://dx.doi.org/10.1016/j.bbamem.2014.10.032.
-%would this be a correct refence here?}
-A bilayer with 288 POPC lipids was hydrated with 12600 TIP3P water~\cite{jorgensen83} molecules ($\sim$44/lipid) and simulated for 100~ns with a time step of 2~fs. Data was saved 
-every 10~ps and the first 20~ns of the trajectory was discarded from the analysis. 
-All bond lengths were constrained with LINCS~\cite{hess97,hess07}. The temperatures of the lipids and the solvent were separately coupled to the Nos\'{e}--Hoover thermostat~\cite{nose84,hoover85} 
-with a target temperature of 310~K and a time constant of 0.4~ps. Semi-isotropical pressure coupling to 1~bar was obtained with the Parrinello--Rahman 
-barostat~\cite{parrinello81} with a time constant of 1~ps. PME~\cite{darden93,essman95} was employed to calculate the long-range electrostatic interactions. Lennard-Jones interactions were cut off 
-at 1~nm and the dispersion correction was applied to both energy and pressure. A neighbour list with a radius of 1~nm was updated every step.
-
-Identical parameters were employed for both full hydration and for the dehydration simulations. The dehydration simulations were also run for 100~ns 
-with data saved every 10~ps.
-
-The initial structures for the simulations with 10, 40, 50 and 60 mol\% of cholesterol were obtained by replacing 14, 56, 64 or 72 POPC molecules 
-with cholesterol molecules in the initial structure containing 128 POPC molecules. These systems were simulated for 400 ns and the first 200 ns was 
-discarded from analysis. Data was saved every 100 ps.
-
-\subsubsection{GAFFLipid}
-The initial structure in Lipidbook \cite{domanski10} had different glycerol backbone isomers in different leaflets. 
-To generate the initial structure we took the structure delivered by Slipids developers~\cite{jambeck12b}. Also this structure
-had one lipid with different glycerol backbone isomer. This lipid and one lipid from opposite leaflet were removed
-after the system was equilibrated.
-
-The force field parameters were generated using files obtained from the Lipidbook website (\url{http://lipidbook.bioch.ox.ac.uk/package/show/id/150.html})~\cite{domanski10}. 
-The conversion to GROMACS compatible formats was performed using the acpype tool~\cite{silva12}. The accuracy of the conversion was checked by calculating 
-the total energy of a single POPC lipid molecule using the sander program which is part of the AmberTools14 package~\cite{ferrer13} and version 4.6.5 of GROMACS. 
-A difference of 0.002 kcal/mol was obtained between the two programs.
-
-Timestep of 2~fs was used in Langevin dynamics with zero friction term and collision frequency of 1.0~ps$^{-1}$. 
-Covalent bonds with hydrogens were constrained with the LINCS algorithm~\cite{hess97,hess07}.
-Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with a real space cut-off at 1.0~nm was used 
-for electrostatics. Plain cut-off with 1~nm was used for Lennard-Jones interactions. 
-The neighbour lists with a cut-off of 1.0~nm were updated every 5 steps. 
-Pressure was semi-isotropically coupled to a pressure of 1~bar with the Berendsen method~\cite{berendsen84}.
-
-It should be noted that the area per molecule with these settings for the GAFFlipid model was 61.6~\AA$^2$,
-while the original publication reported 63.9~\AA$^2$~\cite{dickson12}. However, the same parameters and Amber to Gromacs
-conversion reproduced the area per molecule from original publication for the lipid14 model (see next section).
-
-\subsubsection{Lipid14}
-The initial structure was taken directly from the Lipidbook~\cite{domanski10}.
-The Amber compatible force field parameters were generated using the tleap program which is integrated in the AmberTools14 package~\cite{ferrer13}. 
-A workflow similar to the one used previously for the conversion and validation of the GAFFLipid parameters was followed here. 
-As before, a negligible energy difference of 0.003 kcal/mol was obtained between the two programs.
-
-Timestep of 2~fs was used in Langevin dynamics with zero friction term and collision frequency of 1.0~ps$^{-1}$. 
-Covalent bonds with hydrogens were constrained with LINCS algorithm~\cite{hess97,hess07}.
-Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with real space cut-off of 1.0~nm was used 
-for electrostatics. Plain cut-off with 1~nm was used for Lennard-Jones interactions. Dispersion correction
-was applied for both energy and pressure. The neighbor lists with a cut-off of 1.0~nm were updated every 5 steps. 
-Pressure was semi-isotropically coupled to a pressure of 1~bar with the Berendsen method~\cite{berendsen84}.
-
-The area per molecule with these settings was 65.4~\AA$^2$ which is in agreement with the value reported in the original publication 65.6$\pm$0.5~\AA$^2$~\cite{dickson14}.
-
-\subsubsection{Poger et al.}
-The Poger lipids are derived from GROMOS G53A6~\cite{poger10} and were initially coined 53A6-L (L for lipids). They are now part of GROMOS G54A7~\cite{poger12} 
-and parametrized to work with the SPC water model~\cite{berendsen81}. The initial hydrated bilayer structure of 128 DPPC and 5841 water molecules as well as force field parameters were downloaded 
-from David Poger's web site (\url{http://compbio.chemistry.uq.edu.au/~david/}) on April 2012. 
-We noticed that the same files downloaded in October 2013 appear to lack two dihedral angles in the choline headgroup (only one dihedral of type gd\_29 allowing 
-the rotation of the 3 choline methyls) compared to the April 2012 version (3 dihedrals of type gd\_29 for the 3 choline methyls). This should not affect the 
-bilayer structure and only change the kinetics of the choline methyls rotation. However the October 2013 version has not been tested in this study.
-
-MD Simulations (two repetitions with independent initial velocities) were run for 100~ns using a 2~fs time step and the analysis 
-was performed on the last 50~ns. Coordinates were saved every 50~ps for analysis. All bond lengths were constrained with the LINCS algorithm~\cite{hess97,hess07}. Temperature was kept 
-at 323~K employing the velocity-rescale~\cite{bussi07} thermostat with a time constant of 0.1~ps (DPPC and water coupled separetly). Pressure was maintained semi-isotropically at 1~bar using 
-the Parrinello--Rahman barostat~\cite{parrinello81} using a 4~ps time constant and a compressibility of 4.5e-5~bar$^{-1}$. For non-bonded interactions, two conditions were tested:
-i) A 0.8--1.4 nm twin-range cut-off with the neighbor list updated every 5 steps for both electrostatics and Lennard-Jones (LJ) interactions 
-(simulation files available at~\cite{pogerFILESrf1,pogerFILESrf2}). For the former the generalized reaction 
-field (RF) with a dielectric permittivity of 62 was used beyond the 1.4~nm cut-off~\cite{tironi95}. This is the original setup that Poger et al.~\cite{poger10} used.
-ii) PME~\cite{darden93,essman95} electrostatics with a real space cut-off of 1.0~nm, a Fourier spacing of 0.12~nm and an interpolation order of 4, LJ interactions computed with a 1.0--1.4~nm twin-range cut-off,
-neighbor list updated every 5 steps (simulation files available at~\cite{pogerFILESpme1,pogerFILESpme2}). 
-Note that Poger and Mark tested the effect of PME vs RF in ref.~\cite{poger12}, but used a 1.0~nm cut-off with PME and 1.4~nm with RF for LJ
-interactions. Since 0.8--1.4~nm twin-range cut-off for LJ interations is used in the parametrization of GROMOS force field, we decided to use that
-also in the simulations with PME.
-
-Since Poger lipids come from the GROMOS force field, it is important to note that GROMOS uses the RF scheme for computing electrostatics (this is the method used for the 
-force field parameterization). Using setup i) based on RF, we were able to reproduce the results (i.e. area per lipid value of 0.63~nm$^2$) from the original work only with 
-GROMACS versions 4.0.X and earlier (the original authors~\cite{poger10} used GROMACS version 3.3.3). When switching to versions 4.5.X and above, the area per lipid dropped to below 0.58~nm$^2$. 
-The GROMACS developers were contacted and a redmine issue opened (\url{http://redmine.gromacs.org/issues/1400}). The difference comes from the new Trotter decomposition 
-introduced in versions 4.5.X. A fix has been introduced in version 4.6.6 that allows a recovery of an area per lipid value of 0.615~nm$^2$. The results in terms of area per lipid using the different 
-GROMACS versions are available at~\cite{pogerFILESrf2}.
-Thus we decided to use only the PME setup ii) for computing the order parameter since it gives stable results regardless of the GROMACS version. We obtained an area per 
-lipid of 0.615~nm$^2$, below 0.648 nm$^2$ found by the original authors with their PME setup (see~\cite{poger12}). We explained that by the fact that we used 
-a 1.4~nm for the LJ cut-off whereas a value of 1.0~nm was used in the original publication.
-
-\subsubsection{Slipids}
-Initial coordinates for a hydrated DPPC (at 323~K) and POPC (at 310~K) bilayers (30 and 40 waters/lipid, respectively) were taken directly from the 
-Slipids home page \url{http://mmkluster.fos.su.se/slipids/Downloads.html}.  The Slipids force field~\cite{jambeck12,jambeck12b} was used for the the all atom descriptions of DPPC and POPC, and
-water was described with the TIP3P water model~\cite{jorgensen83}. Simulations were performed within the NPT ensemble using the GROMACS 4.6.X simulation
-package~\cite{hess08}. The Nos\'{e}--Hoover thermostat~\cite{nose84,hoover85} was used with reference temperatures of
-323~K (DPPC) and 310~K (POPC) and a relaxation time constant of 0.5~ps. Water and lipids were coupled separately to
-the heat bath. Pressure was kept constant at 1.013~bar using a semi--isotropic Parrinello--Rahman
-barostat~\cite{parrinello81} with a time constant of 10.0~ps. Equations of motion were
-integrated with the leapfrog algorithm using a timestep of 2~fs. Long range
-electrostatic interactions were calculated using the PME method~\cite{darden93,essman95}, with a fourth order
-smoothing spline. A real space cut-off of 1.0~nm was employed with grid spacing of 0.12~nm in the reciprocal space.
-Lennard-Jones potentials were cut off at 1.4~nm, with a dispersion correction applied to both energy and pressure.
-All covalent bonds in lipids were constrained using the LINCS algorithm~\cite{hess97},
-whereas water molecules were constrained using SETTLE~\cite{miyamoto92}. Twin-range cutoffs,
-1.0~nm and 1.6~nm, were used for the neighbor lists with the longrange neighbor list updated every
-10 steps. This simulation protocol corresponds to the protocol used in Ref~\cite{jambeck13}.
-
-\subsubsection{Kukol}
-A bilayer patch with 512 POPC lipids was constructed and hydrated with $\sim$40 SPC water molecules per lipid. 
-The force field parameters were obtained from Lipidbook \cite{domanski10}.
-This bilayer was simulated with a 2~fs time step for a total of 50~ns and coordinates were saved every 100~ps. 
-All bonds were constrained with LINCS~\cite{hess97,hess07}. PME~\cite{darden93,essman95} was employed for the long-range electrostatics. Lennard-Jones interactions 
-were cut off at 1.4~nm. A neighbour list with a radius of 0.8~nm was updated every 5~steps. The constant temperature of 298~K 
-was maintained with the Berendsen thermostat \cite{berendsen84} with a time constant of 0.1~ps. The Berendsen barostat \cite{berendsen84} 
-was employed for semi-isotropical pressure coupling at 1~bar.
-
-\subsubsection{Chiu et al.}
-The force field parameters and the initial configuration were available through the Lipidbook~\cite{domanski10}.
-Timestep of 2~fs was used with leapfrog integrator. Covalent bond lengths were constrained with LINCS algorithm~\cite{hess97,hess07}. 
-Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with real space cut-off of 1.0~nm was used 
-for electrostatics. Twin range cut-off was used for the Lennard-Jones interactions with short and long cut-offs of 1.0~nm and 1.6~nm, respectively.
-The neighbour lists with a cut-off of 1.0~nm were updated every 5 steps. Temperature was coupled separately
-for lipids and water to 298~K with the velocity-rescale method~\cite{bussi07} with a coupling constant 0.2~ps.
-Pressure was semi-isotropically coupled to the atmospheric pressure with the Parrinello--Rahman method~\cite{parrinello81}.
-
-\subsubsection{Ulmschneider}
-The initial structure containing 128 POPC molecules with 3328 TIP3P water~\cite{jorgensen83} molecules (26 per lipid) was downloaded from Lipidbook \cite{domanski10} 
-together with the topologies. This bilayer was simulated for 100~ns with a time step of 2~fs and the data was saved every 10~ps. The bonds involving hydrogen atoms were 
-constrained with LINCS~\cite{hess97,hess07}. The temperature was kept at 298~K with the Berendsen thermostat~\cite{berendsen84}. The pressure was semi-isotropically coupled to the Berendsen 
-barostat~\cite{berendsen84} with a time constant of 1~ps and a target pressure of 1~bar. PME~\cite{darden93,essman95} was employed for long range electrostatics and a cut-off of 1~ns was employed for 
-the Lennard-Jones interactions. A neighbour list with a radius of 1~nm was updated every 10~steps.
-
-Additionally, the simulations were repeated with the dispersion correction applied to pressure and temperature. Even though the area per lipid decrease
-d slightly, the headgroup order parameters were only slightly affected.
-
-\subsubsection{Tj\"ornhammar et al.}
-The gel phase DPPC bilayer structure delivered by Tj\"ornhammar  and Edholm~\cite{tjornhammar14} was ran for 5~ns at 343~K in order to destroy the 
-ordered gel configuration. This was followed by a 200~ns simulation at 323~K, i.e. in the fluid phase. The last 100~ns of this simulation was used for analysis. 
-The same mdp file as in the Supplementary Information section of the original paper ~\cite{tjornhammar14} was used except for the simulation temperature.
-
-\subsubsection{Lee-CHARMM36-UA}
-A hydrated bilayer consisting of 72 DLPC lipids and 2189 water molecules
- is constructed using the model by Lee et al. ~\cite{lee14}. This model describes 
- the important all-atom CHARMM36 character of the lipid headgroup but 
-reduces the details of the lipid chains into a united-atom model.
-The initial  equilibrated structure was downloaded from the web page of J. Klauda, 
-Department of Chemical and Biomolecular Engineering University of Maryland.
- The parameter files are taken from the Supplementary Material of Ref. \citenum{lee14}.  
-This bilayer was equilibrated for 20~ns and the production run was 50~ns long,  with data was saved every 20~ps. 
- The equations of motion were integrated using the multiple time step Verlet r-RESPA algorithm with
-a time step of 2 fs, and a calculation of the electrostatic forces only every two timesteps.
- Covalent bonds between heavy and hydrogen atoms were constrained 
-using SHAKE/RATTLE algorithm~\cite{miyamoto92}.
-
-The temperature was kept at 323~K with a Langevin thermostat
-with a damping coefficient of 5~ps.  The modified NAMD version of the
-Nose-Hoover barostat with Langevin dynamics (piston period of
-0.1 ps and piston decay time of 0.05 ps)  was used semi-isotropically 
-to reach the averaged target pressure of 1~bar and an averaged zero surface tension.
-PME~\cite{darden93,essman95} was employed for long range electrostatics. 
-A cut-off of 1.2~nm was employed for  the Lennard-Jones interactions, with 
-a force-based switching function  for distances beyond 1~nm.
-A neighbour list with a radius of 1.4~nm was updated every 10~timesteps.
-
-NAMD was developed by the Theoretical and Computational Biophysics Group in the
- Beckman Institute for Advanced Science and Technology at the University
- of Illinois at Urbana-Champaign\cite{phillips05}.
-
-\subsubsection{Botan-CHARMM36-UA}
-This model is similar to the Lee-CHARMM36-UA model with a few  differences.
-A hydrated bilayer consisting of 128 DLPC lipids and 3840 water
-molecules is described by a model derived from the CHARMM27-UA model by H\'enin et al.~\cite{henin08}.
-
-The distribution into all-atom  (AA) and united-atom (UA) parts within the lipids is the same
-as in the original CHARMM27-UA model by H\'enin et al.~\cite{henin08}.  
-This distribution differs from the one by Lee-CHARMM36-UA model solely
-for the first methylene groups of the two acyl chains (C22 and C32 in CHARMM36 topology): 
-their hydrogens are merged into  united atoms, whereas in Lee's model their hydrogens are described explicitely.
-The  united-atom Berger model ~\cite{berger97} was used for the tails and the UA-AA interactions as in Ref. ~\citenum{henin08}. 
-The difference between this model and the original one by H\'enin~\cite{henin08} is the replacement of the 
-AA parameters for the heads by  the parameters of the all-atom
- CHARMM36 force-field~\cite{klauda10}. Contrarily to the model by Lee et al.~\cite{lee14}, no reparametrization was done.
-
-The non-bonded interactions are calculated using an atom-based switching function with short and long cut-offs of 0.8 and 1.2~nm. 
-Long range electrostatic interactions are implemented using the particle-particle particle-mesh solver with a relative accuracy of $10^{-4}$. The system 
-is first equilibrated for 30~ns in the NP$\gamma$T ensemble (Nos\'{e}--Hoover~\cite{nose84,hoover85} style thermostat and barostat with anisotropic pressure coupling) 
-at 323~K and 1~bar with timestep of 1~fs. The following 20~ns of dynamics are taken for calculation of configurational averages. 
-Simulations were carried out by using the LAMMPS package~\cite{plimpton95}. %( http://doi.org/10.1006/jcph.1995.1039 )., the input files are available ( http://dx.doi.org/10.5281/zenodo.13821 ).
-
-\subsection{Dihedral angle distributions as a function of cholesterol in CHARMM36}
-\begin{figure*}[]
-  \centering
-  \includegraphics[width=17.2cm]{../Fig/dihsCHOLcharmm.pdf}
-  \caption{\label{dihsCHOLcharmm}
-    The effect of cholesterol content on the POPC glycerol backbone and choline dihedral angles in CHARMM36 model (T=303~K).}
-\end{figure*}
-
-\subsection{Author Contributions}
-{\it Alexandru Botan} provided simulation results for charmm36-UA\\
-{\it Fernando Favela} prepared, performed and analyzed simulations
-with popc and cholesterol for the CHARMM36 All-Atom FF. \\
-{\it Patrick F. J. Fuchs} Ran and analyzed the Poger simulations. Provided scientific information which significantly advanced the project (signs and forking of order parameters). \\
-{\it Matti Javanainen} prepared and performed simulations with multiple lipid models and analyzed the results. Supervised the work of JT.\\
-{\it Matej Kanduc} provided simulation results for the Berger DLPC model \\
-{\it Waldemar Kulig} prepared the MD simulations with cholesterol for the MacRog FF \\
-{\it Antti Lamberg} Prepared and performed simulations with Berger and its variants to show the importance of the signs and stereospecific labeling of the order parameters. \\
-{\it Claire Loison} provided simulation results for charmm36-UA. \\
-{\it Alexander Lyubartsev} provided simulation results for Högberg08 force field.\\
-{\it Markus S. Miettinen} Co-designed the project with OHSO and supported in the work management. Provided trajectories for Berger DMPC; prepared parameter files for 72 lipids \\
-{\it Luca Monticelli} Critical discussions in all phases of the project. Collaboration with Jukka  M{\"a}{\"a}tt{\"a}. \\
-{\it Jukka M{\"a}{\"a}tt{\"a}}  prepared and performed simulations with Berger and Slipids models and analyzed the results.\\
-{\it O. H. Samuli Ollila}Co-designed the project with MSM and managed the work. Ran and analyzed several simulations. Wrote the manuscript. \\
-{\it Marius Retegan} Prepared and validated the GROMACS compatible parameter files for GAFFlipid and Lipid14 force fields.\\
-{\it Tomasz Rog} Provided the topologies for the MacRog model and also the full hydration simulation data with this model. \\
-{\it Hubert Santuz} prepared and performed the cholesterol simulations with CHARMM36 and analyzed the results. 
-Made significant contribution in the data management of the project.\\
-{\it Joona Tynkkynen} prepared and performed the dehydration simulations with the MacRog FF.\\
-
-%\newpage
-
 \bibliography{refs}

### 2015-05-15 HGmodelMANUSCRIPT/HGmodel_ACSsupplementary.tex
Supplementary -> Supporting

@@
 \title[An \textsf{achemso} demo]
-  {SUPPLEMENTARY INFORMATION: Towards atomistic resolution structure of phosphatidylcholine headgroup and glycerol backbone at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}}
+  {SUPPORTING INFORMATION: Towards atomistic resolution structure of phosphatidylcholine headgroup and glycerol backbone at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}}
 

### 2015-05-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Supplementary -> Supporting

@@
 the acceptance of the offer was based on authors' self-assesment of their scientific contribution.
-These contributions are summarized in the Supplementary Information. %(the location depends on the journal probably).
+These contributions are summarized in the Supporting Information. %(the location depends on the journal probably).
 
@@
 are listed in Tables~\ref{systems},~\ref{systemsDEHYD} and~\ref{systemsCHOL}. Full simulation
-details are given in the Supplementary Information (SI) or in the original publications in case the
+details are given in the Supporting Information (SI) or in the original publications in case the
 data is used previously therein. Additionally, the files related to the simulations and the resulting trajectories are publicly
@@
      the original publication is cited if simulation data from previously published work has been directly used,
-     for other systems the simulation details are given in the Supplementary Information.
+     for other systems the simulation details are given in the Supporting Information.
 $^h$ Magnitudes from Fig.~S4 of Klauda et al.~\cite{klauda10}, signs matched to our simulations.
@@
 we calculated from CHARMM36 the glycerol backbone dihedral angle distributions
-at various cholesterol contents (Supplementary material): The only detectable change
+at various cholesterol contents (Supporting material): The only detectable change
 is the small decrease of gauche(-) and increase of gaughe(+) probability of the g$_3$-g$_2$-g$_1$-O(\textit{sn}-1) dihedral.

### 2015-05-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex

--- 2015-05-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
@@
 However, only $^1$H-$^{13}$C NMR techniques allow also the measurement of the sign of the order parameter~\cite{hong95a,hong95b,gross97}. 
-The measured sign is negative for almost all the carbons discussed in this work, except for $\alpha$ which is positive~\cite{hong95a,hong95b,gross97}. 
-For more detailed discussion about the determination of the sign of the order parameters see Ref.~\citenum{signpost}.
-
-For most CH$_2$ segments in fluid phosholipid bilayer the order parameters are equal for both hydrogens attached to the same carbon.
-However, in some cases (e.g. g$_1$, g$_3$ and  C$_2$ carbon in the \textit{sn}-2 chain) the order parameters are not equal and this 
-can be observed with both $^2$H NMR and $^1$H-$^{13}$C NMR techniques. In the present work we call the phenomena of unequal order parameters 
-for hydrogens attached to the same carbon as {\it forking} to avoid confusion with dipolar and quadrupolar splitting in NMR terminology. 
+The measured sign is negative for almost all the carbons discussed in this work, only the choline $\alpha$ is positive~\cite{hong95a,hong95b,gross97}. 
+For more detailed discussion about the sign measurements of the order parameters, see Ref.~\citenum{signpost}.
+
+For most CH$_2$ segments in a fluid phospholipid bilayer, the order parameters of both hydrogens are equal.
+However, in some cases (e.g., g$_1$, g$_3$, and  C$_2$ carbon in the \textit{sn}-2 chain) the two order parameters are not equal;
+this  can be observed with both $^2$H NMR and $^1$H-$^{13}$C NMR techniques. In the present work,
+to avoid confusion with the dipolar and quadrupolar splittings in NMR terminology,
+we call the phenomenon of unequal order parameters for hydrogens attached to the same carbon {\it forking}. 
 Forking has been studied in detail with $^2$H NMR techniques by deuterating the R or S position in CH$_2$ segment, and
-the studies show that the forking arises from differently sampled orientations of the two C--H bonds, not from two 
+the studies show that it arises from differently sampled orientations of the two C--H bonds, not from two 
 separate populations of lipid conformations~\cite{engel81,gally81}.
@@
 For more details, see Ref.~\citenum{accuracyPOST}. 
-Motivated by the high experimental reproducibility, we have highlighted in 
+Motivated by the high experimental repeatability, we have highlighted in 
 Fig.~\ref{HGorderparameters} the subjective sweet spots (light blue areas), within which we expect the calculated absolute 
@@
 between simulations and experiments in different conditions can be used to assess the quality of the force field 
-in different situations, and, if the quality is good, to interpret the structural changes in experiments.
+in different situations, and, if the quality is good, to potentially interpret the structural changes in experiments.
 Here we exemplify such comparison for a lipid bilayer under low hydration levels and when varying amounts of cholesterol is included in the bilayer. 
@@
   \includegraphics[width=7.3cm]{../DATAreportediINblog/dehydration_lp.pdf}
-%    \includegraphics[width=7.3cm]{../DATAreportediINblog/dehydration_smooth.pdf}
+    \includegraphics[width=7.3cm]{../DATAreportediINblog/dehydration_smooth.pdf}
     \caption{\label{ordPhydr}
@@
   }
-%  	\todo{Markus prepared two versions of this plot. The one on left has all the data points joint by lines;
-%	the one on right shows smooth trend lines (cubic polynomials) to ease qualitative comparison.
-%	Samuli likes the one on left, Markus the one on right. Do others have opinions?}
+  	\todo{Markus prepared two versions of this plot. The one on left has all the data points joint by lines;
+	the one on right shows smooth trend lines (cubic polynomials) to ease qualitative comparison.
+	Samuli likes the one on left, Markus the one on right. Do others have opinions?}
 \end{figure}
@@
   \includegraphics[width=7.3cm]{../DATAreportediINblog/cholesterolization_lp.pdf}
-%    \includegraphics[width=7.3cm]{../DATAreportediINblog/cholesterolization_smooth.pdf}
+    \includegraphics[width=7.3cm]{../DATAreportediINblog/cholesterolization_smooth.pdf}
   \caption{\label{ordPchol}
@@
 }
-%	\todo{Markus prepared two versions of this plot. The one on left has all the data points joint by lines;
-%	the one on right shows smooth trend lines (cubic polynomials) to ease qualitative comparison.
-%	Samuli likes the one on left, Markus the one on right. Do others have opinions?}
+	\todo{Markus prepared two versions of this plot. The one on left has all the data points joint by lines;
+	the one on right shows smooth trend lines (cubic polynomials) to ease qualitative comparison.
+	Samuli likes the one on left, Markus the one on right. Do others have opinions?}
 \end{figure}
@@
 {\it Alexander Lyubartsev} provided simulation results for Högberg08 force field.\\
-{\it Markus S. Miettinen} Co-designed the project with OHSO and supported in the work management. Provided trajectories for Berger DMPC; prepared parameter files for 72 lipids \\
+{\it Markus S. Miettinen} \todo{Please write a text summarizing your contribution} \\
 {\it Luca Monticelli} Critical discussions in all phases of the project. Collaboration with Jukka  M{\"a}{\"a}tt{\"a}. \\
 {\it Jukka M{\"a}{\"a}tt{\"a}}  prepared and performed simulations with Berger and Slipids models and analyzed the results.\\
-{\it O. H. Samuli Ollila}Co-designed the project with MSM and managed the work. Ran and analyzed several simulations. Wrote the manuscript. \\
+{\it O. H. Samuli Ollila} Designed and managed the work. Ran and analyzed several simulations. Wrote the manuscript.  \\
 {\it Marius Retegan} Prepared and validated the GROMACS compatible parameter files for GAFFlipid and Lipid14 force fields.\\

### 2015-05-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Cleaning comments from manuscript tex

--- 2015-05-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
@@
 and (2) cholesterol induces only minor changes to the PC (glycerol backbone) structure.
-%None of the current models is not accurately enough to resolve the structure with experimental accuracy.
-%However, closer inspection of three best performing models (CHARMM36, GAFFlipid and MacRog) suggest
-%that improvements in the sampled dihedral angle distributions would potentilly lead to the model which
-%would resolve the structure. Despite of the inaccuracy in the fully hydrated
-%structures, the response to dehydration, i.e. the tilting of the P--N vector more parallel to membrane normal, 
-%is qualitatively correct in all models. The CHARMM36 and MacRog models describe the interactions
-%between lipids and cholesterol better than the Berger/H\"oltje model.
 This work has been done as a fully open collaboration, using \url{nmrlipids.blogspot.fi} as a
@@
 attempts to derive lipid structures from experimental data~\cite{seelig77b,skarjune79,jacobs80,davis83,akutsu91,hong95b,semchyschyn04}.
-Notably, the structures sampled in a simulation that reproduces these %and other experimental
+Notably, the structures sampled in a simulation that reproduces these 
 parameters will automatically comprise an interpretation of the experiments. In other words, such simulations can be considered as an accurate atomistic resolution description of
@@
 structures between the most often used (Berger-based) lipid model~\cite{berger97} 
-%\todo{If there are no objections, we will leave the naming convention of different Berger versions as it is now.}
 and the best performing models, to demonstrate that by using the 
@@
 \subsection{Open collaboration}
-%\todo{Should we write more about this?}
 
@@
 the acceptance of the offer was based on authors' self-assesment of their scientific contribution.
-These contributions are summarized in the Supporting Information. %(the location depends on the journal probably).
+These contributions are summarized in the Supporting Information.
 
@@
 The order parameters from simulations were calculated directly using the definition as in Eq.~\ref{orderP}.
-%$S_{{\rm CH}}=\frac{3}{2} \langle \cos^2 \theta-1 \rangle$. 
 For the united atom models the hydrogen positions were generated 
@@
   1,2-dimyristoyl-sn-glycero-3-phosphocholine (DMPC),
-  %1,2-dioleoyl-sn-glycero-3-phosphocholine (DOPC),
   dilauroylphosphatidylcholine (DLPC),
@@
 CHARMM36}~\cite{klauda10} & POPC & 128 & 5120 & 303 & 200 & 100 & [\citenum{charmm36files}]$^*$      & SI \\
-%CHARMM36\cite{klauda10}\todoi{Not included in the Fig.
-2.
-Should we remove or include?}        & DOPC   & 128 &  5120    & 303 & 150 & 100  & ?\todoi{Permanent link in progress by Hubert Santuz}  & SI \\
-%\todoi{Permanent link for DOPC CHARMM36 simulation by Hubert Santuz is not required now. However, it would be probably useful in the future.}
 MacRog~\cite{kulig15b}  & POPC & 288 & 12600 & 310 & 100 &  80 & [\citenum{macrogFILES}]$^*$      & SI  \\
@@
 \newline
-%\todo{There is also the interactive version by Hubert Santuz now in https://plot.ly/~HubertSantuz/72/lipid-force-field-comparison/
-%we should figure out which is the most practical way to put that behind permalink once it finalized (Zenodo, figshare or something else?) and then put a citation in the paper.}
   \caption{\label{HGorderparameters}
@@
 values of the order parameters of a well-performing force field to fall.
-
-%In addition to the phosphatidylcholine lipids, similar values of the glycerol backbone order parameters have been measured
-%for phophatidylethanolamine (PE) and phopshatidylglycerol (PG) in E. Coli extract~\cite{gally81},
-%indicating that the glycerol backbone structure is similar independent of the headgroup chemistry and lipid environment.
-%Further, choline order parameters measured from mouse fibroblast L-M cell are similar to the ones in model
-%membranes~\cite{scherer87}.
 
@@
 
-%Interestingly, the probability of the gaughe conformations correlates with the order parameter difference between the $\beta$ and $\alpha$ segments:
-%the larger the gaughe fraction the larger the order parameter difference. This suggestion together with the results in 
-%Fig.~\ref{HGorderparameters2} would indicate that the correct gaughe-trans fraction for N-$\beta$-$\alpha$-O dihedral is 
-%larger than in GAFFlipid but smaller than in CHARMM36. This happens to be quite close to the gaughe-trans fraction in
-%the Berger model. However, there is significant forking and numerical values are off from the experiments in the Berger
-%model suggesting that it has some other inaccuracies.
 \begin{figure}[]
@@
   }
-%  	\todo{Markus prepared two versions of this plot. The one on left has all the data points joint by lines;
-%	the one on right shows smooth trend lines (cubic polynomials) to ease qualitative comparison.
-%	Samuli likes the one on left, Markus the one on right. Do others have opinions?}
 \end{figure}
@@
 their entropy and dynamics (``superlattice model'')~\cite{somerharju09}. 
-%\todo{It has been suggested that we should use direct quotations from original papers here for clarity.}
-%Both of these suggestions 
 Molecular dynamics simulations have supported both
@@
 }
-%	\todo{Markus prepared two versions of this plot. The one on left has all the data points joint by lines;
-%	the one on right shows smooth trend lines (cubic polynomials) to ease qualitative comparison.
-%	Samuli likes the one on left, Markus the one on right. Do others have opinions?}
 \end{figure}
@@
 
-%None of the tested models (13 different models) produces the order parameters with the experimental
-%accuracy for a fully hydrated phoshatidylcholine lipid bilayer. However, the CHARMM36, GAFFlipid and MacRog 
-%models are relatively close. The structures of these models together with the most used lipid model (Berger based) 
-%were subjected to more careful studies. The results revealed that the current models are not accurate
-%enough to resolve the atomistic resolution structures sampled by glycerol backbone and choline headgroup.  
-%However, the correlation between dihedral angle distributions and order parameter differences was found, 
-%suggesting that careful adjustment of dihedral potentials would potentially lead to the model with a correct
-%structure.
-
 (3) The review of previous experimental results revealed that when a bilayer is dehydrated the 
@@
 
-%Independent of the accuracy for a fully hydrated lipid bilayer, all the models reproduced the choline response
-%to dehydration. This can be explained by the change in the P--N vector tilting more parallel to the membrane,
-%which leads to the increase of order parameters despite of the initial configuration. It should be noted, however,
-%that the correct qualitative response does not necessarily indicate correct energetics.
-
 (4) Only modest changes of glycerol backbone and choline order parameters are observed experimentally with increasing cholesterol content.
@@
 dihedral of the glycerol backbone, in other words, there is no major conformational change of the lipid.
-
-%- The response of glycerol backbone and choline headgroup to the cholesterol content is described more
-%realistically in CHARMM36 and MacRog models than in the Berger based model.
 
@@
 platform. All the scientific contributions have been communicated publicly through this blog~\cite{nmrlipids}.
-
-%This work has been, and continues to be, progressed and discussed through a blog at \url{nmrlipids.blogspot.fi}. 
-%Everyone is invited to join the discussion and make contributions through the blog. 
-%The manuscript will be eventually submitted to an appropriate scientific journal. 
-%Everyone who has contributed to the work through the blog will be offered 
-%coauthorship. For more details, see \url{nmrlipids.blogspot.fi}.
-
-%\listoftodos
 

### 2015-05-15 HGmodelMANUSCRIPT/coverLETTER.tex
Update and shorten cover letter

@@
 
-In addition to its main message, the manuscript also provides an extensive comparison on the quality of the head group 
-structures of numerous widely used lipid models in varying biologically relevant conditions. The behavior
-of the lipid head group region plays a key role in multiple processes, such as partitioning of drugs or 
-nanoparticles into membranes and interactions of peripheral proteins with membranes, both of which are 
-currently intensively studied. Therefore the manuscript provides crucial information for a large group
-of computational scientists who are struggling to select the best possible lipid model for their research.
+In addition to its main message, the manuscript also provides an extensive comparison on the structural quality 
+of the numerous widely used computational lipid models. Therefore the manuscript provides crucial information 
+for a large group scientists who are producing or reading the literature involving computational lipid models.
 
@@
 all the simulation raw data in a format that allows easy reanalysis and therefore reuse of the data for other purposes. 
-In addition to the scientic content, we believe that the success of this fresh approach in the field of chemistry 
+In addition to the scientific content, we believe that the success of this fresh approach in the field of chemistry 
 is highly interesting to the audience of JACS.

### 2015-05-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Universitat->Universität
Include pull request https://github.com/NMRLipids/nmrlipids.blogspot.fi/pull/53 by hand due to the incompetence in GitHub usage

@@
 The measured sign is negative for almost all the carbons discussed in this work, except for $\alpha$ which is positive~\cite{hong95a,hong95b,gross97}. 
-For more detailed discussion about the determination of the sign of the order parameters see Ref.~\citenum{signpost}.
+For more detailed discussion about the determination of the sign of the order parameters, see Ref.~\citenum{signpost}.
 
-For most CH$_2$ segments in fluid phosholipid bilayer the order parameters are equal for both hydrogens attached to the same carbon.
-However, in some cases (e.g. g$_1$, g$_3$ and  C$_2$ carbon in the \textit{sn}-2 chain) the order parameters are not equal and this 
-can be observed with both $^2$H NMR and $^1$H-$^{13}$C NMR techniques. In the present work we call the phenomena of unequal order parameters 
-for hydrogens attached to the same carbon as {\it forking} to avoid confusion with dipolar and quadrupolar splitting in NMR terminology. 
+For most CH$_2$ segments in a fluid phospholipid bilayer, the order parameters of both hydrogens are equal.
+However, in some cases (e.g., g$_1$, g$_3$, and  C$_2$ carbon in the \textit{sn}-2 chain) the two order parameters are not equal;
+this  can be observed with both $^2$H NMR and $^1$H-$^{13}$C NMR techniques. In the present work,
+to avoid confusion with the dipolar and quadrupolar splittings in NMR terminology,
+we call the phenomenon of unequal order parameters for hydrogens attached to the same carbon {\it forking}. 
 Forking has been studied in detail with $^2$H NMR techniques by deuterating the R or S position in CH$_2$ segment, and
-the studies show that the forking arises from differently sampled orientations of the two C--H bonds, not from two 
+the studies show that it arises from differently sampled orientations of the two C--H bonds, not from two 
 separate populations of lipid conformations~\cite{engel81,gally81}.

### 2015-05-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Update to the manuscript version to be submitted

@@
 \usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
-\usepackage[obeyFinal]{easy-todo}
+%\usepackage[obeyFinal]{easy-todo}
 \usepackage{rotating}
@@
 Further information, such as scripts, figures, and manuscript text files, are available through our GitHub 
-repository (\url{https://github.com/NMRlipids}). \todo{I will not assign a doi yet. We can wait for the journal decision.}.
+repository (\url{https://github.com/NMRlipids}).
 

### 2015-05-15 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Update to the manuscript version to be submitted

@@
     and Gross et al.~\cite{gross97}.
-    Note that to elucidate the relative change as a function of hydration level,
-    the simulation results are shifted such that the (smaller) $S_\mathrm{CH}$
-    matches (within $\pm$0.01) the experimental value at full hydration;
-    the shift magnitudes for each of the force fields are listed ($S_\mathrm{CH}$+shift) in the y-label.
+    Note that to elucidate the relative change as a function of hydration level, the simulation results were vertically shifted; 
+    the shift magnitudes for each of the force fields are listed (SCH+shift) in the y-label.
   }
@@
     and Gross et al.~\cite{gross97}.  Most order parameters from Berger-POPC-07/H\"oltje-CHOL-13 model for g$_1$ are beoynd the y-axis scale.
-    In order to elucidate the relative change as a function of cholesterol content,
-    the simulation results are shifted such that the (smaller) $S_\mathrm{CH}$
-    matches (within $\pm$0.01) the experimental value without cholesterol;
-    the shift magnitudes for each of the force fields are listed ($S_\mathrm{CH}$+shift) in the y-label.
+    In order to elucidate the relative change as a function of cholesterol content, the simulation results were vertically shifted; 
+    the shift magnitudes for each of the force fields are listed (SCH+shift) in the y-label.
 }

### 2015-05-15 HGmodelMANUSCRIPT/coverLETTER.tex
Final version of cover letter

@@
 existing experimental Nuclear Magnetic Resonance (NMR) data is reviewed and combined with an exceptionally comprehensive 
-collection of generated molecular dynamics (MD) simulation data in order to understand the atomistic resolution structure 
+collection of molecular dynamics (MD) simulation data in order to understand the atomistic resolution structure 
 of biologically abundant phospholipid molecules and their assemblies.
@@
 We believe that the manuscript is interesting to the wide and diverse contemporary audience of
-JACS as it demonstrates how the combination of experimental NMR data with MD simulations is shown 
-to be successful in resolving the atomistic resolution structure of biomolecules in various biologically 
-relevant conditions. The approach is demonstrated for lipids, however, the extension to, e.g. membrane 
+JACS as it demonstrates how the combination of experimental NMR data with MD simulations can 
+be successful in resolving the atomistic resolution structure of biomolecules in various biologically 
+relevant conditions. The approach is demonstrated for lipids, however, the extension to, e.g., membrane 
 proteins is straightforward. Such an atomistic resolution of biomolecular structure significantly advances 
@@
 In addition to its main message, the manuscript also provides an extensive comparison on the structural quality 
-of the numerous widely used computational lipid models. Therefore the manuscript provides crucial information 
-for a large group scientists who are producing or reading the literature involving computational lipid models.
+of the numerous widely used computational lipid models. This is crucial information 
+for the large group scientists who are producing or reading the literature involving computational lipid models.
 
@@
 enormous amount of simulation work was required for this work. The only practical
-solution to collect this data was the open collaboration concept.
+solution to collect these data was the open collaboration concept.
 As result, we have produced original chemical insight for the atomistic resolution 

### 2015-05-15 HGmodelMANUSCRIPT/coverLETTER.tex
Final version of cover letter

@@
 Please find attached to this message a manuscript entitled `Towards atomistic resolution structure of phosphatidylcholine 
-glycerol backbone and choline headgroup at different ambient conditions' by A. Botan et al., which we would like to submit 
+headgroup and glycerol backbone at different ambient conditions' by A. Botan et al., which we would like to submit 
 for consideration of publication in \textit{Journal of the American Chemical Society} (JACS). In the manuscript a large amount of

### 2015-05-18 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Added 43.75 w/l to MacRog dehydration data.
MacRog 44 w/l to Table 2.

@@
 MacRog\cite{kulig15b}     & POPC   & 50 & 288  & 14400 & 310 & 90 & 40 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
+                               & POPC   & 44 & 288  & 12600 & 310 & 100 & 80 & [\citenum{macrogFILES}]$^*$ & SI \\
                                & POPC   & 25 & 288  & 7200 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
@@
   \centering
-  \includegraphics[width=7.3cm]{../DATAreportediINblog/dehydration_lp.pdf}
-%    \includegraphics[width=7.3cm]{../DATAreportediINblog/dehydration_smooth.pdf}
+  \includegraphics[width=7.3cm]{../DATAreportediINblog/dehydration.pdf}
     \caption{\label{ordPhydr}

### 2015-05-19 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Text edits to Methods: Order parameters from simulations

@@
 \affiliation[Freie Universit\"{a}t Berlin] 
-{Fachbereich Physik, Freie Universitat Berlin, Berlin, Germany}
+{Fachbereich Physik, Freie Universit\"at Berlin, Berlin, Germany}
 \author{Luca Monticelli}
@@
 \subsection{Order parameters from simulations}
-The order parameters from simulations were calculated directly using the definition as in Eq.~\ref{orderP}.
-For the united atom models the hydrogen positions were generated 
-in the trajectories post-simulationally using the positions of the heavy atoms and the known hydrocarbon geometries.
-For the statistical error estimates, the time average of order parameters were first calculated separately
-for each lipid in the system. Then it was assumed that different lipids are statistically independent 
-entities (which should be the case in fluid phase) and the error of the mean for the average over individual 
-lipids in the system was calculated and used as error bar for the order parameters.
+The order parameters from simulations were calculated directly using the definition of Eq.~\ref{orderP}.
+For the united atom models, the hydrogen positions were generated 
+post-simulationally from the positions of the heavy atoms and the known hydrocarbon geometries.
+The statistical error was estimated based on the assumption that different lipids are statistically independent 
+entities (which should be the case in fluid phase):
+The time average of a given order parameter was first calculated separately for each lipid,
+and the standard error of the mean over these time averages then taken as the error bar for this order parameter.
 
-It has been recently pointed out that the sampling of individual dihedral angles might be very
-slow compared to the typical simulation timescales~\cite{vogel12}. On the other hand, another recent
-study shows that the slowest rotational correlation functions of a C--H bond (g$_1$) reaches a plateau (S$_{CH}^2$)
-after 200~ns in the Berger-POPC-07~\cite{ollila07a} model, and that the dynamics of this segment is significantly too slow in simulations
-compared to the experiments~\cite{ferreira15}. In practise, less than 200~ns of simulation data is enough for the order parameter
-calculation due to the average over different lipid molecules. In conclusion, if the sampling with typical simulation times
-is not enough for the convergence of the order parameters, then the simulation models has significantly too slow dynamics.
+It has been pointed out that the sampling of individual dihedral angles might be very
+slow compared to the typical (100~ns) simulation timescales~\cite{vogel12}.
+After 200~ns, however, even the slowest rotational correlation function of a C--H bond  (g$_1$) reaches
+a plateau ($S_\mathrm{CH}^2$) in the Berger-POPC-07 model~\cite{ollila07a}---and, notably,
+the dynamics of this segment have been shown to be significantly slower in simulations
+than in experiments~\cite{ferreira15}.
+In practise,
+due to averaging over different lipid molecules,
+less than 200~ns of simulation data should be enough for the order parameter calculation;
+if the sampling within typical simulation times
+is not enough for the convergence of the order parameters,
+then the simulation model in question has unphysically slow dynamics.
 

### 2015-05-21 HGmodelMANUSCRIPT/coverLETTER.tex
Changes for second submission to the JPCB

@@
 headgroup and glycerol backbone at different ambient conditions' by A. Botan et al., which we would like to submit 
-for consideration of publication in \textit{Journal of the American Chemical Society} (JACS). In the manuscript a large amount of
+for consideration of publication in \textit{Journal of Physical Chemistry} (JPC). In the manuscript a large amount of
 existing experimental Nuclear Magnetic Resonance (NMR) data is reviewed and combined with an exceptionally comprehensive 
@@
 We believe that the manuscript is interesting to the wide and diverse contemporary audience of
-JACS as it demonstrates how the combination of experimental NMR data with MD simulations can 
+JPC as it demonstrates how the combination of experimental NMR data with MD simulations can 
 be successful in resolving the atomistic resolution structure of biomolecules in various biologically 
@@
 In addition to the scientific content, we believe that the success of this fresh approach in the field of chemistry 
-is highly interesting to the audience of JACS.
+is highly interesting to the audience of JPC.
 
@@
 
-We hope that you can consider this manuscript to be published as an article in the Journal of the American Chemical Society. \\
+We hope that you can consider this manuscript to be published as an article in the Journal of Physical Chemistry. \\
 

### 2015-07-19 HGmodelMANUSCRIPT/REPLYtoREVIEWER.txt
Add reply to reviewer's comments

REPLY TO 1: We were trying to resolve the atomistic resolution structure of glycerol backbone 
and choline by using the available atomistic resolution models. In all of the used models the 
most recent version published by the developers was intended to be used in NPT ensemble with 
zero tension. The improvements of the models and testing of influence of changes in parameters, 
including the ensemble, is beyond the scope of the work. Further, it is not very likely that the 
change in the ensemble would significantly change the glycerol backbone or choline structure.

REPLY TO 2: It is known experimentally that the glycerol backbone order parameters are essentially 
the same for PC, PE and PG lipids, as written in the manuscript (page 15): 
"Since the glycerol backbone order parameters are very similar irrespective of the headgroup 
chemistry (PC,PE and PG) or lipid environment,26 it is reasonable to assume that the stereospecifity 
measured for the PE lipids holds also for the PC lipids." 
It is extremely unlikely scenario that the stereospecific labeling would be different while the 
order parameters would be numerically equal. Thus the assumption has quite strong experimental justification.

REPLY TO 3: The quantification used is explained in the figure caption. MAYBE MARKUS SHOULD CLARIFY THIS A LITTLE BIT? 
It is based on the distance between simulated and experimental order parameters. The figure is made 
and color coding is used to make the message in Fig. 2 more easily readable, thus all the details are 
also available for the readers who are willing to take a closer look. If the force fields are ranked based 
on the distance of the results from the experimental values, the order of the force fields in the figure will 
not change. The color coding may change if another researcher prefers some other coding. Since all the data 
is available, anyone is free to do that.

REPLY TO 4: As written in the manuscript there is actually also previous experimental support, in addition 
to the current results, for the dihedral distributions suggested in the manuscript. About g3-g2-g1-O dihedral 
it is written on page 19: 
"In conclusion, models preferring the anti conformation for this dihedral give more 
realistic order parameters and this is in agreement with previous crystal structure and 1H NMR studies.19–21,23–25" 
and about N-b-a-O dihedral it is writte on page 20: 
"This result is again in agreement with previous crystal structure, 1H NMR and Raman spectroscopy studies19–22 which suggest that this dihedral is in the gauche conformation in the absence of ions."

REPLY TO 4 (a): We do not know that the other models have exatcly the correct dihedral distribution 
and we do not claim this in the manuscript. Actually they cannot be all correct since they are not 
the exactly same and also the order parameters are not the same and none of them really satisfy the 
experiments, as discussed in the manuscript. However, as said in the manuscript and in the above reply, 
there is strong evidence that dihedrals in other models are significantly more realistic than in Berger.

The justification for usage of phrase "biologically relevant conditions" in the manuscript is given 
in the sentence on page 4-5: 
"Importantly, the structural parameters for the glycerol backbone are similar for various biologically 
relevant lipid species (phosphatidylcholine (PC), phosphatidylethanolamine (PE) and phosphatidylglycerol (PG)) 
in various environments,26 and the structural parameters for the choline headgroup are similar in model 
membranes and real cells (mouse fibroblast L-M cell).27". 
Since the order parameters for these parts are the same in real cells (at least E. Coli and mouse 
fibroblast L-M cell) as in model systems, most likely the structure is the same as well.

REPLY TO 4 (b): Yes, they can. However, if molecular dynamics simulation model would reproduce all 
the experimental parameters correctly (including order parameters) it would be correct with very high 
probability, and would give a valid interpretation for experiments, as stated several times in the 
manuscript, for example on page 21: 
"If the improved atomistic molecular dynamics simulation model reproduced the order parameters 
and other experimental observables (like chemical shift anisotropy) with experimental accuracy, 
it would give an interpretation for the atomistic resolution structure of the glycerol backbone and choline.10–13,15,16,18".

REPLY TO 4 (c): As stated in the replies above, there is also previous experimental evidence for 
the conclusions made about dihedral distributions and these conclusions can be extended for the 
real cells as well. However, it is true that, in principle one cannot exclude 
other dihedrals than these two without further evidence. However, the previous experiments supports 
the conclusions and the main purpose of this analysis here is to  (from page 17-18) 
"demonstrate that 1) the differences in order parameters indicate significantly different 
structural sampling strongly correlating with the dihedral angles of the related bonds, 
and that 2) the comparison between experimental and simulated order parameters can be used 
to exclude nonrealistic structural samping in molecular dynamics simulations." 
And further from page 19 
"We note that performing a similar comparison through all the dihedrals in all the 13 
models would probably give highly useful information on how to improve the accuracy of the models 
yet this is beyond the scope of the current report."

REPLY TO 5: This is indeed relevant work and we were aware of it. The reference was accidentally 
left out from the manuscript but it WILL BE ADDED. We thank the reviewer for pointing this out.

REPLY TO 6: This is indeed relevant for the simulations with cholesterol since the glycerol tilt 
may also explain the minor changes in the order parameters when adding cholesterol. 
WE SHOULD CALCULATE THIS FROM CHOLESTEROL SIMULATIONS WITH CHARMM SINCE THOSE ARE USED FOR THE 
DIHEDRAL ANALYSIS IN SI. MAYBE HUBERT COULD DO THIS? 
As a function of hydration level unfortunately (from page 24): 
"The response of the glycerol backbone seems to be more subtle than that of the choline 
headgroup; none of the four models reproduced the experimental trends upon dehydration 
with enough accuracy to invite a structural interpretation."

REPLY TO 7: The scope of this manuscript has been limited on purpose on the glycerol backbone 
and choline segments to avoid exhausting the reader, thus this analysis is beyond the scope 
of the work. The acyl chain region will be discussed in the context of the NMRLipids III 
project (see http://nmrlipids.blogspot.fi/p/about.html).

### 2015-08-03 HGmodelMANUSCRIPT/REPLYtoREVIEWER.txt
Add reply to reviewer's comments
Add reference to the work by Prakash et al. (Ref. 43)

@@
 and about N-b-a-O dihedral it is writte on page 20: 
-"This result is again in agreement with previous crystal structure, 1H NMR and Raman spectroscopy studies19–22 which suggest that this dihedral is in the gauche conformation in the absence of ions."
+"This result is again in agreement with previous crystal structure, 1H NMR and Raman spectroscopy studies19–22 
+which suggest that this dihedral is in the gauche conformation in the absence of ions."
 
@@
 "If the improved atomistic molecular dynamics simulation model reproduced the order parameters 
-and other experimental observables (like chemical shift anisotropy) with experimental accuracy, 
-it would give an interpretation for the atomistic resolution structure of the glycerol backbone and choline.10–13,15,16,18".
+and other experimental observables (like 31P chemical shift anisotropy 36 and 31P-13C dipolar couplings 43) 
+with experimental accuracy, it would give an interpretation for the atomistic resolution structure of the 
+glycerol backbone and choline.10–13,15,16,18".
 
@@
 
-REPLY TO 5: This is indeed relevant work and we were aware of it. The reference was accidentally 
-left out from the manuscript but it WILL BE ADDED. We thank the reviewer for pointing this out.
+REPLY TO 5: This is indeed relevant work and we were aware of it. However, the reference was accidentally 
+left out from the manuscript. We thank the reviewer for pointing this out.
+This work is now mentioned in page 5:
+"In addition to the order parameters, dihedral angles for the glycerol
+backbone and headgroup estimated from experiments,28,38-42 31P chemical shift anisotropy 36
+and 31P-13C dipolar couplings 43 have been used to assess the quality of a simulation model."
+where Ref. 43 is the mentioned work. We also mention the work now also in page 21: 
+"...(like 31P chemical shift anisotropy 36 and 31P-13C dipolar couplings 43)..."
 

### 2015-08-03 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Changes for second submission to the JPCB
Add reference to the work by Prakash et al. (Ref. 43)

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[journal=jacsat,manuscript=article]{achemso}
+\documentclass[journal=jpcbfk,manuscript=article]{achemso}
 
@@
 \usepackage{rotating}
+\usepackage{epstopdf}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 and choline headgroups are scattered over many publications and published in a format that is difficult to understand without some NMR expertise. 
-In addition to the order parameters, dihedral angles for the glycerol backbone and headgroup estimated from experiments have sometimes been used to 
-assess the quality of a simulation model~\cite{robinson94,essex94,kothekar96,hyvonen97,shinoda97,duong99}.
+In addition to the order parameters, dihedral angles for the glycerol backbone and headgroup estimated 
+from experiments~\cite{robinson94,essex94,kothekar96,hyvonen97,shinoda97,duong99}, 
+$^{31}$P chemical shift anisotropy~\cite{chowdhary13} and $^{31}$P-$^{13}$C dipolar couplings~\cite{prakash10}
+have been used to assess the quality of a simulation model.
 
@@
 If the improved atomistic
-molecular dynamics simulation model reproduced the order parameters and other experimental observables (like chemical shift anisotropy)
+molecular dynamics simulation model reproduced the order parameters and other experimental observables 
+(like $^{31}$P chemical shift anisotropy~\cite{chowdhary13} and $^{31}$P-$^{13}$C dipolar couplings~\cite{prakash10})
 with experimental accuracy, it would give an interpretation for the atomistic resolution structure of the glycerol backbone and 

### 2015-08-09 HGmodelMANUSCRIPT/REPLYtoREVIEWER.txt
Updating the answer to reviewer's point 1 according to issue https://github.com/NMRLipids/nmrlipids.blogspot.fi/issues/57

@@
 most recent version published by the developers was intended to be used in NPT ensemble with 
-zero tension. The improvements of the models and testing of influence of changes in parameters, 
-including the ensemble, is beyond the scope of the work. Further, it is not very likely that the 
-change in the ensemble would significantly change the glycerol backbone or choline structure.
+zero tension, indicating that the general opinion in the field consider the zero tension ensemble 
+being more realistic. Most notably, the two popular lipid force fields that required the use of either 
+a fixed membrane area or a nonzero surface tension, Charmm27 and Lipid11, were recently updated 
+to Charmm36 and Lipid14, which removed this requirement. The older versions of the force fields
+ment to be ran in non-zero tension are not discussed in the manuscript. Thus, the more detailed discussion
+on this issue is beyond the scope of the work. Also, the improvements of the models and testing 
+of influence of changes in parameters, including the ensemble, is beyond the scope of the work. 
+Further, it is not very likely that the change in the ensemble would significantly change the 
+glycerol backbone or choline structure. We have clarified that we only reproduce the simulations
+with original parametrizations by the sentence: "All simulation parameters were set as close to the original 
+parametrization works as possible." in the page 10.
 

### 2015-08-12 HGmodelMANUSCRIPT/HGmodel_ACSsupplementary.tex
Add glycerol orientation analysis and discussion to the manuscript, related to the reviewer's point 6

@@
 
-\section{Dihedral angle distributions as a function of cholesterol in CHARMM36}
+\section{The effect of cholesterol on glycerol backbone structure}
 \begin{figure*}[]
@@
 \end{figure*}
+
+\begin{figure*}[]
+  \centering
+  \includegraphics[height=17.2cm]{../Fig/GlycerolOrientation.eps}
+  \caption{\label{GlycerolOrientation}
+    Orientation of g$_1$-g$_3$ (a) and g$_1$-g$_2$ (b) respect to membrane normal as a function of cholesterol content
+    calculated from CHARMM36 simulations in 298K and 303K.}
+\end{figure*}
+
+\newpage
 

### 2015-08-12 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Updating the answer to reviewer's point 1 according to issue https://github.com/NMRLipids/nmrlipids.blogspot.fi/issues/57
Add glycerol orientation analysis and discussion to the manuscript, related to the reviewer's point 6

@@
 details are given in the Supporting Information (SI) or in the original publications in case the
-data is used previously therein. Additionally, the files related to the simulations and the resulting trajectories are publicly
+data is used previously therein. All simulation parameters were set as close to the original 
+parametrization works as possible.
+Additionally, the files related to the simulations and the resulting trajectories are publicly
 available for almost all systems in the Zenodo collection \url{https://zenodo.org/collection/user-nmrlipids}. 
@@
 Therefore, to resolve the nature of cholesterol-induced structural changes,
-we calculated from CHARMM36 the glycerol backbone dihedral angle distributions
-at various cholesterol contents (Supporting material): The only detectable change
-is the small decrease of gauche(-) and increase of gaughe(+) probability of the g$_3$-g$_2$-g$_1$-O(\textit{sn}-1) dihedral.
-In conclusion, our results suggest that the significant effects of cholesterol on lipid conformations observed in simulations~\cite{zhu07,rog09,alwarawrah12} are overestimated by the computational models used;
+we calculated from CHARMM36 the glycerol backbone orientation and dihedral angle distributions
+at various cholesterol contents (Supporting material). The only detectable changes
+are the small decrease of gauche(-) and increase of gaughe(+) probability of 
+the g$_3$-g$_2$-g$_1$-O(\textit{sn}-1) dihedral and slight (less than 5 degrees) change
+in the glycerol backbone orientation. In conclusion, our results suggest that the significant 
+effects of cholesterol on lipid conformations observed in simulations~\cite{zhu07,rog09,alwarawrah12} 
+are overestimated by the computational models used;
 cholesterol only induces very small structural changes in the glycerol backbone.

### 2015-08-12 HGmodelMANUSCRIPT/REPLYtoREVIEWER.txt
Add glycerol orientation analysis and discussion to the manuscript, related to the reviewer's point 6

@@
 
-REPLY TO 6: This is indeed relevant for the simulations with cholesterol since the glycerol tilt 
-may also explain the minor changes in the order parameters when adding cholesterol. 
-WE SHOULD CALCULATE THIS FROM CHOLESTEROL SIMULATIONS WITH CHARMM SINCE THOSE ARE USED FOR THE 
-DIHEDRAL ANALYSIS IN SI. MAYBE HUBERT COULD DO THIS? 
+REPLY TO 6: This is a relevant point from the reviwer since the glycerol tilt 
+may indeed also explain the minor changes in the order parameters when adding cholesterol. 
+We have now added the analysis in the supplementary information and there is indeed
+a detectable change in the tilt. However, as in the dihedral angles this change is very small,
+thus our main conclusions are not changed. The related text in the manuscript is now rewritten:
+"Therefore, to resolve the nature of
+cholesterol-induced structural changes, we calculated from CHARMM36 the glycerol backbone
+orientation and dihedral angle distributions at various cholesterol contents (Supporting
+material). The only detectable changes are the small decrease of gauche(-) and increase of
+gaughe(+) probability of the g3-g2-g1-O(sn-1) dihedral and slight (less than 5 degrees) change
+in the glycerol backbone orientation. In conclusion, our results suggest that the significant
+effects of cholesterol on lipid conformations observed in simulations153-155 are overestimated
+by the computational models used; cholesterol only induces very small structural changes in
+the glycerol backbone."
 As a function of hydration level unfortunately (from page 24): 

### 2015-08-15 HGmodelMANUSCRIPT/HGmodel_ACSsupplementary.tex
Fix typo g_1-g_2 -> g_2-g_3

@@
   \caption{\label{GlycerolOrientation}
-    Orientation of g$_1$-g$_3$ (a) and g$_1$-g$_2$ (b) respect to membrane normal as a function of cholesterol content
+    Orientation of g$_1$-g$_3$ (a) and g$_2$-g$_3$ (b) respect to membrane normal as a function of cholesterol content
     calculated from CHARMM36 simulations in 298K and 303K.}

### 2015-08-15 HGmodelMANUSCRIPT/REPLYtoREVIEWER.txt
Update REPLYtoREVIEWER.txt
A typo in 4 (a) "We do not know that the other models have exatcly" was corrected.
Small updates to the reply to reviewer

@@
 a fixed membrane area or a nonzero surface tension, Charmm27 and Lipid11, were recently updated 
-to Charmm36 and Lipid14, which removed this requirement. The older versions of the force fields
+to Charmm36 and GAFFlipid/Lipid14, which removed this requirement. The older versions of the force fields
 ment to be ran in non-zero tension are not discussed in the manuscript. Thus, the more detailed discussion
 on this issue is beyond the scope of the work. Also, the improvements of the models and testing 
-of influence of changes in parameters, including the ensemble, is beyond the scope of the work. 
+the influence of changes in parameters, including the ensemble, is beyond the scope of the work. 
 Further, it is not very likely that the change in the ensemble would significantly change the 
@@
 realistic order parameters and this is in agreement with previous crystal structure and 1H NMR studies.19–21,23–25" 
-and about N-b-a-O dihedral it is writte on page 20: 
+and about N-b-a-O dihedral it is written on page 20: 
 "This result is again in agreement with previous crystal structure, 1H NMR and Raman spectroscopy studies19–22 
@@
 
-REPLY TO 4 (a): We do not know that the other models have exatcly the correct dihedral distribution 
+REPLY TO 4 (a): We do not know that the other models have exactly the correct dihedral distribution 
 and we do not claim this in the manuscript. Actually they cannot be all correct since they are not 
-the exactly same and also the order parameters are not the same and none of them really satisfy the 
+the exactly same and also the order parameters are not the same, and none of them really satisfy the 
 experiments, as discussed in the manuscript. However, as said in the manuscript and in the above reply, 
@@
 
-The justification for usage of phrase "biologically relevant conditions" in the manuscript is given 
+The justification for usage of the phrase "biologically relevant conditions" in the manuscript is given 
 in the sentence on page 4-5: 
@@
 the experimental parameters correctly (including order parameters) it would be correct with very high 
-probability, and would give a valid interpretation for experiments, as stated several times in the 
+probability, and would give a valid interpretation to experiments, as stated several times in the 
 manuscript, for example on page 21: 
@@
 and 31P-13C dipolar couplings 43 have been used to assess the quality of a simulation model."
-where Ref. 43 is the mentioned work. We also mention the work now also in page 21: 
+where Ref. 43 is the work by Prakash and Sankararamakrishnan. We also mention the work now also in page 21: 
 "...(like 31P chemical shift anisotropy 36 and 31P-13C dipolar couplings 43)..."

### 2015-08-19 HGmodelMANUSCRIPT/REPLYtoREVIEWER.txt

@@
-REPLY TO 1: We were trying to resolve the atomistic resolution structure of glycerol backbone 
-and choline by using the available atomistic resolution models. In all of the used models the 
-most recent version published by the developers was intended to be used in NPT ensemble with 
-zero tension, indicating that the general opinion in the field consider the zero tension ensemble 
-being more realistic. Most notably, the two popular lipid force fields that required the use of either 
+[I suggest reporting the question/comment before the reply, to make this more readable. LM]
+
+REPLY TO 1: In the present work, we analyze the atomic-level structure of the glycerol backbone 
+and the choline group of phosphatidylcholine lipids using available simulation models with atomistic resolution. 
+All the models used here were intended to be used in the NPT ensemble with 
+zero surface tension, indicating that the zero tension ensemble is currently considered the most appropriate ensemble for lipid blayer simulations by most experts in this field. 
+Notably, two popular lipid force fields that required the use of either 
 a fixed membrane area or a nonzero surface tension, Charmm27 and Lipid11, were recently updated 
-to Charmm36 and GAFFlipid/Lipid14, which removed this requirement. The older versions of the force fields
-ment to be ran in non-zero tension are not discussed in the manuscript. Thus, the more detailed discussion
-on this issue is beyond the scope of the work. Also, the improvements of the models and testing 
-the influence of changes in parameters, including the ensemble, is beyond the scope of the work. 
-Further, it is not very likely that the change in the ensemble would significantly change the 
-glycerol backbone or choline structure. We have clarified that we only reproduce the simulations
-with original parametrizations by the sentence: "All simulation parameters were set as close to the original 
-parametrization works as possible." in the page 10.
+(to Charmm36 and GAFFlipid/Lipid14), which removed this requirement. The older versions of the force fields
+meant to be used in non-zero tension are not discussed in the manuscript. Thus, we believe a detailed discussion
+of this issue is beyond the scope of the present work. 
+Also, the improvements of the models and testing the influence of changes in parameters, including the ensemble, is beyond the scope of the work. 
+Finally, it is not very likely that the change in the ensemble would significantly change the 
+glycerol backbone or choline structure. We clarified that we only reproduce the simulations
+with original parametrizations in the sentence: "All simulation parameters were set as close to the original 
+parametrization works as possible." (page 10).
 
@@
 "Since the glycerol backbone order parameters are very similar irrespective of the headgroup 
-chemistry (PC,PE and PG) or lipid environment,26 it is reasonable to assume that the stereospecifity 
-measured for the PE lipids holds also for the PC lipids." 
-It is extremely unlikely scenario that the stereospecific labeling would be different while the 
-order parameters would be numerically equal. Thus the assumption has quite strong experimental justification.
+chemistry (PC,PE, and PG) or lipid environment,(ref. 26) it is reasonable to assume that the stereospecifity 
+measured for the PE lipids holds also for PC lipids." 
+With numerically very similar order parameters, a difference in the stereospecific labeling represents a very unlikely scenario. Thus, our assumption has quite strong experimental justification.
 
-REPLY TO 3: The quantification used is explained in the figure caption. MAYBE MARKUS SHOULD CLARIFY THIS A LITTLE BIT? 
+REPLY TO 3: The quantification used is explained in the figure caption. MAYBE MARKUS SHOULD CLARIFY THIS A LITTLE BIT? AGREED, PLEASE ADD HERE MORE DETAILS (LM).
 It is based on the distance between simulated and experimental order parameters. The figure is made 
@@
 on the distance of the results from the experimental values, the order of the force fields in the figure will 
-not change. The color coding may change if another researcher prefers some other coding. Since all the data 
-is available, anyone is free to do that.
+not change.
 
-REPLY TO 4: As written in the manuscript there is actually also previous experimental support, in addition 
-to the current results, for the dihedral distributions suggested in the manuscript. About g3-g2-g1-O dihedral 
-it is written on page 19: 
+REPLY TO 4: As explained in the manuscript, there is actually also previous experimental support, in addition 
+to the current results, for the dihedral distributions suggested in the manuscript. 
+About g3-g2-g1-O dihedral, it is written on page 19: 
 "In conclusion, models preferring the anti conformation for this dihedral give more 
@@
 and we do not claim this in the manuscript. Actually they cannot be all correct since they are not 
-the exactly same and also the order parameters are not the same, and none of them really satisfy the 
-experiments, as discussed in the manuscript. However, as said in the manuscript and in the above reply, 
+the exactly same and also the order parameters are not the same, and none of them really satisfies the 
+experiments, as discussed in the manuscript. However, as explained in the manuscript and in the above reply, 
 there is strong evidence that dihedrals in other models are significantly more realistic than in Berger.
@@
 
-REPLY TO 4 (b): Yes, they can. However, if molecular dynamics simulation model would reproduce all 
-the experimental parameters correctly (including order parameters) it would be correct with very high 
+REPLY TO 4 (b): Yes, they can. However, if molecular dynamics simulations reproduced all 
+the experimental parameters correctly (including order parameters), it would be correct with very high 
 probability, and would give a valid interpretation to experiments, as stated several times in the 
@@
 real cells as well. However, it is true that, in principle one cannot exclude 
-other dihedrals than these two without further evidence. However, the previous experiments supports 
+other dihedrals than these two without further evidence. However, previous experiments support
 the conclusions and the main purpose of this analysis here is to  (from page 17-18) 
@@
 "We note that performing a similar comparison through all the dihedrals in all the 13 
-models would probably give highly useful information on how to improve the accuracy of the models 
+models would probably give highly useful information on how to improve the accuracy of the models, 
 yet this is beyond the scope of the current report."
@@
 and 31P-13C dipolar couplings 43 have been used to assess the quality of a simulation model."
-where Ref. 43 is the work by Prakash and Sankararamakrishnan. We also mention the work now also in page 21: 
+where Ref. 43 is the work by Prakash and Sankararamakrishnan. We also mention the work now also at page 21: 
 "...(like 31P chemical shift anisotropy 36 and 31P-13C dipolar couplings 43)..."
 
-REPLY TO 6: This is a relevant point from the reviwer since the glycerol tilt 
+REPLY TO 6: This is a very relevant point, since the glycerol tilt 
 may indeed also explain the minor changes in the order parameters when adding cholesterol. 
@@
 the glycerol backbone."
-As a function of hydration level unfortunately (from page 24): 
+As a function of hydration level, unfortunately (from page 24): 
 "The response of the glycerol backbone seems to be more subtle than that of the choline 
@@
 REPLY TO 7: The scope of this manuscript has been limited on purpose on the glycerol backbone 
-and choline segments to avoid exhausting the reader, thus this analysis is beyond the scope 
+and choline segments to avoid exhausting the reader, thus we believe this analysis is beyond the scope 
 of the work. The acyl chain region will be discussed in the context of the NMRLipids III 

### 2015-08-19 HGmodelMANUSCRIPT/REPLYtoREVIEWER.txt
Minor change in REPLYtoREVIEWER.txt

@@
 [I suggest reporting the question/comment before the reply, to make this more readable. LM]
+[This will be done finally, however, we we are not allowed to this publicly:
+http://nmrlipids.blogspot.com/2015/07/revision-requested-for-first-manuscript.html?showComment=1438341838923#c5455772628464564266]
 
 REPLY TO 1: In the present work, we analyze the atomic-level structure of the glycerol backbone 
-and the choline group of phosphatidylcholine lipids using available simulation models with atomistic resolution. 
+and the choline group of phosphatidylcholine lipids using available state of the art simulation models with atomistic resolution. 
 All the models used here were intended to be used in the NPT ensemble with 
-zero surface tension, indicating that the zero tension ensemble is currently considered the most appropriate ensemble for lipid blayer simulations by most experts in this field. 
+zero surface tension, indicating that the zero tension ensemble is currently considered the most appropriate 
+ensemble for lipid blayer simulations by most experts in this field. 
 Notably, two popular lipid force fields that required the use of either 

### 2015-08-22 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Updated caption of FF comparison table (Fig. 4).

@@
   \caption{\label{FullHydrationComparisonTable}
-Rough ranking of force fields based on data of Fig.~\ref{HGorderparameters}.
-"M" indicates a magnitude problem,
-"F" a forking problem.
-Letter size shows the level (0-4) of severity;
-the $\Sigma$-column shows the sum of these, i.e., the "total severity".
+Rough subjective ranking of force fields based on Fig.~\ref{HGorderparameters}.
+%
+Here "M" indicates a magnitude problem, "F" a forking problem;
+letter size increases with problem severity.
+%
 Color scheme:
@@
 "major deviation from experiments" (dark red).
+%
+The $\Sigma$-column shows the total deviation of the force field,
+when individual carbons are given weights of 0 (matches experiment), 1, 2, and 4 (major deviation).
+%
+For full details of the assessment, see Supporting Information.
 } 
@@
 we calculated from CHARMM36 the glycerol backbone orientation and dihedral angle distributions
-at various cholesterol contents (Supporting material). The only detectable changes
+at various cholesterol contents (Supporting Information). The only detectable changes
 are the small decrease of gauche(-) and increase of gaughe(+) probability of 
@@
 
-Simulation details, one figure and author contributions.
+Simulation and analysis details, two figures, and author contributions.
 

### 2015-08-22 HGmodelMANUSCRIPT/HGmodel_ACSsupplementary.tex
Added description of FF ranking. Added S in front of figure numbers. Fixed some typos.

@@
 \newcommand*\mycommand[1]{\texttt{\emph{#1}}}
+
+\renewcommand{\thefigure}{S\arabic{figure}}
 
@@
 
+\section{Details of the rough subjective force field ranking (Figure 4)}
+
+The assessment was based fully on the Fig. 2.
+%
+First, for each carbon (the columns in Fig. 4) in each force field (the rows),
+we looked separately at deviations in magnitude and forking.
+
+{\bf Magnitude} deviations, i.e., how close to the experimentally obtained C--H order parameters (OPs)
+the force-field-produced OPs were.
+%
+For each carbon, the following 5-step scale was used:
+%
+\begin{description}
+\item [0 (~):] \noindent {More than half of all the calculated OPs (that is, of all different hydrogens in all different lipids) were within the {\it subjective sweet spots} (SSP, blue-shaded areas in Fig.
+2).}
+%
+\item [1 ({\textsf{\tiny M}}):] \noindent {All the calculated OPs  were $< 0.03$ units away from the SSP.}
+%
+\item [2  ({\textsf{\small M}}):] \noindent {All the calculated OPs  were $< 0.05$ units away from the SSP.}
+%
+\item [3 ({\textsf{\large M}}):] \noindent {All the calculated OPs  were $< 0.10$ units away from the SSP.}
+%
+\item [4 ({\textsf{\Large M}}):] \noindent {Some of the calculated OPs  were $> 0.10$ units  away from the SSP.}
+\end{description}
+
+{\bf Forking} deviations, i.e., how well the difference in order parameters of two hydrogens attached to a given carbon matched that obtained experimentally.
+Note that this is not relevant for $\mathrm{g_2}$, which has only one hydrogen.
+For the $\beta$ and $\alpha$ carbons, for which no forking is experimentally seen, the following 5-step scale was used:
+\begin{description}
+\item [0 (~):] \noindent {The distance $D$ between the dots (that mark the measurement-time-weighted averages in Fig.~2) was $< 0.005$ units for all the calculated OPs (that is, for all different lipids).}
+%
+\item [1 ({\textsf{\tiny F}}):] \noindent {$D < 0.015$.}
+%
+\item [2  ({\textsf{\small F}}):] \noindent {$ 0.015 < D < 0.030$.}
+%
+\item [3 ({\textsf{\large F}}):] \noindent {$ 0.030 < D < 0.050$.}
+%
+\item [4 ({\textsf{\Large F}}):] \noindent {$ 0.050 < D $.}
+\end{description}
+%
+For the $\mathrm{g_3}$ carbon, for which a small forking of 0.02 is indicated by experiments, the following 5-step scale was used:
+%
+\begin{description}
+\item [0 (~):] \noindent {$0.010 < D< 0.030$.}
+%
+\item [1 ({\textsf{\tiny F}}):] \noindent {$(D <0.010)$  OR $(0.030 < D < 0.040)$.}
+%
+\item [2  ({\textsf{\small F}}):] \noindent {$0.040 < D < 0.055$.}
+%
+\item [3 ({\textsf{\large F}}):] \noindent {$0.055 < D < 0.070$.}
+%
+\item [4 ({\textsf{\Large F}}):] \noindent {$0.070 < D$.}
+\end{description}
+%
+For the $\mathrm{g_1}$ carbon, for which a considerable forking of 0.15 is experimentally seen, the following 5-step scale was used:
+%
+\begin{description}
+\item [0 (~):] \noindent {$0.13 < D < 0.17$.}
+%
+\item [1 ({\textsf{\tiny F}}):] \noindent {$(0.11 < D < 0.13)$ OR $(0.17 < D < 0.19)$.}
+%
+\item [2  ({\textsf{\small F}}):] \noindent {$(0.09 < D < 0.11)$ OR $(0.19 < D < 0.21)$.}
+%
+\item [3 ({\textsf{\large F}}):] \noindent {$(0.07 < D < 0.09)$ OR $(0.21 < D < 0.23)$.}
+%
+\item [4 ({\textsf{\Large F}}):] \noindent {$(D<0.07)$ OR $(0.23<D)$.}
+\end{description}
+
+Based on these assessments of magnitude and forking deviations,
+each carbon was then assigned to one of the following groups:
+"within experimental error"
+(magnitude and forking deviations both on step 0 of the scales described above),
+"almost within experimental error"
+(sum of the magnitude and forking deviation steps 1 or 2),
+"clear deviation from experiments"
+(sum of magnitude and forking deviation steps from 3 to 5), and
+"major deviation from experiments"
+(sum of magnitude and forking deviation steps from 6 to 8).
+These groups are indicated by colors in Fig.~2.
+(Note that for $\mathrm{g_2}$, for which there can be no forking,
+the corresponding group assigment limits were: 0, 1, 2, and 3.)
+
+Finally, the total ability of the force field to describe the headgroup and
+glycerol structure was estimated.
+To this end, the groups were given the following weights:
+0 (within experimental error),
+1 (almost within experimental error),
+2 (clear deviation from experiments),
+4 (major deviation from experiments),
+and the weights of the five carbons were summed up.
+The sum, given in the $\Sigma$-column of Fig.~2,
+was then used to (roughly and subjectively, as should be clear from the
+above description) rank the force fields.
+
 \section{The effect of cholesterol on glycerol backbone structure}
@@
 with popc and cholesterol for the CHARMM36 All-Atom FF. \\
-{\it Patrick F. J. Fuchs} Ran and analyzed the Poger simulations. Provided scientific information which significantly advanced the project (signs and forking of order parameters). \\
+{\it Patrick F. J. Fuchs} ran and analyzed the Poger simulations. Provided scientific information which significantly advanced the project (signs and forking of order parameters). \\
 {\it Matti Javanainen} prepared and performed simulations with multiple lipid models and analyzed the results. Supervised the work of JT.\\
@@
 {\it Waldemar Kulig} prepared the MD simulations with cholesterol for the MacRog FF \\
-{\it Antti Lamberg} Prepared and performed simulations with Berger and its variants to show the importance of the signs and stereospecific labeling of the order parameters. \\
+{\it Antti Lamberg} prepared and performed simulations with Berger and its variants to show the importance of the signs and stereospecific labeling of the order parameters. \\
 {\it Claire Loison} provided simulation results for charmm36-UA. \\
 {\it Alexander Lyubartsev} provided simulation results for Högberg08 force field.\\
-{\it Markus S. Miettinen} Co-designed the project with OHSO and supported in the work management. Provided trajectories for Berger DMPC; prepared parameter files for 72 lipids \\
-{\it Luca Monticelli} Critical discussions in all phases of the project. Collaboration with Jukka  M{\"a}{\"a}tt{\"a}. \\
+{\it Markus S. Miettinen} co-designed the project with OHSO and supported in the work management. Provided Berger DMPC trajectory; prepared parameter files for 72 lipids %POPC \& DPPC
+CHARMM36. Made Figures 2--4 and 7--9.\\
+{\it Luca Monticelli}: Critical discussions in all phases of the project. Collaboration with Jukka  M{\"a}{\"a}tt{\"a}. \\
 {\it Jukka M{\"a}{\"a}tt{\"a}}  prepared and performed simulations with Berger and Slipids models and analyzed the results.\\
-{\it O. H. Samuli Ollila} Co-designed the project with MSM and managed the work. Ran and analyzed several simulations. Wrote the manuscript. \\
-{\it Marius Retegan} Prepared and validated the GROMACS compatible parameter files for GAFFlipid and Lipid14 force fields.\\
-{\it Tomasz Rog} Provided the topologies for the MacRog model and also the full hydration simulation data with this model. \\
+{\it O. H. Samuli Ollila} co-designed the project with MSM and managed the work. Ran and analyzed several simulations. Wrote the manuscript. \\
+{\it Marius Retegan} prepared and validated the GROMACS compatible parameter files for GAFFlipid and Lipid14 force fields.\\
+{\it Tomasz Rog} provided the topologies for the MacRog model and also the full hydration simulation data with this model. \\
 {\it Hubert Santuz} prepared and performed the cholesterol simulations with CHARMM36 and analyzed the results. 

### 2015-08-22 HGmodelMANUSCRIPT/REPLYtoREVIEWER.txt
Rewrote parts discussing Fig 4. Fixed a few minor typos.

@@
 
-REPLY TO 1: In the present work, we analyze the atomic-level structure of the glycerol backbone 
-and the choline group of phosphatidylcholine lipids using available state of the art simulation models with atomistic resolution. 
-All the models used here were intended to be used in the NPT ensemble with 
-zero surface tension, indicating that the zero tension ensemble is currently considered the most appropriate 
+REPLY TO 1: In the present work, we analyze the atomic-level structure of the glycerol backbone and the choline group of phosphatidylcholine lipids using available state of the art simulation models with atomistic resolution.
+All the models used here were intended to be used in the NPT ensemble with zero surface tension, indicating that the zero tension ensemble is currently considered the most appropriate
 ensemble for lipid blayer simulations by most experts in this field.
-Notably, two popular lipid force fields that required the use of either 
-a fixed membrane area or a nonzero surface tension, Charmm27 and Lipid11, were recently updated 
-(to Charmm36 and GAFFlipid/Lipid14), which removed this requirement. The older versions of the force fields
-meant to be used in non-zero tension are not discussed in the manuscript. Thus, we believe a detailed discussion
-of this issue is beyond the scope of the present work. 
+
+Notably, two popular lipid force fields that required the use of either a fixed membrane area or a nonzero surface tension, Charmm27 and Lipid11, were recently updated (to Charmm36 and GAFFlipid/Lipid14), which removed this requirement.
+The older versions of the force fields, meant to be used in non-zero tension, are not discussed in the manuscript.
+Thus, we believe a detailed discussion of this issue is beyond the scope of the present work.
 Also, the improvements of the models and testing the influence of changes in parameters, including the ensemble, is beyond the scope of the work.
-Finally, it is not very likely that the change in the ensemble would significantly change the 
-glycerol backbone or choline structure. We clarified that we only reproduce the simulations
-with original parametrizations in the sentence: "All simulation parameters were set as close to the original 
-parametrization works as possible." (page 10).
 
-REPLY TO 2: It is known experimentally that the glycerol backbone order parameters are essentially 
-the same for PC, PE and PG lipids, as written in the manuscript (page 15): 
-"Since the glycerol backbone order parameters are very similar irrespective of the headgroup 
-chemistry (PC,PE, and PG) or lipid environment,(ref. 26) it is reasonable to assume that the stereospecifity 
-measured for the PE lipids holds also for PC lipids." 
-With numerically very similar order parameters, a difference in the stereospecific labeling represents a very unlikely scenario. Thus, our assumption has quite strong experimental justification.
+Finally, it is not very likely that the change in the ensemble would significantly change the glycerol backbone or choline structure.
+We clarified that we only reproduce the simulations with original parametrizations in the sentence: "All simulation parameters were set as close to the original parametrization works as possible." (page 10).
 
-REPLY TO 3: The quantification used is explained in the figure caption. MAYBE MARKUS SHOULD CLARIFY THIS A LITTLE BIT? AGREED, PLEASE ADD HERE MORE DETAILS (LM).
-It is based on the distance between simulated and experimental order parameters. The figure is made 
-and color coding is used to make the message in Fig. 2 more easily readable, thus all the details are 
-also available for the readers who are willing to take a closer look. If the force fields are ranked based 
-on the distance of the results from the experimental values, the order of the force fields in the figure will 
-not change.
+REPLY TO 2: It is known experimentally that the glycerol backbone order parameters are essentially the same for PC, PE and PG lipids, as written in the manuscript (page 15): "Since the glycerol backbone order parameters are very similar irrespective of the headgroup chemistry (PC,PE, and PG) or lipid environment,(ref.
+26) it is reasonable to assume that the stereospecifity measured for the PE lipids holds also for PC lipids." With numerically very similar order parameters, a difference in the stereospecific labeling represents a very unlikely scenario.
+Thus, our assumption has quite strong experimental justification.
 
-REPLY TO 4: As explained in the manuscript, there is actually also previous experimental support, in addition 
-to the current results, for the dihedral distributions suggested in the manuscript. 
-About g3-g2-g1-O dihedral, it is written on page 19: 
-"In conclusion, models preferring the anti conformation for this dihedral give more 
-realistic order parameters and this is in agreement with previous crystal structure and 1H NMR studies.19–21,23–25" 
-and about N-b-a-O dihedral it is written on page 20: 
-"This result is again in agreement with previous crystal structure, 1H NMR and Raman spectroscopy studies19–22 
-which suggest that this dihedral is in the gauche conformation in the absence of ions."
+REPLY TO 3: We wish to emphasize the discussive nature of Fig. 4. That is, the actual results are given in Fig. 2, and Fig. 4 is just a (by
+necessity subjective) interpretation of the key features of the flood of information that Fig.
+2 contains.
+We have made this distinction clearer to the reader by adding explicitly the word "subjective" to the caption of Fig.
+4.
 
-REPLY TO 4 (a): We do not know that the other models have exactly the correct dihedral distribution 
-and we do not claim this in the manuscript. Actually they cannot be all correct since they are not 
-the exactly same and also the order parameters are not the same, and none of them really satisfies the 
-experiments, as discussed in the manuscript. However, as explained in the manuscript and in the above reply, 
-there is strong evidence that dihedrals in other models are significantly more realistic than in Berger.
+We also understand the reviewer's worry that Fig.
+4 might be taken too literally, if the basis for assessment is not spelled out in quantitative terms.
+To this end, we have added a detailed description of the assessment to the Supplementary Information.
+Please note that as we needed to make the quantitative limits as simple as possible (e.g., using the same magnitude scheme for all carbons) to ease their description, Fig.
+4 has slightly changed.
+Its key message, coming from the data of Fig.
+2, of course remains unchanged.
 
-The justification for usage of the phrase "biologically relevant conditions" in the manuscript is given 
-in the sentence on page 4-5: 
-"Importantly, the structural parameters for the glycerol backbone are similar for various biologically 
-relevant lipid species (phosphatidylcholine (PC), phosphatidylethanolamine (PE) and phosphatidylglycerol (PG)) 
-in various environments,26 and the structural parameters for the choline headgroup are similar in model 
-membranes and real cells (mouse fibroblast L-M cell).27". 
-Since the order parameters for these parts are the same in real cells (at least E. Coli and mouse 
-fibroblast L-M cell) as in model systems, most likely the structure is the same as well.
+REPLY TO 4: As explained in the manuscript, there is actually also previous experimental support, in addition to the current results, for the dihedral distributions suggested in the manuscript.
+Concerning the g3-g2-g1-O dihedral, it is written on page 19 that "In conclusion, models preferring the anti conformation for this dihedral give more realistic order parameters and this is in agreement with previous crystal structure and 1H NMR studies.
+19–21,23–25".
+Concerning the N-b-a-O dihedral it is written on page 20 that "This result is again in agreement with previous crystal structure, 1H NMR and Raman spectroscopy studies 19–22 which suggest that this dihedral is in the gauche conformation in the absence of ions."
 
-REPLY TO 4 (b): Yes, they can. However, if molecular dynamics simulations reproduced all 
-the experimental parameters correctly (including order parameters), it would be correct with very high 
-probability, and would give a valid interpretation to experiments, as stated several times in the 
-manuscript, for example on page 21: 
-"If the improved atomistic molecular dynamics simulation model reproduced the order parameters 
-and other experimental observables (like 31P chemical shift anisotropy 36 and 31P-13C dipolar couplings 43) 
-with experimental accuracy, it would give an interpretation for the atomistic resolution structure of the 
-glycerol backbone and choline.10–13,15,16,18".
+REPLY TO 4 (a): We do not know that the other models have exactly the correct dihedral distribution and we do not claim this in the manuscript.
+Actually they cannot be all correct since they are not the exactly same, and also the order parameters are not the same.
+In fact, none of the models really satisfies the experiments, as discussed in the manuscript.
+However, as explained in the manuscript and in the above reply, there is strong evidence that dihedrals in other models are significantly more realistic than in Berger.
 
-REPLY TO 4 (c): As stated in the replies above, there is also previous experimental evidence for 
-the conclusions made about dihedral distributions and these conclusions can be extended for the 
-real cells as well. However, it is true that, in principle one cannot exclude 
-other dihedrals than these two without further evidence. However, previous experiments support
-the conclusions and the main purpose of this analysis here is to  (from page 17-18) 
-"demonstrate that 1) the differences in order parameters indicate significantly different 
-structural sampling strongly correlating with the dihedral angles of the related bonds, 
-and that 2) the comparison between experimental and simulated order parameters can be used 
-to exclude nonrealistic structural samping in molecular dynamics simulations." 
-And further from page 19 
-"We note that performing a similar comparison through all the dihedrals in all the 13 
-models would probably give highly useful information on how to improve the accuracy of the models, 
-yet this is beyond the scope of the current report."
+We note that the justification for usage of the phrase "biologically relevant conditions" in the manuscript is given in the sentence on page 4-5: "Importantly, the structural parameters for the glycerol backbone are similar for various biologically relevant lipid species (phosphatidylcholine (PC), phosphatidylethanolamine (PE) and phosphatidylglycerol (PG)) in various environments,26 and the structural parameters for the choline headgroup are similar in model membranes and real cells (mouse fibroblast L-M cell).27".
+Since the order parameters for these parts are the same in real cells (at least E.
+Coli and mouse fibroblast L-M cell) as in model systems, most likely the structure is the same as well.
 
-REPLY TO 5: This is indeed relevant work and we were aware of it. However, the reference was accidentally 
-left out from the manuscript. We thank the reviewer for pointing this out.
-This work is now mentioned in page 5:
-"In addition to the order parameters, dihedral angles for the glycerol
-backbone and headgroup estimated from experiments,28,38-42 31P chemical shift anisotropy 36
-and 31P-13C dipolar couplings 43 have been used to assess the quality of a simulation model."
-where Ref. 43 is the work by Prakash and Sankararamakrishnan. We also mention the work now also at page 21: 
-"...(like 31P chemical shift anisotropy 36 and 31P-13C dipolar couplings 43)..."
+REPLY TO 4 (b): Yes, they can.
+However, if molecular dynamics simulations reproduced all the experimental parameters correctly (including order parameters), it would be correct with very high probability, and would give a valid interpretation of experiments, as stated several times in the manuscript, for example on page 21: "If the improved atomistic molecular dynamics simulation model reproduced the order parameters and other experimental observables (like 31P chemical shift anisotropy 36 and 31P-13C dipolar couplings 43) with experimental accuracy, it would give an interpretation for the atomistic resolution structure of the glycerol backbone and choline.10–13,15,16,18".
 
-REPLY TO 6: This is a very relevant point, since the glycerol tilt 
-may indeed also explain the minor changes in the order parameters when adding cholesterol. 
-We have now added the analysis in the supplementary information and there is indeed
-a detectable change in the tilt. However, as in the dihedral angles this change is very small,
-thus our main conclusions are not changed. The related text in the manuscript is now rewritten:
-"Therefore, to resolve the nature of
-cholesterol-induced structural changes, we calculated from CHARMM36 the glycerol backbone
-orientation and dihedral angle distributions at various cholesterol contents (Supporting
-material). The only detectable changes are the small decrease of gauche(-) and increase of
-gaughe(+) probability of the g3-g2-g1-O(sn-1) dihedral and slight (less than 5 degrees) change
-in the glycerol backbone orientation. In conclusion, our results suggest that the significant
-effects of cholesterol on lipid conformations observed in simulations153-155 are overestimated
-by the computational models used; cholesterol only induces very small structural changes in
-the glycerol backbone."
-As a function of hydration level, unfortunately (from page 24): 
-"The response of the glycerol backbone seems to be more subtle than that of the choline 
-headgroup; none of the four models reproduced the experimental trends upon dehydration 
-with enough accuracy to invite a structural interpretation."
+REPLY TO 4 (c): As stated in the replies above, there is also previous experimental evidence for the conclusions made concerning the dihedral distributions and these conclusions can be extended for the real cells as well.
+However, it is true that in principle one cannot exclude other dihedrals than these two without further evidence.
+However, previous experiments support the conclusions, and the main purpose of this analysis here is to (from pages 17-18) "demonstrate that 1) the differences in order parameters indicate significantly different structural sampling strongly correlating with the dihedral angles of the related bonds, and that 2) the comparison between experimental and simulated order parameters can be used to exclude nonrealistic structural samping in molecular dynamics simulations." And further, from page 19,
+"We note that performing a similar comparison through all the dihedrals in all the 13 models would probably give highly useful information on how to improve the accuracy of the models, yet this is beyond the scope of the current report."
 
-REPLY TO 7: The scope of this manuscript has been limited on purpose on the glycerol backbone 
-and choline segments to avoid exhausting the reader, thus we believe this analysis is beyond the scope 
-of the work. The acyl chain region will be discussed in the context of the NMRLipids III 
-project (see http://nmrlipids.blogspot.fi/p/about.html).
+REPLY TO 5: This is indeed relevant work and we were aware of it.
+However, the reference was accidentally left out from the manuscript.
+We thank the reviewer for pointing this out.
+This work is now mentioned in page 5: "In addition to the order parameters, dihedral angles for the glycerol backbone and headgroup estimated from experiments,28,38-42 31P chemical shift anisotropy 36 and 31P-13C dipolar couplings 43 have been used to assess the quality of a simulation model." where Ref.
+43 is the work by Prakash and Sankararamakrishnan.
+We also mention the work now also at page 21:  "...(like 31P chemical shift anisotropy 36 and 31P-13C dipolar couplings 43)..."
+
+REPLY TO 6: This is a very relevant point, since the glycerol tilt may indeed also explain the minor changes in the order parameters when adding cholesterol.
+We have now added the analysis in the supplementary information and there is indeed a detectable change in the tilt.
+However, as in the dihedral angles this change is very small, thus our main conclusions are not changed.
+The related text in the manuscript is now rewritten: "Therefore, to resolve the nature of cholesterol-induced structural changes, we calculated from CHARMM36 the glycerol backbone orientation and dihedral angle distributions at various cholesterol contents (Supporting Information).
+The only detectable changes are the small decrease of gauche(-) and increase of gaughe(+) probability of the g3-g2-g1-O(sn-1) dihedral and slight (less than 5 degrees) change in the glycerol backbone orientation.
+In conclusion, our results suggest that the significant effects of cholesterol on lipid conformations observed in simulations153-155 are overestimated by the computational models used; cholesterol only induces very small structural changes in the glycerol backbone."
+
+As a function of hydration level, unfortunately (from page 24): "The response of the glycerol backbone seems to be more subtle than that of the choline headgroup; none of the four models reproduced the experimental trends upon dehydration with enough accuracy to invite a structural interpretation."
+
+REPLY TO 7: The scope of this manuscript has been limited on purpose on the glycerol backbone and choline segments to avoid exhausting the reader, thus we believe this analysis is beyond the scope of the work.
+The acyl chain region will be discussed in the context of the NMRLipids III project (see http://nmrlipids.blogspot.fi/p/about.html).

### 2015-08-22 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Add doi created for GitHub repo

@@
 Further information, such as scripts, figures, and manuscript text files, are available through our GitHub 
-repository (\url{https://github.com/NMRlipids}).
+repository~\cite{NMRLipidsGit}.% (\url{https://github.com/NMRlipids}).
 

### 2015-10-11 HGmodelMANUSCRIPT/REPLYtoREVIEWER2.txt
English corrections by the reviewer and typo from https://github.com/NMRLipids/nmrlipids.blogspot.fi/issues/65 were fixed. The first draft of the second reply to the reviewers added

Reviewer 2.

We thank for the very positive feedback. The errors in english are
now fixed as suggested.

Reviewer1.

1.
Author reply:

2.
Author reply:
The subjective sweet spots span +-0.02 AROUND WHAT? IS THIS THE AVERAGE
OF EXPERIMENTAL VALUE?

3.
Author reply:
The title of our manuscript is:
"Towards atomistic resolution structure of
phosphatidylcholine headgroup and glycerol
backbone at different ambient conditions"
in the introduction we write:
"Phospholipid molecules are composed of hydrophobic acyl chains connected
by a glycerol backbone to a hydrophilic headgroup ..."
This is widely used naming convention for phospholipid molecules.
According to the title and the naming convetion used in the manuscript, 
the beginning of acyl chains are not the part of headgroup and are 
beoynd the scope defined in the title. Their presence at the interface
or their chemical environment do not change this.

We are not willing to extend the scope of the manuscript since it
would signifincantly increase already massive information content and 
length. Instead, the acyl chains order parameters will be discussed in 
the context of the NMRLipids III project, see 
http://nmrlipids.blogspot.fi/p/about.html.

### 2015-10-11 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Add TOC in the manuscript
English corrections by the reviewer and typo from https://github.com/NMRLipids/nmrlipids.blogspot.fi/issues/65 were fixed. The first draft of the second reply to the reviewers added

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\begin{tocentry}
+\begin{tocentry}
 
+%\begin{figure}[]
+%  \centering
+  \includegraphics[width=9cm]{../Fig/TOCfinal.png}
+
+%  \caption{\label{POPCstructure}
+%    Chemical structure of  1-palmitoyl-2-oleoylphosphatidylcholine (POPC).}
+
+%\end{figure}
 %Some journals require a graphical entry for the Table of Contents.
@@
 
-%\end{tocentry}
+\end{tocentry}
 
@@
 are different. The absolute values of order parameters from the measured quadrupolar splitting $\Delta \nu_Q$ ($^2$H NMR) are calculated using 
-the equation $|S_{{\rm CD}}|=\frac{4}{3} \frac{e^2qQ}{h} \Delta \nu_Q$, where the value for the static quadrupole
-splitting constant is estimated from various experiments to be 170~kHz leading to a numerical relation $|S_{{\rm CD}}|=0.00784 \times \Delta \nu_Q$~\cite{seelig77c}. 
+the equation $|S_{{\rm CD}}|=\frac{4}{3} \frac{h}{e^2qQ} \Delta \nu_Q$, where the value for the static quadrupole
+splitting constant is estimated from various experiments to be 170~kHz leading to a numerical relation 
+$|S_{{\rm CD}}|=0.00784 \times \Delta \nu_Q$~\cite{seelig77c}. 
 The absolute values of order parameters from the effective dipolar coupling $d_\mathrm{CH}$ ($^1$H-$^{13}$C NMR) are calculated using the equation
@@
 should be preferably inside to the subjective sweet spots determined from experiments (blue shaded regions in Fig.~\ref{HGorderparameters}).
-The results for each force field in respect to the above criteria are summarized in Figure~\ref{FullHydrationComparisonTable}.
+The results for each force field with respect to the above criteria are summarized in Figure~\ref{FullHydrationComparisonTable}.
 \begin{figure}[]
@@
 
-None of the studied force fields fulfils these criteria completely, however CHARMM36 is pretty close. 
-This is not surprising since the dihedral potentials in this model are tuned to reproduce these parameters better against experiments~\cite{klauda10}.
+None of the studied force fields fulfils these criteria completely, however CHARMM36 is close. 
+This is not surprising since the dihedral potentials in this model are tuned to reproduce these 
+parameters better against experiments~\cite{klauda10}.
 The next models in the list are CHARMM36-UA~\cite{henin08,lee14} and H\"ogberg08~\cite{hogberg08} which is also not surprising since
@@
 These models are selected for more detailed studies since they are the best representatives of different dihedral potential parametrization techniques 
-(CHARMM36, MacRog, GAFFlipid), and the Berger based models are the most used lipid model in the literature.
+(CHARMM36, MacRog, GAFFlipid), and the Berger based models are the most used lipid models in the literature.
 

### 2015-10-11 HGmodelMANUSCRIPT/REPLYtoREVIEWER2.txt
Update REPLYtoREVIEWER2.txt

@@
 
-We thank for the very positive feedback. The errors in english are
+We thank for the very positive feedback. The errors in English are
 now fixed as suggested.
 
-Reviewer1.
+Reviewer 1.
 
@@
 Author reply:
-The subjective sweet spots span +-0.02 AROUND WHAT? IS THIS THE AVERAGE
-OF EXPERIMENTAL VALUE?
+
+All the "subjective sweet spots" in Fig. 2 have exactly the same size: 0.04 units.
+
+This size was agreed on based on the high experimental repeatibility of the order
+parameters that we discuss in the text (p12): "Almost all the reported values are
+within a variation of ±0.02 [...] for all fully hydrated PC bilayers, regardless
+of the variation in their acyl chain composition and temperature."
+
+To make the (equal) size of the sweet spots clear to the reader, we now explicitly
+point out this size when introcuding them in the text (p15):
+
+Motivated by the high experimental reproducibility, we have highlighted in Fig. 2
+the subjective sweet spots (light blue areas **spanning 0.04 units**), within which
+we expect the calculated absolute values of the order parameters of a well-performing
+force field to fall.
+
+The added words are those between the stars (**).
 

### 2015-10-11 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Added explicitly the size of the subjective sweet spots.

@@
 reported values are within a variation of $\pm$0.02 (which is also the error estimate given by Gross et al.~\cite{gross97}) 
-for all fully hydrated PC bilayer, regardless of the variation in their acyl chain composition and temperature.
+for all fully hydrated PC bilayers, regardless of the variation in their acyl chain composition and temperature.
 Exceptions are the somewhat lower order parameters sometimes reported from measurements using $^1$H-$^{13}$C NMR~\cite{hong95a,hong95b,warschawski05}.
@@
 Motivated by the high experimental reproducibility, we have highlighted in 
-Fig.~\ref{HGorderparameters} the subjective sweet spots (light blue areas), within which we expect the calculated absolute 
+Fig.~\ref{HGorderparameters} the subjective sweet spots (light blue areas spanning 0.04 units), within which we expect the calculated absolute 
 values of the order parameters of a well-performing force field to fall.

### 2015-10-11 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Fixed typos.

@@
 the ends of the bars mark the extreme values from the sets, and the dot marks their measurement-time-weighted average. 
-The interactive version of this figure is available at  https://plot.ly/$\sim$HubertSantuz/72/lipid-force-field-comparison/.
+An interactive version of this figure is available at  https://plot.ly/$\sim$HubertSantuz/72/lipid-force-field-comparison/.
 } 
@@
 Exceptions are the somewhat lower order parameters sometimes reported from measurements using $^1$H-$^{13}$C NMR~\cite{hong95a,hong95b,warschawski05}.
-These experiments are not shown in Fig.~\ref{HGorderparameters} as the reported error bars are either relatively large~\cite{hong95a,hong95b}, 
-or the spectral resolution is quite low and the numerical lineshape simulations have not been used in the analysis~\cite{warschawski05}.
-Due to this end, it is highly likely that these reported lower order parameters are due to lower experimental 
-accuracy and therefore we exclude them from our discussion. 
-For more details, see Ref.~\citenum{accuracyPOST}. 
+In these experiments, however, the reported error bars are either relatively large~\cite{hong95a,hong95b}, 
+or the spectral resolution is quite low and numerical lineshape simulations have not been used in the analysis~\cite{warschawski05}.
+As it, therefore, is highly likely that the reported lower order parameters are due to lower experimental 
+accuracy, we exclude them from the present discussion;
+for more details, see Ref.~\citenum{accuracyPOST}. 
 Motivated by the high experimental reproducibility, we have highlighted in 
-Fig.~\ref{HGorderparameters} the subjective sweet spots (light blue areas spanning 0.04 units), within which we expect the calculated absolute 
+Fig.~\ref{HGorderparameters} subjective sweet spots (light blue areas spanning 0.04 units), within which we expect the calculated  
 values of the order parameters of a well-performing force field to fall.

### 2015-10-11 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Fixed language.

@@
 The specific deuteration of $\alpha$-, $\beta$- and g$_3$- segments of DPPC has been successful, 
-allowing the absolute value order parameter measurements for these segments by $^2$H NMR~\cite{gally75,brown77,brown78,akutsu81}.
+allowing the absolute values of the order parameters for these segments to be measured  by $^2$H NMR~\cite{gally75,brown77,brown78,akutsu81}.
 In addition, the absolute values of order parameters for all glycerol backbone and choline headgroup segments in egg yolk lecithin~\cite{hong95a},
-DMPC~\cite{hong95b,gross97,dvinskikh05a}, DOPC~\cite{warschawski05} and POPC~\cite{warschawski05,ferreira13}
-have been measured with several different implementations of $^1$H-$^{13}$C NMR experiments. In addition, the signs of order parameters in some systems
-are measured with $^1$H-$^{13}$C NMR techniques~\cite{hong95a,hong95b,gross97}. The experimental values of glycerol backbone 
-and choline order parameters from various publications~\cite{gally75,akutsu81,gross97,dvinskikh05a,ferreira13} with the signs measured in~\cite{hong95a,hong95b,gross97} are shown in Fig.~\ref{HGorderparameters}.
+DMPC~\cite{hong95b,gross97,dvinskikh05a}, DOPC~\cite{warschawski05}, and POPC~\cite{warschawski05,ferreira13}
+have been measured with several different implementations of $^1$H-$^{13}$C NMR experiments. Furthermore, for some systems the signs of the order parameters 
+have been measured with $^1$H-$^{13}$C NMR techniques~\cite{hong95a,hong95b,gross97}. The experimental values of the glycerol backbone 
+and choline order parameters from various publications~\cite{gally75,akutsu81,gross97,dvinskikh05a,ferreira13}, with the signs measured in Refs.~\citenum{hong95a,hong95b,gross97}, are shown in Fig.~\ref{HGorderparameters}.
 \begin{figure}[]
@@
 reported values are within a variation of $\pm$0.02 (which is also the error estimate given by Gross et al.~\cite{gross97}) 
-for all fully hydrated PC bilayers, regardless of the variation in their acyl chain composition and temperature.
-Exceptions are the somewhat lower order parameters sometimes reported from measurements using $^1$H-$^{13}$C NMR~\cite{hong95a,hong95b,warschawski05}.
+for all fully hydrated PC bilayers, regardless of variation in their acyl chain composition and temperature.
+Exceptions are the somewhat lower order parameters reported from some measurements using $^1$H-$^{13}$C NMR~\cite{hong95a,hong95b,warschawski05}.
 In these experiments, however, the reported error bars are either relatively large~\cite{hong95a,hong95b}, 
@@
 In addition to the numerical values, an important feature of the glycerol backbone is the 
-forking (see section~\ref{expORDp}) of the order parameters in g$_1$ and g$_3$ segments, in contrast to the choline segments $\alpha$ and $\beta$. 
+forking (see Sec.~\ref{expORDp}) of the order parameters in g$_1$ and g$_3$ segments, in contrast to the choline segments $\alpha$ and $\beta$. 
 The forking in glycerol backbone g$_3$ segment is small ($\approx$ 0.02) 
-and some experiments only report the larger value or the average value~\cite{akutsu81,ferreira13}. 
+and some experiments only report the larger or the average value~\cite{akutsu81,ferreira13}. 
 In contrast, forking is significant for the glycerol backbone g$_1$ segment, whose lower order parameter is close to zero and the
@@
 extract. This experiment gave the lower order parameter when deuterium was in the S position of g$_1$ or R position for g$_3$.
-Since the glycerol backbone order parameters are very similar irrespective of the headgroup chemistry (PC,PE and PG) or lipid 
+Since the glycerol backbone order parameters are very similar irrespective of the headgroup chemistry (PC, PE and PG) or lipid 
 environment~\cite{gally81}, it is reasonable to assume that the stereospecifity measured for the PE lipids
@@
 segments of POPC bilayer is collected by taking the absolute values from~\cite{ferreira13}, the signs from~\cite{hong95a,hong95b,gross97} 
-and the stereospecific labeling from~\cite{gally81}, and shown in Fig.~\ref{HGorderparameters2}.
+and the stereospecific labeling from~\cite{gally81}, and is shown in Fig.~\ref{HGorderparameters2}.
 \begin{figure*}[]

### 2015-10-11 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Tuned language.

@@
 
-In addition to the numerical values, an important feature of the glycerol backbone is the 
-forking (see Sec.~\ref{expORDp}) of the order parameters in g$_1$ and g$_3$ segments, in contrast to the choline segments $\alpha$ and $\beta$. 
-The forking in glycerol backbone g$_3$ segment is small ($\approx$ 0.02) 
+In addition to the numerical values, forking (see Sec.~\ref{expORDp}) is an important feature of the order parameters.
+In contrast to the lack of forking in the choline segments $\alpha$ and $\beta$, both CH$_2$ segments of the glycerol backbone fork.
+In the g$_3$ segment forking is small ($\approx$ 0.02), 
 and some experiments only report the larger or the average value~\cite{akutsu81,ferreira13}. 
-In contrast, forking is significant for the glycerol backbone g$_1$ segment, whose lower order parameter is close to zero and the
+However, forking is significant for the g$_1$ segment, whose lower order parameter is close to zero and the
 larger one has an absolute value of approximately 0.13--0.15. Forking was studied in detail by Gally et al.~\cite{gally81}, who used E. Coli to 
@@
 extract. This experiment gave the lower order parameter when deuterium was in the S position of g$_1$ or R position for g$_3$.
-Since the glycerol backbone order parameters are very similar irrespective of the headgroup chemistry (PC, PE and PG) or lipid 
+Since the glycerol backbone order parameters are very similar irrespective of the headgroup chemistry (PC, PE, or PG) or lipid 
 environment~\cite{gally81}, it is reasonable to assume that the stereospecifity measured for the PE lipids
@@
 The most detailed experimentally available order parameter information for the glycerol backbone and choline 
-segments of POPC bilayer is collected by taking the absolute values from~\cite{ferreira13}, the signs from~\cite{hong95a,hong95b,gross97} 
-and the stereospecific labeling from~\cite{gally81}, and is shown in Fig.~\ref{HGorderparameters2}.
+segments of POPC bilayer is collected by taking the absolute values from Ref.~\citenum{ferreira13}, the signs from Refs.~\citenum{hong95a,hong95b,gross97} 
+and the stereospecific labeling from Ref.~\citenum{gally81}, and is shown in Fig.~\ref{HGorderparameters2}.
 \begin{figure*}[]

### 2015-10-11 HGmodelMANUSCRIPT/REPLYtoREVIEWER2.txt
Update REPLYtoREVIEWER2.txt

--- 2015-10-11 HGmodelMANUSCRIPT/REPLYtoREVIEWER2.txt
@@
 Author reply:
-The subjective sweet spots span +-0.02 AROUND WHAT? IS THIS THE AVERAGE
-OF EXPERIMENTAL VALUE?
+
+All the "subjective sweet spots" in Fig. 2 have exactly the same size: 0.04 units.
+
+This size was agreed on based on the high experimental repeatibility of the order
+parameters that we discuss in the text (p12): "Almost all the reported values are
+within a variation of ±0.02 [...] for all fully hydrated PC bilayers, regardless
+of variation in their acyl chain composition and temperature."
+
+To make the (equal) size of the sweet spots clear to the reader, we now explicitly
+point out this size when introcuding them in the text (p15):
+
+"Motivated by the high experimental reproducibility, we have highlighted in Fig. 2
+subjective sweet spots (light blue areas **spanning 0.04 units**), within which
+we expect the calculated values of the order parameters of a well-performing
+force field to fall."
+
+The added words are those between the stars (**).
 

### 2015-10-11 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Tuned language.

@@
 MacRog~\cite{maciejewski14} force fields perform better for the glycerol backbone and headgroup structures than the GROMOS based models~\cite{castro08,kapla12,poger12,ferreira13}.
-However, none of the studies exploits the full potential of the available experimental data discussed in previous section, i.e. the quantitative accuracy, known signs and stereospecific labeling of
+However, none of the studies exploits the full potential of the available experimental data discussed in the previous section, i.e.
+the quantitative accuracy, known signs and stereospecific labeling of
 the experimental order parameters.
@@
 plotted the results together with experimental values in Fig.~\ref{HGorderparameters}.
-Two criteria were used to judge the quality of the model: there must not be significant  {\bf forking} in the $\alpha$ and $\beta$ carbons,
-there must be only moderate forking in the g$_3$ carbon and there must be significant forking in the g$_1$ carbon, the {\bf magnitude}
+Two criteria were used to judge the quality of the model: 1) there must not be significant  {\bf forking} in the $\alpha$ and $\beta$ carbons,
+there must be only moderate forking in the g$_3$ carbon and there must be significant forking in the g$_1$ carbon, and 2) the {\bf magnitude}
 should be preferably inside to the subjective sweet spots determined from experiments (blue shaded regions in Fig.~\ref{HGorderparameters}).
@@
 None of the studied force fields fulfils these criteria completely, however CHARMM36 is close. 
-This is not surprising since the dihedral potentials in this model are tuned to reproduce these 
-parameters better against experiments~\cite{klauda10}.
-The next models in the list are CHARMM36-UA~\cite{henin08,lee14} and H\"ogberg08~\cite{hogberg08} which is also not surprising since
-these models are using CHARMM bonded potentials for glycerol backbone and choline. The fourth and the fifth models in the list, MacRog~\cite{maciejewski14} and
+This is not surprising since the dihedral potentials in this model are tuned against experiments to better reproduce these 
+order parameters~\cite{klauda10}.
+The next models in the list are CHARMM36-UA~\cite{henin08,lee14} and H\"ogberg08~\cite{hogberg08}, which is also not surprising since
+these models are using the CHARMM bonded potentials for glycerol backbone and choline. The fourth and the fifth models in the list, MacRog~\cite{maciejewski14} and
 GAFFlipid~\cite{dickson12}, have independently determined dihedral potentials. All the models based on Gromos potentials and Slipids perform less well.
-In the present work we subject the CHARMM36, MacRog, GAFFlipid and Berger-POPC-07 to a more careful comparison including the sterospecific labeling  
-(Fig.~\ref{HGorderparameters2}), and atomistic level structure and responses to the dehydration and cholesterol content in the following sections.
-These models are selected for more detailed studies since they are the best representatives of different dihedral potential parametrization techniques 
+In the following sections we subject CHARMM36, MacRog, GAFFlipid and Berger-POPC-07 to a more careful comparison including the stereospecific labeling  
+(Fig.~\ref{HGorderparameters2}), atomistic level structure, and responses to dehydration and cholesterol content.
+These models are selected for more detailed study since they are the best representatives of different dihedral potential parametrization techniques 
 (CHARMM36, MacRog, GAFFlipid), and the Berger based models are the most used lipid models in the literature.

### 2015-10-11 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Tuned language.

@@
 structure is not unique. In this section we demonstrate that 1) the differences in order parameters
-indicate significantly different structural sampling strongly correlating with the dihedral angles of the related bonds,
+indicate significantly different structural sampling, which is strongly correlated with the dihedral angles of the related bonds;
 and that 2) the comparison between experimental and simulated order parameters can be used to exclude
-nonrealistic structural samping in molecular dynamics simulations. The demonstration is done for 
+nonrealistic structural sampling in molecular dynamics simulations. The demonstration is done for 
 the dihedral angles defined by the g$_3$-g$_2$-g$_1$-O(\textit{sn}-1) segments in the glycerol backbone and 
@@
 We note that performing a similar comparison through all the dihedrals in all the 13 models would probably give highly useful
-information on how to improve the accuracy of the models yet this is beyond the scope of the current report.
+information on how to improve the accuracy of the models; yet this is beyond the scope of the current report.
 
@@
 for the g$_1$ segment in the Berger-POPC-07 model (Fig.~\ref{HGorderparameters2}). 
-In conclusion, models preferring the anti conformation for this dihedral give more realistic order parameters and
+In conclusion, models preferring the anti conformation for this dihedral give more realistic order parameters;
 this is in agreement with previous crystal structure and $^1$H NMR studies~\cite{hauser80,hauser81,hauser81b,hauser88,pascher92,marsh06}.
@@
 shown in Fig.~\ref{dihDISTS2}. Also for this dihedral there are significant differences in the gauche--anti fractions.
-The gauche conformations are dominant in the CHARMM36, in MacRog there are only anti conformations present,
-and in the Berger-POPC-07 and GAFFlipid gauche and anti conformations have equal probabilities. 
+The gauche conformations are dominant in CHARMM36, in MacRog there are only anti conformations present,
+and in Berger-POPC-07 and GAFFlipid the gauche and anti conformations have equal probabilities. 
 On the other hand, comparison of $\alpha$ and $\beta$ order parameters in Fig.~\ref{HGorderparameters2}
@@
 sign (negative) for the $\beta$ order parameter. This result is again in agreement with previous 
-crystal structure, $^1$H NMR and Raman spectroscopy studies~\cite{hauser80,hauser81,hauser81b,akutsu81b} which suggest that
+crystal structure, $^1$H NMR, and Raman spectroscopy studies~\cite{hauser80,hauser81,hauser81b,akutsu81b}, which suggest that
 this dihedral is in the gauche conformation in the absence of ions.
@@
 
-The used examples show that the glycerol backbone and headgroup order parameters reflect the atomistic resolution structure
+These examples show that the glycerol backbone and headgroup order parameters reflect the atomistic resolution structure
 and that the comparison with experiments allows the assessment of the quality of the suggested structure. We were able to pinpoint
 specific problems in the structures in different models and suggest potential improvement strategies.
-If the improved atomistic
-molecular dynamics simulation model reproduced the order parameters and other experimental observables 
+If an improved atomistic
+molecular dynamics simulation model would reproduce the order parameters and other experimental observables 
 (like $^{31}$P chemical shift anisotropy~\cite{chowdhary13} and $^{31}$P-$^{13}$C dipolar couplings~\cite{prakash10})
-with experimental accuracy, it would give an interpretation for the atomistic resolution structure of the glycerol backbone and 
+within experimental accuracy, it would give an interpretation for the atomistic resolution structure of the glycerol backbone and 
 choline~\cite{seelig77b,skarjune79,jacobs80,davis83,akutsu91,hong95b,semchyschyn04}. The research along these lines is left, however,

### 2015-10-11 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Tuned language.

@@
 even very small order parameter changes resulting from the varying conditions can be measured 
-(see Ref.~\citenum{accuracyPOST} for more discussion.) 
-However, as already discussed above, it is not simple to deduce 
-the structural changes from order parameter changes~\cite{akutsu91,semchyschyn04}. Consequently, comparison of the order parameters
+(see Ref.~\citenum{accuracyPOST} for more discussion)---but,
+as already discussed above, it is not simple to deduce 
+the structural changes from order parameter changes~\cite{akutsu91,semchyschyn04}. However, comparison of the order parameters
 between simulations and experiments in different conditions can be used to assess the quality of the force field 
@@
     Note that to elucidate the relative change as a function of hydration level, the simulation results were vertically shifted; 
-    the shift magnitudes for each of the force fields are listed (SCH+shift) in the y-label.
+    the shift magnitudes for each of the force fields are listed (S$_\mathrm{CH}$+shift) in the y-label.
   }
@@
     with the Berger-POPC-07/H\"oltje-CHOL-13, CHARMM36 and MacRog force fields. The signs in the experimental values are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
-    and Gross et al.~\cite{gross97}.  Most order parameters from Berger-POPC-07/H\"oltje-CHOL-13 model for g$_1$ are beoynd the y-axis scale.
+    and Gross et al.~\cite{gross97}.  %Most order parameters from Berger-POPC-07/H\"oltje-CHOL-13 for g$_1$ are beyond the y-axis scale.
     In order to elucidate the relative change as a function of cholesterol content, the simulation results were vertically shifted; 
-    the shift magnitudes for each of the force fields are listed (SCH+shift) in the y-label.
+    the shift magnitudes for each of the force fields are listed (S$_\mathrm{CH}$+shift) in the y-label.
 }
@@
 When interpreted using the computational lipid model that we found to have the most realistic response to cholesterol,
-this observation means that cholesterol induces only minor changes in (the g$_3$-g$_2$-g$_1$-O(sn-1)
+this observation means that cholesterol induces only minor changes in the g$_3$-g$_2$-g$_1$-O(sn-1)
 dihedral of the glycerol backbone, in other words, there is no major conformational change of the lipid.
@@
 
-We acknowledge all the discussion participants at nmrlipids.blogspot.fi.
-AB and CL acknowledges financial support from the French National Research Agency
-(ANR: “Biolubrication by phospholipid membranes” Bioub2012) and
+We acknowledge all the discussion participants at \url{nmrlipids.blogspot.fi}.
+AB and CL acknowledge financial support from the French National Research Agency
+(ANR: Biolubrication by phospholipid membranes, Biolub2012) and
 computing time allocation from P{\^o}le Scientifique de Mod{\'e}lisation Num{\'e}rique from the ENS Lyon (PSMN),
@@
 for financial support, Cluster H\'ibrido de Superc\'omputo Xiuhcoatl - CINVESTAV and Miztli - UNAM for computational resources. 
-MJ and JT acknowledge and CSC - IT Center for Science for computational resources (project number tty3979).
+MJ and JT acknowledge and CSC -- IT Center for Science for computational resources (project number tty3979).
 MJ also acknowledges the Finnish Doctoral Programme in Computational Sciences (FICS) for funding.

### 2015-10-12 HGmodelMANUSCRIPT/REPLYtoREVIEWER2.txt
Update text and reply related to the issue https://github.com/NMRLipids/nmrlipids.blogspot.fi/issues/66

@@
+The original reviewer comments are not written in this publicly available
+file we do not have permission to publish those. The original comments
+will be added to the version which will be submitted to the JPCB.
+
 Reviewer 2.
@@
 
-To make the (equal) size of the sweet spots clear to the reader, we now explicitly
-point out this size when introcuding them in the text (p15):
+To make the definition of the sweet spots clear to the reader, we now explicitly
+point this out when introcuding them in the text (p15):
 
 "Motivated by the high experimental reproducibility, we have highlighted in Fig. 2
-subjective sweet spots (light blue areas **spanning 0.04 units**), within which
+subjective sweet spots (light blue areas **spanning 0.04 units around the average
+of the extremal experimental values**), within which
 we expect the calculated values of the order parameters of a well-performing

### 2015-10-12 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Update text and reply related to the issue https://github.com/NMRLipids/nmrlipids.blogspot.fi/issues/66

@@
 Motivated by the high experimental reproducibility, we have highlighted in 
-Fig.~\ref{HGorderparameters} subjective sweet spots (light blue areas spanning 0.04 units), within which we expect the calculated  
+Fig.~\ref{HGorderparameters} subjective sweet spots (light blue areas spanning 0.04 units around the average
+of the extremal experimental values), within which we expect the calculated  
 values of the order parameters of a well-performing force field to fall.

### 2015-10-13 HGmodelMANUSCRIPT/REPLYtoREVIEWER2.txt
Update REPLYtoREVIEWER2.txt
Add reply to comment 1 by the reviewer

@@
 Author reply:
+In the publications mentioned by the reviewer it is shown that the
+the membrane properties are affected by interfacial tension rather than the ensemble. 
+This is also explicitly said in the abstract of
+J. Chem. Phys. 111, 1281-1287
+"Equivalence between the constant area and constant surface tension ensembles is
+investigated by comparing the present simulations with earlier work from our laboratories and we
+find simulation results to depend much more strongly on the specified surface area or surface tension
+than on the ensemble employed"
+It should be also noted that the total lengths of the simulations are around 1ns 
+in the mentioned publications while our simulations are 1-2 orders of magnitude longer. 
+Thus, even though the dependence on the ensemble would be observed in these short simulations
+it could very well be due to insufficient sampling.
+
+As we have already pointed out we have restricted our work to the available models and
+the improvement of these models has been left for further studies. Since all the developers
+of the tested models have used NPT ensemble in zero tension, there seems to be consensus on 
+this.
+
+It is also much more likely that the force field parameters affect the glycerol backbone and
+choline structures much more than the used ensemble or interfacial tension. We have 
+left the more detailed discussion on the force field parameters for further studies, 
+thus the same should be done for the discussion about ensemble and surface tension which
+is most likely less relevant.
 
@@
 Author reply:
-
 All the "subjective sweet spots" in Fig. 2 have exactly the same size: 0.04 units.
@@
 To make the definition of the sweet spots clear to the reader, we now explicitly
-point this out when introcuding them in the text (p15):
+point this out when introducing them in the text (p15):
 

### 2015-10-13 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Change to title case

@@
 \title[An \textsf{achemso} demo]
-  {Towards atomistic resolution structure of phosphatidylcholine headgroup and glycerol backbone at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}}
+  {Towards Atomistic Resolution Structure of Phosphatidylcholine Headgroup and Glycerol Backbone at Different Ambient Conditions\footnote{Publication about results presented in the NMRlipids project}}
 

### 2015-10-14 HGmodelMANUSCRIPT/REPLYtoREVIEWER2.txt
Update REPLYtoREVIEWER2.txt

--- 2015-10-12 HGmodelMANUSCRIPT/REPLYtoREVIEWER2.txt
@@
 Author reply:
+In the publications mentioned by the reviewer it is shown that the
+the membrane properties are affected by the interfacial tension rather than the ensemble. 
+This is also explicitly said in the abstract of
+J. Chem. Phys. 111, 1281-1287:
+"Equivalence between the constant area and constant surface tension ensembles is
+investigated by comparing the present simulations with earlier work from our laboratories and we
+find simulation results to depend much more strongly on the specified surface area or surface tension
+than on the ensemble employed."
+
+In addition, it should be noted that in the mentioned publications the total lengths of the simulations are around 1ns,
+while our simulations are 1-2 orders of magnitude longer. 
+Thus, even if the dependence on the ensemble would be observed in these short simulations
+it would likely be due to insufficient sampling.
+
+As we have already pointed out, we have restricted our work to the available models and
+the improvement of these models has been left for further studies. Since all the developers
+of the tested models have used NPT ensemble in zero tension, there seems to be consensus on 
+this.
+
+It is also much more likely that the force field parameters affect the glycerol backbone and
+choline structures much more than the used ensemble or interfacial tension. As we have 
+left the more detailed discussion of the force field parameters for further studies, 
+the same should be done for the (most likely less relevant) discussion of the ensemble and surface tension.
 

### 2015-10-14 HGmodelMANUSCRIPT/REPLYtoREVIEWER2.txt
Update REPLYtoREVIEWER2.txt

@@
 phosphatidylcholine headgroup and glycerol
-backbone at different ambient conditions"
+backbone at different ambient conditions", and
 in the introduction we write:
@@
 This is widely used naming convention for phospholipid molecules.
-According to the title and the naming convetion used in the manuscript, 
-the beginning of acyl chains are not the part of headgroup and are 
-beoynd the scope defined in the title. Their presence at the interface
+According to the title and the naming convention used in the manuscript, 
+the beginnings of the acyl chains are not the part of headgroup and are 
+beyond the scope defined in the title. Their presence at the interface
 or their chemical environment do not change this.
@@
 We are not willing to extend the scope of the manuscript since it
-would signifincantly increase already massive information content and 
-length. Instead, the acyl chains order parameters will be discussed in 
+would significantly increase the already massive information content and 
+length. Instead, the acyl chain order parameters will be discussed in 
 the context of the NMRLipids III project, see 

### 2015-10-14 scripts/POPCcharmm/calcPNangle_CHARMM36_40eperl.sh
Add data files asked in issue https://github.com/NMRLipids/nmrlipids.blogspot.fi/issues/55

@@ -1,0 +1,24 @@
+#!/bin/bash
+tmpDIRname=CHARMM36_40wPERl_PNANGLEtmp
+mkdir $tmpDIRname
+cd $tmpDIRname
+cp /home/ollilas1/Dropbox/codes/gro_PNangle.awk ./
+trajname=/wrk/ollilas1/POPCdata/POPCcharmm/40wPERpc/POPC_0CHL.xtc
+tprname=/wrk/ollilas1/POPCdata/POPCcharmm/40wPERpc/POPC_0CHL.tpr
+mappingFILE=../../../../NmrLipidsCholXray/MAPPING/mappingPOPCcharmm.txt
+outFILE=../../../DATAreportediINblog/POPC/dehydration/PN-vector/CHARMM36-303K-40wPERl.dat
+groFILE=/wrk/ollilas1/POPCdata/POPCcharmm/40wPERpc/POPC_0CHL.gro
+#trjcat -f popcCHOL50molPER0-25ns.trr popcCHOL50molPER25-50ns.trr
+LIPIDname=$(grep M_POPC_M $mappingFILE | awk '{printf "%5s\n",$2}')
+LIPIDindexNR=$(echo q | make_ndx -f $tprname | grep $LIPIDname | awk '{if(NR==1)print $1}')
+Pname=$(grep M_G3P2_M $mappingFILE | awk '{printf "%5s\n",$2}')
+Nname=$(grep M_G3N6_M $mappingFILE | awk '{printf "%5s\n",$2}')
+echo $LIPIDindexNR System | trjconv -f $trajname -s $tprname -fit progressive -o FITtraj.xtc
+#Ztrans=$(tail -n 1 $groFILE | awk '{print $3/2}')
+#echo System | trjconv -f FITtraj.xtc -s $tprname -trans 0 0 $Ztrans -o ANALtraj.xtc
+echo System | trjconv -f FITtraj.xtc -s $tprname -o INBOXtraj.gro -pbc mol
+echo $LIPIDindexNR | g_traj -f INBOXtraj.gro -s $tprname -ox com.xvg -com -xvg none
+Zcom=$(cat com.xvg | awk '{sumZ=sumZ+$4; sum=sum+1}END{print sumZ/sum}')
+awk -v Cname="$Nname" -v Hname="$Pname" -v comz=$Zcom -f gro_PNangle.awk INBOXtraj.gro > $outFILE
+rm $tmpDIRname/*
+rmdir $tmpDIRname

### 2015-10-14 scripts/POPCmacrog/calcPNangle_MACROG_44eperl.sh

@@ -1,0 +1,24 @@
+#!/bin/bash
+tmpDIRname=MACROG_44wPERl_PNANGLEtmp
+mkdir $tmpDIRname
+cd $tmpDIRname
+cp /home/ollilas1/Dropbox/codes/gro_PNangle.awk ./
+trajname=/wrk/ollilas1/POPCdata/POPCmacrog/30_110ns_100ps.xtc
+tprname=/wrk/ollilas1/POPCdata/POPCmacrog/run.tpr
+mappingFILE=../../../../NmrLipidsCholXray/MAPPING/mappingPOPCmacrog.txt
+outFILE=../../../DATAreportediINblog/POPC/dehydration/PN-vector/MACROG-310K-44wPERl.dat
+#groFILE=/wrk/ollilas1/POPCdata/POPCcharmm/40wPERpc/POPC_0CHL.gro
+#trjcat -f popcCHOL50molPER0-25ns.trr popcCHOL50molPER25-50ns.trr
+LIPIDname=$(grep M_POPC_M $mappingFILE | awk '{printf "%5s\n",$2}')
+LIPIDindexNR=$(echo q | /home/ollilas1/gromacs/gromacs465/bin/make_ndx -f $tprname | grep $LIPIDname | awk '{if(NR==1)print $1}')
+Pname=$(grep M_G3P2_M $mappingFILE | awk '{printf "%5s\n",$2}')
+Nname=$(grep M_G3N6_M $mappingFILE | awk '{printf "%5s\n",$2}')
+echo $LIPIDindexNR System | /home/ollilas1/gromacs/gromacs465/bin/trjconv -f $trajname -s $tprname -fit progressive -o FITtraj.xtc
+#Ztrans=$(tail -n 1 $groFILE | awk '{print $3/2}')
+#echo System | trjconv -f FITtraj.xtc -s $tprname -trans 0 0 $Ztrans -o ANALtraj.xtc
+echo System | /home/ollilas1/gromacs/gromacs465/bin/trjconv -f FITtraj.xtc -s $tprname -o INBOXtraj.gro -pbc mol
+echo $LIPIDindexNR | /home/ollilas1/gromacs/gromacs465/bin/g_traj -f INBOXtraj.gro -s $tprname -ox com.xvg -com -xvg none
+Zcom=$(cat com.xvg | awk '{sumZ=sumZ+$4; sum=sum+1}END{print sumZ/sum}')
[…]

### 2015-10-14 HGmodelMANUSCRIPT/REPLYtoREVIEWER2.txt
Update REPLYtoREVIEWER2.txt

@@
 According to the title and the naming convention used in the manuscript, 
-the beginnings of the acyl chains are not the part of headgroup and are 
+the beginnings of the acyl chains are not part of the headgroup and are 
 beyond the scope defined in the title. Their presence at the interface

### 2015-10-26 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Small updates to the cholesterol and dehydration figure
Add keywords
Update references and fix a typo from https://github.com/NMRLipids/nmrlipids.blogspot.fi/issues/68

@@
 \abbreviations{IR,NMR,UV}
-\keywords{American Chemical Society, \LaTeX}
+\keywords{lipid, bilayer, molecular dynamics simulation, nuclear magnetic resonance, order parameter}
 
@@
 The absolute values of order parameters from the effective dipolar coupling $d_\mathrm{CH}$ ($^1$H-$^{13}$C NMR) are calculated using the equation
-$|S_{{\rm CH}}|=\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c} d_\mathrm{CH}$, where
-values between 20.2--22.7~kHz are used for $\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c}$,
+$|S_{{\rm CH}}|=2\pi\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c} d_\mathrm{CH}$, where
+values between 20.2--22.7~kHz are used for $(2\pi\frac{4\pi\langle r_\mathrm{CH}^3 \rangle}{\hbar \mu_0 \gamma_h \gamma_c})^{-1}$,
 depending on the original authors~\cite{hong95a,gross97,dvinskikh05a,ferreira13}.
@@
   \centering
-  \includegraphics[width=7.3cm]{../DATAreportediINblog/cholesterolization_lp.pdf}
+  \includegraphics[width=7.3cm]{../DATAreportediINblog/cholesterolization.pdf}
 %    \includegraphics[width=7.3cm]{../DATAreportediINblog/cholesterolization_smooth.pdf}

### 2015-11-23 HGmodelMANUSCRIPT/HGmodel_ACSsupplementary.tex
Compile pdf from changes in commits https://github.com/NMRLipids/nmrlipids.blogspot.fi/commit/a3bc5f847a9a6818302cd8919b06ba6aab8d7542 https://github.com/NMRLipids/nmrlipids.blogspot.fi/commit/c79f4092bfa48f8b9c046fa7a507b513e4b0ff12 https://github.com/NMRLipids/nmrlipids.blogspot.fi/commit/749c796304a5858b555b184e6e67cdfd9b763238 https://github.com/NMRLipids/nmrlipids.blogspot.fi/commit/467d3870a2569038ebf7dfef8e29344ddfbbead0
Proof reading.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
-\usepackage[obeyFinal]{easy-todo}
+%\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
+%\usepackage[obeyFinal]{easy-todo}
 \usepackage{rotating}
 \usepackage{epstopdf}
+\usepackage{setspace}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\author{Alexandru Botan}
+\author{\setstretch{1.05} Alexandru Botan}
 %\affiliation{The authors are listed in alphaphetical order.}
 %\affiliation{The author list is not completed.}
-\affiliation[Lyon CNRS]{Institut Lumi\`ere Mati\`ere, UMR5306 Universit\'e Lyon 1-CNRS, Universit\'e de Lyon 69622 Villeurbanne, France}
+\affiliation[Lyon CNRS]{\small\setstretch{1.0} Institut Lumi\`ere Mati\`ere, UMR5306 Universit\'e Lyon 1-CNRS, Universit\'e de Lyon 69622 Villeurbanne, France}
 %
@@
 \author{Matej Kandu\v{c}}
-\affiliation[Freie Universit\"{a}t Berlin] 
-{Fachbereich Physik, Freie Universitat Berlin, Berlin, Germany}
+\affiliation[Freie Universitat Berlin] 
+{Fachbereich Physik, Freie Universit{\"a}t Berlin, Berlin, Germany}
 \author{Waldemar Kulig}
@@
 \author{Markus S. Miettinen}
-\affiliation[Freie Universit\"{a}t Berlin] 
-{Fachbereich Physik, Freie Universitat Berlin, Berlin, Germany}
+\affiliation[Freie Universitat Berlin] 
+{Fachbereich Physik, Freie Universit{\"a}t Berlin, Berlin, Germany}
 \author{Luca Monticelli}
@@
 \title[An \textsf{achemso} demo]
-  {SUPPORTING INFORMATION: Towards atomistic resolution structure of phosphatidylcholine headgroup and glycerol backbone at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}}
+  {\setstretch{1.05} SUPPORTING INFORMATION:\\
+  Towards atomistic resolution structure of phosphatidylcholine headgroup and glycerol backbone at different ambient conditions\footnote{Publication about results presented in the NMRlipids project (\url{nmrlipids.blogspot.fi})}}
 
@@
 %\end{center}
+\setstretch{1.025}
 \section{Simulation details} 
-\subsection{Berger based models}
-For the Berger based models we use here the following naming convention: 
+\subsection{Berger-based models}
+For the Berger-based models we used the following naming convention: 
 Berger - \{{\it molecule name}\} - \{{\it year when model published first time}\} \{{\it citation}\}.
-The reason is that there are several different molecular topologies which are using the non-bonded parameters originally
-developed by Berger et al.~\cite{berger97}. Thus the common factor in the Berger based models are the non-bonded parameters,
-while the molecule specific parameters might somewhat vary. However, the majority of the molecular level topologies are 
-relying (especially for the glycerol backbone and headgroup) on the parameters originally introduced by Marrink et al.~\cite{marrink98}.
-This is the case for all the Berger based simulations discussed in this work.
+The reason was that there are several different molecular topologies %which are
+using the non-bonded parameters originally
+developed by Berger et al.~\cite{berger97}. Thus the common factor in the Berger-based models are the non-bonded parameters,
+while the molecule-specific parameters might somewhat vary. However, the majority of the molecular-level topologies  
+rely (especially for the glycerol backbone and headgroup) on the parameters originally introduced by Marrink et al.~\cite{marrink98};
+this is the case for all the Berger-based simulations discussed in this work.
 
-POPC simulations at full hydration at 298~K and simulations studying the effect of cholesterol are the same as in previous publications~\cite{ferreira13,ferreira15}.
-In these simulation the POPC parameters introduced by Ollila et al~\cite{ollila07a} are used, which are using the non-bonded parameters of Berger~\cite{berger97}
-and a molecular topology from Tieleman et al.~\cite{tieleman99} with improved double bond dihedrals by Bachar et al.~\cite{bachar04}. 
-Thus they are called Berger-POPC-07~\cite{ollila07a}. The cholesterol model is based on the parameters by H\"oltje et al.~\cite{holtje01} with the
-exception that the atom types were changed from CH2/CH3 to LP2/LP3 to avoid overcondensation of the bilayer as suggested in ref.~\cite{tieleman06}.
-Since this modification was introduced by Ferreira et al.~\cite{ferreira13}, we call the used cholesterol model as H\"oltje-CHOL-13~\cite{ferreira13}.
+POPC simulations at full hydration at 298~K, and simulations studying the effect of cholesterol, were the same as in previous publications~\cite{ferreira13,ferreira15}.
+In these simulations, the POPC parameters introduced in 2007 by Ollila et al.~\cite{ollila07a} were used (the non-bonded parameters of Berger et al.~\cite{berger97}
+and a molecular topology from Tieleman et al.~\cite{tieleman99} with improved double bond dihedrals by Bachar et al.~\cite{bachar04}), thus
+they were called Berger-POPC-07~\cite{ollila07a}. The cholesterol model was based on the parameters by H\"oltje et al.~\cite{holtje01} with the
+exception that the atom types were changed from CH2/CH3 to LP2/LP3 to avoid overcondensation of the bilayer as suggested in Ref.~\citenum{tieleman06}.
+Since this modification was introduced in 2013 by Ferreira et al.~\cite{ferreira13}, we call this cholesterol model H\"oltje-CHOL-13~\cite{ferreira13}.
 
-For the POPC at 323~K and POPC in low hydration the same force field parameters are used.
-For DPPC the implementation of Berger parameters~\cite{berger97} by Peter Tieleman et al. are used~\cite{marrink98}.
-For all of these simulations a timestep of 2~fs was used with a leapfrog integrator. Covalent bond lengths were constrained with the LINCS algorithm~\cite{hess97,hess07}. 
-Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with real space cut-off of 1.0~nm was used 
-for electrostatics. Plain cut-off was used for the Lennard-Jones interactions with a 1.0~nm cut-off.
-The neighbor lists with cut-off of 1.0~nm were updated every 5 steps. Temperature was coupled separately
+For the POPC at 323~K and POPC at low hydration, the same force field parameters were used.
+For DPPC, the implementation of Berger parameters~\cite{berger97} by Peter Tieleman et al. was used~\cite{marrink98}.
+For all  these simulations, a timestep of 2~fs was used with a leapfrog integrator. Covalent bond lengths were constrained with the LINCS algorithm~\cite{hess97,hess07}. 
+Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with a real space cut-off of 1.0~nm was used 
+for electrostatics. Plain cut-off was used for the Lennard-Jones interactions, with a 1.0~nm cut-off.
+The neighbor lists with a cut-off of 1.0~nm were updated every 5 steps. Temperature was coupled separately
 for lipids and water to 298~K using the velocity-rescale method~\cite{bussi07} with coupling constant 0.1~ps.
@@
 (\url{http://mackerell.umaryland.edu/CHARMM_ff_params.html}) were used.
-Timestep of 1~fs was used with the leapfrog integrator. Covalent bonds with hydrogens were constrained with LINCS algorithm~\cite{hess97,hess07}. 
+Timestep of 1~fs was used with the leapfrog integrator. Covalent bonds with hydrogens were constrained with the LINCS algorithm~\cite{hess97,hess07}. 
 Coordinates were written every 5~ps. PME~\cite{darden93,essman95} with real space cut-off of 1.4~nm was used 
 for electrostatics. Lennard-Jones interactions were switched to zero between 0.8~nm and 1.2~nm.
-The neighbour lists with a cut-off of 1.4~nm were updated every 5 steps. Temperature was coupled separately
-for lipids and water to 303~K using the velocity-rescale method~\cite{bussi07} with coupling constant of 0.2~ps.
+The neighbor lists with a cut-off of 1.4~nm were updated every 5 steps. Temperature was coupled separately
+for lipids and water to 303~K using the velocity-rescale method~\cite{bussi07} with a coupling constant of 0.2~ps.
 Pressure was semi-isotropically coupled to the atmospheric pressure with the Berendsen method~\cite{berendsen84}.
 
-{\it POPC with 128 lipids}
-The starting structures for the pure POPC simulations was taken from the Slipids~\cite{jambeck12b} website (\url{http://mmkluster.fos.su.se/slipids/Downloads.html}).
-The starting structures for mixed POPC/Cholesterol simulations were constructed with the CHARMM-GUI website~\cite{jo08}. 
-They contained 100 POPC/24 cholesterol molecules and 80 POPC/80 cholesterol molecules for
-the simulations of 20\% cholesterol and 50\% cholesterol respectively. The TIP3P water model~\cite{jorgensen83} was used to
+{\it POPC with 128 lipids.}
+The starting structures for the pure POPC simulation was taken from the Slipids~\cite{jambeck12b} website (\url{http://mmkluster.fos.su.se/slipids/Downloads.html}).
+The starting structures for the mixed POPC/Cholesterol simulations were constructed with the CHARMM-GUI website~\cite{jo08}. 
+They contained 100 POPC~/~24 cholesterol molecules and 80 POPC~/~80 cholesterol molecules for
+the simulations of 20\% cholesterol and 50\% cholesterol, respectively. The TIP3P water model~\cite{jorgensen83} was used to
 solvate the system.
@@
 by Piggot et al. \cite{piggot12} were used. Cholesterol parameters came
-from Lim et al. \cite{lim12} and were converted into GROMACS format with the PyTopol tool~\cite{salari15}.  
-Single point energy calculation was done to assess the conversion. 
+from Lim et al. \cite{lim12}, and were converted into GROMACS format with the PyTopol tool~\cite{salari15}.  
+Single-point energy calculation was done to assess the conversion. 
 %(At the time I've made the simulations, the charmm36 ff for cholesterol in gromacs format was not available at
 %http://mackerell.umaryland.edu/charmm_ff.shtml#gromacs).
-Simulations were performed for 200~ns and the last 100~ns was used for the calculations. Timestep of 2~fs was
+Simulations were performed for 200~ns and the last 100~ns was used for the analysis. Timestep of 2~fs was
 used with leapfrog integrator. All bond lengths were constrained with LINCS~\cite{hess97,hess07}. Temperature was maintened at
-303~K with the velocity-rescale method~\cite{bussi07} and a time constant of 0.2~ps. Pressure was maintained semiisotropically
-at 1~bar using the Parrinello--Rahman algorithm~\cite{parrinello81} with a time constant of 1.0~ps. The neighbour list with a cut-off of 1.2~nm
+303~K using the velocity-rescale method~\cite{bussi07} with a time constant of 0.2~ps. Pressure was maintained semi-isotropically
+at 1~bar using the Parrinello--Rahman algorithm~\cite{parrinello81} with a time constant of 1.0~ps. The neighbor list with a cut-off of 1.2~nm
 was updated every 10 steps. Lennard-Jones interactions were switched to zero
-between 0.8~nm and 1.2~nm. PME~\cite{darden93,essman95} with real space cut-off of 1.2~nm was used for electrostatics.
+between 0.8~nm and 1.2~nm. PME~\cite{darden93,essman95} with a real space cut-off of 1.2~nm was used for electrostatics.
 
@@
 %
-POPC The starting structures for simulations were constructed with
+The starting structures for simulations were constructed with
 the CHARMM-GUI website~\cite{jo08}. 512 POPC lipids (256 per leafleat) were 
-used for the initial 0\% concentration and they were subsequently 
-substituted by cholesterol molecules to fulfil the desired concentration.
+used for the initial 0\% cholesterol content, and they were subsequently 
+substituted by cholesterol molecules to obtain the desired content.
 The TIP3P~\cite{jorgensen83} water model was used to solvate the system. 
-The publicly available port of the CHARMM36 forcefield parameters (\url{mackerell.umaryland.edu/CHARMM_ff_params.html}) 
-was used for both POPC and cholesterol. 
-Simulations were performed for 170 ns and the last 100 ns was used for the calculations.
-Timestep of 1 fs was used with the leapfrog integrator. Covalent bonds with hydrogen 
-were constrained with LINCS algorithm. Coordinates were written every 10 ps.
-PME with real space cut-off of 1.2 nm was used for electrostatics. Lennard-Jones 
-interactions were switched to zero between 0.8 nm and 1.2 nm. The neighbour lists with a
-cut-off of 1.2 nm were updated every 5 steps. Temperature was coupled separately for cholesterol, lipids
-and water to 298 K using the velocity-rescale method with coupling constant of 0.2 ps.
+The publicly available port of the CHARMM36~\cite{klauda10} forcefield parameters (\url{mackerell.umaryland.edu/CHARMM_ff_params.html}) 
+was used for both the POPC and cholesterol. 
+Simulations were performed for 170~ns, and the last 100~ns was used for the analysis.
+Timestep of 1~fs was used with the leapfrog integrator. Covalent bonds with hydrogen 
+were constrained with the LINCS~\cite{hess97,hess07} algorithm. Coordinates were written every 10~ps.
+PME~\cite{darden93,essman95} with real space cut-off of 1.2~nm was used for electrostatics. Lennard-Jones 
+interactions were switched to zero between 0.8~nm and 1.2~nm. The neighbor lists with a
+cut-off of 1.2~nm were updated every 5 steps. Temperature was coupled separately for cholesterol, lipids
+and water to 298~K using the velocity-rescale method~\cite{bussi07} with coupling constant of 0.2~ps.
 Pressure was semi-isotropically coupled to the atmospheric pressure with the Berendsen
-method.
+method~\cite{berendsen84}.
 
 \subsection{MacRog}
-The lipid force field parameters were obtained from the developers and the parameters are presented in Refs.~\citenum{maciejewski14,kulig15b}. 
+The lipid force field parameters were obtained from the developers; the parameters are presented in Refs.~\citenum{maciejewski14,kulig15b}. 
 %correspond to the published DPPC parameters~\cite{maciejewski14} with the inclusion of the double bond parameters~\cite{kulig15}. 
@@
 
-Identical parameters were employed for both full hydration and for the dehydration simulations. The dehydration simulations were also run for 100~ns 
+Identical parameters were employed for the full hydration and the dehydration simulations. The dehydration simulations were also run for 100~ns 
 with data saved every 10~ps.
 
-The initial structures for the simulations with 10, 40, 50 and 60 mol\% of cholesterol were obtained by replacing 14, 56, 64 or 72 POPC molecules 
+The initial structures for the simulations with 10, 40, 50 and 60 mol-\% of cholesterol were obtained by replacing 14, 56, 64 or 72 POPC molecules 
 with cholesterol molecules in the initial structure containing 128 POPC molecules. These systems were simulated for 400 ns and the first 200 ns was 
@@
 
-\subsection{GAFFLipid}
+\subsection{GAFFlipid}
 The initial structure in Lipidbook \cite{domanski10} had different glycerol backbone isomers in different leaflets. 
-To generate the initial structure we took the structure delivered by Slipids developers~\cite{jambeck12b}. Also this structure
+To generate the initial structure, we took the structure delivered by the Slipids developers~\cite{jambeck12b}. Also this structure
 had one lipid with different glycerol backbone isomer. This lipid and one lipid from opposite leaflet were removed
@@
 The force field parameters were generated using files obtained from the Lipidbook website (\url{http://lipidbook.bioch.ox.ac.uk/package/show/id/150.html})~\cite{domanski10}. 
-The conversion to GROMACS compatible formats was performed using the acpype tool~\cite{silva12}. The accuracy of the conversion was checked by calculating 
-the total energy of a single POPC lipid molecule using the sander program which is part of the AmberTools14 package~\cite{ferrer13} and version 4.6.5 of GROMACS. 
+The conversion to GROMACS compatible formats was performed using the acpype tool~\cite{silva12}. Accuracy of the conversion was checked by calculating 
+the total energy of a single POPC lipid molecule using the sander program, which is part of the AmberTools14 package~\cite{ferrer13}, and version 4.6.5 of GROMACS. 
 A difference of 0.002 kcal/mol was obtained between the two programs.
@@
 Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with a real space cut-off at 1.0~nm was used 
-for electrostatics. Plain cut-off with 1~nm was used for Lennard-Jones interactions. 
-The neighbour lists with a cut-off of 1.0~nm were updated every 5 steps. 
-Pressure was semi-isotropically coupled to a pressure of 1~bar with the Berendsen method~\cite{berendsen84}.
+for electrostatics. Plain cut-off at 1~nm was used for Lennard-Jones interactions. 
+The neighbor lists with a cut-off of 1.0~nm were updated every 5 steps. 
+Pressure was semi-isotropically coupled to 1~bar with the Berendsen method~\cite{berendsen84}.
 
 It should be noted that the area per molecule with these settings for the GAFFlipid model was 61.6~\AA$^2$,
-while the original publication reported 63.9~\AA$^2$~\cite{dickson12}. However, the same parameters and Amber to Gromacs
-conversion reproduced the area per molecule from original publication for the lipid14 model (see next section).
+while the original publication~\cite{dickson12} reported 63.9~\AA$^2$. Notably, however, the same parameters and Amber-to-Gromacs
+conversion procedure reproduced the area per molecule from the original publication~\cite{dickson14} for the Lipid14 model (see the Subsec.~Lipid14).
 
@@
 The initial structure was taken directly from the Lipidbook~\cite{domanski10}.
-The Amber compatible force field parameters were generated using the tleap program which is integrated in the AmberTools14 package~\cite{ferrer13}. 
-A workflow similar to the one used previously for the conversion and validation of the GAFFLipid parameters was followed here. 
-As before, a negligible energy difference of 0.003 kcal/mol was obtained between the two programs.
+The Amber-compatible force field parameters were generated using the tleap program, which is integrated in the AmberTools14 package~\cite{ferrer13}. 
+A workflow similar to the one used for the conversion and validation of the GAFFlipid parameters was followed;
+here a negligible energy difference of 0.003 kcal/mol was obtained between sander and GROMACS.
 
-Timestep of 2~fs was used in Langevin dynamics with zero friction term and collision frequency of 1.0~ps$^{-1}$. 
-Covalent bonds with hydrogens were constrained with LINCS algorithm~\cite{hess97,hess07}.
-Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with real space cut-off of 1.0~nm was used 
+Timestep of 2~fs was used in Langevin dynamics with zero friction term and a collision frequency of 1.0~ps$^{-1}$. 
+Covalent bonds with hydrogens were constrained with the LINCS algorithm~\cite{hess97,hess07}.
+Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with real space cut-off at 1.0~nm was used 
 for electrostatics. Plain cut-off with 1~nm was used for Lennard-Jones interactions. Dispersion correction
@@
 
-The area per molecule with these settings was 65.4~\AA$^2$ which is in agreement with the value reported in the original publication 65.6$\pm$0.5~\AA$^2$~\cite{dickson14}.
+The area per molecule with these settings was 65.4~\AA$^2$, which is in agreement with the value 65.6$\pm$0.5~\AA$^2$ reported in the original publication~\cite{dickson14}.
 
-\subsection{Poger et al.}
+\subsection{Poger}
 The Poger lipids are derived from GROMOS G53A6~\cite{poger10} and were initially coined 53A6-L (L for lipids). They are now part of GROMOS G54A7~\cite{poger12} 
@@
 the rotation of the 3 choline methyls) compared to the April 2012 version (3 dihedrals of type gd\_29 for the 3 choline methyls). This should not affect the 
-bilayer structure and only change the kinetics of the choline methyls rotation. However the October 2013 version has not been tested in this study.
+bilayer structure and only change the kinetics of the choline methyls rotation. However, the October 2013 version has not been tested in this study.
 
-MD Simulations (two repetitions with independent initial velocities) were run for 100~ns using a 2~fs time step and the analysis 
+Simulations (two repetitions with independent initial velocities) were run for 100~ns using a 2~fs time step and the analysis 
 was performed on the last 50~ns. Coordinates were saved every 50~ps for analysis. All bond lengths were constrained with the LINCS algorithm~\cite{hess97,hess07}. Temperature was kept 
-at 323~K employing the velocity-rescale~\cite{bussi07} thermostat with a time constant of 0.1~ps (DPPC and water coupled separetly). Pressure was maintained semi-isotropically at 1~bar using 
-the Parrinello--Rahman barostat~\cite{parrinello81} using a 4~ps time constant and a compressibility of 4.5e-5~bar$^{-1}$. For non-bonded interactions, two conditions were tested:
-i) A 0.8--1.4 nm twin-range cut-off with the neighbor list updated every 5 steps for both electrostatics and Lennard-Jones (LJ) interactions 
-(simulation files available at~\cite{pogerFILESrf1,pogerFILESrf2}). For the former the generalized reaction 
+at 323~K employing the velocity-rescale~\cite{bussi07} thermostat with a time constant of 0.1~ps (DPPC and water coupled separately). Pressure was maintained semi-isotropically at 1~bar using 
+the Parrinello--Rahman barostat~\cite{parrinello81} with a 4~ps time constant and a compressibility of 4.5e-5~bar$^{-1}$. For non-bonded interactions, two conditions were tested:
+\begin{description}
+\item[i)]
+{\noindent A 0.8--1.4 nm twin-range cut-off with the neighbor list updated every 5 steps for both electrostatics and Lennard-Jones (LJ) interactions 
+(simulation files available at Refs.~\citenum{pogerFILESrf1,pogerFILESrf2}). For the former, the generalized reaction 
 field (RF) with a dielectric permittivity of 62 was used beyond the 1.4~nm cut-off~\cite{tironi95}. This is the original setup that Poger et al.~\cite{poger10} used.
-ii) PME~\cite{darden93,essman95} electrostatics with a real space cut-off of 1.0~nm, a Fourier spacing of 0.12~nm and an interpolation order of 4, LJ interactions computed with a 1.0--1.4~nm twin-range cut-off,
-neighbor list updated every 5 steps (simulation files available at~\cite{pogerFILESpme1,pogerFILESpme2}). 
-Note that Poger and Mark tested the effect of PME vs RF in ref.~\cite{poger12}, but used a 1.0~nm cut-off with PME and 1.4~nm with RF for LJ
-interactions. Since 0.8--1.4~nm twin-range cut-off for LJ interations is used in the parametrization of GROMOS force field, we decided to use that
-also in the simulations with PME.
+}
+\item[ii)]
+{\noindent PME~\cite{darden93,essman95} electrostatics with a real space cut-off of 1.0~nm, a Fourier spacing of 0.12~nm and an interpolation order of 4, LJ interactions computed with a 1.0--1.4~nm twin-range cut-off,
+neighbor list updated every 5 steps (simulation files available at Refs.~\citenum{pogerFILESpme1,pogerFILESpme2}).
+}
+\end{description}
+Note that Poger and Mark tested the effect of PME vs RF in Ref.~\citenum{poger12}, but used for LJ interactions a 1.0~nm cut-off with PME and 1.4~nm with RF.
+Since 0.8--1.4~nm twin-range cut-off for LJ interactions is used in the parametrization of the GROMOS force field, we decided to use that also in the simulations with PME.
 
@@
 GROMACS versions 4.0.X and earlier (the original authors~\cite{poger10} used GROMACS version 3.3.3). When switching to versions 4.5.X and above, the area per lipid dropped to below 0.58~nm$^2$. 
-The GROMACS developers were contacted and a redmine issue opened (\url{http://redmine.gromacs.org/issues/1400}). The difference comes from the new Trotter decomposition 
+The GROMACS developers were contacted and a redmine issue opened (\url{http://redmine.gromacs.org/issues/1400}). Outcome: the difference comes from the new Trotter decomposition 
 introduced in versions 4.5.X. A fix has been introduced in version 4.6.6 that allows a recovery of an area per lipid value of 0.615~nm$^2$. The results in terms of area per lipid using the different 
-GROMACS versions are available at~\cite{pogerFILESrf2}.
-Thus we decided to use only the PME setup ii) for computing the order parameter since it gives stable results regardless of the GROMACS version. We obtained an area per 
-lipid of 0.615~nm$^2$, below 0.648 nm$^2$ found by the original authors with their PME setup (see~\cite{poger12}). We explained that by the fact that we used 
-a 1.4~nm for the LJ cut-off whereas a value of 1.0~nm was used in the original publication.
+GROMACS versions are available at Ref.~\citenum{pogerFILESrf2}.
+
+We decided to use only the PME setup ii) for computing the order parameters, since it gives stable results regardless of the GROMACS version. We obtained an area per 
+lipid of 0.615~nm$^2$, which is below the 0.648 nm$^2$ found by the original authors with their PME setup (see Ref.~\citenum{poger12}). We explain this by the fact that we used
+1.4~nm for the LJ cut-off, whereas a value of 1.0~nm was used in the original publication.
 
 \subsection{Slipids}
-Initial coordinates for a hydrated DPPC (at 323~K) and POPC (at 310~K) bilayers (30 and 40 waters/lipid, respectively) were taken directly from the 
-Slipids home page \url{http://mmkluster.fos.su.se/slipids/Downloads.html}.  The Slipids force field~\cite{jambeck12,jambeck12b} was used for the the all atom descriptions of DPPC and POPC, and
+Initial coordinates for hydrated DPPC (at 323~K) and POPC (at 310~K) bilayers (30 and 40 waters/lipid, respectively) were taken directly from the 
+Slipids home page (\url{http://mmkluster.fos.su.se/slipids/Downloads.html}).  The Slipids force field~\cite{jambeck12,jambeck12b} was used for the the all atom descriptions of DPPC and POPC, and
 water was described with the TIP3P water model~\cite{jorgensen83}. Simulations were performed within the NPT ensemble using the GROMACS 4.6.X simulation
 package~\cite{hess08}. The Nos\'{e}--Hoover thermostat~\cite{nose84,hoover85} was used with reference temperatures of
-323~K (DPPC) and 310~K (POPC) and a relaxation time constant of 0.5~ps. Water and lipids were coupled separately to
+323~K (DPPC) and 310~K (POPC) and a relaxation time constant of 0.5~ps; water and lipids were coupled separately to
 the heat bath. Pressure was kept constant at 1.013~bar using a semi--isotropic Parrinello--Rahman
 barostat~\cite{parrinello81} with a time constant of 10.0~ps. Equations of motion were
-integrated with the leapfrog algorithm using a timestep of 2~fs. Long range
+integrated with the leapfrog algorithm using a timestep of 2~fs. Long-range
 electrostatic interactions were calculated using the PME method~\cite{darden93,essman95}, with a fourth order
@@
 whereas water molecules were constrained using SETTLE~\cite{miyamoto92}. Twin-range cutoffs,
-1.0~nm and 1.6~nm, were used for the neighbor lists with the longrange neighbor list updated every
-10 steps. This simulation protocol corresponds to the protocol used in Ref~\cite{jambeck13}.
+1.0~nm and 1.6~nm, were used for the neighbor lists with the long-range neighbor list updated every
+10 steps. This simulation protocol corresponds to the protocol used in Ref~\citenum{jambeck13}.
 
 \subsection{Kukol}
-A bilayer patch with 512 POPC lipids was constructed and hydrated with $\sim$40 SPC water molecules per lipid. 
+A bilayer patch with 512 POPC lipids was constructed and hydrated with $\sim$40 SPC~\cite{berendsen81} water molecules per lipid. 
 The force field parameters were obtained from Lipidbook \cite{domanski10}.
@@
 All bonds were constrained with LINCS~\cite{hess97,hess07}. PME~\cite{darden93,essman95} was employed for the long-range electrostatics. Lennard-Jones interactions 
-were cut off at 1.4~nm. A neighbour list with a radius of 0.8~nm was updated every 5~steps. The constant temperature of 298~K 
+were cut off at 1.4~nm. A neighbor list with a radius of 0.8~nm was updated every 5~steps. Temperature of 298~K 
 was maintained with the Berendsen thermostat \cite{berendsen84} with a time constant of 0.1~ps. The Berendsen barostat \cite{berendsen84} 
@@
 
-\subsection{Chiu et al.}
+\subsection{Chiu}
 The force field parameters and the initial configuration were available through the Lipidbook~\cite{domanski10}.
-Timestep of 2~fs was used with leapfrog integrator. Covalent bond lengths were constrained with LINCS algorithm~\cite{hess97,hess07}. 
-Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with real space cut-off of 1.0~nm was used 
+Timestep of 2~fs was used with leapfrog integrator. Covalent bond lengths were constrained with the LINCS algorithm~\cite{hess97,hess07}. 
+Coordinates were written every 10~ps. PME~\cite{darden93,essman95} with a real space cut-off of 1.0~nm was used 
 for electrostatics. Twin range cut-off was used for the Lennard-Jones interactions with short and long cut-offs of 1.0~nm and 1.6~nm, respectively.
-The neighbour lists with a cut-off of 1.0~nm were updated every 5 steps. Temperature was coupled separately
+The neighbor lists with a cut-off of 1.0~nm were updated every 5 steps. Temperature was coupled separately
 for lipids and water to 298~K with the velocity-rescale method~\cite{bussi07} with a coupling constant 0.2~ps.
@@
 
-\subsection{Ulmschneider}
+\subsection{Ulmschneiders}
 The initial structure containing 128 POPC molecules with 3328 TIP3P water~\cite{jorgensen83} molecules (26 per lipid) was downloaded from Lipidbook \cite{domanski10} 
@@
 constrained with LINCS~\cite{hess97,hess07}. The temperature was kept at 298~K with the Berendsen thermostat~\cite{berendsen84}. The pressure was semi-isotropically coupled to the Berendsen 
-barostat~\cite{berendsen84} with a time constant of 1~ps and a target pressure of 1~bar. PME~\cite{darden93,essman95} was employed for long range electrostatics and a cut-off of 1~ns was employed for 
-the Lennard-Jones interactions. A neighbour list with a radius of 1~nm was updated every 10~steps.
+barostat~\cite{berendsen84} with a time constant of 1~ps and a target pressure of 1~bar. PME~\cite{darden93,essman95} was employed for long-range electrostatics and a cut-off of 1~ns was employed for 
+the Lennard-Jones interactions. A neighbor list with a radius of 1~nm was updated every 10~steps.
 
-Additionally, the simulations were repeated with the dispersion correction applied to pressure and temperature. Even though the area per lipid decrease
-d slightly, the headgroup order parameters were only slightly affected.
+Additionally, the simulations were repeated with the dispersion correction applied to pressure and temperature.
+Even though the area per lipid decreased slightly, the headgroup order parameters were only slightly affected.
 
-\subsection{Tj\"ornhammar et al.}
+\subsection{Tj\"ornhammar14}
 The gel phase DPPC bilayer structure delivered by Tj\"ornhammar  and Edholm~\cite{tjornhammar14} was ran for 5~ns at 343~K in order to destroy the 
 ordered gel configuration. This was followed by a 200~ns simulation at 323~K, i.e. in the fluid phase. The last 100~ns of this simulation was used for analysis. 
-The same mdp file as in the Supplementary Information section of the original paper ~\cite{tjornhammar14} was used except for the simulation temperature.
+Except for the simulation temperature, the same mdp file as in the Supplementary Information of the original paper~\cite{tjornhammar14} was used.
 
@@
 A hydrated bilayer consisting of 72 DLPC lipids and 2189 water molecules
- is constructed using the model by Lee et al. ~\cite{lee14}. This model describes 
- the important all-atom CHARMM36 character of the lipid headgroup but 
+ was constructed using the model by Lee et al.~\cite{lee14}. This model describes 
+ the important all-atom CHARMM36 character of the lipid headgroup, but 
 reduces the details of the lipid chains into a united-atom model.
-The initial  equilibrated structure was downloaded from the web page of J. Klauda, 
-Department of Chemical and Biomolecular Engineering University of Maryland.
- The parameter files are taken from the Supplementary Material of Ref. \citenum{lee14}.  
-This bilayer was equilibrated for 20~ns and the production run was 50~ns long,  with data was saved every 20~ps. 
+The initial  equilibrated structure was downloaded from the web page of J.~Klauda, 
+Department of Chemical and Biomolecular Engineering, University of Maryland.
+ The parameter files were taken from the Supplementary Material of Ref.~\citenum{lee14}.  
+This bilayer was equilibrated for 20~ns and the production run was 50~ns long,  with data saved every 20~ps. 
  The equations of motion were integrated using the multiple time step Verlet r-RESPA algorithm with
-a time step of 2 fs, and a calculation of the electrostatic forces only every two timesteps.
+a time step of 2 fs, and electrostatic forces calculated only every two timesteps.
  Covalent bonds between heavy and hydrogen atoms were constrained 
@@
 with a damping coefficient of 5~ps.  The modified NAMD version of the
-Nose-Hoover barostat with Langevin dynamics (piston period of
+Nos\'{e}--Hoover barostat with Langevin dynamics (piston period of
 0.1 ps and piston decay time of 0.05 ps)  was used semi-isotropically 
@@
 a force-based switching function  for distances beyond 1~nm.
-A neighbour list with a radius of 1.4~nm was updated every 10~timesteps.
+Neighbor lists with a radius of 1.4~nm were updated every 10~timesteps.
 
@@
 \subsection{Botan-CHARMM36-UA}
-This model is similar to the Lee-CHARMM36-UA model with a few  differences.
+This model was similar to the Lee-CHARMM36-UA model with a few  differences.
 A hydrated bilayer consisting of 128 DLPC lipids and 3840 water
-molecules is described by a model derived from the CHARMM27-UA model by H\'enin et al.~\cite{henin08}.
+molecules was described by a model derived from the CHARMM27-UA model by H\'enin et al.~\cite{henin08}.
 
-The distribution into all-atom  (AA) and united-atom (UA) parts within the lipids is the same
+The distribution into all-atom  (AA) and united-atom (UA) parts within the lipids was the same
 as in the original CHARMM27-UA model by H\'enin et al.~\cite{henin08}.  
@@
 for the first methylene groups of the two acyl chains (C22 and C32 in CHARMM36 topology): 
-their hydrogens are merged into  united atoms, whereas in Lee's model their hydrogens are described explicitely.
-The  united-atom Berger model ~\cite{berger97} was used for the tails and the UA-AA interactions as in Ref. ~\citenum{henin08}. 
-The difference between this model and the original one by H\'enin~\cite{henin08} is the replacement of the 
-AA parameters for the heads by  the parameters of the all-atom
- CHARMM36 force-field~\cite{klauda10}. Contrarily to the model by Lee et al.~\cite{lee14}, no reparametrization was done.
+Their hydrogens are merged into  united atoms, whereas in Lee's model their hydrogens are described explicitly.
+The UA Berger model~\cite{berger97} was used for the tails and the UA-AA interactions as in Ref.~\citenum{henin08}. 
+The difference between this model and the original one by H\'enin~\cite{henin08} was the replacement of the 
+AA parameters for the heads by  the parameters of the AA
+ CHARMM36 force-field~\cite{klauda10}. Contrarily to the model by Lee et al.~\cite{lee14}, no reparameterization was done.
 
-The non-bonded interactions are calculated using an atom-based switching function with short and long cut-offs of 0.8 and 1.2~nm. 
-Long range electrostatic interactions are implemented using the particle-particle particle-mesh solver with a relative accuracy of $10^{-4}$. The system 
-is first equilibrated for 30~ns in the NP$\gamma$T ensemble (Nos\'{e}--Hoover~\cite{nose84,hoover85} style thermostat and barostat with anisotropic pressure coupling) 
-at 323~K and 1~bar with timestep of 1~fs. The following 20~ns of dynamics are taken for calculation of configurational averages. 
-Simulations were carried out by using the LAMMPS package~\cite{plimpton95}. %( http://doi.org/10.1006/jcph.1995.1039 )., the input files are available ( http://dx.doi.org/10.5281/zenodo.13821 ).
+The non-bonded interactions were calculated using an atom-based switching function with short and long cut-offs of 0.8 and 1.2~nm. 
+Long-range electrostatic interactions were implemented using the particle-particle particle-mesh solver with a relative accuracy of $10^{-4}$. The system 
+was first equilibrated for 30~ns in the NP$\gamma$T ensemble (Nos\'{e}--Hoover~\cite{nose84,hoover85} style thermostat and barostat with anisotropic pressure coupling) 
+at 323~K and 1~bar with a timestep of 1~fs. The following 20~ns of dynamics were taken for calculation of configurational averages. 
+Simulations were carried out using the LAMMPS package~\cite{plimpton95}. %( http://doi.org/10.1006/jcph.1995.1039 )., the input files are available ( http://dx.doi.org/10.5281/zenodo.13821 ).
 
-\section{Details of the rough subjective force field ranking (Figure 4)}
+\pagebreak
+\section{Details of the rough subjective force field ranking (Fig.~4)}
 
@@
 (sum of magnitude and forking deviation steps from 6 to 8).
-These groups are indicated by colors in Fig.~2.
+These groups are indicated by colors in Fig.~4.
 (Note that for $\mathrm{g_2}$, for which there can be no forking,
@@
 and the weights of the five carbons were summed up.
-The sum, given in the $\Sigma$-column of Fig.~2,
+The sum, given in the $\Sigma$-column of Fig.~4,
 was then used to (roughly and subjectively, as should be clear from the
@@
 
-\section{The effect of cholesterol on glycerol backbone structure}
+\newpage
+
+\section{Effect of cholesterol on glycerol backbone structure}
 \begin{figure*}[]
@@
   \caption{\label{dihsCHOLcharmm}
-    The effect of cholesterol content on the POPC glycerol backbone and choline dihedral angles in CHARMM36 model (T=303~K).}
+    The effect of cholesterol content on the POPC glycerol backbone and choline dihedral angles in the CHARMM36 model (T=303~K).}
 \end{figure*}
@@
   \caption{\label{GlycerolOrientation}
-    Orientation of g$_1$-g$_3$ (a) and g$_2$-g$_3$ (b) respect to membrane normal as a function of cholesterol content
-    calculated from CHARMM36 simulations in 298K and 303K.}
+    Orientations of the (a) g$_1$--g$_3$ and (b) g$_2$--g$_3$ vectors with respect to the membrane normal
+    as a function of cholesterol content.
+    Calculated from the CHARMM36 simulations at 298 and 303~K.}
 \end{figure*}
@@
 \section{Author Contributions}
-{\it Alexandru Botan} provided simulation results for charmm36-UA\\
+
+{\it Alexandru Botan} provided simulation results for CHARMM36-UA.\\[0.7ex]
 {\it Fernando Favela} prepared, performed and analyzed simulations
-with popc and cholesterol for the CHARMM36 All-Atom FF. \\
-{\it Patrick F. J. Fuchs} ran and analyzed the Poger simulations. Provided scientific information which significantly advanced the project (signs and forking of order parameters). \\
-{\it Matti Javanainen} prepared and performed simulations with multiple lipid models and analyzed the results. Supervised the work of JT.\\
-{\it Matej Kanduc} provided simulation results for the Berger DLPC model \\
-{\it Waldemar Kulig} prepared the MD simulations with cholesterol for the MacRog FF \\
-{\it Antti Lamberg} prepared and performed simulations with Berger and its variants to show the importance of the signs and stereospecific labeling of the order parameters. \\
-{\it Claire Loison} provided simulation results for charmm36-UA. \\
-{\it Alexander Lyubartsev} provided simulation results for Högberg08 force field.\\
-{\it Markus S. Miettinen} co-designed the project with OHSO and supported in the work management. Provided Berger DMPC trajectory; prepared parameter files for 72 lipids %POPC \& DPPC
-CHARMM36. Made Figures 2--4 and 7--9.\\
-{\it Luca Monticelli}: Critical discussions in all phases of the project. Collaboration with Jukka  M{\"a}{\"a}tt{\"a}. \\
-{\it Jukka M{\"a}{\"a}tt{\"a}}  prepared and performed simulations with Berger and Slipids models and analyzed the results.\\
-{\it O. H. Samuli Ollila} co-designed the project with MSM and managed the work. Ran and analyzed several simulations. Wrote the manuscript. \\
-{\it Marius Retegan} prepared and validated the GROMACS compatible parameter files for GAFFlipid and Lipid14 force fields.\\
-{\it Tomasz Rog} provided the topologies for the MacRog model and also the full hydration simulation data with this model. \\
+with POPC and cholesterol for the CHARMM36 all-atom force field. \\[0.7ex]
+{\it Patrick F. J. Fuchs} ran and analyzed the Poger simulations. Provided scientific information which significantly advanced the project (signs and forking of order parameters). \\[0.7ex]
+{\it Matti Javanainen} prepared and performed simulations with multiple lipid models and analyzed the results. Supervised the work of JT.\\[0.7ex]
+{\it Matej Kandu\v{c}} provided simulation results for the Berger DLPC model. \\[0.7ex]
+{\it Waldemar Kulig} prepared the simulations with cholesterol for the MacRog force field. \\[0.7ex]
+{\it Antti Lamberg} prepared and performed simulations with Berger and its variants to show the importance of the signs and stereospecific labeling of the order parameters. \\[0.7ex]
+{\it Claire Loison} provided simulation results for CHARMM36-UA. \\
+{\it Alexander Lyubartsev} provided simulation results for H\"ogberg08 force field.\\[0.7ex]
+{\it Markus S. Miettinen} co-designed the project with OHSO and supported in the work management. Provided the Berger DMPC trajectory; prepared parameter files for 72 lipids %POPC \& DPPC
+CHARMM36. Made the Figs. 2--4 and 7--9.\\[0.7ex]
+{\it Luca Monticelli} contributed critical discussions in all phases of the project. Collaborated with JM. \\[0.7ex]
+{\it Jukka M{\"a}{\"a}tt{\"a}}  prepared and performed simulations with Berger and Slipids models and analyzed the results.\\[0.7ex]
+{\it O. H. Samuli Ollila} co-designed the project with MSM and managed the work. Ran and analyzed several simulations. Wrote the manuscript. \\[0.7ex]
+{\it Marius Retegan} prepared and validated the GROMACS compatible parameter files for GAFFlipid and Lipid14 force fields.\\[0.7ex]
+{\it Tomasz R\'og} provided the topologies for the MacRog model, as well as the full hydration simulation data with this model. \\[0.7ex]
 {\it Hubert Santuz} prepared and performed the cholesterol simulations with CHARMM36 and analyzed the results. 
-Made significant contribution in the data management of the project.\\
-{\it Joona Tynkkynen} prepared and performed the dehydration simulations with the MacRog FF.\\
+Made significant contribution in the data management of the project.\\[0.7ex]
+{\it Joona Tynkkynen} prepared and performed the dehydration simulations with the MacRog force field.\\
 
-%\newpage
+\newpage
 

### 2015-11-24 HGmodelMANUSCRIPT/HGmodel_ACSsupplementary.tex
Minor changes in the proof stage

@@
 force field parameterization). Using setup i) based on RF, we were able to reproduce the results (i.e. area per lipid value of 0.63~nm$^2$) from the original work only with 
-GROMACS versions 4.0.X and earlier (the original authors~\cite{poger10} used GROMACS version 3.3.3). When switching to versions 4.5.X and above, the area per lipid dropped to below 0.58~nm$^2$. 
+GROMACS versions 4.0.X and earlier (the original authors~\cite{poger10} used GROMACS version 3.3.3). When switching to versions 4.5.X and above, the area per lipid dropped below 0.58~nm$^2$. 
 The GROMACS developers were contacted and a redmine issue opened (\url{http://redmine.gromacs.org/issues/1400}). Outcome: the difference comes from the new Trotter decomposition 
@@
 \subsection{Botan-CHARMM36-UA}
-This model was similar to the Lee-CHARMM36-UA model with a few  differences.
+This model was similar to the Lee-CHARMM36-UA model with a few  differences~\cite{botan15}.
 A hydrated bilayer consisting of 128 DLPC lipids and 3840 water
